---
title: "Outbreak Data Preparation"
author: "Data Analysis Team"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(tidyr)
```

# Introduction

This document implements a methodology to identify unique outbreaks from a dataset that may contain duplicate entries across different administrative levels. The approach focuses on:

1. Excluding outbreaks reported at the country level
2. Examining outbreaks at admin1 level and excluding them if there are outbreaks at admin2 or admin3 that:
   - Fall under the same admin1 region
   - Have start and end dates within the timeframe of the admin1 outbreak
3. Examining outbreaks at admin2 level and excluding them if there are outbreaks at admin3 that:
   - Fall under the same admin2 region
   - Have start and end dates within the timeframe of the admin2 outbreak

The result is a deduplicated dataset that contains only the most granular (specific) outbreak information available.

# Data Preparation

## Reading and Preprocessing Data

```{r}
# Function to read and preprocess the outbreak data
read_outbreak_data <- function(file_path) {
  # Read the CSV file
  outbreak_data <- fread(file_path, stringsAsFactors = FALSE)
  # outbreak_data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert date columns to Date objects
  outbreak_data$start_date <- as.Date(outbreak_data$start_date)
  outbreak_data$end_date <- as.Date(outbreak_data$end_date)
  
  # Filter out country-level outbreaks as specified
  outbreak_data <- outbreak_data[spatial_scale != "country"] 
  

  # Ensure proper ordering by spatial scale
  outbreak_data$spatial_scale <- factor(outbreak_data$spatial_scale, 
                                        levels = c("admin1", "admin2", "admin3"))
  
  return(outbreak_data)
}
```

## Extracting Administrative Hierarchy

```{r}
# Function to extract administrative hierarchy from location string
extract_admin_hierarchy <- function(location) {
  parts <- strsplit(location, "::")[[1]]
  
  # The first element is the region (e.g., "afr"), the second is the country code
  if (length(parts) < 3) {
    return(list(admin1 = NA, admin2 = NA, admin3 = NA))
  }
  
  admin1 <- paste(parts[1:3], collapse = "::")
  admin2 <- if (length(parts) >= 4) paste(parts[1:4], collapse = "::") else NA
  admin3 <- if (length(parts) >= 5) location else NA
  
  return(list(admin1 = admin1, admin2 = admin2, admin3 = admin3))
}

# Function to identify if an outbreak is within another outbreak's timeframe
is_within_timeframe <- function(child_start, child_end, parent_start, parent_end) {
  # Check if child outbreak timeframe is within parent outbreak timeframe
  # Use vectorized approach to handle multiple dates
  if(length(child_start) > 1) {
    return(child_start >= parent_start & child_end <= parent_end)
  } else {
    return(child_start >= parent_start && child_end <= parent_end)
  }
}
```

# Deduplication Algorithm

## Main Deduplication Function

```{r}
# Main function to deduplicate outbreaks
deduplicate_outbreaks <- function(outbreak_data) {
  # Add columns to store hierarchy information
  outbreak_data <- outbreak_data %>%
    rowwise() %>%
    mutate(
      admin_hierarchy = list(extract_admin_hierarchy(location)),
      admin1_location = admin_hierarchy$admin1,
      admin2_location = admin_hierarchy$admin2,
      admin3_location = admin_hierarchy$admin3
    ) %>%
    ungroup()
  
  # Initialize a column to mark outbreaks for exclusion
  outbreak_data$exclude <- FALSE
  
  # Process admin1 outbreaks
  admin1_outbreaks <- outbreak_data %>% 
    filter(spatial_scale == "admin1")
  
  # For each admin1 outbreak
  for (i in 1:nrow(admin1_outbreaks)) {
    admin1_loc <- admin1_outbreaks$location[i]
    admin1_start <- admin1_outbreaks$start_date[i]
    admin1_end <- admin1_outbreaks$end_date[i]
    
    # Find potential children (admin2 or admin3) that have this admin1 as parent
    potential_children <- outbreak_data %>%
      filter(
        (spatial_scale == "admin2" | spatial_scale == "admin3") &
        startsWith(location, paste0(admin1_loc, "::"))
      )
    
    # Check if any children are within the timeframe of this admin1 outbreak
    children_within_timeframe <- potential_children %>%
      filter(
        start_date >= admin1_start & end_date <= admin1_end
      )
    
    # If there are children within the timeframe, mark this admin1 for exclusion
    if (nrow(children_within_timeframe) > 0) {
      outbreak_data$exclude[outbreak_data$location == admin1_loc & 
                            outbreak_data$start_date == admin1_start] <- TRUE
    }
  }
  
  # Process admin2 outbreaks
  admin2_outbreaks <- outbreak_data %>% 
    filter(spatial_scale == "admin2")
  
  # For each admin2 outbreak
  for (i in 1:nrow(admin2_outbreaks)) {
    admin2_loc <- admin2_outbreaks$location[i]
    admin2_start <- admin2_outbreaks$start_date[i]
    admin2_end <- admin2_outbreaks$end_date[i]
    
    # Find potential children (admin3) that have this admin2 as parent
    potential_children <- outbreak_data %>%
      filter(
        spatial_scale == "admin3" &
        startsWith(location, paste0(admin2_loc, "::"))
      )
    
    # Check if any children are within the timeframe of this admin2 outbreak
    children_within_timeframe <- potential_children %>%
      filter(
        start_date >= admin2_start & end_date <= admin2_end
      )
    
    # If there are children within the timeframe, mark this admin2 for exclusion
    if (nrow(children_within_timeframe) > 0) {
      outbreak_data$exclude[outbreak_data$location == admin2_loc & 
                            outbreak_data$start_date == admin2_start] <- TRUE
    }
  }
  
  # Return the unique outbreaks (not marked for exclusion)
  unique_outbreaks <- outbreak_data %>%
    filter(!exclude)
  
  return(unique_outbreaks)
}
```

# Validation Methods

## Basic Validation

```{r}
# Function to validate the deduplication results
validate_deduplication <- function(original_data, deduplicated_data) {
  # Calculate basic statistics
  total_original <- nrow(original_data)
  total_after_country_filter <- nrow(original_data %>% filter(spatial_scale != "country"))
  total_deduplicated <- nrow(deduplicated_data)
  
  excluded_count <- total_after_country_filter - total_deduplicated
  
  # Count by spatial scale
  original_by_scale <- table(original_data$spatial_scale)
  deduplicated_by_scale <- table(deduplicated_data$spatial_scale)
  
  # Create a validation report
  validation_report <- list(
    total_original = total_original,
    total_after_country_filter = total_after_country_filter,
    total_deduplicated = total_deduplicated,
    excluded_count = excluded_count,
    excluded_percentage = round(excluded_count / total_after_country_filter * 100, 2),
    original_by_scale = original_by_scale,
    deduplicated_by_scale = deduplicated_by_scale
  )
  
  return(validation_report)
}
```

## Extended Validation

```{r}
# Extended validation function
validate_deduplication_extended <- function(original_data, deduplicated_data) {
  # Basic validation
  basic_validation <- validate_deduplication(original_data, deduplicated_data)
  
  # Identify the excluded outbreaks
  original_ids <- paste0(original_data$location, "_", original_data$start_date)
  deduplicated_ids <- paste0(deduplicated_data$location, "_", deduplicated_data$start_date)
  excluded_ids <- setdiff(original_ids, deduplicated_ids)
  
  excluded_outbreaks <- original_data[original_ids %in% excluded_ids, ]
  
  # Analyze exclusions by administrative level
  exclusions_by_admin <- table(excluded_outbreaks$spatial_scale)
  
  # Analyze exclusions by country
  exclusions_by_country <- excluded_outbreaks %>%
    group_by(country) %>%
    summarise(count = n()) %>%
    arrange(desc(count))
  
  # Time range analysis
  time_range_excluded <- range(excluded_outbreaks$start_date)
  time_range_kept <- range(deduplicated_data$start_date)
  
  # Check if any admin1 outbreaks with children were NOT excluded
  false_negatives <- 0
  admin1_kept <- deduplicated_data %>% filter(spatial_scale == "admin1")
  
  for (i in 1:nrow(admin1_kept)) {
    admin1_loc <- admin1_kept$location[i]
    admin1_start <- admin1_kept$start_date[i]
    admin1_end <- admin1_kept$end_date[i]
    
    # Find potential children in the deduplicated data
    potential_children <- deduplicated_data %>%
      filter(
        (spatial_scale == "admin2" | spatial_scale == "admin3") &
        startsWith(location, paste0(admin1_loc, "::")) &
        start_date >= admin1_start & end_date <= admin1_end
      )
    
    if (nrow(potential_children) > 0) {
      false_negatives <- false_negatives + 1
    }
  }
  
  # Check the same for admin2 outbreaks
  admin2_kept <- deduplicated_data %>% filter(spatial_scale == "admin2")
  
  for (i in 1:nrow(admin2_kept)) {
    admin2_loc <- admin2_kept$location[i]
    admin2_start <- admin2_kept$start_date[i]
    admin2_end <- admin2_kept$end_date[i]
    
    # Find potential children in the deduplicated data
    potential_children <- deduplicated_data %>%
      filter(
        spatial_scale == "admin3" &
        startsWith(location, paste0(admin2_loc, "::")) &
        start_date >= admin2_start & end_date <= admin2_end
      )
    
    if (nrow(potential_children) > 0) {
      false_negatives <- false_negatives + 1
    }
  }
  
  # Create a comprehensive validation report
  extended_validation <- c(basic_validation, list(
    exclusions_by_admin = exclusions_by_admin,
    top_countries_with_exclusions = head(exclusions_by_country, 10),
    time_range_excluded = time_range_excluded,
    time_range_kept = time_range_kept,
    false_negatives = false_negatives
  ))
  
  return(extended_validation)
}

# Function to check for remaining duplicates
check_remaining_duplicates <- function(unique_outbreaks) {
  duplicates_found <- 0
  
  # Check admin1 outbreaks
  admin1_outbreaks <- unique_outbreaks %>% filter(spatial_scale == "admin1")
  
  for (i in 1:nrow(admin1_outbreaks)) {
    admin1_loc <- admin1_outbreaks$location[i]
    admin1_start <- admin1_outbreaks$start_date[i]
    admin1_end <- admin1_outbreaks$end_date[i]
    
    # Look for potential children in the unique outbreaks
    potential_children <- unique_outbreaks %>%
      filter(
        (spatial_scale == "admin2" | spatial_scale == "admin3") &
        startsWith(location, paste0(admin1_loc, "::")) &
        start_date >= admin1_start & end_date <= admin1_end
      )
    
    if (nrow(potential_children) > 0) {
      duplicates_found <- duplicates_found + 1
      cat(paste0("Found admin1 with children still in dataset: ", admin1_loc, "\n"))
    }
  }
  
  # Check admin2 outbreaks
  admin2_outbreaks <- unique_outbreaks %>% filter(spatial_scale == "admin2")
  
  for (i in 1:nrow(admin2_outbreaks)) {
    admin2_loc <- admin2_outbreaks$location[i]
    admin2_start <- admin2_outbreaks$start_date[i]
    admin2_end <- admin2_outbreaks$end_date[i]
    
    # Look for potential children in the unique outbreaks
    potential_children <- unique_outbreaks %>%
      filter(
        spatial_scale == "admin3" &
        startsWith(location, paste0(admin2_loc, "::")) &
        start_date >= admin2_start & end_date <= admin2_end
      )
    
    if (nrow(potential_children) > 0) {
      duplicates_found <- duplicates_found + 1
      cat(paste0("Found admin2 with children still in dataset: ", admin2_loc, "\n"))
    }
  }
  
  if (duplicates_found == 0) {
    cat("Validation successful: No remaining duplicates found.\n")
  } else {
    cat(paste0("Validation failed: Found ", duplicates_found, " cases where potential duplicates remain.\n"))
  }
  
  return(duplicates_found)
}
```

# Testing the Algorithms

## Test on Sample Data

```{r}
# Function to test the deduplication on a sample
test_deduplication <- function(outbreak_data, sample_size = 100) {
  # Create a subset of the data for testing
  set.seed(123) # For reproducibility
  sample_indices <- sample(1:nrow(outbreak_data), min(sample_size, nrow(outbreak_data)))
  sample_data <- outbreak_data[sample_indices, ]
  
  # Run deduplication on sample
  cat("Testing deduplication on a sample of", nrow(sample_data), "records...\n")
  deduplicated_sample <- deduplicate_outbreaks(sample_data)
  
  # Validate results
  validation <- validate_deduplication(sample_data, deduplicated_sample)
  
  cat("\nTest Validation Report:\n")
  cat(paste0("Sample size: ", nrow(sample_data), "\n"))
  cat(paste0("Unique outbreaks after deduplication: ", nrow(deduplicated_sample), "\n"))
  cat(paste0("Number of outbreaks excluded: ", validation$excluded_count, 
             " (", validation$excluded_percentage, "%)\n"))
  
  # Example of checking specific cases
  if (nrow(deduplicated_sample) < nrow(sample_data)) {
    cat("\nExample of excluded outbreak:\n")
    excluded <- sample_data %>% 
      filter(!location %in% deduplicated_sample$location)
    
    if (nrow(excluded) > 0) {
      print(excluded[1, c("location", "spatial_scale", "start_date", "end_date")])
      
      # Try to find why it was excluded (find its potential children)
      if (excluded$spatial_scale[1] == "admin1") {
        cat("\nPotential children that caused exclusion:\n")
        children <- sample_data %>%
          filter(
            (spatial_scale == "admin2" | spatial_scale == "admin3") &
            startsWith(location, paste0(excluded$location[1], "::")) &
            start_date >= excluded$start_date[1] & end_date <= excluded$end_date[1]
          )
        
        if (nrow(children) > 0) {
          print(children[, c("location", "spatial_scale", "start_date", "end_date")])
        } else {
          cat("No children found in the sample that would explain exclusion.\n")
        }
      }
    }
  }
  
  return(deduplicated_sample)
}
```

## Full Deduplication Process

```{r}
# Function to run the full deduplication process
run_deduplication <- function(file_path) {
  cat("Reading and preprocessing outbreak data...\n")
  outbreak_data <- read_outbreak_data(file_path)
  
  cat("Running deduplication algorithm...\n")
  unique_outbreaks <- deduplicate_outbreaks(outbreak_data)
  
  cat("Validating results...\n")
  validation <- validate_deduplication(outbreak_data, unique_outbreaks)
  
  cat("\nValidation Report:\n")
  cat(paste0("Total outbreaks (after filtering country level): ", validation$total_after_country_filter, "\n"))
  cat(paste0("Total unique outbreaks after deduplication: ", validation$total_deduplicated, "\n"))
  cat(paste0("Number of outbreaks excluded as potential duplicates: ", validation$excluded_count, 
             " (", validation$excluded_percentage, "%)\n"))
  
  cat("\nOutbreaks by spatial scale (before deduplication):\n")
  print(validation$original_by_scale)
  
  cat("\nOutbreaks by spatial scale (after deduplication):\n")
  print(validation$deduplicated_by_scale)
  
  # Return the deduplicated data
  return(unique_outbreaks)
}
```

# Execute Deduplication Process

```{r}
# Set the file path to your data
file_path <- "data/outbreak_data_summary.csv"

# Read in the data
outbreak_data <- read_outbreak_data(file_path)

# Show basic information about the dataset
cat("Dataset Information:\n")
cat(paste0("Total number of outbreaks: ", nrow(outbreak_data), "\n"))
cat("Outbreaks by spatial scale:\n")
print(table(outbreak_data$spatial_scale))

# Test the method on a sample
sample_results <- test_deduplication(outbreak_data, 200)
```

```{r}
# Run the full deduplication
unique_outbreaks <- run_deduplication(file_path)

# Verify there are no remaining duplicates
cat("\nVerifying deduplication results...\n")
remaining_duplicates <- check_remaining_duplicates(unique_outbreaks)
```

# Data Visualization

## Geographic Distribution

```{r}
# Geographic distribution of unique outbreaks
country_counts <- unique_outbreaks %>%
  group_by(country) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Plot the top 10 countries by outbreak count
ggplot(head(country_counts, 10), aes(x = reorder(country, count), y = count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Top 10 Countries by Number of Unique Outbreaks",
    x = "Country",
    y = "Number of Outbreaks"
  ) +
  theme_minimal()
```

## Temporal Distribution

```{r}
# Add year column
unique_outbreaks$year <- lubridate::year(unique_outbreaks$start_date)

# Create temporal distribution
year_counts <- unique_outbreaks %>%
  group_by(year) %>%
  summarise(count = n()) %>%
  arrange(year)

# Plot temporal distribution
ggplot(year_counts, aes(x = year, y = count)) +
  geom_line(group = 1, color = "steelblue", size = 1) +
  geom_point(color = "steelblue", size = 3) +
  labs(
    title = "Temporal Distribution of Outbreaks",
    x = "Year",
    y = "Number of Outbreaks"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(year_counts$year), max(year_counts$year), by = 1))
```

## Distribution by Administrative Level

```{r}
# Distribution by administrative level
admin_level_counts <- as.data.frame(table(unique_outbreaks$spatial_scale))
names(admin_level_counts) <- c("admin_level", "count")

# Plot distribution by administrative level
ggplot(admin_level_counts, aes(x = admin_level, y = count, fill = admin_level)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Distribution of Outbreaks by Administrative Level",
    x = "Administrative Level",
    y = "Number of Outbreaks"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues")
```

# Save Results

```{r}
# Save the deduplicated data to a CSV file
output_file <- "unique_outbreaks.csv"
write.csv(unique_outbreaks, output_file, row.names = FALSE)
cat(paste0("Deduplicated outbreaks saved to: ", output_file, "\n"))
```

# Conclusion

This analysis has identified unique outbreaks by filtering out potential duplicates across different administrative levels. The deduplication process:

1. Started with `r nrow(outbreak_data)` outbreaks (excluding country-level)
2. Identified `r nrow(outbreak_data) - nrow(unique_outbreaks)` potential duplicates (outbreaks at higher administrative levels that had more granular data available)
3. Resulted in `r nrow(unique_outbreaks)` unique outbreaks

The methodology provides a systematic approach to remove duplicates while preserving the most granular outbreak information available. This ensures that each unique outbreak is counted only once in subsequent analyses.

# Appendix: Function Reference

## Complete List of Functions

- `read_outbreak_data()`: Reads and preprocesses outbreak data
- `extract_admin_hierarchy()`: Extracts administrative hierarchy from location strings
- `is_within_timeframe()`: Checks if a date range is within another date range
- `deduplicate_outbreaks()`: Main deduplication algorithm
- `validate_deduplication()`: Basic validation of deduplication results
- `validate_deduplication_extended()`: Extended validation with detailed analysis
- `check_remaining_duplicates()`: Verifies that no duplicates remain in the dataset
- `test_deduplication()`: Tests the deduplication on a sample of data
- `run_deduplication()`: Runs the complete deduplication process
