---
title: "Data derivation"
format: html
editor_options: 
  chunk_output_type: console
---


## Load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(lubridate)
library(stringr)
library(data.table)
library(janitor)
library(tidyverse)
source("utils.R")
save_csv <- FALSE
```

## Outbreak data sets

Exclude the following outbreaks:

 1. Outbreaks too small or too large
  - Exclude outbreaks that were reported in populations that are too small or
  too large considering historic OCV campaigns from the global OCV stockpile

 2. Duplicate outbreaks across administrative levels
  - Leave the outbreak reported at the highest admin unit and remove those that
    were encapsulated in terms of time and location. 

 3. OCV-affected outbreaks
  - Exclude outbreaks from regions where OCV campaigns occurred within 3 years
    prior to the outbreak (as vaccine protection typically lasts 3+ years)

```{r}
# Step 0: Load initial data
outbreaks0 <- clean_names(fread("data/outbreak_files_Oct2024/outbreak_summary.csv"))
outbreaks0 |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> outbreaks0

setDT(outbreaks0) 

# Step 1: Filter by size (too small or too large)
ocv_use <- fread("outputs/ocv_use_GTFCC_data_20241120.csv")
max_ocv_dose <- 1e3 * max(ocv_use$DOSE_ROUND_ONE, na.rm = TRUE)
min_ocv_dose <- 1e3 * min(ocv_use$DOSE_ROUND_ONE, na.rm = TRUE)

outbreaks1 <- outbreaks0[outbreak_pop <= max_ocv_dose & outbreak_pop >= min_ocv_dose]
excluded1 <- 
  outbreaks0[!id_outbreak %in% outbreaks1$id_outbreak][, filter_reason := "Too small or too large"]

# Step 2: Remove nested outbreaks
outbreaks2 <- remove_nested_outbreaks(outbreaks1, 
                                      selection_method = "size_duration_specificity")
included2 <- outbreaks2$deduplicated
excluded2 <- outbreaks1[!id_outbreak %in% included2$id_outbreak][, filter_reason := "Nested outbreak"]

# Step 3: Remove outbreaks after 2020 (OCV data unavailable)
included3 <- included2[start_date <= as.Date("2021-01-01")]
excluded3 <- 
  included2[!id_outbreak %in% included3$id_outbreak][, filter_reason := "Outbreak after 2020"]

# Step 4: Remove OCV-affected outbreaks
docv_ <- clean_names(fread("data/outbreak_files_Oct2024/ocv_long_dataset.csv"))
docv <- docv_[(ocv_before_1 == TRUE | ocv_before_2 == TRUE | 
               ocv_before_3 == TRUE | ocv_during == TRUE)]

included4 <- included3[!id_outbreak %in% docv$id_outbreak]
excluded4 <- included3[id_outbreak %in% docv$id_outbreak][, filter_reason := "OCV-affected outbreak"]

# Combine all excluded
# excluded3 and excluded4 inherit the duplicate_group column from included2 
excluded1[, duplicate_group := NA]
excluded2[, duplicate_group := NA]
# adjust the column order
setcolorder(excluded1, c(setdiff(names(excluded1), c("duplicate_group", "filter_reason")), "duplicate_group", "filter_reason"))
setcolorder(excluded2, c(setdiff(names(excluded2), c("duplicate_group", "filter_reason")), "duplicate_group", "filter_reason"))
excluded1$start_date <- as.Date(excluded1$start_date)
excluded1$end_date <- as.Date(excluded1$end_date)
excluded2$start_date <- as.Date(excluded2$start_date)
excluded2$end_date <- as.Date(excluded2$end_date)

excluded_all <- rbindlist(list(excluded1, excluded2, excluded3, excluded4), fill = TRUE)

# Save data sets
if (save_rds) {
  saveRDS(outbreaks0, paste0("outputs/outbreaks_initial_", tstamp(), ".rds"))
  saveRDS(included4, paste0("outputs/outbreaks_included_", tstamp(), ".rds"))
  saveRDS(excluded_all, paste0("outputs/outbreaks_excluded_", tstamp(), ".rds"))
}
# Create summary log
filter_log <- data.frame(
  Step = c("Initial", "Size filter", "Nested filter", "OCV cutoff", "OCV-affected"),
  Included = c(nrow(outbreaks0), nrow(outbreaks1), nrow(included2), nrow(included3), nrow(included4)),
  Excluded = c(0, nrow(excluded1), nrow(excluded2), nrow(excluded3), nrow(excluded4))
)

if (save_csv) {
  fwrite(filter_log, paste0("outputs/outbreak_filter_summary_", tstamp(),".csv"), 
          row.names = FALSE)
}
```


Outbreak characteristics depending on the exclusion criteria
```{r}
outbreak_data_all <- as.data.frame(fread("outputs/outbreak_data_all_20250228.csv"))

# Load outbreaks not affected by OCVs (for modeling)
outbreak_data <- fread("outputs/outbreak_data_20250228.csv") 
outbreaks_filtered <-
  readRDS("outputs/outbreak_summary_filtered_20250418.rds")

outbreak_data %<>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 


# Identify outbreaks that have been excluded or included in analysis
ids_excl <- outbreak_data_all |> 
  filter(!(id_outbreak %in% outbreak_data$id_outbreak)) |>  
  pull(id_outbreak) |>
  unique()

# Identify confirmed outbreaks (with laboratory confirmation)
ids_conf <- outbreak_data |> 
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique()

# Identify suspected (not confirmed) outbreaks
ids_susp <- outbreak_data |>
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |>
  filter(!(c_ch_sum > 0)) |>
  pull(id_outbreak) |>
  unique()

# data on the subregion name
d <- fread("data/africa_subregion_country_names.csv")
d$iso3c <- sapply(d$Country, function(x) countrycode::countrycode(x, origin="country.name.en", destination = "iso3c"))

outbreak_data_all$iso3c <- outbreak_data_all$country
d0 <- left_join(outbreak_data_all, d, by="iso3c")
d0$outbreak_group <- ifelse(d0$id_outbreak %in% ids_excl, "Excluded", 
                            ifelse(d0$id_outbreak %in% ids_conf, "Confirmed",
                                   "Suspected"))

d0$ocv_affected <- ifelse(d0$id_outbreak %in% docv$id_outbreak, "Yes", "No")
d0$excluded_step <- ifelse(d0$id_outbreak %in% excluded1$id_outbreak, "1",
                          ifelse(d0$id_outbreak %in% excluded2$id_outbreak, "2", ifelse(d0$id_outbreak %in% excluded3$id_outbreak, "3", ifelse(d0$id_outbreak %in% excluded4$id_outbreak, "4", "NA"))))
```


```{r}
# Get outbreak summary data
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Oct2024/outbreak_summary.csv")))

# Create unique IDs for each outbreak
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# Join outbreak data with summary details
outbk_data_summary <- 
  left_join(filter(outbreaks, week == 1), 
            ds[, c("outbreak_pop", "total_s_ch", "total_c_ch", 
                   "total_deaths", "duration_weeks", 
                  "peak_week_incidence", "time_to_peak_week",
                  "peak_week_incidence_per_100percent", 
                  "peak_week_incidence_per_10e3", "mean_r0_1st_week", 
                  "area_per_1km2", "rural_urban", "ar_per_10e3",
                  "cfr_per_100percent", "id_outbreak")], 
            by = "id_outbreak")
```

```{r}
# Get time series data
outbreak_ts <- 
  read_rds("data/outbreak_files_Oct2024/time_series_outbreak_extraction.rds")
```

```{r}
ds <- fread("data/outbreak_data_summary.csv")
ds$ID_outbreak <- paste0(ds$location, "-", ds$start_date, "-", ds$end_date)
ds$ID <- 1:nrow(ds)
# data for the outbreaks affected by OCV
docv <- fread("data/ocv_long_dataset.csv")
# exclude outbreaks that were affected by OCV campaigns using IDs 
dat <- ds[!(ds$ID_outbreak %in% docv$ID_outbreak),]
ts <- readRDS("data/outbreak_data_ts.rds")

fits <- fread("outputs/data_fit_eval_20240409.csv")
good_fits <- fits[PARAM_GLOBAL == 1 & PARAM_SANITY_5pct == 1 & 
                    FIT_SANITY == 1 & R2_adj >= 0.5]
```


Combine vaccine impact 
```{r}
vacc_cov <- 1:9
vacc_week <- 0:12
vacc_impact_list <- vector('list', length(vacc_cov))
for (i in 1:length(vacc_cov)) {
  vc <- vacc_cov[i]
  cols = paste0("inc_C", vc, "T", vacc_week)
  d = read.csv(paste0("G:\\My Drive\\Projects\\VIMC\\VIMC 2.0\\CholeraOutbreakModel\\julia\\outputs\\vacc_4p_cov", vc, "_20240425.csv"))
  # NA values are replaced with baseline incidence values when no vaccines were
  # introduced. The decision was made because NA values were created when
  # the vaccination was not implemented because outbreak ended earlier than
  # the vaccination start + duration
  for (col in cols) {
    nas = is.na(d[, col])
    d[nas, col] = d[nas, "inc_novacc"]
  }
  d |>
    group_by(data_id) |>
    summarise(data_id = data_id[1],
              ID_outbreak = ID_outbreak[1],
              across(contains("inc"), sum)) -> d2
  
  col_id = 1:ncol(d2)
  id_iter <- col_id[grepl("inc_C", names(d2))]
  d3 <- d2
  for (k in id_iter) {
    d3[,k] = 100 * (d2[,"inc_novacc"] - d2[,k]) / d2[,"inc_novacc"]
  }
  
  d4 = data.frame(data_id = rep(d3$data_id, length(vacc_week)),
                  ID_outbreak = rep(d3$ID_outbreak, length(vacc_week)),
                  vacc_eff = unlist(d3[, id_iter]), 
                  vacc_cov = vc,
                  vacc_week = rep(vacc_week, each=nrow(d3)))
 
  
  vacc_impact_list[[i]] <- d4
}

vacc_impact <- do.call('rbind', vacc_impact_list)
# vacc_impact_0425 based on parameters comparable to the static model
# i.e., no waning of vaccine-induced immunity and vaccine efficacy of [30, 64]
# fwrite(vacc_impact, paste0("outputs/dynamic_vim", tstamp(), ".csv"))
```

## Life expectancy data
```{r}
library(readxl)
estimates <- 
  readxl::read_xlsx("data/WPP2024_MORT_F05_1_LIFE_EXPECTANCY_BY_AGE_BOTH_SEXES.xlsx",
                     "Estimates")
nms <- estimates[12,]
estimates <- estimates[-(1:12),]
names(estimates) <- nms

data.table::fwrite(estimates, paste0("outputs/wpp_life_expectancy_", tstamp(), ".csv"))

# # for years 2022 and onwards
# med_variant <- 
#   readxl::read_xlsx("data/WPP2022_MORT_F05_1_LIFE_EXPECTANCY_BY_AGE_BOTH_SEXES.xlsx",
#                      "Medium variant")
# nms_ <- med_variant[12,]
# med_variant <- med_variant[-(1:12),]
# names(med_variant) <- nms_
# 
# life_expectancy <- rbind(estimates, med_variant)
# data.table::fwrite(life_expectancy, 
#                    paste0("outputs/wpp_life_expectancy_", tstamp(), ".csv"))
```
 

## Population by age to compute the proportion of population under 5
```{r}
library(readxl)
estimates <-
  readxl::read_xlsx("data/WPP2024_POP_F01_1_POPULATION_SINGLE_AGE_BOTH_SEXES.xlsx",
                     "Estimates")
nms <- estimates[12,]
estimates <- estimates[-(1:12),]
names(estimates) <- nms

data.table::fwrite(estimates, 
                   paste0("outputs/wpp_pop_by_age_", tstamp(), ".csv"))

# # for years 2022 and onwards
# med_variant <- 
#   readxl::read_xlsx("data/WPP2022_POP_F01_1_POPULATION_SINGLE_AGE_BOTH_SEXES.xlsx",
#                      "Medium variant")
# nms_ <- med_variant[12,]
# med_variant <- med_variant[-(1:12),]
# names(med_variant) <- nms_
# 
# pop_by_age <- rbind(estimates, med_variant)
pop_by_age <- estimates
pop_by_age$prop_u5 <- NA
for (i in 1:nrow(pop_by_age)) {
 pop_by_age$prop_u5[i] <- sum(as.double(pop_by_age[i, 12:16])) / 
   sum(as.double(pop_by_age[i, 12:112]))
}
data.table::fwrite(pop_by_age,
                   paste0("outputs/wpp_pop_by_age_", tstamp(), ".csv"))
```

  
## Indirect vaccine efficacy (IVE)

IVE data come from a Bayesian beta regression fit
```{r}
fit <- read_rds("outputs/fit_ive_matlab_20241008.rds")
# print(fit, pars=c("beta","phi","lp__"), probs=c(0.025, 0.5, 0.975))
ive_yrep <- list(yrep = as.data.frame(rstan::extract(fit)$yrep),
                vacc_cov = round(expit(rstan::extract(fit)$new_X[1,,2]), digits=2))
ive_yrep$vacc_cov
saveRDS(ive_yrep, paste0("outputs/ive_yrep_", tstamp(), ".rds"))

# plot(ive_yrep$eff_vacc_cov, ive_yrep$yrep[sample(1:nrow(ive_yrep$yrep),1), ], type="l", col='grey90')
# for(i in 1:100){
#   lines(ive_yrep$eff_vacc_cov, 
#         ive_yrep$yrep[sample(1:nrow(ive_yrep$yrep),1), ], col="grey90")
#   invisible(readline(prompt="Press [enter] to continue"))
# }
```

## Number of days lost due to illness

### Patients

1. Ilboudo PG, Huang XX, Ngwira B, Mwanyungwe A, Mogasale V, Mengel MA, et al. Cost-of-illness of cholera to households and health facilities in rural Malawi. PLoS One. 2017;12(9):e0185041. 
8 days for 100 people

2. Awalime DK, Davies-Teye BBK, Vanotoo LA, Owoo NS, Nketiah-Amponsah E. Economic evaluation of 2014 cholera outbreak in Ghana: a household cost analysis. Health Econ Rev. 2017 Dec 4;7(1):45. 


3. Tembo T, Simuyandi M, Chiyenu K, Sharma A, Chilyabanyama ON, Mbwili-Muleya C, et al. Evaluating the costs of cholera illness and cost-effectiveness of a single dose oral vaccination campaign in Lusaka, Zambia. PLoS One. 2019;14(5):e0215972. 


4. Figuereo S, Yoon I, Kaddu SS, Lubogo M, Baruch J, Hossain AA, et al. Cost of Cholera for Households and Health Facilities, Somalia. J Epidemiol Glob Health. 2024 Sep 1;14(3):1219–30. 

The study only account for the stay at the hospital. It may be reasonable to assume that twice as many days would be required as in 
Tembo et al. study

5. Poulos C, Riewpaiboon A, Stewart JF, Clemens J, Guh S, Agtini M, et al. Costs of Illness Due to Endemic Cholera. Epidemiol Infect. 2012 Mar;140(3):10.1017/S0950268811000513. 


6. Schaetti C, Weiss MG, Ali SM, Chaignat CL, Khatib AM, Reyburn R, et al. Costs of illness due to cholera, costs of immunization and cost-effectiveness of an oral cholera mass vaccination campaign in Zanzibar. PLoS Negl Trop Dis. 2012;6(10):e1844. 


```{r}
df <- data.frame(study = "Ilboudo (2017)", 
                 location = "Machinga and Zomba, Malawi", 
                 days_lost = 8.0, n = 100, 
                 description="patient inactive period")
df <- rbind(df, 
            data.frame(study = "Awalime (2017)", 
                       location = "La-Dadekotopon and Shai-Osudoku, Ghana", 
                       days_lost = 6.2, n = 439, 
                       description="days missed"))
df <- rbind(df,
            data.frame(study = "Tembo (2019)", 
                       location = "Zambia", 
                       days_lost = 6.49, n = 189, 
                       description="days of hospitalization plus days of inactivity"))
df <- rbind(df,
            data.frame(study = "Figuereo (2024)", 
                       location = "Somalia",
                       days_lost = 3.08, n = 400, 
                       description="length of stay at the cholera treatment center"))
df <- rbind(df, 
            data.frame(study = "Poulos (2012)", 
                       location = "Beira, Mozambique", 
                       days_lost = 6.85, n = 400, 
                       description="days ill for > 18 yo and adults"))
df <- rbind(df, 
            data.frame(study = "Schaetti (2012)", 
                       location = "Zanzibar, Tanzania", 
                       days_lost = 5, n = 95, 
                       description="duration of illness episode"))

df %>% filter(!(study %in% c("Tembo (2019)", "Figuereo (2024)"))) -> df

weighted.mean(df$days_lost, df$n)
with(df, sum(days_lost * n) / sum(n))
```

Individual studies
```{r}
#-------------------------------------------------------------------------------
study <- "Ilboudo (2017)"
days_inactive_patient <- c(8.0)
days_inactive_care_giver <- c(4.0)
n <- c(100)

#-------------------------------------------------------------------------------
study <- "Awalime (2017)"
# In Table 3: La-Dadekotopon: Total patient days missed = 2071, Average = 6.9 days. This implies that around 300 (2071 / 6.9 ≈ 300.14) cholera cases were included. While the text says that total number of days missed 48.03 for La-Dadekopon and 24.79 days for Shai-Osudoku, it isn't clear that how they got these numbers. Summing across all four categories--Days missed (Patient),  Days missed (Caregiver),  Admission Days,  and Travel Time (Days)--does not produce this number. It appears that Days missed include the days at the hospital (admission days) and the days missed after discharge. We therefore use this number.

# From the following text:
# ... The time components include; travel time to facility, time spent at facility
# till discharge and work or school days lost after discharge ...
# the days lost column seems to imply days lost after discharge, which is a bit ig 

n_pat_la <-round(2071/6.9) # patients for La-Dadekotopon
n_cg_la <- round(1451/4.84) # caregiver for La-Dadekotopon
n_pat_shai <- round(655/4.71) # patients for Shai-Osudoku
n_cg_shai <- round(388/1.41) # caregiver for Shai-Osudoku
n_pat_hosp_la <- round(823/2.74)
n_pat_hosp_shai <- round(357/2.57)

# Updated average days missed and inferred sample sizes
days_missed_patient <- c(6.90, 4.71)
n_patient <- c(n_pat_la, n_pat_shai) 
days_missed_patient_wt <- sum(days_missed_patient * n_patient) / sum(n_patient)

days_hosp <- c(2.74, 2.57)
n_pat <-c(n_pat_hosp_la, n_pat_hosp_shai)

days_hosp_wt <- sum(days_hosp * n_pat) / sum(n_pat)

days_missed_caregiver <- c(4.84, 1.41)
n_caregiver <- c(n_cl, n_cs)
# Compute weighted average
days_missed_caregiver_wt <- sum(days_missed_caregiver * n_caregiver) / sum(n_caregiver)

travel_time <- 0.02*24 # hours

#-------------------------------------------------------------------------------
study <- "Tembo (2019)"
# Days of hospitalisation (Tembo et al. 2019, Table 5)
counts_hosp <- c(134, 47, 8)
days_hosp_midpoints <- c(2, 5.5, 9)
days_hosp_mean <- sum(counts_hosp * days_hosp_midpoints) / sum(counts_hosp)

# Days of inactivity (Tembo et al. 2019, Table 5)
counts_inact <- c(29, 85, 62, 13)
days_inact_midpoints <- c(0, 2, 5.5, 9)
days_inactivity_mean <- sum(counts_inact * days_inact_midpoints) / sum(counts_inact)

(mean_days_lost <- days_hosp_mean + days_inactivity_mean)

#-------------------------------------------------------------------------------
study <- "Figuereo (2024)"
# based on the text lines 10- p.1224
days_hosp <- 3.08
n <- 400

#-------------------------------------------------------------------------------
# days ill appear to include days at the hospital and the missed school(or work) days 
study <- "Poulos (2012)"
# Table 4
days_missed_school <- 2.8 # sd(1.2)
days_missed_work <- 3.7 # sd (2.7)
days_missed <- c(days_missed_school, days_missed_work) # no work for under 18?, no school for >= 18?

days_ill_child <- 5.8 # sd(1.9) # < 18 yo
days_ill_adult <- 7.5 # sd(4.5)
days_ill <- c(days_ill_child, days_ill_adult)
n_child <- 19
n_adult <- 31
n <- c(n_child, n_adult)

days_ill_wt <- sum(days_ill * n) / sum(n)
days_missed_wt <- sum(days_missed * n) / sum(n)

####
study <- "Schaetti (2012)"
days_ill <- 5 # median from patient data
n <- 95
```

#### Care-givers

```{r}
df <- data.frame(study = "Ilboudo (2017)", 
                 location = "Machinga and Zomba, Malawi", 
                 days_lost = 4.0, n = 100, 
                 description="caregiver inactive period")
df <- rbind(df, 
            data.frame(study = "Awalime (2017)", 
                       location = "La-Dadekotopon and Shai-Osudoku, Ghana", 
                       days_lost = 3.2, n = 575, 
                       description="days missed"))

weighted.mean(df$days_lost, df$n)
with(df, sum(days_lost * n) / sum(n))
```

Individual studies
```{r}
#-------------------------------------------------------------------------------
study <- "Ilboudo (2017)"
days_inactive_care_giver <- c(4.0)
n <- c(100)

#-------------------------------------------------------------------------------
study <- "Awalime (2017)"
# In Table 3: La-Dadekotopon: Total patient days missed = 2071, Average = 6.9 days. This implies that around 300 (2071 / 6.9 ≈ 300.14) cholera cases were included. While the text says that total number of days missed 48.03 for La-Dadekopon and 24.79 days for Shai-Osudoku, it isn't clear that how they got these numbers. Summing across all four categories--Days missed (Patient),  Days missed (Caregiver),  Admission Days,  and Travel Time (Days)--does not produce this number. It appears that Days missed include the days at the hospital (admission days) and the days missed after discharge. We therefore use this number.

# From the following text:
# ... The time components include; travel time to facility, time spent at facility
# till discharge and work or school days lost after discharge ...
# the days lost column seems to imply days lost after discharge, which is a bit ig 

days_missed_caregiver <- c(4.84, 1.41)
n_caregiver <- c(n_cl, n_cs)
# Compute weighted average
days_missed_caregiver_wt <- sum(days_missed_caregiver * n_caregiver) / sum(n_caregiver)

```


## Severity 

Mbewe et al. (2025) in Lusaka, Zambia 

The number of cases by symptoms at admission across four
different patients outcomes. We assume cases with no dehydration and
some dehydration belong to the moderate case definition. This is a simplification
in that patients with no or some dehydration have died, which indicates that
those with moderate symptoms might have progressed to severe states

```{r}
## the third number indicates deaths
no_dehydra <- 1607+381+4+1603
some_dehydra <- 1955+691+22+1933
severe_dehydra <- 1031+252+20+1011

prop <- c(no_dehydra-4, some_dehydra-22, severe_dehydra+26) / 
  sum(c(no_dehydra, some_dehydra, severe_dehydra))
```


