---
title: "Data derivation"
format: html
editor_options: 
  chunk_output_type: console
---


## Load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(lubridate)
library(stringr)
library(data.table)
library(janitor)
library(tidyverse)
source("utils.R")
```

## Outbreak data sets

Outbreak Selection Criteria
Exclude the following outbreaks:

 1. Outbreaks too small or too large
  - Exclude outbreaks that were reported in populations that are too small or
  too large considering historic OCV campaigns from the global OCV stockpile

 2. Duplicate outbreaks across administrative levels
  - Leave the outbreak reported at the highest admin unit and remove those that
    were encapsulated in terms of time and location. 

 3. OCV-affected outbreaks
  - Exclude outbreaks from regions where OCV campaigns occurred within 3 years
    prior to the outbreak (as vaccine protection typically lasts 3+ years)

```{r}
# Step 0: Load initial data
outbreaks0 <- clean_names(fread("data/outbreak_files_Oct2024/outbreak_summary.csv"))
outbreaks0 |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> outbreaks0

setDT(outbreaks0) 

# Step 1: Filter by size (too small or too large)
ocv_use <- fread("outputs/ocv_use_GTFCC_data_20241120.csv")
max_ocv_dose <- 1e3 * max(ocv_use$DOSE_ROUND_ONE, na.rm = TRUE)
min_ocv_dose <- 1e3 * min(ocv_use$DOSE_ROUND_ONE, na.rm = TRUE)

outbreaks1 <- outbreaks0[outbreak_pop <= max_ocv_dose & outbreak_pop >= min_ocv_dose]
excluded1 <- 
  outbreaks0[!id_outbreak %in% outbreaks1$id_outbreak][, filter_reason := "Too small or too large"]

# Step 2: Remove nested outbreaks
outbreaks2 <- remove_nested_outbreaks(outbreaks1, selection_method = "size_duration_specificity")
included2 <- outbreaks2$deduplicated
excluded2 <- outbreaks1[!id_outbreak %in% included2$id_outbreak][, filter_reason := "Nested outbreak"]

# Step 3: Remove outbreaks after 2020 (OCV data unavailable)
included3 <- included2[start_date <= as.Date("2021-01-01")]
excluded3 <- 
  included2[!id_outbreak %in% included3$id_outbreak][, filter_reason := "Outbreak after 2020"]

# Step 4: Remove OCV-affected outbreaks
docv_ <- clean_names(fread("data/outbreak_files_Oct2024/ocv_long_dataset.csv"))
docv <- docv_[(ocv_before_1 == TRUE | ocv_before_2 == TRUE | 
               ocv_before_3 == TRUE | ocv_during == TRUE)]

included4 <- included3[!id_outbreak %in% docv$id_outbreak]
excluded4 <- included3[id_outbreak %in% docv$id_outbreak][, filter_reason := "OCV-affected outbreak"]

# Combine all excluded
# excluded3 and excluded4 inherit the duplicate_group column from included2 
excluded1[, duplicate_group := NA]
excluded2[, duplicate_group := NA]
# adjust the column order
setcolorder(excluded1, c(setdiff(names(excluded1), c("duplicate_group", "filter_reason")), "duplicate_group", "filter_reason"))
setcolorder(excluded2, c(setdiff(names(excluded2), c("duplicate_group", "filter_reason")), "duplicate_group", "filter_reason"))
excluded1$start_date <- as.Date(excluded1$start_date)
excluded1$end_date <- as.Date(excluded1$end_date)
excluded2$start_date <- as.Date(excluded2$start_date)
excluded2$end_date <- as.Date(excluded2$end_date)

excluded_all <- rbindlist(list(excluded1, excluded2, excluded3, excluded4), fill = TRUE)

# Save datasets
saveRDS(outbreaks0, paste0("outputs/outbreaks_initial_", tstamp(), ".rds"))
saveRDS(included4, paste0("outputs/outbreaks_included_", tstamp(), ".rds"))
saveRDS(excluded_all, paste0("outputs/outbreaks_excluded_", tstamp(), ".rds"))

# Create summary log
filter_log <- data.frame(
  Step = c("Initial", "Size filter", "Nested filter", "OCV cutoff", "OCV-affected"),
  Included = c(nrow(outbreaks0), nrow(outbreaks1), nrow(included2), nrow(included3), nrow(included4)),
  Excluded = c(0, nrow(excluded1), nrow(excluded2), nrow(excluded3), nrow(excluded4))
)

write.csv(filter_log, paste0("outputs/outbreak_filter_summary_", tstamp(),".csv"), 
          row.names = FALSE)
```


Outbreak characteristics depending on the exclusion criteria
```{r}
outbreak_data_all <- as.data.frame(fread("outputs/outbreak_data_all_20250228.csv"))

# Load outbreaks not affected by OCVs (for modeling)
outbreak_data <- fread("outputs/outbreak_data_20250228.csv") 
outbreaks_filtered <-
  readRDS("outputs/outbreak_summary_filtered_20250418.rds")

outbreak_data %<>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 


# (A) Outbreak time series by group
# Identify outbreaks that have been excluded or included in analysis
ids_excl <- outbreak_data_all |> 
  filter(!(id_outbreak %in% outbreak_data$id_outbreak)) |>  
  pull(id_outbreak) |>
  unique()

# Identify confirmed outbreaks (with laboratory confirmation)
ids_conf <- outbreak_data |> 
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique()

# Identify suspected (not confirmed) outbreaks
ids_susp <- outbreak_data |>
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |>
  filter(!(c_ch_sum > 0)) |>
  pull(id_outbreak) |>
  unique()

# data on the subregion name
d <- fread("data/africa_subregion_country_names.csv")
d$iso3c <- sapply(d$Country, function(x) countrycode::countrycode(x, origin="country.name.en", destination = "iso3c"))

outbreak_data_all$iso3c <- outbreak_data_all$country
d0 <- left_join(outbreak_data_all, d, by="iso3c")
d0$outbreak_group <- ifelse(d0$id_outbreak %in% ids_excl, "Excluded", 
                            ifelse(d0$id_outbreak %in% ids_conf, "Confirmed",
                                   "Suspected"))

d0$ocv_affected <- ifelse(d0$id_outbreak %in% docv$id_outbreak, "Yes", "No")
d0$excluded_step <- ifelse(d0$id_outbreak %in% excluded1$id_outbreak, "1",
                          ifelse(d0$id_outbreak %in% excluded2$id_outbreak, "2", ifelse(d0$id_outbreak %in% excluded3$id_outbreak, "3", ifelse(d0$id_outbreak %in% excluded4$id_outbreak, "4", "NA"))))
```


```{r}
# Get outbreak summary data
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Oct2024/outbreak_summary.csv")))

# Create unique IDs for each outbreak
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# Join outbreak data with summary details
outbk_data_summary <- 
  left_join(filter(outbreaks, week == 1), 
            ds[, c("outbreak_pop", "total_s_ch", "total_c_ch", 
                   "total_deaths", "duration_weeks", 
                  "peak_week_incidence", "time_to_peak_week",
                  "peak_week_incidence_per_100percent", 
                  "peak_week_incidence_per_10e3", "mean_r0_1st_week", 
                  "area_per_1km2", "rural_urban", "ar_per_10e3",
                  "cfr_per_100percent", "id_outbreak")], 
            by = "id_outbreak")
```

```{r}
# Get time series data
outbreak_ts <- 
  read_rds("data/outbreak_files_Oct2024/time_series_outbreak_extraction.rds")
```

```{r}
ds <- fread("data/outbreak_data_summary.csv")
ds$ID_outbreak <- paste0(ds$location, "-", ds$start_date, "-", ds$end_date)
ds$ID <- 1:nrow(ds)
# data for the outbreaks affected by OCV
docv <- fread("data/ocv_long_dataset.csv")
# exclude outbreaks that were affected by OCV campaigns using IDs 
dat <- ds[!(ds$ID_outbreak %in% docv$ID_outbreak),]
ts <- readRDS("data/outbreak_data_ts.rds")

fits <- fread("outputs/data_fit_eval_20240409.csv")
good_fits <- fits[PARAM_GLOBAL == 1 & PARAM_SANITY_5pct == 1 & 
                    FIT_SANITY == 1 & R2_adj >= 0.5]
```


Combine vaccine impact 
```{r}
vacc_cov <- 1:9
vacc_week <- 0:12
vacc_impact_list <- vector('list', length(vacc_cov))
for (i in 1:length(vacc_cov)) {
  vc <- vacc_cov[i]
  cols = paste0("inc_C", vc, "T", vacc_week)
  d = read.csv(paste0("G:\\My Drive\\Projects\\VIMC\\VIMC 2.0\\CholeraOutbreakModel\\julia\\outputs\\vacc_4p_cov", vc, "_20240425.csv"))
  # NA values are replaced with baseline incidence values when no vaccines were
  # introduced. The decision was made because NA values were created when
  # the vaccination was not implemented because outbreak ended earlier than
  # the vaccination start + duration
  for (col in cols) {
    nas = is.na(d[, col])
    d[nas, col] = d[nas, "inc_novacc"]
  }
  d |>
    group_by(data_id) |>
    summarise(data_id = data_id[1],
              ID_outbreak = ID_outbreak[1],
              across(contains("inc"), sum)) -> d2
  
  col_id = 1:ncol(d2)
  id_iter <- col_id[grepl("inc_C", names(d2))]
  d3 <- d2
  for (k in id_iter) {
    d3[,k] = 100 * (d2[,"inc_novacc"] - d2[,k]) / d2[,"inc_novacc"]
  }
  
  d4 = data.frame(data_id = rep(d3$data_id, length(vacc_week)),
                  ID_outbreak = rep(d3$ID_outbreak, length(vacc_week)),
                  vacc_eff = unlist(d3[, id_iter]), 
                  vacc_cov = vc,
                  vacc_week = rep(vacc_week, each=nrow(d3)))
 
  
  vacc_impact_list[[i]] <- d4
}

vacc_impact <- do.call('rbind', vacc_impact_list)
# vacc_impact_0425 based on parameters comparable to the static model
# i.e., no waning of vaccine-induced immunity and vaccine efficacy of [30, 64]
# fwrite(vacc_impact, paste0("outputs/dynamic_vim", tstamp(), ".csv"))
```

## Life expectancy data
```{r}
library(readxl)
estimates <- 
  readxl::read_xlsx("data/WPP2024_MORT_F05_1_LIFE_EXPECTANCY_BY_AGE_BOTH_SEXES.xlsx",
                     "Estimates")
nms <- estimates[12,]
estimates <- estimates[-(1:12),]
names(estimates) <- nms

data.table::fwrite(estimates, paste0("outputs/wpp_life_expectancy_", tstamp(), ".csv"))

# # for years 2022 and onwards
# med_variant <- 
#   readxl::read_xlsx("data/WPP2022_MORT_F05_1_LIFE_EXPECTANCY_BY_AGE_BOTH_SEXES.xlsx",
#                      "Medium variant")
# nms_ <- med_variant[12,]
# med_variant <- med_variant[-(1:12),]
# names(med_variant) <- nms_
# 
# life_expectancy <- rbind(estimates, med_variant)
# data.table::fwrite(life_expectancy, 
#                    paste0("outputs/wpp_life_expectancy_", tstamp(), ".csv"))
```
 

## Population by age to compute the proportion of population under 5
```{r}
library(readxl)
estimates <-
  readxl::read_xlsx("data/WPP2024_POP_F01_1_POPULATION_SINGLE_AGE_BOTH_SEXES.xlsx",
                     "Estimates")
nms <- estimates[12,]
estimates <- estimates[-(1:12),]
names(estimates) <- nms

data.table::fwrite(estimates, 
                   paste0("outputs/wpp_pop_by_age_", tstamp(), ".csv"))

# # for years 2022 and onwards
# med_variant <- 
#   readxl::read_xlsx("data/WPP2022_POP_F01_1_POPULATION_SINGLE_AGE_BOTH_SEXES.xlsx",
#                      "Medium variant")
# nms_ <- med_variant[12,]
# med_variant <- med_variant[-(1:12),]
# names(med_variant) <- nms_
# 
# pop_by_age <- rbind(estimates, med_variant)
pop_by_age <- estimates
pop_by_age$prop_u5 <- NA
for (i in 1:nrow(pop_by_age)) {
 pop_by_age$prop_u5[i] <- sum(as.double(pop_by_age[i, 12:16])) / 
   sum(as.double(pop_by_age[i, 12:112]))
}
data.table::fwrite(pop_by_age,
                   paste0("outputs/wpp_pop_by_age_", tstamp(), ".csv"))
```

  
## Indirect vaccine efficacy (IVE)

IVE data come from a Bayesian beta regression fit
```{r}
fit <- read_rds("outputs/fit_ive_matlab_20241008.rds")
# print(fit, pars=c("beta","phi","lp__"), probs=c(0.025, 0.5, 0.975))
ive_yrep <- list(yrep = as.data.frame(rstan::extract(fit)$yrep),
                vacc_cov = round(expit(rstan::extract(fit)$new_X[1,,2]), digits=2))
ive_yrep$vacc_cov
saveRDS(ive_yrep, paste0("outputs/ive_yrep_", tstamp(), ".rds"))

# plot(ive_yrep$eff_vacc_cov, ive_yrep$yrep[sample(1:nrow(ive_yrep$yrep),1), ], type="l", col='grey90')
# for(i in 1:100){
#   lines(ive_yrep$eff_vacc_cov, 
#         ive_yrep$yrep[sample(1:nrow(ive_yrep$yrep),1), ], col="grey90")
#   invisible(readline(prompt="Press [enter] to continue"))
# }
```


