---
title: "Exploratory data analysis"
format: html
editor_options: 
  chunk_output_type: console
---

## Load packages, source `utils.R`, and outbreak data sets

```{r}
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(scales)
library(viridis)
library(sf)
library(metR)
library(googlesheets4)
library(conflicted)
library(MASS)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
source("utils.R") # utility functions
save_plot <- FALSE
save_table <- FALSE
# figure size
fgw1 <- figure_size[figure_size$journal =="Lancet",]$single
fgw2 <- figure_size[figure_size$journal =="Lancet",]$double
set1 <- RColorBrewer::brewer.pal(9,"Set1") # color sets
library(patchwork)
# remove the space between the plot and the legend and the space around the plot
# representative probability values for quantile 
probs <- c(0.025,0.25,0.5,0.75,0.975)
VW <- 9 # representative vaccination week
VW_seq <- seq(1,15,2)
VC <- 0.75 # representative vaccination coverage rate
VC3 <- c(0.6,0.75,0.9) # three representative vaccination coverage rates

shp <- st_read("data/shapefiles/total_shp_0427.shp")
africa <- st_read("data/shapefiles/afr_g2014_2013_0.shp")
```

## Outbreak data and delay
```{r}
outbreak_data_all <- as.data.frame(fread("outputs/outbreak_data_all_20241024.csv"))
outbreak_data <- fread("outputs/outbreak_data_20241121.csv") # for vaccine impact modeling

# fwrite(ocv_use, paste0("outputs/ocv_use_GTFCC_data_", tstamp(), ".csv"))
ocv_use <- fread("outputs/ocv_use_GTFCC_data_20241120.csv")
delay_req_r1 <- as.integer(ocv_use$DATE_ROUND_ONE - ocv_use$DATE_REQUEST)
# max_pop_camp <- max(ocv_use$DOSE_DELIVERED_ROUND_ONE*1000, na.rm=T)
min_pop_camp <- min(ocv_use$DOSE_DELIVERED_ROUND_ONE*1000, na.rm=T)
ocv_delay <- quantile(delay_req_r1, probs=c(0.25,0.5,0.75,0.9), na.rm=TRUE)

# OCV use data
docv <- clean_names(fread("data/outbreak_files_Jun2024/ocv_long_dataset.csv"))
# consumer price index for cost adjustment (2015 to 2023)
dis_params <- as.data.frame(data.table::fread("data/parameters.csv"))
# IVE posterior predictive values from Bayesian beta regression (see parameters.qmd)
ive_fit <- read_rds("outputs/ive_yrep_20241010.rds")
ive_data <- ive_fit$yrep
rm(ive_fit) # large file size

# extract ids for the outbreaks not affected by OCV campaigns
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Oct2024/outbreak_summary.csv")))
paste0(length(unique(ds$country)), " countries")
# create 'id_outbreak' variables to link with the 
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds
```

```{r}
outbreak_data <- fread("outputs/outbreak_data_20241121.csv") # for vaccine impact modeling
outbreak_data[, data_id := .GRP, by = id_outbreak]

ggplot(outbreak_data)+
  geom_line(aes(week,s_ch,group = data_id), color="grey50")+
  theme_light()
ggplot(outbreak_data[data_id == 5])+
    geom_line(aes(week, s_ch, group = data_id), color="grey50")+
    theme_light()

d <- outbreak_data[data_id == 13]
plot(d$week, d$s_ch, type="l", col="grey50", ylim=c(0,200), xlim=c(0,50))

for (i in 2:length(unique(outbreak_data$data_id))) {
  cat("i = ", i, "\n")
  d <- outbreak_data[data_id == i]
  lines(d$week, d$s_ch, col="grey50")
  
  readline(prompt="Press [enter] to continue")

}
```


```{r}
ds |>
  ggplot()+
  geom_histogram(aes(x=ar_per_10e3))+
  scale_x_log10()

quantile(ds$ar_per_10e3/10, probs=c(0.025,0.25,0.6,0.74,0.975))


fit <- fread("julia/outputs/fit_4p_fA010_20241130T00.csv")
fit$data_id <- 1:nrow(fit)
fit[outbreak_data, on = 'data_id', id_outbreak := i.id_outbreak]
fit <- left_join(fit,
                 ds[,c("id_outbreak","ar_per_10e3","rural_urban","duration_weeks","total_s_ch")], 
                  by="id_outbreak")

# fit <- fit, on="id_outbreak"]
# fit[ds, on = 'id_outbreak', ar_per_10e3 := i.ar_per_10e3]

fwrite(fit, "julia/outputs/fit_4p_fA010_20241130T00_ds.csv")
```


## Outbreak summary dataset 

```{r}
# extract ids for the outbreaks not affected by OCV campaigns
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Jun2024/outbreak_data_summary.csv")))
ds <- ds[!is.na(ds$country), ] # why NA's? 
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds
# ds |>
#   group_by(id_outbreak) |>
#   summarize(n=n()) |>
#   arrange(desc(n))
# 
# ds |>
#   group_by(id_outbreak) |>
#   summarize(n=n()) |>
#   arrange(desc(n)) -> ds_ids
# 
# ds_ids_2 <- ds_ids |> filter(n > 1)
# 
# print(ds_ids, n=30)
# ds[ds$id_outbreak %in% ds_ids_2$id_outbreak, ]

# data for the outbreaks affected by OCV
docv <- clean_names(fread("data/outbreak_files_Jun2024/ocv_long_dataset.csv"))
ts <- clean_names(readRDS("data/outbreak_files_Jun2024/outbreak_data_ts.rds")) # time-series data
# fwrite(ts, "data/outbreak_files_Jun2024/outbreak_data_ts.csv") # to see in the spreadsheet
ts |>
  group_by(location, outbreak_number) |> # identify unique outbreaks
  mutate(week = row_number(),
         date = tl + 3, # pick the date for the midpoint for each week
         start_date_outbreak = min(tl), 
         end_date_outbreak = max(tr),
         id_outbreak = paste0(location, "-", start_date_outbreak, "-",
                              end_date_outbreak)) |>
  ungroup() -> ts # time series with unique ids

# plot to check the sanity of the time series
# ids <- unique(ts$id_outbreak)
# d1 <- ts[ts$id_outbreak == ids[1],  ]
# plot(d1$week, d1$s_ch, type="l", xlim=c(0, 100), ylim=c(0,10000))
# for (id in ids) {
#   d <- ts[ts$id_outbreak == id,  ]
#   lines(d$week, d$s_ch)
# }
length(unique(ts$id_outbreak))
sum(unique(ts$id_outbreak) %in% ds$id_outbreak)

ts |> 
  group_by(id_outbreak) |> 
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique() -> ids_conf
length(ids_conf)

# data.table::fwrite(ts, 
#                    paste0("outputs/outbreak_data_all_", tstamp(), ".csv"))
```


## Create subsets of outbreak data set to compute vaccine impact

### 1. Population size
```{r}
library(googlesheets4)
# OCV use during campaigns
ocv_use <- read_sheet('https://docs.google.com/spreadsheets/d/1g9Ddrp0Ft14BtmAlZ-j7r8YCEAddro6H2jcmLkn7u84/edit?gid=0#gid=0', 
                      sheet='Sheet1',
                      col_types = "cccDDDDDDDdddddddccccc")

# Infer population size by the number of vaccine doses  
max_ocv_dose <- max(ocv_use$DOSE_ROUND_ONE*1000, na.rm=T)
min_ocv_dose <- min(ocv_use$DOSE_ROUND_ONE*1000, na.rm=T)
# summary(ocv_use$DOSE_REQUESTED*1000)
# hist(ocv_use$DOSE_ROUND_ONE*1000, breaks=1000)

# ts |> 
#   mutate(admin1 = ifelse(admin1 == "", NA, admin1), 
#          admin2 = ifelse(admin2 == "", NA, admin2),
#          admin3 = ifelse(admin3 == "", NA, admin3),
#          admin4 = ifelse(admin4 == "", NA, admin4),
#          admin5 = ifelse(admin5 == "", NA, admin5),) -> ts
length(unique(ts$id_outbreak))

ts |> filter(pop <= max_ocv_dose & pop >= min_ocv_dose) -> ts_pop 
 
ts_pop |> 
  group_by(id_outbreak) |> 
  summarize(pop = pop[1]) |> 
  pull(pop) |> 
  summary()

length(unique(ts_pop$id_outbreak))

# length(unique(ts$id_outbreak))
# ts_adm1 <- ts |> filter(!is.na(admin1))
# length(unique(ts_adm1$id_outbreak))
# ts_adm12 <- ts |> filter(!is.na(admin1) & !is.na(admin2))
# length(unique(ts_adm12$id_outbreak))
# 
# ts_adm1 |> 
#   group_by(id_outbreak) |> 
#   summarize(pop = pop[1]) |> 
#   pull(pop) |> 
#   summary()
# 
# ts_adm1 <- ts |> filter(spatial_scale == "admin1")
# ts_adm1 |> 
#   group_by(id_outbreak) |> 
#   summarize(pop = pop[1]) |> 
#   pull(pop) |> 
#   summary()
# 
# ts_adm2 <- ts |> filter(spatial_scale == "admin2")
# ts_adm2 |> 
#   group_by(id_outbreak) |> 
#   summarize(pop = pop[1]) |> 
#   pull(pop) |> 
#   summary()
```


### 2. Exclude outbreaks affected by OCVs
```{r}
# exclude outbreaks that were affected by OCV campaigns using IDs 
# length(unique(ts_pop$id_outbreak))
# ts_pop_ocv <- ts_pop[!(ts_pop$id_outbreak %in% docv$id_outbreak),]
ts_pop_ocv <- ts_pop

length(unique(ts_pop_ocv$id_outbreak))
sum(unique(ts_pop_ocv$id_outbreak) %in% ids_conf)

sum(!unique(ts_pop_ocv$id_outbreak) %in% ds$id_outbreak)
```

### 2.1 Exclude outbreaks with no information on deaths

For vaccine impact, we conduct analyses for the outbreaks even if information
on death is not available.
```{r}
length(unique(ts_pop_ocv$id_outbreak))
ts_pop_ocv |> 
  group_by(id_outbreak) |> 
  mutate(total_deaths = sum(deaths)) |>
  filter(total_deaths >= 0) -> ts_pop_ocv_death # for YLD and DALY

ts_pop_ocv -> ts_pop_ocv_death

length(unique(ts_pop_ocv_death$id_outbreak))
```

### 3 Set aside the culture-confirmed outbreaks

```{r}
length(unique(ts_pop_ocv$id_outbreak))

ts_pop_ocv |> 
  group_by(id_outbreak) |> 
  mutate(total_confirmed_cases = sum(c_ch)) |>
  filter(total_confirmed_cases > 0) |> 
  ungroup() -> ts_pop_ocv_confirmed # for YLD and DALY 

length(unique(ts_pop_ocv_confirmed$id_outbreak))
```


### Outbreak data for modeling

Combine the time-series data with the summary data, which has information on
attack rate, case fatality ratio, outbreak_pop, rural_urban, etc.

```{r}
# unfortunately, three outbreaks from time series don't have corresponding summary
ts_pop_ocv <- ts
all(ts_pop_ocv$id_outbreak %in% ds$id_outbreak)
unique(ts_pop_ocv[!(ts_pop_ocv$id_outbreak %in% ds$id_outbreak),]$id_outbreak)

ts_pop_ocv_ds <- ts_pop_ocv[(ts_pop_ocv$id_outbreak %in% ds$id_outbreak),]
  
  
ds_ <- ds[, c("ar_per_10e3", "cfr_per_100percent", 
              "outbreak_pop", "rural_urban", "id_outbreak")]
names(ds_) <- c("attack_rate", "cfr", "population", "rural_urban", 
                "id_outbreak")


outbreak_data_modeling <- left_join(ts_pop_ocv_ds, ds_, by="id_outbreak")

outbreak_data_modeling |> 
  group_by(id_outbreak) |> 
  summarize(pop = pop[1], 
            population = population[1]) -> d

# for some outbreaks, population size is not the same between ts and ds
# in which case, we will use values from ts as ds has NA's as well.
# outbreak_data_modeling |> 
#  group_by(id_outbreak) |> 
#  summarize(pop = pop, 
#            population = population) -> d
# summary(abs(d$pop - d$population))
# which(abs(d$pop - d$population) > 1000)

data.table::fwrite(outbreak_data_modeling,
                   paste0("outputs/outbreak_data_", tstamp(), ".csv"))

# ts_test <- ts_test |>
#   mutate(outbk_yesno = ifelse(s_ch / population * 1e5 > threshold, 1, 0))

# parameter values to explore
vacc_covs <- seq(0.6, 0.95, by=0.05)
vacc_weeks <- 1:20
case_triggers <- c(5, 10, 20, 30, 50, 70, 100, 150, 250, 400, 600, 1000)
# to save and run by batch (batch_size = parameter sample size to go through all outbreaks, vaccination weeks, and vaccination coverage rates)
batch_size <- 4
```


Data of modeling the OCV use during outbreaks

```{r, eval=F}
library(googlesheets4)
#Read google sheets data into R
# outbreak modeling papers
review1 <- read_sheet('https://docs.google.com/spreadsheets/d/1ZN6M4tx21-8BSRMZ692MOU3M5yDYK7bLqRbYGVOZGtQ/edit#gid=959832095', sheet="Outbreaks")
review2 <- read_sheet('https://docs.google.com/spreadsheets/d/1ZN6M4tx21-8BSRMZ692MOU3M5yDYK7bLqRbYGVOZGtQ/edit?gid=1840356702#gid=1840356702', sheet="Outbreaks 2")
outbks2 <- filter(review2, `Defined outbreak vs surveillance` == "Outbreak")

ocv_use <- read_sheet('https://docs.google.com/spreadsheets/d/1g9Ddrp0Ft14BtmAlZ-j7r8YCEAddro6H2jcmLkn7u84/edit?gid=0#gid=0', 
                      sheet='Sheet1',
                      col_types = "cccDDDDDDDdddddddccccc")

# Infer population size by the number of vaccine doses  
summary(ocv_use$DOSE_ROUND_ONE*1000)
summary(ocv_use$DOSE_REQUESTED*1000)
hist(ocv_use$DOSE_ROUND_ONE*1000, breaks=10)
```

## Read in data sets
```{r}
outbreak_data_all <- fread("outputs/outbreak_data_all_20241024.csv")
outbreak_data <- fread("outputs/outbreak_data_20241024.csv") # for vaccine impact modeling
save_plot <- FALSE
library(googlesheets4)
# OCV use during campaigns
ocv_use <- read_sheet('https://docs.google.com/spreadsheets/d/1g9Ddrp0Ft14BtmAlZ-j7r8YCEAddro6H2jcmLkn7u84/edit?gid=0#gid=0', sheet='Sheet1',
                      col_types = "cccDDDDDDDdddddddccccc")
delay_req_R1 <- as.integer(ocv_use$DATE_ROUND_ONE - ocv_use$DATE_REQUEST)
# summary(delay_req_R1)
hist(ocv_use$DOSE_DELIVERED_ROUND_ONE)
summary(ocv_use$DOSE_DELIVERED_ROUND_ONE)

cutoffs <- quantile(ocv_use$DOSE_DELIVERED_ROUND_ONE, probs=seq(0, 1, 0.1), na.rm=TRUE)

grp_levels <- c(paste0(cutoffs[1],"-", number(cutoffs[2], big.mark = ",")),
             paste0(cutoffs[2], "-", number(cutoffs[3], big.mark = ",")),
             paste0(cutoffs[3], "-", number(cutoffs[4], big.mark = ",")),
             paste0(cutoffs[4], "-", number(cutoffs[5], big.mark = ",")),
             paste0(cutoffs[5], "-", number(cutoffs[6], big.mark = ",")),
             paste0(cutoffs[6], "-", number(cutoffs[7], big.mark = ",")),
             paste0(cutoffs[7], "-", number(cutoffs[8], big.mark = ",")),
             paste0(cutoffs[8], "-", number(cutoffs[9], big.mark = ",")),
             paste0(cutoffs[9], "-", number(cutoffs[10], big.mark = ",")))

svim$pop_size_grp <- ifelse(svim$s_ch_tot < cutoffs[2], 
  paste0(cutoffs[1],"-", cutoffs[2]),
  ifelse(svim$population < cutoffs[3], 
  paste0(cutoffs[2],"-", cutoffs[3]),
  ifelse(svim$population < cutoffs[4], 
  paste0(cutoffs[3],"-", cutoffs[4]),
  ifelse(svim$population < cutoffs[5], 
  paste0(cutoffs[4],"-", cutoffs[5]),
  ifelse(svim$population < cutoffs[6], 
  paste0(cutoffs[5],"-", cutoffs[6]),
  ifelse(svim$population < cutoffs[7], 
  paste0(cutoffs[6],"-", cutoffs[7]),
  ifelse(svim$population < cutoffs[8], 
  paste0(cutoffs[7],"-", cutoffs[8]),
  ifelse(svim$population < cutoffs[9], 
  paste0(cutoffs[8],"-", cutoffs[9]), 
  paste0(cutoffs[9],"-", number(cutoffs[10], big.mark = ","))))))))))

svim$pop_size_grp <- factor(svim$pop_size_grp, levels=grp_levels)
```

Resampling approach

We sample outbreaks based on the size of the population based on the empirical 
ditribution

```{r}
n_sample <- 1000
obs_ocv_dose <- ocv_use$DOSE_DELIVERED_ROUND_ONE[!is.na(ocv_use$DOSE_DELIVERED_ROUND_ONE)]
min_dose <- min(obs_ocv_dose)
max_dose <- max(obs_ocv_dose)

df <- data.frame(num_dose = obs_ocv_dose)
linspace <- seq(min_dose, max_dose, length.out=101)
df$grp <- NA
# Integer Division:
df$grp <- df$num_dose %/% 49.899 + 1 

quantile(ocv_use$DOSE_DELIVERED_ROUND_ONE, probs=seq(0, 1, 0.1), na.rm=TRUE)

svim$pop_size_grp <- (svim$pop/1000) %/% 49.899 + 1 
  
n <- length(historical_changes)
prob_vector <- rep(1/n, n)
resamples <- sample(x=historical_changes, size = num_samples, replace = TRUE, prob = prob_vector)
svim |> 
  
names(outbreak_data)  
svim <- left_join(svim, outbreak_data[,c("id_outbreak", "pop")], by="id_outbreak")# future stuff
```


### Summary statistics
```{r, eval=FALSE}
length(unique(outbreak_data_all$id_outbreak))
length(unique(outbreak_data$id_outbreak))

summary(outbreak_data_all$pop)
filter(outbreak_data_all, pop > 5000000) |> pull(id_outbreak) |> unique() |> length()

outbreak_data_all[outbreak_data_all$pop > 1e7,]
outbreak_data_all |> 
  mutate(admin1 = ifelse(admin1 == "", NA, admin1), 
         admin2 = ifelse(admin2 == "", NA, admin2),
         admin3 = ifelse(admin3 == "", NA, admin3),
         admin4 = ifelse(admin4 == "", NA, admin4),
         admin5 = ifelse(admin5 == "", NA, admin5),) -> outbreak_data_all

# outbreak_data_all |> 
#   mutate(admin1 = gsub("", NA, admin1), 
#          admin2 = gsub("", NA, admin2),
#          admin3 = gsub("", NA, admin3),
#          admin4 = gsub("", NA, admin4),
#          admin5 = gsub("", NA, admin5)) -> outbreak_data_all

summary(outbreak_data_all$pop)
summary(outbreak_data_all[!is.na(outbreak_data_all$admin1),]$pop)

```


```{r}
# svim <- fread("outputs/mean_svim_vacc_week_20241021.csv")
svim <- fread("outputs/svim_res_vacc_week_20241029.csv")
str <- names(svim)
names(svim) <- gsub("mean_", "", str)
# for cost-associated analyses
# bod = burden of disease
svim_cost <- fread("outputs/svim_bod_res_vacc_week_20241029.csv")
str <- names(svim_cost)
names(svim_cost) <- gsub("mean_", "", str)
# svim <- clean_names(fread("outputs/mean_svim_vacc_week_20241010.csv"))
# svim$vacc_week <- svim$vacc_week
# This is needed because outbreak data has been updated
# svim <- svim[svim$id_outbreak %in% outbreak_data$id_outbreak,]

svim_conf <- svim[svim$id_outbreak %in% ids_conf,]
week_vacc_start <- round((median(delay_req_R1, na.rm=T) + 10)/7)
svim |> filter(vacc_cov == 0.8, vacc_week == 7) -> svim_v9w4
svim |> filter(vacc_cov == 0.8, vacc_week == 7) -> svim_v8w7

library(moments)
summary(round(svim_v9w4$s_ch_tot))
sd(svim_v9w4$s_ch_tot)
skewness(svim_v9w4$s_ch_tot)

summary(svim_v9w4$s_ch_averted_tot)
sd(svim_v9w4$s_ch_averted_tot)
skewness(svim_v9w4$s_ch_averted_tot)

```

```{r}
outbreak_data |> 
  group_by(id_outbreak) |> 
  summarize(duration = max(tr) - min(tl),
            s_ch_tot = sum(s_ch),
            death_tot = sum(deaths), 
            cfr = death_tot/s_ch_tot) -> d

plot(d$duration, d$cfr)

fit <- lm(cfr ~ duration, data=d)
newd <- data.frame(duration=0:1000)
pred <- predict(fit, newdata=newd)

lines(newd$duration, coef(fit)[1] + coef(fit)[2] * newd$duration, col=2)
```

