---
title: "Hierarchical beta regression for the indirect vaccine effectiveness"
format: html
editor_options: 
  chunk_output_type: console
---


##  IVE data

Load packages and source utility functions

```{r}
library(rstan)
library(cmdstanr)
source("utils.R")
```

Incidence rate based on the study by Ali (2005) Matlab, Bangladesh
```{r, eval=FALSE}
# Individuals targeted for the trial, 89 596
# this is the first year of 4,035 in surveillance and therefore
# vaccine waning is ignored
vc_cutoff <- c(0, 0.28, 0.35, 0.40, 0.50, 0.65) # vaccine coverage cutoff, Table 1
vc_mid <- vc_cutoff[1:5] + diff(vc_cutoff)/2 # mid-point used a coverage
ir <- c(7.01, 5.87, 4.72, 4.65, 1.47) # incidence rate
# baseline incidence rate (i.e., incidence rate at zero vaccination rate)
dat <- data.frame(ir = ir, vc = vc_mid, region = "Matlab")
```

Baseline incidence rate, incidence rate at vaccine coverage of 0, was adopted based on the study
by Longini et al.(2007) PLoS Med (Fig. 3A), which is around 11.2 per 1,000 person-years. Using the following equation, we add a column on the indirect vaccine effectiveness (IVE),

$$\text{VE}_{I,vc} = 1 - \left( \frac{\textrm{IR}_{vc}}{\textrm{IR}_{base}}\right)$$

```{r}
ir_baseline <- 11.2 # Longini et al.(2007) PLoS Med (Fig. 3A)
dat$ive <- 1 - dat$ir / ir_baseline
```

Longini et al. (2007) study provides IVE from their individual-based model.
```{r}
ive_longini <- 
  data.frame(vc = seq(0.1,0.9,by=0.2),
             ive = c(0.3,0.7,0.89,0.97,0.99),
             lb = c(-0.39,0.31,0.72,0.91,0.98),
             ub = c(0.83,0.93,0.98,0.99,1))
```

### Fitting

```{r, eval=F}
N <- nrow(dat)
X <- cbind(1, logit(dat$vc))
K <- ncol(X)
y <- dat$ive

fit <- stan(file = "ir_ive_beta_matlab.stan",
            data = list(N=N, X=X, K=K, y=y),
            iter = 1e4, cores = 4,
            control = list(adapt_delta=0.9))

# saveRDS(fit, paste0("outputs/fit_ive_matlab_", tstamp(), ".rds"))
print(fit, pars=c("beta","phi","lp__"), probs=c(0.025, 0.5, 0.975))
```

### Posterior predictive check

```{r, eval=FALSE}
# fit <- read_rds("outputs/fit_ive_matlab_20241008.rds")
# print(fit, pars=c("beta","phi","lp__"), probs=c(0.025, 0.5, 0.975))
yrep <- as.data.frame(rstan::extract(fit)$yrep)
pr <- c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975) 
pred <- as.data.frame(t(apply(yrep, 2, quantile, probs=pr, na.rm=T)))
pred$x <- expit(rstan::extract(fit)$new_X[1,,2])
# pred$x <- extract(fit)$new_X[1,,2]
# pred$x <- sort(unique(extract(fit1)$new_X[1,,2]))

# pred$`95%` <- ifelse(pred$`95%` > 1, 1, pred$`95%`)
# pred$`5%` <- ifelse(pred$`5%` < 0, 0, pred$`5%`)
# pred$`97.5%` <- ifelse(pred$`97.5%` > 1, 1, pred$`97.5%`)
# pred$`2.5%` <- ifelse(pred$`2.5%` < 0, 0, pred$`2.5%`)
# pred$`75%` <- ifelse(pred$`75%` > 1, 1, pred$`75%`)
# pred$`25%` <- ifelse(pred$`25%` < 0, 0, pred$`25%`)

ggplot(ive_longini)+
  geom_point(aes(vc, ive, color="Dynamic model"), shape=15) +
  geom_linerange(aes(vc, ymin=lb, ymax=ub, color="Dynamic model"))+
  geom_point(data=dat, aes(vc, ive, color="Data"), shape=17)+
  geom_linerange(data=pred, aes(x, ymin=`5%`, ymax=`95%`, color="Beta regression"), linewidth=0.4)+
  geom_linerange(data=pred, aes(x, ymin=`25%`, ymax=`75%`, color="Beta regression"), linewidth=0.8)+
  geom_point(data=pred, aes(x, `50%`, color="Beta regression"))+
  scale_y_continuous(limits=c(0,1))+
  scale_color_manual("", values=c("Dynamic model"="firebrick",
                                  "Data"="steelblue",
                                  "Beta regression"="black"))+  
  theme_light()+
  # labs(x="Vaccine coverage", y="Indirect vaccine effectiveness",
  #      color="", title="Posterior predictive values")+
    labs(x="Vaccine coverage", y="Indirect vaccine effectiveness",
       color="")+
  theme(legend.position="top") -> plt

plt
fac <- 1.4
ggsave(paste0("plots/ppv_ive_matlab_", tstamp(), ".pdf"), plt, height=2.7*fac, width=3.4*fac,
        units="in", device=cairo_pdf)
```
