---
title: "Parameters"
format: html
editor_options: 
  chunk_output_type: console
---

## Parameter samples for simulation
```{r}
source("utils.R")
dat_params <- data.table::fread("data/parameters.csv")
nruns <- 200
params <- draw_parameter_samples(nruns=nruns, parameter_data=dat_params)
data.table::fwrite(params, paste0("outputs/params_", tstamp(), ".csv"))
summary(params)
summary(params$dur_campaign)
```


Vaccine coverage data from Table 2 in Pezzoli et al. 2020
We chose 0.6, 0.75 and 0.9 as the three vaccine coverage rates
```{r}
vacc_cov_data <- 
  data.frame(setting = c("Endemic","Humanitarian crisis","Outbreak"),
           admin_cov_first_dose = c(93.6, 82.9, 98.8)/100,
           admin_cov_second_dose = c(92.4, 82.0, 93.0)/100, 
           estimate_cov_at_least_one = c(85.1, 73.2, 86.3)/100)

vacc_cov_data |> 
  filter(setting != "Endemic") |>
  group_by(setting) |>
  summarize(frac_effect = estimate_cov_at_least_one/(1-(1-admin_cov_first_dose)*(1-admin_cov_second_dose)),
            vacc_cov_r1 = admin_cov_first_dose*frac_effect) -> vacc_cov_r1
vacc_cov_r1
# we take the vaccine coverage for round 1 for Outbreak rather than Humanitarian crises 

```



## Compute the impact of vaccine

The impact expected from the pre-emptive vaccination is computed assuming all vaccine recipients experience the direct vaccine efficacy, DVE. The proportion of the population, age group $i$, at vaccination coverage proportion, $pi$, is given as belows: 

$$P_{i,\pi} = 1-\left[\pi(1-\text{DVE}_i)(1-\text{IVE}) + (1-\pi)(1-\text{IVE})\right]$$

## Compare IVE from the indirect effectiveness


