---
title: "Parameters"
format: html
editor_options: 
  chunk_output_type: console
---

## Beta regression for the indirect vaccine effectiveness

### IVE data

Load packages and source utility functions

```{r}
library(rstan)
library(cmdstanr)
source("utils.R")
```

Incidence rate based on the study by Ali (2005) Matlab, Bangladesh
```{r, eval=FALSE}
# Individuals targeted for the trial, 89 596
# this is the first year of 4,035 in surveillance and therefore
# vaccine waning is ignored
vc_cutoff <- c(0, 0.28, 0.35, 0.40, 0.50, 0.65) # vaccine coverage cutoff, Table 1
vc_mid <- vc_cutoff[1:5] + diff(vc_cutoff)/2 # mid-point used a coverage
ir <- c(7.01, 5.87, 4.72, 4.65, 1.47) # incidence rate
# baseline incidence rate (i.e., incidence rate at zero vaccination rate)
dat <- data.frame(ir = ir, vc = vc_mid, region = "Matlab")
```

Baseline incidence rate, incidence rate at vaccine coverage of 0, was adopted based on the study
by Longini et al.(2007) PLoS Med (Fig. 3A), which is around 11.2 per 1,000 person-years. Using the following equation, we add a column on the indirect vaccine effectiveness (IVE),

$$\text{VE}_{I,vc} = 1 - \left( \frac{\textrm{IR}_{vc}}{\textrm{IR}_{base}}\right)$$

```{r}
ir_baseline <- 11.2 # Longini et al.(2007) PLoS Med (Fig. 3A)
dat$ive <- 1 - dat$ir / ir_baseline
```

Longini et al. (2007) study provides IVE from their indivdiual-based model.
```{r}
ive_longini <- 
  data.frame(vc = seq(0.1,0.9,by=0.2),
             ive = c(0.3,0.7,0.89,0.97,0.99),
             lb = c(-0.39,0.31,0.72,0.91,0.98),
             ub = c(0.83,0.93,0.98,0.99,1))
```

### Fitting

```{r, eval=F}
N <- nrow(dat)
X <- cbind(1, logit(dat$vc))
K <- ncol(X)
y <- dat$ive

fit <- stan(file = "ir_ive_beta_matlab.stan",
            data = list(N=N, X=X, K=K, y=y),
            iter = 1e4, cores = 4,
            control = list(adapt_delta=0.9))

# saveRDS(fit, paste0("outputs/fit_ive_matlab_", tstamp(), ".rds"))
print(fit, pars=c("beta","phi","lp__"), probs=c(0.025, 0.5, 0.975))
```

### Posterior predictive check

```{r, eval=FALSE}
# fit <- read_rds("outputs/fit_ive_matlab_20241008.rds")
# print(fit, pars=c("beta","phi","lp__"), probs=c(0.025, 0.5, 0.975))
yrep <- as.data.frame(rstan::extract(fit)$yrep)
pr <- c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975) 
pred <- as.data.frame(t(apply(yrep, 2, quantile, probs=pr, na.rm=T)))
pred$x <- expit(rstan::extract(fit)$new_X[1,,2])

ggplot(ive_longini)+
  geom_point(aes(vc, ive, color="Dynamic model"), shape=15) +
  geom_linerange(aes(vc, ymin=lb, ymax=ub, color="Dynamic model"))+
  geom_point(data=dat, aes(vc, ive, color="Data"), shape=17)+
  geom_linerange(data=pred, aes(x, ymin=`5%`, ymax=`95%`, 
                                color="Beta regression"), linewidth=0.4)+
  geom_linerange(data=pred, aes(x, ymin=`25%`, ymax=`75%`, 
                                color="Beta regression"), linewidth=0.8)+
  geom_point(data=pred, aes(x, `50%`, color="Beta regression"))+
  scale_y_continuous(limits=c(0,1))+
  scale_color_manual("", values=c("Dynamic model"="firebrick",
                                  "Data"="steelblue",
                                  "Beta regression"="black"))+  
  theme_light()+
  labs(x="Vaccine coverage", y="Indirect vaccine effectiveness", color="")+
  theme(legend.position="top") -> plt

plt
fac <- 1.4
ggsave(paste0("plots/ppv_ive_matlab_", tstamp(), ".pdf"), plt, 
       height=2.7*fac, width=3.4*fac, units="in", device=cairo_pdf)
```

## Parameter samples for simulation
```{r}
dat_params <- data.table::fread("data/parameters.csv")
nruns <- 200
params <- draw_parameter_samples(nruns=nruns, parameter_data=dat_params)
data.table::fwrite(params, paste0("outputs/params_", tstamp(), ".csv"))
summary(params)
summary(params$dur_campaign)
```


Vaccine coverage data from Table 2 in Pezzoli et al. 2020
We chose 0.6, 0.75 and 0.9 as the three vaccine coverage rates
```{r}
vac_cov_data <- 
  data.frame(setting = c("Endemic","Humanitarian crisis","Outbreak"),
           admin_cov_first_dose = c(93.6, 82.9, 98.8)/100,
           admin_cov_second_dose = c(92.4, 82.0, 93.0)/100, 
           estimate_cov_at_least_one = c(85.1, 73.2, 86.3)/100)

vac_cov_data |> 
  filter(setting != "Endemic") |>
  group_by(setting) |>
  summarize(frac_effect = estimate_cov_at_least_one/(1-(1-admin_cov_first_dose)*(1-admin_cov_second_dose)),
            vacc_cov_r1 = admin_cov_first_dose*frac_effect) |>
  summarize(mean_vacc_cov_r1 = mean(vacc_cov_r1))
```



## Compute the impact of vaccine

The impact expected from the pre-emptive vaccination is computed assuming all vaccine recipients experience the direct vaccine efficacy, DVE. The proportion of the population, age group $i$, at vaccination coverage proportion, $pi$, is given as belows: 

$$P_{i,\pi} = 1-\left[\pi(1-\text{DVE}_i)(1-\text{IVE}) + (1-\pi)(1-\text{IVE})\right]$$

## Compare IVE from the indirect effectiveness


