---
title: "Parameter Preparation for Cholera Vaccine Impact Analysis"
author: "Your Name"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
editor_options: 
  chunk_output_type: console
---

# Introduction

This document handles the preparation of parameters for the cholera vaccine impact simulation. 
It generates parameter samples for multiple simulation runs, defines vaccine coverage values, 
and calculates impact metrics that will be used in the main analysis pipeline.


# Setup

## Load Packages and Utility Functions

```{r setup}
#| warning: false
#| message: false

# Core packages
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(ggplot2)
library(ggridges)
library(patchwork)
library(conflicted)
library(scales)
library(viridis)
library(doParallel)  # For parallel processing
library(MASS)

# Resolve conflicts in favor of dplyr functions
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflicts_prefer(data.table::first)
conflicts_prefer(data.table::year)

# Load utility functions
source("utils.R")  # Custom utility functions for vaccine impact calculations
```

# Parameter Sample Generation

We draw parameter samples from distributions to account for uncertainty in our model parameters.

```{r setup}
# Load required utility functions
source("utils.R")

# Load parameter definitions from file
dat_params <- fread("data/parameters.csv")

# Generate multiple parameter sets for simulation runs
nruns <- 200  # Number of simulation runs
params <- draw_parameter_samples(nruns = nruns, parameter_data = dat_params)

# Save parameter samples for later use
# fwrite(params, paste0("outputs/params_", tstamp(), ".csv"))

# Display summary statistics for the parameter distributions
summary(params)

# Check distribution of campaign duration parameter
summary(params$dur_campaign)
```

# Vaccine Coverage Configuration

We define vaccine coverage values based on observed field data from Pezzoli et al. 2020.

```{r coverage}
# Vaccine coverage data from Table 2 in Pezzoli et al. 2020
# We choose 0.6, 0.75 and 0.9 as representative vaccine coverage rates
vacc_cov_data <- 
  data.frame(setting = c("Endemic", "Humanitarian crisis", "Outbreak"),
           admin_cov_first_dose = c(93.6, 82.9, 98.8)/100,  # Convert percentages to proportions
           admin_cov_second_dose = c(92.4, 82.0, 93.0)/100, 
           estimate_cov_at_least_one = c(85.1, 73.2, 86.3)/100)

# Calculate effective first-round coverage for different settings
# We adjust for the relationship between administrative coverage and survey-estimated coverage
vacc_cov_data |> 
  filter(setting != "Endemic") |>  # Exclude endemic settings for emergency response
  group_by(setting) |>
  summarize(
    # Calculate fraction of people reached by either dose
    frac_effect = estimate_cov_at_least_one / 
                 (1 - (1 - admin_cov_first_dose) * (1 - admin_cov_second_dose)),
    # Apply this fraction to first dose coverage
    vacc_cov_r1 = admin_cov_first_dose * frac_effect
  ) -> vacc_cov_r1

# We use vaccine coverage for outbreak settings rather than humanitarian crises
print(vacc_cov_r1)
```

# Vaccine Impact Calculation

This section describes how we calculate the expected vaccine impact from the combination of direct 
and indirect vaccine effectiveness.

The impact expected from the pre-emptive vaccination is computed assuming all vaccine recipients experience the direct vaccine efficacy, DVE. The proportion of the population, age group $i$, at vaccination coverage proportion, $pi$, is given as belows: 

$$P_{i,\pi} = 1-\left[\pi(1-\text{DVE}_i)(1-\text{IVE}) + (1-\pi)(1-\text{IVE})\right]$$

```{r impact}
# The impact expected from vaccination is computed assuming all vaccine recipients 
# experience the direct vaccine efficacy (DVE), plus indirect vaccine effectiveness (IVE).

# For a population with age group i, at vaccination coverage proportion π, 
# the overall protective effect is given by:
# P_{i,π} = 1-[π(1-DVE_i)(1-IVE) + (1-π)(1-IVE)]

# Where:
# - DVE_i is the direct vaccine efficacy for age group i
# - IVE is the indirect vaccine effectiveness at coverage level π
# - The first term represents vaccinated people getting protection from both direct and indirect effects
# - The second term represents unvaccinated people getting protection only from indirect effects
```

# Indirect Vaccine Effectiveness Model

We load and process the indirect vaccine effectiveness model results from a previously-run
Bayesian beta regression.

```{r ive}
# Load IVE model posterior predictive samples
# This comes from a Bayesian beta regression model of indirect effectiveness vs. coverage
ive_fit <- read_rds("outputs/ive_yrep_20241010.rds")
ive_data <- ive_fit$yrep
rm(ive_fit)  # Remove large object from memory to save space

# The ive_data matrix contains posterior predictive samples of IVE for different coverage levels
# Rows are simulation runs, columns are coverage levels from 0 to 1 in small increments
# We will use these values to model how indirect protection varies with coverage
```

# Compare IVE from Indirect Effectiveness Studies

```{r ive-compare}
# The Longini et al. (2007) study provides IVE estimates from their individual-based model
ive_longini <- 
  data.frame(vc = seq(0.1, 0.9, by = 0.2),
             ive = c(0.3, 0.7, 0.89, 0.97, 0.99),
             lb = c(-0.39, 0.31, 0.72, 0.91, 0.98),
             ub = c(0.83, 0.93, 0.98, 0.99, 1))

# We can also analyze empirical data from field studies
# Incidence rate based on the study by Ali (2005) in Matlab, Bangladesh
# Individuals targeted for the trial: 89,596
ir_baseline <- 11.2  # Baseline incidence rate per 1000 person-years
vc_cutoff <- c(0, 0.28, 0.35, 0.40, 0.50, 0.65)  # Vaccine coverage cutoff points
vc_mid <- vc_cutoff[1:5] + diff(vc_cutoff)/2     # Mid-point used as coverage value
ir <- c(7.01, 5.87, 4.72, 4.65, 1.47)            # Observed incidence rates

# Create data frame of empirical observations
dat <- data.frame(ir = ir, vc = vc_mid, region = "Matlab")

# Calculate indirect effectiveness from incidence rates
dat$ive <- 1 - dat$ir / ir_baseline

# Visualize relationship between coverage and indirect effectiveness
# ggplot() +
#   geom_point(data = ive_longini, aes(vc, ive, color = "Model"), shape = 15) +
#   geom_linerange(data = ive_longini, aes(vc, ymin = lb, ymax = ub, color = "Model")) +
#   geom_point(data = dat, aes(vc, ive, color = "Empirical"), shape = 17) +
#   scale_y_continuous(limits = c(0, 1)) +
#   scale_color_manual("Data source", values = c("Model" = "firebrick", "Empirical" = "steelblue")) +
#   labs(x = "Vaccine coverage", y = "Indirect vaccine effectiveness") +
#   theme_light()
```

## Age distribution of cholera cases: average age of infection

```{r}
# Eastern Mediterranean Health Journal Volume 30 2024  
# Retrospective analysis of a large-scale cholera outbreak in Sudan
# Table 1.
N <- 983 + 985
mean <- (30.9*983+29.5*985)/(983+985)

# BMC Infectious Diseases volume 19, Article number: 4 (2019)
# Epidemiological description of a protracted cholera outbreak in Tonj East and Tonj North counties, former Warrap State, South Sudan, May-Oct 2017

N <- c(N, 1451)
mean <- c(mean, 24)

# Descriptive characterization of the 2010 cholera outbreak in Nigeria
# BMC Public Health volume 14, Article number: 1167 (2014) 

N <- c(N, 20688)
median <- 12
Q1 <- 3
Q3 <- 17
IQR <- Q3 - Q1
# Quick estimation

# Simple average of the quartiles
mean_est1 <- (Q1 + median + Q3) / 3

# Weighted toward median
mean_est2 <- (Q1 + 2*median + Q3) / 4
mean_est3 <- (Q1 + Q3) / 2

# Estimate mean as integral approximation
mean_est4 <- Q1*0.25 + median*0.25 + Q3*0.25 + 
            (median-Q1)*0.125 + (Q3-median)*0.125

assumed_min <- Q1 - 1.5*IQR
assumed_max <- Q3 + 1.5*IQR
mean_est5 <- (assumed_min + Q1 + median + Q3 + assumed_max) / 5

# For right-skewed: mean > median
# For left-skewed: mean < median
skewness_factor <- (Q3 - median) - (median - Q1)
mean_est6 <- median + skewness_factor * 0.1  # Empirical adjustment

mean <- c(mean, mean_est2) 

# Clin Infect Dis. 2024 Jul 12;79(Suppl 1):S8–S19. doi: 10.1093/cid/ciae236
# Retrospective Analysis of Cholera/Acute Watery Diarrhea Outbreaks in Ethiopia From 2001 To 2023: Incidence, Case Fatality Rate, and Seasonal and Multiyear Epidemic Patterns

# Age groups and case counts
age_groups <- data.frame(
  age_range = c("<5", "5-14", "15-44", "≥45"),
  lower_bound = c(0, 5, 15, 45),
  upper_bound = c(5, 14, 44, 75),  # Assumed 75 for ≥45 group
  midpoint = c(2.5, 9.5, 29.5, 60), # Assumed 60 for ≥45 group
  cases = c(11460, 18273, 53990, 16215)
)

total_cases <- sum(age_groups$cases)
weighted_sum <- sum(age_groups$midpoint * age_groups$cases)
weighted_mean <- weighted_sum / total_cases

N <- c(N, total_cases)
mean <- c(mean, weighted_mean)

# Bwire PLoS Negl Trop Dis. 2017 Mar 13;11(3):e0005407. doi: 10.1371/journal.pntd.0005407
# in the fishing villages of Uganda: 2011-2015
# test which of the options provided closest value
median <- 20
Q1 <- 7
Q3 <- 30
(median + Q1 + Q3)/3
(2*median + Q1 + Q3)/4
(Q1 + Q3) / 2
# Estimate mean as integral approximation
Q1*0.25 + median*0.25 + Q3*0.25 + (median-Q1)*0.125 + (Q3-median)*0.125
assumed_min <- Q1 - 1.5*IQR
assumed_max <- Q3 + 1.5*IQR
(assumed_min + Q1 + median + Q3 + assumed_max) / 5

N <- c(N, 1827)
mean <- c(mean, 21.2)

##
# Bwire. PLoS Negl Trop Dis. 2013 Dec 5;7(12):e2545. doi: 10.1371/journal.pntd.0002545
# The Burden of Cholera in Uganda
# District-specific data from the Uganda Ministry of Health, Health Management Information System disease surveillance database for the period 2005–2010.
age_groups <- data.frame(
  age_range = c("<1", "1-4", "5-14", "15-30", "31-59","60+"),
  lower_bound = c(0, 1, 5,  15, 31, 60),
  upper_bound = c(1, 5, 15, 31, 60, 75),  # Assumed 75 for ≥45 group
  midpoint = c(0.5, 3, 10, 23, 45.5, 67.5), # Assumed 60 for ≥45 group
  cases = c(5, 10, 17, 22, 31, 15)
)
total_cases <- sum(age_groups$cases)
weighted_sum <- sum(age_groups$midpoint * age_groups$cases)
weighted_mean <- weighted_sum / total_cases


N <- c(N, 2760)
mean <- c(mean, weighted_mean)

## Summary
N 
mean
paste0("weighted mean = ", round(sum(mean*N)/sum(N), digits=2))
```

# Conclusion

This document has prepared the parameter sets needed for the cholera vaccine impact simulation.
The parameters include:

1. Vaccine efficacy parameters (DVE)
2. Indirect vaccine effectiveness (IVE) estimates
3. Vaccine coverage targets
4. Campaign logistics parameters (duration, delay)

These will be used in the main simulation analysis to model the impact of different
vaccination strategies on cholera outbreaks.
