---
title: "Cholera Vaccine Impact Visualization and Tables"
author: "Jong-Hoon Kim"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
editor_options: 
  chunk_output_type: console
---

# Introduction

This document creates visualizations and tables to summarize the results from 
our cholera vaccine impact simulations. It processes results from different vaccination strategies (outbreak onset-triggered, case-triggered, and preemptive) and coverage levels, and produces figures and tables for the final report.

# Setup

```{r setup}
#| warning: false
#| message: false

# Load required packages
source("R/utils.R")  # utility functions
library(data.table)
library(tidyverse)
library(magrittr)
library(purrr)
library(broom)
library(effsize)
library(randomForest)
library(rstatix)
library(ggridges)
library(readxl)
library(janitor)
library(scales)
library(viridis)
library(sf)
library(metR)
library(googlesheets4)
library(MASS)
library(patchwork)
library(ggpattern)
library(gridExtra)
library(corrr)
library(RColorBrewer)
library(countrycode)
library(conflicted)
library(xdvir)

# Resolve package conflicts in favor of dplyr
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflicts_prefer(data.table::first)
conflicts_prefer(ggplot2::margin)
# Output control flags
save_plot <- FALSE
save_table <- FALSE

# Configure figure size settings
figure_size <- data.frame(
  journal = "Lancet",
  single = 85,   # Width for single-column figures (mm)
  double = 178   # Width for double-column figures (mm)
)
fgw1 <- figure_size[figure_size$journal == "Lancet",]$single
fgw2 <- figure_size[figure_size$journal == "Lancet",]$double

# Color palette for consistent visualization
set1 <- brewer.pal(9, "Set1")  # color sets
# my_discrete_colors <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00")

# Define standard probability values for quantiles
probs <- c(0.025, 0.25, 0.5, 0.75, 0.975)  # 2.5%, 25%, 50%, 75%, 97.5% quantiles

# Reference vaccination week
VW <- 9  # representative vaccination week from GTFCC data

# Sequence of vaccination weeks to evaluate
VW_seq <- seq(1, 15, 2)

# Representative vaccination coverage rate
VC <- 0.75  # 75% coverage

# Three representative vaccination coverage rates
VC3 <- c(0.6, 0.75, 0.9)  # 60%, 75%, 90%

# Thresholds for case-triggered vaccination to analyze
cts_sel <- c(1, seq(10, 100, 10))  # case triggers selected

# Latest date for which OCV campaign data is available
YEAR <- as.Date("2020-12-28")

# Load geographic data for mapping
shp <- st_read("data/paper/total_shp_0427.shp", quiet = TRUE)
africa <- st_read("data/paper/afr_g2014_2013_0.shp", quiet = TRUE)

# only a subset of sample runid's is used for plotting for convenience
set.seed(42)
sample_runids <- sample(1:200, size=30)

# Plot parameters
line_width_50 <- 1.2 
line_width_95 <- 0.5 
line_interval <- 0.7
point_size <- 0.5
x_size <- 1.5
x_stroke <- 1
```

# Load Data

```{r load-data}
#| warning: false
# Load outbreak data
# All outbreaks including those affected by OCVs
outbreak_data_all <- as.data.frame(fread("outputs/paper/outbreak_data_all_20250228.csv"))

# Outbreaks included in vaccine impact modeling (not affected by OCVs)
outbreak_data <- fread("outputs/paper/outbreak_data_20250228.csv") 

# outbreaks filtered
# This includes outbreaks that are finally analyzed and excludes
# outbreaks that were duplicated
outbreaks_filtered <-
  readRDS("outputs/paper/outbreak_summary_filtered_20250418.rds")

outbreak_data %<>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 
# outbreaks excluded
outbreaks_excluded <- read_rds("outputs/paper/outbreaks_excluded_20250612.rds")

# Load OCV campaign data
ocv_use <- fread("outputs/paper/ocv_use_GTFCC_data_20241120.csv")

# Calculate delay statistics between request and round 1 vaccination
delay_req_r1 <- as.integer(ocv_use$DATE_ROUND_ONE - ocv_use$DATE_REQUEST)
# Compute key percentiles of the delay distribution
ocv_delay <- quantile(delay_req_r1, probs = c(0.25, 0.5, 0.75, 0.9), na.rm = TRUE)

# Get historical OCV deployment data
dat <- clean_names(fread("outputs/paper/OCV_Use_GTFCC_Dashboard.csv"))
dat |> 
  filter(shipped == "Shipped", type == "Reactive") |> 
  group_by(year) |> 
  summarize(num_ocv = sum(dose)) |>
  setDT() -> ocv_year

ocv_year$cum_num_ocv <- cumsum(ocv_year$num_ocv)

# Load OCV campaign dataset
docv_ <- clean_names(fread("data/paper/outbreak_files_Oct2024/ocv_long_dataset.csv"))
docv <- docv_[(ocv_before_1 == TRUE | ocv_before_2 == TRUE | 
               ocv_before_3 == TRUE | ocv_during == TRUE)]

# Load model parameters
dis_params <- as.data.frame(data.table::fread("data/paper/parameters.csv"))

# Load IVE posterior predictive values from Bayesian beta regression
# ive_fit <- read_rds("outputs/ive_yrep_20241010.rds")
ive_fit <- readRDS("outputs/paper/ive_yhat_yrep_20250919.rds")
ive_data <- ive_fit$yrep
rm(ive_fit)  # Free memory by removing large object

# Load outbreak summary data
ds <- clean_names(as.data.frame(fread("data/paper/outbreak_files_Oct2024/outbreak_summary.csv")))
message(paste0(length(unique(ds$country)), " countries"))

# Create unique outbreak identifiers
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# Join outbreak data with summary information
outbk_data_summary <- 
  left_join(
    filter(outbreak_data, week == 1), 
    ds[, c("outbreak_pop", "total_s_ch", "total_c_ch", "total_deaths", "duration_weeks", 
           "peak_week_incidence", "time_to_peak_week", "peak_week_incidence_per_100percent", 
           "peak_week_incidence_per_10e3", "mean_r0_1st_week", "area_per_1km2", 
           "rural_urban", "ar_per_10e3", "cfr_per_100percent", "id_outbreak")], 
    by = "id_outbreak"
  )

outbk_data_summary %<>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 
# Summarize attack rates and population sizes
summary(outbk_data_summary$ar_per_10e3)
summary(outbk_data_summary$outbreak_pop)
```

# Recent cholera outbreak analysis

This section analyzes recent cholera data from WHO Global Cholera Dashboard.

```{r eval=FALSE recent-outbreaks}
# Load recent cholera data
c2023 <- fread("data/cholera_adm0_public_2023.csv")
c2024 <- fread("data/cholera_adm0_public_2024.csv")
c2023$year <- 2023
c2024$year <- 2024
chol <- rbind(c2023, c2024)

# Analyze 2023 data by WHO region
c2023 |> 
  group_by(year, who_region) |>
  summarize(case_region_total = sum(case_total), 
            death_region_total = sum(death_total)) |> 
  mutate(case_pct = 100 * case_region_total / sum(case_region_total),
         death_pct = 100 * death_region_total / sum(death_region_total),
         case_tot = sum(case_region_total),
         death_tot = sum(death_region_total))

# Analyze 2024 data by WHO region
c2024 |> 
  group_by(year, who_region) |>
  summarize(case_region_total = sum(case_total), 
            death_region_total = sum(death_total)) |> 
  mutate(case_pct = 100 * case_region_total / sum(case_region_total),
         death_pct = 100 * death_region_total / sum(death_region_total),
         case_tot = sum(case_region_total),
         death_tot = sum(death_region_total))

# Analyze all data by WHO region
chol |> 
  group_by(who_region) |>
  summarize(case_region_total = sum(case_total), 
            death_region_total = sum(death_total)) |> 
  mutate(case_pct = 100 * case_region_total / sum(case_region_total),
         death_pct = 100 * death_region_total / sum(death_region_total),
         case_tot = sum(case_region_total),
         death_tot = sum(death_region_total))
```

# Load Vaccine Impact Simulation Results

```{r load-results}
# Outbreak onset-triggered vaccination results
cea_files <- c(
  "outputs/paper/cea_0.6_onset_trigger_20250919.rds",
  "outputs/paper/cea_0.75_onset_trigger_20250919.rds",
  "outputs/paper/cea_0.9_onset_trigger_20250919.rds"
)
cea <- map_dfr(cea_files, read_rds)

cea_ori <- cea %>% filter(ori_occurred == TRUE)
# Case-triggered vaccination results
# cea_ct <- read_rds("outputs/cea_0.75_case_trigger_20250716.rds")
cea_ct <- read_rds("outputs/paper/cea_0.75_case_trigger_20250921.rds")
cea_ct_ori <- cea_ct %>% filter(ori_occurred == TRUE)

# excluded outbreaks for sensitivity analysis
# cea_excl <- read_rds("outputs/cea_0.6_onset_trigger_excl_20250716.rds")
# cea_excl <- rbind(cea_excl, read_rds("outputs/cea_0.75_onset_trigger_excl_20250716.rds"))
# cea_excl <- rbind(cea_excl, read_rds("outputs/cea_0.9_onset_trigger_excl_20250716.rds"))
# cea_ct_excl <- read_rds("outputs/cea_0.75_case_trigger_excl_20250716.rds")

# weekly decline
fit_decay <- read_rds("outputs/paper/fit_decay_model_20250923.rds")

# Load subset analyses reflecting global OCV stockpile
# cea_stockpile_subsets_q1 <-
#   read_rds("outputs/cea_stockpile_0_onset_trigger_by_year_20250828.rds")
# cea_stockpile_subsets_q2 <-
#   read_rds("outputs/cea_stockpile_0.25_onset_trigger_by_year_20250828.rds")
# cea_stockpile_subsets_q3 <-
#   read_rds("outputs/cea_stockpile_0.5_onset_trigger_by_year_20250828.rds")
# cea_stockpile_subsets_q4 <-
#   read_rds("outputs/cea_stockpile_0.75_onset_trigger_by_year_20250828.rds")
```

# Abstract

Values used in the abstract
```{r}
onset_impact <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, pct_reduc_case, 
                      case_averted_per_1000_OCV, cost_per_case_averted, 
                      cost_per_daly_averted)

onset_impact
# At week 1
# cases averted for vaccine coverage of 75%
format_median_iqr(onset_impact, digits=1, row=1)
# percent case averted
format_median_iqr(onset_impact, digits=1, row=3)
# case averted per 1000 OCVs
format_median_iqr(onset_impact, digits=2, row=5)
# cost per case averted
format_median_iqr(onset_impact, digits=0, row=7)
# At week 9
format_median_iqr(onset_impact, digits=1, row=2)
format_median_iqr(onset_impact, digits=1, row=4)
format_median_iqr(onset_impact, digits=2, row=6)

format_median_iqr(onset_impact, digits=0, row=8)
# cost per DALY averted
format_median_iqr(onset_impact, digits=0, row=9)
format_median_iqr(onset_impact, digits=0, row=10)


# weekly reduction rate
fit_decay[["5"]] %>% 
  filter(vacc_cov == VC) %>% 
  mutate(rate_reduc = 1 - exp(-decay_rate_k)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# half_life                  
fit_decay[["5"]] %>% 
  filter(vacc_cov == VC) %>%  
  summarize_quantiles(half_life_weeks) %>% 
  format_median_iqr(digits=1)


# stockpile impact
stockpile_impact <- cea_stockpile_subsets_q1 %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, pct_reduc_case, 
                      case_averted_per_1000_OCV, cost_per_case_averted, 
                      cost_per_daly_averted)
  
stockpile_impact
# cases averted for vaccine coverage of 75%
format_median_iqr(stockpile_impact, row=1)
# percentage cases averted
format_median_iqr(stockpile_impact, row=2)


format_median_iqr(stockpile_impact, digits=2, row=3)
format_median_iqr(stockpile_impact, row=4)

stockpile_impact_q4 <- cea_stockpile_subsets_q4 %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, pct_reduc_case, 
                      case_averted_per_1000_OCV, cost_per_case_averted, 
                      cost_per_daly_averted)
# cases averted for vaccine coverage of 75%
format_median_iqr(stockpile_impact_q4, row=1)
# percentage cases averted
format_median_iqr(stockpile_impact_q4, row=2)

rm(onset_impact, stockpile_impact, stockpile_impact_q4)
```

# Plots

## Figure 1. Outbreaks and % case reduction 

The right panel of Figure 1 in the main text was generated using the following code. The left panel, a flowchart illustrating outbreak identification and selection for vaccine impact modeling, was created in PowerPoint.

```{r}
set.seed(42)

# Global parameters ----------------------------------------------------
save_plot <- TRUE
text_size <- 5          # Default text size for annotations
line_height <- 0.95     # Line spacing for text annotation
pattern_dens <- 0.1     # Density of pattern filling
pattern_spacing <- 0.1  # Spacing between patterns
step_linewidth <- 0.7   # Width of step lines
arrow_vertical_len <- 14 # Vertical length of arrows

# Axis labels
ylabel <- "Weekly suspected cholera cases"
xlabel <- "Weeks since outbreak onset"

# Key timeline parameters
week_or <- 3      # Outbreak response at week 3  
week_vax <- 6.5   # Vaccination started
week_vax_eff <- 8.5 # Vaccine takes effect

# Rectangle dimensions configuration ------------------------------------
# This section makes it easy to adjust all rectangles at once

# Base rectangle dimensions
rect_width <- 0.8
rect_height <- 6

# Spacing between rectangles
gap_horizontal <- 1.9   # Horizontal gap between rectangle groups
gap_vertical <- 12      # Vertical gap between top and bottom rectangles

# Font sizes for elements
eq_symbol_size <- 8     # Size for equation symbols (<, +)
paren_size <- 9         # Size for parentheses
plus_size <- 5          # Size for plus symbols

# Vertical offset for second equation from first
eq2_vertical_offset <- 24

# Helper functions ----------------------------------------------------
create_filled_square <- function(x1, y1, x2, y2, fill_color, 
                                 color='black', alpha=0.5) {
  geom_rect(xmin=x1, xmax=x2, ymin=y1, ymax=y2,
            color=color, fill=fill_color, alpha=alpha, inherit.aes = FALSE)
}

create_striped_square <- function(x1, y1, x2, y2, fill_color, color='black',
                                  dens=0.1, angle=30) {
  geom_rect_pattern(
    xmin=x1, xmax=x2, ymin=y1, ymax=y2,
    pattern = 'stripe',
    fill = NA,
    colour = color,
    pattern_density = dens,
    pattern_spacing = 0.025,
    pattern_fill = fill_color,
    pattern_colour = NA,
    pattern_angle = angle,
    inherit.aes = FALSE)
}

create_divider <- function(x1, x2, y, extension=0.2, linewidth=0.7) {
  annotate("segment", 
           x = x1 - extension, 
           y = y, 
           xend = x2 + extension, 
           yend = y, 
           linewidth = linewidth)
}

theme_upper_panel <- function() {
  theme_classic() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 16),
      axis.text.y = element_text(size = 14),
      plot.caption = element_text(size = 14),
      legend.position = "none"
    )
}

# Data preparation ----------------------------------------------------
# Create synthetic example of an outbreak trajectory with and without vaccination
d <- data.frame(
  date = seq(as.Date("2010-06-15"), by = "week", length.out = 24),
  week = 0:23,
  # No vaccination scenario
  s_ch = c(2, 2, 3, 4, 10, 12, 29, 100, 120, 150, 140, 
           75, 68, 47, 24, 5, 12, 5, 3, 2, 4, 1, 0, 0)
)

# Add vaccination scenario (40% of cases after vaccine takes effect)
d$s_ch_vacc <- c(
  2, 2, 3, 4, 10, 12, 29, 100, 120, 
  round(0.4 * c(150, 140, 75, 68, 47, 24, 5, 12, 5, 3, 2, 4, 1, 0, 0))
)
d$s_ch_cumul <- cumsum(d$s_ch)

# Rename columns for clarity
names(d) <- c("date", "week", "No vacc", "Vacc", "No vacc cumulative")

# Reshape data for plotting
dlong <- pivot_longer(d, cols = c("No vacc", "Vacc"))

# Define time weeks for additional scenarios
weeks <- 0:25

# Generate additional outbreak scenarios
# Smallest outbreak (ended before vaccination)
smallest_no_vax <- rpois(length(weeks), lambda=30 * exp(-0.1 * (weeks - 2)^2))
smallest_no_vax[(week_vax-1):length(smallest_no_vax)] <- 0

# Smaller outbreak (ended at vaccine takes effect)
smaller_no_vax <- rpois(length(weeks), lambda=50 * exp(-0.1 * (weeks - 3)^2))
smaller_no_vax[week_vax_eff:length(smaller_no_vax)] <- 0

# Small outbreak (peak at week 8)
small_no_vax <- rpois(length(weeks), lambda=70 * exp(-0.08 * (weeks - 8)^2))
small_with_vax <- small_no_vax
small_with_vax[weeks > week_vax_eff] <- round(small_with_vax[weeks > week_vax_eff] * 0.5)  # 50% reduction

# Large outbreak (peak at week 12)
large_no_vax <- rpois(length(weeks), lambda=120 * exp(-0.04 * (weeks - 12)^2))
large_with_vax <- large_no_vax
large_with_vax[weeks > week_vax_eff] <- round(large_with_vax[weeks > week_vax_eff] * 0.5)  # 50% reduction

# Combine into a data frame
df <- tibble(
  week = rep(weeks, 6),
  cases = c(smallest_no_vax, smaller_no_vax, small_no_vax, small_with_vax, 
            large_no_vax, large_with_vax),
  outbreak = rep(c("Smallest", "Smaller", "Small", "Small", "Large", "Large"), 
                 each = length(weeks)),
  scenario = c(rep("No Vaccine", 2*length(weeks)), 
               rep(c("No Vaccine", "With Vaccine"), times=2, each=length(weeks)))
)

# Create wide version of the data
df_wide <- df %>%
  pivot_wider(names_from = scenario, values_from = cases) %>%
  mutate(reduction = `No Vaccine` - `With Vaccine`)

# Rectangle positions for equation 1 ----------------------------------
# Starting position for first equation
base_x1 <- 15
base_y1 <- 65

# a/b fraction
a_x1 <- base_x1
a_y1 <- base_y1
a_x2 <- a_x1 + rect_width
a_y2 <- a_y1 + rect_height

b_x1 <- a_x1
b_y1 <- a_y1 - gap_vertical
b_x2 <- b_x1 + rect_width
b_y2 <- b_y1 + rect_height

ab_div_y <- (a_y1 + b_y2) / 2

# c/d fraction
c_x1 <- a_x1 + gap_horizontal
c_y1 <- a_y1
c_x2 <- c_x1 + rect_width
c_y2 <- c_y1 + rect_height

d_x1 <- c_x1
d_y1 <- b_y1
d_x2 <- d_x1 + rect_width
d_y2 <- d_y1 + rect_height

cd_div_y <- (c_y1 + d_y2) / 2

# Inequality symbol
eq1_lt_x <- (a_x2 + c_x1) / 2
eq1_lt_y <- (a_y1 + b_y2) / 2

# Equation label
eq1_label_x <- a_x1 - 1.2
eq1_label_y <- cd_div_y

# Rectangle positions for equation 2 ----------------------------------
# Left side: 1/2(a/b+c/d)
# a/b part
eq2_a_x1 <- a_x1 + 0.2
eq2_a_y1 <- a_y1 - eq2_vertical_offset
eq2_a_x2 <- eq2_a_x1 + rect_width
eq2_a_y2 <- eq2_a_y1 + rect_height

eq2_b_x1 <- eq2_a_x1
eq2_b_y1 <- eq2_a_y1 - gap_vertical
eq2_b_x2 <- eq2_b_x1 + rect_width
eq2_b_y2 <- eq2_b_y1 + rect_height

# c/d part
eq2_c_x1 <- eq2_a_x1 + gap_horizontal
eq2_c_y1 <- eq2_a_y1
eq2_c_x2 <- eq2_c_x1 + rect_width
eq2_c_y2 <- eq2_c_y1 + rect_height

eq2_d_x1 <- eq2_c_x1
eq2_d_y1 <- eq2_b_y1
eq2_d_x2 <- eq2_d_x1 + rect_width
eq2_d_y2 <- eq2_d_y1 + rect_height

# Parentheses
left_paren_x <- eq2_a_x1 - 0.3
right_paren_x <- eq2_c_x2 + 0.3
paren_y <- (eq2_a_y1 + eq2_b_y2) / 2

# 1/2 fraction position
eq2_frac_x <- right_paren_x + 0.6
eq2_frac_y_num <- (eq2_a_y1 + eq2_a_y2)/2
eq2_frac_y_denom <- (eq2_b_y1 + eq2_b_y2)/2
eq2_frac_y_line <- (eq2_a_y1 + eq2_b_y2)/2

# Plus symbol position
eq2_plus1_x <- (eq2_a_x2 + eq2_c_x1) / 2
eq2_plus1_y <- (eq2_a_y1 + eq2_b_y2) / 2

# Less than symbol for second equation
eq2_lt_x <- eq2_frac_x + 0.8
eq2_lt_y <- (eq2_a_y1 + eq2_b_y2)/2

# Right side: (a+c)/(b+d)
# Numerator (a+c)
eq2_right_a_x1 <- eq2_lt_x + 0.8
eq2_right_a_y1 <- eq2_a_y1
eq2_right_a_x2 <- eq2_right_a_x1 + rect_width
eq2_right_a_y2 <- eq2_right_a_y1 + rect_height

eq2_right_c_x1 <- eq2_right_a_x1 + gap_horizontal
eq2_right_c_y1 <- eq2_a_y1
eq2_right_c_x2 <- eq2_right_c_x1 + rect_width
eq2_right_c_y2 <- eq2_right_c_y1 + rect_height

eq2_plus2_x <- (eq2_right_a_x2 + eq2_right_c_x1)/2
eq2_plus2_y <- (eq2_right_a_y1 + eq2_right_a_y2)/2

# Denominator (b+d)
eq2_right_b_x1 <- eq2_right_a_x1
eq2_right_b_y1 <- eq2_b_y1
eq2_right_b_x2 <- eq2_right_b_x1 + rect_width
eq2_right_b_y2 <- eq2_right_b_y1 + rect_height

eq2_right_d_x1 <- eq2_right_c_x1
eq2_right_d_y1 <- eq2_right_b_y1
eq2_right_d_x2 <- eq2_right_d_x1 + rect_width
eq2_right_d_y2 <- eq2_right_d_y1 + rect_height

eq2_plus3_x <- eq2_plus2_x
eq2_plus3_y <- (eq2_right_b_y1 + eq2_right_b_y2)/2

# Division line
eq2_right_div_x1 <- eq2_right_a_x1
eq2_right_div_x2 <- eq2_right_c_x2
eq2_right_div_y <- (eq2_right_a_y1 + eq2_right_b_y2) / 2

# Equation 2 label
eq2_label_x <- eq1_label_x
eq2_label_y <- a_y1 - eq2_vertical_offset

# Create plots --------------------------------------------------------
# Panel 1: Main example with vaccine effect
plt1a <- ggplot(dlong, aes(week, value, linetype = name, shape = name, group = name)) +
  # Background stripes for no vaccination scenario
  geom_rect_pattern(
    data = subset(dlong, name == "No vacc"),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=value),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_colour = NA,
    pattern_density = pattern_dens,
    pattern_fill = 'steelblue',
    pattern_angle = 30) +
  
  # Vaccination impact rectangle
  geom_rect(
    data = d,
    aes(xmin=week, xmax=lead(week), ymin=`Vacc`, ymax=`No vacc`), 
    fill='steelblue', 
    alpha=0.5,
    color=NA,
    inherit.aes = FALSE) +
  
  # Step lines for clarity
  geom_step(linewidth=step_linewidth) +
  
  # Outbreak response triggered arrow and label
  geom_segment(
    x=week_or, y=10+arrow_vertical_len, xend=week_or, yend=10,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_or-0.5, y=38, 
    label = "Outbreak\nresponse triggered", 
    hjust=0.5, size=text_size, lineheight=line_height
  ) +
  
  # Vaccination started arrow and label
  geom_segment(
    x=week_vax, y=32+arrow_vertical_len, xend=week_vax, yend=32,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_vax, y=60, 
    label = "Vaccination\nstarted", hjust=1,
    size=text_size, lineheight=line_height
  ) +
  
  # Vaccine takes effect arrow and label
  geom_segment(
    x=week_vax_eff, y=122+arrow_vertical_len, xend=week_vax_eff, yend = 122,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_vax_eff, y=147, 
    label="Vaccine\ntakes effect", hjust=1,
    size=text_size, lineheight=line_height
  ) +
  
  # Scenario labels
  annotate(
    "text", x=11.5, y=120, 
    label = "Without\nvaccination", hjust=0, 
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=9.5, y=12, 
    label = "With\nvaccination", hjust=0,
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=16, y=75, 
    label = "% case\nreduction", 
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=19.2, y=75, 
    label = bquote("= 100 \u00D7"), 
    size=text_size, lineheight=line_height
  ) +
  
  # Fraction expressed using squares
  create_divider(x1=21.1-0.2, x2=22.1+0.2, y=75) +
  create_filled_square(
    x1=21.1, x2=22.1, y1=77, y2=85, 
    color='black', fill="steelblue", alpha=0.5
  ) +
  create_striped_square(
    x1=21.1, x2=22.1, y1=64, y2=72,
    fill_color='steelblue', color='black'
  ) +
  
  # Colored lines on the horizontal axis marking periods
  annotate("segment", x=0, y=0, xend=3, yend=0, linewidth=2, color="firebrick") +
  annotate("segment", x=3, y=0, xend=7, yend=0, linewidth=2, color="forestgreen") +
  annotate("segment", x=7, y=0, xend=9, yend=0, linewidth=2, color="purple") +
  
  scale_x_continuous(limits=c(0,25)) +
  labs(x="", y=ylabel) +
  theme_upper_panel()

# Panel 2: Smaller outbreaks
plt1b <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot smaller outbreaks
  geom_step(
    data = subset(df, outbreak == "Smallest"),
    aes(linetype=scenario),
    linewidth=step_linewidth, 
    color='grey'
  ) +
  geom_step(
    data = subset(df, outbreak == "Smaller"),
    aes(linetype=scenario),
    linewidth=step_linewidth, 
    color='black'
  ) +
  
  labs(x=NULL, y=ylabel) +
  theme_upper_panel()

# Panel 3: Small outbreak with equations
plt1c <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot small outbreak
  geom_step(
    data = subset(df, outbreak == "Small"), 
    aes(linetype=scenario),
    linewidth=step_linewidth
  ) +
  
  # Pattern for no vaccination scenario
  geom_rect_pattern(
    data = subset(df, (scenario == "No Vaccine" & outbreak == "Small")),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=cases),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_density = pattern_dens*0.5,
    pattern_fill = 'steelblue',
    pattern_colour = NA,
    pattern_angle = 30
  ) +
  
  # Vaccination impact area
  geom_rect(
    data=subset(df_wide, outbreak == "Small"),
    aes(xmin=week, xmax=lead(week), ymin=`With Vaccine`, ymax=`No Vaccine`),
    fill="steelblue", 
    alpha=0.5, 
    inherit.aes=FALSE
  ) +
  
  # EQUATION 1: a/b < c/d
  # a/b fraction
  create_filled_square(a_x1, a_y1, a_x2, a_y2, "steelblue") +
  create_striped_square(b_x1, b_y1, b_x2, b_y2, "steelblue", dens=pattern_dens*0.5) +
  create_divider(a_x1, a_x2, ab_div_y) +
  
  # Less than symbol
  annotate("text", x=eq1_lt_x, y=eq1_lt_y, label= "<", size=eq_symbol_size) +
  
  # c/d fraction
  create_filled_square(c_x1, c_y1, c_x2, c_y2, "firebrick") +
  create_striped_square(d_x1, d_y1, d_x2, d_y2, "firebrick", dens=pattern_dens*0.5) +
  create_divider(c_x1, c_x2, cd_div_y) +
  
  # Label for equation 1
  annotate("text", x=eq1_label_x, y=eq1_label_y, label="1.", size=5, hjust=0) +
  
  # EQUATION 2: 1/2(a/b+c/d) < (a+c)/(b+d)
  # Left side: 1/2(a/b+c/d)
  
  # a/b part
  create_filled_square(eq2_a_x1, eq2_a_y1, eq2_a_x2, eq2_a_y2, "steelblue") +
  create_striped_square(eq2_b_x1, eq2_b_y1, eq2_b_x2, eq2_b_y2, "steelblue", dens=pattern_dens*0.5) +
  create_divider(eq2_a_x1, eq2_a_x2, (eq2_a_y1 + eq2_b_y2)/2) +
  
  # Plus symbol
  annotate("text", x=eq2_plus1_x, y=eq2_plus1_y, label="+", size=plus_size) +
  
  # c/d part
  create_filled_square(eq2_c_x1, eq2_c_y1, eq2_c_x2, eq2_c_y2, "firebrick") +
  create_striped_square(eq2_d_x1, eq2_d_y1, eq2_d_x2, eq2_d_y2, "firebrick", dens=pattern_dens*0.5) +
  create_divider(eq2_c_x1, eq2_c_x2, (eq2_c_y1 + eq2_d_y2)/2) +
  
  # Parentheses
  annotate("text", x=left_paren_x, y=paren_y, label="(", size=paren_size) +
  annotate("text", x=right_paren_x, y=paren_y, label=")", size=paren_size) +
  
  # 1/2 fraction
  annotate("text", x=eq2_frac_x, y=eq2_frac_y_num, label="1", size=5) +
  annotate("text", x=eq2_frac_x, y=eq2_frac_y_denom, label="2", size=5) +
  annotate("segment", 
           x=eq2_frac_x-0.3, y=eq2_frac_y_line, 
           xend=eq2_frac_x+0.3, yend=eq2_frac_y_line, 
           linewidth=0.8) +
  
  # Less than symbol for second equation
  annotate("text", x=eq2_lt_x, y=eq2_lt_y, label="<", size=eq_symbol_size) +
  
  # Right side: (a+c)/(b+d)
  # Numerator (a+c)
  create_filled_square(eq2_right_a_x1, eq2_right_a_y1, eq2_right_a_x2, eq2_right_a_y2, "steelblue") +
  annotate("text", x=eq2_plus2_x, y=eq2_plus2_y, label="+", size=plus_size) +
  create_filled_square(eq2_right_c_x1, eq2_right_c_y1, eq2_right_c_x2, eq2_right_c_y2, "firebrick") +
  
  # Division line
  create_divider(eq2_right_div_x1, eq2_right_div_x2, eq2_right_div_y) +
  
  # Denominator (b+d)
  create_striped_square(eq2_right_b_x1, eq2_right_b_y1, eq2_right_b_x2, eq2_right_b_y2, 
                        "steelblue", dens=pattern_dens*0.5) +
  annotate("text", x=eq2_plus3_x, y=eq2_plus3_y, label="+", size=plus_size) +
  create_striped_square(eq2_right_d_x1, eq2_right_d_y1, eq2_right_d_x2, eq2_right_d_y2, 
                        "firebrick", dens=pattern_dens*0.5) +
  
  # Label for equation 2
  annotate("text", x=eq2_label_x, y=eq2_label_y, label="2.", size=5, hjust=0) +
  
  labs(x=NULL, y=ylabel) +
  theme_upper_panel()

# Panel 4: Large outbreak  
plt1d <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot large outbreak
  geom_step(
    data = subset(df, outbreak == "Large"), 
    aes(linetype=scenario),
    linewidth=step_linewidth
  ) +
  
  # Pattern for no vaccination scenario
  geom_rect_pattern(
    data = subset(df, (scenario == "No Vaccine" & outbreak == "Large")),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=cases),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_density = pattern_dens*0.5,
    pattern_fill = 'firebrick',
    pattern_colour = NA,
    pattern_angle = 30
  ) +
  
  # Vaccination impact area
  geom_rect(
    data=subset(df_wide, outbreak == "Large"),
    aes(xmin=week, xmax=lead(week), ymin=`With Vaccine`, ymax=`No Vaccine`),
    fill="firebrick", 
    alpha=0.5, 
    inherit.aes=FALSE
  ) +
  
  # Axis and labels
  labs(x=xlabel, y=ylabel) +
  scale_x_continuous(limits=c(0,25)) +
  theme_upper_panel() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    legend.position = "none"
  )

# Combine all plots
# Right panel of Figure 1
combined_plot <- (plt1a / plt1b / plt1c / plt1d) + 
  plot_layout(
    heights = c(0.8, 0.4, 0.4, 0.4), 
    axis_titles = 'collect',
    guides = 'collect'
  )

# Save plot
if (save_plot) {
  fac <- 1.8
  ggsave(paste0("plots/fig1_vacc_eff_schematic_", tstamp(), ".pdf"),
        combined_plot,
        device = cairo_pdf,
        width=1.25*fgw1*fac,
        height=1.6*fgw1*fac,
        units = "mm")
}
```

## Figure 2: Outbreak Description

Outbreak ids by confirmation status
```{r}
ids_excl <- outbreak_data_all |> 
  filter(!(id_outbreak %in% outbreak_data$id_outbreak)) |>  
  pull(id_outbreak) |>
  unique()

# Identify confirmed outbreaks (with laboratory confirmation)
ids_conf <- outbreak_data |> 
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique()

# Identify suspected (not confirmed) outbreaks
ids_susp <- outbreak_data |>
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |>
  filter(!(c_ch_sum > 0)) |>
  pull(id_outbreak) |>
  unique()
```

```{r}
# Load and prepare country-region mapping data
country_region <- fread("data/paper/africa_subregion_country_names.csv")
dt <- country_region
country_region <- country_region[Subregion == "Middle Africa", Subregion := "Central Africa"]
# change Middle to Central
country_region %>% mutate()
country_region$iso3c <- sapply(country_region$Country, 
                               function(x) countrycode(x, origin = "country.name.en", 
                                                     destination = "iso3c"))

# Join outbreak data with regional information
outbreak_data$iso3c <- outbreak_data$country
outbreak_data_region <- left_join(outbreak_data, country_region, by = "iso3c")

# =============================================================================
# PLOT 2A: TEMPORAL TRENDS BY SUBREGION (STACKED AREA CHART)
# =============================================================================

# Aggregate data by month and subregion
outbreak_month <- outbreak_data_region %>% 
  mutate(date = as.Date(date),
         date_bin = floor_date(date, "1 month"))

# Calculate total cases per subregion per month
outbreak_month_summary <- outbreak_month %>%
  group_by(date_bin, Subregion) %>%
  summarise(total_cases = sum(s_ch, na.rm = TRUE), .groups = "drop")

# Create complete time series grid to fill gaps
all_subregions <- unique(outbreak_month_summary$Subregion)
all_date_bins <- seq(min(outbreak_month_summary$date_bin), 
                     max(outbreak_month_summary$date_bin), 
                     by = "1 month")
full_grid <- expand_grid(date_bin = all_date_bins, Subregion = all_subregions)

# Merge and fill missing values with zero
df_complete <- full_grid %>%
  left_join(outbreak_month_summary, by = c("date_bin", "Subregion")) %>%
  mutate(total_cases = replace_na(total_cases, 0))

# Create stacked area plot
plt_2a <- ggplot(df_complete, aes(x = date_bin, y = total_cases, 
                                  fill = Subregion)) +
  geom_area(position = "stack", alpha = 0.8) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  scale_y_continuous(labels = comma_format()) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  labs(x = "", y = "Monthly suspected cases", fill = "") +
  theme_light() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    plot.margin = margin(0,0,0,0,"pt")
  )

# =============================================================================
# PLOT 2B: SPATIAL DISTRIBUTION MAP
# =============================================================================
# Prepare outbreak data for mapping
dt <- outbreak_data %>% rename(lctn_pr = location_period_id)
first_rows_df <- dt %>% 
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

# Ensure consistent data types for joining
shp$lctn_pr <- as.character(shp$lctn_pr)

# Join outbreak data with spatial data
map_data <- left_join(first_rows_df, shp, by = "lctn_pr")
map_data %<>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE))

# Convert to spatial features
map_data <- st_as_sf(map_data)

islands_to_exclude <- c("CPV",  # Cape Verde
                        "SYC",  # Seychelles
                        "MYT",  # Mayotte
                        "REU",  # Réunion
                        "SHN")  # Saint Helena, Ascension and Tristan da Cunha

# Exclude these from your Africa shapefile
africa_main <- africa %>% dplyr::filter(!ISO3 %in% islands_to_exclude)
bbox   <- st_bbox(africa_main)       # much tighter now

plt_2b <- ggplot() +
  geom_sf(data = africa_main,  fill = NA, colour = "black", size = .3) +
  geom_sf(data = map_data, aes(colour = is_confirmed, fill = is_confirmed),
          alpha = .7, size = .8) +
  coord_sf(xlim = c(bbox["xmin"], bbox["xmax"]-10),
           ylim = c(bbox["ymin"]+12, bbox["ymax"]),
           expand = FALSE, clip = "on") +    # clip cuts everything outside
  scale_colour_manual("", values = c("steelblue4", "firebrick4"),
                      labels  = c("Not confirmed", "Confirmed")) +
  scale_fill_manual("",   values = c("steelblue4", "firebrick4"),
                    labels  = c("Not confirmed", "Confirmed")) +
  theme_void() +                                   
  theme(legend.position = "top",
        plot.margin     = margin(0,0,0,0,"pt"),
        panel.spacing = unit(0, "pt"))

# plt_2b <- plt_2b + theme(aspect.ratio = 0.7)

# =============================================================================
# PLOT 2C: EPIDEMIC CURVES WITH INSET
# =============================================================================

# Prepare weekly case data
week_sch <- outbreak_data[, c("id_outbreak", "week", "s_ch")]
setDT(week_sch)
max_week <- week_sch[, max(week)]

# Create complete grid for all outbreaks up to max week
all_weeks <- CJ(id_outbreak = unique(week_sch$id_outbreak), week = 1:max_week)

# Merge and fill missing values with zero
week_sch_full <- merge(all_weeks, week_sch, 
                       by = c("id_outbreak", "week"), all.x = TRUE)
week_sch_full[is.na(s_ch), s_ch := 0]

# Calculate summary statistics across all outbreaks
week_sch_summary <- week_sch_full[, .(
  sum_sch = sum(s_ch, na.rm = TRUE),
  median = median(s_ch, na.rm = TRUE),
  p025 = quantile(s_ch, 0.025, na.rm = TRUE),
  p25  = quantile(s_ch, 0.25, na.rm = TRUE),
  p75  = quantile(s_ch, 0.75, na.rm = TRUE),
  p975 = quantile(s_ch, 0.975, na.rm = TRUE)
), by = week]


# Main epidemic curve plot
plt2c <- ggplot(week_sch_summary, aes(x = week)) +
  geom_ribbon(aes(ymin = p025, ymax = p975), fill = set1[3], alpha = 0.45) +
  geom_ribbon(aes(ymin = p25, ymax = p75), fill = set1[3], alpha = 0.6) +
  geom_line(aes(y = median), color = "black", linewidth = 1) +
  labs(title = "", x = "Epidemic week", y = "Suspected cholera cases") +
  scale_x_continuous(limits = c(0, 50), breaks = seq(0, 50, 10)) +
  theme_light() +
  theme(
    plot.margin = margin(0,1,0,1, "pt")
  )

# time to peak
outbk_data_summary %>% 
  summarize(mean_time_to_peak = mean(time_to_peak_week),
            median_time_to_peak = median(time_to_peak_week))

# percentage of samples that are shorter than or equal to 10 weeks
sum(outbk_data_summary$time_to_peak_week <= 10) /
  length(outbk_data_summary$time_to_peak_week)


plt2c_inset <- ggplot(outbk_data_summary, aes(x = time_to_peak_week)) +
  stat_ecdf(geom = "step", size = 0.8) +
  labs(x = "Time to peak (weeks)", 
       y = expression(P(X <= x)))+
  scale_x_continuous(limits=c(0,30), breaks = seq(0,30,10), labels = seq(0,30,10))+
  scale_y_continuous(limits=c(0,1), breaks = seq(0,1,0.5), labels = seq(0,1,0.5))+
  theme_light()+
  theme_tight()+
  theme(plot.margin = margin(0,0,0,0,"pt"))


# Convert inset to grob and add to main plot with better positioning
grob_inset <- ggplotGrob(plt2c_inset)

plt_2c_combined <- plt2c +
  annotation_custom(
    grob = grob_inset,
    xmin = 20, xmax = 50,  # Moved to right side with more space
    ymin = 100, ymax = 340   # Increased height for better visibility
  )

# =============================================================================
# PLOT 2D: OUTBREAK SIZE vs DURATION
# =============================================================================

plt_2d <- outbk_data_summary %>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE)) %>% 
  ggplot() +
  geom_point(aes(x = total_s_ch, y = duration_weeks, color = is_confirmed), 
             alpha = 0.7, size = 1.2) +
  scale_y_log10(labels = comma_format()) +
  scale_x_log10(labels = comma_format()) +
  scale_color_manual("", 
                     labels = c("Not confirmed", "Confirmed"),
                     values = c("steelblue4", "firebrick4")) +
  labs(title = "", x = "Total suspected cases", y = "Outbreak duration (weeks)") +
  theme_light() +
  theme(legend.position = "top", plot.margin = margin(0,1,0,0,"pt"))

# =============================================================================
# PLOT 2E: OUTBREAK SIZE vs ATTACK RATE
# =============================================================================

summary(outbk_data_summary$duration_weeks) 
summary(outbk_data_summary$total_s_ch)
summary(outbk_data_summary$ar_per_10e3)

summary(outbk_data_summary[id_outbreak %in% ids_conf]$duration_weeks)
summary(outbk_data_summary[id_outbreak %in% ids_conf]$total_s_ch)
summary(outbk_data_summary[id_outbreak %in% ids_conf]$ar_per_10e3)

plt_2e <- outbk_data_summary %>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE)) %>% 
  ggplot() +
  geom_point(aes(x = total_s_ch, y = ar_per_10e3, color = is_confirmed), 
             alpha = 0.7, size = 1.2) +
  scale_y_log10(labels = comma_format()) +
  scale_x_log10(labels = comma_format()) +
  scale_color_manual("", 
                     labels = c("Not confirmed", "Confirmed"),
                     values = c("steelblue4", "firebrick4")) +  
  labs(title="", x = "Total suspected cases", y = "Attack rate (per 10,000)") +
  theme_light() +
  theme(legend.position="top",  plot.margin = margin(0,0,0,1,"pt"))



# =============================================================================
# FINAL COMBINED PLOT WITH IMPROVED LAYOUT
# =============================================================================

# Create combined plot with better proportions
# Row 1: Temporal trends (full width)
# Row 2: Map (60%) + Epidemic curve (40%) 
# Row 3: Scatter plots (50% each)
combined_plot <- plt_2a /
  (plt_2b + plt_2c_combined + plot_layout(widths = c(0.6, 1))) /
  (plt_2d + plt_2e + plot_layout(widths = c(1, 1), guides="collect") +
     plot_annotation(theme = theme(legend.position = "top")))+
  plot_layout(heights = c(1,1.3,1)) +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.margin = margin(0,0,0,0,"pt"),
        plot.tag.position = c(0.2, 1),
        plot.tag = element_text(size = 13))

# Display the combined plot
# combined_plot
if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/fig2_outbreak_summary_", tstamp(), ".pdf"), combined_plot,
         device = cairo_pdf, width = fgw2 * fac, height = fgw2 * fac, units = "mm")
}
```

## Figure S1. Indirect Vaccine Effectiveness 

## Figure S2. All outbreaks (eligible and ineligible)

```{r figS1, eval=FALSE}
# Create time series plot with different colors by outbreak type
plt_S1a <- ggplot() +
  geom_line(data = filter(outbreak_data_all, id_outbreak %in% ids_excl),
            aes(date, s_ch, group = id_outbreak, color = "Excluded"), 
            alpha = 0.5) +
  geom_line(data = filter(outbreak_data, !(id_outbreak %in% ids_conf)),
            aes(date, s_ch, group = id_outbreak, color = "Suspected"), 
            alpha = 0.5) +
  geom_line(data = filter(outbreak_data, id_outbreak %in% ids_conf), 
            aes(date, s_ch, group = id_outbreak, color = "Confirmed"),
            alpha = 0.5) +
  scale_color_manual("", values = c("Excluded" = "yellow4",
                                  "Suspected" = "steelblue4", 
                                  "Confirmed" = "firebrick4")) +
  labs(x = "", y = "Weekly suspected cases") +
  theme_light() +
  theme_tight() +
  theme(legend.position = "top")

# (B) Outbreak size vs. duration by type
outbreak_data_all |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                  as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d_all

outbreak_data |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                  as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d

if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1a_outbreak_ts_", tstamp(), ".pdf"), plt_S1a,
         device = cairo_pdf, width = fgw2 * fac, height = fgw1 * fac, units = "mm")
}

plt_S1b <- ggplot() +
  geom_point(data = filter(d_all, id_outbreak %in% ids_excl), 
             aes(duration, total_s_ch, color = "Excluded"), alpha = 0.5) +
  geom_point(data = filter(d, id_outbreak %in% ids_susp), 
             aes(duration, total_s_ch, color = "Suspected"), alpha = 0.5) +
  geom_point(data = filter(d, id_outbreak %in% ids_conf), 
             aes(duration, total_s_ch, color = "Confirmed"), alpha = 0.5) +
  scale_x_log10() +
  scale_y_log10() +
  theme_light() +
  theme_tight() +
  scale_color_manual("", values = c("Excluded" = "yellow4", 
                                  "Suspected" = "steelblue4", 
                                  "Confirmed" = "firebrick4")) +
  labs(x = "Outbreak duration (days)", y = "Total suspected cases") +
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1b_outbreak_size_dur_", tstamp(), ".pdf"), plt_S1b,
         device = cairo_pdf, width = fgw1 * fac, height = fgw1 * fac, units = "mm")
}

# (C) Map of outbreaks
dt <- outbreak_data 
dt_all <- outbreak_data_all 

# Prepare data for mapping
dt <- dt %>% rename(lctn_pr = location_period_id)
dt_all <- dt_all %>% rename(lctn_pr = location_period_id)

first_rows_df <- dt %>% 
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

first_rows_df_all <- dt_all %>%
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

first_rows_df$lctn_pr <- as.numeric(first_rows_df$lctn_pr)
first_rows_df_all$lctn_pr <- as.numeric(first_rows_df_all$lctn_pr)

# Join with spatial data
shp$lctn_pr <- as.character(shp$lctn_pr)
first_rows_df$lctn_pr <- as.character(first_rows_df$lctn_pr)
first_rows_df_all$lctn_pr <- as.character(first_rows_df_all$lctn_pr)

d <- left_join(first_rows_df, shp, by = "lctn_pr")
d_all <- left_join(first_rows_df_all, shp, by = "lctn_pr")

# Create indicator variables for outbreak types
d <- d %>%
  mutate(id_conf = ifelse(id_outbreak %in% ids_conf, 1, 0),
         id_sus = ifelse(!id_outbreak %in% ids_conf, 1, 0))

# Subset data by confirmation status
d_conf <- d %>% filter(id_conf == 1)
d_sus <- d %>% filter(id_sus == 1)

# Convert to spatial data
d_conf <- st_as_sf(d_conf)
d_sus <- st_as_sf(d_sus)
d_all <- d_all %>% mutate(id_excl = ifelse(id_outbreak %in% ids_excl, 1, 0))
d <- st_as_sf(d)
d_all <- st_as_sf(d_all)

# Create map
plt_S1c <- ggplot() +
  geom_sf(data = africa, fill = NA, color = "black") +  # Country outlines
  geom_sf(data = d_all, aes(fill = id_excl), color = "black", 
          fill = "yellow4", alpha = 0.5) +
  geom_sf(data = d_sus, aes(fill = id_sus), color = "steelblue4", 
          fill = "steelblue4", alpha = 0.5) +
  geom_sf(data = d_conf, aes(fill = id_conf), color = "firebrick4", 
          fill = "firebrick4", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Map of Cholera Outbreaks")


if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1c_outbreak_map_", tstamp(), ".pdf"), plt_S1c,
         device = cairo_pdf, width = fgw1 * fac, height = fgw1 * fac, units = "mm")
}
```

## Figure 3

### Summary values referred in main text
```{r fig3, eval=FALSE}
onset_impact <- cea %>%
  filter(ori_occurred) %>%
  filter(vacc_cov %in% c(0.6, 0.75, 0.9), vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, pct_reduc_case, 
                      case_averted_per_1000_OCV, cost_per_case_averted)

# cases averted for vaccine coverage of 75%
format_median_iqr(onset_impact, row=3)
format_median_iqr(onset_impact, row=4)
# percentage cases averted for vaccine coverage of 75%
format_median_iqr(onset_impact, row=9)
format_median_iqr(onset_impact, row=10)
# cases averted per 1,000 OCVs
format_median_iqr(onset_impact, digits=3, row=15)
format_median_iqr(onset_impact, digits=3, row=16)
# cost per case averted
format_median_iqr(onset_impact, row=21)
format_median_iqr(onset_impact, row=22)


# outbreaks missed by delayed vaccination
wks_delay <- c(1, 5, 9, 14) # de
dur_missed_id <- vector('list', length(wks_delay))
vax_missed_id <- vector('list', length(wks_delay))

for (i in 1:length(wks_delay)) {
  outbk_data_summary %>% 
    filter(duration_weeks <= wks_delay[i]) %>%
    pull(id_outbreak) -> dur_missed_id[[i]]
  
  cea %>% 
    filter(vacc_week == wks_delay[i] & runid == 1 & vacc_cov == 0.75) %>% 
    filter(!ori_occurred) %>% 
    pull(id_outbreak) -> vax_missed_id[[i]]
}
n_outbreaks_missed1 <- map_vec(dur_missed_id, function(i) length(i))
n_outbreaks_missed2 <- map_vec(vax_missed_id, function(i) length(i))

n_outbreaks_missed1
100 * n_outbreaks_missed1 / nrow(outbk_data_summary)
100 * n_outbreaks_missed2 / nrow(outbk_data_summary)

# 60%, 75%, 90% at week 1
onset_impact %>% filter(vacc_week == 1) %>% format_median_iqr(row=1:3, digits=0)
# 60%, 75%, 90% at week 9
onset_impact %>% filter(vacc_week == 9) %>% format_median_iqr(row=1:3, digits=0)

# percentage case reduction by size group
probs <- seq(0, 1, by = 0.1)
unq_data <- filter(cea, runid == 1 & vacc_cov == 0.75 & vacc_week == 1)
size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)

d <- cea
d <- left_join(d, outbk_data_summary[,
                      c("id_outbreak", "ar_per_10e3","duration_weeks")],
                 by = "id_outbreak")

# d$size_grp <- sapply(d$s_ch_tot, function(x) get_group_id(x, size_grp_cutoff))
# 
# d <- d[ori_occurred == T]
# res <- rbindlist(lapply(1:10, function(i) calculate_impact_by_subset(d, i)))
# res %>% filter(vacc_week %in% c(1,5,9), subset_id %in% c(1,10), vacc_cov == VC)

# rate of decrease in percentage case reduction
# by size, attack rate, and duration
ar_grp_cutoff <- quantile(outbk_data_summary$ar_per_10e3, probs = probs)
dur_grp_cutoff <- quantile(outbk_data_summary$duration_weeks, probs = probs)

d %<>%
  group_by(runid, id_outbreak, vacc_cov) %>%
  mutate(pct_reduc_case_frac = 1-frac_change(pct_reduc_case, lag=1),
         pct_reduc_case_diff = c(1, -diff(pct_reduc_case, lag=1)))

d %>% filter(vacc_cov == 0.75) -> d1
d1$grp <- "All"
d %>% filter(ar_per_10e3 > ar_grp_cutoff[10]) -> d2 # top 10%
d2$grp <- "AR"
d %>% filter(duration_weeks > dur_grp_cutoff[10]) -> d3 # top 10%
d3$grp <- "Dur"
d %>% filter(s_ch_tot > size_grp_cutoff[10]) -> d4 # top 10%
d4$grp <- "Size"

d <- rbind(d1, d2)
d_ <- rbind(d3, d4)
d <- rbind(d, d_)

d %>%
  filter(vacc_week > 1, vacc_week <= 15) %>% 
  mutate(half_life = log(2)/abs(pct_reduc_case_frac)) %>% 
  group_by(grp) %>% 
  summarize_quantiles(pct_reduc_case_frac*100, half_life) -> res

res$grp <- factor(res$grp, c("All", "AR", "Dur", "Size"))
for (i in 1:4) print(format_median_iqr(res, row=i))
for (i in 5:8) print(format_median_iqr(res, digits=3, row=i))

# percentage reduction in an original scale
# 1 - exp(-k_hat)
names(fit_decay)
fit_decay[["5"]] %>%  # outbreaks 5 weeks or longer
  filter(vacc_cov == VC) %>% 
  mutate(rate_reduc = 100*(1 - exp(- decay_rate_k))) %>% 
  summarize_group(vars=c("rate_reduc"), group_name="> 5 weeks") %>% 
  format_median_iqr(digits=1)

fit_decay[["9"]] %>%  # outbreaks 5 weeks or longer
  filter(vacc_cov == VC) %>% 
  mutate(rate_reduc = 100*(1 - exp(- decay_rate_k))) %>% 
  summarize_group(vars=c("rate_reduc"), group_name="> 9 weeks") %>% 
  format_median_iqr(digits=1)

fit_decay[["13"]] %>%  # outbreaks 5 weeks or longer
  filter(vacc_cov == VC) %>% 
  mutate(rate_reduc = 100*(1 - exp(- decay_rate_k))) %>% 
  summarize_group(vars=c("rate_reduc"), group_name="> 13 weeks") %>% 
  format_median_iqr(digits=1)

rm(d, d1, d2, d3, d4, res, onset_impact)
```

### Plots 
```{r}
cea_ori_sum <- read_rds("outputs/paper/cea_ori_plot_summary_20251001.rds")

ymean  <- 95 
ymedian<- 87.5
y50    <- c(82, 93)
y95    <- c(75, 100)

VW_plot_seq <- seq(1,15,2)
# percentage case averted 
p1 <- cea_ori_sum %>%
  filter(vacc_week %in% VW_plot_seq) %>%
  ggplot(aes(x = vacc_week, color = as.factor(vacc_cov), 
             group = as.factor(vacc_cov))) +
  ci_layers_from_summary(
    dodge_width = 1.4,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = 2.4,
    mean_size = x_size,
    mean_stroke = x_stroke
  )+
  add_summary_braces(
    x0 = 13,
    ymean = ymean, ymedian = ymedian,
    y50 = y50, y95 = y95,
    pt_size_median = 3,
    pt_size_mean = x_size,
    pt_stroke_mean = x_stroke,
    dx50 = 0.5, dx95 = 0.8,
    text_50_x = 13.6,
    text_50_y = y50[2],
    text_95_x = 14.85,
    text_95_y = y95[2],
    brace_50_x = 13.2,
    brace_95_x = 14.4,
    text_offset_mean = 0.3,
    lw_50 = 0.6, lw_95 = 0.6,
    whisker_50 = line_width_50, 
    whisker_95 = line_width_95
  ) +
  scale_y_continuous("% cases averted",
                     labels = scales::percent_format(scale = 1), 
                     limits=c(0,100)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[1:3],
    labels = c(paste0(format_num(VC3*100, 1), "%")))+
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))

# plot 2
# exponential decay curve
set.seed(42)
## ---- sampling / filtering (unchanged except vacc_cov) ----
runid_sample <- 1
# decay model was fit using the first 5, 9, 13 points and they 
ids_wk_5 <- fit_decay$`5` %>% 
  filter(!(id_outbreak %in% unique(fit_decay$`9`$id_outbreak))) %>%
  pull(id_outbreak) %>% unique()

id_outbreak_sample_wk_5 <- sample(ids_wk_5, size = 10)

fitted_models <- fit_decay$`5` %>% 
  filter(runid %in% runid_sample,
         id_outbreak %in% id_outbreak_sample_wk_5,
         vacc_cov == 0.75)
# x-axis grid
x_range <- seq(1, 15, by = 0.2)
t0 <- 1  # baseline week for normalization

## ---- normalized "spaghetti" using fitted params ----
# On the normalized scale, A0 cancels; z_hat(t) = exp(-k * (t - t0))
spaghetti_data_norm_wk_5 <- fitted_models %>%
  mutate(vacc_week = list(x_range)) %>%
  unnest(vacc_week) %>%
  mutate(
    z_hat = exp(-decay_rate_k * (vacc_week - t0))
  )

ids_wk_9 <- fit_decay$`9` %>% 
  filter(!(id_outbreak %in% unique(fit_decay$`13`$id_outbreak))) %>%
  pull(id_outbreak) %>% unique()

id_outbreak_sample_wk_9 <- sample(ids_wk_9, size = 10)

fitted_models <- fit_decay$`9` %>% 
  filter(runid %in% runid_sample,
         id_outbreak %in% id_outbreak_sample_wk_9,
         vacc_cov == 0.75)

## ---- normalized "spaghetti" using fitted params ----
# On the normalized scale, A0 cancels; z_hat(t) = exp(-k * (t - t0))
spaghetti_data_norm_wk_9 <- fitted_models %>%
  mutate(vacc_week = list(x_range)) %>%
  unnest(vacc_week) %>%
  mutate(
    z_hat = exp(-decay_rate_k * (vacc_week - t0))
  )

ids_wk_13 <- unique(fit_decay$`13`$id_outbreak)
id_outbreak_sample_wk_13 <- sample(ids_wk_13, size = 10)

fitted_models <- fit_decay$`13` %>% 
  filter(runid %in% runid_sample,
         id_outbreak %in% id_outbreak_sample_wk_13,
         vacc_cov == 0.75)

## ---- normalized "spaghetti" using fitted params ----
# On the normalized scale, A0 cancels; z_hat(t) = exp(-k * (t - t0))
spaghetti_data_norm_wk_13 <- fitted_models %>%
  mutate(vacc_week = list(x_range)) %>%
  unnest(vacc_week) %>%
  mutate(
    z_hat = exp(-decay_rate_k * (vacc_week - t0))
  )

# k_hat and half_life colums are generated during fitting
# It, however, occurs only for the first 5 weeks (weeks 1-5) 
# and therefore set to NULL. before joining with weeks 1-9 and weeks 1-13 
# results
spaghetti_data_norm_wk_5$k_hat <- NULL
spaghetti_data_norm_wk_5$half_life <- NULL 
 
spaghetti_data_norm <- rbindlist(list(spaghetti_data_norm_wk_5, 
                                      spaghetti_data_norm_wk_9,
                                      spaghetti_data_norm_wk_13))
id_outbreak_sample <- c(id_outbreak_sample_wk_5, id_outbreak_sample_wk_9, 
                        id_outbreak_sample_wk_13)
## ---- raw data, normalized by week-1 value per series ----
# Expecting 'cea_ori' with columns: vacc_cov, ori_occurred, vacc_week, runid, id_outbreak, pct_reduc_case
raw_data_points_norm <- cea_ori %>%
  filter(vacc_cov == 0.75,
         vacc_week %in% 1:15, 
         runid %in% runid_sample,
         id_outbreak %in% id_outbreak_sample) %>%
  arrange(runid, id_outbreak, vacc_week) %>%
  group_by(runid, id_outbreak) %>%
  mutate(
    y0 = first(pct_reduc_case),
    z  = ifelse(y0 > 0, pct_reduc_case / y0, NA_real_)
  ) %>%
  ungroup()

## ---- summary curves from k only (median + IQR) ----
# Use the full data set
# k_q <- quantile(fit_decay$`5`$decay_rate_k, 
#                 probs = c(0.025, .25, .5, .75, 0.975), na.rm = TRUE)
k_q <- quantile(fit_decay$`9`$decay_rate_k, 
                probs = c(0.025, .25, .5, .75, 0.975), na.rm = TRUE)
k_q025 <- k_q[[1]]; k_q25 <- k_q[[2]]; k_med <- k_q[[3]]; 
k_q75 <- k_q[[4]]; k_q975 <- k_q[[5]]

summary_curves_norm <- tibble(
  vacc_week = x_range,
  med = exp(-k_med * (x_range - t0)),
  lo  = exp(-k_q75 * (x_range - t0)),  # faster decay => lower envelope
  hi  = exp(-k_q25 * (x_range - t0)),   # slower decay => higher envelope
  q975   = exp(-k_q975 * (x_range - t0)),  # faster decay => lower envelope
  q025  = exp(-k_q025 * (x_range - t0))   # slower decay => higher envelope
)

## ---- pretty label (normalized) ----
format_k <- paste0(round(k_med, digits = 3))
eq_label_norm <- paste0("$\\hat{z}(t)=\\exp\\{-", format_k, "(t-1)\\}$")

## ---- plot (normalized scale 0-1) ----
p2 <- ggplot() +
  # Individual normalized fitted curves
  geom_line(
    data = spaghetti_data_norm,
    aes(x = vacc_week, y = z_hat, group = id_outbreak, color = id_outbreak),
    alpha = 0.60
  ) +
  # Raw points, normalized by their own week-1
  geom_point(
    data = raw_data_points_norm,
    aes(x = vacc_week, y = z, color = id_outbreak),
    alpha = 0.60, size = 1
  ) +
  # Median + IQR summary on normalized scale
  geom_ribbon(
    data = summary_curves_norm,
    aes(x = vacc_week, ymin = lo, ymax = hi),
    inherit.aes = FALSE,
    fill = "firebrick",
    alpha = 0.30
  ) +
  geom_ribbon(
    data = summary_curves_norm,
    aes(x = vacc_week, ymin = q025, ymax = q975),
    inherit.aes = FALSE,
    fill = "firebrick",
    alpha = 0.20
  ) +
  geom_line(
    data = summary_curves_norm,
    aes(x = vacc_week, y = med),
    color = "firebrick",
    linewidth = 1.2
  ) +
  annotate(
    "latex",
    x = 15, y = 1,
    label = eq_label_norm,
    size = 4.5, hjust = 1, vjust = 1
  ) +
  # scale_y_continuous("Normalized reduction", limits = c(0, 1)) +
  scale_y_continuous("Normalized reduction") +
  guides(color = "none") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))


# plot 3
# Calculate proportion of outbreaks included
# outbreaks missed by delayed vaccination
wks_delay <- 1:15
dur_missed_id <- vector('list', length(wks_delay))
vax_missed_id <- vector('list', length(wks_delay))

for (i in 1:length(wks_delay)) {
  outbk_data_summary %>% 
    filter(duration_weeks <= wks_delay[i]) %>%
    pull(id_outbreak) -> dur_missed_id[[i]]
  
  cea %>% 
    filter(vacc_week == wks_delay[i] & runid == 1 & vacc_cov == 0.75) %>% 
    filter(!ori_occurred) %>% 
    pull(id_outbreak) -> vax_missed_id[[i]]
}
n_outbreaks_missed1 <- map_vec(dur_missed_id, function(i) length(i))
n_outbreaks_missed2 <- map_vec(vax_missed_id, function(i) length(i))

n_outbreaks <- nrow(outbk_data_summary)
prop_df <- data.frame(
  vacc_week = 1:15, 
  prop_included_dur = (1 - n_outbreaks_missed1 / n_outbreaks),
  prop_included_vax = (1 - n_outbreaks_missed2 / n_outbreaks))

df <- prop_df %>%
  select(vacc_week, prop_included_vax) %>%
  filter(vacc_week >= 4)

# --- fit linear model: percent-points per week ---
m <- lm(prop_included_vax ~ vacc_week, data = df)
co <- coef(m)
alpha <- unname(co[1])   # intercept
beta  <- unname(co[2])   # slope (expected to be negative)
r2    <- summary(m)$r.squared

# --- prediction grid & values ---
x_grid <- seq(min(df$vacc_week), max(df$vacc_week), length.out = 200)
pred   <- predict(m, newdata = data.frame(vacc_week = x_grid))

# --- a tidy equation label using LaTeX (percent scale) ---
# round for display
alpha_lab <- format(round(alpha, 2), nsmall = 2)
beta_lab  <- format(round(beta, 3), nsmall = 3)
r2_lab    <- format(round(r2, 3), nsmall = 3)

eq_label <- paste0(
  "$\\hat{y}(t) = ", alpha_lab, beta_lab, "\\, t$"
)

p3 <- ggplot() +
  # raw data (black)
  geom_point(
    data = prop_df,
    aes(vacc_week, prop_included_vax),
    color = "black", size = 2
  ) +
  # fitted line (firebrick)
  geom_line(
    aes(x_grid, pred),
    color = "firebrick", linewidth = 1.2
  ) +
  # equation label (LaTeX)
  annotate(
    "latex",
    x = 15, y = 1,
    label = eq_label,
    size = 4,
    hjust = 1, vjust = 1
  ) +
  # scale_x_continuous("Delay in ORI implementation (weeks)", 
  #                    breaks = seq(1, 15, 2), limits=c(1,15)) +
  scale_y_continuous("Proportion included") +
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0))

```

Combine plots

```{r}
hj <- 3.4
vj <- -0.3
sz <- 5

# A, B, and C annotations
p1_tag <- p1 +
  annotate("text", x = -Inf, y = Inf, label = "A",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2), limits=c(0.2,15.5))+
  coord_cartesian(clip = "off")

p2_tag <- p2 +
  annotate("text", x = -Inf, y = Inf, label = "B",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2), limits=c(0.2,15.5))+
  coord_cartesian(clip = "off")

p3_tag <- p3 +
  annotate("text", x = -Inf, y = Inf, label = "C",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Delay in ORI implementation (weeks)",
                     breaks=seq(1,15,2), limits=c(0.2,15.5))+
  coord_cartesian(clip = "off")

plt3 <- (p1_tag / p2_tag / p3_tag) + plot_layout(ncol=1) 

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig3_onset_triggered_", tstamp(), ".pdf"),
         plt3,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

## Figure 4

### Plots

Heatmap of cases averted per 1,000 OCVs

```{r}
capd_files <- c(
  "outputs/paper/cea_onset_trigger_10m_doses_ar_cutoff_20250922.rds",
  "outputs/paper/cea_onset_trigger_10m_doses_size_cutoff_20250922.rds",
  "outputs/paper/cea_onset_trigger_10m_doses_dur_cutoff_20250922.rds"
)
capd_data <- map(capd_files, read_rds)

# 1. Create the color palette function (your code was correct here)
myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))

# 3. Build the complete, polished plot
plts <- vector('list', length(capd_data))

y_labs <- c("Attack rate percentile", "Outbreak duration percentile", 
            "Outbreak size percentile")
y_annotations <- c("Prioritizing high AR outbreaks", 
                   "Prioritizing longer outbreaks",
                   "Prioritizing larger outbreaks")
  
p1  <- capd_data[[1]] %>% filter(vacc_cov == 0.75) %>%
  
  ggplot(aes(x = vacc_week, y = size_prob_cutoff, 
               fill = case_averted_per_1000_OCV_q500)) +
  geom_tile(color = "white", linewidth = 0.25) +
  scale_fill_gradientn(
    trans = scales::pseudo_log_trans(),
    colours = myPalette(100), # Generate a smooth ramp of 100 colors
    breaks = c(1, 2, 5, 10),
    limits = c(0.2, 15), 
    name = "Cases Averted\nper 1,000 OCVs" # Legend title
  ) +
  scale_y_continuous(y_labs[1],
    expand = c(0, 0), 
    breaks = seq(0, 0.8, 0.2),
    labels = c(expression(">="*P[0*"%"]),
               expression(">="*P[20*"%"]),
               expression(">="*P[40*"%"]),
               expression(">="*P[60*"%"]),
               expression(">="*P[80*"%"]))) +
    
  scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    # 1. Add the arrows
    #    The origin point for all arrows is (x=9, y=0)
  annotate(
    "segment",
    x = 9, y = 0,
    xend = c(9, 1), # End points for the three arrows
    yend = c(0.9, 0),
    arrow = arrow(length = unit(0.3, "cm")),
    color = "black",
    linewidth = 0.4
  ) +
  # 2. Add the four points on top of the arrows
  #    Using shape 21 allows for a fill and a border color
  annotate(
    "point",
    x = c(9), # The four key coordinates
    y = c(0),
    size = 3,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  annotate(
    "segment",
    x = 8, y = 0.5,
    xend = c(8, 1), # End points for the three arrows
    yend = c(0.9, 0.5),
    arrow = arrow(length = unit(0.3, "cm")),
    color = "black",
    linewidth = 0.4,
    linetype = "dashed" # Changed to dashed lines
  ) +
  annotate(
    "point",
    x = c(8), # The four key coordinates
    y = c(0.5),
    size = 3,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  # 3. Add text labels next to the arrows
  #    Coordinates are chosen to be midway along the arrow, with a slight offset
  annotate(
    "text",
    x = c(9.5, 5), # x-positions for the two labels
    y = c(0.5, 0.05), # y-positions for the two labels
    label = c(y_annotations[1], "Reducing ORI delay"),
    angle = c(90, 0), # Rotate the vertical label
    color = "black",
    # fontface = "bold",
    size = 4,
    hjust = 0.5
  ) +
  # labs(
  #   x = "vaccination week<br>(*slower Implementation →*)",
  #   y = "attack rate % cutoff<br>(*prioritizing high AR outbreaks →*)"
  # ) +
  labs(
    x = NULL,
    y = y_labs[i]
  ) +
  theme_light() +
  theme_tight() +
  theme(
    legend.position = "top",
    # plot.margin = margin(0,0.03,0,0)
  ) + 
  theme_x_blank()
 

p2  <- 
    capd_data[[2]] %>% filter(vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = size_prob_cutoff, 
               fill = case_averted_per_1000_OCV_q500)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = c(1, 2, 5, 10),
      limits = c(0.2, 15), 
      name = "Cases Averted\nper 1,000 OCVs" # Legend title
    ) +
    scale_y_continuous(y_labs[2], 
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = c(expression(">="*P[0*"%"]),
                 expression(">="*P[20*"%"]),
                 expression(">="*P[40*"%"]),
                 expression(">="*P[60*"%"]),
                 expression(">="*P[80*"%"]))) +
    
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    # 1. Add the arrows
    #    The origin point for all arrows is (x=9, y=0)
    labs( x = NULL, y = y_labs[2]) +
    guides(fill='none')+
    theme_light() +
    theme_tight() +
    theme_x_blank()
   

p3  <- capd_data[[3]] %>% filter(vacc_cov == 0.75) %>%
  ggplot(aes(x = vacc_week, y = size_prob_cutoff, 
             fill = case_averted_per_1000_OCV_q500)) +
  geom_tile(color = "white", linewidth = 0.25) +
  scale_fill_gradientn(
    trans = scales::pseudo_log_trans(),
    colours = myPalette(100), # Generate a smooth ramp of 100 colors
    breaks = c(1, 2, 5, 10),
    limits = c(0.2, 15), 
    # Handles values outside the limits gracefully
    # oob = scales::squish, 
    name = "Cases Averted\nper 1,000 OCVs" # Legend title
  ) +
  scale_y_continuous(y_labs[3],
    expand = c(0, 0), 
    breaks = seq(0, 0.8, 0.2),
    labels = c(expression(">="*P[0*"%"]),
               expression(">="*P[20*"%"]),
               expression(">="*P[40*"%"]),
               expression(">="*P[60*"%"]),
               expression(">="*P[80*"%"]))) +
  
  scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
  labs(
    x = "Delay in ORI implementation (weeks)",
    y = y_labs[3]
  ) +
  guides(fill = "none")+
  theme_light() +
  theme_tight()
```

Combine plots

```{r}
hj <- 3.4
vj <- -0.3
sz <- 5

# A, B, and C annotations
p1_tag <- p1 +
  annotate("text", x = -Inf, y = Inf, label = "A",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

p2_tag <- p2 +
  annotate("text", x = -Inf, y = Inf, label = "B",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

p3_tag <- p3 +
  annotate("text", x = -Inf, y = Inf, label = "C",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Delay in ORI implementation (weeks)",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

plt4 <- (p1_tag / p2_tag / p3_tag) + plot_layout(ncol=1) 

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig4_capd_10m_prioritization_", tstamp(), ".pdf"),
         plt4,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

## Figures 3 & 4

```{r}
hj <- 3.4
vj <- -0.3
sz <- 5

# A, B, and C annotations
p1_tag <- p1 +
  annotate("text", x = -Inf, y = Inf, label = "D",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

p2_tag <- p2 +
  annotate("text", x = -Inf, y = Inf, label = "E",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

p3_tag <- p3 +
  annotate("text", x = -Inf, y = Inf, label = "F",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Delay in ORI implementation (weeks)",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

plt4 <- (p1_tag / p2_tag / p3_tag) + plot_layout(ncol=1) 


plts_3_4 <- (plt3 | plt4) + plot_layout(ncol=2) + 
  plot_layout(axis_titles = "collect_x")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/figs_3_4_pcr_capd_", tstamp(), ".pdf"),
         plts_3_4,
         device = cairo_pdf, width = fgw2*fac*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

## Figure S3: Case Reduction in Case-Triggered Scenario

### Codes to produce the values

```{r}
res <- cea_ct %>%
  filter(ori_occurred) %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1, 9), case_trigger %in% c(1,10,100)) %>%
  group_by(case_trigger, vacc_week) %>%
  summarize_quantiles(pct_reduc_case)

# cases averted for vaccine coverage of 75%
format_median_iqr(res, row=c(1,5))
format_median_iqr(res, row=c(2,6))

cea_ct %>%
  filter(ori_occurred) %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1, 9), case_trigger %in% c(1,10,100)) %>%
  group_by(case_trigger, vacc_week) %>%
  summarize(max_val = max(pct_reduc_case))

```

### Left panel
 
```{r}
# Calculate proportion of outbreaks included
# outbreaks missed by delayed vaccination
wks_delay <- VW_seq
# ct <- c(1, 5, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)
ct_values <- c(1, 10, 100)

vax_missed_id <- vector("list", length(wks_delay))

for (w in seq_along(wks_delay)) {
  # Inner list: one element per ct
  vax_missed_id[[w]] <- vector("list", length(ct_values))
  
  for (i in seq_along(ct_values)) {
    vax_missed_id[[w]][[i]] <- cea_ct %>%
      filter(vacc_week == wks_delay[w] & runid == 1 & vacc_cov == 0.75) %>%
      filter(case_trigger == ct_values[i]) %>%
      filter(!ori_occurred) %>%
      pull(id_outbreak)
  }
}

# Count number of missed outbreaks for each [wks_delay, ct] pair
n_outbreaks_missed <- map(vax_missed_id, function(inner_list) {
  map_int(inner_list, length)   # get length of each vector inside
})

# Total number of outbreaks
n_outbreaks <- nrow(outbk_data_summary)

# Build a data frame in long format
prop_df <- map2_df(
  n_outbreaks_missed,
  wks_delay,
  function(inner_counts, delay_val) {
    tibble(
      wks_delay = delay_val,
      case_trigger = ct_values,
      prop_included_vax = 100 * (1 - inner_counts / n_outbreaks)
    )
  }
)

# description in the main text
prop_df %>% filter(wks_delay == 9, case_trigger %in% c(1,100)) %>% 
  mutate(prop_missed = 100 - prop_included_vax)

p1 <- cea_ct %>% 
  filter(ori_occurred == TRUE) %>%
  filter(vacc_cov == VC, vacc_week %in% wks_delay, case_trigger %in% ct_values) %>%
  ggplot(aes(x = as.factor(vacc_week), 
             y = pct_reduc_case,
             color = as.factor(case_trigger))) +
  # 95% central range (thin line + median point)
  stat_summary(
    fun = median,
    fun.min = function(z) quantile(z, 0.025),
    fun.max = function(z) quantile(z, 0.975),
    geom = "pointrange",
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thick line only, no point)
  stat_summary(
    fun.min = function(z) quantile(z, 0.25),
    fun.max = function(z) quantile(z, 0.75),
    geom = "errorbar",
    position = position_dodge(width = line_interval),
    linewidth = line_width_50,
    width = 0  # vertical line only, no horizontal cap
  ) +
  # Mean marked with bold red X
  stat_summary(
    aes(group = as.factor(case_trigger), color = as.factor(case_trigger)),
    fun = mean,
    geom = "point",
    shape = 4,        # X shape
    size = x_size,
    stroke = x_stroke,
    position = position_dodge(width = line_interval)
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:6],
    labels = c("1 case", "10 cases", "100 cases")) +
  labs(x = NULL, y = "% cases averted") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))

# Bottom plot: proportion of outbreaks included
p2 <- prop_df %>% 
  ggplot(aes(x = as.factor(wks_delay), y = prop_included_vax, 
             group = as.factor(case_trigger), color = as.factor(case_trigger))) +
  geom_line() +
  geom_point() +
  # geom_col(fill="steelblue") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(x = "Delay in ORI implementation (weeks) after the trigger",
       y = "% included") +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:6],
    labels = c("1 case", "10 cases", "100 cases")) +
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0))

cea_ct %>%
  filter(ori_occurred) %>% 
  filter(vacc_cov %in% VC, vacc_week %in% wks_delay, case_trigger %in% c(1,100)) %>% 
  group_by(vacc_week, case_trigger) %>% 
  summarize_quantiles(pct_reduc_case)
```

### Right panel

Plot decay
```{r}
# weekly decline
res <- read_rds("outputs/fit_decay_model_20250903.rds")

# res <- read_rds("outputs/fit_decay_model_case_trigger_cut_wk_5_20250822.rds")
```

```{r}
# across weeks for case thresholds of 1, 10, and 100
# res <- read_rds("outputs/fit_decay_model_case_trigger_20250811.rds")
res <- read_rds("outputs/fit_decay_model_case_trigger_cut_wk_5_20250822.rds")
# # across case thresholcs for weeks of delay of 1, 5, 9, 13
# res2 <- read_rds("outputs/fit_decay_model_case_trigger_20250809.rds")
# sum(res2$`1`$k_hat == 0)
# sum(res2$`5`$k_hat == 0)
# sum(res2$`9`$k_hat == 0)
# sum(res2$`13`$k_hat == 0)

res_df <- rbindlist(res) 

# number of outbreaks that were analyzed are different by case trigger thresholds
res_df %>% 
  group_by(case_trigger) %>% 
  filter(runid == 1) %>% 
  count() %>% 
  mutate(prop_included = n / nrow(outbk_data_summary))

khat <- res_df %>% 
  # filter(id_outbreak %in% ids) %>% 
  mutate(rate_reduc = 100*(1 - exp(-k_hat))) %>%
  group_by(case_trigger) %>% 
  summarize_quantiles(rate_reduc) # across runid's

# number of outbreaks that were analyzed are different
khat %>% group_by(case_trigger) %>% tally()

df <- res_df %>%
  mutate(rate_reduc = 100*(1 - exp(-k_hat))) %>%  # compute first
  group_by(case_trigger, id_outbreak) %>%
  summarize_quantiles(rate_reduc, half_life) %>%
  ungroup() %>%
  select(variable, id_outbreak, case_trigger, q500) %>%
  pivot_wider(names_from = variable, values_from = q500) %>% 
  select(c("id_outbreak", "case_trigger", "rate_reduc", "half_life")) %>%
  left_join(outbk_data_summary[, c("id_outbreak", "ar_per_10e3", 
                                   "duration_weeks", "spatial_scale", 
                                   "rural_urban", "pop", "total_s_ch")], 
            by = "id_outbreak")

names(df) <- c(names(df)[1:(ncol(df)-1)], "s_ch_tot")

ct_values <- unique(df$case_trigger)

lst <- vector("list", length(ct_values)) 
vars <- c("rate_reduc", "half_life")

lst_ids <- vector("list", length(ct_values)) 

for (i in 1:length(ct_values)) {
  ct <- ct_values[i]
  
  df_ct <- df %>% filter(case_trigger == ct)
  qn <- floor(0.25*nrow(df_ct))
  
  lst_ids[[i]] <- df_ct %>% pull(id_outbreak)
   
  top_size_ids <- df %>% filter(case_trigger == ct) %>% 
    arrange(desc(s_ch_tot)) %>% slice(1:qn) %>% pull(id_outbreak)
  top_ar_ids <- df %>% filter(case_trigger == ct) %>% 
    arrange(desc(ar_per_10e3)) %>% slice(1:qn) %>% pull(id_outbreak)
  top_dur_ids <- df %>% filter(case_trigger == ct) %>% 
    arrange(desc(duration_weeks)) %>% slice(1:qn) %>% pull(id_outbreak)
  
  all <- df %>%
    filter(case_trigger == ct) %>% 
    summarize_group(vars = vars, 
                    group_name = "All")
  top_ar <- df %>%
    filter(case_trigger == ct) %>% 
    filter(id_outbreak %in% top_ar_ids) %>%
    summarize_group(vars = vars, 
                    group_name = "Top AR")
  top_dur <- df %>% 
    filter(case_trigger == ct) %>% 
    filter(id_outbreak %in% top_ar_ids) %>%
    summarize_group(vars=vars, group_name = "Top duration")
  
  top_size <- df %>%
    filter(case_trigger == ct) %>%
    filter(id_outbreak %in% top_size_ids) %>% 
    summarize_group(vars=vars, group_name = "Top size")
  
  dfall <- rbindlist(list(all, top_ar, top_dur, top_size))
  dfall$case_trigger <- ct
  lst[[i]] <- dfall
}

# case trigger threshold of 1
# duration >=9 week 
for (i in c(1,3)) print(format_median_iqr(lst[[i]], row=c(1,3,5,7), digits=1))


plot_df <- rbindlist(lst)

plot_df %>% 
  filter(variable == "rate_reduc", group == "All") %>%
  format_median_iqr(row=c(1,3), digits=1)

# % decline per week by group
plot_df %>% 
  filter(variable == "rate_reduc", case_trigger == 1) %>%
  group_by(group) %>% 
  format_median_iqr(row=1:4, digits=1)

# to ensure that the same set of outbreaks
# outbreaks that were included in case_trigger == 100

df_ct100 <- df %>% filter(id_outbreak %in% lst_ids[[3]])
           
lst <- vector("list", length(ct_values)) 

for (i in 1:length(ct_values)) {
  ct <- ct_values[i]
  
  df_ct <- df_ct100 %>% filter(case_trigger == ct)
  qn <- floor(0.25*nrow(df_ct))
  
  top_size_ids <- df_ct100 %>% filter(case_trigger == ct) %>% 
    arrange(desc(s_ch_tot)) %>% slice(1:qn) %>% pull(id_outbreak)
  top_ar_ids <- df_ct100 %>% filter(case_trigger == ct) %>% 
    arrange(desc(ar_per_10e3)) %>% slice(1:qn) %>% pull(id_outbreak)
  top_dur_ids <- df_ct100 %>% filter(case_trigger == ct) %>% 
    arrange(desc(duration_weeks)) %>% slice(1:qn) %>% pull(id_outbreak)
  
  all <- df_ct100 %>%
    filter(case_trigger == ct) %>% 
    summarize_group(vars = vars, 
                    group_name = "All")
  top_ar <- df_ct100 %>%
    filter(case_trigger == ct) %>% 
    filter(id_outbreak %in% top_ar_ids) %>%
    summarize_group(vars = vars, 
                    group_name = "Top AR")
  top_dur <- df_ct100 %>% 
    filter(case_trigger == ct) %>% 
    filter(id_outbreak %in% top_ar_ids) %>%
    summarize_group(vars=vars, group_name = "Top duration")
  
  top_size <- df_ct100 %>%
    filter(case_trigger == ct) %>%
    filter(id_outbreak %in% top_size_ids) %>% 
    summarize_group(vars=vars, group_name = "Top size")
  
  dfall <- rbindlist(list(all, top_ar, top_dur, top_size))
  dfall$case_trigger <- ct
  lst[[i]] <- dfall
}

plot_df_sub <- rbindlist(lst)

plot_df %>% filter(variable == "rate_reduc", group == "All") %>% 
  format_median_iqr(row=c(1,2,3),digits=1) 


plot_df_sub %>% filter(variable == "rate_reduc", group == "All")
plot_df_sub %>% filter(variable == "rate_reduc", group == "All") %>% 
  format_median_iqr(row=c(1,2,3),digits=1) 

plot_df_sub %>% filter(variable == "rate_reduc", case_trigger == 1)
plot_df_sub %>% filter(variable == "rate_reduc", case_trigger == 100)

plot_df_sub %>% filter(variable == "half_life", group == "All") 
plot_df %>% filter(variable == "half_life", group == "All") %>% 
  format_median_iqr(row=c(1,2,3), digits=1) 
plot_df_sub %>% filter(variable == "half_life", group == "All") %>% 
  format_median_iqr(row=c(1,3), digits=1) 


plot_df_sub %>% filter(variable == "half_life", case_trigger == 1)
plot_df_sub %>% filter(variable == "half_life", case_trigger == 1) %>% 
  format_median_iqr(row=1:4, digits=1) 

plot_df_sub %>% filter(variable == "half_life", case_trigger == 100)
  
# p3 <- plot_df %>% 
p3 <- plot_df_sub %>% 
  filter(variable == "rate_reduc") %>%
  ggplot(aes(x = as.factor(case_trigger), color = group)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = group, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_color_manual("", 
          values = c("All" = "steelblue", 
                     "Top AR" = "purple", 
                     "Top duration" = "firebrick", 
                     "Top size" = "forestgreen"), 
          labels = c("All", "Top AR", "Top duration", "Top size")) +
  labs(y = "Weekly % decline in cases averted", x = NULL) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_light() +
  theme_tight() +
  theme_x_blank() + 
  theme(plot.margin = margin(0,0,0,0)) 

p3

# p4 <- plot_df %>% 
p4 <- plot_df_sub %>%   
  filter(variable == "half_life") %>%
  ggplot(aes(x = as.factor(case_trigger), color = group)) +
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = group, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_color_manual("", 
          values = c("All" = "steelblue", 
                     "Top AR" = "purple", 
                     "Top duration" = "firebrick", 
                     "Top size" = "forestgreen"), 
          labels = c("All", "Top AR", "Top duration", "Top size")) +
  labs(y = "Halving time (weeks)", x = "Case trigger threshold") +
  scale_y_continuous() +
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0)) 
p4
```

#### Combine panels
```{r}
# Combine left column plots with equal heights for a,b,c
save_plot <- TRUE

left_col <- p1 / p2 + 
  plot_layout(guides = "collect", heights = c(3, 1)) & theme(legend.position = "top")
left_col <- left_col + plot_layout()

# Combine right column plots first with equal heights
right_col <- (p3 / p4) + plot_layout(guides = "collect") & theme(legend.position = "top")

# Now bind the two columns side-by-side
final_plot <- (left_col | right_col) + plot_layout(widths = c(1, 1)) +
  plot_annotation(tag_levels = 'A') +
  theme(plot.tag.position = c(0.05, 1),
        plot.tag = element_text(size = 13))

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig4_case_triggered_", tstamp(), ".pdf"),
         final_plot,
         device = cairo_pdf, width = fgw2*fac,
         height = fgw2*fac, units = "mm")
}
```

# Aggregated Outbreaks

## Figure S4. OCV Impact aggregated across multiple oubreaks

```{r fig4, eval=FALSE}
# subsets with varying size
# cumulative (i.e., q1 = all, q2 = all except the first quartile)
cea_subsets_by_size_q1 <-
  read_rds("outputs/cea_onset_trigger_subsets_0_by_size_20250827.rds")
cea_subsets_by_size_q2 <-
  read_rds("outputs/cea_onset_trigger_subsets_0.25_by_size_20250827.rds")
cea_subsets_by_size_q3 <-
  read_rds("outputs/cea_onset_trigger_subsets_0.5_by_size_20250827.rds")
cea_subsets_by_size_q4 <-
  read_rds("outputs/cea_onset_trigger_subsets_0.75_by_size_20250827.rds")

cea_subsets_by_size_q3 %>%
  filter(vacc_cov == 0.75 & sample_size %in% c(1,10,100) &
           vacc_week %in% c(9)) %>% 
  group_by(sample_size, vacc_week, vacc_cov) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> pcr1

format_median_iqr(pcr1, row=1:3)

cea_subsets_by_size_q3 %>%
  filter(vacc_cov == 0.75 & sample_size %in% c(1,10,100) &
           vacc_week %in% c(9)) %>% 
  group_by(sample_size, vacc_week, vacc_cov) %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) -> capd1

format_median_iqr(capd1, row=1:3, digits=3)


plt5a <- cea_subsets_by_size_q4 %>% 
  filter(vacc_cov == 0.75 & vacc_week %in% seq(1,15,2) & 
           sample_size %in% c(1, 10, 50)) %>% 
  ggplot(aes(x = as.factor(vacc_week), y = pct_reduc_case, 
             color = as.factor(sample_size))) +
  # 95% central range (thin line + median point)
  stat_summary(
    fun = median,
    fun.min = function(z) quantile(z, 0.025),
    fun.max = function(z) quantile(z, 0.975),
    geom = "pointrange",
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thick line only, no point)
  stat_summary(
    fun.min = function(z) quantile(z, 0.25),
    fun.max = function(z) quantile(z, 0.75),
    geom = "errorbar",
    position = position_dodge(width = line_interval),
    linewidth = line_width_50,
    width = 0  # vertical line only, no horizontal cap
  ) +
  # Mean marked with bold red X
  stat_summary(
    aes(group = as.factor(sample_size), color = as.factor(sample_size)),
    fun = mean,
    geom = "point",
    shape = 4,        # X shape
    size = x_size,
    stroke = x_stroke,
    position = position_dodge(width = line_interval)
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:6],
    labels = c("1 outbreak", "10 outbreaks", "100 outbreaks")) +
  labs(x = "Delay in ORI implementation (week)", y = "% cases averted") +
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0))


plt5b <- cea_subsets_by_size_q4 %>%
  filter(vacc_cov == 0.75 & vacc_week %in% seq(1,15,2) &
           sample_size %in% c(1, 10, 50)) %>%
  ggplot(aes(x = as.factor(vacc_week), y = case_averted_per_1000_OCV,
             color = as.factor(sample_size))) +
  stat_summary(
    fun = median,
    fun.min = function(z) quantile(z, 0.025),
    fun.max = function(z) quantile(z, 0.975),
    geom = "pointrange",
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thick line only, no point)
  stat_summary(
    fun.min = function(z) quantile(z, 0.25),
    fun.max = function(z) quantile(z, 0.75),
    geom = "errorbar",
    position = position_dodge(width = line_interval),
    linewidth = line_width_50,
    width = 0  # vertical line only, no horizontal cap
  ) +
  # Mean marked with bold red X
  stat_summary(
    aes(group = as.factor(sample_size), color = as.factor(sample_size)),
    fun = mean,
    geom = "point",
    shape = 4,        # X shape
    size = x_size,
    stroke = x_stroke,
    position = position_dodge(width = line_interval)
  ) +
  scale_y_continuous(limits=c(0,3)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:6],
    labels = c("1 outbreak", "10 outbreaks", "100 outbreaks")) +
  guides(color = "none") +
  labs(x = "Delay in ORI implementation (week)", 
       y = "Cases averted per 1,000 OCVs") + 
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0))

# Combine all four plots into a single figure
plt5 <- (plt5a + plt5b) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              nrow = 1,
              guides = "collect") &
  theme(plot.tag.position = c(0.05, 0.92),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig5_ORI_impact_aggregated_", tstamp(), ".pdf"),
         plt5, device = cairo_pdf,  width = fgw2*fac, 
         height = fgw1*1.15*fac, units = "mm")
}

rm(cea_subsets_by_size_q1, cea_subsets_by_size_q2, 
     cea_subsets_by_size_q3, cea_subsets_by_size_q4)
```

## Figure 4: Cumulative OCV Impact during OCV stockpile deployment

```{r fig4, eval=FALSE}
# Load subset analyses reflecting global OCV stockpile
cea_stockpile_subsets_q1 <-
  read_rds("outputs/cea_stockpile_0_onset_trigger_by_year_20250828.rds")
cea_stockpile_subsets_q2 <-
  read_rds("outputs/cea_stockpile_0.25_onset_trigger_by_year_20250828.rds")
cea_stockpile_subsets_q3 <-
  read_rds("outputs/cea_stockpile_0.5_onset_trigger_by_year_20250828.rds")
cea_stockpile_subsets_q4 <-
  read_rds("outputs/cea_stockpile_0.75_onset_trigger_by_year_20250828.rds")
# Reference values for individual assessment
cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(s_ch_averted_tot) 

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(case_averted_per_1000_OCV)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(cost_per_daly_averted)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(cost_per_case_averted)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week) |>
  summarize(vacc_cov = vacc_cov[1],
          count = n(),
          median_prc = median(pct_reduc_case, na.rm=T),
          median_capd = median(case_averted_per_1000_OCV, na.rm=T),
          median_cpca = median(cost_per_case_averted, na.rm=T)) -> ref_val

# subsets for plotting

cea_stockpile_subsets_q1 %>% 
  group_by(year_max) %>%
  summarize(cumul_case_averted = median(s_ch_averted_tot))

cea_stockpile_subsets_q2 %>% 
  group_by(year_max) %>%
  summarize(cumul_case_averted = median(s_ch_averted_tot))

cea_stockpile_subsets_q3 %>% 
  group_by(year_max) %>%
  summarize(cumul_case_averted = median(s_ch_averted_tot))

cea_stockpile_subsets_q4 %>% 
  group_by(year_max) %>%
  summarize(cumul_case_averted = median(s_ch_averted_tot))

d1 <- cea_stockpile_subsets_q1
d1$group <- "Q1"
d2 <- cea_stockpile_subsets_q2
d2$group <- "Q2"
d3 <- cea_stockpile_subsets_q3
d3$group <- "Q3"
d4 <- cea_stockpile_subsets_q3
d4$group <- "Q4"
rm(cea_stockpile_subsets_q1, cea_stockpile_subsets_q2, 
   cea_stockpile_subsets_q3, cea_stockpile_subsets_q4)

stockpile_impact_data <- rbindlist(list(d1,d2,d3,d4))

rm(d1,d2,d3,d4)
# sample to speed up plotting  
set.seed(42)
subset_ids <- sample(1:200, 30)
plot_data <- stockpile_impact_data[subset_id %in% subset_ids,]
# cases averted
plot_data %>%
  filter(vacc_cov == 0.75 & vacc_week == 9 & year_max == 2020) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(variable=s_ch_averted_tot) 

# percent case reduction
plot_data %>%
  filter(vacc_cov == 0.75 & vacc_week == 9 & year_max == 2020) %>% 
  summarize_quantiles(variable=pct_reduc_case) %>% 
  format_median_interval(interval = "IQR")

# case averted per 1,000 doses
plot_data %>%
  filter(vacc_cov == 0.75 & vacc_week == 9 & year_max == 2020) %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) %>% 
  format_median_interval(interval = "IQR")

plot_data %>%
  group_by(subset_id) %>% 
  summarize(max_id = max(runid))
  
# weekly % reduction in percentage case reduction
plot_data_diff <- plot_data %>%
  group_by(subset_id, runid, vacc_cov) %>%
  mutate(pct_reduc_case_frac = c(1-frac_change(pct_reduc_case, lag=1)),
         pct_reduc_case_diff = c(1, -diff(pct_reduc_case, lag=1)))

weekly_reduc <- plot_data_diff %>%
  # filter(subset_id %in% 1:10) %>% 
  filter(vacc_cov == 0.75 & vacc_week > 1 & vacc_week <= 15 & year_max == 2020) %>% 
  summarize_quantiles(variable=pct_reduc_case_frac*100)

# case averted per 1,000 OCVs
plot_data %>%
  filter(vacc_cov == 0.75 & vacc_week == 9 & year_max == 2020) %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) %>% 
  format_median_iqr(digits=2)

# cost ase averted per 1,000 OCVs
plot_data %>%
  filter(vacc_cov == 0.75 & vacc_week == 9 & year_max == 2020) %>% 
  summarize_quantiles(variable=cost_per_case_averted) %>% 
  format_median_iqr(digits=0)

# (A) OCV deployment over time# (A) Oq500CV deployment over time
plt_a1 <- ocv_year |>
  filter(year <= 2020) |> 
  ggplot(aes(x = as.factor(year), y = cum_num_ocv)) +
  geom_point() +
  labs(x = "", y = "Cumulative\nOCVs deployed") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))


# (B) Cases averted
plt_b1 <- plot_data |>
  ggplot(aes(x = as.factor(year_max), y = s_ch_averted_tot, 
             color=group)) +
  ci_layers(
    group = group, color = group,
    dodge_width = line_interval,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = point_size,
    mean_size = x_size,
    mean_stroke = x_stroke
  )+
  scale_y_continuous() +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:7],
    labels = c("All", "Top 75%", "Top 50%", "Top 25%")) +
  labs(x = NULL, y = "Cumulative cases averted") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))

  
# (C) Percent case reduction
plt_c1 <- plot_data |>
  ggplot(aes(x = as.factor(year_max), y = pct_reduc_case, color = group)) +
  ci_layers(
    group = group, color = group,
    dodge_width = line_interval,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = point_size,
    mean_size = x_size,
    mean_stroke = x_stroke
  )+
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:7],
    labels = c("All", "Top 75%", "Top 50%", "Top 25%")) +
  labs(x = NULL, y = "% cases averted") +
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0))
  
# (D) Cases averted per 1000 OCVs
plt_d1 <- plot_data |>
  ggplot(aes(x = as.factor(year_max), y = case_averted_per_1000_OCV, 
             color = group)) +
  ci_layers(
    group = group, color = group,
    dodge_width = line_interval,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = point_size,
    mean_size = x_size,
    mean_stroke = x_stroke
  )+
  geom_hline(data = ref_val, aes(yintercept = median_capd), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_continuous(limits = c(0,3)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:7],
    labels = c("All", "Top 75%", "Top 50%", "Top 25%")) +
  labs(x = NULL, y = "Cases averted\nper 1,000 OCVs") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))

# (E) Cost per case averted
plt_e1 <- plot_data |>
  ggplot(aes(x = as.factor(year_max), y = cost_per_case_averted, 
             color = group)) +
  ci_layers(
    group = group, color = group,
    dodge_width = line_interval,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = point_size,
    mean_size = x_size,
    mean_stroke = x_stroke
  )+
  scale_y_log10() +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:7],
    labels = c("All", "Top 75%", "Top 50%", "Top 25%")) +
  labs(x = "", y = "Cost per case\naverted (US$)") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))

# (F) Cost per DALY averted
plt_f1 <- plot_data |>
  ggplot(aes(x = as.factor(year_max), y = cost_per_daly_averted, 
             color = group)) +
  ci_layers(
    group = group, color = group,
    dodge_width = line_interval,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = point_size,
    mean_size = x_size,
    mean_stroke = x_stroke
  )+
  scale_y_log10() +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:7],
    labels = c("All", "Top 75%", "Top 50%", "Top 25%")) +
  labs(x = "", y = "Cost per DALY\naverted (US$)") +
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0))

# Combine all panels into a single figure
plt6 <- wrap_plots(list(plt_a1, plt_d1, plt_b1, plt_e1,
                        plt_c1, plt_f1), ncol = 2,
                   guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0.07, 0.88),
        plot.tag = element_text(size = 12),
        legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig6_cumulative_impact_", tstamp(), ".pdf"),
         plt6, device = cairo_pdf, width=fgw2*fac*1.2, 
         height=fgw2*fac, units="mm")
}

stockpile_impact_data %>%
  filter(group %in% c("Q1","Q4")) %>%
  filter(vacc_week == 9, vacc_cov == 0.9, year_max %in% c(2020)) %>%
  group_by(group) %>%
  summarize_quantiles(variable=s_ch_averted_tot) %>%
  format_median_iqr(row=c(1,2), digits=1)

stockpile_impact_data %>%
  filter(group == "Q1") %>%
  filter(vacc_week == 9, vacc_cov == 0.9, year_max %in% c(2020)) %>%
  group_by(year_max) %>%
  summarize_quantiles(variable=pct_reduc_case) %>% 
  format_median_iqr(row=1)

stockpile_impact_data %>%
  filter(group == "Q1") %>%
  filter(vacc_week == 9, vacc_cov == 0.9, year_max %in% c(2020)) %>%
  group_by(year_max) %>%
  summarize_quantiles(variable=case_averted_per_1000_OCV) %>% 
  format_median_iqr(row=1, digits=0)

stockpile_impact_data %>%
  filter(group == "Q1") %>%
  filter(vacc_week == 9, vacc_cov == 0.9, year_max %in% c(2020)) %>%
  group_by(year_max) %>%
  summarize_quantiles(variable=cost_per_case_averted) %>%
  format_median_iqr(row=1, digits=0)

stockpile_impact_data %>%
  filter(group == "Q1") %>%
  filter(vacc_week == 9, vacc_cov == 0.9, year_max %in% c(2020)) %>%
  group_by(year_max) %>%
  summarize_quantiles(variable=cost_per_daly_averted) %>%
  format_median_iqr(row=1, digits=0)

cea %>%
  filter(vacc_week == 9, vacc_cov == 0.9) %>%
  summarize_quantiles(case_averted_per_1000_OCV)

cea %>%
  filter(vacc_week == 9, vacc_cov == 0.9) %>%
  summarize_quantiles(cost_per_daly_averted) %>%
  format_median_iqr(row=1, digits=0)
```

```{r fig4, eval=FALSE}
# Load subset analyses reflecting global OCV stockpile
cea_stockpile_subsets_q1 <-
  read_rds("outputs/cea_stockpile_0_onset_trigger_by_year_20250828.rds")
cea_stockpile_ar_subsets <-
  read_rds("outputs/cea_stockpile_ar_cutoff_onset_trigger_by_year_20250924.rds")
# Reference values for individual assessment

# subsets for plotting
cea_stockpile_subsets_q1 %>% 
  group_by(year_max) %>%
  summarize(cumul_case_averted = median(s_ch_averted_tot))

cea_stockpile_ar_subsets[["1"]] %>% 
  group_by(year_max) %>%
  summarize(cumul_case_averted = median(s_ch_averted_tot))

cea_stockpile_ar_subsets[["2"]] %>% 
  group_by(year_max) %>%
  summarize(cumul_case_averted = median(s_ch_averted_tot))

cea_stockpile_ar_subsets[["3"]] %>% 
  group_by(year_max) %>%
  summarize(cumul_case_averted = median(s_ch_averted_tot))

d1 <- cea_stockpile_subsets_q1
d1$group <- "AR0"
d2 <- cea_stockpile_ar_subsets[["1"]]
d2$group <- "AR1"
d3 <- cea_stockpile_ar_subsets[["2"]]
d3$group <- "AR2"
d4 <- cea_stockpile_ar_subsets[["3"]]
d4$group <- "AR3"
rm(cea_stockpile_subsets_q1, cea_stockpile_ar_subsets)

stockpile_impact_data <- rbindlist(list(d1,d2,d3,d4))
rm(d1,d2,d3,d4)

stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 9),] %>% 
  filter(group == "AR0") %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) %>% 
  format_median_interval(interval = "IQR", digits=2)

stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 1),] %>% 
  filter(group == "AR0") %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) %>% 
  format_median_interval(interval = "IQR", digits=2)

stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 9),] %>% 
  filter(group == "AR3") %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) %>% 
  format_median_interval(interval = "IQR", digits=4)

# overall impact of 38M doses
c(2.4524, 1.9078, 2.9837)*38000000/1000

stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 5),] %>% 
  filter(group == "AR3") %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) %>% 
  format_median_interval(interval = "IQR", digits=4)
c(4.7478, 3.7228, 5.7029)*38000000/1000

stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 1),] %>% 
  filter(group == "AR1") %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) %>% 
  format_median_interval(interval = "IQR", digits=2)



d1 <- stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 9 & group == "AR0"),] 
d1$plot_grp <- "Baseline"
d2 <- stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 9 & group == "AR3"),] 
d2$plot_grp <- "Higher AR"
d3 <- stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 1 & group == "AR0"),] 
d3$plot_grp <- "Faster ORI"

plot_data_all <- rbindlist(list(d1,d2,d3))
rm(d1,d2,d3)

# sample to speed up plotting  
set.seed(42)
id_sample <- sample(1:200, 30)
plot_data <- plot_data_all[subset_id %in% id_sample, ]

# grp_labels <- c("All", ">1 per 10,000", ">2 per 10,000", ">3 per 10,000")
grp_labels <- c("Baseline", "Faster ORI", "Higher AR")
# (A) OCV deployment over time# (A) Oq500CV deployment over time
plt_a <- ocv_year |>
  filter(year <= 2020) |> 
  ggplot(aes(x = as.factor(year), y = cum_num_ocv)) +
  geom_point() +
  labs(x = "", y = "Cumulative\nOCVs deployed") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))

# (c) Cases averted per 1000 OCVs
plt_b <- plot_data |>
  ggplot(aes(x = as.factor(year_max), y = case_averted_per_1000_OCV, 
             color = plot_grp)) +
  ci_layers(
    group = plot_grp, color = plot_grp,
    dodge_width = line_interval,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = point_size,
    mean_size = x_size,
    mean_stroke = x_stroke
  )+
  scale_y_continuous(limits = c(0,3)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[c(4,5,7)],
    labels = grp_labels) +
  labs(x = NULL, y = "Cases averted\nper 1,000 OCVs") +
  theme_light() +
  theme_tight() +
  # theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))


# Combine all panels into a single figure
plt3 <- wrap_plots(list(plt_a, plt_b), ncol = 1, guides = "collect") &
  theme(legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/stockpile_cumulative_impact_baseline1_", tstamp(), ".pdf"),
         plt3, device = cairo_pdf, width=fgw1*fac, 
         height=fgw1*fac*1.5, units="mm")
}

d1 <- stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 9 & group == "AR1"),] 
d1$plot_grp <- "Baseline"
d2 <- stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 9 & group == "AR3"),] 
d2$plot_grp <- "Higher AR"
d3 <- stockpile_impact_data[(vacc_cov == 0.75 & vacc_week == 1 & group == "AR1"),] 
d3$plot_grp <- "Faster ORI"

plot_data_all <- rbindlist(list(d1,d2,d3))
rm(d1,d2,d3)

# sample to speed up plotting  
set.seed(42)
id_sample <- sample(1:200, 30)
plot_data <- plot_data_all[subset_id %in% id_sample, ]

# grp_labels <- c("All", ">1 per 10,000", ">2 per 10,000", ">3 per 10,000")
grp_labels <- c("Baseline", "Faster ORI", "Higher AR")
# (A) OCV deployment over time# (A) Oq500CV deployment over time
plt_a <- ocv_year |>
  filter(year <= 2020) |> 
  ggplot(aes(x = as.factor(year), y = cum_num_ocv)) +
  geom_point() +
  labs(x = "", y = "Cumulative\nOCVs deployed") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))

# (c) Cases averted per 1000 OCVs
plt_b <- plot_data |>
  ggplot(aes(x = as.factor(year_max), y = case_averted_per_1000_OCV, 
             color = plot_grp)) +
  ci_layers(
    group = plot_grp, color = plot_grp,
    dodge_width = line_interval,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = point_size,
    mean_size = x_size,
    mean_stroke = x_stroke
  )+
  scale_y_continuous(limits = c(0,3)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[c(4,5,7)],
    labels = grp_labels) +
  labs(x = NULL, y = "Cases averted\nper 1,000 OCVs") +
  theme_light() +
  theme_tight() +
  # theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))


# Combine all panels into a single figure
plt3 <- wrap_plots(list(plt_a, plt_b), ncol = 1, guides = "collect") &
  theme(legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/stockpile_cumulative_impact_baseline2_", tstamp(), ".pdf"),
         plt3, device = cairo_pdf, width=fgw1*fac, 
         height=fgw1*fac*1.5, units="mm")
}
```

# Result Tables

## Table 3. Impact under the outbreak onset-triggered scenario

```{r table3}
# Create summary for time-triggered vaccination at different times

cea_ori[(vacc_cov == VC & vacc_week %in% c(1, 13))] %>% 
  group_by(vacc_week) %>% 
  summarize_quantiles(case_averted_per_1000_OCV) -> capd

format_median_iqr(capd, row=1, digits=3)
format_median_iqr(capd, row=2, digits=3)

cea_ori[(vacc_cov == VC & vacc_week %in% c(1, 13))] %>% 
  group_by(vacc_week) %>% 
  summarize_quantiles(cost_per_case_averted) -> cpca

format_median_iqr(cpca, row=1, digits=0)
format_median_iqr(cpca, row=2, digits=0)

##
wks <- c(1, 5, 9, 13)
tab <- data.frame()

for (wk in wks) {
  d <- cea_ori[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- nrow(d[d$runid == 1])
  tab <- rbind(tab, tab_temp)
}

# top quartile
df <- outbk_data_summary[, c("id_outbreak", "ar_per_10e3", 
                                   "duration_weeks", "spatial_scale", 
                                   "rural_urban", "pop", "total_s_ch")]

qn <- floor(0.25 * nrow(df))

top_ar_ids <- df %>% 
  arrange(desc(ar_per_10e3)) %>% slice(1:qn) %>% pull(id_outbreak)
top_dur_ids <- df %>% 
  arrange(desc(duration_weeks)) %>% slice(1:qn) %>% pull(id_outbreak)
top_size_ids <- df %>% 
  arrange(desc(total_s_ch)) %>% slice(1:qn) %>% pull(id_outbreak)

# top quartile in attack rate
dat <- cea_ori %>% filter(id_outbreak %in% top_ar_ids)
for (wk in wks) {
  d <- dat[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- nrow(d[d$runid == 1])
  tab <- rbind(tab, tab_temp) 
}

# top quartile in duration
dat <- cea_ori %>% filter(id_outbreak %in% top_dur_ids)
for (wk in wks) {
  d <- dat[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- nrow(d[d$runid == 1])
  tab <- rbind(tab, tab_temp)
}

# top quartile in size
dat <- cea_ori %>% filter(id_outbreak %in% top_size_ids)
for (wk in wks) {
  d <- dat[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- nrow(d[d$runid == 1])
  tab <- rbind(tab, tab_temp)
}


tab$week <- rep(wks, 4)
tab$group <- rep(c("All","Top 25% attack rate",
                    "Top 25% duration","Top 25% size"), each = 4)

# format number of outbreaks with percentages
tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/1192*100, digits=1), "%)") 

if (save_table) {
  fwrite(tab, paste0("outputs/table3_onset_triggered_impact_IQR_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```

## Table S1. Impact under the outbreak onset-triggered scenario

Broader vaccine coverage and ORI implementation delays

```{r tableS1}
# Create summary for time-triggered vaccination at different times
wks <- c(1, 5, 9, 13)
tab <- data.frame()

cea_ori <- cea[ori_occurred == TRUE]
for (vc in VC3) {
  for (wk in wks) {
    d <- cea_ori[(vacc_cov == vc & vacc_week == wk)]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- nrow(d[d$runid == 1])
    tab <- rbind(tab, tab_temp)
  }  
}
tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/1192*100, digits=1), "%)")

tab$week <- rep(wks, length(VC3))
tab$vc <- rep(VC3, each=length(wks))

if (save_table) {
  fwrite(tab, paste0("outputs/tableS1_onset_triggered_impact_IQR_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```

## Table S2. Impact under the case-triggered scenario

```{r}
# Create summary for case-triggered vaccination at different thresholds
vwks <- c(1, 5, 9, 13)
ctgs <- c(1, 10, 50, 100)
tab <- data.frame()

cea_ct_ori <- cea_ct[ori_occurred == TRUE]

for (wk in vwks) {
  # identify a set of outbreaks across case triggers 
  ids0 <- cea_ct %>% 
    filter(vacc_week == wk & runid == 1, case_trigger == max(ctgs)) %>% 
    filter(ori_occurred) %>% 
    pull(id_outbreak)
  
  for (ctg in ctgs) {
    d <- cea_ct_ori[(vacc_cov == VC & vacc_week == wk & 
                       case_trigger == ctg & id_outbreak %in% ids0)]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- nrow(d[d$runid == 1])
    tab <- rbind(tab, tab_temp)
  }
}

tab$case_trigger <- rep(ctgs, length(vwks))
tab$vacc_week <- rep(vwks, each=length(ctgs))

tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/1192*100, digits=1), "%)")

if (save_table) {
  fwrite(tab, paste0("outputs/table4_case_triggered_impact_IQR_", tstamp(), ".csv"),
            row.names = FALSE)
}
```

## Global OCV stockpile 

```{r tableS3}
cea_stockpile_subsets_q1 <-
  read_rds("outputs/cea_stockpile_0_onset_trigger_by_year_20250828.rds")
cea_stockpile_subsets_q2 <-
  read_rds("outputs/cea_stockpile_0.25_onset_trigger_by_year_20250828.rds")
cea_stockpile_subsets_q3 <-
  read_rds("outputs/cea_stockpile_0.5_onset_trigger_by_year_20250828.rds")
cea_stockpile_subsets_q4 <-
  read_rds("outputs/cea_stockpile_0.75_onset_trigger_by_year_20250828.rds")

# cea_stockpile_subsets_q2 is assumed to be more representative than q1, ..., q4 
cea_stockpile_subsets_q1 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(s_ch_averted_tot) %>% 
  format_median_iqr(digits=0)

# year 2020
cea_stockpile_subsets_q1 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(s_ch_averted_tot) %>% 
  format_median_iqr(digits=0, row=8)

cea_stockpile_subsets_q1 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(pct_reduc_case)

# year 2020
cea_stockpile_subsets_q4 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(s_ch_averted_tot) %>% 
  format_median_iqr(digits=0, row=8)

cea_stockpile_subsets_q4%>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(pct_reduc_case) %>% 
  format_median_iqr(digits=0, row=8)

cea_ori %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  summarize_quantiles(pct_reduc_case) %>% 
  format_median_iqr(digits=1)

cea_stockpile_subsets_q1 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(case_averted_per_1000_OCV) %>% 
  format_median_iqr(digits=3, row=8)

cea_stockpile_subsets_q4 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(case_averted_per_1000_OCV) %>% 
  format_median_iqr(digits=3, row=8)

cea_ori %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  summarize_quantiles(case_averted_per_1000_OCV) %>% 
  format_median_iqr(digits=3)

cea_stockpile_subsets_q1 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(cost_per_case_averted) %>% 
  format_median_iqr(digits=0, row=8)

cea_stockpile_subsets_q4 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(cost_per_case_averted) %>% 
  format_median_iqr(digits=0, row=8)

cea_ori %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  summarize_quantiles(cost_per_case_averted) %>% 
  format_median_iqr(digits=0)

cea_stockpile_subsets_q1 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(cost_per_daly_averted) %>% 
  format_median_iqr(digits=0, row=8)

cea_stockpile_subsets_q4 %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(cost_per_daly_averted) %>% 
  format_median_iqr(digits=0, row=8)

cea_ori %>%
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  summarize_quantiles(cost_per_daly_averted) %>% 
  format_median_iqr(digits=0)
#
```


## Table S3. Impact under the global OCV stockpile, 2013-2020, scenario

```{r tableS3}
wks <- c(1, 5, 9, 13)
probs <- c(0.25, 0.5, 0.75)

# Calculate summary statistics
tab1 <- data.frame()
for (vc in VC3) {
  for (wk in wks) {
    d <- cea_stockpile_subsets_q1[year_max == 2020 & vacc_week == wk & vacc_cov == vc]
    tab1 <- rbind(tab1, impact_summary(d, probs=probs))
  }  
}
tab1$week <- rep(wks, length(VC3))
tab1$vc <- rep(VC3, each=length(wks))

tab2 <- data.frame()
for (vc in VC3) {
  for (wk in wks) {
    d <- cea_stockpile_subsets_q2[year_max == 2020 & vacc_week == wk & vacc_cov == vc]
    tab2 <- rbind(tab2, impact_summary(d, probs=probs))
  }  
}
tab2$week <- rep(wks, length(VC3))
tab2$vc <- rep(VC3, each=length(wks))

# top 50% outbreaks 
tab3 <- data.frame()
for (vc in VC3) {
  for (wk in wks) {
    d <- cea_stockpile_subsets_q3[year_max == 2020 & vacc_week == wk & vacc_cov == vc]
    tab3 <- rbind(tab3, impact_summary(d, probs=probs))
  }  
}
tab3$week <- rep(wks, length(VC3))
tab3$vc <- rep(VC3, each=length(wks))

# top 25% outbreaks
tab4 <- data.frame()
for (vc in VC3) {
  for (wk in wks) {
    d <- cea_stockpile_subsets_q4[year_max == 2020 & vacc_week == wk & vacc_cov == vc]
    tab4 <- rbind(tab4, impact_summary(d, probs=probs))
  }  
}
tab4$week <- rep(wks, length(VC3))
tab4$vc <- rep(VC3, each=length(wks))

tab <- rbindlist(list(tab1, tab2, tab3, tab4))
tab$group <- rep(c("All","Top 75%", 
                   "Top 50%","Top 25%"), each = 12)


if (save_table) {
  write.csv(tab, paste0("outputs/tableS3_stockpile_outbreak_subsets_IQR_", 
                        tstamp(), ".csv"), 
            row.names = FALSE)
}
```


## Table S4: Sensitivity Analysis of the Impact of Outbreak Onset-Triggered Vaccination

```{r tableS1}
cea_0.75 <- read_rds("outputs/cea_0.75_onset_trigger_20250716.rds")
vc <- c(0.75) 
wks <- c(1, 5, 9, 13)

tab <- data.frame()
cea_0.75_ori <- cea_0.75[ori_occurred == TRUE]
for (wk in wks) {
  d <- cea_0.75_ori[(vacc_cov == vc & vacc_week == wk)]
  tab <- rbind(tab, impact_summary(d, probs=c(0.25, 0.5, 0.75)))
}

# impact under the scenario where ORIs DO occur even when outbreaks have ended 
for (wk in wks) {
  d <- cea_0.75[(vacc_cov == vc & vacc_week == wk)]
  tab <- rbind(tab, impact_summary(d, probs=c(0.25, 0.5, 0.75)))
}

outbreaks_alt <- read_rds("outputs/outbreaks_alt_included_20250912.rds")

deaexcl <- read_rds("outputs/outbreaks_excluded_20250716.rds")

cea_incl_excl <- rbindlist(list(cea, cea_excl))
cea_alt <- cea_incl_excl %>% filter(id_outbreak %in% outbreaks_alt$id_outbreak)
  
# Create summary for time-triggered vaccination at different times
 
wks <- c(1, 5, 9, 13)

# # impact under the scenario where ORIs DO NOT occur even when outbreaks have ended 
tab <- data.frame()

# impact under the scenario where ORIs DO occur even when outbreaks have ended 
for (vc in VC3) {
  for (wk in wks) {
    d <- cea_alt[(vacc_cov == vc & vacc_week == wk)]
    tab <- rbind(tab, impact_summary(d, probs=c(0.25, 0.5, 0.75)))
  }
}

tab$week <- rep(wks, 3)

if (save_table) {
  fwrite(tab, paste0("outputs/tableS4_outbreak_data_sens_anal_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```
