---
title: "Cholera Vaccine Impact Visualization and Tables"
author: "Jong-Hoon Kim"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
editor_options: 
  chunk_output_type: console
---

# Introduction

This document creates visualizations and tables to summarize the results from 
our cholera vaccine impact simulations. It processes results from different vaccination strategies (outbreak onset-triggered, case-triggered, and preemptive) and coverage levels, and produces figures and tables for the final report.

# Setup

```{r setup}
#| warning: false
#| message: false

# Load required packages
source("utils.R")  # utility functions
library(data.table)
library(tidyverse)
library(magrittr)
library(ggridges)
library(readxl)
library(janitor)
library(scales)
library(viridis)
library(sf)
library(metR)
library(googlesheets4)
library(MASS)
library(patchwork)
library(ggpattern)
library(gridExtra)
library(corrr)
library(RColorBrewer)
library(countrycode)
library(conflicted)
# Resolve package conflicts in favor of dplyr
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# Output control flags
save_plot <- FALSE
save_table <- FALSE

# Configure figure size settings
figure_size <- data.frame(
  journal = "Lancet",
  single = 85,   # Width for single-column figures (mm)
  double = 178   # Width for double-column figures (mm)
)
fgw1 <- figure_size[figure_size$journal == "Lancet",]$single
fgw2 <- figure_size[figure_size$journal == "Lancet",]$double

# Color palette for consistent visualization
set1 <- brewer.pal(9, "Set1")  # color sets
my_discrete_colors <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00")

# Define standard probability values for quantiles
probs <- c(0.025, 0.25, 0.5, 0.75, 0.975)  # 2.5%, 25%, 50%, 75%, 97.5% quantiles

# Reference vaccination week
VW <- 9  # representative vaccination week from GTFCC data

# Sequence of vaccination weeks to evaluate
VW_seq <- seq(1, 15, 2)

# Representative vaccination coverage rate
VC <- 0.75  # 75% coverage

# Three representative vaccination coverage rates
VC3 <- c(0.6, 0.75, 0.9)  # 60%, 75%, 90%

# Thresholds for case-triggered vaccination to analyze
cts_sel <- c(1, seq(10, 100, 10))  # case triggers selected

# Latest date for which OCV campaign data is available
YEAR <- as.Date("2020-12-28")

# Load geographic data for mapping
shp <- st_read("data/shapefiles/total_shp_0427.shp", quiet = TRUE)
africa <- st_read("data/shapefiles/afr_g2014_2013_0.shp", quiet = TRUE)

# only a subset of sample runid's is used for plotting for convenience
set.seed(42)
sample_runids <- sample(1:200, size=30)
```

# Load Data

```{r load-data}
#| warning: false
# Load outbreak data
# All outbreaks including those affected by OCVs
outbreak_data_all <- as.data.frame(fread("outputs/outbreak_data_all_20250228.csv"))

# Outbreaks included in vaccine impact modeling (not affected by OCVs)
outbreak_data <- fread("outputs/outbreak_data_20250228.csv") 

# outbreaks filtered
# This includes outbreaks that are finally analyzed and excludes
# outbreaks that were duplicated
outbreaks_filtered <-
  readRDS("outputs/outbreak_summary_filtered_20250418.rds")

outbreak_data %<>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 
# outbreaks excluded
outbreaks_excluded <- read_rds("outputs/outbreaks_excluded_20250612.rds")

# Load OCV campaign data
ocv_use <- fread("outputs/ocv_use_GTFCC_data_20241120.csv")

# Calculate delay statistics between request and round 1 vaccination
delay_req_r1 <- as.integer(ocv_use$DATE_ROUND_ONE - ocv_use$DATE_REQUEST)
# Compute key percentiles of the delay distribution
ocv_delay <- quantile(delay_req_r1, probs = c(0.25, 0.5, 0.75, 0.9), na.rm = TRUE)

# Get historical OCV deployment data
dat <- clean_names(fread("data/OCV_Use_GTFCC_Dashboard.csv"))
dat |> 
  filter(shipped == "Shipped", type == "Reactive") |> 
  group_by(year) |> 
  summarize(num_ocv = sum(dose)) |>
  setDT() -> ocv_year

ocv_year$cum_num_ocv <- cumsum(ocv_year$num_ocv)

# Load OCV campaign dataset
docv_ <- clean_names(fread("data/outbreak_files_Oct2024/ocv_long_dataset.csv"))
docv <- docv_[(ocv_before_1 == TRUE | ocv_before_2 == TRUE | 
               ocv_before_3 == TRUE | ocv_during == TRUE)]

# Load model parameters
dis_params <- as.data.frame(data.table::fread("data/parameters.csv"))

# Load IVE posterior predictive values from Bayesian beta regression
ive_fit <- read_rds("outputs/ive_yrep_20241010.rds")
ive_data <- ive_fit$yrep
rm(ive_fit)  # Free memory by removing large object

# Load outbreak summary data
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Oct2024/outbreak_summary.csv")))
message(paste0(length(unique(ds$country)), " countries"))

# Create unique outbreak identifiers
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# Join outbreak data with summary information
outbk_data_summary <- 
  left_join(
    filter(outbreak_data, week == 1), 
    ds[, c("outbreak_pop", "total_s_ch", "total_c_ch", "total_deaths", "duration_weeks", 
           "peak_week_incidence", "time_to_peak_week", "peak_week_incidence_per_100percent", 
           "peak_week_incidence_per_10e3", "mean_r0_1st_week", "area_per_1km2", 
           "rural_urban", "ar_per_10e3", "cfr_per_100percent", "id_outbreak")], 
    by = "id_outbreak"
  )

outbk_data_summary %<>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 
# Summarize attack rates and population sizes
summary(outbk_data_summary$ar_per_10e3)
summary(outbk_data_summary$outbreak_pop)
```

# Recent Cholera Outbreak Analysis

This section analyzes recent cholera data from WHO Global Cholera Dashboard.

```{r recent-outbreaks}
# Load recent cholera data
c2023 <- fread("data/cholera_adm0_public_2023.csv")
c2024 <- fread("data/cholera_adm0_public_2024.csv")
c2023$year <- 2023
c2024$year <- 2024
chol <- rbind(c2023, c2024)

# Analyze 2023 data by WHO region
c2023 |> 
  group_by(year, who_region) |>
  summarize(case_region_total = sum(case_total), 
            death_region_total = sum(death_total)) |> 
  mutate(case_pct = 100 * case_region_total / sum(case_region_total),
         death_pct = 100 * death_region_total / sum(death_region_total),
         case_tot = sum(case_region_total),
         death_tot = sum(death_region_total))

# Analyze 2024 data by WHO region
c2024 |> 
  group_by(year, who_region) |>
  summarize(case_region_total = sum(case_total), 
            death_region_total = sum(death_total)) |> 
  mutate(case_pct = 100 * case_region_total / sum(case_region_total),
         death_pct = 100 * death_region_total / sum(death_region_total),
         case_tot = sum(case_region_total),
         death_tot = sum(death_region_total))

# Analyze all data by WHO region
chol |> 
  group_by(who_region) |>
  summarize(case_region_total = sum(case_total), 
            death_region_total = sum(death_total)) |> 
  mutate(case_pct = 100 * case_region_total / sum(case_region_total),
         death_pct = 100 * death_region_total / sum(death_region_total),
         case_tot = sum(case_region_total),
         death_tot = sum(death_region_total))
```

# Load Vaccine Impact Simulation Results

```{r load-results}
# Time-triggered vaccination results
cea <- read_rds("outputs/cea_onset_trigger_20250520.rds")

# cea_vc85 <- read_rds("outputs/cea_0.85_time_trigger_20250313.rds")
# Combine results from different coverage levels
# cea <- rbind(cea, cea_vc85)
# rm(cea_vc85)

# Case-triggered vaccination results
cea_ct <- read_rds("outputs/cea_0.75_case_trigger_20250523.rds")
# cea_ct <- rbind(cea_ct, read_rds("outputs/cea_0.75_case_trigger_20250523.rds"))

# Preemptive vaccination results
cea_pe_vc60 <- read_rds("outputs/cea_0.6_preemptive_20250314.rds")
cea_pe_vc75 <- read_rds("outputs/cea_0.75_preemptive_20250314.rds")
cea_pe_vc90 <- read_rds("outputs/cea_0.9_preemptive_20250314.rds")

# Combine preemptive results
cea_pe <- rbind(cea_pe_vc60, cea_pe_vc75)
rm(cea_pe_vc60, cea_pe_vc75)  # Free memory
cea_pe <- rbind(cea_pe, cea_pe_vc90)
rm(cea_pe_vc90)
cea_pe <- cea_pe %>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak)

# Load subset analyses
cea_subsets <-
  read_rds("outputs/cea_0.75_onset_trigger_random_subsets_by_year_20250606.rds")

cea_subsets_by_size <-
  read_rds("outputs/cea_onset_trigger_subsets_by_size_20250607.rds")

cea <- left_join(cea, outbk_data_summary[,                                 c("id_outbreak","ar_per_10e3","duration_weeks")], 
                 by = "id_outbreak")

```

# Result Figures

## Figure 1. Outbreaks and % case reduction 
```{r}
# Setup --------------------------------------------------------------
set.seed(42)
# Global parameters ----------------------------------------------------
save_plot <- TRUE
text_size <- 5          # Default text size for annotations
line_height <- 0.95     # Line spacing for text annotation
pattern_dens <- 0.1     # Density of pattern filling
pattern_spacing <- 0.1  # Spacing between patterns
step_linewidth <- 0.7   # Width of step lines
arrow_vertical_len <- 14 # Vertical length of arrows

# Axis labels
ylabel <- "Weekly suspected cholera cases"
xlabel <- "Weeks since outbreak onset"

# Key timeline parameters
week_or <- 3      # Outbreak response at week 3  
week_vax <- 6.5   # Vaccination started
week_vax_eff <- 8.5 # Vaccine takes effect

# Rectangle dimensions configuration ------------------------------------
# This section makes it easy to adjust all rectangles at once

# Base rectangle dimensions
rect_width <- 0.8
rect_height <- 6

# Spacing between rectangles
gap_horizontal <- 1.9   # Horizontal gap between rectangle groups
gap_vertical <- 12      # Vertical gap between top and bottom rectangles

# Font sizes for elements
eq_symbol_size <- 8     # Size for equation symbols (<, +)
paren_size <- 9         # Size for parentheses
plus_size <- 5          # Size for plus symbols

# Vertical offset for second equation from first
eq2_vertical_offset <- 24

# Helper functions ----------------------------------------------------
create_filled_square <- function(x1, y1, x2, y2, fill_color, 
                                 color='black', alpha=0.5) {
  geom_rect(xmin=x1, xmax=x2, ymin=y1, ymax=y2,
            color=color, fill=fill_color, alpha=alpha, inherit.aes = FALSE)
}

create_striped_square <- function(x1, y1, x2, y2, fill_color, color='black',
                                  dens=0.1, angle=30) {
  geom_rect_pattern(
    xmin=x1, xmax=x2, ymin=y1, ymax=y2,
    pattern = 'stripe',
    fill = NA,
    colour = color,
    pattern_density = dens,
    pattern_spacing = 0.025,
    pattern_fill = fill_color,
    pattern_colour = NA,
    pattern_angle = angle,
    inherit.aes = FALSE)
}

create_divider <- function(x1, x2, y, extension=0.2, linewidth=0.7) {
  annotate("segment", 
           x = x1 - extension, 
           y = y, 
           xend = x2 + extension, 
           yend = y, 
           linewidth = linewidth)
}

theme_upper_panel <- function() {
  theme_classic() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 16),
      axis.text.y = element_text(size = 14),
      plot.caption = element_text(size = 14),
      legend.position = "none"
    )
}

# Data preparation ----------------------------------------------------
# Create synthetic example of an outbreak trajectory with and without vaccination
d <- data.frame(
  date = seq(as.Date("2010-06-15"), by = "week", length.out = 24),
  week = 0:23,
  # No vaccination scenario
  s_ch = c(2, 2, 3, 4, 10, 12, 29, 100, 120, 150, 140, 
           75, 68, 47, 24, 5, 12, 5, 3, 2, 4, 1, 0, 0)
)

# Add vaccination scenario (40% of cases after vaccine takes effect)
d$s_ch_vacc <- c(
  2, 2, 3, 4, 10, 12, 29, 100, 120, 
  round(0.4 * c(150, 140, 75, 68, 47, 24, 5, 12, 5, 3, 2, 4, 1, 0, 0))
)
d$s_ch_cumul <- cumsum(d$s_ch)

# Rename columns for clarity
names(d) <- c("date", "week", "No vacc", "Vacc", "No vacc cumulative")

# Reshape data for plotting
dlong <- pivot_longer(d, cols = c("No vacc", "Vacc"))

# Define time weeks for additional scenarios
weeks <- 0:25

# Generate additional outbreak scenarios
# Smallest outbreak (ended before vaccination)
smallest_no_vax <- rpois(length(weeks), lambda=30 * exp(-0.1 * (weeks - 2)^2))
smallest_no_vax[(week_vax-1):length(smallest_no_vax)] <- 0

# Smaller outbreak (ended at vaccine takes effect)
smaller_no_vax <- rpois(length(weeks), lambda=50 * exp(-0.1 * (weeks - 3)^2))
smaller_no_vax[week_vax_eff:length(smaller_no_vax)] <- 0

# Small outbreak (peak at week 8)
small_no_vax <- rpois(length(weeks), lambda=70 * exp(-0.08 * (weeks - 8)^2))
small_with_vax <- small_no_vax
small_with_vax[weeks > week_vax_eff] <- round(small_with_vax[weeks > week_vax_eff] * 0.5)  # 50% reduction

# Large outbreak (peak at week 12)
large_no_vax <- rpois(length(weeks), lambda=120 * exp(-0.04 * (weeks - 12)^2))
large_with_vax <- large_no_vax
large_with_vax[weeks > week_vax_eff] <- round(large_with_vax[weeks > week_vax_eff] * 0.5)  # 50% reduction

# Combine into a data frame
df <- tibble(
  week = rep(weeks, 6),
  cases = c(smallest_no_vax, smaller_no_vax, small_no_vax, small_with_vax, 
            large_no_vax, large_with_vax),
  outbreak = rep(c("Smallest", "Smaller", "Small", "Small", "Large", "Large"), 
                 each = length(weeks)),
  scenario = c(rep("No Vaccine", 2*length(weeks)), 
               rep(c("No Vaccine", "With Vaccine"), times=2, each=length(weeks)))
)

# Create wide version of the data
df_wide <- df %>%
  pivot_wider(names_from = scenario, values_from = cases) %>%
  mutate(reduction = `No Vaccine` - `With Vaccine`)

# Rectangle positions for equation 1 ----------------------------------
# Starting position for first equation
base_x1 <- 15
base_y1 <- 65

# a/b fraction
a_x1 <- base_x1
a_y1 <- base_y1
a_x2 <- a_x1 + rect_width
a_y2 <- a_y1 + rect_height

b_x1 <- a_x1
b_y1 <- a_y1 - gap_vertical
b_x2 <- b_x1 + rect_width
b_y2 <- b_y1 + rect_height

ab_div_y <- (a_y1 + b_y2) / 2

# c/d fraction
c_x1 <- a_x1 + gap_horizontal
c_y1 <- a_y1
c_x2 <- c_x1 + rect_width
c_y2 <- c_y1 + rect_height

d_x1 <- c_x1
d_y1 <- b_y1
d_x2 <- d_x1 + rect_width
d_y2 <- d_y1 + rect_height

cd_div_y <- (c_y1 + d_y2) / 2

# Inequality symbol
eq1_lt_x <- (a_x2 + c_x1) / 2
eq1_lt_y <- (a_y1 + b_y2) / 2

# Equation label
eq1_label_x <- a_x1 - 1.2
eq1_label_y <- cd_div_y

# Rectangle positions for equation 2 ----------------------------------
# Left side: 1/2(a/b+c/d)
# a/b part
eq2_a_x1 <- a_x1 + 0.2
eq2_a_y1 <- a_y1 - eq2_vertical_offset
eq2_a_x2 <- eq2_a_x1 + rect_width
eq2_a_y2 <- eq2_a_y1 + rect_height

eq2_b_x1 <- eq2_a_x1
eq2_b_y1 <- eq2_a_y1 - gap_vertical
eq2_b_x2 <- eq2_b_x1 + rect_width
eq2_b_y2 <- eq2_b_y1 + rect_height

# c/d part
eq2_c_x1 <- eq2_a_x1 + gap_horizontal
eq2_c_y1 <- eq2_a_y1
eq2_c_x2 <- eq2_c_x1 + rect_width
eq2_c_y2 <- eq2_c_y1 + rect_height

eq2_d_x1 <- eq2_c_x1
eq2_d_y1 <- eq2_b_y1
eq2_d_x2 <- eq2_d_x1 + rect_width
eq2_d_y2 <- eq2_d_y1 + rect_height

# Parentheses
left_paren_x <- eq2_a_x1 - 0.3
right_paren_x <- eq2_c_x2 + 0.3
paren_y <- (eq2_a_y1 + eq2_b_y2) / 2

# 1/2 fraction position
eq2_frac_x <- right_paren_x + 0.6
eq2_frac_y_num <- (eq2_a_y1 + eq2_a_y2)/2
eq2_frac_y_denom <- (eq2_b_y1 + eq2_b_y2)/2
eq2_frac_y_line <- (eq2_a_y1 + eq2_b_y2)/2

# Plus symbol position
eq2_plus1_x <- (eq2_a_x2 + eq2_c_x1) / 2
eq2_plus1_y <- (eq2_a_y1 + eq2_b_y2) / 2

# Less than symbol for second equation
eq2_lt_x <- eq2_frac_x + 0.8
eq2_lt_y <- (eq2_a_y1 + eq2_b_y2)/2

# Right side: (a+c)/(b+d)
# Numerator (a+c)
eq2_right_a_x1 <- eq2_lt_x + 0.8
eq2_right_a_y1 <- eq2_a_y1
eq2_right_a_x2 <- eq2_right_a_x1 + rect_width
eq2_right_a_y2 <- eq2_right_a_y1 + rect_height

eq2_right_c_x1 <- eq2_right_a_x1 + gap_horizontal
eq2_right_c_y1 <- eq2_a_y1
eq2_right_c_x2 <- eq2_right_c_x1 + rect_width
eq2_right_c_y2 <- eq2_right_c_y1 + rect_height

eq2_plus2_x <- (eq2_right_a_x2 + eq2_right_c_x1)/2
eq2_plus2_y <- (eq2_right_a_y1 + eq2_right_a_y2)/2

# Denominator (b+d)
eq2_right_b_x1 <- eq2_right_a_x1
eq2_right_b_y1 <- eq2_b_y1
eq2_right_b_x2 <- eq2_right_b_x1 + rect_width
eq2_right_b_y2 <- eq2_right_b_y1 + rect_height

eq2_right_d_x1 <- eq2_right_c_x1
eq2_right_d_y1 <- eq2_right_b_y1
eq2_right_d_x2 <- eq2_right_d_x1 + rect_width
eq2_right_d_y2 <- eq2_right_d_y1 + rect_height

eq2_plus3_x <- eq2_plus2_x
eq2_plus3_y <- (eq2_right_b_y1 + eq2_right_b_y2)/2

# Division line
eq2_right_div_x1 <- eq2_right_a_x1
eq2_right_div_x2 <- eq2_right_c_x2
eq2_right_div_y <- (eq2_right_a_y1 + eq2_right_b_y2) / 2

# Equation 2 label
eq2_label_x <- eq1_label_x
eq2_label_y <- a_y1 - eq2_vertical_offset

# Create plots --------------------------------------------------------
# Panel 1: Main example with vaccine effect
plt1a <- ggplot(dlong, aes(week, value, linetype = name, shape = name, group = name)) +
  # Background stripes for no vaccination scenario
  geom_rect_pattern(
    data = subset(dlong, name == "No vacc"),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=value),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_colour = NA,
    pattern_density = pattern_dens,
    pattern_fill = 'steelblue',
    pattern_angle = 30) +
  
  # Vaccination impact rectangle
  geom_rect(
    data = d,
    aes(xmin=week, xmax=lead(week), ymin=`Vacc`, ymax=`No vacc`), 
    fill='steelblue', 
    alpha=0.5,
    color=NA,
    inherit.aes = FALSE) +
  
  # Step lines for clarity
  geom_step(linewidth=step_linewidth) +
  
  # Outbreak response triggered arrow and label
  geom_segment(
    x=week_or, y=10+arrow_vertical_len, xend=week_or, yend=10,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_or-0.5, y=38, 
    label = "Outbreak\nresponse triggered", 
    hjust=0.5, size=text_size, lineheight=line_height
  ) +
  
  # Vaccination started arrow and label
  geom_segment(
    x=week_vax, y=32+arrow_vertical_len, xend=week_vax, yend=32,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_vax, y=60, 
    label = "Vaccination\nstarted", hjust=1,
    size=text_size, lineheight=line_height
  ) +
  
  # Vaccine takes effect arrow and label
  geom_segment(
    x=week_vax_eff, y=122+arrow_vertical_len, xend=week_vax_eff, yend = 122,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_vax_eff, y=147, 
    label="Vaccine\ntakes effect", hjust=1,
    size=text_size, lineheight=line_height
  ) +
  
  # Scenario labels
  annotate(
    "text", x=11.5, y=120, 
    label = "Without\nvaccination", hjust=0, 
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=9.5, y=12, 
    label = "With\nvaccination", hjust=0,
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=16, y=75, 
    label = "% case\nreduction", 
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=19.2, y=75, 
    label = bquote("= 100 \u00D7"), 
    size=text_size, lineheight=line_height
  ) +
  
  # Fraction expressed using squares
  create_divider(x1=21.1-0.2, x2=22.1+0.2, y=75) +
  create_filled_square(
    x1=21.1, x2=22.1, y1=77, y2=85, 
    color='black', fill="steelblue", alpha=0.5
  ) +
  create_striped_square(
    x1=21.1, x2=22.1, y1=64, y2=72,
    fill_color='steelblue', color='black'
  ) +
  
  # Colored lines on the horizontal axis marking periods
  annotate("segment", x=0, y=0, xend=3, yend=0, linewidth=2, color="firebrick") +
  annotate("segment", x=3, y=0, xend=7, yend=0, linewidth=2, color="forestgreen") +
  annotate("segment", x=7, y=0, xend=9, yend=0, linewidth=2, color="purple") +
  
  scale_x_continuous(limits=c(0,25)) +
  labs(x="", y=ylabel) +
  theme_upper_panel()

# Panel 2: Smaller outbreaks
plt1b <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot smaller outbreaks
  geom_step(
    data = subset(df, outbreak == "Smallest"),
    aes(linetype=scenario),
    linewidth=step_linewidth, 
    color='grey'
  ) +
  geom_step(
    data = subset(df, outbreak == "Smaller"),
    aes(linetype=scenario),
    linewidth=step_linewidth, 
    color='black'
  ) +
  
  labs(x=NULL, y=ylabel) +
  theme_upper_panel()

# Panel 3: Small outbreak with equations
plt1c <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot small outbreak
  geom_step(
    data = subset(df, outbreak == "Small"), 
    aes(linetype=scenario),
    linewidth=step_linewidth
  ) +
  
  # Pattern for no vaccination scenario
  geom_rect_pattern(
    data = subset(df, (scenario == "No Vaccine" & outbreak == "Small")),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=cases),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_density = pattern_dens*0.5,
    pattern_fill = 'steelblue',
    pattern_colour = NA,
    pattern_angle = 30
  ) +
  
  # Vaccination impact area
  geom_rect(
    data=subset(df_wide, outbreak == "Small"),
    aes(xmin=week, xmax=lead(week), ymin=`With Vaccine`, ymax=`No Vaccine`),
    fill="steelblue", 
    alpha=0.5, 
    inherit.aes=FALSE
  ) +
  
  # EQUATION 1: a/b < c/d
  # a/b fraction
  create_filled_square(a_x1, a_y1, a_x2, a_y2, "steelblue") +
  create_striped_square(b_x1, b_y1, b_x2, b_y2, "steelblue", dens=pattern_dens*0.5) +
  create_divider(a_x1, a_x2, ab_div_y) +
  
  # Less than symbol
  annotate("text", x=eq1_lt_x, y=eq1_lt_y, label= "<", size=eq_symbol_size) +
  
  # c/d fraction
  create_filled_square(c_x1, c_y1, c_x2, c_y2, "firebrick") +
  create_striped_square(d_x1, d_y1, d_x2, d_y2, "firebrick", dens=pattern_dens*0.5) +
  create_divider(c_x1, c_x2, cd_div_y) +
  
  # Label for equation 1
  annotate("text", x=eq1_label_x, y=eq1_label_y, label="1.", size=5, hjust=0) +
  
  # EQUATION 2: 1/2(a/b+c/d) < (a+c)/(b+d)
  # Left side: 1/2(a/b+c/d)
  
  # a/b part
  create_filled_square(eq2_a_x1, eq2_a_y1, eq2_a_x2, eq2_a_y2, "steelblue") +
  create_striped_square(eq2_b_x1, eq2_b_y1, eq2_b_x2, eq2_b_y2, "steelblue", dens=pattern_dens*0.5) +
  create_divider(eq2_a_x1, eq2_a_x2, (eq2_a_y1 + eq2_b_y2)/2) +
  
  # Plus symbol
  annotate("text", x=eq2_plus1_x, y=eq2_plus1_y, label="+", size=plus_size) +
  
  # c/d part
  create_filled_square(eq2_c_x1, eq2_c_y1, eq2_c_x2, eq2_c_y2, "firebrick") +
  create_striped_square(eq2_d_x1, eq2_d_y1, eq2_d_x2, eq2_d_y2, "firebrick", dens=pattern_dens*0.5) +
  create_divider(eq2_c_x1, eq2_c_x2, (eq2_c_y1 + eq2_d_y2)/2) +
  
  # Parentheses
  annotate("text", x=left_paren_x, y=paren_y, label="(", size=paren_size) +
  annotate("text", x=right_paren_x, y=paren_y, label=")", size=paren_size) +
  
  # 1/2 fraction
  annotate("text", x=eq2_frac_x, y=eq2_frac_y_num, label="1", size=5) +
  annotate("text", x=eq2_frac_x, y=eq2_frac_y_denom, label="2", size=5) +
  annotate("segment", 
           x=eq2_frac_x-0.3, y=eq2_frac_y_line, 
           xend=eq2_frac_x+0.3, yend=eq2_frac_y_line, 
           linewidth=0.8) +
  
  # Less than symbol for second equation
  annotate("text", x=eq2_lt_x, y=eq2_lt_y, label="<", size=eq_symbol_size) +
  
  # Right side: (a+c)/(b+d)
  # Numerator (a+c)
  create_filled_square(eq2_right_a_x1, eq2_right_a_y1, eq2_right_a_x2, eq2_right_a_y2, "steelblue") +
  annotate("text", x=eq2_plus2_x, y=eq2_plus2_y, label="+", size=plus_size) +
  create_filled_square(eq2_right_c_x1, eq2_right_c_y1, eq2_right_c_x2, eq2_right_c_y2, "firebrick") +
  
  # Division line
  create_divider(eq2_right_div_x1, eq2_right_div_x2, eq2_right_div_y) +
  
  # Denominator (b+d)
  create_striped_square(eq2_right_b_x1, eq2_right_b_y1, eq2_right_b_x2, eq2_right_b_y2, 
                        "steelblue", dens=pattern_dens*0.5) +
  annotate("text", x=eq2_plus3_x, y=eq2_plus3_y, label="+", size=plus_size) +
  create_striped_square(eq2_right_d_x1, eq2_right_d_y1, eq2_right_d_x2, eq2_right_d_y2, 
                        "firebrick", dens=pattern_dens*0.5) +
  
  # Label for equation 2
  annotate("text", x=eq2_label_x, y=eq2_label_y, label="2.", size=5, hjust=0) +
  
  labs(x=NULL, y=ylabel) +
  theme_upper_panel()

# Panel 4: Large outbreak  
plt1d <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot large outbreak
  geom_step(
    data = subset(df, outbreak == "Large"), 
    aes(linetype=scenario),
    linewidth=step_linewidth
  ) +
  
  # Pattern for no vaccination scenario
  geom_rect_pattern(
    data = subset(df, (scenario == "No Vaccine" & outbreak == "Large")),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=cases),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_density = pattern_dens*0.5,
    pattern_fill = 'firebrick',
    pattern_colour = NA,
    pattern_angle = 30
  ) +
  
  # Vaccination impact area
  geom_rect(
    data=subset(df_wide, outbreak == "Large"),
    aes(xmin=week, xmax=lead(week), ymin=`With Vaccine`, ymax=`No Vaccine`),
    fill="firebrick", 
    alpha=0.5, 
    inherit.aes=FALSE
  ) +
  
  # Axis and labels
  labs(x=xlabel, y=ylabel) +
  scale_x_continuous(limits=c(0,25)) +
  theme_upper_panel() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    legend.position = "none"
  )

# Combine all plots
combined_plot <- (plt1a / plt1b / plt1c / plt1d) + 
  plot_layout(
    heights = c(0.8, 0.4, 0.4, 0.4), 
    axis_titles = 'collect',
    guides = 'collect'
  )

# Save plot
if (save_plot) {
  fac <- 1.8
  ggsave(paste0("plots/fig1_vacc_eff_schematic_", tstamp(), ".pdf"),
        combined_plot,
        device = cairo_pdf,
        width=1.25*fgw1*fac,
        height=1.6*fgw1*fac,
        units = "mm")
}

```

## Figure 2: Outbreak Description

Outbreak ids by confirmation status
```{r}
ids_excl <- outbreak_data_all |> 
  filter(!(id_outbreak %in% outbreak_data$id_outbreak)) |>  
  pull(id_outbreak) |>
  unique()

# Identify confirmed outbreaks (with laboratory confirmation)
ids_conf <- outbreak_data |> 
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique()

# Identify suspected (not confirmed) outbreaks
ids_susp <- outbreak_data |>
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |>
  filter(!(c_ch_sum > 0)) |>
  pull(id_outbreak) |>
  unique()
```

```{r}
# Load and prepare country-region mapping data
country_region <- fread("data/africa_subregion_country_names.csv")
country_region$iso3c <- sapply(country_region$Country, 
                               function(x) countrycode(x, origin = "country.name.en", 
                                                     destination = "iso3c"))

# Join outbreak data with regional information
outbreak_data$iso3c <- outbreak_data$country
outbreak_data_region <- left_join(outbreak_data, country_region, by = "iso3c")

# =============================================================================
# PLOT 2A: TEMPORAL TRENDS BY SUBREGION (STACKED AREA CHART)
# =============================================================================

# Aggregate data by month and subregion
outbreak_month <- outbreak_data_region %>% 
  mutate(date = as.Date(date),
         date_bin = floor_date(date, "1 month"))

# Calculate total cases per subregion per month
outbreak_month_summary <- outbreak_month %>%
  group_by(date_bin, Subregion) %>%
  summarise(total_cases = sum(s_ch, na.rm = TRUE), .groups = "drop")

# Create complete time series grid to fill gaps
all_subregions <- unique(outbreak_month_summary$Subregion)
all_date_bins <- seq(min(outbreak_month_summary$date_bin), 
                     max(outbreak_month_summary$date_bin), 
                     by = "1 month")
full_grid <- expand_grid(date_bin = all_date_bins, Subregion = all_subregions)

# Merge and fill missing values with zero
df_complete <- full_grid %>%
  left_join(outbreak_month_summary, by = c("date_bin", "Subregion")) %>%
  mutate(total_cases = replace_na(total_cases, 0))

# Create stacked area plot
plt_2a <- ggplot(df_complete, aes(x = date_bin, y = total_cases, fill = Subregion)) +
  geom_area(position = "stack", alpha = 0.8) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  scale_y_continuous(labels = comma_format()) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  labs(x = "", y = "Monthly suspected cases", fill = "") +
  theme_light() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    plot.margin = margin(0,1,0,1,"pt")
  )

# =============================================================================
# PLOT 2B: SPATIAL DISTRIBUTION MAP
# =============================================================================

# Prepare outbreak data for mapping
dt <- outbreak_data %>% rename(lctn_pr = location_period_id)
first_rows_df <- dt %>% 
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

# Ensure consistent data types for joining
shp$lctn_pr <- as.character(shp$lctn_pr)

# Join outbreak data with spatial data
map_data <- left_join(first_rows_df, shp, by = "lctn_pr")
map_data %<>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE))

# Convert to spatial features
map_data <- st_as_sf(map_data)

islands_to_exclude <- c("CPV",  # Cape Verde
                        "SYC",  # Seychelles
                        "MYT",  # Mayotte
                        "REU",  # Réunion
                        "SHN")  # Saint Helena, Ascension and Tristan da Cunha

# Exclude these from your Africa shapefile
africa_main <- africa %>% dplyr::filter(!ISO3 %in% islands_to_exclude)
bbox   <- st_bbox(africa_main)       # much tighter now

plt_2b <- ggplot() +
  geom_sf(data = africa_main,  fill = NA, colour = "black", size = .3) +
  geom_sf(data = map_data, aes(colour = is_confirmed, fill = is_confirmed),
          alpha = .7, size = .8) +
  coord_sf(xlim = c(bbox["xmin"], bbox["xmax"]-10),
           ylim = c(bbox["ymin"]+12, bbox["ymax"]),
           expand = FALSE, clip = "on") +    # clip cuts everything outside
  scale_colour_manual("", values = c("steelblue4", "firebrick4"),
                      labels  = c("Not confirmed", "Confirmed")) +
  scale_fill_manual("",   values = c("steelblue4", "firebrick4"),
                    labels  = c("Not confirmed", "Confirmed")) +
  theme_void() +                                   
  theme(legend.position = "top",
        plot.margin     = margin(0,0,0,0,"pt"),
        panel.spacing = unit(0, "pt"))

# plt_2b <- plt_2b + theme(aspect.ratio = 0.7)

# =============================================================================
# PLOT 2C: EPIDEMIC CURVES WITH INSET
# =============================================================================

# Prepare weekly case data
week_sch <- outbreak_data[, c("id_outbreak", "week", "s_ch")]
setDT(week_sch)
max_week <- week_sch[, max(week)]

# Create complete grid for all outbreaks up to max week
all_weeks <- CJ(id_outbreak = unique(week_sch$id_outbreak), week = 1:max_week)

# Merge and fill missing values with zero
week_sch_full <- merge(all_weeks, week_sch, 
                       by = c("id_outbreak", "week"), all.x = TRUE)
week_sch_full[is.na(s_ch), s_ch := 0]

# Calculate summary statistics across all outbreaks
week_sch_summary <- week_sch_full[, .(
  sum_sch = sum(s_ch, na.rm = TRUE),
  median = median(s_ch, na.rm = TRUE),
  p025 = quantile(s_ch, 0.025, na.rm = TRUE),
  p25  = quantile(s_ch, 0.25, na.rm = TRUE),
  p75  = quantile(s_ch, 0.75, na.rm = TRUE),
  p975 = quantile(s_ch, 0.975, na.rm = TRUE)
), by = week]

# Main epidemic curve plot
plt2c <- ggplot(week_sch_summary, aes(x = week)) +
  geom_ribbon(aes(ymin = p025, ymax = p975), fill = set1[3], alpha = 0.45) +
  geom_ribbon(aes(ymin = p25, ymax = p75), fill = set1[3], alpha = 0.6) +
  geom_line(aes(y = median), color = "black", size = 1) +
  labs(title = "", x = "Epidemic week", y = "Suspected cholera cases") +
  scale_x_continuous(limits = c(0, 50), breaks = seq(0, 50, 10)) +
  theme_light() +
  theme(
    plot.margin = margin(0,1,0,1, "pt")
  )

plt2c_inset <- ggplot(outbk_data_summary, aes(x = time_to_peak_week)) +
  stat_ecdf(geom = "step", size = 0.8) +
  labs(x = "Time to peak (weeks)", 
       y = expression(P(X <= x)))+
  scale_x_continuous(limits=c(0,30), breaks = seq(0,30,10), labels = seq(0,30,10))+
  scale_y_continuous(limits=c(0,1), breaks = seq(0,1,0.5), labels = seq(0,1,0.5))+
  theme_light()+
  theme_tight()+
  theme(plot.margin = margin(0,0,0,0,"pt"))


# Convert inset to grob and add to main plot with better positioning
grob_inset <- ggplotGrob(plt2c_inset)

plt_2c_combined <- plt2c +
  annotation_custom(
    grob = grob_inset,
    xmin = 20, xmax = 50,  # Moved to right side with more space
    ymin = 100, ymax = 340   # Increased height for better visibility
  )

# =============================================================================
# PLOT 2D: OUTBREAK SIZE vs DURATION
# =============================================================================

plt_2d <- outbk_data_summary %>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE)) %>% 
  ggplot() +
  geom_point(aes(x = total_s_ch, y = duration_weeks, color = is_confirmed), 
             alpha = 0.7, size = 1.2) +
  scale_y_log10(labels = comma_format()) +
  scale_x_log10(labels = comma_format()) +
  scale_color_manual("", 
                     labels = c("Not confirmed", "Confirmed"),
                     values = c("steelblue4", "firebrick4")) +
  labs(title = "", x = "Total suspected cases", y = "Outbreak duration (weeks)") +
  theme_light() +
  theme(legend.position = "top", plot.margin = margin(0,3,0,1,"pt"))

# =============================================================================
# PLOT 2E: OUTBREAK SIZE vs ATTACK RATE
# =============================================================================
 
plt_2e <- outbk_data_summary %>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE)) %>% 
  ggplot() +
  geom_point(aes(x = total_s_ch, y = ar_per_10e3, color = is_confirmed), 
             alpha = 0.7, size = 1.2) +
  scale_y_log10(labels = comma_format()) +
  scale_x_log10(labels = comma_format()) +
  scale_color_manual("", 
                     labels = c("Not confirmed", "Confirmed"),
                     values = c("steelblue4", "firebrick4")) +  
  labs(title="", x = "Total suspected cases", y = "Attack rate (per 10,000)") +
  theme_light() +
  theme(legend.position="top",  plot.margin = margin(0,1,0,3,"pt"))

# =============================================================================
# FINAL COMBINED PLOT WITH IMPROVED LAYOUT
# =============================================================================

# Create combined plot with better proportions
# Row 1: Temporal trends (full width)
# Row 2: Map (60%) + Epidemic curve (40%) 
# Row 3: Scatter plots (50% each)
combined_plot <- plt_2a /
  (plt_2b + plt_2c_combined + plot_layout(widths = c(0.6, 1))) /
  (plt_2d + plt_2e + plot_layout(widths = c(1, 1), guides="collect") +
     plot_annotation(theme = theme(legend.position = "top")))+
  plot_layout(heights = c(1,1.3,1)) +
  plot_annotation(
    theme = theme(plot.margin = margin(0,0,0,0,"pt")), 
    tag_levels = 'A')

# Display the combined plot
# combined_plot
if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/fig2_outbreak_summary_", tstamp(), ".pdf"), combined_plot,
         device = cairo_pdf, width = fgw2 * fac, height = fgw2 * fac, units = "mm")
}
```

## Figure S1: Excluded outbreaks were also plotted 

```{r figS1, eval=FALSE}
save_plot <- TRUE
# (A) Outbreak time series by group
# Identify outbreaks that have been excluded or included in analysis
ids_excl <- outbreak_data_all |> 
  filter(!(id_outbreak %in% outbreak_data$id_outbreak)) |>  
  pull(id_outbreak) |>
  unique()

# Identify confirmed outbreaks (with laboratory confirmation)
ids_conf <- outbreak_data |> 
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique()

# Identify suspected (not confirmed) outbreaks
ids_susp <- outbreak_data |>
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |>
  filter(!(c_ch_sum > 0)) |>
  pull(id_outbreak) |>
  unique()

# Create time series plot with different colors by outbreak type
plt_S1a <- ggplot() +
  geom_line(data = filter(outbreak_data_all, id_outbreak %in% ids_excl),
            aes(date, s_ch, group = id_outbreak, color = "Excluded"), 
            alpha = 0.5) +
  geom_line(data = filter(outbreak_data, !(id_outbreak %in% ids_conf)),
            aes(date, s_ch, group = id_outbreak, color = "Suspected"), 
            alpha = 0.5) +
  geom_line(data = filter(outbreak_data, id_outbreak %in% ids_conf), 
            aes(date, s_ch, group = id_outbreak, color = "Confirmed"),
            alpha = 0.5) +
  scale_color_manual("", values = c("Excluded" = "yellow4",
                                  "Suspected" = "steelblue4", 
                                  "Confirmed" = "firebrick4")) +
  labs(x = "", y = "Weekly suspected cases") +
  theme_light() +
  theme_tight() +
  theme(legend.position = "top")

# (B) Outbreak size vs. duration by type
outbreak_data_all |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                  as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d_all

outbreak_data |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                  as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d


if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1a_outbreak_ts_", tstamp(), ".pdf"), plt_S1a,
         device = cairo_pdf, width = fgw2 * fac, height = fgw1 * fac, units = "mm")
}

plt_S1b <- ggplot() +
  geom_point(data = filter(d_all, id_outbreak %in% ids_excl), 
             aes(duration, total_s_ch, color = "Excluded"), alpha = 0.5) +
  geom_point(data = filter(d, id_outbreak %in% ids_susp), 
             aes(duration, total_s_ch, color = "Suspected"), alpha = 0.5) +
  geom_point(data = filter(d, id_outbreak %in% ids_conf), 
             aes(duration, total_s_ch, color = "Confirmed"), alpha = 0.5) +
  scale_x_log10() +
  scale_y_log10() +
  theme_light() +
  theme_tight() +
  scale_color_manual("", values = c("Excluded" = "yellow4", 
                                  "Suspected" = "steelblue4", 
                                  "Confirmed" = "firebrick4")) +
  labs(x = "Outbreak duration (days)", y = "Total suspected cases") +
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1b_outbreak_size_dur_", tstamp(), ".pdf"), plt_S1b,
         device = cairo_pdf, width = fgw1 * fac, height = fgw1 * fac, units = "mm")
}

# (C) Map of outbreaks
dt <- outbreak_data 
dt_all <- outbreak_data_all 

# Prepare data for mapping
dt <- dt %>% rename(lctn_pr = location_period_id)
dt_all <- dt_all %>% rename(lctn_pr = location_period_id)

first_rows_df <- dt %>% 
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

first_rows_df_all <- dt_all %>%
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

first_rows_df$lctn_pr <- as.numeric(first_rows_df$lctn_pr)
first_rows_df_all$lctn_pr <- as.numeric(first_rows_df_all$lctn_pr)

# Join with spatial data
d <- left_join(first_rows_df, shp, by = "lctn_pr")
d_all <- left_join(first_rows_df_all, shp, by = "lctn_pr")

# Create indicator variables for outbreak types
d <- d %>%
  mutate(id_conf = ifelse(id_outbreak %in% ids_conf, 1, 0),
         id_sus = ifelse(!id_outbreak %in% ids_conf, 1, 0))

# Subset data by confirmation status
d_conf <- d %>% filter(id_conf == 1)
d_sus <- d %>% filter(id_sus == 1)

# Convert to spatial data
d_conf <- st_as_sf(d_conf)
d_sus <- st_as_sf(d_sus)
d_all <- d_all %>% mutate(id_excl = ifelse(id_outbreak %in% ids_excl, 1, 0))
d <- st_as_sf(d)
d_all <- st_as_sf(d_all)

# Create map
plt_S1c <- ggplot() +
  geom_sf(data = africa, fill = NA, color = "black") +  # Country outlines
  geom_sf(data = d_all, aes(fill = id_excl), color = "black", 
          fill = "yellow4", alpha = 0.5) +
  geom_sf(data = d_sus, aes(fill = id_sus), color = "steelblue4", 
          fill = "steelblue4", alpha = 0.5) +
  geom_sf(data = d_conf, aes(fill = id_conf), color = "firebrick4", 
          fill = "firebrick4", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Map of Cholera Outbreaks")


if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1c_outbreak_map_", tstamp(), ".pdf"), plt_S1c,
         device = cairo_pdf, width = fgw1 * fac, height = fgw1 * fac, units = "mm")
}
```

## Figure 3: Outbreak Onset-Triggered Scenario

```{r fig3, eval=FALSE}
cea %>%
  filter(vacc_cov == 0.75, vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(variable = s_ch_averted_tot) -> ca1

format_median_iqr(ca1, row=1)
format_median_iqr(ca1, row=2)

cea %>%
  filter(vacc_cov == 0.75, vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(variable = pct_reduc_case) -> pcr1

format_median_iqr(pcr1, row=1)
format_median_iqr(pcr1, row=2)

cea %>%
  filter(vacc_cov == 0.9, vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(variable = pct_reduc_case) -> pcr2

format_median_iqr(pcr2, row=1)
format_median_iqr(pcr2, row=2)

# sampling because large dataset makes plotting slow
cea_s <- filter(cea, runid %in% sample_runids) # cea subset

plt3a <- cea_s |>
  filter(vacc_cov %in% VC3, vacc_week %in% VW_seq) |>
  ggplot(aes(x = as.factor(vacc_week), y = s_ch_averted_tot,
             color = as.factor(vacc_cov))) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  scale_y_continuous(limits=c(0,1000)) +
  scale_color_manual(name = "Vaccine coverage",
                     values = my_discrete_colors[1:3],
                     labels = c(paste0(format_num(VC3*100, 1), "%"))) +
  labs(x = "Delay in ORI implementation (weeks)", y = "Number of cases averted") +
  theme_light() +
  theme_tight()

cea_s |> 
  filter(vacc_cov ==0.75, vacc_week == 1) |>
  summarize(pct = median(pct_reduc_case))
             
plt3b <- cea_s |> 
  filter(vacc_cov %in% VC3, vacc_week %in% VW_seq) |>
  ggplot(aes(x = as.factor(vacc_week), y = pct_reduc_case,
             color = as.factor(vacc_cov))) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  scale_y_continuous(labels = scales::percent_format(scale = 1),
                     limits=c(0,100)) +
  scale_color_manual(name = "Vaccine coverage",
                     values = my_discrete_colors[1:3],
                     labels = c(paste0(format_num(VC3*100, 1), "%"))) +
  labs(x = "Delay in ORI implementation (weeks)", y = "Cases averted (%)") +
  theme_light() +
  theme_tight()

# (C) Impact by outbreak size for outbreak onset-triggered vaccination
# Calculate percentiles for outbreak size groups
probs <- seq(0, 1, by = 0.1)
runids <- unique(cea_s$runid)
unq_data <- filter(cea_s, runid == runids[1] & vacc_cov == 0.75 & vacc_week == 1)
size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)
cea_s$size_grp <- sapply(cea_s$s_ch_tot, function(x) get_group_id(x, size_grp_cutoff))

res <- rbindlist(lapply(1:10, function(i) calculate_impact_by_subset(cea_s, i)))
res %>% filter(vacc_week %in% c(1,9), subset_id %in% c(1,10), vacc_cov == VC)

rbindlist(lapply(1:10, function(i) calculate_impact_by_subset(cea_s, i, summary_fun = "quantile", quantile_prob = c(0.25)))) %>% 
  filter(vacc_week %in% c(1,9), subset_id %in% c(1,10), vacc_cov == VC)-> res_lb 

rbindlist(lapply(1:10, function(i) calculate_impact_by_subset(cea_s, i, summary_fun = "quantile", quantile_prob = c(0.75)))) %>% 
  filter(vacc_week %in% c(1,9), subset_id %in% c(1,10), vacc_cov == VC)-> res_ub

# Create heatmap of impact by outbreak size and vaccination week
plt3c <- res |>
  filter(vacc_cov %in% VC, vacc_week %in% VW_seq) |>
  ggplot(aes(x = vacc_week, y = subset_id)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case_median)) +
  metR::geom_contour2(aes(z = pct_reduc_case_median, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)  ", limits = c(0, 100),
                       breaks = seq(0, 100, 50),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Delay in ORI implementation (weeks)", 
       y = "Total case percentile") +
  scale_x_continuous(breaks = seq(1, 15, 2))+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() +
  theme(legend.position = "bottom")

# probs <- seq(0, 1, by = 0.1)
# ar_grp_cutoff <- quantile(outbk_data_summary$ar_per_10e3, probs = probs)
# dur_grp_cutoff <- quantile(outbk_data_summary$duration_weeks, probs = probs)

cea_s %<>%
  group_by(runid, id_outbreak, vacc_cov) %>%
  mutate(pct_reduc_case_frac = c(1, 1-frac_diff(pct_reduc_case, lag=1)),
         pct_reduc_case_diff = c(1, -diff(pct_reduc_case, lag=1)))

cea_s %>% filter(vacc_cov == 0.75) -> d1
d1$grp <- "All"
cea_s %>% filter(ar_per_10e3 > ar_grp_cutoff[10]) -> d2 # top 10%
d2$grp <- "AR"
cea_s %>% filter(duration_weeks > dur_grp_cutoff[10]) -> d3 
d3$grp <- "Dur"
cea_s %>% filter(s_ch_tot > size_grp_cutoff[10]) -> d4 
d4$grp <- "Size"
d <- rbind(d1,d2)
d_ <- rbind(d3,d4)
d <- rbind(d, d_)
d

d %>%
  filter(vacc_week > 1, vacc_week <= 15) %>% 
  group_by(grp) %>% 
  summarize_quantiles(pct_reduc_case_frac*100) -> prpcr

format_median_iqr(prpcr, row=1)
format_median_iqr(prpcr, row=2)
format_median_iqr(prpcr, row=3)
format_median_iqr(prpcr, row=4)

d %>%
  filter(vacc_week > 1, vacc_week <= 15) %>% 
  mutate(half_life = log(2)/abs(pct_reduc_case_frac)) %>% 
  group_by(grp) %>% 
  summarize_quantiles(half_life) -> hlife

format_median_iqr(hlife, row=1)
format_median_iqr(hlife, row=2)
format_median_iqr(hlife, row=3)
format_median_iqr(hlife, row=4)

half_life <- log(2)/abs(rate)  # Half-life in weeks


plt3d <- d %>% 
  filter(vacc_cov == VC, vacc_week > 1, vacc_week <= 15) %>%
  ggplot(aes(x = as.factor(grp), y = 100*pct_reduc_case_frac)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_x_discrete(labels = c(expression("All"),
                              expression(">="*P[90*"%"]*"(AR)"),
                              expression(">="*P[90*"%"]*"(Dur)"),
                              expression(">="*P[90*"%"]*"(Size)"))) +
  labs(x = "", y = "Weekly reduction in % case averted") +
  theme_light() +
  theme_tight()

# Combine all four plots into a single figure
plt3 <- (plt3a + plt3b + plt3c + plt3d) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              guides = "collect", 
              heights = c(1, 1)) &
  theme(plot.tag.position = c(0.05, 1),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Adjust tag positions for top panels
for (i in c(1, 2)) {
  plt3[[i]] <- plt3[[i]] + theme(plot.tag.position = c(0.05, 0.9))
}

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig3_onset_triggered_", tstamp(), ".pdf"),
         plt3,
         device = cairo_pdf, width = fgw2*fac, 
         height = fgw2*1.15*fac, units = "mm")
}

d %>%
  filter(vacc_cov == VC, vacc_week > 1, vacc_week <= 15) %>%
  group_by(grp) %>%
  summarize_quantiles(variable = 100 * pct_reduc_case_frac) -> prc_frac_summary

rate <- as.double(prc_frac_summary[1, 3:5]) / 100
half_life <- log(2)/abs(rate)  # Half-life in weeks

# Print correlation coefficients
cat("Reduction of percentage case reduction:\n")
for (i in 1:3) {
  cat("Weekly reduction rate averaged over the first 15 weeks:", 
      round(rate[i], 3), "\n")
  cat("Percentage reduction decreases by half every ", round(half_life[i], 3),  "weeks\n")
}

rate <- as.double(prc_frac_summary[3, 3:5]) / 100
half_life <- log(2)/abs(rate)  # Half-life in weeks

# Print correlation coefficients
cat("Reduction of percentage case reduction:\n")
for (i in 1:3) {
  cat("Weekly reduction rate averaged over the first 15 weeks:", 
      round(rate[i], 3), "\n")
  cat("Percentage reduction decreases by half every ", round(half_life[i], 3),  "weeks\n")
}
```

## Figure S2: Case reduction across attack rate and duration

```{r fig3, eval=FALSE}
# (A) Onset-triggered
set.seed(42)
# sampling as large dataset makes plotting slow
sampled_runids <- sample(1:200, size=30)
  
cea_s <- filter(cea, runid %in% sampled_runids) # cea subset
save_plot <- TRUE

# Impact by duration size for outbreak onset-triggered vaccination
# Calculate percentiles for outbreak size groups
probs <- seq(0, 1, by = 0.1)
runids <- unique(cea_s$runid)
unq_data <- filter(cea_s, runid == runids[1] & vacc_cov == 0.75 & vacc_week == 1)
# size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)
ar_grp_cutoff <- quantile(outbk_data_summary$ar_per_10e3, probs = probs)
dur_grp_cutoff <- quantile(outbk_data_summary$duration_weeks, probs = probs)

# Prepare data by outbreak size group
d <- cea_s |>
  group_by(vacc_week, vacc_cov, id_outbreak) |> 
  summarize(ar_per_10e3 = ar_per_10e3[1],
            duration_weeks = duration_weeks[1],
            pct_reduc_case = median(pct_reduc_case)) # median across simulation runs

# Assign size group to each outbreak
d$ar_grp <- sapply(d$ar_per_10e3, function(x) get_group_id(x, ar_grp_cutoff))
d$dur_grp <- sapply(d$duration_weeks, function(x) get_group_id(x, dur_grp_cutoff))

# Create nested data by size group
lst <- vector("list", length(ar_grp_cutoff) - 1)
d <- setDT(d)
for(i in 1:length(lst)){
  lst[[i]] <- d[ar_grp >= i]
}

# Calculate median impact by size group
res_list <- vector('list', length(ar_grp_cutoff) - 1)
for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    ar_grp = min(ar_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm = TRUE)),
    by = .(vacc_week, vacc_cov)]
}
res_rbind <- rbindlist(res_list)

# Create heatmap of impact by outbreak size and vaccination week
plt3sa <- res_rbind |>
  filter(vacc_cov %in% VC, vacc_week %in% VW_seq) |>
  ggplot(aes(x = vacc_week, y = ar_grp)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case)) +
  metR::geom_contour2(aes(z = pct_reduc_case, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)", limits = c(0, 100),
                       breaks = seq(0, 100, 50),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Delay in ORI implementation (weeks)", 
       y = "Attack rate percentile") +
  scale_x_continuous(breaks = seq(1, 15, 2))+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() +
  theme(legend.position = "bottom")

# Create nested data by size group
lst <- vector("list", length(dur_grp_cutoff) - 1)
d <- setDT(d)
for(i in 1:length(lst)){
  lst[[i]] <- d[dur_grp >= i]
}

# Calculate median impact by size group
res_list <- vector('list', length(dur_grp_cutoff) - 1)
for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    dur_grp = min(dur_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm = TRUE)),
    by = .(vacc_week, vacc_cov)]
}
res_rbind <- rbindlist(res_list)

# Create heatmap of impact by outbreak size and vaccination week
plt3sb <- res_rbind |>
  filter(vacc_cov %in% VC, vacc_week %in% VW_seq) |>
  ggplot(aes(x = vacc_week, y = dur_grp)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case)) +
  metR::geom_contour2(aes(z = pct_reduc_case, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)", limits = c(0, 100),
                       breaks = seq(0, 100, 50),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Delay in ORI implementation (weeks)", 
       y = "Duration percentile") +
  scale_x_continuous(breaks = seq(1, 15, 2))+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() +
  theme(legend.position = "bottom")


# Combine all four plots into a single figure
plt3 <- (plt3sa + plt3sb) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              nrow = 1,
              guides = "collect") &
  theme(plot.tag.position = c(0.05, 1),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Adjust tag positions for top panels
for (i in c(1, 2)) {
  plt3[[i]] <- plt3[[i]] + theme(plot.tag.position = c(0.05, 0.9))
}

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/figS2_onset_triggered_AR_Dur_", tstamp(), ".pdf"),
         plt3,
         device = cairo_pdf, width = fgw2*fac, 
         height = fgw1*1.15*fac, units = "mm")
}

```

## Figure 4: Case Reduction in Case-Triggered Scenario

```{r}
save_plot <- TRUE
set.seed(42)
sample_runid <- sample(1:200, 30)
cea_ct_s <- filter(cea_ct, runid %in% sample_runid)

wks <- c(1,5,9)

cea_ct_s %>%
  filter(vacc_cov %in% VC, vacc_week %in% wks, case_trigger %in% c(1,100)) %>% 
  group_by(vacc_week, case_trigger) %>% 
  summarize_quantiles(pct_reduc_case)

plt4a <- cea_ct_s %>%
  filter(vacc_cov %in% VC, vacc_week %in% wks, 
         case_trigger %in% c(1, seq(20,100,20))) %>%
  ggplot() +
  geom_boxplot(aes(x = as.factor(case_trigger), y = pct_reduc_case,
             color = as.factor(vacc_week)),
             width = 0.6, position = position_dodge(width = 0.7),
             outlier.size = 0.6) +
  scale_y_continuous(limits=c(0,100)) +
  scale_color_manual(name = "Delay in ORI implementation (weeks)",
                     values = my_discrete_colors[4:6],
                     labels = as.character(wks)) + 
  labs(x = "Number of cases that triggered the ORI", y = "Cases averted (%)") + 
  theme_light() +
  theme_tight() 

# Impact by outbreak size for case-triggered vaccination
# Prepare data for case-triggered analysis
d <- cea_ct_s |> 
  group_by(vacc_week, case_trigger, vacc_cov, id_outbreak) |> 
  summarize(s_ch_tot = s_ch_tot[1],
            med_pct_reduc_case = median(pct_reduc_case))  # median across runid's

probs <- seq(0, 1, by = 0.1)
runids <- unique(cea_ct_s$runid)
unq_data <- filter(cea_ct_s, runid == runids[1] 
                   & vacc_cov == 0.75 & vacc_week == 1)
# size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)
size_grp_cutoff <- quantile(unq_data$s_ch_averted_tot, probs = probs)
# Assign size group to each outbreak
d$size_grp <- sapply(d$s_ch_tot, function(x) get_group_id(x, size_grp_cutoff))

# Create nested data by size group
lst <- vector("list", length(size_grp_cutoff) - 1)
d <- setDT(d)
for(i in 1:length(lst)){
  lst[[i]] <- d[size_grp >= i]
}

# Calculate median impact by size group
res_list <- vector('list', length(size_grp_cutoff) - 1)
for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    size_grp = min(size_grp),
    med_pct_reduc_case = median(med_pct_reduc_case, na.rm = TRUE)),
    by = .(vacc_week, case_trigger, vacc_cov)]
}
res_rbind <- rbindlist(res_list)
res_rbind %>%
  filter(size_grp >2 ) %>% 
  group_by(vacc_week, vacc_cov, size_grp) %>%
  mutate(reference = med_pct_reduc_case[case_trigger == 1]) %>%
  mutate(rel_med_pct_reduc_case = med_pct_reduc_case/reference) -> tmp

tmp %>% 
  filter(vacc_week %in% c(5), vacc_cov ==0.75, size_grp %in% c(3, 10), 
         case_trigger %in% c(1,100)) %>%
  group_by(size_grp, case_trigger) %>%
  summarize(mprc = med_pct_reduc_case)

z# Create heatmap of impact by outbreak size and case trigger
plt4b <- res_rbind |>
  filter(vacc_cov %in% VC, vacc_week %in% c(5)) |>
  ggplot(aes(x = case_trigger, y = size_grp)) +
  metR::geom_contour_fill(aes(z = med_pct_reduc_case)) +
  metR::geom_contour2(aes(z = med_pct_reduc_case, label = stat(level)),
                     linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)", limits = c(0, 60),
                       breaks = seq(0, 60, 20),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Number of cases that triggered the ORI", y = "Total cases percentile") +
  scale_x_continuous()+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() 

# Combine all four plots into a single figure
plt4 <- (plt4a + plt4b) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              nrow = 1,
              guides = "collect") &
  theme(plot.tag.position = c(0, 0.92),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig4_case_trigger_outbk_size_", tstamp(), ".pdf"),
         plt4,
         device = cairo_pdf, width = fgw2*fac, 
         height = fgw1*1.15*fac, units = "mm")
}
```

## Figure S4: Case Reduction in Case-Triggered Scenario across attack rate and duration

```{r}
save_plot <- TRUE
set.seed(42)
sample_runid <- sample(1:200, 30)
cea_ct_s <- filter(cea_ct, runid %in% sample_runid)

wks <- c(1,5,9)

cea_ct_s %>%
  filter(vacc_cov %in% VC, vacc_week %in% wks, case_trigger %in% c(1,100)) %>% 
  group_by(vacc_week, case_trigger) %>% 
  summarize_quantiles(pct_reduc_case)

cea_ct_s <- left_join(cea_ct_s,
                 outbk_data_summary[,c("id_outbreak",
                                       "ar_per_10e3","duration_weeks")], 
                 by = "id_outbreak")

# Impact by duration size for outbreak onset-triggered vaccination
# Calculate percentiles for attack rate groups
probs <- seq(0, 1, by = 0.1)
runids <- unique(cea_ct_s$runid)
unq_data <- filter(cea_ct_s, runid == runids[1] & vacc_cov == VC & vacc_week == 1)
ar_grp_cutoff <- quantile(outbk_data_summary$ar_per_10e3, probs = probs)
dur_grp_cutoff <- quantile(outbk_data_summary$duration_weeks, probs = probs)

# Prepare data by outbreak size group
d <- cea_ct_s |> 
  group_by(vacc_week, case_trigger, vacc_cov, id_outbreak) |> 
  summarize(ar_per_10e3 = ar_per_10e3[1],
            duration_weeks = duration_weeks[1],
            med_pct_reduc_case = median(pct_reduc_case)) # median across simulation runs

# Assign size group to each outbreak
d$ar_grp <- sapply(d$ar_per_10e3, function(x) get_group_id(x, ar_grp_cutoff))
d$dur_grp <- sapply(d$duration_weeks, function(x) get_group_id(x, dur_grp_cutoff))


# Create nested data by size group
lst <- vector("list", length(ar_grp_cutoff) - 1)
d <- setDT(d)
for(i in 1:length(lst)){
  lst[[i]] <- d[ar_grp >= i]
}

d %>%
  filter(vacc_cov %in% VC, vacc_week %in% c(5), case_trigger %in% c(1,100), 
         ar_grp %in% c(1,10)) %>% 
  group_by(vacc_week, case_trigger) %>% 
  summarize_quantiles(med_pct_reduc_case)

# Calculate median impact by size group
res_list <- vector('list', length(ar_grp_cutoff) - 1)
for (i in 1:length(lst)) {
  el <- lst[[i]]
  res_list[[i]] <- el[, .(
    ar_grp = min(ar_grp),
    med_pct_reduc_case = median(med_pct_reduc_case, na.rm = TRUE)),
    by = .(vacc_week, case_trigger, vacc_cov)]
}
res_rbind <- rbindlist(res_list)

res_rbind %>%
  group_by(vacc_week, vacc_cov, ar_grp) %>%
  mutate(reference = med_pct_reduc_case[case_trigger == 1]) %>%
  mutate(rel_med_pct_reduc_case = med_pct_reduc_case/reference) -> tmp

tmp %>% 
  filter(vacc_week %in% c(5), vacc_cov ==0.75, ar_grp %in% c(1, 10), 
         case_trigger %in% c(1,100)) %>%
  group_by(ar_grp, case_trigger) %>%
  summarize(mprc = med_pct_reduc_case)

# Create heatmap of impact by outbreak size and case trigger
plt4as <- res_rbind |>
  filter(vacc_cov %in% VC, vacc_week %in% c(5)) |>
  ggplot(aes(x = case_trigger, y = ar_grp)) +
  metR::geom_contour_fill(aes(z = med_pct_reduc_case)) +
  metR::geom_contour2(aes(z = med_pct_reduc_case, label = stat(level)),
                     linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)", limits = c(0, 70),
                       breaks = seq(0, 60, 20),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Number of cases that triggered the ORI", y = "Attack rate percentile") +
  scale_x_continuous()+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() 


# Create nested data by size group
lst <- vector("list", length(dur_grp_cutoff) - 1)
for(i in 1:length(lst)){
  lst[[i]] <- d[dur_grp >= i]
}

lst[[1]] %>%
   filter(vacc_cov %in% VC, vacc_week %in% c(5), case_trigger %in% c(1,100), 
          ar_grp %in% c(1,10)) %>% 
   group_by(vacc_week, case_trigger, ar_grp) %>% 
   summarize_quantiles(med_pct_reduc_case)


# Calculate median impact by dur group
res_list <- vector('list', length(dur_grp_cutoff) - 1)
for (i in 1:length(lst)) {
  el <- lst[[i]]
  res_list[[i]] <- el[, .(
    dur_grp = min(dur_grp),
    med_pct_reduc_case = median(med_pct_reduc_case, na.rm = TRUE)),
    by = .(vacc_week, case_trigger, vacc_cov)]
}
res_rbind <- rbindlist(res_list)

# Create heatmap of impact by outbreak size and case trigger
plt4bs <- res_rbind |>
  filter(vacc_cov %in% VC, vacc_week %in% c(5)) |>
  ggplot(aes(x = case_trigger, y = dur_grp)) +
  metR::geom_contour_fill(aes(z = med_pct_reduc_case)) +
  metR::geom_contour2(aes(z = med_pct_reduc_case, label = stat(level)),
                     linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)", limits = c(0,70),
                       breaks = seq(0, 60, 20),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Number of cases that triggered the ORI", y = "Duration percentile") +
  scale_x_continuous()+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() 

# Combine all four plots into a single figure
plt4s <- (plt4as + plt4bs) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              nrow = 1,
              guides = "collect") &
  theme(plot.tag.position = c(0, 0.9),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig4S_case_trigger_ar_dur_grp_", tstamp(), ".pdf"),
         plt4s,
         device = cairo_pdf, width = fgw2*fac, 
         height = fgw1*1.15*fac, units = "mm")
}
```

# Aggregated Outbreaks

## Figure 5: OCV Impact aggregated across $n$ oubreaks

```{r fig4, eval=FALSE}
cea_subsets_by_size %>%
  filter(vacc_cov == 0.75 & sample_size %in% c(1,10,50,100) &
           vacc_week %in% c(1,9)) %>% 
  group_by(sample_size, vacc_week, vacc_cov) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> pcr1

format_median_iqr(pcr1)
format_median_iqr(pcr1, row=7)

pcr1 %>% 
  filter(vacc_week == 9) %>%
  group_by(sample_size) %>% 
  summarize(median_pcr = q500)

cea_subsets_by_size %>%
  filter(vacc_cov == 0.75 & sample_size %in% c(1,10,50,100) &
           vacc_week %in% c(1,9)) %>% 
  group_by(sample_size, vacc_week, vacc_cov) %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) -> capd1

format_median_iqr(capd1, digits=3)
format_median_iqr(capd1, row=7, digits=3)

capd1 %>% 
  filter(vacc_week == 9) %>%
  group_by(sample_size) %>% 
  summarize(median_pcr = q500)

plt5a <- cea_subsets_by_size %>%
  filter(vacc_cov == 0.75 & vacc_week %in% seq(1,15,2) & 
           sample_size %in% c(1, 10, 50)) %>% 
  ggplot(aes(x = as.factor(vacc_week), y = pct_reduc_case, 
             color = as.factor(sample_size))) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6)+
  # scale_y_continuous(labels = NULL) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_color_manual(name = "# outbreaks",
                     values = my_discrete_colors[1:3],
                     labels = c(1,10,50)) +
  labs(x = "Delay in ORI implementation (week)", y = "Cases averted (%)") + 
  theme_light()+
  theme_tight_legend()+
  theme(legend.position = "top")

plt5b <- cea_subsets_by_size %>%
  filter(vacc_cov == 0.75 & vacc_week %in% seq(1,15,2) &
           sample_size %in% c(1, 10, 50)) %>% 
  ggplot(aes(x = as.factor(vacc_week), y = case_averted_per_1000_OCV, 
             color = as.factor(sample_size))) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6)+
  # scale_y_continuous(labels = NULL) +
  scale_y_continuous(limits=c(0,3)) +
  scale_color_manual(name = "# outbreaks",
                     values = my_discrete_colors[1:3],
                     labels = c(1,10,50)) +
  
  labs(x = "Delay in ORI implementation (week)", 
       y = "Cases averted per 1,000 OCVs") + 
  theme_light() +
  theme_tight_legend()+
  theme(legend.position = "top")

# Combine all four plots into a single figure
plt5 <- (plt5a + plt5b) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              nrow = 1,
              guides = "collect") &
  theme(plot.tag.position = c(0, 0.92),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig5_ORI_impact_aggregated_", tstamp(), ".pdf"),
         plt5, device = cairo_pdf,  width = fgw2*fac, 
         height = fgw1*1.15*fac, units = "mm")
}
```

## Figure 6: Cumulative OCV Impact during OCV stockpile deployment

```{r fig4, eval=FALSE}
# Reference values for individual assessment
cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(s_ch_averted_tot) 

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(case_averted_per_1000_OCV)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(cost_per_daly_averted)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(cost_per_case_averted)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week) |>
  summarize(vacc_cov = vacc_cov[1],
          count = n(),
          median_prc = median(pct_reduc_case, na.rm=T),
          median_capd = median(case_averted_per_1000_OCV, na.rm=T),
          median_cpca = median(cost_per_case_averted, na.rm=T)) -> ref_val

# cases averted
cea_subsets %>%
  filter(vacc_cov == 0.75 & vacc_week == 9 & year_max %in% c(2013, 2020)) %>% 
  group_by(year_max) %>% 
  summarize_quantiles(variable=s_ch_averted_tot) %>% 
  format_median_interval(nrow=2)

# percent case reduction
cea_subsets %>%
  filter(vacc_cov == 0.75 & vacc_week == 9 & year_max == 2020) %>% 
  summarize_quantiles(variable=pct_reduc_case) %>% 
  format_median_interval()

cea_subsets %>%
  group_by(subset_id) %>% 
  summarize(max_id = max(runid))
  
# weekly % reduction in percentage case reduction
cea_subsets_diff <- cea_subsets %>%
  group_by(subset_id, runid, vacc_cov) %>%
  mutate(pct_reduc_case_frac = c(1, 1-frac_diff(pct_reduc_case, lag=1)),
         pct_reduc_case_diff = c(1, -diff(pct_reduc_case, lag=1)))

weekly_reduc <- cea_subsets_diff %>%
  # filter(subset_id %in% 1:10) %>% 
  filter(vacc_cov == 0.75 & vacc_week > 1 & vacc_week <= 15 & year_max == 2020) %>% 
  summarize_quantiles(variable=pct_reduc_case_frac*100) 

weekly_reduc %>% 
  format_median_interval(digits=1)

weekly_reduc %>%  # Half-life in weeks
  summarize_quantiles(variable = log(2)/abs(1/as.numeric(weekly_reduc$q500))) %>% 
  format_median_interval(digits=1)

# case averted per 1,000 OCVs
cea_subsets %>%
  filter(vacc_cov == 0.75 & vacc_week == 9 & year_max == 2020) %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) %>% 
  format_median_interval(digits=3)

# cost ase averted per 1,000 OCVs
cea_subsets %>%
  filter(vacc_cov == 0.75 & vacc_week == 9 & year_max == 2020) %>% 
  summarize_quantiles(variable=cost_per_case_averted) %>% 
  format_median_interval(digits=3)

# (A) OCV deployment over time# (A) Oq500CV deployment over time
plt_a1 <- ocv_year |>
  filter(year <= 2020) |> 
  ggplot(aes(x = as.factor(year), y = cum_num_ocv)) +
  geom_point() +
  labs(x = "", y = "Cumulative\nOCVs deployed") +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

cea_subsets_s <- cea_subsets[cea_subsets$runid %in% sample_runids,] #subsetting for plotting
# (B) Cases averted
plt_b1 <- cea_subsets_s |>
  ggplot(aes(x = as.factor(year_max), y = s_ch_averted_tot + 1)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  labs(x = "", y = "Cumulative\ncases averted") +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (C) Percent case reduction
plt_c1 <- cea_subsets_s |>
  ggplot(aes(x = as.factor(year_max), y = pct_reduc_case)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_prc), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "% reduction\nin cases") +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (D) Cases averted per 1000 OCVs
plt_d1 <- cea_subsets_s |>
  ggplot(aes(x = as.factor(year_max), y = case_averted_per_1000_OCV)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_capd), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_continuous(limits = c(0,3)) +
  scale_color_manual(name = "",
                   values = c("firebrick", "steelblue"),
                   labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "Cases averted\nper 1,000 OCVs") +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (E) Cost per case averted
plt_e1 <- cea_subsets_s |>
  ggplot(aes(x = as.factor(year_max), y = cost_per_case_averted)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_cpca), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_log10(limits = c(1, 1e5)) +
  scale_color_manual(name = "",
                   values = c("firebrick", "steelblue"),
                   labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "Cost per case\naverted (US$)") +
  theme_light() +
  theme_tight()

# Combine all panels into a single figure
plt6 <- wrap_plots(list(plt_a1, plt_b1, plt_c1, plt_d1, plt_e1), ncol = 1,
                  guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0.13, 0.92),
        plot.tag = element_text(size = 12),
        legend.position = "top")

# Save figure
if (save_plot) {
  ggsave(paste0("plots/fig6_cumulative_impact_", tstamp(), ".pdf"),
         plt6, device = cairo_pdf, width=fgw2*fac*1.2, 
         height=fgw2*fac, units="mm")
}

cea_subsets %>% 
  filter(vacc_week == 9, vacc_cov == 0.9, year_max %in% c(2013, 2020)) %>%
  group_by(year_max) %>% 
  summarize_quantiles(variable=s_ch_averted_tot) -> ca1

format_median_iqr(ca1, row=1)
format_median_iqr(ca1, row=2)

cea_subsets %>% 
  filter(vacc_week == 9, vacc_cov == 0.9, year_max %in% c(2013, 2020)) %>%
  group_by(year_max) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> pcr1

format_median_iqr(pcr1, row=1)
format_median_iqr(pcr1, row=2)

cea_subsets %>% 
  filter(vacc_week == 9, vacc_cov == 0.9, year_max %in% c(2013, 2020)) %>%
  group_by(year_max) %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) -> capd1

format_median_iqr(capd1, row=1, digits=3)
format_median_iqr(capd1, row=2, digits=3)

cea_subsets %>% 
  filter(vacc_week == 9, vacc_cov == 0.9, year_max %in% c(2013, 2020)) %>%
  group_by(year_max) %>% 
  summarize_quantiles(variable=cost_per_daly_averted) -> cpda1

format_median_iqr(cpda1, row=1, digits=3)
format_median_iqr(cpda1, row=2, digits=3)

cea %>% 
  filter(vacc_week == 9, vacc_cov == 0.9) %>%
  summarize_quantiles(case_averted_per_1000_OCV)

cea %>% 
  filter(vacc_week == 9, vacc_cov == 0.9) %>%
  summarize_quantiles(cost_per_daly_averted)
 
```


```{r fig4, eval=FALSE}
# Filter subset analysis results for easier visualization
# cea_subsets |> filter(vacc_week == 9, vacc_cov == 0.75) -> dat
# cea_subsets_case_wt |> filter(vacc_week == 9, vacc_cov == 0.75) -> dat_wt

# Reference values for individual assessment
cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(s_ch_averted_tot) 

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(case_averted_per_1000_OCV)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(cost_per_daly_averted)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(cost_per_case_averted)


# Combine random and case-weighted results
dat$sampling <- "Random"
dat_wt$sampling <- "Case-weighted"
d <- rbind(dat, dat_wt)

d <- left_join(subsets, cea, by="id_outbreak")
# (A) OCV deployment over time
plt_a1 <- ocv_year |>
  filter(year <= 2022) |> 
  ggplot(aes(x = as.factor(year), y = cum_num_ocv)) +
  geom_point() +
  labs(x = "", y = "Cumulative\nOCVs deployed") +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (B) Cases averted
plt_b1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = s_ch_averted_tot + 1, color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  labs(x = "", y = "Cumulative\ncases averted") +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (C) Percent case reduction
plt_c1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = pct_reduc_case, color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_prc), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "% reduction\nin cases") +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (D) Cases averted per 1000 OCVs
plt_d1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = case_averted_per_1000_OCV, 
             color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_capd), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  # scale_y_continuous(limits = c(0, 2)) +
  scale_color_manual(name = "",
                   values = c("firebrick", "steelblue"),
                   labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "Cases averted\nper 1,000 OCVs") +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (E) Cost per case averted
plt_e1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = cost_per_case_averted, color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_cpca), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_log10(limits = c(1, 1e5)) +
  scale_color_manual(name = "",
                   values = c("firebrick", "steelblue"),
                   labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "Cost per case\naverted (US$)") +
  theme_light() +
  theme_tight()

# Combine all panels into a single figure
plt6 <- wrap_plots(list(plt_a1, plt_b1, plt_c1, plt_d1, plt_e1), ncol = 1,
                  guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0.13, 0.92),
        plot.tag = element_text(size = 12),
        legend.position = "top")

# Save figure
if (save_plot) {
  ggsave(paste0("plots/fig4_cumulative_impact_", tstamp(), ".pdf"),
         plt6, device = cairo_pdf, width = 70*2.4, height = 70*2.4, units = "mm")
}

d %>% 
  filter(year_max %in% c(2013, 2022)) %>%
  group_by(year_max, sampling) %>% 
  summarize_quantiles(variable=s_ch_averted_tot) -> ca1

format_median_iqr(ca1, row=2)
format_median_iqr(ca1, row=4)

d %>% 
  filter(year_max %in% c(2013, 2022)) %>%
  group_by(year_max, sampling) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> pcr1

format_median_iqr(pcr1, row=2)
format_median_iqr(pcr1, row=4)

d %>% 
  filter(year_max %in% c(2013, 2022)) %>%
  group_by(year_max, sampling) %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) -> capd1

format_median_iqr(capd1, row=2)
format_median_iqr(capd1, row=4)

d %>% group_by(year_max, sampling) %>% 
  summarize(case_averted_per_1000_OCV = median(case_averted_per_1000_OCV))
  
```

# Result Tables

## Table 3: Impact of Outbreak Onset-Triggered Vaccination

```{r table3}
# Create summary for time-triggered vaccination at different times
wks <- c(1, 5, 9, 13)
tab <- data.frame()
interval_est <- "IQR"
for (wk in wks) {
  d <- cea[(vacc_cov == VC & vacc_week == wk)]
  tab <- rbind(tab, impact_summary(d, probs=c(0.25, 0.5, 0.75)))
}  
tab$week <- wks

if (save_table) {
  fwrite(tab, paste0("outputs/table3_onset_triggered_impact_IQR_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```

## Table S1: Impact of Outbreak Onset-Triggered Vaccination

```{r tableS1}
# Create summary for time-triggered vaccination at different times
wks <- c(1, 5, 9, 13)
tab <- data.frame()

for (vc in VC3) {
  for (wk in wks) {
    d <- cea[(vacc_cov == vc & vacc_week == wk)]
    tab <- rbind(tab, impact_summary(d, probs=c(0.25, 0.5, 0.75)))
  }  
}

tab$week <- rep(wks, length(VC3))
tab$vc <- rep(VC3, each=length(wks))

if (save_table) {
  fwrite(tab, paste0("outputs/tableS1_onset_triggered_impact_IQR_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```

## Table 4: Impact of Case-Triggered Vaccination

```{r table3}
# Create summary for case-triggered vaccination at different thresholds
vwks <- c(1, 5, 9, 13)
ctgs <- c(1, 10, 50, 100)
tab <- data.frame()

for (wk in vwks) {
  for (ctg in ctgs) {
    d <- cea_ct[(vacc_cov == VC & vacc_week == wk & case_trigger == ctg)]
    tab <- rbind(tab, impact_summary(d, probs=c(0.25, 0.5, 0.75)))
  }
}
tab$case_trigger <- rep(ctgs, length(vwks))
tab$vacc_week <- rep(vwks, each=length(ctgs))

if (save_table) {
  fwrite(tab, paste0("outputs/table4_case_triggered_impact_IQR_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```

## Table S2: Impact of global OCV stockpile, 2013-2020
```{r tableS3}

cea_ocv_stockpile_w9vc75 <- cea_subsets[vacc_week == 9 & vacc_cov == 0.75]

cea_subsets %>% 
  filter(vacc_week == 9 & vacc_cov == 0.75) %>% 
  group_by(year_max) %>% 
  summarize(capd_med = median(case_averted_per_1000_OCV, na.rm=T))

# Filter the subset data for year 2022 (most recent data)
dat_2020 <- cea_subsets[year_max == 2020 & vacc_week == 9 & vacc_cov == 0.75]
impact_summary(dat_2020)
# dat_wt_2022 <- cea_subsets_case_wt[year_max == 2022 & vacc_week == 9 & vacc_cov == 0.75]

vwks <- c(1, 5, 9, 13)
# Calculate summary statistics
tab <- data.frame()
for (vc in VC3) {
  for (wk in wks) {
    d <- cea_subsets[year_max == 2020 & vacc_week == wk & vacc_cov == vc]
    tab <- rbind(tab, impact_summary(d, probs=c(0.25, 0.5, 0.75)))
  }  
}
tab$week <- rep(wks, length(VC3))
tab$vc <- rep(VC3, each=length(wks))

if (save_table) {
  write.csv(tab, paste0("outputs/tableS2_outbreak_subsets_IQR_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```

