---
title: "Plots and tables"
format: html
editor_options: 
  chunk_output_type: console
---

## Load packages, source `utils.R`, and outbreak data sets

```{r}
library(data.table)
library(tidyverse)
library(ggridges)
library(readxl)
library(janitor)
library(scales)
library(viridis)
library(sf)
library(metR)
library(googlesheets4)
library(conflicted)
library(MASS)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
source("utils.R") # utility functions
save_plot <- FALSE
save_table <- FALSE
# figure size
fgw1 <- figure_size[figure_size$journal =="Lancet",]$single
fgw2 <- figure_size[figure_size$journal =="Lancet",]$double
set1 <- RColorBrewer::brewer.pal(9,"Set1") # color sets
library(patchwork)
# remove the space between the plot and the legend and the space around the plot
# representative probability values for quantile 
probs <- c(0.025, 0.25, 0.5, 0.75, 0.975)
VW <- 9 # representative vaccination week
VW_seq <- seq(1, 15, 2)
VC <- 0.75 # representative vaccination coverage rate
VC3 <- c(0.6, 0.75, 0.9) # three representative vaccination coverage rates

theme_set(theme_gray(base_size=14))

shp <- st_read("data/shapefiles/total_shp_0427.shp")
africa <- st_read("data/shapefiles/afr_g2014_2013_0.shp")
```

## Outbreak data and delay
```{r}
outbreak_data_all <- as.data.frame(fread("outputs/outbreak_data_all_20241121.csv"))
outbreak_data <- fread("outputs/outbreak_data_20241121.csv") # for vaccine impact modeling

# fwrite(ocv_use, paste0("outputs/ocv_use_GTFCC_data_", tstamp(), ".csv"))
ocv_use <- fread("outputs/ocv_use_GTFCC_data_20241120.csv")
delay_req_r1 <- as.integer(ocv_use$DATE_ROUND_ONE - ocv_use$DATE_REQUEST)
# max_pop_camp <- max(ocv_use$DOSE_DELIVERED_ROUND_ONE*1000, na.rm=T)
min_pop_camp <- min(ocv_use$DOSE_DELIVERED_ROUND_ONE*1000, na.rm=T)
ocv_delay <- quantile(delay_req_r1, probs=c(0.25,0.5,0.75,0.9), na.rm=TRUE)

# OCV use data
docv <- clean_names(fread("data/outbreak_files_Oct2024/ocv_long_dataset.csv"))
# consumer price index for cost adjustment (2015 to 2023)
dis_params <- as.data.frame(data.table::fread("data/parameters.csv"))
# IVE posterior predictive values from Bayesian beta regression (see parameters.qmd)
ive_fit <- read_rds("outputs/ive_yrep_20241010.rds")
ive_data <- ive_fit$yrep
rm(ive_fit) # large file size

# extract ids for the outbreaks not affected by OCV campaigns
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Oct2024/outbreak_summary.csv")))
paste0(length(unique(ds$country)), " countries")
# create 'id_outbreak' variables to link with the 
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

outbk_data_summary <- left_join(filter(outbreak_data, week == 1), ds, by="id_outbreak")
summary(outbk_data_summary$ar_per_10e3)
summary(outbk_data_summary$outbreak_pop)

```

## VIM results
```{r}
svim <- fread("outputs/svim_mean_across_runid_20241126.csv")
svim_ct <- fread("outputs/svim_ct_mean_across_runid_20241126.csv")
svim_pe <- fread("outputs/svim_pe_mean_across_runid_20241126.csv")
svim_pe_sum <- fread("outputs/svim_pe_sum_across_week_by_runid_20241126.csv")

svim_pe2 <- fread("outputs/svim_pe_mean_across_runid_20241217.csv")
svim_pe_sum2 <- fread("outputs/svim_pe_sum_across_week_by_runid_20241217.csv")

svim_cea <- fread("outputs/svim_mean_cea_across_runid_20241126.csv")
svim_ct_cea <- fread("outputs/svim_ct_mean_cea_across_runid_20241126.csv")

subset_random <- fread("outputs/svim_mean_random_subset_ids_20241126.csv")
subset_attack_rate <- fread("outputs/svim_mean_attack_rate_subset_ids_20241126.csv")
subset_cases <- fread("outputs/svim_mean_cases_subset_ids_20241126.csv")
subset_duration <- fread("outputs/svim_mean_duration_subset_ids_20241126.csv")
```

## Figure 2 Outbreak description

### (A) Outbreak time series by group
```{r, eval=FALSE}
# outbreaks that have been excluded
ids_excl <- outbreak_data_all |> 
  filter(!(id_outbreak %in% outbreak_data$id_outbreak)) |>  
  pull(id_outbreak) |>
  unique()

# id's of the outbreaks that were confirmed
ids_conf <- outbreak_data |> 
  group_by(id_outbreak) |> 
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique()

# id's of the outbreaks that were suspected but not confirmed
ids_susp <- outbreak_data |>
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |>
  filter(!(c_ch_sum > 0)) |>
  pull(id_outbreak) |>
  unique()

# id's of the outbreaks that were affected by OCVs
ids_ocv <- outbreak_data_all |> 
  filter(!(id_outbreak %in% docv$id_outbreak)) |>  
  pull(id_outbreak) |>
  unique()

ggplot()+
  geom_line(data=filter(outbreak_data_all, id_outbreak %in% ids_excl),
            aes(date, s_ch, group=id_outbreak, color="Excluded"), 
            alpha=0.5)+
  geom_line(data=filter(outbreak_data, !(id_outbreak %in% ids_conf)),
            aes(date, s_ch, group=id_outbreak, color="Suspected"), 
            alpha=0.5)+
  geom_line(data=filter(outbreak_data, id_outbreak %in% ids_conf), 
            aes(date, s_ch, group=id_outbreak, color="Confirmed"),
            alpha=0.5)+
  scale_color_manual("", values=c("Excluded"="yellow4",
                                  "Suspected"="steelblue4", 
                                  "Confirmed"="firebrick4"))+
  labs(x="", y="Weekly suspected cases")+
  theme_light()+
  theme(legend.position="top") -> plt

plt

if (save_plot) {
  ggsave(paste0("plots/outbk_ts_case_type_", tstamp(), ".pdf"), plt,
         device = cairo_pdf, width=fgw2, height=fgw1*1.1, units="mm")
}

outbk_conf_summary <- outbk_data_summary[id_outbreak %in% ids_conf]
summary(outbk_data_summary$ar_per_10e3)
summary(outbk_data_summary$outbreak_pop)
```

### (B) Outbreak size vs. duration by type 

```{r, eval=FALSE}
# outbreak_data$end_date_outbreak <- as.Date(outbreak_data$end_date_outbreak)
# outbreak_data$start_date_outbreak <- as.Date(outbreak_data$start_date_outbreak)

outbreak_data_all |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                    as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d_all

outbreak_data |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                    as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d

ggplot()+
  geom_point(data=filter(d_all, id_outbreak %in% ids_excl), 
             aes(duration, total_s_ch, color="Excluded"), alpha=0.5)+
  geom_point(data=filter(d, id_outbreak %in% ids_susp), 
             aes(duration, total_s_ch, color="Suspected"), alpha=0.5)+
  geom_point(data=filter(d, id_outbreak %in% ids_conf), 
             aes(duration, total_s_ch, color="Confirmed"), alpha=0.5)+
  scale_x_log10()+
  scale_y_log10()+
  theme_light()+
  scale_color_manual("", values=c("Excluded" = "yellow4", 
                                  "Suspected" = "steelblue4", 
                                  "Confirmed" = "firebrick4"))+
  labs(x="Outbreak duration (week)", y="Total suspected cases")+
  theme(legend.position="top") -> plt
plt

if (save_plot){
  fac <- 1.25
  ggsave(paste0("plots/outbk_size_dur_case_type_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}
```

### (C) Map of outbreaks
```{r, eval=F}
dt <- outbreak_data
dt_all <- outbreak_data_all 

dt <- dt %>% rename(lctn_pr = location_period_id)

dt_all <- dt_all %>% rename(lctn_pr = location_period_id)

first_rows_df <- dt %>% 
  group_by(id_outbreak) %>% 
  slice(1) %>%   # Extracts the first row of each group
  ungroup()
length(unique(dt$id_outbreak))

first_rows_df <- dt %>%
  group_by(id_outbreak) %>%
  slice(1) %>%   # Extracts the first row of each group
  ungroup()

first_rows_df_all <- dt_all %>%
  group_by(id_outbreak) %>%
  slice(1) %>%   # Extracts the first row of each group
  ungroup()

length(unique(dt$id_outbreak))

ids_excl <- dt_all |>
  filter(!(id_outbreak %in% dt$id_outbreak)) |> 
  pull(id_outbreak) |>
  unique()

ids_conf <- dt |>
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |>
  filter(c_ch_sum > 0) |>
  pull(id_outbreak) |>
  unique()

d <- left_join(first_rows_df, shp, by = "lctn_pr")
d1 <- d_all %>% filter((id_outbreak %in% ids_excl))

d_all <- left_join(first_rows_df_all, shp, by = "lctn_pr")


d <- d %>%
  mutate(id_conf = ifelse(id_outbreak %in% ids_conf, 1, 0))
d <- d %>%
  mutate(id_sus = ifelse(!id_outbreak %in% ids_conf, 1, 0))

d_conf <- d %>%
  filter(id_conf == 1)
d_sus <- d %>% 
  filter(id_sus == 1)

d_conf <- st_as_sf(d_conf)
d_sus <- st_as_sf(d_sus)

d_all <- d_all %>%
  mutate(id_excl = ifelse(id_outbreak %in% ids_excl, 1, 0))
d <- st_as_sf(d)

d2 <- d
d_all <- st_as_sf(d_all)

a <- ggplot() +
  geom_sf(data = africa, fill = NA, color = "black") +  # Plot country outlines
  geom_sf(data = d_all, aes(fill = id_excl), color = "black", 
          fill = "yellow4", alpha=0.5) +
  geom_sf(data = d_sus, aes(fill = id_sus), color = "steelblue4", 
          fill = "steelblue4", alpha=0.5) +
  geom_sf(data = d_conf, aes(fill = id_conf), color = "firebrick4", 
          fill = "firebrick4", alpha=0.5) +
  theme_minimal() +
  labs( title = "Map of Outbreaks", fill = "Outbreak Level") # Change to your legend label)

a

if (save_plot) {
  ggsave(paste0("plots/new_map_outbreaks_", tstamp(), ".pdf"), a)
}
```

## Figure 3 Percent case reduction

### Example outbreak time series (no vaccine and vaccine) and vaccine impact

```{r, eval=FALSE}
# week_lst <- read_rds("outputs/svim_week_list_20240822.rds")
# lst <- read_rds("outputs/svim_list_20240822.rds")
week_lst <- read_rds("outputs/svim_week_list_20240813.rds")
lst <- read_rds("outputs/svim_list_20240620.rds")

for (id in unique(week_lst[[runid]]$data_id)) {
  cat("id = ", id, "\n")
  runid <- 1
  outbk_num_id <- id
  vacc_wk <- 4
  
  vacc_cv <- 0.9
  d <- week_lst[[runid]] |> 
    filter(data_id == outbk_num_id, vacc_week == vacc_wk, vacc_cov == vacc_cv)
    
  plot(d$week, d$s_ch, type="l")
  points(d$week, d$s_ch)
  lines(d$week, d$s_ch - d$s_ch_averted, lty="dotted")
  
  readline(prompt="Press [enter] to proceed")
}
# id = 22, 25, 30, 99, 105, 106, 131, 133, 209, 213-4, 259, 285
for (runid in 1:200) {
  cat("run_id = ", runid, "\n")
  d <- week_lst[[runid]] |> 
    filter(data_id == outbk_num_id, vacc_week == vacc_wk, vacc_cov == vacc_cv)
    
  plot(d$week, d$s_ch, type="l")
  points(d$week, d$s_ch)
  lines(d$week, d$s_ch - d$s_ch_averted, lty="dotted")
  
  readline(prompt="Press [enter] to proceed")
}
# runid = 57, 59, 70
d <- week_lst[[70]] |> 
  filter(data_id == 25, vacc_week == 4, vacc_cov == 0.9)
thr <- ds[ds$ID == 25, ]$threshold
d <- d[, c("date", "s_ch", "s_ch_averted")]
d$s_ch_under_vacc <- d$s_ch - d$s_ch_averted
names(d) <- c("date", "No vacc", "Averted", "Vacc")
# dlong <- pivot_longer(d, cols=c("s_ch","s_ch_under_vacc"))
dlong <- pivot_longer(d, cols=c("No vacc", "Vacc"))

dar <- data.frame(x=c(1,1), y=c(20,10)) # data for arrow
vdate <- as.Date("2010-07-11")


ggplot(dlong, aes(date, value, linetype=name, shape=name, group=name)) +
  geom_line()+
  geom_point()+
  geom_ribbon(data=subset(d, date>= as.Date("2010-08-08")), 
          aes(x=date, ymin=`Vacc`, ymax=`No vacc`), fill="steelblue", 
          alpha=0.5, inherit.aes = FALSE) +
  geom_segment(aes(x=vdate+7,y=65,xend=vdate+7,yend=20),
               arrow=arrow(type="closed", angle=10, length=unit(5,"mm")),
               inherit.aes = FALSE)+
  annotate("text", x=vdate+7, y=75, label= "Vaccination \nstarted", hjust=0.8) +
  geom_segment(aes(x=vdate+24,y=135,xend=vdate+24,yend=115),
               arrow=arrow(type="closed", angle=10, length=unit(5,"mm")),
               inherit.aes = FALSE)+
  annotate("text", x=vdate+24, y=144, label= "Vaccine\ntakes effect", hjust=0.8) +
  
  geom_segment(aes(x=vdate-12, y=35, xend=vdate-12, yend=1.8),
               arrow=arrow(type="closed", angle=10, length=unit(5,"mm")),
               inherit.aes=FALSE)+
  annotate("text", x=vdate-12, y=43, label= "Outbreak\nthreshold", hjust=0.3) +
  annotate("text", x=as.Date("2010-08-28"), y=120, 
           label= "No\nvaccination", hjust=0) +
  annotate("text", x=as.Date("2010-09-03"), y=6, 
           label= "Vaccination", hjust=1) +
  geom_segment(aes(x=as.Date("2010-09-30"),
                   y=100,xend=as.Date("2010-08-21"),
                   yend=80), arrow=arrow(type="closed", angle=10, 
                                         length=unit(5,"mm")), 
               inherit.aes = FALSE)+
  annotate("text", x=as.Date("2010-10-18"), y=110, 
           label= "Cases reduced \ndue to the vaccine") +
  geom_hline(yintercept=thr - 1, color="firebrick", linetype="dashed")+
  geom_vline(xintercept=min(dlong$date), color="steelblue",linetype="dashed")+
  geom_vline(xintercept=max(dlong$date), color="steelblue", linetype="dashed")+
  geom_vline(xintercept=as.Date("2010-09-26"), color="steelblue", linetype="dashed")+

  geom_segment(aes(x=min(dlong$date),y=158,xend=max(dlong$date),yend=158),
               arrow=arrow(type="closed", ends="both", angle=10,
                           length=unit(5,"mm")), inherit.aes = FALSE)+
  geom_segment(aes(x=min(dlong$date),y=153,xend=as.Date("2010-09-26"),
                   yend=153),  arrow=arrow(type="closed", ends="both", angle=10,
                           length=unit(5,"mm")), inherit.aes = FALSE)+
  
  geom_segment(aes(x=as.Date("2010-09-26"),y=153,xend=max(dlong$date),
                   yend=153), color="steelblue",
               arrow=arrow(type="closed", ends="both", angle=10,
                           length=unit(5,"mm")), inherit.aes = FALSE)+
  annotate("text", x=as.Date("2010-10-18"), y=140, 
           label= "Duration reduced\ndue to the vaccine") +
  labs(x="", y="Number of suspected cases", linetype="")+
  theme_light()+
  theme(legend.position = "none") -> plt

fac <- 2.0
if (save_plot) {
  ggsave(paste0("plots/vacc_eff_example_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}
```

###  Impact across VW at three different VCs

Impact from preemptive vaccination
```{r}
svim_pe_sum |> 
  filter(vacc_cov %in% VC3) |> 
  group_by(vacc_cov) |> 
  summarize(mean_pct_reduc_case = mean(pct_reduc_case),
            q500_pct_reduc_case = quantile(pct_reduc_case, probs=c(0.5)),
            q025_pct_reduc_case = quantile(pct_reduc_case, probs=c(0.025)),
            q975_pct_reduc_case = quantile(pct_reduc_case, probs=c(0.975)),
            max_pct_reduc_case = max(pct_reduc_case),
            min_pct_reduc_case = min(pct_reduc_case)) -> svim_pe_summary

paste0("maximum percent case reduction from the pre-emptive vaccination are ", 
       format_num(svim_pe_summary$max_pct_reduc_case[1], digits=1), "%, ", 
       format_num(svim_pe_summary$max_pct_reduc_case[2], digits=1), "%, and ",  
       format_num(svim_pe_summary$max_pct_reduc_case[3], digits=1), "% for vaccine coverage rate of  60%, 75%, and 90%, respectively.")

# lower vaccine coverage levels to compare with Longini (2007) et al. results
svim_pe_sum2 |> 
  group_by(vacc_cov) |> 
  summarize(mean_pct_reduc_case = mean(pct_reduc_case),
            q500_pct_reduc_case = quantile(pct_reduc_case, probs=c(0.5)),
            q025_pct_reduc_case = quantile(pct_reduc_case, probs=c(0.025)),
            q975_pct_reduc_case = quantile(pct_reduc_case, probs=c(0.975)),
            max_pct_reduc_case = max(pct_reduc_case),
            min_pct_reduc_case = min(pct_reduc_case)) 
```

Figure 3
```{r}
# time-triggered
svim |>
  filter(vacc_cov %in% VC3, vacc_week %in% VW_seq) |>  
  ggplot(aes(x=as.factor(vacc_week), y=pct_reduc_case,
             color=as.factor(vacc_cov)))+
  geom_boxplot(width=0.6, position=position_dodge(width=0.7), 
               outlier.size=0.6)+
  geom_hline(data=svim_pe_summary, aes(yintercept=max_pct_reduc_case, 
                 color=as.factor(vacc_cov)), linetype="dashed")+
  scale_y_continuous(labels = scales::percent_format(scale=1))+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:3],
                     labels=c(paste0(format_num(VC3*100,1),"%")))+
  # labs(x="Vaccination week since the outbreak began", 
  #      y="Case reduction")+ 
  labs(x="", 
       y="Case reduction")+ 
  theme_light()+
  theme_tight_legend()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  theme(legend.position = "bottom") -> plt1a


# Add a proportion of outbreaks that are missed by ORI 

svim_df <- fread("outputs/svim_sum_across_week_by_runid_20241122.csv")
svim_df |>
  filter(vacc_cov == VC, vacc_week %in% VW_seq) |>
  select(runid, id_outbreak, vacc_week, outbk_dur, week_delay_to_vacc_effect) |> 
  group_by(vacc_week, runid) |> 
  summarize(
    n_outbk = sum(outbk_dur > 0),
    n_outbk_missed = sum(outbk_dur < vacc_week),
    prop_outbk_missed = n_outbk_missed / n_outbk,
    n_outbk_missed_vacc_effect = sum(outbk_dur < week_delay_to_vacc_effect),
    prop_outbk_missed_vacc_effect = n_outbk_missed_vacc_effect / n_outbk) ->
  outbk_missed_vacc 

svim_df |>
  filter(vacc_cov == VC, vacc_week %in% VW_seq) |>
  select(id_outbreak, vacc_week, outbk_dur, week_delay_to_vacc_effect) |> 
  group_by(vacc_week) |> 
  summarize(
    n_outbk = sum(outbk_dur > 0),
    n_outbk_missed = sum(outbk_dur < vacc_week),
    prop_outbk_missed = n_outbk_missed / n_outbk,
    n_outbk_missed_vacc_effect = sum(outbk_dur < week_delay_to_vacc_effect),
    prop_outbk_missed_vacc_effect = n_outbk_missed_vacc_effect / n_outbk) ->
  outbk_missed_vacc_agg

outbk_missed_vacc$vacc_week <- as.factor(outbk_missed_vacc$vacc_week)
outbk_missed_vacc_agg$vacc_week <- 
  as.factor(outbk_missed_vacc_agg$vacc_week)

ggplot(outbk_missed_vacc) +
  geom_step(aes(vacc_week, 
                prop_outbk_missed, group=runid, color="By campaign"), 
                alpha=0.1)+
  geom_step(data=outbk_missed_vacc_agg, 
            aes(vacc_week, prop_outbk_missed, group=1,
                color="By campaign"), 
            linewidth=0.6)+
  geom_step(aes(vacc_week, prop_outbk_missed_vacc_effect, group=runid,
                color="By vaccine effect"), alpha=0.1)+
  geom_step(data=outbk_missed_vacc_agg, 
            aes(vacc_week, prop_outbk_missed_vacc_effect, 
                group=1,
                color="By vaccine effect"), linewidth=0.6)+
  scale_color_manual("", values=c("By campaign"= my_discrete_colors[4], 
                                  "By vaccine effect" = my_discrete_colors[5]))+
  scale_y_continuous("Outbreaks missed",
                     labels = scales::percent_format(scale=100),
                     breaks=seq(0,1,by=0.25),
                     limits=c(0,0.75))+
  scale_x_discrete("") +
  theme_light()+
  theme_tight_legend()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank()) -> plt1c

# outbreak 
outbreak_data |>
  group_by(week) |>
  summarize(s_ch_tot = sum(s_ch)) -> data_tot 

data_tot |> 
  mutate(s_ch_tot_all = sum(s_ch_tot), 
         s_ch_tot_pct = 100 * s_ch_tot / s_ch_tot_all, 
         s_ch_tot_pct_cumul = cumsum(s_ch_tot_pct)) -> data_tot

outbreak_data |>
  filter(week %in% VW_seq) |> 
  ggplot(aes(x=s_ch+1, y=as.factor(week))) +
  geom_density_ridges(fill="steelblue", alpha=0.5, scale=0.5) +
  # geom_point(data=filter(data_tot, week %in% VW_seq), 
  #            aes(cumsum(s_ch_tot), as.factor(week)), shape=15) +
  # geom_line(data=filter(data_tot, week %in% VW_seq), 
  #           aes(cumsum(s_ch_tot), as.factor(week)),
  #           linetype="solid", linewidth=0.7, group=1)+
  scale_x_log10() +
  coord_flip() +
  geom_line(data=filter(data_tot, week %in% VW_seq), 
            aes(s_ch_tot, as.factor(week)), 
            linetype="solid", linewidth=1, group=1) +
  geom_point(data=filter(data_tot, week %in% VW_seq), 
             aes(s_ch_tot, as.factor(week)), shape=15) +
  theme_light()+
  theme_tight_legend()+
  labs(y="Vaccination week since the outbreak began", 
       x="Suspected cases") -> plt1e


# case-triggered 
cts <- unique(svim_ct$case_trigger)
cts_ids <- c(21,25,29,2,6,10,14,18)
svim_ct |>
 filter(vacc_cov %in% VC3, case_trigger %in% cts[cts_ids]) -> dd

dd$case_trigger <- factor(dd$case_trigger, levels=c(5,seq(40,280,40)))
dd$vacc_cov <- factor(dd$vacc_cov, levels=c(0.6,0.75,0.9))

dd |>
  ggplot()+
  geom_boxplot(aes(x=case_trigger, y=pct_reduc_case,
             group=interaction(case_trigger, vacc_cov),
             color=vacc_cov),
             width=0.6, position=position_dodge(width=0.7),
             outlier.size=0.6) +
  geom_hline(data=svim_pe_summary, aes(yintercept= max_pct_reduc_case, 
                 color=as.factor(vacc_cov)), linetype="dashed")+
  scale_y_continuous(labels = scales::percent_format(scale=1))+
  # scale_y_continuous(labels = NULL)+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:3],
                     labels=c("60%","75%","90%"))+ 
  labs(x="Cumulative cases at vaccination", 
       y="Case reduction")+ 
  theme_light()+
  theme_tight_legend()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  theme(legend.position = "bottom") -> plt1b


# Add a proportion of outbreaks that are missed by ORI 
svim_df <- fread("outputs/svim_ct_sum_across_week_by_runid_20241123.csv")
# case_trigger <= cts[19] just to make consistent with boxplots of pct_reduc_case
svim_df |>
  # filter(week_vaccination > 10) |> # this is simply due to the limit of object size in R
  filter(vacc_cov == VC, case_trigger %in% cts[cts_ids]) |>
  select(runid, id_outbreak, case_trigger, vacc_week, outbk_dur, week_delay_to_vacc_effect) |> 
  group_by(case_trigger, runid) |> 
  summarize(
    n_outbk = sum(outbk_dur > 0),
    n_outbk_missed = sum(outbk_dur < vacc_week),
    prop_outbk_missed = n_outbk_missed / n_outbk,
    n_outbk_missed_vacc_effect = sum(outbk_dur < week_delay_to_vacc_effect),
    prop_outbk_missed_vacc_effect = n_outbk_missed_vacc_effect / n_outbk) ->
  outbk_missed_vacc 

svim_df |>
  filter(vacc_cov == VC, case_trigger %in% cts[cts_ids]) |>
  select(id_outbreak, vacc_week, case_trigger, outbk_dur, week_delay_to_vacc_effect) |> 
  group_by(case_trigger) |> 
  summarize(
    n_outbk = sum(outbk_dur > 0),
    n_outbk_missed = sum(outbk_dur < vacc_week),
    prop_outbk_missed = n_outbk_missed / n_outbk,
    n_outbk_missed_vacc_effect = sum(outbk_dur < week_delay_to_vacc_effect),
    prop_outbk_missed_vacc_effect = n_outbk_missed_vacc_effect / n_outbk) ->
  outbk_missed_vacc_agg

outbk_missed_vacc$case_trigger <- as.factor(outbk_missed_vacc$case_trigger) 
outbk_missed_vacc_agg$case_trigger <- 
  as.factor(outbk_missed_vacc_agg$case_trigger) 
  
ggplot(outbk_missed_vacc) +
  geom_step(aes(case_trigger, prop_outbk_missed, group=runid, 
                color="By campaign"), 
                alpha=0.1)+
  geom_step(data=outbk_missed_vacc_agg, 
            aes(case_trigger, prop_outbk_missed, group=1, color="By campaign"), 
            linewidth=0.6)+
  geom_step(aes(case_trigger, prop_outbk_missed_vacc_effect, group=runid,
                color="By vaccine effect"), alpha=0.1)+
  geom_step(data=outbk_missed_vacc_agg, 
            aes(case_trigger, prop_outbk_missed_vacc_effect,
                group=1, color="By vaccine effect"), linewidth=0.6)+
  scale_color_manual("", values=c("By campaign"= my_discrete_colors[4], 
                                  "By vaccine effect" = my_discrete_colors[5]))+
  scale_y_continuous("Outbreaks missed",
                     labels = scales::percent_format(scale=100),
                     breaks=seq(0,1,by=0.25),
                     limits=c(0,0.75))+
  theme_light()+
  theme_tight_legend()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank()) -> plt1d # the margin around each legend)


target_case <- cts[cts_ids]
lst <- vector('list', length(target_case))
outbreak_data[, cum_s_ch := cumsum(s_ch), by=id_outbreak]

for (i in 1:length(target_case)) {
  outbreak_data |>
    group_by(id_outbreak) |>
    filter(max(cum_s_ch) > target_case[i]) |>
    ungroup() |>
    filter(cum_s_ch >= target_case[i]) |>
    group_by(id_outbreak) |>
    summarize(min_week = min(week)) |>
    pull(min_week) -> lst[[i]]
}

df <- data.frame()
for (i in 1:length(target_case)) {
  d <- data.frame(cumul_case = target_case[i], week = lst[[i]])
  df <- rbind(df, d)
}

df$cumul_case <- as.factor(df$cumul_case) 

df |> 
  ggplot(aes(x=week, y=cumul_case)) +
  geom_density_ridges(fill="steelblue", alpha=0.5, scale=0.5) +
  coord_flip() +
  labs(y="Cumulative cases at vaccination", x="Weeks to reach the cumulative case")+
  theme_light()+
  theme_tight_legend()+
  theme()-> plt1f

# combine plot        
# plt1cd <- plt1d / plt1c + plot_layout(ncol = 1, heights = c(0.3,1))  
# plt1 <- plt1ab + plt1cd + plot_layout(ncol=2, axis_titles = "collect")  

# plt1 <- (plt1b + plt1d + plt1a + plt1c) +
#   plot_annotation(tag_levels = 'A') +
#   plot_layout(ncol=2,
#               axis_titles="collect",
#               guides="collect", heights=c(0.4,1)) &
#     theme(plot.tag.position  = c(0.98,0.96),
#         plot.tag = element_text(size=13)) &
#   theme(legend.position = "top")
# # 
# for (i in c(1,2)) {
#   plt1[[i]] <- plt1[[i]] + theme(plot.tag.position = c(0.98,0.68))
# }


# plt1 <- (plt1b + plt1d + plt1e + plt1f + plt1a + plt1c) +
#   plot_annotation(tag_levels = 'A') +
#   plot_layout(ncol=2,
#               # axis_titles="collect",
#               guides="collect", heights=c(0.4,1,1)) &
#   theme(plot.tag.position  = c(0.1,0.8),
#         plot.tag = element_text(size=13)) &
#   theme(legend.position = "top")

plt1 <- (plt1a + plt1b + plt1c + plt1d + plt1e + plt1f) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol=2,
              # axis_titles="collect",
              guides="collect", heights=c(1,0.4,1)) &
  theme(plot.tag.position  = c(0.1,0.8),
        plot.tag = element_text(size=13)) &
  theme(legend.position = "top")
  # theme(plot.margin = margin(0,0,0,0.1))

# 
for (i in c(1,2)) {
  plt1[[i]] <- plt1[[i]] + theme(plot.tag.position = c(0.97,0.86))
}
for (i in c(3,4)) {
  plt1[[i]] <- plt1[[i]] + theme(plot.tag.position = c(0.97,0.93))
}
for (i in c(5,6)) {
  plt1[[i]] <- plt1[[i]] + theme(plot.tag.position = c(0.97,0.97))
}

fac <- 1.3
  ggsave(paste0("plots/fig3_pcr_outbreaks_missed_", tstamp(hour=T), ".pdf"), 
         plt1,
         device = cairo_pdf, width=fgw2*fac, height=fgw2*1.15*fac, units="mm")
  
if (save_plot) {
  ggsave(paste0("plots/fig3_pcr_outbreaks_missed_", tstamp(hour=T), ".pdf"), 
         plt1,
         device = cairo_pdf, width=fgw2*fac, height=fgw2*1.15*fac, units="mm")
}
```

### Exponential model fit
```{r}
svim |>
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1:19)) |>
  group_by(vacc_week) |>
  summarize(vacc_cov = vacc_cov[1],
          count = n(),
          mean_prc = mean(pct_reduc_case),
          sd_prc = sd(pct_reduc_case), 
          median_prc = quantile(pct_reduc_case, probs=c(0.5)),
          q1_prc = quantile(pct_reduc_case, probs=c(0.25)),
          q3_prc = quantile(pct_reduc_case, probs=c(0.75)),
          min_prc = min(pct_reduc_case),
          max_prc = max(pct_reduc_case)) -> res 

res$t <- 1:19
f <- function(t, Asym=0, lrc, y0) {
    Asym + (y0 - Asym) * exp(-exp(lrc) * t)
}

fit <- nls(median_prc ~ f(t=t, Asym=0, lrc, y0),  
           data = res, start = list(lrc = -1, y0=86))
fit
# weekly reduction rate
exp(as.numeric(coef(fit)["lrc"]))
# half-life (weeks)
log(2)/abs(exp(as.numeric(coef(fit)["lrc"])))

ggplot(res)+
  geom_point(aes(x=t, y=median_prc, color="Data")) +
  geom_line(aes(x=t, y=fitted(fit), color="Fitted")) +
  labs(y="Median case averted", x="Vaccinatino week since the outbreak start")+
  theme_light()+
  scale_color_manual("",values=c("Data"="steelblue","Fitted"="firebrick"))+
  theme_tight() -> plt

fac <- 1.5
if (save_plot)
  ggsave(paste0("plots/median_ca_exp_decay_", tstamp(), ".pdf"), plt,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")

# case-triggered vaccination
svim_ct |>
  filter(vacc_cov %in% c(0.75)) |>
  group_by(case_trigger) |>
  summarize(vacc_cov = vacc_cov[1],
          count = n(),
          mean_prc = mean(pct_reduc_case),
          sd_prc = sd(pct_reduc_case), 
          median_prc = quantile(pct_reduc_case, probs=c(0.5)),
          q1_prc = quantile(pct_reduc_case, probs=c(0.25)),
          q3_prc = quantile(pct_reduc_case, probs=c(0.75)),
          min_prc = min(pct_reduc_case),
          max_prc = max(pct_reduc_case)) -> res 

res$t <- res$case_trigger
f <- function(t, Asym=0, lrc, y0) {
    Asym + (y0 - Asym) * exp(-exp(lrc) * t)
}
res_sub <- filter(res, res$case_trigger <= 200)
fit <- nls(median_prc ~ f(t=t, Asym=0, lrc, y0),
           data = res_sub, start = list(lrc=log(0.01), y0=100))
fit
# reduction rate by unit (week or a single case)
a <- exp(as.numeric(coef(fit)["lrc"]))
# half-life (weeks)
b <- log(2)/abs(exp(as.numeric(coef(fit)["lrc"])))
paste0("Medican case reduction (%) decreases with increasing number of cumulative cases with ", format_num(a*100, digits=1) , "% by one case increase, and decreasing by half every ",  format_num(b, digits=1), " additional cases.")

ggplot(res_sub)+
  geom_point(aes(x=t, y=median_prc, color="Data")) +
  geom_line(aes(x=t, y=fitted(fit), color="Fitted")) +
  labs(y="Median case averted", x="Cumulative cases at vaccination")+
  theme_light()+
  scale_color_manual("",values=c("Data"="steelblue","Fitted"="firebrick"))+
  theme_tight() -> plt

if (save_plot)
  ggsave(paste0("plots/ct_median_ca_exp_decay_", tstamp(), ".pdf"), plt,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
```


## Figure 4 Outbreak size, duration, and attack rate groups

```{r}
# ensure that the svim has size_grp by running the relevant function in the 
# vacc_impact.qmd
lst <- vector("list", 10)
for(i in 1:length(lst)){
  lst[[i]] <- svim[size_grp >=i]
}

res_list <- vector('list', 10)
for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    size_grp = min(size_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm=T),
    pct_reduc_dur = median(pct_reduc_dur, na.rm=T)),
    by = .(vacc_week, vacc_cov)]
}
res <- rbindlist(res_list)

res |>
  filter(vacc_cov %in% VC, vacc_week %in% seq(1,19,2)) |>
  ggplot(aes(x=vacc_week, y=size_grp))+
  metR::geom_contour_fill(aes(z=pct_reduc_case)) +
  metR::geom_contour2(aes(z=pct_reduc_case, label = stat(level)),
                      , linetype="dotted")+
  scale_fill_viridis_c("Case reduction", limits=c(0,100),
                       breaks=seq(0,100,50),
                       labels = scales::percent_format(scale=1))+
  labs(x="Vaccination week since the outbreak began", 
       y="Total cases")+
  scale_x_continuous(breaks=seq(1,19,2), labels=NULL) +
  scale_y_continuous(breaks=c(1,4,7,10),
                     labels=c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"])))+
  theme_light()+
  theme_tight()+
  theme(legend.position="top") -> plt4a

# case_triggered

lst <- vector("list", 10)
for(i in 1:length(lst)){
  lst[[i]] <- svim_ct[size_grp >=i]
}

res_list <- vector('list', 10)
for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    size_grp = min(size_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm=T),
    pct_reduc_dur = median(pct_reduc_dur, na.rm=T)),
    by = .(case_trigger, vacc_cov)]
}
res <- rbindlist(res_list)

cts <- unique(svim_ct$case_trigger)
cts_ids <- c(21,25,29,2,6,10,14,18)

res |>
  # filter(vacc_cov %in% c(0.75), case_trigger %in% cts[cts_ids]) |>
  filter(vacc_cov %in% c(0.75)) |>
  ggplot(aes(x=case_trigger, y=size_grp))+
  metR::geom_contour_fill(aes(z=pct_reduc_case)) +
  metR::geom_contour2(aes(z=pct_reduc_case, label = stat(level)),
                      , linetype="dotted")+
  scale_fill_viridis_c("Case reduction", limits=c(0,100),
                       breaks=seq(0,100,50),
                       labels=scales::percent_format(scale=1))+
  labs(x="Cumulative cases at vaccination", 
       y="Total cases")+
  scale_x_continuous(breaks=seq(200,1000,200), 
                     labels=NULL) +
  scale_y_continuous(breaks=c(1,4,7,10), labels=NULL)+
  theme_light()+
  theme_tight()+
  theme(legend.position="top") -> plt4b

fac <- 1.4

# Outbreak duration groups

lst <- vector("list", 10)

for(i in 1:length(lst)){
  lst[[i]] <- svim[dur_grp >=i]
}

res_list <- vector('list', 10)

for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    dur_grp = min(dur_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm=T),
    pct_reduc_dur = median(pct_reduc_dur, na.rm=T)),
    by = .(vacc_week, vacc_cov)]
}
res <- rbindlist(res_list)

res |>
  filter(vacc_cov %in% VC, vacc_week %in% c(1:19)) |>
  ggplot(aes(x=vacc_week, y=dur_grp))+
  metR::geom_contour_fill(aes(z=pct_reduc_case)) +
  metR::geom_contour2(aes(z=pct_reduc_case, label = stat(level)),
                      , linetype="dotted")+
  scale_fill_viridis_c("Case reduction", limits=c(0,100),
                       breaks=seq(0,100,50),
                       labels = scales::percent_format(scale=1))+
  labs(x="Vaccination week since the outbreak began", 
       y="Outbreak duration")+
  scale_x_continuous(breaks=seq(1,19,2), labels=NULL) +
  scale_y_continuous(breaks=c(1,4,7,10),
                     labels=c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"])))+
  theme_light()+
  theme_tight()+
  theme(legend.position="top") -> plt4c


# case trigger
lst <- vector("list", 10)
for(i in 1:length(lst)){
  lst[[i]] <- svim_ct[dur_grp >=i]
}

res_list <- vector('list', 10)

for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    dur_grp = min(dur_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm=T),
    pct_reduc_dur = median(pct_reduc_dur, na.rm=T)),
    by = .(case_trigger, vacc_cov)]
}
res <- rbindlist(res_list)
tgs <- unique(svim_ct$case_trigger)

res |>
  filter(vacc_cov %in% VC) |>
  ggplot(aes(x=case_trigger, y=dur_grp))+
  metR::geom_contour_fill(aes(z=pct_reduc_case)) +
  metR::geom_contour2(aes(z=pct_reduc_case, label = stat(level)),
                      , linetype="dotted")+
  scale_fill_viridis_c("Case reduction", limits=c(0,100),
                       breaks=seq(0,100,50),
                       labels = scales::percent_format(scale=1))+
  labs(x="Cumulative cases at vaccination", 
       y="Outbreak duration")+
  scale_x_continuous(breaks=seq(200,1000,200), 
                     labels=NULL) +
  scale_y_continuous(breaks=c(1,4,7,10), labels=NULL)+
  theme_light()+
  theme_tight()+
  theme(legend.position="top") -> plt4d

# Outbreak attack rate groups

lst <- vector("list", 10)
for (i in 1:length(lst)) {
  lst[[i]] <- svim[ar_grp >=i]
}
res_list <- vector('list', 10)

for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    ar_grp = min(ar_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm=T),
    pct_reduc_dur = median(pct_reduc_dur, na.rm=T)),
    by = .(vacc_week, vacc_cov)]
}
res <- rbindlist(res_list)

res |>
  filter(vacc_cov %in% VC, vacc_week %in% c(1:19)) |>
  ggplot(aes(vacc_week, ar_grp))+
  metR::geom_contour_fill(aes(z=pct_reduc_case)) +
  metR::geom_contour2(aes(z=pct_reduc_case, label = stat(level)),
                      , linetype="dotted")+
  scale_fill_viridis_c("Case reduction", , limits=c(0,100),
                       breaks=seq(0,100,50),
                       labels = scales::percent_format(scale=1))+
  labs(x="Vaccination week since the outbreak began", 
       y="Outbreak attack rate")+
  scale_x_continuous(breaks=seq(1,19,2), labels=seq(1,19,2)) +
  scale_y_continuous(breaks=c(1,4,7,10),
                     labels=c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"])))+
  theme_light()+
  theme_tight()+
  theme(legend.position="top") -> plt4e


# case-triggered vaccination
lst <- vector("list", 10)
for(i in 1:length(lst)){
  lst[[i]] <- svim_ct[ar_grp >=i]
}

res_list <- vector('list', 10)

for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    ar_grp = min(ar_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm=T),
    pct_reduc_dur = median(pct_reduc_dur, na.rm=T)),
    by = .(case_trigger, vacc_cov)]
}

res <- rbindlist(res_list)
res |>
  filter(vacc_cov %in% VC) |>
  ggplot(aes(x=case_trigger, y=ar_grp))+
  metR::geom_contour_fill(aes(z=pct_reduc_case)) +
  metR::geom_contour2(aes(z=pct_reduc_case, label = stat(level)),
                      , linetype="dotted")+
  scale_fill_viridis_c("Case reduction", limits=c(0,100),
                       breaks=seq(0,100,50),
                       labels = scales::percent_format(scale=1))+
  labs(x="Cumulative cases at vaccination", 
       y="Outbreak attack rate")+
  scale_x_continuous(breaks=seq(200,1000,200), 
                     labels=seq(200,1000,200)) +
  scale_y_continuous(breaks=c(1,4,7,10), labels=NULL)+
  theme_light()+
  theme_tight()+
  theme(legend.position="top") -> plt4f

plt4 <- (plt4a + plt4b + plt4c + plt4d + plt4e + plt4f) +
  plot_annotation(tag_levels = 'A') + 
  plot_layout(byrow=T, ncol=2, axis_titles="collect", guides="collect") &
  theme(plot.tag.position  = c(.98, .96),
        plot.tag = element_text(size=13)) &
  theme(legend.position = "bottom")

fac <- 1.3

if (save_plot) {
  ggsave(paste0("plots/fig4_pcr_subgroups_contour_", 
                tstamp(hour=T,minute=T), ".pdf"), plt4,
         device = cairo_pdf, width=fgw2*fac, height=fgw2*fac, units="mm")
}
```

## Probability that the OCVs are delivered after the outbreak ended
```{r}
# just extract one set of outbreaks
svim_cea |> filter(vacc_cov == VC, vacc_week == VW) -> d
dur <- 9 # arbitrary delay to test
sum(d$duration_weeks > 9) / nrow(d)
D <- round(delay_req_r1[!is.na(delay_req_r1)]/7) # delay in weeks
X <- rep(NA, length(D))
for (i in 1:length(D)) {
  X[i] <- sum(d$duration_weeks > D[i]) /nrow(d)
}
# this assumes that OCVs are requested for all outbreaks regardless of duration and size
sum(X)/length(X)

svim_cea |> 
  filter(vacc_cov == VC & vacc_week == VW & 
         duration_weeks >= round((26+ocv_delay[[4]])/7) & 
         outbreak_pop >= min_pop_camp) -> dat

prob_outbreaks_longer <- function(outbreaks, delays){
  # prob for outbreaks longer than the given delay 
  X <- rep(NA, length(delays))
  for (i in 1:length(delays)) {
    X[i] <- sum(outbreaks$duration_weeks > delays[i]) / nrow(outbreaks)
  }
  return(X)
}

svim_cea |> 
  filter(vacc_cov == VC & vacc_week == VW & 
         duration_weeks >= round((26+ocv_delay[[4]])/7) & 
         outbreak_pop >= min_pop_camp) -> dat

pr <- prob_outbreaks_longer(dat, D)
paste0("At given delays, ", format_num(sum(pr)/length(pr)*100, 1), "% of the outbreaks in this data set last longer the time required for OCVs to be delivered to the site.")
```


## Table 3. Cases/deaths/DALYs averted, costs, and cost-effectiveness analysis 

```{r, eval=F}
# In emergency situations, the median time from the event (i.e.
# first laboratory confirmation of cholera or occurrence of humanitarian emergency) to the receipt of the official OCV request was 26 days (range 12–206 days). 

# 26 days
# request to round 1 Median (IQR)	45 (25, 66)
# delay from the start to round 1
# delay_median <- (26+45)/7 
# delay_lower <- (12+45)/7 
# delay_upper <- (206+45)/7
tab <- data.frame()
wks <- c(1, 5, 9, 13, 17)

d <- svim_cea |> filter(vacc_cov == VC, vacc_week == VW)
# tab <- create_impact_table(d)
for (wk in wks) {
  d <- svim_cea[(vacc_cov == VC & vacc_week == wk)]
  tab <- rbind(tab, impact_summary(d))
}  
tab$week <- wks

if (save_table) {
  fwrite(tab, paste0("outputs/svim_time_trigger_tab_", tstamp(), ".csv"))
}

tab_ct <- data.frame()
ctgs <- c(5,  40,  100, 300, 500, 1000)
# tab <- create_impact_table(d)
for (ctg in ctgs) {
  d <- svim_ct_cea[(vacc_cov == VC & case_trigger == ctg)]
  tab_ct <- rbind(tab_ct, impact_summary(d))
}  
tab_ct$case_trigger <- ctgs
if (save_table) {
  fwrite(tab_ct, paste0("outputs/svim_case_trigger_tab_", tstamp(), ".csv"))
}
```

## Table S1. Cases/deaths/DALYs averted, costs, and cost-effectiveness analysis 

```{r, eval=F}
tab <- data.frame()
wks <- c(1, 5, 9, 13, 17)
vcs <- c(0.6, 0.75, 0.9)
vcw <- expand.grid(vacc_week = wks, vacc_cov = vcs)
for (i in 1:nrow(vcw)) {
  d <- svim_cea[(vacc_cov == vcw$vacc_cov[i] & vacc_week == vcw$vacc_week[i])]
  tab <- rbind(tab, impact_summary(d))
}
tab <- cbind(vcw, tab)

# tat$vacc_cov
# tab$week <- wks

if (save_table) {
  fwrite(tab, paste0("outputs/svim_time_trigger_tab_full_", tstamp(), ".csv"))
}

tab_ct <- data.frame()
ctgs <- c(5,  40,  100, 300, 500, 1000)
vct <- expand.grid(case_trigger = ctgs, vacc_cov = vcs)
# tab <- create_impact_table(d)
for (i in 1:nrow(vct)) {
  d <- svim_ct_cea[(vacc_cov == vct$vacc_cov[i] & case_trigger == vct$case_trigger[i])]
  tab_ct <- rbind(tab_ct, impact_summary(d))
}  

tab_ct <- cbind(vct, tab_ct)

if (save_table) {
  fwrite(tab_ct, paste0("outputs/svim_case_trigger_tab_full_", tstamp(), ".csv"))
}
```


## Confirmed outbreaks

```{r}
length(unique(svim_cea$id_outbreak))
svim_cea |> 
  group_by(id_outbreak) |> 
  filter(c_ch_tot > 0) |> 
  ungroup() -> d

d <- filter(d, vacc_cov == VC, vacc_week == VW)
length(unique(d$id_outbreak))

tab <- create_impact_table(d)

if (save_table) {
  fwrite(tab, paste0("outputs/svim_confirmed_tab_", tstamp(), ".csv"))
}
```

## Figure 4. Overall percentage case reduction and case and deaths averted 
by year across campaign timing

```{r}
svim_preemptive$vacc_dose <- svim_preemptive$vacc_cov * svim_preemptive$population 
svim_preemptive$death_tot_averted <- svim_preemptive$death_tot * svim_preemptive$pct_reduc_case / 100

svim_preemptive |> 
  filter(vacc_cov %in% VC3) |> 
  group_by(vacc_cov) |> 
  summarize(sum_case = sum(s_ch_tot),
            sum_death = sum(death_tot, na.rm=T),
            sum_vacc = sum(vacc_dose, na.rm=T),
            sum_CA = sum(s_ch_averted_tot),
            sum_DA = sum(death_tot_averted, na.rm=T),
            sum_remaining_case = sum_case - sum_CA,
            sum_remaining_death = sum_death - sum_DA,
            pct_reduc_case = sum_CA / sum_case,
            pct_reduc_death = sum_DA / sum_death,
            case_averted_per_1000_OCVs = 1000 * sum_CA / sum_vacc,
            death_verted_per_1000_OCVs  = 1000 * sum_DA / sum_vacc) -> cada_pe

svim_cea |>
  filter(vacc_cov %in% VC3) |> 
  group_by(vacc_cov, vacc_week) |>
  summarize(sum_case = sum(s_ch_tot),
            sum_death = sum(death_tot, na.rm=T),
            sum_vacc = sum(vacc_dose),
            sum_CA = sum(s_ch_averted_tot),
            sum_DA = sum(death_tot_averted, na.rm=T),
            sum_remaining_case = sum_case - sum_CA,
            sum_remaining_death = sum_death - sum_DA,
            pct_reduc_case = sum_CA / sum_case,
            pct_reduc_death = sum_DA / sum_death,
            case_averted_per_1000_OCVs = 1000 * sum_CA / sum_vacc,
            death_verted_per_1000_OCVs = 1000 * sum_DA / sum_vacc) -> cada

cada
cada_by_year_wide <- 
  pivot_wider(ca_da_by_year,  names_from = vacc_week, 
              names_sep = ".", 
              values_from = -c(year, vacc_week, vacc_cov, sum_case, sum_death))

ggplot(cada)+
  geom_point(aes(vacc_week, pct_reduc_case, color=as.factor(vacc_cov), 
                 group=as.factor(vacc_cov)))+
  geom_line(aes(vacc_week, pct_reduc_case, color=as.factor(vacc_cov), 
                 group=as.factor(vacc_cov)))+
  # geom_hline(data=cada_pe, aes(yintercept=pct_reduc_case, 
  #                color=as.factor(vacc_cov)), linetype="dashed")+
  
  annotate("rect", xmin=6, xmax=14, ymin=0, ymax = 1, 
           alpha=0.1, fill="firebrick")+
  scale_y_continuous(labels = scales::percent_format(scale=100))+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:3],
                     labels=c(paste0(format_num(VC3*100,1),"%")))+
  labs(x="Vaccination week since the outbreak start", 
       y="Case reduction")+ 
  theme_light()+
  theme_tight() -> plt1

ggplot(svim_all_outbreaks)+
  geom_point(aes(vacc_week, q500_PCR/100, color=as.factor(vacc_cov), 
                 group=as.factor(vacc_cov)), position=position_dodge(width=0.5))+  
  geom_errorbar(aes(vacc_week, ymin=q025_PCR/100, ymax=q975_PCR/100,
                    color=as.factor(vacc_cov), group=as.factor(vacc_cov)),
                width=0, position=position_dodge(width=0.5))+
  geom_line(aes(vacc_week, q500_PCR/100, color=as.factor(vacc_cov), 
                 group=as.factor(vacc_cov)), position=position_dodge(width=0.5))+
  geom_hline(data=svim_pe_summary_across_runid, aes(yintercept=q975_PCR/100, 
                 color=as.factor(vacc_cov)), linetype="dashed")+
  annotate("rect", xmin=6, xmax=14, ymin=0, ymax = 1, 
           alpha=0.1, fill="firebrick")+
  scale_y_continuous(labels = scales::percent_format(scale=100))+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:3],
                     labels=c(paste0(format_num(VC3*100,1),"%")))+
  labs(x="Vaccination week since the outbreak start", 
       y="Case reduction")+ 
  theme_light()+
  theme_tight() -> plt1
plt1

fac <- 1.4
if (save_plot) {
  ggsave(paste0("plots/fig4_overall_pcr_", tstamp(), ".pdf"), 
         plt1,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}

res <- cada_75 <- filter(cada, vacc_cov == VC)

f <- function(t, Asym=0, lrc, y0) {
    Asym + (y0 - Asym) * exp(-exp(lrc) * t)
}
# fit <- nls(median_ca ~ SSasymp(t, yf, y0, log_alpha), data=res2)
fit <- nls(pct_reduc_case* ~ f(t=vacc_week, Asym=0, lrc, y0),  
           data=res, start=list(lrc=-1, y0=0.86))
fit
# weekly reduction rate
a <- exp(as.numeric(coef(fit)["lrc"]))
# half-life (weeks)
b <- log(2)/abs(exp(as.numeric(coef(fit)["lrc"])))

paste0("Percentage case reduction (%) decreases with delay in vaccination with ", format_num(a*100, digits=1) , "% by week of delay in campaign vaccination, and decreasing by half every ", format_num(b, digits=1), " weeks.")

svim_ct_cea |>
  filter(vacc_cov %in% VC3) |> 
  group_by(vacc_cov, case_trigger) |>
  summarize(sum_case = sum(s_ch_tot),
            sum_death = sum(death_tot, na.rm=T),
            sum_vacc = sum(vacc_dose),
            sum_CA = sum(s_ch_averted_tot),
            sum_DA = sum(death_tot_averted, na.rm=T),
            sum_remaining_case = sum_case - sum_CA,
            sum_remaining_death = sum_death - sum_DA,
            pct_reduc_case = sum_CA / sum_case,
            pct_reduc_death = sum_DA / sum_death,
            case_averted_per_1000_OCVs = 1000 * sum_CA / sum_vacc,
            death_verted_per_1000_OCVs = 1000 * sum_DA / sum_vacc) -> cada_ct

cts <- unique(cada_ct$case_trigger)
 
cada_ct |>
  filter(vacc_cov %in% VC3, case_trigger %in% cts[seq(1,19,2)]) |>
  ggplot()+
  geom_point(aes(x=case_trigger, y=pct_reduc_case, group=vacc_cov, 
             color=as.factor(vacc_cov)))+
  geom_line(aes(x=case_trigger, y=pct_reduc_case, group=vacc_cov,
             color=as.factor(vacc_cov)))+
  geom_hline(data=cada_pe, aes(yintercept=pct_reduc_case, 
                 color=as.factor(vacc_cov)), linetype="dashed")+
  scale_y_continuous(labels = scales::percent_format(scale=100))+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:3],
                     labels=c("60%","75%","90%"))+ 
  labs(x="Cumulative cases at vaccination", 
       y="Case reduction")+ 
  theme_light()+
  theme_tight() -> plt2

plt2

fac <- 1.4
if (save_plot) { 
  ggsave(paste0("plots/fig4_ct_overall_pcr_", tstamp(), 
                ".pdf"), plt2,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}

res <- filter(cada_ct, vacc_cov == VC)

f <- function(t, Asym=0, lrc, y0) {
    Asym + (y0 - Asym) * exp(-exp(lrc) * t)
}
# fit <- nls(median_ca ~ SSasymp(t, yf, y0, log_alpha), data=res2)
fit <- nls(pct_reduc_case ~ f(t=case_trigger, Asym=0, lrc, y0),  
           data=res[], start=list(lrc=-1, y0=0.86))
fit
# reduction rate by one additional case
a <- exp(as.numeric(coef(fit)["lrc"]))
# half-life (in terms of cumulative cases)
b <- log(2)/abs(exp(as.numeric(coef(fit)["lrc"])))

paste0("Percentage case reduction (%) decreases with delay in vaccination with ", format_num(a*100, digits=3) , "% by one additional case in campaign vaccination, and decreasing by half every ", format_num(b, digits=1), " cases")

# case averted by year
delay_outbk_request <- 26
delay_request_r1 <- quantile(delay_req_r1, probs=probs, na.rm=TRUE)

med_delay <- round((delay_outbk_request + delay_request_r1[3])/7)
ub_delay <- round((delay_outbk_request + delay_request_r1[4])/7)
lb_delay <- round((delay_outbk_request + delay_request_r1[2])/7)

svim_cea |>
  filter(vacc_week %in% c(med_delay, lb_delay, ub_delay)) |> 
  group_by(year, vacc_cov, vacc_week) |>
  summarize(sum_case = sum(s_ch_tot),
            sum_death = sum(death_tot, na.rm=T),
            sum_vacc = sum(vacc_dose),
            sum_CA = sum(s_ch_averted_tot),
            sum_DA = sum(death_tot_averted, na.rm=T),
            sum_remaining_case = sum_case - sum_CA,
            sum_remaining_death = sum_death - sum_DA,
            pct_reduc_case = sum_CA / sum_case,
            pct_reduc_death = sum_DA / sum_death,
            case_averted_per_1000_OCVs = 1000 * sum_CA / sum_vacc,
            death_verted_per_1000_OCVs  = 1000 * sum_DA / sum_vacc) -> cada_by_year

svim_gtfcc |>
  filter(vacc_week %in% c(med_delay, lb_delay, ub_delay)) |> 
  group_by(year, vacc_cov, vacc_week) |>
  summarize(sum_case = sum(s_ch_tot),
            sum_death = sum(death_tot, na.rm=T),
            sum_vacc = sum(vacc_dose),
            sum_CA = sum(s_ch_averted_tot),
            sum_DA = sum(death_tot_averted, na.rm=T),
            sum_remaining_case = sum_case - sum_CA,
            sum_remaining_death = sum_death - sum_DA,
            pct_reduc_case = sum_CA / sum_case,
            pct_reduc_death = sum_DA / sum_death,
            case_averted_per_1000_OCVs = 1000 * sum_CA / sum_vacc,
            death_verted_per_1000_OCVs = 1000 * sum_DA / sum_vacc) -> gtfcc_cada_by_year


cada_by_year_wide <- 
  pivot_wider(cada_by_year,  names_from = vacc_week, 
              names_sep = ".", 
              values_from = -c(year, vacc_week, vacc_cov, sum_case, sum_death))

gftcc_cada_by_year_wide <-
  pivot_wider(gtfcc_cada_by_year,
              names_from = vacc_week,
              names_sep = ".",
              values_from = -c(year, vacc_week, vacc_cov, sum_case, sum_death))

df <- filter(cada_by_year_wide, vacc_cov == VC, year <= 2022)
df_gtfcc <- filter(gftcc_cada_by_year_wide, vacc_cov == VC, year <= 2022)

ggplot(data=df)+
  geom_point(aes(year, sum_case, color="No vaccination"))+
  geom_line(aes(year, sum_case, color="No vaccination"))+
  geom_point(aes(year, sum_remaining_case.10, color="Vaccination"))+
  geom_line(aes(year, sum_remaining_case.10, color="Vaccination"))+
  geom_errorbar(aes(year, ymin=sum_remaining_case.8, 
                    ymax=sum_remaining_case.14), width=0, color="firebrick")+
  geom_ribbon(aes(x=year, ymin=pmax(sum_remaining_case.10,0),
                  ymax=sum_case), fill="firebrick", 
              col="firebrick", alpha=0.5)+
  scale_color_manual("", values=c("black","firebrick"))+
  scale_x_continuous(breaks=seq(2010,2022,by=2),
                     labels = seq(2010,2022,by=2))+
  labs(x="Year", 
       y="Cases from outbreaks by year")+ 
  theme_light()+
  theme_tight() -> plt_ca

plt_ca
fac <- 1.4
if (save_plot) {
  ggsave(paste0("plots/fig4_cases_averted_year_", tstamp(), ".pdf"), 
         plt_ca,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}

# deaths
ggplot(data=df)+
  geom_point(aes(year, sum_death, color="No vaccination"))+
  geom_line(aes(year, sum_death, color="No vaccination"))+
  geom_point(aes(x=year, y=sum_remaining_death.10, color="Vaccination"))+
  geom_line(aes(x=year, y=sum_remaining_death.10, color="Vaccination"))+
  geom_errorbar(aes(x=year, ymin=sum_remaining_death.8, 
                    ymax=sum_remaining_death.14), width=0, color="steelblue")+
  geom_ribbon(aes(x=year, ymin=pmax(sum_remaining_death.10,0),
                  ymax=sum_death), fill="steelblue", 
              col="steelblue", alpha=0.5)+
  scale_color_manual("", values=c("black","steelblue"))+
  # scale_color_manual("", values = c("No vaccination" = "black",
  #                                   "Vaccination" = "steelblue"),
  #                    breaks = c("Vaccination"))+
  scale_x_continuous(breaks=seq(2010,2022,by=2),
                     labels = seq(2010,2022,by=2))+
  labs(x="Year", 
       y="Deaths from outbreaks by year")+ 
  # scale_color_manual("", values=my_discrete_colors[1:13])+
  
  theme_light()+
  theme_tight() -> plt_da

plt_da
if (save_plot) {
  ggsave(paste0("plots/fig4_deaths_averted_year_", tstamp(), ".pdf"), 
         plt_da,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}

plt_ca_da <- plt_ca + plt_da + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(axis_titles = "collect_x") + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')

fac <- 1.8
if (save_plot) {
  ggsave(paste0("plots/fig4ab_cases_deaths_averted_af_year_",
                tstamp(), ".pdf"), plt_ca_da,
         device=cairo_pdf, width=fgw2*fac/1.2, height=fgw1*fac, units="mm")
}


paste0("Overall case reduction ", round(sum(df$sum_CA.10)), "[IQR: ", round(sum(df$sum_CA.14)), " - ",  round(sum(df$sum_CA.8)),"]", " of ", round(sum(df$sum_case)))

paste0("Overall percentage case reduction ", sum(df$sum_CA.10)/sum(df$sum_case)
, "[IQR: ", sum(df$sum_CA.14)/sum(df$sum_case), " - ",  sum(df$sum_CA.8)/sum(df$sum_case),"]")

paste0("Overall percentage death reduction ", sum(df$sum_DA.10)/sum(df$sum_death)
, "[IQR: ", sum(df$sum_DA.14)/sum(df$sum_death), " - ",  sum(df$sum_DA.8)/sum(df$sum_death),"]")

paste0("Overall case averted per 1,000 OCVs are ",  format_num(1000*sum(df$sum_CA.8)/sum(df$sum_vacc.8), digits=3), ", ",   format_num(1000*sum(df$sum_CA.10)/sum(df$sum_vacc.10), digits=3), ", and ",  format_num(1000*sum(df$sum_CA.14)/sum(df$sum_vacc.14), digits=3), " for 25th, 50th, and 75th percentiles of delays.")

pcr <- summary(df$pct_reduc_case.10[1:13])
paste0("Percentage case reduction vary from ", pcr[1], " to ", 
       pcr[6], " with a mean of ", pcr[4])

pdr <- summary(df$pct_reduc_death.10[1:13])
paste0("Percentage case reduction vary from ", pdr[1], " to ", 
       pdr[6], " with a mean of ", pdr[4])
df$pct_reduc_death.10[1:13]

paste0("Overall percentage case reduction ", sum(df_gtfcc$sum_CA.10)/sum(df_gtfcc$sum_case)
, "[IQR: ", sum(df_gtfcc$sum_CA.14)/sum(df_gtfcc$sum_case), " - ",  sum(df_gtfcc$sum_CA.8)/sum(df_gtfcc$sum_case),"]")

paste0("Overall percentage death reduction ", sum(df_gtfcc$sum_DA.10)/sum(df_gtfcc$sum_death)
, "[IQR: ", sum(df_gtfcc$sum_DA.14)/sum(df_gtfcc$sum_death), " - ",  sum(df_gtfcc$sum_DA.8)/sum(df_gtfcc$sum_death),"]")

pcr_gtfcc <- summary(df_gtfcc$pct_reduc_case.10[1:13])
paste0("For GTFCC subsets, percentage case reduction vary from ", 
       pcr_gtfcc[1], " to ", pcr_gtfcc[6], " with a mean of ", pcr_gtfcc[4])

pdr_gtfcc <- summary(df_gtfcc$pct_reduc_death.10[1:13])
paste0("Percentage case reduction vary from ", pdr_gtfcc[1], " to ", 
       pdr_gtfcc[6], " with a mean of ", pdr_gtfcc[4])
```

## Figure 4. Subsets overall percentage case reduction and case and deaths averted 
by year across campaign timing

```{r}
rand <- subset_random[vacc_cov == VC & vacc_week %in% seq(1,20,by=2)]
rand$subset_method <- "random"
ar <- subset_attack_rate[vacc_cov == VC & vacc_week %in% seq(1,20,by=2)]
ar$subset_method <- "attack rate"
cases <- subset_cases[vacc_cov == VC & vacc_week %in% seq(1,20,by=2)]
cases$subset_method <- "cases"
duration <- subset_duration[vacc_cov == VC & vacc_week %in% seq(1,20,by=2)]
duration$subset_method <- "duration"

subsets <- rbind(rand, ar, cases, duration)
  
svim_pe$vacc_dose <- svim_pe$vacc_cov * svim_pe$population 
svim_pe$death_tot_averted <- svim_pe$death_tot * svim_pe$pct_reduc_case / 100

svim_pe |> 
  filter(vacc_cov %in% VC3) |> 
  group_by(vacc_cov) |> 
  summarize(sum_case = sum(s_ch_tot),
            sum_death = sum(death_tot, na.rm=T),
            sum_vacc = sum(vacc_dose, na.rm=T),
            sum_CA = sum(s_ch_averted_tot),
            sum_DA = sum(death_tot_averted, na.rm=T),
            sum_remaining_case = sum_case - sum_CA,
            sum_remaining_death = sum_death - sum_DA,
            pct_reduc_case = sum_CA / sum_case,
            pct_reduc_death = sum_DA / sum_death,
            case_averted_per_1000_OCVs = 1000 * sum_CA / sum_vacc,
            death_verted_per_1000_OCVs  = 1000 * sum_DA / sum_vacc) -> cada_pe

subsets |>
  group_by(subset_method, vacc_cov, vacc_week) |>
  summarize(q500_cases = quantile(s_ch_tot, prob=0.5),
            q025_cases = quantile(s_ch_tot, prob=0.025),
            q975_cases = quantile(s_ch_tot, prob=0.975),
            q500_deaths = quantile(death_tot, prob=0.5),
            q025_deaths = quantile(death_tot, prob=0.025),
            q975_deaths = quantile(death_tot, prob=0.975),
            q500_CA = quantile(s_ch_averted_tot, prob=0.5),
            q025_CA = quantile(s_ch_averted_tot, prob=0.025),
            q975_CA = quantile(s_ch_averted_tot, prob=0.975),
            q500_DA = quantile(death_averted_tot, prob=0.5),
            q025_DA = quantile(death_averted_tot, prob=0.025),
            q975_DA = quantile(death_averted_tot, prob=0.975),
            q500_PCR = quantile(pct_reduc_case, prob=0.5),
            q025_PCR = quantile(pct_reduc_case, prob=0.025),
            q975_PCR = quantile(pct_reduc_case, prob=0.975),
            q500_PDR = quantile(pct_reduc_death, prob=0.5),
            q025_PDR = quantile(pct_reduc_death, prob=0.025),
            q975_PDR = quantile(pct_reduc_death, prob=0.975),
            q500_CAPD = quantile(case_averted_per_1000_OCV, prob=0.5),
            q025_CAPD = quantile(case_averted_per_1000_OCV, prob=0.025),
            q975_CAPD = quantile(case_averted_per_1000_OCV, prob=0.975),
            q500_DAPD = quantile(death_averted_per_1000_OCV, prob=0.5),
            q025_DAPD = quantile(death_averted_per_1000_OCV, prob=0.025),
            q975_DAPD = quantile(death_averted_per_1000_OCV, prob=0.975)) -> 
  subsets_q

wd <- 0.8   
# svim_subsets_sum_q |>
#   filter(vacc_cov == VC, vacc_week %in% seq(1,20,by=2)) -> dp

theme_4_panels <- function() { 
  theme(plot.margin = unit(c(0,0.3,0.15,0), "cm"))
}

ggplot(subsets_q)+
  geom_point(aes(vacc_week, q500_PCR, color=subset_method), 
             position=position_dodge(width=wd))+
  geom_line(aes(vacc_week, q500_PCR, color=subset_method),
            position=position_dodge(width=wd))+
  geom_errorbar(aes(vacc_week, ymin=q025_PCR, ymax=q975_PCR,
                    color=subset_method), width=0,
             position=position_dodge(width=wd))+
  # geom_hline(data=cada_pe, aes(yintercept=pct_reduc_case, 
  #                color=as.factor(vacc_cov)), linetype="dashed")+
  
  annotate("rect", xmin=6, xmax=14, ymin=0, ymax=max(subsets_q$q975_PCR), 
           alpha=0.1, fill="firebrick")+
  scale_y_continuous(labels = scales::percent_format(scale=1))+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:4],
                     labels=c("Attack rate", "Cases", "Duration", "Random"))+
  labs(x="Vaccination week since the outbreak start", 
       y="Case reduction")+ 
  theme_light()+
  theme_tight()+
  theme_4_panels() -> plt1


ggplot(subsets_q)+
  geom_point(aes(vacc_week, q500_PDR, color=subset_method), 
             position=position_dodge(width=wd))+
  geom_line(aes(vacc_week, q500_PDR, color=subset_method),
            position=position_dodge(width=wd))+
  geom_errorbar(aes(vacc_week, ymin=q025_PDR, ymax=q975_PDR,
                    color=subset_method), width=0,
             position=position_dodge(width=wd))+
  # geom_hline(data=cada_pe, aes(yintercept=pct_reduc_case, 
  #                color=as.factor(vacc_cov)), linetype="dashed")+
  
  annotate("rect", xmin=6, xmax=14, ymin=0, ymax=max(subsets_q$q975_PDR), 
           alpha=0.1, fill="firebrick")+
  scale_y_continuous(labels = scales::percent_format(scale=1))+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:4],
                     labels=c("Attack rate", "Cases", "Duration", "Random"))+
  labs(x="Vaccination week since the outbreak start", 
       y="Death reduction")+ 
  theme_light()+
  theme_tight() -> plt2

# plt <- plt1 + plt2 + 
#   plot_layout(guides = "collect", axis_titles = "collect") & theme(legend.position="top") 
  

ggplot(subsets_q)+
  geom_point(aes(vacc_week, q500_CAPD, color=subset_method), 
             position=position_dodge(width=wd))+
  geom_line(aes(vacc_week, q500_CAPD, color=subset_method),
            position=position_dodge(width=wd))+
  geom_errorbar(aes(vacc_week, ymin=q025_CAPD, ymax=q975_CAPD,
                    color=subset_method), width=0,
             position=position_dodge(width=wd))+
  # geom_hline(data=cada_pe, aes(yintercept=pct_reduc_case, 
  #                color=as.factor(vacc_cov)), linetype="dashed")+
  annotate("rect", xmin=6, xmax=14, ymin=0, 
           ymax=max(filter(subsets_q, vacc_cov == VC)$q975_CAPD),
           alpha=0.1, fill="firebrick")+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:4],
                     labels=c("Attack rate", "Cases", "Duration", "Random"))+
  labs(x="Vaccination week since the outbreak start", 
       y="Case averted per 1,000 OCVs")+ 
  theme_light()+
  theme_tight() -> plt3

ggplot(subsets_q)+
  geom_point(aes(vacc_week, q500_DAPD, color=subset_method), 
             position=position_dodge(width=wd))+
  geom_line(aes(vacc_week, q500_DAPD, color=subset_method),
            position=position_dodge(width=wd))+
  geom_errorbar(aes(vacc_week, ymin=q025_DAPD, ymax=q975_DAPD,
                    color=subset_method), width=0,
             position=position_dodge(width=wd))+
  # geom_hline(data=cada_pe, aes(yintercept=pct_reduc_case, 
  #                color=as.factor(vacc_cov)), linetype="dashed")+
  annotate("rect", xmin=6, xmax=14, ymin=0, 
           ymax=max(filter(subsets_q, vacc_cov == VC)$q975_DAPD),
           alpha=0.1, fill="firebrick")+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:4],
                     labels=c("Attack rate", "Cases", "Duration", "Random"))+
  labs(x="Vaccination week since the outbreak start", 
       y="Death averted per 1,000 OCVs")+ 
  theme_light()+
  theme_tight() -> plt4

plt <- plt1 + plt2 + plt3 + plt4 + 
  # plot_annotation(tag_levels = "A") + 
  plot_layout(ncol=2, byrow=T, guides="collect", axis_titles="collect") &
  theme(legend.position="top")

fac <- 2
if (save_plot) {
  ggsave(paste0("plots/fig4_subsets_pcdr_cada_", tstamp(), ".pdf"), 
         plt,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}

res <- cada_75 <- filter(cada, vacc_cov == VC)

f <- function(t, Asym=0, lrc, y0) {
    Asym + (y0 - Asym) * exp(-exp(lrc) * t)
}
# fit <- nls(median_ca ~ SSasymp(t, yf, y0, log_alpha), data=res2)
fit <- nls(pct_reduc_case* ~ f(t=vacc_week, Asym=0, lrc, y0),  
           data=res, start=list(lrc=-1, y0=0.86))
fit
# weekly reduction rate
a <- exp(as.numeric(coef(fit)["lrc"]))
# half-life (weeks)
b <- log(2)/abs(exp(as.numeric(coef(fit)["lrc"])))

paste0("Percentage case reduction (%) decreases with delay in vaccination with ", format_num(a*100, digits=1) , "% by week of delay in campaign vaccination, and decreasing by half every ", format_num(b, digits=1), " weeks.")

svim_ct_cea |>
  filter(vacc_cov %in% VC3) |> 
  group_by(vacc_cov, case_trigger) |>
  summarize(sum_case = sum(s_ch_tot),
            sum_death = sum(death_tot, na.rm=T),
            sum_vacc = sum(vacc_dose),
            sum_CA = sum(s_ch_averted_tot),
            sum_DA = sum(death_tot_averted, na.rm=T),
            sum_remaining_case = sum_case - sum_CA,
            sum_remaining_death = sum_death - sum_DA,
            pct_reduc_case = sum_CA / sum_case,
            pct_reduc_death = sum_DA / sum_death,
            case_averted_per_1000_OCVs = 1000 * sum_CA / sum_vacc,
            death_verted_per_1000_OCVs = 1000 * sum_DA / sum_vacc) -> cada_ct

cts <- unique(cada_ct$case_trigger)
 
cada_ct |>
  filter(vacc_cov %in% VC3, case_trigger %in% cts[seq(1,19,2)]) |>
  ggplot()+
  geom_point(aes(x=case_trigger, y=pct_reduc_case, group=vacc_cov, 
             color=as.factor(vacc_cov)))+
  geom_line(aes(x=case_trigger, y=pct_reduc_case, group=vacc_cov,
             color=as.factor(vacc_cov)))+
  geom_hline(data=cada_pe, aes(yintercept=pct_reduc_case, 
                 color=as.factor(vacc_cov)), linetype="dashed")+
  scale_y_continuous(labels = scales::percent_format(scale=100))+
  scale_color_manual(name="",
                     values=my_discrete_colors[1:3],
                     labels=c("60%","75%","90%"))+ 
  labs(x="Cumulative cases at vaccination", 
       y="Case reduction")+ 
  theme_light()+
  theme_tight() -> plt2

plt2

fac <- 1.4
if (save_plot) { 
  ggsave(paste0("plots/fig4_ct_overall_pcr_", tstamp(), 
                ".pdf"), plt2,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}

res <- filter(cada_ct, vacc_cov == VC)

f <- function(t, Asym=0, lrc, y0) {
    Asym + (y0 - Asym) * exp(-exp(lrc) * t)
}
# fit <- nls(median_ca ~ SSasymp(t, yf, y0, log_alpha), data=res2)
fit <- nls(pct_reduc_case ~ f(t=case_trigger, Asym=0, lrc, y0),  
           data=res[], start=list(lrc=-1, y0=0.86))
fit
# reduction rate by one additional case
a <- exp(as.numeric(coef(fit)["lrc"]))
# half-life (in terms of cumulative cases)
b <- log(2)/abs(exp(as.numeric(coef(fit)["lrc"])))

paste0("Percentage case reduction (%) decreases with delay in vaccination with ", format_num(a*100, digits=3) , "% by one additional case in campaign vaccination, and decreasing by half every ", format_num(b, digits=1), " cases")

# case averted by year
delay_outbk_request <- 26
delay_request_r1 <- quantile(delay_req_r1, probs=probs, na.rm=TRUE)

med_delay <- round((delay_outbk_request + delay_request_r1[3])/7)
ub_delay <- round((delay_outbk_request + delay_request_r1[4])/7)
lb_delay <- round((delay_outbk_request + delay_request_r1[2])/7)

svim_cea |>
  filter(vacc_week %in% c(med_delay, lb_delay, ub_delay)) |> 
  group_by(year, vacc_cov, vacc_week) |>
  summarize(sum_case = sum(s_ch_tot),
            sum_death = sum(death_tot, na.rm=T),
            sum_vacc = sum(vacc_dose),
            sum_CA = sum(s_ch_averted_tot),
            sum_DA = sum(death_tot_averted, na.rm=T),
            sum_remaining_case = sum_case - sum_CA,
            sum_remaining_death = sum_death - sum_DA,
            pct_reduc_case = sum_CA / sum_case,
            pct_reduc_death = sum_DA / sum_death,
            case_averted_per_1000_OCVs = 1000 * sum_CA / sum_vacc,
            death_verted_per_1000_OCVs  = 1000 * sum_DA / sum_vacc) -> cada_by_year

svim_gtfcc |>
  filter(vacc_week %in% c(med_delay, lb_delay, ub_delay)) |> 
  group_by(year, vacc_cov, vacc_week) |>
  summarize(sum_case = sum(s_ch_tot),
            sum_death = sum(death_tot, na.rm=T),
            sum_vacc = sum(vacc_dose),
            sum_CA = sum(s_ch_averted_tot),
            sum_DA = sum(death_tot_averted, na.rm=T),
            sum_remaining_case = sum_case - sum_CA,
            sum_remaining_death = sum_death - sum_DA,
            pct_reduc_case = sum_CA / sum_case,
            pct_reduc_death = sum_DA / sum_death,
            case_averted_per_1000_OCVs = 1000 * sum_CA / sum_vacc,
            death_verted_per_1000_OCVs = 1000 * sum_DA / sum_vacc) -> gtfcc_cada_by_year


cada_by_year_wide <- 
  pivot_wider(cada_by_year,  names_from = vacc_week, 
              names_sep = ".", 
              values_from = -c(year, vacc_week, vacc_cov, sum_case, sum_death))

gftcc_cada_by_year_wide <-
  pivot_wider(gtfcc_cada_by_year,
              names_from = vacc_week,
              names_sep = ".",
              values_from = -c(year, vacc_week, vacc_cov, sum_case, sum_death))

df <- filter(cada_by_year_wide, vacc_cov == VC, year <= 2022)
df_gtfcc <- filter(gftcc_cada_by_year_wide, vacc_cov == VC, year <= 2022)

ggplot(data=df)+
  geom_point(aes(year, sum_case, color="No vaccination"))+
  geom_line(aes(year, sum_case, color="No vaccination"))+
  geom_point(aes(year, sum_remaining_case.10, color="Vaccination"))+
  geom_line(aes(year, sum_remaining_case.10, color="Vaccination"))+
  geom_errorbar(aes(year, ymin=sum_remaining_case.8, 
                    ymax=sum_remaining_case.14), width=0, color="firebrick")+
  geom_ribbon(aes(x=year, ymin=pmax(sum_remaining_case.10,0),
                  ymax=sum_case), fill="firebrick", 
              col="firebrick", alpha=0.5)+
  scale_color_manual("", values=c("black","firebrick"))+
  scale_x_continuous(breaks=seq(2010,2022,by=2),
                     labels = seq(2010,2022,by=2))+
  labs(x="Year", 
       y="Cases from outbreaks by year")+ 
  theme_light()+
  theme_tight() -> plt_ca

plt_ca
fac <- 1.4
if (save_plot) {
  ggsave(paste0("plots/fig4_cases_averted_year_", tstamp(), ".pdf"), 
         plt_ca,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}

# deaths
ggplot(data=df)+
  geom_point(aes(year, sum_death, color="No vaccination"))+
  geom_line(aes(year, sum_death, color="No vaccination"))+
  geom_point(aes(x=year, y=sum_remaining_death.10, color="Vaccination"))+
  geom_line(aes(x=year, y=sum_remaining_death.10, color="Vaccination"))+
  geom_errorbar(aes(x=year, ymin=sum_remaining_death.8, 
                    ymax=sum_remaining_death.14), width=0, color="steelblue")+
  geom_ribbon(aes(x=year, ymin=pmax(sum_remaining_death.10,0),
                  ymax=sum_death), fill="steelblue", 
              col="steelblue", alpha=0.5)+
  scale_color_manual("", values=c("black","steelblue"))+
  # scale_color_manual("", values = c("No vaccination" = "black",
  #                                   "Vaccination" = "steelblue"),
  #                    breaks = c("Vaccination"))+
  scale_x_continuous(breaks=seq(2010,2022,by=2),
                     labels = seq(2010,2022,by=2))+
  labs(x="Year", 
       y="Deaths from outbreaks by year")+ 
  # scale_color_manual("", values=my_discrete_colors[1:13])+
  
  theme_light()+
  theme_tight() -> plt_da

plt_da
if (save_plot) {
  ggsave(paste0("plots/fig4_deaths_averted_year_", tstamp(), ".pdf"), 
         plt_da,
         device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
}

plt_ca_da <- plt_ca + plt_da + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(axis_titles = "collect_x") + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')

fac <- 1.8
if (save_plot) {
  ggsave(paste0("plots/fig4ab_cases_deaths_averted_af_year_",
                tstamp(), ".pdf"), plt_ca_da,
         device=cairo_pdf, width=fgw2*fac/1.2, height=fgw1*fac, units="mm")
}


paste0("Overall case reduction ", round(sum(df$sum_CA.10)), "[IQR: ", round(sum(df$sum_CA.14)), " - ",  round(sum(df$sum_CA.8)),"]", " of ", round(sum(df$sum_case)))

paste0("Overall percentage case reduction ", sum(df$sum_CA.10)/sum(df$sum_case)
, "[IQR: ", sum(df$sum_CA.14)/sum(df$sum_case), " - ",  sum(df$sum_CA.8)/sum(df$sum_case),"]")

paste0("Overall percentage death reduction ", sum(df$sum_DA.10)/sum(df$sum_death)
, "[IQR: ", sum(df$sum_DA.14)/sum(df$sum_death), " - ",  sum(df$sum_DA.8)/sum(df$sum_death),"]")

paste0("Overall case averted per 1,000 OCVs are ",  format_num(1000*sum(df$sum_CA.8)/sum(df$sum_vacc.8), digits=3), ", ",   format_num(1000*sum(df$sum_CA.10)/sum(df$sum_vacc.10), digits=3), ", and ",  format_num(1000*sum(df$sum_CA.14)/sum(df$sum_vacc.14), digits=3), " for 25th, 50th, and 75th percentiles of delays.")

pcr <- summary(df$pct_reduc_case.10[1:13])
paste0("Percentage case reduction vary from ", pcr[1], " to ", 
       pcr[6], " with a mean of ", pcr[4])

pdr <- summary(df$pct_reduc_death.10[1:13])
paste0("Percentage case reduction vary from ", pdr[1], " to ", 
       pdr[6], " with a mean of ", pdr[4])
df$pct_reduc_death.10[1:13]

paste0("Overall percentage case reduction ", sum(df_gtfcc$sum_CA.10)/sum(df_gtfcc$sum_case)
, "[IQR: ", sum(df_gtfcc$sum_CA.14)/sum(df_gtfcc$sum_case), " - ",  sum(df_gtfcc$sum_CA.8)/sum(df_gtfcc$sum_case),"]")

paste0("Overall percentage death reduction ", sum(df_gtfcc$sum_DA.10)/sum(df_gtfcc$sum_death)
, "[IQR: ", sum(df_gtfcc$sum_DA.14)/sum(df_gtfcc$sum_death), " - ",  sum(df_gtfcc$sum_DA.8)/sum(df_gtfcc$sum_death),"]")

pcr_gtfcc <- summary(df_gtfcc$pct_reduc_case.10[1:13])
paste0("For GTFCC subsets, percentage case reduction vary from ", 
       pcr_gtfcc[1], " to ", pcr_gtfcc[6], " with a mean of ", pcr_gtfcc[4])

pdr_gtfcc <- summary(df_gtfcc$pct_reduc_death.10[1:13])
paste0("Percentage case reduction vary from ", pdr_gtfcc[1], " to ", 
       pdr_gtfcc[6], " with a mean of ", pdr_gtfcc[4])

```


## Table 4 Subsets

```{r}
# In emergency situations, the median time from the event (i.e.
# first laboratory confirmation of cholera or occurrence of humanitarian emergency) to the receipt of the official OCV request was 26 days (range 12–206 days). 

# 26 days
# request to round 1 Median (IQR)	45 (25, 66)
# delay from the start to round 1
# delay_median <- (26+45)/7 
# delay_lower <- (12+45)/7 
# delay_upper <- (206+45)/7
svim_subsets_sum |>
  filter(vacc_cov == VC, vacc_week == VW, subset_method == "random") -> d1 
svim_subsets_sum |>
  filter(vacc_cov == VC, vacc_week == VW, subset_method == "attack_rate") -> d2 
svim_subsets_sum |>
  filter(vacc_cov == VC, vacc_week == VW, subset_method == "cases") -> d3 
svim_subsets_sum |>
  filter(vacc_cov == VC, vacc_week == VW, subset_method == "duration") -> d4 

tab1 <- create_impact_table2(d1)
tab2 <- create_impact_table2(d2)
tab3 <- create_impact_table2(d3)
tab4 <- create_impact_table2(d4)
tab <- rbind(tab1, tab2, tab3, tab4)
tab$subset_method <- c("random","attack_rate","cases","duration")

if (save_table) {
  fwrite(tab, paste0("outputs/svim_subset_tab_", tstamp(), ".csv"))
}

tab_all <- create_impact_table2(svim)
```


