---
title: "Plots for Gavi"
format: html
editor_options: 
  chunk_output_type: console
---

## Load packages

```{r}
# library(CholeraOutbreakModel)
# devtools::load_all() # works only for me who is writing a package
library(data.table)
library(tidyverse)
library(readxl)
source("R/functions.R")
set1 <- RColorBrewer::brewer.pal(9,"Set1")
# figure size
fgw1 <- figure_size[figure_size$journal =="Lancet",]$single
fgw2 <- figure_size[figure_size$journal =="Lancet",]$double
```

## Load outbreak datasets

```{r}
fits <- fread("outputs/data_fit_eval_20240409.csv")
good_fits <- fits[PARAM_GLOBAL == 1 & PARAM_SANITY_5pct == 1 & 
                   FIT_SANITY == 1 & R2_adj >= 0.5]
# extract ids for the outbreaks not affected by OCV campaigns
ds <- as.data.frame(fread("inst/extdata/outbreak_data_summary.csv"))
ds$ID_outbreak <- paste0(ds$location, "-", ds$start_date, "-", ds$end_date)
ds$ID <- 1:nrow(ds)
# data for the outbreaks affected by OCV
docv <- fread("inst/extdata/ocv_long_dataset.csv")
# exclude outbreaks that were affected by OCV campaigns using IDs 
dat <- ds[!(ds$ID_outbreak %in% docv$ID_outbreak),]
ts <- readRDS("inst/extdata/outbreak_data_ts.rds")
# see data_manipulation.Rmd to understand how these data are created
ts824_wk <- as.data.frame(fread("outputs/ts824_week.csv"))
ts999_wk <- as.data.frame(fread("outputs/ts999_week.csv"))
ts113_wk <- as.data.frame(fread("outputs/ts113_week.csv"))
```

## Figures -- Outbreak description

Outbreak time series by group (All, No OCV, and No OCV & Confirmed)
```{r, eval=FALSE}
# time series across dates
library(colorRamps)
colfunc <- colorRampPalette(c("white", "darkred"))
cols <- colfunc(9)
ggplot()+
  geom_line(data=ts999_wk, aes(date, sCh, group=data_id, color="All"), 
            alpha=0.5)+
  geom_line(data=ts824_wk, 
            aes(date, sCh, group=data_id, color="No OCV"),alpha=0.5)+
  geom_line(data=ts113_wk, 
            aes(date, sCh, group=data_id, color="No OCV & Confirmed"), 
            alpha=0.5)+
  scale_color_manual("", values=c("All" = "yellow4", "No OCV" = "steelblue4",
                                  "No OCV & Confirmed" = "firebrick4"))+
  labs(x="", y="Weekly suspected cases")+
  theme_light()+
  # theme(legend.position=c(0.1,0.8))+
  theme(legend.position="top") -> plt

ggsave(paste0("plots/outbk_ts_case_type_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw2, height=fgw1*1.1, units="mm")
```

Outbreak size vs. duration by type 
```{r, eval=FALSE}
ggplot()+
  geom_point(data=ds, aes(duration, total_suspected_cases,
             color="All"), alpha=0.5)+
  geom_point(data=dat, aes(duration, total_suspected_cases,
             color="No OCV"), alpha=0.5)+
  geom_point(data=subset(dat, !is.na(total_confirmed_cases), 
                         total_confirmed_cases > 0), 
             aes(duration, total_suspected_cases,
             color="No OCV & Confirmed"), alpha=0.5)+
  scale_x_log10()+
  scale_y_log10()+
  theme_light()+
  scale_color_manual("", values=c("All" = "yellow4", "No OCV" = "steelblue4",
                                  "No OCV & Confirmed" = "firebrick4"))+
  labs(x="Outbreak duration (week)", y="Total suspected cases")+
  theme(legend.position="top") -> plt

fac <- 1.25
ggsave(paste0("plots/outbk_size_dur_case_type_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
```

Functions

```{r}
# extract admin information from id
get_adm <- function(x) {
  # d <- grep("[a-zA-Z+]", strsplit(x, "::|-")[[1]], value=TRUE)
  d <- lapply(x, function(z) grep("[a-zA-Z+]", strsplit(z, "::|-")[[1]], value=TRUE))
  return (d)
}

# extract admin information from id
# this produces the correct outbreak start dates (when compared with the "start_date" column)
# but not used because the 
# ds file already has the start_date column
get_outbreak_start_date <- function(x) {
  # d <- grep("[^0-9+]", strsplit(x, "::|-")[[1]], value=TRUE)
  # ymd <- strsplit(x, "-")[[1]]
  ymd <- lapply(x, function(z) strsplit(z, "-"))
  dt <- lapply(ymd, function(z) as.Date(paste0(z[[1]][2], "-", z[[1]][3], "-", z[[1]][4])))
  return (dt)
}

# extract outbreak ids that may not occur because of the lasting immunity 
find_ghost_outbreaks <- function(x, lb=0, ub=1000, ...) {
  group_var <- enquos(...)
  x |> 
    # group_by(admin0, admin1, admin2) |>
    group_by(!!!group_var) |>
    reframe(id = id,
            data_id = data_id,
            date = new_date,
            min_date = min(date),
            date_diff = date - min_date) -> x1

  ghost_id <- x1[x1$date_diff > lb & x1$date_diff <= ub,]$data_id
  x1 <- x1[x1$date_diff > ub, ]
  
  while (!is.null(x1) & nrow(x1) > 1) {
    x1 |> 
      # group_by(admin0, admin1, admin2) |>
      group_by(!!!group_var) |>
      reframe(id = id,
            data_id = data_id,
            date = date,
            min_date = min(date),
            date_diff = date - min_date) -> x1
    ghost_id <- c(ghost_id, x1[x1$date_diff > lb & x1$date_diff <= ub,]$data_id)
    x1 <- x1[x1$date_diff > ub,]
  }
  return (ghost_id)
}
```

```{r}
 svim1 |>
    group_by(admin0, admin1, admin2) |>
    reframe(id = id,
            data_id = data_id,
            date = new_date,
            min_date = min(date),
            date_diff = date - min_date) -> x1

  ghost_id <- x1[x1$date_diff > lb & x1$date_diff <= ub,]$data_id
  x1 <- x1[x1$date_diff > ub, ]

  while (!is.null(x1) & nrow(x1) > 1) {
    x1 |>
      # group_by(admin0, admin1, admin2) |>
      group_by(!!!group_var) |>
      reframe(id = id,
              data_id = data_id,
              date = date,
              min_date = min(date),
              date_diff = date - min_date) -> x1
    ghost_id <- c(ghost_id, x1[x1$date_diff > lb & x1$date_diff <= ub,]$data_id)
    x1 <- x1[x1$date_diff > ub,]
  }
```


Vaccine impact simulation results
```{r}
# create grouping variables (size, duration) to analyze outbreaks accordingly
ds$size_grp <- ifelse(ds$total_suspected_cases < 200, "small", "large")
ds$dur_grp <- ifelse(ds$duration < 10, "short", "long")
ds_ <- ds[, c("threshold", "attack_rate", "cfr", "population","rural_urban", 
              "start_date","end_date", "total_confirmed_cases","ID_outbreak",
              "size_grp", "dur_grp")]
names(ds_) <- c("threshold", "attack_rate", "cfr", "population","rural_urban", 
              "start_date","end_date", "total_confirmed_cases","id",
              "size_grp", "dur_grp")

svim_list <- readRDS("outputs/svim_list_20240620.rds")

svim1 <- left_join(svim_list[[1]], ds_, by="id")
```

Add vaccine impact parameters
```{r}
parms <- as.data.frame(fread("data/parameters.csv"))
pr_asymp <- parms[parms$Parameter == "Prop_Asymptomatic", "Value"]
pr_mild <- parms[parms$Parameter == "Prop_Mild", "Value"]
pr_moderate <- parms[parms$Parameter == "Prop_Moderate", "Value"]
pr_severe <- parms[parms$Parameter == "Prop_Severe", "Value"]
dur <- parms[parms$Parameter == "Duration_Illness", "Value"]
wt_asymp <- parms[parms$Parameter == "Disability_Weight_Asymptomatic", "Value"]
wt_mild <- parms[parms$Parameter == "Disability_Weight_Mild", "Value"]
wt_moderate <- parms[parms$Parameter == "Disability_Weight_Moderate", "Value"]
wt_severe <- parms[parms$Parameter == "Disability_Weight_Severe", "Value"]

vacc_cost_per_dose <- parms[parms$Parameter == "Vaccine_Cost", "Value"]
vacc_delivery_cost <- parms[parms$Parameter == "Vaccine_Delivery_Cost", "Value"]
patient_cost_hosp <- parms[parms$Parameter == "Patient_Cost_Hosp", "Value"]
patient_cost_outpt <-parms[parms$Parameter == "Patient_Cost_Outpt", "Value"]
public_cost_hosp <- parms[parms$Parameter == "Public_Cost_Hosp", "Value"]
public_cost_outpt <-parms[parms$Parameter == "Public_Cost_Outpt", "Value"] 
patient_workday_lost <- parms[parms$Parameter == "Pt_Workdays_Lost", "Value"]
caregiver_workday_lost <- parms[parms$Parameter == "Caregiver_Workdays_Lost", "Value"]
mean_age_inf <- parms[parms$Parameter == "Mean_Age_Infection", "Value"]

dose_regimen <- 1
vacc_cost_per_person <- vacc_cost_per_dose * dose_regimen

svim1$year <- data.table::year(svim1$date)

# combine life expectancy, GDP, and percent workforce data
life_exp_data <- as.data.frame(fread("outputs/wpp2022_life_expectancy.csv"))
names(life_exp_data)[names(life_exp_data) == "ISO3 Alpha-code"] <- "country"
names(life_exp_data)[names(life_exp_data) == "Year"] <- "year"
avg_age_inf <- 26
svim1 <- left_join(svim1, 
                   life_exp_data[, c("country", "year", as.character(avg_age_inf))], 
                   by = c("country", "year"))


names(svim1)[names(svim1) == as.character(avg_age_inf)] <- "life_exp"
svim1$life_exp <- as.numeric(svim1$life_exp)

gdp <- read_xls("data/GDP_WorldBank.xls")
names(gdp)[names(gdp) == "Country Code"] <- "country"
gdp_long <- gdp %>%
  pivot_longer(
    cols = `1960`:`2023`, 
    names_to = "year",     
    values_to = "gdp"      
  )

gdp_long$year <- as.integer(gdp_long$year)

svim1 <- left_join(svim1, 
                  gdp_long[, c("country", "year", "gdp")],
                  by = c("country", "year"))

workforce <- read_xls("data/Workforce_Worldbank.xls")
names(workforce)[names(workforce) == "Country Code"] <- "country"
workforce_long <- workforce %>%
  pivot_longer(
    cols = `1960`:`2023`,
    names_to = "year",
    values_to = "pct_workforce"
  )
workforce_long$year <- as.integer(workforce_long$year)

svim1 <- left_join(svim1, 
                   workforce_long[, c("country", "year", "pct_workforce")],
                   by = c("country", "year"))

# svim1$population <- sapply(1:nrow(svim1), function(i) dat[dat$ID == svim1$data_id[i], "population"])
```

## Compute vaccine impact summary 

```{r}
# vaccine impact
# 100% for the svim1_sub and normal weekly impact for the rest
# cost-of-illness is based on 
coi_per_patient <- (pr_mild * patient_cost_outpt + 
      pr_moderate * patient_cost_outpt + pr_severe * patient_cost_hosp)

svim1 |>
  # filter(week_vaccination == 1, vacc_cov == 0.9) |> 
  mutate(pct_reduc_case = 100 * sCh_averted_tot / sCh_tot,
         yld = compute_yld(sCh_tot, parms=parms),
         yld_averted = compute_yld(sCh_averted_tot, parms=parms), 
         yll = compute_yll(death_tot, life_exp, parms),
         yll_averted = yll * (sCh_averted_tot / sCh_tot),
         daly_averted = yld_averted + yll_averted,
         coi_averted = sCh_averted_tot * coi_per_patient,
         cod_averted = death_tot * (sCh_averted_tot / sCh_tot) * gdp * life_exp,
         productivity_lost_averted = 
           sCh_averted_tot * gdp * ((patient_workday_lost/365)*(pct_workforce/100) +
                                 (caregiver_workday_lost/365)),
         vacc_dose = population * dose_regimen,
         vacc_cost = vacc_cost_per_dose * vacc_dose,
         net_cost = vacc_cost - coi_averted - cod_averted,
         cost_per_daly_averted = net_cost / daly_averted,
         case_averted_per_dose = sCh_averted_tot / vacc_dose,
         death_averted_per_dose = death_tot * (sCh_averted_tot / sCh_tot) / vacc_dose) -> svim1
```

Add weekly summary information (e.g., duration of the outbreak shorted via vaccination)

```{r}
svim_week_list <- readRDS("outputs/svim_week_list_20240627.rds")
svim1_week <- svim_week_list[[1]]
ds_ <- ds[, c("threshold", "attack_rate", "cfr", "population", "rural_urban", 
              "start_date","end_date", "total_confirmed_cases","ID_outbreak",
              "size_grp", "dur_grp")]
names(ds_) <- c("threshold", "attack_rate", "cfr", "population","rural_urban", 
              "start_date","end_date", "total_confirmed_cases","id",
              "size_grp", "dur_grp")
svim1_week <- left_join(svim1_week, ds_, by="id")

svim1_week <- svim1_week |>
  # filter(vacc_cov == 0.9, week_vaccination == 1) |> 
  # group_by(id) |>
  mutate(sCh_vacc = sCh_tot - sCh_averted_tot,
         outbk_yesno = ifelse(1e5 * sCh_tot / population > threshold, 1, 0),
         outbk_vacc_yesno = ifelse(1e5 * sCh_vacc / population > threshold, 1, 0))

svim1_week_summary <- svim1_week |> 
  group_by(id, week_vaccination, vacc_cov) |> 
  summarise(.groups = "drop",
            data_id = data_id[1],
            sCh_tot_sum = sum(sCh_tot),
            sCh_averted_tot_sum = sum(sCh_averted_tot),
            pct_reduc_case_wk = 100 * sCh_averted_tot_sum / sCh_tot_sum,
            threshold = threshold[1],
            outbk_dur = sum(outbk_yesno),
            outbk_dur_vacc = sum(outbk_vacc_yesno),
            pct_reduc_dur = 100 * (outbk_dur - outbk_dur_vacc) / outbk_dur)

svim1 <- 
  left_join(svim1, 
            svim1_week_summary[,c("id", "week_vaccination", 
                                  "vacc_cov",
                                  "pct_reduc_case_wk", 
                                  "threshold", 
                                  "outbk_dur",
                                  "outbk_dur_vacc",
                                  "pct_reduc_dur")],
            by = c("id", "week_vaccination", "vacc_cov"))
# 0 if computations are consistent across both ways
# sum(svim1$pct_reduc_case != svim1$pct_reduc_case_wk)
```

## Figures - Percent case reduction for individual outbreaks 

```{r}
vcov_interest <- c(0.9)
vweek_interest <- c(4)

svim1 |>
  filter(vacc_cov == vcov_interest, 
         week_vaccination == vweek_interest) -> svim1_sub

svim1_sub |>
  ggplot()+
  geom_histogram(aes(x=pct_reduc_case/100, after_stat(count/sum(count))),
                 binwidth = 0.02,
                 fill=set1[2])+ # /100 because of `scales::percent` function
  labs(x="Percent case reduction", y="Percent outbreaks", 
       subtitle = "90% vaccination at week 4")+ 
  scale_x_continuous(labels = scales::percent)+
  scale_y_continuous(labels = scales::percent)+
  theme_light() -> plt

fac <- 1.25
ggsave(paste0("plots/pct_case_reduc_hist_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
svim1 |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,3,5,7,9,11,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduc_case/100))+
  geom_boxplot(aes(color=as.factor(vacc_cov)))+
  scale_color_manual("Vaccine coverage", values=c(set1[3],set1[2]), 
                     labels=c("60%","90%"))+
  scale_y_continuous(labels = scales::percent)+
  theme_light()+
  theme(legend.position="top")+
  labs(x="Week of vaccination", y="Percent case reduction") -> plt

fac <- 1.25
ggsave(paste0("plots/pct_case_red_vc_60_90_camp_timing_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
  
svim1 |> 
  filter(vacc_cov %in% c(0.9), week_vaccination %in% c(1:13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduc_case))+
  geom_boxplot(color=set1[2])+
  theme_light()+
  theme(legend.position="top")+
  labs(x="Week of vaccination", y="Percent case reduction",
       title="90% vaccination coverage") -> plt

fac <- 1.25
ggsave(paste0("plots/pct_case_reduc_camp_timing_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")

svim1 |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduc_case))+
  geom_boxplot(aes(color=as.factor(size_grp)))+
  scale_color_manual("Outbreak size", values=c(set1[3],set1[4]),
                     labels=c(">=200 cases","<200 cases"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="Week of vaccination", y="Percent case reduction",
       title="90% vaccine coverage") -> plt 

fac <- 1.25
ggsave(paste0("plots/pct_case_reduc_outbrk_size_vc_60_90_camp_timing_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")

svim1 |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduc_case))+
  geom_boxplot(aes(color=as.factor(dur_grp)))+
  scale_color_manual("Outbreak duration", values=c(set1[3],set1[4]),
                     labels=c(">=10 weeks", "<10 weeks"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="Week of vaccination", y="Percent case reduction",
       title="90% vaccine coverage") -> plt

fac <- 1.25
ggsave(paste0("plots/pct_case_reduc_outbrk_dur_vc_60_90_camp_timing_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
```


## Figures - Other metrics (percent duration reduction, DALYs averted, 
ICER, for individual outbreaks 

```{r}
vcov_interest <- c(0.9)
vweek_interest <- c(4)

svim1 |>
  filter(vacc_cov == vcov_interest, 
         week_vaccination == vweek_interest) -> svim1_sub

svim1_sub |>
  filter(outbk_dur > 5) |> # this is just for now, 
  ggplot()+
  geom_histogram(aes(x=pct_reduc_dur/100, after_stat(count/sum(count))),
                 binwidth = 0.02,
                 fill=set1[2])+ # /100 because of `scales::percent` function
  labs(x="Percent duration reduction", y="Percent outbreaks", 
       subtitle = "90% vaccination at week 4")+ 
  scale_x_continuous(labels = scales::percent)+
  scale_y_continuous(labels = scales::percent)+
  theme_light() -> plt

fac <- 1.25
ggsave(paste0("plots/pct_case_reduc_hist_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
svim1 |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,3,5,7,9,11,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduc_case/100))+
  geom_boxplot(aes(color=as.factor(vacc_cov)))+
  scale_color_manual("Vaccine coverage", values=c(set1[3],set1[2]), 
                     labels=c("60%","90%"))+
  scale_y_continuous(labels = scales::percent)+
  theme_light()+
  theme(legend.position="top")+
  labs(x="Week of vaccination", y="Percent case reduction") -> plt

fac <- 1.25
ggsave(paste0("plots/pct_case_red_vc_60_90_camp_timing_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
  
svim1 |> 
  filter(vacc_cov %in% c(0.9), week_vaccination %in% c(1:13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduc_case))+
  geom_boxplot(color=set1[2])+
  theme_light()+
  theme(legend.position="top")+
  labs(x="Week of vaccination", y="Percent case reduction",
       title="90% vaccination coverage") -> plt

fac <- 1.25
ggsave(paste0("plots/pct_case_reduc_camp_timing_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")

svim1 |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduc_case))+
  geom_boxplot(aes(color=as.factor(size_grp)))+
  scale_color_manual("Outbreak size", values=c(set1[3],set1[4]),
                     labels=c(">=200 cases","<200 cases"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="Week of vaccination", y="Percent case reduction",
       title="90% vaccine coverage") -> plt 

fac <- 1.25
ggsave(paste0("plots/pct_case_reduc_outbrk_size_vc_60_90_camp_timing_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")

svim1 |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduc_case))+
  geom_boxplot(aes(color=as.factor(dur_grp)))+
  scale_color_manual("Outbreak duration", values=c(set1[3],set1[4]),
                     labels=c(">=10 weeks", "<10 weeks"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="Week of vaccination", y="Percent case reduction",
       title="90% vaccine coverage") -> plt

fac <- 1.25
ggsave(paste0("plots/pct_case_reduc_outbrk_dur_vc_60_90_camp_timing_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
```

### Vaccine impact summary across outbreaks
This gives one estimate for the dataset and we can explore the variation across
parameter values.

```{r}
svim1 |>
  # filter(!(data_id %in% ghost_ids)) |>
  summarize(
    sCh_tot_sum = sum(sCh_tot),
    sCh_averted_tot_sum = sum(sCh_averted_tot),
    prop_averted = sCh_averted_tot_sum / sCh_tot_sum,
    yld_tot = compute_yld(sCh_tot_sum, parms=parms),
    yld_averted_tot = compute_yld(sCh_averted_tot_sum, parms=parms),
    yll_tot = sum(compute_yll(death_tot, life_exp, parms), na.rm=T),
    yll_averted_tot = yll_tot * prop_averted,
    daly_averted_tot = yld_averted_tot + yll_averted_tot,
    coi_averted_tot =
      sCh_averted_tot_sum * (pr_mild * patient_cost_outpt +
      pr_moderate * patient_cost_outpt + pr_severe * patient_cost_hosp),
    death_tot_sum = sum(death_tot, na.rm=T),
    cod_averted_tot = sum(death_tot*prop_averted*gdp*life_exp, na.rm=T),
    productivity_lost_averted =
      sum(sCh_averted_tot*gdp*((patient_workday_lost/365)*(pct_workforce/100) +
                                 (caregiver_workday_lost/365)), na.rm=T),
    vacc_dose_tot = sum(as.numeric(vacc_cov) * dose_regimen * population),
    vacc_cost = vacc_dose_tot * vacc_cost_per_dose,
    net_cost_tot = vacc_cost - coi_averted_tot - cod_averted_tot,
    cost_per_daly_averted = net_cost_tot / daly_averted_tot,
    gdp_sum = sum(gdp, na.rm=T),
    death_tot_sum = sum(death_tot, na.rm=T),
    case_averted_per_dose = sCh_averted_tot_sum / vacc_dose_tot,
    death_averted_per_dose = death_tot_sum * prop_averted / vacc_dose_tot)
          
```


Similar figure to the above but with uncertainty across parameter values

```{r}
df_all <- data.table::rbindlist(svim_list) # ma
nrep <- nrow(df_all)/nrow(dd) # 200
df_all$pct_reduction <- 100 * df_all$sCh_averted_tot / df_all$sCh_tot
df_all$size_grp <- rep(dd$size_grp, nrep)
df_all$dur_grp <- rep(dd$size_grp, nrep)

df_all |>
  filter(vacc_cov %in% c(0.9), week_vaccination %in% c(4)) |>
  ggplot()+
  geom_histogram(aes(x=pct_reduction/100, after_stat(count/sum(count))),
                 binwidth = 0.02, fill="steelblue4")+
  labs(x="Case reduction", y="Percent outbreaks", 
       subtitle = "90% vaccination at week 4")+ 
  scale_x_continuous(labels = scales::percent)+
  scale_y_continuous(labels = scales::percent)+
  theme_light()
  
df_all |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduction))+
  geom_boxplot(aes(color=as.factor(vacc_cov)))+
  scale_color_manual("", values=c("orange4","green4"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="week of vaccination", y="percent case reduction")

df_all |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduction))+
  geom_boxplot(aes(color=as.factor(size_grp)))+
  scale_color_manual("", values=c("orange4","green4"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="week of vaccination", y="percent case reduction")  

df_all |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduction))+
  geom_boxplot(aes(color=as.factor(dur_grp)))+
  scale_color_manual("", values=c("orange4","green4"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="week of vaccination", y="percent case reduction")
```

Mean by id across simulation runs

```{r}
# svim_list <- read_rds("outputs/svim_list_20240620.rds")
# compute statistics across 200 runs stored in a list
nc <- length(svim_list)

df <- data.frame(vacc_cov = svim1_c$vacc_cov, vacc_week = svim1_c$week_vaccination, data_id = svim1_c$data_id)
sCh_averted <- sCh <- pct <- cbind(df, data.frame(matrix(NA, nrow=nrow(svim1_c), ncol=nc)))

lst <- vector("list", nc)
for (i in 1:nc) {
  svim_list[[i]] |> 
    group_by(vacc_cov, week_vaccination, data_id) |>
    summarize(.groups = 'drop',
              pct_reduction = 100 * sCh_averted_tot / sCh_tot) -> lst[[i]]
}

d <- rbindlist(lst)
d |> 
  group_by(vacc_cov, week_vaccination, data_id) |>
  summarize(.groups = 'drop',
            mean_pct_reduction = mean(pct_reduction)) -> res


nrow(svim_list[[1]])
nrow(res)
names(res) <- c("vacc_cov", "week_vaccination", "data_id", 
                "pct_reduction", "size_grp", "dur_grp")  
res$size_grp <- dd$size_grp
res$dur_grp <- dd$dur_grp

res |>
  filter(vacc_cov %in% c(0.9), week_vaccination %in% c(4)) |>
  ggplot()+
  geom_histogram(aes(x=pct_reduction/100, after_stat(count/sum(count))),
                 binwidth = 0.02, fill="steelblue4")+
  labs(x="Case reduction", y="Percent outbreaks", 
       subtitle = "90% vaccination at week 4")+ 
  scale_x_continuous(labels = scales::percent)+
  scale_y_continuous(labels = scales::percent)+
  theme_light() -> plt

fac <- 1.2
ggsave(paste0("plots/ind_red_hist_by_week", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
res |> 
  filter(vacc_cov %in% c(0.6,0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduction))+
  geom_boxplot(aes(color=as.factor(vacc_cov)))+
  scale_color_manual(labels=c("60%","90%"), values=c("orange4","green4"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="week of vaccination", y="percent case reduction", color="")

res |>
  filter(vacc_cov %in% c(0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduction))+
  geom_boxplot(aes(color=as.factor(size_grp)))+
  scale_color_manual(labels=c("Outbreaks > 200 cases","Smaller outbreaks"), 
                     values=c("orange4","green4"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="week of vaccination", y="percent case reduction", color="",
       subtitle = "90% vaccination coverage") -> plt 

fac <- 1.2
ggsave(paste0("plots/red_box_by_week_size_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")

res |> 
  filter(vacc_cov %in% c(0.9), week_vaccination %in% c(1,4,7,10,13)) |>  
  ggplot(aes(x=as.factor(week_vaccination), y=pct_reduction))+
  geom_boxplot(aes(color=as.factor(dur_grp)))+
  scale_color_manual(labels=c("Outbreaks > 10 weeks","Shorter outbreaks"), 
                     values=c("orange4","green4"))+
  theme_light()+
  theme(legend.position="top")+
  labs(x="week of vaccination", y="percent case reduction", color="",
       subtitle = "90% vaccination coverage") -> plt 

fac <- 1.2
ggsave(paste0("plots/red_box_by_week_dur_", tstamp(), ".pdf"), plt,
       device = cairo_pdf, width=fgw1*fac, height=fgw1*fac, units="mm")
```


## Interepidemic period

First, we get the date and location names from id, whose shape
```{r}
svim1$location <- sapply(svim1$id, function(x) strsplit(x, "-")[[1]][1])
adms <- lapply(svim1$id, function(x) get_adm(x))
# create admin variables to classify outbreaks accordingly
svim1$admin0 <- sapply(adms, function(x) x[[1]][2])
svim1$admin1 <- sapply(adms, function(x) x[[1]][3])
svim1$admin2 <- sapply(adms, function(x) x[[1]][4])
svim1$admin3 <- sapply(adms, function(x) x[[1]][5])
```


We first identify outbreaks that have happened in each of the admin regions (there must be some overlap) first group by bigger admin units first, then measure the time until the next epidemic. Of X1 countries (states, counties, district), X2 (X3%) experienced recurrent outbreaks over the next X4 period.

Admin 0
```{r}
svim1 |> 
  # group_by(eval(parse(text="admin0"))) |>
  group_by(admin0) |>
  reframe(adm = admin0, 
          id = id,
          date = new_date,
          min_date = min(date), 
          min_date_id = id[which(date == min_date)[1]],
          date_diff = date - min_date) |>
  arrange(-desc(date_diff)) -> res

table(res$admin0)
res[res$admin0 =="gha",]
print(res[res$admin0 =="ner",], n=25)

res <- res |> 
  group_by(admin0) |>
  mutate(inter_epid_period = c(0, diff(date_diff)))

summary(res$inter_epid_period)
 
print(res[res$admin0 =="ner",], n=25)

res |> 
  ggplot(aes(x=admin0))+
  geom_point(aes(y=inter_epid_period))+
  # geom_boxplot(aes(y=inter_epid_period))+
  labs(x="", y="Interepidemic period (day)")+
  theme_light()+
  scale_y_continuous(trans='log10')

# create a dot plot by admin regions - this will help tell the story of 

# length(unique(res[(res$inter_epid_period > 0 & !is.na(res$admin0)),]$adm)) / 
#   length(unique(res[!is.na(res$admin0),]$adm))
summary(res$inter_epid_period)

res |> 
  filter(!is.na(admin0)) |>
  group_by(adm) |>
  summarize(
    n_iep = sum(ifelse(inter_epid_period > 0, 1, 0)),
            min_iep = min(inter_epid_period),
            mean_iep = mean(inter_epid_period),
            median_iep = mean(inter_epid_period),
            max_iep = max(inter_epid_period)) -> res2


outbreak_freq_dist_adm <- function(x) {
  df <- data.frame(matrix(NA, nrow=4, ncol=2))
  names(df) <- c("n", "percent")
  tot <- sum(table(res2$n_iep))
  n1 <- sum(ifelse(tb$niep >= 1, tb$freq, 0))
  df$n[1] <- n1
  df$percent[1] <- n1 / tot * 100
  n2 <- sum(ifelse(tb$niep >= 2, tb$freq, 0))
  df$n[2] <- n2
  df$percent[2] <- n2 / tot*100
  n3 <- sum(ifelse(tb$niep >= 3, tb$freq, 0))
  df$n[3] <- n3
  df$percent[3] <- n3 / tot*100
  n4 <- sum(ifelse(tb$niep >= 5, tb$freq, 0))
  df$n[4] <- n4
  df$percent[4] <- n4 / tot*100
  return(df)
}
table(res2$n_iep)
outbreak_freq_dist_adm(res2)

# tb <- table(res2$n_iep)
# tb <- data.frame(niep = as.integer(names(tb))+1, # 0 means at one outbreak
#                  freq = as.integer(tb))
# (tot <- sum(table(res2$n_iep)))
# (n1 <- sum(ifelse(tb$niep >= 1, tb$freq, 0)))
# n1/tot*100
# (n2 <- sum(ifelse(tb$niep >= 2, tb$freq, 0)))
# n2/tot*100
# (n3 <- sum(ifelse(tb$niep >= 3, tb$freq, 0)))
# n3/tot*100
# (n4 <- sum(ifelse(tb$niep >= 5, tb$freq, 0)))
# n4/tot*100

```

Admin 0 and Admin 1
```{r}
# d_adm0 <- d_adm0[!(d_adm0$id %in% d_adm0$min_date_id), ]
# d_adm0$cumul_prob <- 1:nrow(d_adm0) / nrow(svim1_confirmed)

svim1 |> 
  group_by(admin0, admin1) |>
  reframe(id = id,
          adm = paste0(admin0, "::", admin1),
          date = new_date,
          min_date = min(date),
          min_date_id = id[which(date == min_date)[1]],
          date_diff = date - min_date) |>
  arrange(-desc(date_diff)) -> res

res |> 
  group_by(admin0, admin1) |>
  mutate(inter_epid_period = c(0, diff(date_diff))) -> res

res <- res[order(res$admin0, res$admin1, res$inter_epid_period),]

summary(res$inter_epid_period)

# length(unique(res[(res$inter_epid_period > 0 & !is.na(res$admin1)),]$adm)) / 
#   length(unique(res[!is.na(res$admin1),]$adm))
draw_iep_cdf(res)
res |> 
  filter(!is.na(admin1)) |>
  group_by(adm) |>
  summarize(
    n_iep = sum(ifelse(inter_epid_period > 0, 1, 0)),
            min_iep = min(inter_epid_period),
            mean_iep = mean(inter_epid_period),
            median_iep = mean(inter_epid_period),
            max_iep = max(inter_epid_period)) -> res2

table(res2$n_iep)

outbreak_freq_dist_adm(res2)
```

Admin 0, Admin 1, and Admin 2
```{r}
svim1 |> 
  group_by(admin0, admin1, admin2) |>
  reframe(id = id,
          adm = paste0(admin0, "::", admin1, "::", admin2),
          date = new_date,
          min_date = min(date),
          min_date_id = id[which(date == min_date)[1]],
          date_diff = date - min_date) |>
  arrange(-desc(date_diff)) -> res

res <- res |> 
  group_by(admin0, admin1, admin2) |>
  mutate(inter_epid_period = c(0, diff(date_diff)))

res <- 
  res[order(res$admin0,res$admin1,res$admin2, 
                   res$inter_epid_period),] 

# 
# length(unique(res[(res$inter_epid_period > 0 & !is.na(res$admin2)),]$adm)) / 
#   length(unique(res[!is.na(res$admin2),]$adm))

summary(res$inter_epid_period)
res |> 
  filter(inter_epid_period > 0) |> 
  ggplot(aes(x=inter_epid_period))+
  geom_histogram(binwidth = 10)+
  theme_light() +
  labs(x="Interepidemic period (day)", y="Frequency")

# plot(ecdf(res$inter_epid_period[res$inter_epid_period > 0]))

draw_iep_cdf <- function(x){
  d <- x$inter_epid_period[x$inter_epid_period > 0]
  d <- data.frame(inter_epid_period = sort(d))
  d$cnt <- rep(1, nrow(d))
  d$cdf <- cumsum(d$cnt)/nrow(d)
  d |> 
    ggplot(aes(x=inter_epid_period, y=cdf))+
    geom_step()+
    theme_light()+
    labs(x="Interepidemic period (day)", y="Cumulative prob") -> p
  return(p)
}
draw_iep_cdf(res)
d <- res$inter_epid_period[res$inter_epid_period > 0]
sum(d<90)/length(d)*100
# % outbreaks occur within 3 month since the previous outbreak

sum(d<365*3)/length(d)*100
# % outbreaks occur within 3 years since the previous outbreak
# res |> filter(!is.na(admin2)) |> count()
res |> 
  # filter(inter_epid_period > 0, !is.na(admin2)) |>
  filter(!is.na(admin2)) |>
  group_by(adm) |>
  summarize(
    n_iep = sum(ifelse(inter_epid_period > 0, 1, 0)), # this takes care of outbreaks that have the same dates (i.e., iep=0) because of overlapping admins
    n_iep2 = length(adm), # number of recurrent outbreaks
            min_iep = min(inter_epid_period),
            mean_iep = mean(inter_epid_period),
            median_iep = mean(inter_epid_period),
            max_iep = max(inter_epid_period)) -> res2


# 441 unique admins 
# n_iep correctly captures the number of unique admins
table(res2$n_iep)
outbreak_freq_dist_adm(res2)
```

Interepidemic period (IEP) ~XX percent 

Admin 0, Admin 1, Admin 2, and Admin 3 
```{r}
svim1 |> 
  group_by(admin0, admin1, admin2, admin3) |>
  reframe(id = id,
          adm = paste0(admin0, "::", admin1, "::", admin2, "::", admin3),
          date = new_date,
          min_date = min(date),
          min_date_id = id[which(date == min_date)[1]],
          date_diff = date - min_date) |>
  arrange(-desc(date_diff)) -> res


res |> 
  group_by(admin0, admin1, admin2, admin3) |>
  mutate(inter_epid_period = c(0, diff(date_diff))) -> res

res <- 
  res[order(res$admin0, res$admin1, res$admin2,
                   res$admin3, res$inter_epid_period),]

length(unique(res[res$inter_epid_period != 0,]$adm)) / length(unique(res$adm))
length(unique(res[(res$inter_epid_period != 0 & !is.na(res$admin3)),]$adm)) / 
  length(unique(res[!is.na(res$admin3),]$adm))

# table(res[(!is.na(res$admin3) & res$inter_epid_period > 0),]$inter_epid_period)
# 
# hist(res[(!is.na(res$admin3) & res$inter_epid_period > 0),]$inter_epid_period,
#      breaks=100)
summary(res$inter_epid_period)
res |> 
  filter(!is.na(admin3)) |>
  group_by(adm) |>
  summarize(
    n_iep = sum(ifelse(inter_epid_period > 0, 1, 0)),
            min_iep = min(inter_epid_period),
            mean_iep = mean(inter_epid_period),
            median_iep = mean(inter_epid_period),
            max_iep = max(inter_epid_period)) -> res2

table(res2$n_iep)

tb <- table(res2$n_iep)
tb <- data.frame(niep = as.integer(names(tb))+1, # 0 means at one outbreak
                 freq = as.integer(tb))

outbreak_freq_dist_adm(res2)
```


## Vaccine impact taking lasting immunity into account

We assume some of the outbreaks could be removed if the outbreaks occurred in the same area and the start time of the outbreak is within the duration of vaccine-derived protection

Summarize vaccine impact computation
```{r}
lb <- 180 # days  This could be random delay to vaccine effectiveness
ub <- 365*3
ghost_ids <- find_ghost_outbreaks(x=svim1, lb=lb, ub=ub, admin0, admin1, admin2)
svim1_sub <- svim1[!(svim1$data_id %in% ghost_ids), ]

# vaccine impact
# 100% for the svim1_sub and normal weekly impact for the rest
# cost-of-illness is based on 
coi_per_patient <- (pr_mild * patient_cost_outpt + 
      pr_moderate * patient_cost_outpt + pr_severe * patient_cost_hosp)

svim1_ <- svim1
# for individual outbreak
# removed outbreaks are the same as 100% prevented
svim1 <- mutate(svim1, sCh_averted_tot = 
                  ifelse(data_id %in% ghost_ids,sCh_tot, sCh_averted_tot))  
svim1 |>
  mutate(pct_reduction = 100 * sCh_averted_tot / sCh_tot,
         yld = compute_yld(sCh_tot, parms=parms),
         yld_averted = compute_yld(sCh_averted_tot, parms=parms), 
         yll = compute_yll(death_tot, life_exp, parms),
         yll_averted = yll * (sCh_averted_tot / sCh_tot),
         daly_averted = yld_averted + yll_averted,
         coi_averted = sCh_averted_tot * coi_per_patient,
         cod_averted = death_tot * (sCh_averted_tot / sCh_tot) * gdp * life_exp,
         productivity_lost_averted = 
           sCh_averted_tot * gdp * ((patient_workday_lost/365)*(pct_workforce/100) +
                                 (caregiver_workday_lost/365)),
         vacc_dose = population * dose_regimen,
         vacc_cost = vacc_cost_per_dose * vacc_dose,
         net_cost = vacc_cost - coi_averted - cod_averted,
         cost_per_daly_averted = net_cost / daly_averted,
         case_averted_per_dose = sCh_averted_tot / vacc_dose,
         death_averted_per_dose = death_tot * (sCh_averted_tot / sCh_tot) / vacc_dose) -> svim1

# across outbreaks
svim1 |>
  filter(!(data_id %in% ghost_ids)) |>
  summarize(
    sCh_tot_sum = sum(sCh_tot),
    sCh_averted_tot_sum = sum(sCh_averted_tot),
    prop_averted = sCh_averted_tot_sum / sCh_tot_sum,
    yld_tot = compute_yld(sCh_tot_sum, parms=parms),
    yld_averted_tot = compute_yld(sCh_averted_tot_sum, parms=parms),
    yll_tot = sum(compute_yll(death_tot, life_exp, parms), na.rm=T),
    yll_averted_tot = yll_tot * prop_averted,
    daly_averted_tot = yld_averted_tot + yll_averted_tot,
    coi_averted_tot =
      sCh_averted_tot_sum * (pr_mild * patient_cost_outpt +
      pr_moderate * patient_cost_outpt + pr_severe * patient_cost_hosp),
    death_tot_sum = sum(death_tot, na.rm=T),
    cod_averted_tot = sum(death_tot*prop_averted*gdp*life_exp, na.rm=T),
    productivity_lost_averted =
      sum(sCh_averted_tot*gdp*((patient_workday_lost/365)*(pct_workforce/100) +
                                 (caregiver_workday_lost/365)), na.rm=T),
    vacc_dose_tot = sum(as.numeric(vacc_cov) * dose_regimen * population),
    vacc_cost = vacc_dose_tot * vacc_cost_per_dose,
    net_cost_tot = vacc_cost - coi_averted_tot - cod_averted_tot,
    cost_per_daly_averted = net_cost_tot / daly_averted_tot,
    gdp_sum = sum(gdp, na.rm=T),
    death_tot_sum = sum(death_tot, na.rm=T),
    case_averted_per_dose = sCh_averted_tot_sum / vacc_dose_tot,
    death_averted_per_dose = death_tot_sum * prop_averted / vacc_dose_tot)
          
```


```{r}
svim1_s
# compute statistics across 200 runs stored in a list
vc <- 0.9
wk <- 6
nc <- length(svim_list)
df <- data.frame(vacc_cov = rep(vc, length(vacc_protect_period)), 
                 vacc_week = rep(wk, length(vacc_protect_period)),
                 vacc_protect_period = vacc_protect_period)

death_averted_per_dose <- 
  case_averted_per_dose <- cbind(df, data.frame(matrix(NA, nrow=1, ncol=nc)))

for (i in 1:length(vacc_protect_period)) {
  d_adm2 |> 
    filter(lb < date_diff & date_diff < vacc_protect_period[i]) -> d
  
  for (j in 1:nc){
    svim_list[[j]] |> 
      filter(!is.na(confirmed_tot) & confirmed_tot > 0) |>
      filter(vacc_cov == vc, week_vaccination == wk) -> v9w6
    
    v9w6 |> filter(id %in% d$id) -> sd
    v9w6$population <- sapply(1:nrow(v9w6), function(i) dat[dat$ID == v9w6$data_id[i], "population"])
    
    v9w6 |>
      filter(!(id %in% d$id)) |> 
      summarize(.groups = "drop",
                vacc_doses_tot = sum(vacc_cov * dose_regimen * population),
                sCh_tot_sum = sum(sCh_tot),
                sCh_averted_tot_sum = sum(sCh_averted_tot) + sum(sd$sCh_tot),
                prop_averted = sCh_averted_tot_sum / sCh_tot_sum,
                death_tot_sum = sum(death_tot, na.rm=T),
                case_averted_per_dose = sCh_averted_tot_sum / vacc_doses_tot,
                death_averted_per_dose = death_tot_sum * prop_averted / vacc_doses_tot) -> svim1_s

    case_averted_per_dose[i, j+3] <- svim1_s$case_averted_per_dose
    death_averted_per_dose[i, j+3] <- svim1_s$death_averted_per_dose
  }
}

pr <- c(0.025,0.25,0.5,0.75,0.975)
case_averted_per_dose_qnt <- apply(case_averted_per_dose[,4:203], 1, quantile, 
                     probs=pr)
case_averted_per_dose_qnt <- cbind(case_averted_per_dose[,1:3], 
                                   t(case_averted_per_dose_qnt))
death_averted_per_dose_qnt <- apply(death_averted_per_dose[,4:203], 1, quantile, 
                             probs=pr)
death_averted_per_dose_qnt <- cbind(death_averted_per_dose[,1:3],
                                    t(death_averted_per_dose_qnt))

case_averted_per_dose_qnt |>
  ggplot(aes(x=vacc_protect_period/365, y=`50%`*1000))+
  geom_pointrange(aes(x=vacc_protect_period/365,ymin=`2.5%`*1000, ymax=`97.5%`*1000),
  size=0.5) + 
  labs(x="Period of vaccine-derived protection (year)",
       y="Case averted per 1,000 dose of OCV")+
  ggtitle('Vaccine coverage = 90%, Week of vaccination=6')+
  theme_bw() -> plt
  
plt

ggsave(paste0("plots/case_averted_per_dose_vacc_prot_vc",vc*10, "wk", wk,"_", tstamp(), ".pdf"), plt,
       width=fgw2, height=fgw2, units="mm")
```







```{r}
d_adm1 <- d_adm1[!(d_adm1$id %in% d_adm1$min_date_id), ]
d_adm1$cumul_prob <- 1:nrow(d_adm1) / nrow(svim1_confirmed)

svim1_confirmed |> 
  group_by(admin0, admin1, admin2) |>
  reframe(adm = "Admin 2", 
          id = id,
          date = as.Date(date),
          min_date = min(date),
          min_date_id = id[which(date == min_date)[1]],
          date_diff = date - min_date) |>
  arrange(-desc(date_diff)) -> d_adm2

d_adm2 <- d_adm2[!(d_adm2$id %in% d_adm2$min_date_id), ]
d_adm2 <- d_adm2[!is.na(d_adm2$admin2),]
d_adm2$cumul_prob <- 1:nrow(d_adm2) / nrow(svim1_confirmed)

svim1_confirmed |> 
  group_by(admin0, admin1, admin2, admin3) |>
  reframe(adm = "Admin 3", 
          id = id,
          date = as.Date(date),
          min_date = min(date),
          min_date_id = id[which(date == min_date)[1]],
          date_diff = date - min_date) |>
  arrange(-desc(date_diff)) -> d_adm3

d_adm3 <- d_adm3[!(d_adm3$id %in% d_adm3$min_date_id), ]
d_adm3 <- d_adm3[!is.na(d_adm3$admin3),]
d_adm3$cumul_prob <- 1:nrow(d_adm3) / nrow(svim1_confirmed)

cols <- c("adm", "cumul_prob", "date_diff")
d <- rbind(d_adm0[,cols], d_adm1[,cols], d_adm2[,cols])

d |> group_by(adm) |> summarize(max_cumulprob = max(cumul_prob))

d |> 
  ggplot(aes(as.numeric(date_diff)/365, cumul_prob, color=adm, group=adm))+
  geom_step(linewidth=1)+
  labs(x="Year since the first outbreak in the given administrative unit", 
       y="Cumulative probability", color="")+
  theme_bw()+
  # scale_colour_viridis_d(option="H")+
  scale_colour_viridis_d()+
  theme(legend.position = "bottom") -> plt

# ggsave(paste0("plots/recurrent_outbreaks_", tstamp(), ".pdf"), plt,
#         width=fgw2, height=fgw2, units="mm")
```


```{r}
vc_interest<- c(0.6, 0.9)
svim_list <- readRDS("outputs/svim_list_20240620.rds")
# percent change
svim_list[[1]] |> filter(vacc_cov == vc_interest) |> 
  group_by(vacc_cov, week_vaccination) |> 
  summarize(sCh_tot_sum = sum(sCh_tot),
            sCh_averted_tot_sum = sum(sCh_averted_tot),
            pct_reduction = 100 * sCh_averted_tot_sum / sCh_tot_sum) -> svim1_s
# 16 weeks for 9 vaccine coverages
(16*9 == nrow(svim1_s))

# compute statistics across 200 runs stored in a list
nc <- length(svim_list)

df <- data.frame(vacc_cov = svim1_s$vacc_cov, vacc_week = svim1_s$week_vaccination)
pct <- cbind(df, data.frame(matrix(NA, nrow=16*9, ncol=nc)))

for (i in 1:nc){
  svim_list[[i]] |>
    filter(!is.na(confirmed_tot) & confirmed_tot > 0) |>
    group_by(vacc_cov, week_vaccination) |>
    summarize(.groups = 'drop',
      sCh_tot_sum = sum(sCh_tot),
      sCh_averted_tot_sum = sum(sCh_averted_tot),
      pct_reduction = 100 * sCh_averted_tot_sum / sCh_tot_sum) -> svim_s
  
  pct[, i+2] <- svim_s$pct_reduction
}


pct_qnt <- apply(pct[,3:202], 1, quantile, probs=c(0.025,0.25,0.5,0.75,0.975))
pct_qnt <- cbind(pct[,1:2], t(pct_qnt))


pct_qnt |> 
  group_by(vacc_cov, vacc_week) |> 
  summarize(.groups = 'drop',
    vacc_cov = 100 * vacc_cov[1],
    vacc_week = vacc_week,
    vacc_eff = `50%`)  |> 
  ggplot(aes(x=vacc_week, y=vacc_cov, z=vacc_eff))+
  geom_contour_filled()+
  labs(x="Week of vaccination",
       y="Vaccine coverage (%)",
       fill="Case averted (%)") +
  scale_x_continuous(breaks=seq(1,16,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour() -> p1


pct_qnt |> 
  group_by(vacc_cov) |> 
  summarize(.groups='drop',
            vacc_week = vacc_week, 
            pct_change = c(`50%`[1:15] - `50%`[2:16], NA)) -> pct_qnt_change

pct_qnt_change |>
  group_by(vacc_cov, vacc_week) |> 
  summarize(.groups = 'drop',
    vacc_cov = 100 * vacc_cov[1],
    vacc_week = vacc_week,
    vacc_eff_change = pct_change) |> 
  ggplot(aes(x=vacc_week, y=vacc_cov, z=vacc_eff_change))+
  geom_contour_filled()+
  # theme_minimal()+
  labs(x="Week of vaccination",
       y="Vaccine coverage (%)",
       fill="Vaccine impact\nchange (%)") +
  scale_x_continuous(breaks=seq(1,16,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour() -> p2


library(patchwork)     
p <-  p1 + p2 + plot_annotation(tag_levels = list(c("A","B")))
p
ggsave(paste0("plots/pct_change_case_averted_", tstamp(), ".pdf"), p,
        width=fgw2*1.8, height=fgw2*0.9, units="mm")
```


Figure 2: The impact of the vaccine based on the static model
```{r, eval=F}
library(viridis)
library(patchwork)

svim_list <- readRDS("outputs/svim_list_20240620.rds")
# svim_raw <- as.data.frame(fread("outputs/svim_mean_200_20240620.csv"))
# svim_raw_confirmed <- filter(svim_raw, (!is.na(confirmed_tot) & confirmed_tot > 0))

# This is deprecated version of computation and used ONLY for ploting
# example vaccine impact
svim_ex <- as.data.frame(fread("outputs/static_vim20240427.csv"))
# one example of lab-confirmed cholera outbreak
# the shape of the outbreak close to a typical outbreak
id0 <- ds[ds$ID == 412, "ID_outbreak"]
svim_ex |>
  filter(vacc_cov == 0.8, id == id0) |>
  mutate(case_wk_occurred_tot = case_wk_tot - case_wk_averted_tot) -> d

dar <- data.frame(x=c(1,1), y=c(20,10)) # data for arrow
ggplot()+
  geom_point(data=subset(d, week>=4), aes(x=week,y=case_wk_occurred_tot), 
             color="steelblue4")+
  geom_point(data=subset(d, week>=5), aes(x=week,y=case_wk_occurred_tot), 
             color="firebrick4")+
  geom_line(data=d, aes(x=week, y=case_wk_tot))+
  geom_point(data=d, aes(x=week, y=case_wk_tot))+
  geom_ribbon(data=subset(d, week >= 5),
              aes(x=week, ymin=case_wk_occurred_tot,
                  ymax=case_wk_tot), fill="firebrick4", alpha=0.3)+
  geom_ribbon(data=subset(d, (4 <= week & week <= 5)),
              aes(x=week, ymin=case_wk_occurred_tot,
                  ymax=case_wk_tot), fill="steelblue4", alpha=0.3)+
  geom_segment(data=dar, aes(x=1,y=24,xend=1,yend=10),
               arrow=arrow(type="closed", angle=10, length=unit(5,"mm")))+
  annotate("text", x=1, y=26, label= "Vaccination", hjust = 0) + 
  theme_bw()+
  labs(x="Epidemic week", y="Weekly incidence") -> p1


# cumulative percentage of outbreaks across week
svim_ex |> 
  filter(!is.na(confirmed), confirmed > 0) |>
  group_by(week) |> 
  summarize(n = n()) -> n_outbks_per_week

n_outbks_per_week$surv_prob <- n_outbks_per_week$n / n_outbks_per_week$n[1]

ggplot(n_outbks_per_week, aes(x=week, y=surv_prob*100)) +
  geom_step() +
  scale_x_continuous(limits = c(0,16))+
  # labs(x="", y=expression(Prob~(T >~italic(t)))) +
  labs(x="Week", y="Cumulative\npercentage") +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA_character_), 
    # necessary to avoid drawing panel outline
    # panel.grid.major = element_blank(), # get rid of major grid
    # panel.grid.minor = element_blank(), # get rid of minor grid
    plot.background = element_rect(fill = "transparent",
                                   colour = NA_character_), # necessary to avoid drawing plot outline
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent"),
    plot.margin = margin(t=5, r=5, b=0.1, l=5, unit="pt")) -> p2a # reduce the bottom margin

# unique(svim_lab_conf$data_id)

svim_raw_confirmed |> 
  filter(vacc_cov >= 0.799, vacc_cov <= 0.801) |>
  ggplot(aes(x = week_vaccination, y=sCh_averted_tot, group = week_vaccination))+
  # geom_violin(linewidth = 0.6)+
  geom_boxplot(linewidth = 0.6)+
  scale_x_continuous(limits = c(0,16))+
  # geom_jitter(width=0.1, size=0.5, color="steelblue",alpha=0.5)+
  labs(x="Week of vaccination",
       y="Total cases averted\nper outbreak") +
  theme_bw()+
  theme(plot.margin = margin(t=0.1, r=5, b=5, l=5, unit="pt")) -> p2b

p2 <- p2a / p2b + plot_layout(heights = c(1, 2))

# tile plot of % averted cases for each outbreak for each simulation run  
# 113 confirmed outbreaks and 200 simulation runs
df <- data.frame(matrix(NA, nrow=113, ncol=200))
for (i in 1:length(svim_list)){
  svim_list[[i]] |> 
  filter(!is.na(confirmed_tot) & confirmed_tot > 0) |>
  filter(vacc_cov == 0.8 & week_vaccination == 3) |>
  group_by(data_id) |> 
  summarize(pct_reduction = 100 * sCh_averted_tot / sCh_tot ) -> d 
  df[,i] <- d$pct_reduction  
}
    
df <- cbind(data.frame(id = 1:113), df) 
dflong <- pivot_longer(df, cols = 2:201, 
                           names_to = "runid", values_to = "pct_reduction")
dflong$runid <- rep(1:200, 113)
  
dflong |> ggplot(aes(runid, id))+ 
  geom_tile(aes(fill=pct_reduction))+
  theme_minimal()+
  theme(axis.ticks.length = unit(.01, "mm"),
        legend.position = "bottom")+
  scale_fill_viridis(option="magma", limits=c(0,100),
                     direction = -1) +
  labs(x="Simulation run ID", y="Outbreak ID", 
       fill="% case averted") -> p3


# plot of outbreak size vs outbreak impact
svim_list[[i]] |> filter(!is.na(confirmed_tot) & confirmed_tot > 0) ->
  svim1_confirmed

df <- data.frame(matrix(NA, 
                        nrow=nrow(svim1_confirmed), 
                        ncol=length(svim_list)))

for (i in 1:length(svim_list)){
  # only confirmed outbreaks
  svim_list[[i]] |> filter(!is.na(confirmed_tot) & confirmed_tot > 0) -> d
  df[,i] <- 100 * d[, "sCh_averted_tot"] / d[, "sCh_tot"] 
}

d <- data.frame(vacc_cov = svim1_confirmed$vacc_cov, 
           vacc_week = svim1_confirmed$week_vaccination,
           outbreak_size = svim1_confirmed$sCh_tot,
           mean_pct_reduction = rowMeans(df, na.rm=TRUE))


d |> filter(vacc_cov %in% c(0.5, 0.8), vacc_week %in% c(1, 4)) -> d2

d2$grp <- 
  ifelse(d2$vacc_cov == 0.8 & d2$vacc_week == 1, "80%, Week 1", 
  ifelse(d2$vacc_cov == 0.8 & d2$vacc_week == 4, "80%, Week 4",
  ifelse(d2$vacc_cov == 0.5 & d2$vacc_week == 1, "50%, Week 1", 
         "50%, Week 4")))

d2 |> 
  ggplot(aes(x=outbreak_size, y=mean_pct_reduction, 
             color=c(as.factor(grp))))+
  geom_point(size=2, alpha=0.6) +
  geom_smooth(method = "lm")+
  scale_x_log10()+
  scale_color_viridis(option="magma", discrete=TRUE)+
  labs(x="Outbreak size",
       y="Mean case averted (%)",
       color ="")+
  theme_bw()+
  theme(legend.position="bottom") -> p4

library(patchwork)
p <- (p1 | p2) / (p3 | p4) + plot_annotation(tag_levels = list(c("A","B","","C","D")))
p
ggsave(paste0("plots/vi_fig2_", tstamp(), ".pdf"), p,
        width=fgw2*1.8, height=fgw2*0.9, units="mm")
```



Figure S2
All suspected outbreaks
```{r, eval=F}
svim |> 
  filter(vacc_cov == 0.7) |>
  ggplot(aes(x=week,y=case_rem_averted_tot,group=week))+
  geom_violin(linewidth=0.6)+
  geom_jitter(width=0.1,size=0.5,color="steelblue",alpha=0.5)+
  labs(x="Week of protection",
       y="Total cases averted per outbreak") +
  theme_bw() -> p2

svim |> 
  filter(week < maxwk) |>
  group_by(vacc_cov, week) |> 
  summarize(
    vacc_cov = 100*vacc_cov[1],
    vacc_week = week[1],
    vacc_eff = mean(pct_case_averted_tot)) |> 
  ggplot(aes(x=vacc_week, y=vacc_cov, z=vacc_eff))+
  geom_contour_filled()+
  labs(x="Week of protection",
       y="Vaccine coverage (%)",
       fill="Case averted (%)") +
  scale_x_continuous(breaks=seq(1,15,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour() -> p3

svim |> 
  filter(week < 15) |>
  group_by(vacc_cov, week) |> 
  summarize(
    vacc_cov = 100*vacc_cov[1],
    vacc_week = week[1],
    vacc_eff_gain = mean(pct_gain_case_averted_tot,na.rm=T)) |> 
  ggplot(aes(x=vacc_week, y=vacc_cov, z=vacc_eff_gain))+
  geom_contour_filled()+
  labs(x="Week of protection",
       y="Vaccine coverage (%)",
       fill="Percent gain (%)") +
  scale_x_continuous(breaks=seq(1,15,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour() -> p4

library(patchwork)     
p <- (p2/p3/p4) + plot_annotation(tag_levels = "A")

ggsave(paste0("plots/vim_susp_static_contour_", tstamp(), ".pdf"), p,
       width=fgw2, height=fgw2*2.4, units="mm")
```

Figure S3
Vaccine impact for a subset of outbreaks by duration, size, ... 
```{r, eval=F}
# unique(svim_lab_conf$data_id)
case_qnt <- quantile(svim$case_total, probs=seq(0.1,0.9,0.1))
case_quant <- function(x, cutoffs){
  case_qnt <- rep(1,length(x))
  for(k in 1:length(x)){
    for(i in 1:(length(cutoffs)-1)){
      if(cutoffs[i] <= x[k] & x[k] < cutoffs[i+1]){
        case_qnt[k] <- i+1;
        break
      }
    }
    if(x[k]>= cutoffs[9]){
      case_qnt[k] <- 10;
    }
  }
  return(case_qnt)
}

case_quant(c(4,505), case_qnt)
svim$case_percentile <- case_quant(x=svim$case_total, cutoffs=case_qnt)

svim |> filter(!is.na(confirmed), confirmed > 0, week <= maxwk) -> d

d |>
  group_by(week, case_percentile) |> 
  summarize(
    outbk_size_pct = case_percentile*10,
    vacc_cov = 100*vacc_cov[1],
    vacc_eff = mean(pct_case_averted_tot)) -> svim_week_pct

svim_week_pct |> 
  ggplot(aes(x=week, y=outbk_size_pct, z=vacc_eff))+
  geom_contour_filled()+
  labs(x="Week of protection",
       y="Outbreaks size percentile",
       fill="Case averted (%)") +
  scale_x_continuous(breaks=seq(1,15,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour()

# cumulative 
wks <- unique(d$week)
pcts <- unique(d$case_percentile)
df <- expand.grid(week=wks, percentile=pcts)
df$vacc_eff_cum <- NA

for (i in 1:nrow(df)) {
  df$vacc_eff_cum[i] <- 
    mean(d[week==df$week[i] & case_percentile <= df$percentile[i],
           pct_case_averted_tot])
}

df |>
  ggplot(aes(x=week, y=percentile*10, z=vacc_eff_cum))+
  geom_contour_filled()+
  labs(x="Week of protection",
       y="Outbreaks size percentile",
       fill="Case averted (%)") +
  scale_x_continuous(breaks=seq(1,15,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour()+
  theme(legend.position = "bottom") -> p

ggsave(paste0("plots/svim_cum_outbksize_pct", tstamp(), ".pdf"), p,
       width=fgw2, height=fgw2, units="mm", scale=1)
```


Tables S2 and S3
The impact of the vaccine across vaccine coverage and week of vaccine
protection for 113 confirmed and 824 suspected cholera outbreaks.
```{r}
svim_ <- fread("outputs/static_vim20240427.csv")
maxwk <- 16
svim <- adjust_week_pct_effect(svim_, maxwk)
svim |> filter(!is.na(confirmed), confirmed > 0) -> svim_lab_conf

svim |> 
  filter(week <= maxwk) |>
  # filter(!is.na(confirmed), confirmed > 0) |>
  group_by(vacc_cov, week) |> 
  summarize(
    pct_vacc_cov = 100*vacc_cov[1],
    mean_pct_case_averted_tot = mean(pct_case_averted_tot),
    mean_pct_case_averted_U5 = mean(pct_case_averted_U5),
    mean_pct_case_averted_5up = mean(pct_case_averted_5up),
    sd_pct_case_averted_tot = sd(pct_case_averted_tot),
    sd_pct_case_averted_U5 = sd(pct_case_averted_U5),
    sd_pct_case_averted_5up = sd(pct_case_averted_5up),
    max_pct_case_averted_tot = max(pct_case_averted_tot),
    max_pct_case_averted_U5 = max(pct_case_averted_U5),
    max_pct_case_averted_5up = max(pct_case_averted_5up),
    min_pct_case_averted_tot = min(pct_case_averted_tot),
    min_pct_case_averted_U5 = min(pct_case_averted_U5),
    min_pct_case_averted_5up = min(pct_case_averted_5up),
    median_pct_case_averted_tot = median(pct_case_averted_tot),
    median_pct_case_averted_U5 = median(pct_case_averted_U5),
    median_pct_case_averted_5up = median(pct_case_averted_5up),
    lb_pct_case_averted_tot = quantile(pct_case_averted_tot, probs=c(0.025)),
    lb_pct_case_averted_U5 = quantile(pct_case_averted_U5, probs=c(0.025)),
    lb_pct_case_averted_5up = quantile(pct_case_averted_5up, probs=c(0.025)),
    ub_pct_case_averted_tot = quantile(pct_case_averted_tot, probs=c(0.975)),
    ub_pct_case_averted_U5 = quantile(pct_case_averted_U5, probs=c(0.975)),
    ub_pct_case_averted_5up = quantile(pct_case_averted_5up, probs=c(0.975)),
    lb_IQR_pct_case_averted_tot = quantile(pct_case_averted_tot, probs=c(0.25)),
    lb_IQR_pct_case_averted_U5 = quantile(pct_case_averted_U5, probs=c(0.25)),
    lb_IQR_pct_case_averted_5up = quantile(pct_case_averted_5up, probs=c(0.25)),
    ub_IQR_pct_case_averted_tot = quantile(pct_case_averted_tot, probs=c(0.75)),
    ub_IQR_pct_case_averted_U5 = quantile(pct_case_averted_U5, probs=c(0.75)),
    ub_IQR_pct_case_averted_5up = quantile(pct_case_averted_5up, probs=c(0.75))) -> d




write_med_bds <- function(med, lb, ub, digits=1) {
  med <- format(round(med, digits=digits), big.mark=",", trim=TRUE)
  lb <- format(round(lb, digits=digits), big.mark=",", trim=TRUE)
  ub <- format(round(ub, digits=digits), big.mark=",", trim=TRUE)
  paste0(med, " (", lb , " - ", ub, ")")
}

# d |> filter(pct_vacc_cov == 90) -> d2

# med_IQR_tot <- write_med_bds(med = d2$median_pct_case_averted_tot, 
#                     lb = d2$lb_IQR_pct_case_averted_tot,
#                     ub = d2$ub_IQR_pct_case_averted_tot)
# med_IQR_U5 <- write_med_bds(med = d2$median_pct_case_averted_U5, 
#                     lb = d2$lb_IQR_pct_case_averted_U5,
#                     ub = d2$ub_IQR_pct_case_averted_U5)
# med_IQR_5up <- write_med_bds(med = d2$median_pct_case_averted_5up, 
#                     lb = d2$lb_IQR_pct_case_averted_5up,
#                     ub = d2$ub_IQR_pct_case_averted_5up)

# tab <- data.frame(
#   vc = rep(90,16),
#   wp = 1:16, 
#   ca_tot = med_IQR_tot,
#   ca_U5 = med_IQR_U5,
#   ca_5up = med_IQR_5up)
# 
# names(tab) <- c("Vaccine coverage (%)","Week of protection", 
#   "Total case averted (%)", "Case averted for 0-4 yo (%)",
#   "Case averted for 5+ yo (%)")

vcs <-  c(10,30,50,70,90)
lst <- vector("list", length(vcs))
for (i in 1:length(vcs)) {
  d |> filter(pct_vacc_cov == vcs[i] ) -> d2

  med_IQR_tot <- write_med_bds(med = d2$median_pct_case_averted_tot, 
                    lb = d2$lb_IQR_pct_case_averted_tot,
                    ub = d2$ub_IQR_pct_case_averted_tot)
  med_IQR_U5 <- write_med_bds(med = d2$median_pct_case_averted_U5, 
                    lb = d2$lb_IQR_pct_case_averted_U5,
                    ub = d2$ub_IQR_pct_case_averted_U5)
  med_IQR_5up <- write_med_bds(med = d2$median_pct_case_averted_5up, 
                    lb = d2$lb_IQR_pct_case_averted_5up,
                    ub = d2$ub_IQR_pct_case_averted_5up)

  df <- data.frame(
    vc = rep(vcs[i], maxwk),
    wp = 1:maxwk, 
    ca_tot = med_IQR_tot,
    ca_U5 = med_IQR_U5,
    ca_5up = med_IQR_5up)
  
  lst[[i]] <- df
} 
tab <- rbindlist(lst)
names(tab) <- c("Vaccine coverage (%)","Week of protection", 
  "Total case averted (%)", "Case averted for 0-4 yo (%)",
  "Case averted for 5+ yo (%)")

# fwrite(tab, paste0("outputs/static_vim_pct_lab_conf_", tstamp(), ".csv"))
fwrite(tab, paste0("outputs/static_vim_pct_all_", tstamp(), ".csv"))
```


DALY table
The number of cases, deaths, and DALYs averted for 113 confirmed and 824 suspected cholera outbreaks.
```{r}
svim_ <- fread("outputs/static_vim20240503.csv")
maxwk <- 16
svim <- adjust_week_pct_effect(svim_, maxwk)

svim |> 
  filter(week <= maxwk) |>
  # filter(!is.na(confirmed), confirmed > 0) |>
  group_by(vacc_cov, week) |> 
  summarize(
    cases = sum(case_total, na.rm=T),
    case_averted_tot = round(sum(case_rem_averted_tot, na.rm=T)),
    death_averted_tot = round(sum(death_averted, na.rm=T)),
    daly_averted = round(sum(daly_averted, na.rm=T))) -> d

d
names(d) <- c("Vaccine coverage (%)","Week of protection", 
  "Total case expected", "Case averted" , "Death averted", 
  "DALYs averted")

# fwrite(d, paste0("outputs/averted_vim_labconf_", tstamp(), ".csv"))
fwrite(d, paste0("outputs/averted_vim_all_", tstamp(), ".csv"))
```

DALY averted per dose of vaccine table
The number of cases, deaths, and DALYs averted for 113 confirmed and 824 suspected cholera outbreaks.
```{r}
svim_ <- fread("outputs/static_vim20240503.csv")
maxwk <- 16
svim <- adjust_week_pct_effect(svim_, maxwk)

svim_ |> 
  filter(week <= maxwk) |>
  # filter(!is.na(confirmed), confirmed > 0) |>
  group_by(vacc_cov, week) |> 
  summarize(
    vacc_doses = sum(vacc_doses, na.rm=T),
    mean_case_averted_per_dose = mean(case_averted_per_dose, na.rm=T),
    mean_death_averted_per_dose = mean(death_averted_per_dose, na.rm=T),
    mean_dose_per_case_averted = mean(dose_per_case_averted, na.rm=T),
    mean_dose_per_death_averted = mean(dose_per_death_averted, na.rm=T)) -> d_

d
names(d) <- c("Vaccine coverage","Week of protection", 
  "Vaccine doses", "Mean case averted per dose", "Mean death averted per dose" , "Mean dose per case averted", 
  "Mean dose per death averted")

# fwrite(d, paste0("outputs/averted_vim_labconf_", tstamp(), ".csv"))
fwrite(d, paste0("outputs/dose_averted_vim_all_", tstamp(), ".csv"))
```


Table S3 and S4
```{r}
# d <- fread("outputs/averted_vim_labconf_20240503.csv")
d <- fread("outputs/averted_vim_all_20240503.csv")
names(d)
d |> 
  filter(`Vaccine coverage (%)` %in% seq(0.1,0.9,by=0.2)) |> 
  mutate(`Vaccine coverage (%)` = 100 * `Vaccine coverage (%)`) -> d

# fwrite(d, paste0("outputs/averted_vim_simple_labconf_", tstamp(), ".csv"))
fwrite(d, paste0("outputs/averted_vim_simple_all_", tstamp(), ".csv"))
```


Descriptive statistics
Good fits and all outbreaks
Median duration and total suspected cases
```{r}
svim <- fread("outputs/static_vim20240427.csv")
svim |> filter(good_fit == 1) -> svim_a
idx_goodfits <- unique(svim_a$data_id)

summary(dat$duration)
summary(dat[dat$ID %in% idx_goodfits,]$duration)

summary(dat$total_suspected_cases)
summary(dat[dat$ID %in% idx_goodfits,]$total_suspected_cases)

summary(dat$time_to_peak..weeks)
summary(dat[dat$ID %in% idx_goodfits,]$time_to_peak..weeks)
```

Figure 4.
Dynamic model 
```{r}
svim <- fread("outputs/static_vim20240427.csv")
svim |> filter(!is.na(confirmed), confirmed > 0, good_fit == 1) -> svim_a
idx <- unique(svim_a$data_id)
length(idx)
# idx <- 898 # looks good, long enough to check
dvim <- fread("outputs/dynamic_vim20240425.csv")
dvim |> 
  group_by(data_id, vacc_cov) |>
  mutate(pct_gain = c(-diff(vacc_eff),NA)) -> d
         
d |>
  filter(data_id %in% svim_a$data_id) |>
  group_by(vacc_week, vacc_cov) |> 
  summarize(mean_vacc_eff = mean(vacc_eff),
            mean_pct_gain = mean(pct_gain, na.rm=T)) -> d2

p1 <- ggplot(d2)+
  geom_contour_filled(aes(x=vacc_week+1, y=vacc_cov*10, z=mean_vacc_eff))+
  labs(x="Week of vaccination",
       y="Vaccine coverage rate (%)",
       fill="Case\naverted\n(%)")+
  scale_x_continuous(breaks=seq(1,13,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour()

p2 <- ggplot(d2)+
  geom_contour_filled(aes(x=vacc_week+1, y=vacc_cov*10, z=mean_pct_gain))+
  labs(x="Week of vaccination",
       y="Vaccine coverage rate (%)",
       fill="Percent\ndifference")+
  scale_x_continuous(breaks=seq(1,13,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour()


library(patchwork)
p <- (p1 | p2) + plot_annotation(tag_levels = "A")

ggsave(paste0("plots/vim_dynamic_labconf_", tstamp(), ".pdf"), p,
       width=fgw2*1.8, height=fgw2, units="mm")
```


Figure S4
```{r}
# 
idx <- unique(svim_a$data_id)
# idx <- 898 # looks good, long enough to check
dvim <- fread("outputs/dynamic_vim20240425.csv")
svim$vacc_cov <- as.character(svim$vacc_cov)

lst <- vector("list", length(idx))
for (i in 1:length(idx)) {
  id0 <- idx[i]
  cat("id =", id0, "\n")
  df <- expand.grid(id = id0, vacc_cov = seq(0.1,0.9,0.1)*100, 
                    model = c("static","dynamic"), vacc_eff = NA)
  # Week 0 for the dynamic model may correspond to Week 2 in the static model
  for (j in unique(df$vacc_cov)) {
    sv <- svim[(svim$data_id == id0 & svim$vacc_cov == as.character(j/100)),]
    if (max(sv$week) >= 3) {
      df[df$model == "static" & df$vacc_cov==j,]$vacc_eff <- 
        sv[sv$week == 3,]$pct_case_averted_tot
    }
    dv <- dvim[(dvim$data_id == id0 & dvim$vacc_cov == as.integer(j/10)),]
    if (max(dv$vacc_week) >= 0) {
      df[df$model == "dynamic" & df$vacc_cov==j,]$vacc_eff <- 
        dv[dv$vacc_week == 0,]$vacc_eff
    }
  }
  lst[[i]] <- df
}
df <- rbindlist(lst)
df |>
  filter(id %in% c(898,428,930)) |>
  ggplot(aes(x=vacc_cov, y=vacc_eff, color=model))+
  geom_line(aes(linetype = factor(id)), linewidth=0.6)+
  scale_color_manual(labels=c("Static","Dynamic"), 
                     values=c("steelblue","firebrick"))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_x_continuous(breaks=seq(10,90,20))+
  scale_y_continuous(breaks=seq(0,100,20))+
  labs(x="Vaccine coverage", y="Case averted (%)", linetype="", color="")+
  guides(linetype="none")+
  theme_bw()+
  theme(legend.position="bottom") -> p

ggsave(paste0("plots/vim_ex_comparison_static_dynamic_", tstamp(), ".pdf"), p,
       width=fgw2, height=fgw2, units="mm")

```


Figure S4A, S4B.
Dynamic model vaccine impact for all suspected outbreaks
```{r}
svim <- fread("outputs/static_vim20240427.csv")
svim |> filter(good_fit == 1) -> svim_a
idx <- unique(svim_a$data_id)
length(idx)
dvim <- fread("outputs/dynamic_vim20240425.csv")
dvim |> 
  group_by(data_id, vacc_cov) |>
  mutate(pct_gain = c(-diff(vacc_eff),NA)) -> d
         
d |>
  filter(data_id %in% svim_a$data_id) |>
  group_by(vacc_week, vacc_cov) |> 
  summarize(mean_vacc_eff = mean(vacc_eff),
            mean_pct_gain = mean(pct_gain, na.rm=T)) -> d2

p1 <- ggplot(d2)+
  geom_contour_filled(aes(x=vacc_week+1, y=vacc_cov*10, z=mean_vacc_eff))+
  labs(x="Week of vaccination",
       y="Vaccine coverage rate (%)",
       fill="Case\naverted\n(%)")+
  # scale_x_continuous(breaks=seq(1,13,2),expand=c(0,0))+
  # scale_y_continuous(breaks=seq(10,90,20),expand=c(0,0))+
  scale_x_continuous(breaks=seq(1,13,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour()

p2 <- ggplot(d2)+
  geom_contour_filled(aes(x=vacc_week+1, y=vacc_cov*10, z=mean_pct_gain))+
  labs(x="Week of vaccination",
       y="Vaccine coverage rate (%)",
       fill="Percent\ndifference")+
  scale_x_continuous(breaks=seq(1,13,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour()


library(patchwork)
p <- (p1 | p2) + plot_annotation(tag_levels = "A")

ggsave(paste0("plots/vim_dynamic_all_", tstamp(), ".pdf"), p,
       width=fgw2*1.8, height=fgw2, units="mm")
```


Figure 4C, 4D.
Dynamic vs. static vaccine impact for good fits 
```{r}

```


Figure 5.
The number of vaccine doses required to avert a case and death
```{r}
svim_ <- fread("outputs/static_vim20240503.csv")
maxwk <- 16
svim <- adjust_week_pct_effect(svim_, maxwk)
svim |> 
  filter(!is.na(confirmed), confirmed > 0) |> 
  filter(vacc_cov == 0.9) |>
  ggplot(aes(x=week, y=dose_required_per_case_averted, group=week))+
  geom_violin(linewidth=0.6)+
  geom_jitter(width=0.1,size=0.5,color="steelblue",alpha=0.5)+
  labs(x="Week of protection",
       y="Vaccine doses\nrequired to avert a case") +
  theme_bw() -> p1

svim |>
  filter(!is.na(confirmed), confirmed > 0) |> 
  group_by(week, vacc_cov) |> 
  summarize(mean_dose_required_per_case = mean(dose_required_per_case_averted, na.rm=T),
            mean_dose_required_per_death = mean(dose_required_per_death_averted, na.rm=T)) -> d2

p2 <- ggplot(d2)+
  geom_contour_filled(aes(x=week+1, y=vacc_cov*10, 
                          z=mean_dose_required_per_case))+
  labs(x="Week of vaccination",
       y="Vaccine coverage rate (%)",
       fill="Case\naverted\n(%)")+
  # scale_x_continuous(breaks=seq(1,13,2),expand=c(0,0))+
  # scale_y_continuous(breaks=seq(10,90,20),expand=c(0,0))+
  scale_x_continuous(breaks=seq(1,13,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour()

p2 <- ggplot(d2)+
  geom_contour_filled(aes(x=vacc_week+1, y=vacc_cov*10, z=mean_pct_gain))+
  labs(x="Week of vaccination",
       y="Vaccine coverage rate (%)",
       fill="Percent\ndifference")+
  scale_x_continuous(breaks=seq(1,13,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour()


library(patchwork)
p <- (p1 | p2) + plot_annotation(tag_levels = "A")

ggsave(paste0("plots/vim_dynamic_all_", tstamp(), ".pdf"), p,
       width=fgw2*1.8, height=fgw2, units="mm")

```




docv analysis

```{r}
# data for the outbreaks affected by OCV
docv <- fread("inst/extdata/ocv_long_dataset.csv")
length(unique(docv$ID_outbreak))
sum(docv$OCV_during == TRUE,na.rm=T)
docv[docv$OCV_during == TRUE, ] -> ocv_during
ocv_during$Num_doses
ocv_during$Start_date

ocv_during |> 
  group_by(ID_outbreak) |> 
  summarize(start_date = Start_date) -> ocv_d

ts <- fread("inst/extdata/outbreak_data_ts.csv")
ts |>
  filter(ID_outbreak %in% ocv_during$ID_outbreak) |>
  group_by(ID_outbreak) |> 
  summarize(min_TL = min(TL)) -> ts_d
  
ocv_ts <- left_join(ocv_d, ts_d, by="ID_outbreak")
ocv_ts$start_date <- as.Date(ocv_ts$start_date)
ocv_ts$delay <- ocv_ts$start_date - as.Date(ocv_ts$min_TL) 
summary(as.integer(ocv_ts$delay))
as.integer(ocv_ts$delay)
```



Figure 2
1. Outbreak size versus duration
2. Weekly suspected cases versus week.

Some of the weekly cases are not integers 
(e.g., afr::ben::atlantique::abomeycalavi-2013-09-30-2013-12-08)

```{r}
tot <- sum(ts824_wk$sCh)
ts824_wk |>
  group_by(week) |> summarize(cases = sum(sCh)) -> ts824_wk_cum

ts824_wk_cum$cum_cases <- cumsum(ts824_wk_cum$cases)
ts824_wk_cum$case_rem <- tot - ts824_wk_cum$cum_cases
ts824_wk_cum$prop_rem <- ts824_wk_cum$case_rem/tot

p1 <- ggplot(ts824_wk_cum, aes(week, cases))+
  geom_col(fill="steelblue")+
  theme_bw()+
  labs(x="", y="Weekly suspected cases\nsummed across outbreaks")


# initial points where y=1 added
d <- data.frame(x=c(0,ts824_wk_cum$week+1), 
                 y=c(1, ts824_wk_cum$prop_rem))
mod <- y ~ exp(-r*x)
fit <- nls(mod, data=d, start=list(r=0.1))
summary(fit)
# plot(d$x,d$y)
# lines(d$x,fitted(fit),col=2)
dffit <- data.frame(x=d$x, y=predict(fit))

p2 <- ggplot(d, aes(x, y))+
  geom_col(fill="steelblue")+
  geom_line(data=dffit, aes(x,y), color="firebrick")+
  theme_bw()+
  labs(x="Epidemic week", y="Proportion of suspected cases\noccurred after the week")
# p2
# case protection
ts824_wk |>
  group_by(data_id ) |> mutate(totcase = sum(sCh)) -> ts824_wk_totcase

# p <- grid.arrange(p1, p2, nrow=2)


# Impact of vaccine

#Every week the number of cases that can be prevented decrease by 14.6% compared with the number the week before.


(1 - exp(-coef(fit)[3]))

ts824_wk |>
  group_by(id) |> 
  mutate(cum_case = cumsum(sCh),
         total_case = max(cum_case),
         case_rem = total_case - cum_case) -> ts824_wk_cumsum
vc <- 0.9
dve <- 0.65
ive <- 0.8

ts824_wk_cumsum |>
  group_by(id) |>
  mutate(case_protected = case_rem -
           (vc*case_rem*(1-dve)*(1-ive)+(1-vc)*case_rem*(1-ive)),
         vacc_impact = 100*case_protected/total_case) -> ts824_wk_cumsum_vacc
ts824_wk_cumsum_vacc

d <- ts824_wk_cumsum_vacc |> filter(week <= 15)
d$week <- as.numeric(d$week)
library(ggridges)

p3 <- ggplot(d)+
  geom_density_ridges(aes(y=as.factor(week),x=vacc_impact,
                          group=as.factor(week)))+
  geom_point(aes(y=week+1,x=vacc_impact), size=2,color="steelblue", alpha=0.3)+
  # scale_y_continuous(breaks=0:15,labels=0:15)+
  coord_flip()+
  labs(y="Week of protection",
       x="Case reduction (%)") +
  theme_bw()

library(patchwork)
p <- (p1|p2)/p3 + plot_annotation(tag_levels = "A")

ggsave(paste0("plots/case_protect_vacc_week_", tstamp(), ".pdf"), p,
       width=fgw2, height=fgw2, units="mm")
```



```{r}
ts824_wk |>
  group_by(id) |> 
  reframe(
    data_id = data_id,
    week = week,
    week_case = sCh,
    cum_case = cumsum(sCh),
    total_case = max(cum_case),
    case_rem = total_case - cum_case) -> ts824_wk_cumsum

ggplot(ts824_wk_cumsum, aes(week, case_rem, group=data_id))+
  geom_point(color="steelblue", alpha=0.3)+
  geom_line(color="steelblue", alpha=0.3)+
  # scale_x_log10()+
  # scale_y_log10()+
  theme_bw()+
  labs(x="Epidemic week", y="Case occurred after the week")

ts824_wk_cumsum

1 <- ggplot(data=dat)+
  geom_point(aes(duration, total_suspected_cases), color="black", alpha=0.3)+
  scale_x_log10()+
  scale_y_log10()+
  labs(x="Outbreak duration (weeks)", y="Total suspected cases")+
  theme_bw()

plt1 <- ggplot(dat)+
  geom_point(aes(duration, total_suspected_cases), color="black", alpha=0.3)+
  geom_point(data=dat_good, aes(duration, total_suspected_cases), color="steelblue", alpha=0.3)+
  geom_point(data=df, aes(outbk_dur, outbk_size), color="firebrick")+
  scale_x_log10(limits=c(1,max_dur))+
  scale_y_log10(limits=c(1,max_size))+
  geom_vline(data=df_dur,aes(xintercept=outbk_dur), linetype="dotted", color="firebrick")+
  geom_hline(data=df_size,aes(yintercept=outbk_size), linetype="dotted", color="firebrick")+
  labs(x="Outbreak duration (weeks)", y="Total suspected cases")+
  theme_bw()

# ggsave("plots/outbreak_data.pdf", plt, width=3.4*2, height=2.7*2, units="in")

# time series standardized by week
ts_wk  <- fread("outputs/ts_week_all.csv")
ts_wk_good <- ts_week[good_fit == 1]

plt2 <- ggplot()+
  geom_line(data=ts_wk, aes(week, sCh, group=data_id), 
            color="black", alpha=0.3)+
  geom_line(data=ts_wk_good, aes(week, sCh, group=data_id),
            color="steelblue", alpha=1.0)+
  labs(x="Epidemic week", y="Weekly suspected cases")+
  theme_bw()
plt2

library(patchwork)
p <- plt1 + plt2 + plot_annotation(tag_levels = "A")

ggsave(paste0("plots/ts_outbk_goodfits_", tstamp(), ".pdf"), p,
       width=fgw2, height=fgw2*0.8, units="mm")
```


### Outbreak duration, size, and time series

Select good fits

```{r}
# dat0 <- fread("julia/outputs/data_fit_eval_20240323.csv")
# fit_eval was modified to include R^2, RMSE, normalzied RMSE, AIC, and BIC
fits <- fread("outputs/data_fit_eval_20240409.csv")
# nrow(d)
# View(d)
# names(d)
sprintf("Total %d outbreaks", nrow(fits))
fits |> filter(PARAM_GLOBAL == 1) -> d1
sprintf("Global maxima in maximum likelihood estimation: %d outbreaks", nrow(d1)) #
# parameter estimates are >5% apart from the bounds
fits |> filter(PARAM_GLOBAL == 1, PARAM_SANITY_5pct == 1) -> d2
sprintf("Parameter estimation successful in %d outbreaks", nrow(d2))

fits |> filter(PARAM_GLOBAL == 1, PARAM_SANITY_5pct == 1, FIT_SANITY == 1) -> d2
sprintf("Parameter estimation successful in %d outbreaks", nrow(d2))

fits |> filter(PARAM_GLOBAL == 1, 
            PARAM_SANITY_5pct == 1, 
            FIT_SANITY == 1,
            R2_adj >= 0.5) -> good_fits
sprintf("Parameter estimation successful in %d outbreaks", nrow(good_fits))
```

1. Outbreak size versus duration
2. Weekly suspected cases versus week.
```{r}
# in case you need to plot the subset, good_fits from the fits code block
dat_good <- dat[dat$ID %in% good_fits$ID, ]
# data on the outbreaks that have been studies using models so far
library(googlesheets4)
#Read google sheets data into R
outbks_modeled <- read_sheet('https://docs.google.com/spreadsheets/d/1ZN6M4tx21-8BSRMZ692MOU3M5yDYK7bLqRbYGVOZGtQ/edit#gid=959832095')
dur0 <- as.integer(unlist(outbks_modeled$`Outbreak duration (days)`))
size0 <- as.integer(unlist(outbks_modeled$`Outbreak size (num)`))
# create dataframe with duration in terms of a week
df <- data.frame(outbk_dur=dur0/7, outbk_size=size0)
max_dur <- max(df$outbk_dur, na.rm=T)
max_size <- max(df$outbk_size, na.rm=T)
df_dur <- df[is.na(df$outbk_size),]
df_size <- df[is.na(df$outbk_dur),]
plt1 <- ggplot(dat)+
  geom_point(aes(duration, total_suspected_cases), color="black", alpha=0.3)+
  geom_point(data=dat_good, aes(duration, total_suspected_cases), color="steelblue", alpha=0.3)+
  geom_point(data=df, aes(outbk_dur, outbk_size), color="firebrick")+
  scale_x_log10(limits=c(1,max_dur))+
  scale_y_log10(limits=c(1,max_size))+
  geom_vline(data=df_dur,aes(xintercept=outbk_dur), linetype="dotted", color="firebrick")+
  geom_hline(data=df_size,aes(yintercept=outbk_size), linetype="dotted", color="firebrick")+
  labs(x="Outbreak duration (weeks)", y="Total suspected cases")+
  theme_bw()

# ggsave("plots/outbreak_data.pdf", plt, width=3.4*2, height=2.7*2, units="in")

# time series standardized by week
ts_wk  <- fread("outputs/ts_week_all.csv")
ts_wk_good <- ts_week[good_fit == 1]

plt2 <- ggplot()+
  geom_line(data=ts_wk, aes(week, sCh, group=data_id), 
            color="black", alpha=0.3)+
  geom_line(data=ts_wk_good, aes(week, sCh, group=data_id),
            color="steelblue", alpha=1.0)+
  labs(x="Epidemic week", y="Weekly suspected cases")+
  theme_bw()
plt2

library(patchwork)
p <- plt1 + plt2 + plot_annotation(tag_levels = "A")

ggsave(paste0("plots/ts_outbk_goodfits_", tstamp(), ".pdf"), p,
       width=fgw2, height=fgw2*0.8, units="mm")
```

### Weekly suspected cases - consolidated

```{r}
ts_wk  <- fread("outputs/ts_week_all.csv")
ts_wk_good <- ts_week[good_fit == 1]

ggplot(ts_wk, aes(week, sCh, group=data_id)) +
  geom_area(fill="black",alpha=0.3) + 
  geom_area(data=ts_wk_good, 
             aes(week, sCh, group=data_id), fill="steelblue")+
  theme_bw()+
  labs(x="Epidemic week", y="Weekly suspected cases")


ts_wk |> 
  group_by(week) |> 
  summarize(week=week,
            sCh_total=sum(sCh)) -> ts_wk_cum 

sum(ts_wk_cum$sCh_total)
# 252,002,150 cases

ggplot(ts_wk_cum, aes(week, sCh_total+0.1))+
  geom_col(fill="steelblue")+
  scale_y_continuous()+
  theme_bw()+
  labs(x="Epidemic week", y="Weekly suspected cases")


ggplot(ts_wk, aes(date, sCh, group=data_id)) +
  geom_col(aes(fill=data_id),alpha=0.3)+
  geom_col(fill="black",alpha=0.3) + 
  geom_area(data=ts_wk_good, 
             aes(week, sCh, group=data_id), fill="steelblue")+
  theme_bw()+
  labs(x="Epidemic week", y="Weekly suspected cases")         
```


We applied a simple static model to compute the number of cases that can be
averted 

### Time series comparison between model (lines) and data (points)
```{r}
d = fread("julia\\outputs\\fit_inc_s_20240330.csv")
head(d)
d$date <- as.Date(d$date, format="%m/%d/%Y") 
d$data_id <- as.factor(d$data_id)
d$ID_label <- as.factor(paste0("ID = ", d$data_id))
# d$loc_short <- sapply(d$location, function(x) gsub("afr::","",x))

d |> 
  group_by(ID_label) |>
  pivot_longer(cols=contains("inc_"), 
               names_to="model_var", 
               values_to="incidence") -> dlong

cols = dput(names(d)[4:9])
# c("inc_data", "s0001_inc_novacc", "s0026_inc_novacc", "s0050_inc_novacc", 
# "s0074_inc_novacc", "s0099_inc_novacc")

idx_all <- unique(d$data_id)
# ----------------------------------------------------------------
# in case if we want the good fits only
# d2 from the "fits" code block
idx_all <- good_fits$ID
#-----------------------------------------------------------------
iend <- round(length(idx_all)/15) + 1

for (i in 1:iend) {
  cat("i =", i, "\n")
  id0 <- 1 + (i-1)*15
  id1 <- 15 + (i-1)*15
  ids = idx_all[id0:id1] 
  
  d1 <- filter(dlong, data_id %in% ids, model_var %in% cols[c(2)])
  d2 <- filter(dlong, data_id %in% ids, model_var %in% cols[c(1)])
  # labelling
  # d1 |>
  #   group_by(loc_short) |>
  #   summarize(data_id = data_id[1],
  #             loc_short = loc_short[1],
  #             label = paste0("ID = ", data_id[1]),
  #             date = date[length(date)],
  #             incidence = max(incidence)) -> d3
  
  ggplot(d1, aes(date, incidence, group=data_id)) +
    geom_line(color = "steelblue") + 
    geom_point(data=d2,
               aes(date, incidence, group=data_id), color="firebrick") +
    facet_wrap(~ID_label, scales=c("free"), ncol=3) +
    theme_bw() + 
    theme(legend.position = "bottom") + 
    labs(x="", y="", color="") -> plt
  
  # plt + geom_text(data=d3, aes(date, incidence, label = label),
  #                 hjust=1, vjust=1) -> plt
  
  # A4 Size (8.3" x 11.7")
  # ggsave(paste0("plots/fit_data_model_", id0, "_", id1, ".pdf"), plt,
  #        width=7.2, height=7.2*1.4, units="in")
  ggsave(paste0("plots/good_fits_data_model_", id0, "_", id1, ".pdf"), plt,
         width=7.2, height=7.2*1.4, units="in")
}

```

Goodness of fit metrics

```{r, eval=F}
aic <- function(LL, npar, k=2){
  -2*LL + k*npar
}
bic <- function(LL, npar, nobs){
  -2*LL + log(nobs)*npar
}
adj_r2 <- function(r2, nobs, npar) {
   1 - ((1-r2)*(nobs-1)/(nobs-npar-1))
}

library(dplyr)
library(hydroGOF)
d = fread("julia\\outputs\\fit_inc_s_20240330.csv")
d |> group_by(data_id) |>
  summarize(nobs = length(inc_data),
            nse = NSE(sim=s0001_inc_novacc, obs=inc_data),
            R2_adj = adj_r2(r2=nse, nobs=length(inc_data), npar=4),
            pbias = pbias(sim=s0001_inc_novacc, obs=inc_data),
            rmse = rmse(sim=s0001_inc_novacc, obs=inc_data),
            nrmse = nrmse(sim=s0001_inc_novacc, obs=inc_data)) -> gof

names(gof) <- c("ID", names(gof)[2:ncol(gof)])

d <- fread("julia/outputs/data_fit_eval_20240323.csv")
dgof <- left_join(d, gof, by="ID")
dgof$aic <- aic(LL=dgof$f1_obj, npar=4, k=2)
dgof$bic <- bic(LL=dgof$f1_obj, npar=4, nobs=dgof$nobs)

# fwrite(dgof, "outputs/data_fit_eval_20240402.csv")
# following was taken from sim_analysis.jl file
# lower = [0.0001, 0.0001, 1.1000, 0.0001];
# upper = [0.9999, 0.9999, 30.0000, 0.9999];
lb <- c(s0=0.0001, i0=0.0001, R0=1.1000, n0=0.0001)
ub <- c(s0=0.9999, i0=0.9999, R0=30.0000, n0=0.9999)
dat <- fread("outputs/data_fit_eval_20240402.csv")
dat$bounds_safe <- 1
#5% off from the bounds to ensure optimization worked
cut <- 0.01
for (i in 1:nrow(dat)) {
  if (dat$f1_s0[i] > (1-cut)*ub[["s0"]] | dat$f1_i0[i] > (1-cut)*ub[["i0"]] |
      dat$f1_R0[i] > (1-cut)*ub[["R0"]] | dat$f1_n0[i] > (1-cut)*ub[["n0"]] |
      dat$f1_s0[i] < (1+cut)*lb[["s0"]] | dat$f1_i0[i] < (1+cut)*lb[["i0"]] |
      dat$f1_R0[i] < (1+cut)*lb[["R0"]] | dat$f1_n0[i] < (1+cut)*lb[["n0"]] ) {
    dat$bounds_safe[i] <- 0
  } 
  
}
sum(dat$bounds_safe == 0)
dat <- dat[bounds_safe==1]
fwrite(dat, paste0("outputs/data_fit_eval_", tstamp(), ".csv"))
```



### Parameter estimates from good fits

```{r}
# View(good_fits)
d <- good_fits[,.(s0=f1_s0,i0=f1_i0,R0=f1_R0,n0=f1_n0)]
summary(d)
library(GGally)
# names(d) 
# ggparcoord(d, columns = 1:4) 
plt <- ggpairs(d, columnLabels=c("s[0]","i[0]","R[0]","N[eff]"),
               labeller="label_parsed")

# ggsave(paste("plots/param_pairs",tstamp(),".pdf"), plt, width=fgw2, height=fgw2, units="mm")
```

Combine the impact of vaccine
```{r, eval=FALSE}
vacc_cov = 4:9
vacc_week = 3:12
vacc_impact_list <- vector('list', length(vacc_cov))
for (i in 1:length(vacc_cov)) {
  vc <- vacc_cov[i]
  cols = paste0("inc_C", vc, "T", vacc_week)
  d = read.csv(paste0("julia\\outputs\\vacc_4p_cov", vc, "_20240311T13.csv"))
  # NA values are replaced with baseline incidence values when no vaccines were   introduced. The decision was made because NA values were created when the vaccination was not implemented because outbreak ended earlier than the vaccination start + duration
  
  for (col in cols) {
    nas = is.na(d[,col])
    d[nas, col] = d[nas, "inc_novacc"]
  }
  d |>
    group_by(data_id) |>
    summarise(data_id = data_id[1],
              ID_outbreak = ID_outbreak[1],
              across(contains("inc"), sum)) -> d2
  
  col_id = 1:ncol(d2)
  id_iter <- col_id[grepl("inc_C", names(d2))]
  d3 <- d2
  for (k in id_iter) {
    d3[,k] = 100 * (d2[,"inc_novacc"] - d2[,k]) / d2[,"inc_novacc"]
  }
  
  d4 = data.frame(data_id = rep(d3$data_id, length(vacc_week)),
                  ID_outbreak = rep(d3$ID_outbreak, length(vacc_week)),
                  vacc_eff = unlist(d3[, id_iter]), 
                  vacc_cov = vc,
                  vacc_week = rep(vacc_week, each=nrow(d3)))
 
  
  vacc_impact_list[[i]] <- d4
}

vacc_impact <- do.call('rbind', vacc_impact_list)
# fwrite(vacc_impact, "outputs/vacc_impact.csv")
```

### Vaccine impact by coverage rate and timing of vaccination
 
Individual vaccine impact by duration

```{r}
dat <- fread("outputs/data_fit_eval_20240409.csv")
good_fits <- dat[PARAM_GLOBAL == 1 & PARAM_SANITY_5pct == 1 & 
                   FIT_SANITY == 1 & R2_adj >= 0.5]
vacc_impact <- fread("outputs/vacc_impact.csv")
# cntries <- unique(dat$country)

vc <- 7 # target vaccine coverage, 70%
vacc_impact |>
  filter(data_id %in% good_fits$ID & vacc_cov == vc) |> 
  group_by(data_id, vacc_week) |> 
  summarize(vacc_cov = vacc_cov[1],
            vacc_week = vacc_week[1],
            vacc_impact = vacc_eff) -> d

library(ggridges)
plt1 <- ggplot(d)+
  # geom_violin(scale="width", linewidth=1) + 
  # geom_density_ridges(aes(y=as.factor(vacc_week), x=vacc_impact))+
  geom_density_ridges(aes(y=vacc_week,x=vacc_impact,group=vacc_week))+
  # geom_line(aes(x=vacc_week, y=vacc_impact, group=data_id), 
            # color="steelblue", alpha=0.3)+
  geom_point(aes(y=vacc_week,x=vacc_impact),color="steelblue", alpha=0.3)+
  scale_y_continuous(breaks=3:12,labels=3:12)+
  coord_flip()+
  labs(y="Week of vaccination",
       x="Case reduction (%)") +
  theme_minimal()+
  theme(legend.position = "bottom",  
        text = element_text(size=16),
        axis.text = element_text(size=13),
        legend.text = element_text(size=13))


# ggplot(d, aes(x=vacc_week, y=vacc_impact, group=data_id))+
#   geom_line(color="black", alpha=0.3)+
#   geom_point(color="black", alpha=0.3)+
#   labs(x="Epidemic week",
#        y="% case reduction") +
#   # scale_x_continuous(breaks=seq(3,12,2))+
#   # scale_y_continuous(breaks=seq(30,90,10))+
#   theme(legend.position = "bottom",  
#         text = element_text(size=16),
#         axis.text = element_text(size=13),
#         legend.text = element_text(size=13))+
#   theme_bw()
  
# d$vacc_week <- factor(d$vacc_week, ordered=TRUE, levels=3:12)
# 
# library(ggridges)
# ggplot(d, aes(x=vacc_impact, y=vacc_week)) +
#   geom_density_ridges(scale = 4) + 
#   scale_y_discrete(expand=c(0,0)) +
#   scale_x_continuous(expand=c(0,0)) +
#   coord_cartesian(clip = "off") +
#   labs(x="% case duction",
#        y="Epidemic week at vaccination") +
#   theme_bw() -> plt2
# plt2
```

Mean vaccine impact

```{r}
dat <- fread("outputs/data_fit_eval_20240409.csv")
good_fits <- dat[PARAM_GLOBAL == 1 & PARAM_SANITY_5pct == 1 & 
                   FIT_SANITY == 1 & R2_adj >= 0.5]

vacc_impact <- fread("outputs/vacc_impact.csv")

vacc_impact |>
  filter(data_id %in% good_fits$ID) |> 
  group_by(vacc_cov, vacc_week) |> 
  summarize(vacc_cov = vacc_cov[1],
            vacc_week = vacc_week[1],
            vacc_eff = mean(vacc_eff)) -> d
  
plt2 <- ggplot(d, aes(x=vacc_week, y=vacc_cov*10, fill=vacc_eff))+
  geom_raster()+
  geom_contour(aes(z=vacc_eff), breaks=4.99, 
               color="firebrick", linetype="solid")+
  geom_vline(xintercept = 59/7, color="firebrick", linetype="dotted")+
  # geom_contour(aes(z=vacc_eff), breaks=0.99,
  #              color="firebrick",linetype="dotted")+
  # metR::geom_contour2(aes(z=vacc_eff, label=stat(level))) +
  theme_minimal()+
  labs(x="Week of vaccination",
       y="Vaccine coverage (%)",
       fill="% case reduction") +
  scale_x_continuous(breaks=seq(3,12,2))+
  scale_y_continuous(breaks=seq(30,90,10))+
  # scale_fill_continuous(limits=c(0,100))+
  # theme(legend.position = "bottom")

  theme(legend.position = "bottom",
        text = element_text(size=16),
        axis.text = element_text(size=13),
        legend.text = element_text(size=13))
plt2 

library(patchwork)
p <- plt2 + plt1 + plot_annotation(tag_levels = "A")

# ggsave(paste0("plots/vacc_impact_", tstamp(), ".pdf"), p,
#        width=fgw2*1.4, height=fgw2, units="mm")
```


Filter the impact estimates based on various conditions
1. Select good fits
2. Categorize by attack rate, duration, and basic reproduction number

```{r}
dat <- fread("julia/outputs/data_fit_init_vals_20240311.csv")
vacc_impact <- fread("outputs/vacc_impact.csv")
cntries <- unique(dat$country)
dat1 <-  dat |> filter(same == 1) # fits with 4 independent random seeds
dat2 <-  dat1 |> filter(duration > 7) # fits with 4 independent random seeds

cases_susp <- quantile(dat2$total_suspected_cases, 
                       probs=c(0.025, 0.05, 0.5, 0.75, 0.95, 0.975))

dat1_l5 <- dat1 |> filter(total_suspected_cases < cases_susp[2])
dat1_u5 <- dat1 |> filter(total_suspected_cases > cases_susp[5])

vacc_impact |>
  filter(data_id %in% dat1_u5$ID) |> 
  group_by(vacc_cov, vacc_week) |> 
  summarize(vacc_cov = vacc_cov[1],
            vacc_week = vacc_week[1],
            vacc_eff = mean(vacc_eff)) -> d
  
plt <- ggplot(d, aes(x=vacc_week, y=vacc_cov*10, fill=vacc_eff))+
  geom_raster()+
  geom_contour(aes(z=vacc_eff), breaks=4.9, color="red")+
  # metR::geom_contour2(aes(z=vacc_eff, label=stat(level))) +
  theme_minimal()+
  labs(title=paste0("Impact of the vaccine: Upper 5% in suspected cases"),
       x="Epidemic week",
       y="Vaccine coverage (%)",
       fill="% reduction") +
  scale_x_continuous(breaks=seq(3,12,2))+
  scale_y_continuous(breaks=seq(30,90,10))+
  # scale_fill_continuous(limits=c(0,100))+
  theme(legend.position = "bottom",  
        text = element_text(size=16),
        axis.text = element_text(size=13),
        legend.text = element_text(size=13))
    
ggsave(paste0("plots/vacc_impact_u5_cases_susp_", tstamp(), ".pdf"), 
         plt, width = 3.4*2, height = 2.7*2, units = "in")
```

Mean vaccine impact by duration

```{r}
dat <- fread("julia/outputs/data_fit_init_vals_20240311.csv")
vacc_impact <- fread("outputs/vacc_impact.csv")
cntries <- unique(dat$country)
dat1 <-  dat |> filter(same == 1) # fits with 4 independent random seeds
dat2 <-  dat1 |> filter(duration > 7) # fits with 4 independent random seeds

vacc_impact |>
  filter(data_id %in% dat2$ID) |> 
  group_by(vacc_cov, vacc_week) |> 
  summarize(vacc_cov = vacc_cov[1],
            vacc_week = vacc_week[1],
            vacc_eff = mean(vacc_eff)) -> d

plt <- 
  ggplot(d, aes(x=vacc_week, y=vacc_eff, group=vacc_cov, color=vacc_cov))+
  geom_line()+
  geom_point()+
  labs(title=paste0("Impact of the vaccine across vaccination timing"),
       x="Epidemic week",
       y="Vaccine coverage (%)") +
  scale_x_continuous(breaks=seq(3,12,2))+
  scale_y_continuous(breaks=seq(30,90,10))+
  # scale_fill_continuous(limits=c(0,100))+
  theme(legend.position = "bottom",  
        text = element_text(size=16),
        axis.text = element_text(size=13),
        legend.text = element_text(size=13))
plt
```

### Case-triggered vaccination








## Vaccine impact

```{r}
d = read.csv("julia\\outputs\\vacc_4p_cov4_20240311T13.csv")
cols = paste0("inc_C4T", 3:12)
# NA values are replaced with baseline incidence values when no vaccines were introduced. The decision was made because NA values were created when the vaccination was not implemented because outbreak ended earlier than the vaccination start + duration

for (col in cols) {
  nas = is.na(d[,col])
  d[nas, col] = d[nas, "inc_novacc"]
}
sum(is.na(d))
d |>
  group_by(data_id) |>
  summarise(across(contains("inc"), sum)) -> d2
   
d3 <- d2
for (k in 2:ncol(d2)) {
  d3[,k] = 100 * (d2[,"inc_novacc"] - d2[,k]) / d2[,"inc_novacc"]
}

d4 = data.frame(vacc_week = rep(3:12, each=nrow(d2)))
d4$vacc_eff = unlist(d3[,3:ncol(d3)])

ggplot(d4) +
  geom_violin(aes(as.factor(vacc_week), vacc_eff), fill="steelblue", width=1.4)+
  geom_boxplot(aes(as.factor(vacc_week), vacc_eff), fill="grey", width=0.1, color="grey", alpha=0.6)+
  theme_bw() + 
  theme(legend.position = "bottom") + 
  labs(title="Percent reduction in cumulative incidence by vaccination week", x="Vaccination week", y="Percent reduction")
```

## 3D plotting using plotly
```{r}
vacc_impact <- fread("outputs/vacc_impact.csv")
vacc_impact |>
    group_by(vacc_cov, vacc_week) |> 
    summarize(vacc_cov = vacc_cov[1],
              vacc_week = vacc_week[1],
              mean_eff = mean(vacc_eff),
              sd_eff = sd(vacc_eff), 
              cv_eff =  sd(vacc_eff)/mean(vacc_eff)) -> d

library(plotly)
fig <- plot_ly(d, x=~vacc_week, y=~10*vacc_cov, z=~mean_eff,
               size=~sd_eff, marker=list(size=5))
# fig <- plot_ly(mtcars, x = ~wt, y = ~hp, z = ~qsec, color = ~am, colors = c('#BF382A', '#0C4B8E'))
fig <- fig %>% add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Vaccination week'),
                     yaxis = list(title = 'Vaccine coverage (%)'),
                     zaxis = list(title = 'CI reduction (%)')))

fig


##########################################################################
# 3d histogram codes copided from https://stackoverflow.com/questions/64479652/how-to-make-a-3d-histogram-with-plotly-in-r
# The matrix with frequencies from a 3 x 4 cross table
z_mtx <- cbind(c(2,4,6,5), c(1,5,9,6), c(2,4,2,3))
# Define a function to add 3D bars
add_3Dbar <- function(p, x,y,z, width=0.4) {
   w <- width
   add_trace(p, type="mesh3d",
     x = c(x-w, x-w, x+w, x+w, x-w, x-w, x+w, x+w),
     y = c(y-w, y+w, y+w, y-w, y-w, y+w, y+w, y-w),
     z = c(0, 0, 0, 0, z, z, z, z),
     i = c(7, 0, 0, 0, 4, 4, 2, 6, 4, 0, 3, 7),
     j = c(3, 4, 1, 2, 5, 6, 5, 5, 0, 1, 2, 2),
     k = c(0, 7, 2, 3, 6, 7, 1, 2, 5, 5, 7, 6),
     facecolor = rep(toRGB(viridisLite::inferno(6)), each = 2)) 
}

# Draw the 3D histogram
fig <- plot_ly()
for (k1 in 1:nrow(z_mtx)) {
  for (k2 in 1:ncol(z_mtx)) {
     fig <- fig %>% add_3Dbar(k1,k2,z_mtx[k1,k2])
  }
}
fig 
```

Compute information criteria
```{r}
AIC = function(loglik, k){
  2*k - 2*loglik
}

d1 <- read.csv("julia\\outputs\\fit_4p_fA005_20240215T1913.csv")
d2 <- read.csv("julia\\outputs\\fit_4p_fA050_20240216T0147.csv")
d3 <- read.csv("julia\\outputs\\fit_5p_fA005_20240216T1809.csv")
d4 <- read.csv("julia\\outputs\\fit_5p_fA005_20240216T1809.csv")

df = data.frame(data_id = d1$data_id, 
                P4fA005 = d1$objective,
                P4fA050 = d2$objective,
                P5fA005 = d3$objective,
                P5fA050 = d4$objective)
dflong = pivot_longer(df, cols=-(data_id), names_to="model", values_to="loglik")

```


## Vaccination campaign

```{r}
# library(CholeraOutbreakModel)
# devtools::load_all()

fits = readRDS("outputs/AR_over_3/fits_20230714T14.rds")
parlist = lapply(fits, function(x) x$fit$optim$bestmem)
pardf = as.data.frame(do.call('rbind', parlist))
names(pardf) = c("s0", "i0", "R0")
pardf = cbind(data.frame(ID=1:nrow(pardf)), pardf)

ids = c(4,6,8)
id = ids[1]

obs = fits[[id]]$OBS
parm_base = fits[[id]]$PARAMS
x = fits[[id]]$dat
# to reproduce the original plot
output_days = fits[[id]]$PARAMS$output_days
pop = fits[[id]]$PARAMS$population
  
p = fits[[id]]$fit$optim$bestmem
# generate the best fit simulation
parm = c(s0=p[[1]],
         i0=p[[2]],
         R0=p[[3]],
         prop_report=1)

parm_base$obs_length <- 100
parm_base$output_days <- 1

parm_base$tau <- 0.01 # vaccine impact requires a small time step
# vaccination-campaign associated
parm_base$campaign_1_start <- 1e6
inc = incidence(pars=parm, pars_baseline=parm_base)
m = rowSums(inc$inc) # also makes inc$inc into a vector
tot = length(m)

parm_base$campaign_1_start <- 30
inc_vacc = incidence(pars=parm, pars_baseline=parm_base)
mv = rowSums(inc_vacc$inc)

# png(filename = paste0("plots/AR_over_3/vacc_", tstamp(hour=T), "_", id, "_", ".png"))
plot(m, type="l", ylim=c(0,max(m)), xlab="",
       ylab="Suspected cases")
lines(mv, xlab="", lty=2, col=2)
# lines(mv, xlab="", lty=3, col=3)
# dev.off()
```

### Plot
```{r}
# ?data.table::frollsum
df <- data.frame(matrix(NA, nrow=52, ncol=10))
df$week=1:52
for(i in 1:10){
  parm <- c(R0=2, R0W=1, case_threshold=10, alpha=(i-1)/10)
  d <- daily_incidence(pars=parm)
  weekroll <- data.table::frollsum(d$CI, n=7)
  sq <- seq(7, length(weekroll), by=7)
  df[, i] <- weekroll[sq]
}

head(df, n=20)
names(df) <- c(paste0(seq(0,90,10)," %"), "Week")
df <- df[,c(seq(1,9,2),11)]
library(tidyr)
# dflong <- pivot_longer(df, cols=1:ncol(df), names_to = "Reduction")
dflong <- pivot_longer(df, cols=(!Week), names_to = "Reduction")
library(ggplot2)
mycolors <- my_discrete_colors

p <- ggplot(dflong) +
  geom_line(aes(Week, value, color=Reduction), linewidth=1)+
  labs(x="Week", y="Case")+
  scale_color_manual("Intervention", values=mycolors[1:10])+
  theme_classic()+
  theme(panel.grid.major = element_line(colour="grey95", linewidth=0.3),
    text = element_text(size=14), axis.text = element_text(size=14))
p

ggsave("plots/outbreak_intervention.png", p, width=3.4*2, height=2.7*2, unit="in")



# sum(d$CI)
cat("outbreak size:", round(sum(weekly)), "\n")
# df <- data.frame(week = 1:length(weekly), case = weekly)
# library(ggplot2)
# p <- ggplot(df) +
#   geom_col(aes(week, case), fill = "brown") + 
#   labs(x="Week", y="Case")+
#   theme_classic()+
#   theme(panel.grid.major = element_line(colour="grey95", linewidth=0.3),
#     text = element_text(size=14),
#         axis.text = element_text(size=14))
# p
# ggsave("plots/typical_outbreak.png", p, width=3.4*2, height=2.7*2, unit="in")

out <- PARAMETERS$model(PARAMETERS)
day_filter <- seq(1, by=round(1/PARAMETERS$tau), length.out=(PARAMETERS$ndays+1))
out <- out[day_filter, ]

```



### Sanity check

```{r}
devtools::load_all()
PARAMS = initialize_params(
  day1=0,
  prop_immune=0,
  prop_report=1,
  prop_eff_pop=1,
  output_state="CI",
  R0=2,
  output_days=1,
  obs_length=365,
  sigma = 0,
  ndays=365,
  tau=0.01)

# NOT WORKING BECAUSE OF VACC_COV INDEX
# PARAMS$model <- seiarw
# 
# paste("outbreak size (R0) =", final_epidemic_size(R0=PARAMS$R0))
# out <- PARAMS$model(PARAMS)
# paste("outbreak size (seirw_1-S) =", 1 - tail(out[["S"]],1)/PARAMS$population)
# paste("outbreak size (seirw_R) =", tail(out[["R"]],1)/PARAMS$population)
# sum(head(out[c("S","E","I","R")],1))
# sum(tail(out[c("S","E","I","R")],1))

# two age groups
PARAMS$model <- seiarw_2ag_erlang
paste("outbreak size (R0) =", final_epidemic_size(R0=PARAMS$R0))
out <- PARAMS$model(PARAMS)
grep("S", names(out))
paste("outbreak size (seirw_1-S) =", 1 - sum(tail(out[,grep("S", names(out))],1))/PARAMS$population)

paste("outbreak size (seirw_R) =", sum(tail(out[,grep("R", names(out))],1))/PARAMS$population)
sum(head(out[,grep("S|^E|^I|R", names(out))],1))
sum(tail(out[,grep("S|^E|^I|R", names(out))],1)) # ^E|^I to exclude CE and CI
```

### Sanity check 2
```{r}
devtools::load_all()
PARAMS = initialize_params(
  day1=0,
  s0=0.99,
  i0=0.01,
  prop_report=1,
  output_state="CI",
  R0=3,
  output_days=1,
  obs_length=365,
  sigma = 0,
  ndays=365,
  tau=0.01)

# for the single age group model
# paste("outbreak size (R0) =", final_epidemic_size(R0=PARAMS$R0))
# out <- PARAMS$model(PARAMS)
# paste("outbreak size (seirw_1-S) =", 1 - tail(out[["S"]],1)/PARAMS$population)
# paste("outbreak size (seirw_R) =", tail(out[["R"]],1)/PARAMS$population)
# sum(head(out[c("S","E","I","R")],1))
# sum(tail(out[c("S","E","I","R")],1))

# two age groups
PARAMS$susceptible + PARAMS$infectious + PARAMS$asymptomatic +  PARAMS$exposed + PARAMS$recovered

PARAMS$model <- seiarw_2ag_erlang
paste("outbreak size (R0) =", final_epidemic_size(R0=PARAMS$R0))
out <- PARAMS$model(PARAMS)
grep("S", names(out))

paste("outbreak size (S) =", 1 - sum(tail(out[,grep("S", names(out))],1))/PARAMS$population)
paste("outbreak size (seirw_R) =", sum(tail(out[,grep("R", names(out))],1))/PARAMS$population)

sum(head(out[,grep("S|^E|^I|^A|^R", names(out))],1))
sum(tail(out[,grep("S|^E|^I|^A|^R", names(out))],1)) # ^E|^I to exclude CE and CI

PARAMS$model <- seiarw_2ag_erlang_leaky
out <- PARAMS$model(PARAMS)
grep("S", names(out))

paste("outbreak size (S) =", 1 - sum(tail(out[,grep("S", names(out))],1))/PARAMS$population)
paste("outbreak size (seirw_R) =", sum(tail(out[,grep("R", names(out))],1))/PARAMS$population)
sum(head(out[,grep("S|^E|^I|^A|^R", names(out))],1))
sum(tail(out[,grep("S|^E|^I|^A|^R", names(out))],1)) 
```


### Two age group vector check
```{r}
devtools::load_all()
PARAMS <- initialize_params()
seiarw_check(PARAMS)
out = seiarw_2ag_erlang_leaky(PARAMS)
head(out)
# how do I insert this to C++ function such that it automatically 
out = as.data.frame(do.call("cbind", out))
names(out) = c("time","S.1","S.2",
                             "E1.1","E1.2","E2.1","E2.2",
                             "I1.1","I1.2","I2.1","I2.2",
                             "A1.1","A1.2","A2.1","A2.2",
                             "R1.1","R1.2","R2.1","R2.2",
                             "O1.1","O1.2","O2.1","O2.2",
                             "T1.1","T1.2","T2.1","T2.2",
                             "CE.1","CE.2","CI.1","CI.2",
                             "OE1.1","OE1.2","OE2.1","OE2.2",
                             "OI1.1","OI1.2","OI2.1","OI2.2",
                             "OA1.1","OA1.2","OA2.1","OA2.2",
                             "OR1.1","OR1.2","OR2.1","OR2.2",
                             "COE.1","COE.2","COI.1","COI.2",
                             "TE1.1","TE1.2","TE2.1","TE2.2",
                             "TI1.1","TI1.2","TI2.1","TI2.2",
                             "TA1.1","TA1.2","TA2.1","TA2.2",
                             "TR1.1","TR1.2","TR2.1","TR2.2",
                             "CTE.1","CTE.2","CTI.1","CTI.2")

sum(head(out[,grep("S|^E|^I|^A|^R", names(out))],1))
sum(tail(out[,grep("S|^E|^I|^A|^R", names(out))],1))

paste("outbreak size (S) =", 1 - sum(tail(out[,grep("S", names(out))],1))/PARAMS$population)
paste("outbreak size (seirw_R) =", sum(tail(out[,grep("R", names(out))],1))/PARAMS$population)

# check vaccine coverage = O + T
# reduction in incidence in
```

#### Select outbreak for fitting
1. No OCV campaign has been implemented
2. Lasted longer than 7 weeks, which is the median time before the OCV campaign ..  [Pezzoli 2020]
The median time from the event (i.e., first laboratory confirmation of cholera or occurrence of humanitarian emergency) to the receipt of the official OCV request: 26 d (range 12206 days).
The median of time from receipt to request approval: 5 d (range 0180 days)
The median time from approval to arrival of vaccines in the country: 13 d (range, 424 days)
The median time from arrival to the start of vaccination: 15 d (range, - 2 to 87 days)
2. Admin 2 or 3 considering uniform mixing assumption in the model
Attack rate > 3%* (the model could fit outbreaks with AR over ~0.7.  For those with lower attack rate,  it appears that we need to assume that not the entire population in the admin is subject to being transmitted)
The average number of case for the first few weeks is greater than the threshold X


```{r}
# population size data set
dat = data.table::fread("https://raw.githubusercontent.com/kimfinale/cholera_outbreaks_ssa/main/reference_data/outbreak_data.csv")
# to be matched with the output of the ext_outbreak_st function
dat$start_date <- as.Date(dat$start_date, format="%m/%d/%Y")
dat$end_date <- as.Date(dat$end_date, format="%m/%d/%Y")
dat$outbreak_ID  = paste0(dat$location, "-", dat$start_date, "-",
                       dat$end_date)

dat$my_attack_rate = 100 * dat$total_suspected_cases /dat$population

# include the outbreak occurred in admin 2 or smaller
# time series data set for fitting the outbreak
dat_ts = readRDS("inst/extdata/outbreak_data_ts.rds")

# extract ids for the outbreaks not affected by OCV campaigns
d = data.table::fread("inst/extdata/outbreak_data_summary.csv")
d$ID_outbreak = paste0(d$location, "-", d$start_date, "-",
                       d$end_date)
# data for the outbreaks affected by OCV
docv = data.table::fread("inst/extdata/ocv_long_dataset.csv")
# remove ocv outbreaks using IDs
dnoocv = d[!(d$ID_outbreak %in% docv$ID_outbreak),]
# 1. outbreaks in a small region which we may be able to model using a simple ODE
dnoocv <- dnoocv[dnoocv$spatial_scale %in% c("admin2","admin3"),]
# 2. outbreaks that are longer than the median time required for the vaccination to be implemented. This duration is the time required for the first realization of the outbreak plus 7 weeks
# Pezzoli 2020
min_delay_vacc = 3 # weeks min = 12 d + 0 d + 4d + 0
med_delay_vacc = 7 # weeks
dnoocv <- dnoocv[dnoocv$duration > med_delay_vacc,]

# 3. outbreaks that are attack rates are greater than certain threshold, ie, severe enough to require an intervention
min_suscpected_case <- 100
dnoocv <- dnoocv[dnoocv$total_suspected_cases > min_suscpected_case,]
# 4. Outbreaks that happend in the most recent year
# library(dplyr)
# dnoocv = left_join(dnoocv, dat_s0[,c("location","max_start_date","prop_immune")], by ="location")
# dnoocv <- filter(dnoocv, start_date == max_start_date)

# another variable just for convenience
ids_outbks_no_ocv <- dnoocv$ID_outbreak
fitlist <- vector('list', length(ids_outbks_no_ocv))
ppath <- "G:/My Drive/Projects/VIMC/VIMC 2.0/CholeraOutbreakModel/"
for (i in 1:length(ids_outbks_no_ocv)) {
# for (i in 1:124) {
  id <- ids_outbks_no_ocv[i]
  st <- ext_outbreak_st(id)
  pop = dat[(location == st$location &
            start_date == st$start_date &
            end_date == st$end_date), population]

  cat("i =", i, ", id =", id, ", pop =", pop, "\n")

  if (!is.numeric(pop) | !length(pop) > 0) {
    warning("Population size is missing and the simulation is skipped.")
  }

  x = dat_ts[dat_ts$ID_outbreak == id,]
  obs_len = nrow(x)
  output_days = get_output_days(x)

  OBS <- x$sCh
    # plot best fit
  png(filename = paste0(ppath,"plots/s0i0R0/", tstamp(hour=T), "_", i, "_",
                          gsub("::","_", st$location), ".png"))

  date = rev(seq(x$TL[length(x$TL)],
                 by=paste0("-", output_days, " days"), length.out=length(OBS)))
  plot(date, OBS, ylim=c(0,max(OBS)),
       main=ids_outbks_no_ocv[i], xlab="",
       ylab="Suspected cases", col=2)
  text(date[length(date)], 0.95*max(OBS), sprintf("ID=%s, Pop=%.0f", i, round(pop)), pos=2)
  text(date[length(date)], 0.85*max(OBS), sprintf("CI=%.0f, AR=%.2f", sum(OBS), 100*sum(OBS)/pop), pos=2)
  dev.off()
}
```




```{r}
#incidence rate in the no-vaccine group
# 1 Trach DD, Clemens JD, Ke NT, et al. Field trial of a locally produced, killed, oral cholera vaccine in Vietnam. Lancet 1997; 349: 2315.
ir1 <- c(92/67058*100000)
ir0 <- c(0.41*365)
ir1 <- c(0.25)
```

Epidemiologic features of the outbreaks where

```{r}
# 1. outbreaks in a small region which we may be able to model using a simple ODE
dnoocv <- dnoocv[dnoocv$spatial_scale %in% c("admin2","admin3"),]
# 2. outbreaks that are longer than the median time required for the vaccination to be implemented. This duration is the time required for the first realization of the outbreak plus 7 weeks
# Pezzoli 2020
min_delay_vacc = 3 # weeks min = 12 d + 0 d + 4d + 0
med_delay_vacc = 7 # weeks
dnoocv <- dnoocv[dnoocv$duration > med_delay_vacc,]

# 3. outbreaks that are attack rates are greater than certain threshold, ie, severe enough to require an intervention
# simple ODE model
dnoocv <- dnoocv[dnoocv$total_suspected_cases > 100,]
# 4. Outbreaks that happend in the most recent year
library(dplyr)
dnoocv = left_join(dnoocv, dat_s0[,c("location","max_start_date","prop_immune")], by ="location")
dnoocv <- filter(dnoocv, start_date == max_start_date)
# another variable just for convenience
ids_outbks_no_ocv <- dnoocv$ID_outbreak
fitlist <- vector('list', length(ids_outbks_no_ocv))
ppath <- "G:/My Drive/Projects/VIMC/VIMC 2.0/CholeraOutbreakModel/"
for (i in 1:length(ids_outbks_no_ocv)) {
# for (i in 1:124) {
  id <- ids_outbks_no_ocv[i]
  st <- ext_outbreak_st(id)
  pop = dat[(location == st$location & 
            start_date == st$start_date &  
            end_date == st$end_date), population]
  
  cat("i =", i, ", id =", id, ", pop =", pop, "\n")
  
  if (!is.numeric(pop) | !length(pop) > 0) {
    warning("Population size is missing and the simulation is skipped.")
  }
  
  x = dat_ts[dat_ts$ID_outbreak == id,]
  obs_len = nrow(x)
  output_days = get_output_days(x)
  
  OBS <- x$sCh    
    # plot best fit
  png(filename = paste0(ppath,"plots/s0i0R0/", tstamp(hour=T), "_", i, "_", 
                          gsub("::","_", st$location), ".png"))
  
  date = rev(seq(x$TL[length(x$TL)], 
                 by=paste0("-", output_days, " days"), length.out=length(OBS)))
  plot(date, OBS, ylim=c(0,max(OBS)), 
       main=ids_outbks_no_ocv[i], xlab="",
       ylab="Suspected cases", col=2)
  text(date[length(date)], 0.95*max(OBS), sprintf("ID=%s, Pop=%.0f", i, round(pop)), pos=2)
  text(date[length(date)], 0.85*max(OBS), sprintf("CI=%.0f, AR=%.2f", sum(OBS), 100*sum(OBS)/pop), pos=2)  
  dev.off()
}
```

Plot the impact by country
```{r}
dat <- fread("julia/outputs/data_fit_init_vals_20240311.csv")
vacc_impact <- fread("outputs/vacc_impact.csv")
cntries <- unique(dat$country)
# AR_cutoff <- 0.02
# dat |> 
#   filter(attack_rate_naive > AR_cutoff) -> dat_filtered
# R0_cutoff <- 3
for (cntry in cntries) {
  dat |> filter(country == cntry) -> dat_filtered
  
  vacc_impact |>
    filter(data_id %in% dat_filtered$ID) |> 
    group_by(vacc_cov, vacc_week) |> 
    summarize(vacc_cov = vacc_cov[1],
              vacc_week = vacc_week[1],
              vacc_eff = mean(vacc_eff)) -> d
  
  plt <- ggplot(d, aes(x=vacc_week, y=vacc_cov*10, fill=vacc_eff))+
    geom_raster()+
    geom_contour(aes(z=vacc_eff), breaks=4.9, color="red")+
    # metR::geom_contour2(aes(z=vacc_eff, label=stat(level))) +
    theme_minimal()+
    labs(title=paste0("Impact of the vaccine: ", cntry),
         x="Epidemic week",
         y="Vaccine coverage (%)",
         fill="% reduction") +
    scale_x_continuous(breaks=seq(3,12,2))+
    scale_y_continuous(breaks=seq(30,90,10))+
    # scale_fill_continuous(limits=c(0,100))+
    theme(legend.position = "bottom",  
          text = element_text(size=16),
          axis.text = element_text(size=13),
          legend.text = element_text(size=13))
      
  ggsave(paste0("plots/vacc_impact_", cntry, "_", tstamp(), ".pdf"), 
         plt, width = 3.4*2, height = 2.7*2, units = "in")
}
```



Figure 2 Old : The impact of the vaccine based on the static model
```{r, eval=F}
library(patchwork)
svim_lst <- read_rds("outputs/static_vim_lst_parallel_20240613.rds")
svim_ <- svim_lst[[1]]
maxwk <- 16
# rows filled for correct vaccine impact computation
svim <- adjust_week_pct_effect(svim_, maxwk)
svim |> filter(!is.na(confirmed), confirmed > 0) -> svim_lab_conf

# svim_ <- fread("outputs/static_vim20240427.csv")
# one example of lab-confirmed cholera outbreak
# the shape of the outbreak close to a typical outbreak
id0 <- ds[ds$ID == 412, ID_outbreak]
svim_ |>
  filter(vacc_cov >= 0.699, vacc_cov <= 0.701, id == id0) |>
  mutate(case_wk_occurred_tot = case_wk_tot - case_wk_averted_tot) -> d

ggplot()+
  geom_point(data=subset(d, week>=4), aes(x=week,y=case_wk_occurred_tot), 
             color="steelblue")+
  geom_point(data=subset(d, week>=5), aes(x=week,y=case_wk_occurred_tot), 
             color="firebrick")+
  geom_line(data=d, aes(x=week, y=case_wk_tot))+
  geom_point(data=d, aes(x=week, y=case_wk_tot))+
  geom_ribbon(data=subset(d, week >= 5),
              aes(x=week, ymin=case_wk_occurred_tot,
                  ymax=case_wk_tot), fill="firebrick", alpha=0.3)+
  geom_ribbon(data=subset(d, (4 <= week & week <= 5)),
              aes(x=week, ymin=case_wk_occurred_tot,
                  ymax=case_wk_tot), fill="steelblue", alpha=0.3)+
  theme_bw()+
  labs(x="Epidemic week", y="Weekly incidence") -> p1

svim_ |> 
  filter(!is.na(confirmed), confirmed > 0) |>
  group_by(week) |> 
  summarize(n = n()) -> n_outbks_per_week

n_outbks_per_week$surv_prob <- n_outbks_per_week$n / n_outbks_per_week$n[1]

ggplot(n_outbks_per_week, aes(x=week, y=surv_prob*100)) +
  geom_step() +
  # labs(x="", y=expression(Prob~(T >~italic(t)))) +
  labs(x="", y="Cumulative\npercentage") +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA_character_), 
    # necessary to avoid drawing panel outline
    # panel.grid.major = element_blank(), # get rid of major grid
    # panel.grid.minor = element_blank(), # get rid of minor grid
    plot.background = element_rect(fill = "transparent",
                                   colour = NA_character_), # necessary to avoid drawing plot outline
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent"),
    plot.margin = margin(t = 5, r = 5, b = 1, l = 5, unit = "pt")) -> p2a


# unique(svim_lab_conf$data_id)

svim_lab_conf |> 
  filter(vacc_cov >= 0.699, vacc_cov <= 0.701) |>
  ggplot(aes(x=week,y=case_rem_averted_tot,group=week))+
  geom_violin(linewidth=0.6)+
  geom_jitter(width=0.1,size=0.5,color="steelblue",alpha=0.5)+
  labs(x="Week of protection",
       y="Total cases averted per outbreak") +
  theme_bw() -> p2b


p2 <- p2a / p2b + plot_layout(heights = c(1, 2))
 
svim_lab_conf |> 
  filter(week <= maxwk) |>
  group_by(vacc_cov, week) |> 
  summarize(
    vacc_cov = 100*vacc_cov[1],
    vacc_week = week[1],
    vacc_eff = mean(pct_case_averted_tot)) |> 
  ggplot(aes(x=vacc_week, y=vacc_cov, z=vacc_eff))+
  geom_contour_filled()+
  labs(x="Week of protection",
       y="Vaccine coverage (%)",
       fill="Case averted (%)") +
  scale_x_continuous(breaks=seq(1,16,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour() -> p3

  # theme(legend.position = "bottom") -> 

svim_lab_conf |>
  filter(week <= maxwk) |>
  group_by(vacc_cov, week) |> 
  summarize(
    vacc_cov = 100*vacc_cov[1],
    vacc_week = week[1],
    vacc_eff_gain = mean(pct_gain_case_averted_tot,na.rm=T)) |> 
  ggplot(aes(x=vacc_week, y=vacc_cov, z=vacc_eff_gain))+
  geom_contour_filled()+
  # theme_minimal()+
  labs(x="Week of protection",
       y="Vaccine coverage (%)",
       fill="Percent gain (%)") +
  scale_x_continuous(breaks=seq(1,16,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour() -> p4
# theme(legend.position = "bottom") -> p4

library(patchwork)     
p <- (p1 | p2) / (p3 | p4) + plot_annotation(tag_levels = list(c("A","B","","C","D")))
p
ggsave(paste0("plots/vim_static_contour_", tstamp(), ".pdf"), p,
        width=fgw2*1.8, height=fgw2*1.8, units="mm")
```


Figure 2 old: The impact of the vaccine based on the static model
```{r, eval=F}
library(patchwork)

svim_raw <- as.data.frame(fread("outputs/svim_mean_200_20240620.csv"))
svim_raw_confirmed <- filter(svim_raw, (!is.na(confirmed_tot) & confirmed_tot > 0))

# This is deprecated version of computation and used ONLY for ploting
# example vaccine impact
svim_ex <- fread("outputs/static_vim20240427.csv")
# one example of lab-confirmed cholera outbreak
# the shape of the outbreak close to a typical outbreak
id0 <- ds[ds$ID == 412, "ID_outbreak"]
svim_ex |>
  filter(vacc_cov >= 0.899, vacc_cov <= 0.901, id == id0) |>
  mutate(case_wk_occurred_tot = case_wk_tot - case_wk_averted_tot) -> d

ggplot()+
  geom_point(data=subset(d, week>=4), aes(x=week,y=case_wk_occurred_tot), 
             color="steelblue")+
  geom_point(data=subset(d, week>=5), aes(x=week,y=case_wk_occurred_tot), 
             color="firebrick")+
  geom_line(data=d, aes(x=week, y=case_wk_tot))+
  geom_point(data=d, aes(x=week, y=case_wk_tot))+
  geom_ribbon(data=subset(d, week >= 5),
              aes(x=week, ymin=case_wk_occurred_tot,
                  ymax=case_wk_tot), fill="firebrick", alpha=0.3)+
  geom_ribbon(data=subset(d, (4 <= week & week <= 5)),
              aes(x=week, ymin=case_wk_occurred_tot,
                  ymax=case_wk_tot), fill="steelblue", alpha=0.3)+
  theme_bw()+
  labs(x="Epidemic week", y="Weekly incidence") -> p1

svim_ex |> 
  filter(!is.na(confirmed), confirmed > 0) |>
  group_by(week) |> 
  summarize(n = n()) -> n_outbks_per_week

n_outbks_per_week$surv_prob <- n_outbks_per_week$n / n_outbks_per_week$n[1]

ggplot(n_outbks_per_week, aes(x=week, y=surv_prob*100)) +
  geom_step() +
  # labs(x="", y=expression(Prob~(T >~italic(t)))) +
  labs(x="", y="Cumulative\npercentage") +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA_character_), 
    # necessary to avoid drawing panel outline
    # panel.grid.major = element_blank(), # get rid of major grid
    # panel.grid.minor = element_blank(), # get rid of minor grid
    plot.background = element_rect(fill = "transparent",
                                   colour = NA_character_), # necessary to avoid drawing plot outline
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent"),
    plot.margin = margin(t=5, r=5, b=1, l=5, unit="pt")) -> p2a # reduce the bottom margin

# unique(svim_lab_conf$data_id)

svim_raw_confirmed |> 
  filter(vacc_cov >= 0.699, vacc_cov <= 0.701) |>
  ggplot(aes(x = week_vaccination, y=sCh_averted_tot, group = week_vaccination))+
  geom_violin(linewidth = 0.6)+
  geom_jitter(width=0.1, size=0.5, color="steelblue",alpha=0.5)+
  labs(x="Week of vaccination",
       y="Total cases averted\nper outbreak") +
  theme_bw()+
  theme(plot.margin = margin(t=1, r=5, b=5, l=5, unit="pt")) -> p2b

p2 <- p2a / p2b + plot_layout(heights = c(1, 2))
 
svim_raw_confirmed |> group_by(vacc_cov, week_vaccination) |> 
  summarize(
    mean_pct_reduction_50 = quantile(mean_pct_reduction, probs=c(0.5)),
    mean_pct_reduction_025 = quantile(mean_pct_reduction, probs=c(0.025)),
    mean_pct_reduction_975 = quantile(mean_pct_reduction, probs=c(0.975)),
    mean_pct_reduction_25 = quantile(mean_pct_reduction, probs=c(0.25)),
    mean_pct_reduction_75 = quantile(mean_pct_reduction, probs=c(0.75))) -> 
  svim_confirmed


# percent change
svim_confirmed |> 
  group_by(vacc_cov) |> 
  summarize(week_vaccination = week_vaccination, 
            pct_change = 
              c(mean_pct_reduction_50[1:15] - mean_pct_reduction_50[2:16], 
                             NA)) -> svim_confirmed_pctchange

svim_confirmed |> 
  group_by(vacc_cov, week_vaccination) |> 
  summarize(
    vacc_cov = 100 * vacc_cov[1],
    vacc_week = week_vaccination,
    vacc_eff = mean_pct_reduction_50)  |> 
  ggplot(aes(x=vacc_week, y=vacc_cov, z=vacc_eff))+
  geom_contour_filled()+
  labs(x="Week of vaccination",
       y="Vaccine coverage (%)",
       fill="Case averted (%)") +
  scale_x_continuous(breaks=seq(1,16,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour() -> p3

svim_confirmed_pctchange |>
  group_by(vacc_cov, week_vaccination) |> 
  summarize(
    vacc_cov = 100 * vacc_cov[1],
    vacc_week = week_vaccination,
    vacc_eff_change = pct_change) |> 
  ggplot(aes(x=vacc_week, y=vacc_cov, z=vacc_eff_change))+
  geom_contour_filled()+
  # theme_minimal()+
  labs(x="Week of vaccination",
       y="Vaccine coverage (%)",
       fill="Vaccome impact\nchange (%)") +
  scale_x_continuous(breaks=seq(1,16,2))+
  scale_y_continuous(breaks=seq(10,90,20))+
  theme_2D_contour() -> p4


library(patchwork)     
p <- (p1 | p2) / (p3 | p4) + plot_annotation(tag_levels = list(c("A","B","","C","D")))
p
ggsave(paste0("plots/vim_static_contour_", tstamp(), ".pdf"), p,
        width=fgw2*1.8, height=fgw2*1.8, units="mm")
```


```{r}
# case averted
svim_list <- read_rds("outputs/svim_list_20240620.rds")
svim1 <- svim_list[[1]] |> filter(vacc_cov == 0.9, week_vaccination == 1)

get_adm <- function(x) {
  # d <- grep("[a-zA-Z+]", strsplit(x, "::|-")[[1]], value=TRUE)
  d <- lapply(x, function(z) grep("[a-zA-Z+]", strsplit(z, "::|-")[[1]], value=TRUE))
  return (d)
}

get_date <- function(x) {
  # d <- grep("[^0-9+]", strsplit(x, "::|-")[[1]], value=TRUE)
  # ymd <- strsplit(x, "-")[[1]]
  ymd <- lapply(x, function(z) strsplit(z, "-"))
  dt <- lapply(ymd, function(z) as.Date(paste0(z[[1]][2], "-", z[[1]][3], "-", z[[1]][4])))
  return (dt)
}

adms <- lapply(svim1$id, function(x) get_adm(x))

svim1$admin0 <- sapply(adms, function(x) x[[1]][2])
svim1$admin1 <- sapply(adms, function(x) x[[1]][3])
svim1$admin2 <- sapply(adms, function(x) x[[1]][4])
svim1$admin3 <- sapply(adms, function(x) x[[1]][5])

dates <- lapply(svim1$id, function(x) get_date(x))
# we already have date variable and this is based on 

# replace date, which appears to be based on the weekly interval
# such that we may have better 
svim1$new_date <- as.Date(sapply(dates, function(x) x[[1]]), 
                          origin="1970-01-01")
# svim1 |> 
#   group_by(admin0, admin1, admin2, admin3) |>
#   reframe(id = id,
#           adm = paste0(admin0, "::", admin1, "::", admin2, "::", admin3),
#           date = new_date,
#           min_date = min(date),
#           min_date_id = id[which(date == min_date)[1]],
#           date_diff = date - min_date) |>
#   arrange(-desc(date_diff)) -> res

# dur_ve <- 3*365; # duration of vaccine-derived protection
# # res[res$date_diff > 0 & res$date_diff < dur_ve,] 
# # outbreaks with the period of vaccine-derived protection are removed
# res$removed <- ifelse(res$date_diff > 0 & res$date_diff < dur_ve, TRUE, FALSE)
# res <- res[order(res$admin0, res$admin1, res$admin2, res$admin3),]
# 
# res |> 
#   filter(!removed & date_diff > 0) |> 
#   group_by(admin0, admin1, admin2, admin3) |>
#   reframe(id = id,
#           adm = adm,
#           date = date,
#           min_date = min(date),
#           min_date_id = id[which(date == min_date)[1]],
#           date_diff = date - min_date) |>
#   arrange(-desc(date_diff)) -> res2
# 
# res2$removed <- ifelse(res2$date_diff > 0 & res2$date_diff < dur_ve, TRUE, FALSE)
# 
# res2 |> 
#   filter(!removed & date_diff > 0) |> 
#   group_by(admin0, admin1, admin2, admin3) |>
#   reframe(id = id,
#           adm = adm,
#           date = date,
#           min_date = min(date),
#           min_date_id = id[which(date == min_date)[1]],
#           date_diff = date - min_date) |>
#   arrange(-desc(date_diff)) -> res3
# 
# res3$removed <- ifelse(res3$date_diff > 0 & res3$date_diff < dur_ve, TRUE, FALSE)
# 
# res_final <- left_join(res, res2, by="id")
# res_final <- left_join(res_final, res3, by="id")

find_ghost_outbreaks <- function(x, lb=0, ub=1000, ...) {
  group_var <- enquos(...)
  x |> 
    # group_by(admin0, admin1, admin2) |>
    group_by(!!!group_var) |>
    reframe(id = id,
            data_id = data_id,
            date = new_date,
            min_date = min(date),
            date_diff = date - min_date) -> x1

  ghost_id <- x1[x1$date_diff > lb & x1$date_diff <= ub,]$data_id
  x1 <- x1[x1$date_diff > ub, ]
  
  while (!is.null(x1) & nrow(x1) > 1) {
    x1 |> 
      # group_by(admin0, admin1, admin2) |>
      group_by(!!!group_var) |>
      reframe(id = id,
            data_id = data_id,
            date = date,
            min_date = min(date),
            date_diff = date - min_date) -> x1
    ghost_id <- c(ghost_id, x1[x1$date_diff > lb & x1$date_diff <= ub,]$data_id)
    x1 <- x1[x1$date_diff > ub,]
  }
  return (ghost_id)
}

```
