---
title: "Cholera Vaccine Impact Modeling: Visualization and Tables"
author: "Jong-Hoon Kim"
date: "2025-10-31"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

# Introduction

This document creates visualizations and tables to summarize the results from 
cholera vaccine impact simulations. 

# Setup

```{r setup}
#| warning: false
#| message: false

# Load required packages
source("R/utils.R")  # utility functions
library(data.table)
library(tidyverse)
library(magrittr)
library(purrr)
library(broom)
library(effsize)
library(randomForest)
library(rstatix)
library(ggridges)
library(readxl)
library(janitor)
library(scales)
library(viridis)
library(sf)
library(metR)
library(googlesheets4)
library(MASS)
library(patchwork)
library(ggpattern)
library(gridExtra)
library(corrr)
library(RColorBrewer)
library(countrycode)
library(conflicted)
library(xdvir)

# Resolve package conflicts in favor of dplyr
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflicts_prefer(data.table::first)
conflicts_prefer(ggplot2::margin)
# Output control flags
save_plot <- FALSE
save_table <- FALSE

# Figure dimensions
figure_size <- data.table(
  journal = c("Lancet"),
  single = c(85),
  double = c(178)
)
fgw1 <- figure_size[journal == "Lancet", single]
fgw2 <- figure_size[journal == "Lancet", double]

# Color palette for consistent visualization
set1 <- brewer.pal(9, "Set1")  # color sets
# my_discrete_colors <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00")

# Define standard probability values for quantiles
PROBS <- c(0.025, 0.25, 0.5, 0.75, 0.975)  # 2.5%, 25%, 50%, 75%, 97.5% quantiles

# Reference vaccination week
VW <- 9  # representative vaccination week from GTFCC data

# Sequence of vaccination weeks to evaluate
VW_seq <- seq(1, 15, 2)

# Three representative vaccination coverage rates
VC3 <- c(0.6, 0.75, 0.9)  # 60%, 75%, 90%

# Reference vaccination coverage rate
VC <- VC3[2]  # 75% coverage

# Thresholds for case-triggered vaccination to analyze
cts_sel <- c(1, seq(10, 100, 10))  # case triggers selected

# Latest date for which OCV campaign data is available
YEAR <- as.Date("2020-12-28")

# Load geographic data for mapping
shp <- st_read("data/paper/total_shp_0427.shp", quiet = TRUE)
africa <- st_read("data/paper/afr_g2014_2013_0.shp", quiet = TRUE)

# only a subset of sample runid's is used for plotting for convenience
set.seed(42)
# sample_runids <- sample(1:200, size=30)

# Plot parameters
line_width_50 <- 1.2
line_width_95 <- 0.5
line_interval <- 0.7
point_size <- 0.5
x_size <- 1.5
x_stroke <- 1
```

# Load data

```{r load-data}
#| warning: false
# Load outbreak data
# All outbreaks including those affected by OCVs
outbreak_data_all <- 
  as.data.frame(fread("outputs/paper/outbreak_data_all_20250228.csv"))

# Outbreaks included in vaccine impact modeling (not affected by OCVs)
outbreak_data <- fread("outputs/paper/outbreak_data_20250228.csv") 

# outbreaks filtered
# This includes outbreaks that are finally analyzed and excludes
# outbreaks that were duplicated
outbreaks_filtered <-
  readRDS("outputs/paper/outbreak_summary_filtered_20250418.rds")

outbreak_data <- outbreak_data %>%
  filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 
# outbreaks excluded
outbreaks_excluded <- read_rds("outputs/paper/outbreaks_excluded_20250612.rds")

# Historical OCV deployment data
# data were collected based on the website: https://apps.epicentre-msf.org/public/app/gtfcc
# The number of doses shipped by 2020 and 2024 can be obtained by adjusting the shipment period
global_ocv_use <-
  clean_names(fread("data/paper/OCV_Use_GTFCC_Dashboard_20251031.csv"))

# Load model parameters
dis_params <- as.data.frame(fread("data/paper/parameters.csv"))

# Load Bayesian beta regression results for the indirect vaccine effectiveness
ive_fit <- readRDS("outputs/paper/ive_yhat_yrep_20250919.rds")
ive_data <- ive_fit$yrep
rm(ive_fit)  # Free memory by removing large object

# Load outbreak summary data
ds <- clean_names(as.data.frame(fread("data/paper/outbreak_files_Oct2024/outbreak_summary.csv")))
# message(paste0(length(unique(ds$country)), " countries"))

# Create unique outbreak identifiers
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# Join outbreak data with summary information
outbk_data_summary <- 
  left_join(
    filter(outbreak_data, week == 1), 
    ds[, c("outbreak_pop", "total_s_ch", "total_c_ch", "total_deaths", "duration_weeks", 
           "peak_week_incidence", "time_to_peak_week", "peak_week_incidence_per_100percent", 
           "peak_week_incidence_per_10e3", "mean_r0_1st_week", "area_per_1km2", 
           "rural_urban", "ar_per_10e3", "cfr_per_100percent", "id_outbreak")], 
    by = "id_outbreak"
  )

outbk_data_summary %<>% 
  filter(id_outbreak %in% outbreaks_filtered$id_outbreak)

outbk_data_summary <- fread("outputs/paper/outbreak_data_summary_20251022.csv")
# Summarize attack rates and population sizes
# summary(outbk_data_summary$ar_per_10e3)
# summary(outbk_data_summary$outbreak_pop)

# outbreaks in the top quartile by attack rate (ar), duration (dur), and size
n25 <- floor(0.25 * nrow(outbk_data_summary))

top_q_ar_ids <- outbk_data_summary %>% 
  arrange(desc(ar_per_10e3)) %>% slice(1:n25) %>% pull(id_outbreak)
top_q_dur_ids <- outbk_data_summary %>%
  arrange(desc(duration_weeks)) %>% slice(1:n25) %>% pull(id_outbreak)
top_q_size_ids <- outbk_data_summary %>% 
  arrange(desc(total_s_ch)) %>% slice(1:n25) %>% pull(id_outbreak)


n10 <- floor(0.1 * nrow(outbk_data_summary))

top_10_ar_ids <- outbk_data_summary %>% 
  arrange(desc(ar_per_10e3)) %>% slice(1:n10) %>% pull(id_outbreak)
top_10_dur_ids <- outbk_data_summary %>%
  arrange(desc(duration_weeks)) %>% slice(1:n10) %>% pull(id_outbreak)
top_10_size_ids <- outbk_data_summary %>% 
  arrange(desc(total_s_ch)) %>% slice(1:n10) %>% pull(id_outbreak)
```

# Load vaccine impact simulation results

```{r load-results}
# Outbreak onset-triggered vaccination results
cea_files <- c("outputs/paper/cea_onset_trigger_20251031.rds")
cea <- map_dfr(cea_files, read_rds)
cea_ori <- cea %>% filter(ori_occurred == TRUE)

# percentage case averted summary for Figure 3
cea_ori_pca <- read_rds("outputs/paper/cea_pca_summary_20251031.rds")
# impact summary for Figure 3
cea_prioritization <-
  read_rds("outputs/paper/summary_impact_0.75_prioritization_20251031.rds")

cea_weekly <- read_rds("outputs/paper/cea_delay_onset_trigger_20251031.rds")
cea_weekly_delay_2 <- 
  cea_weekly %>% filter(no_ori_delay_outbreak_end == 2)

cea_weekly_delay_2_ori <- cea_weekly %>% 
  filter(no_ori_delay_outbreak_end == 2, ori_occurred == TRUE)

cea_alt_data <- 
  read_rds("outputs/paper/cea_alt_data_onset_trigger_20251031.rds")

# weekly decline
fit_decay <- read_rds("outputs/paper/fit_exp_decay_20251031.rds")
fit_decay_ct <- 
  read_rds("outputs/paper/fit_exp_decay_case_trigger_20251031.rds")
fit_lm <- read_rds("outputs/paper/fit_lm_20251031.rds")
# Case-triggered vaccination results
cea_ct <- read_rds("outputs/paper/cea_case_trigger_20251031.rds")
cea_ct_ori <- cea_ct %>% filter(ori_occurred == TRUE)
# outbreak subsets for given number of OCVs
subset_files <- c(
  "outputs/paper/cea_summary_subsets_vw_1_20251031.rds",
  "outputs/paper/cea_summary_subsets_vw_9_20251031.rds"
)
cea_subsets <- map_dfr(subset_files, read_rds)
# excluded outbreaks for sensitivity analysis
# cea_excl <- read_rds("outputs/cea_0.6_onset_trigger_excl_20250716.rds")
# cea_excl <- rbind(cea_excl, read_rds("outputs/cea_0.75_onset_trigger_excl_20250716.rds"))
# cea_excl <- rbind(cea_excl, read_rds("outputs/cea_0.9_onset_trigger_excl_20250716.rds"))
# cea_ct_excl <- read_rds("outputs/cea_0.75_case_trigger_excl_20250716.rds")
```

# Abstract

Values used in the abstract
```{r}
onset_impact <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, 
                      pct_reduc_case, 
                      case_averted_per_1000_OCV, 
                      daly_averted_per_1000_OCV,
                      cost_per_case_averted, 
                      cost_per_daly_averted)

# cases averted (CA) for vaccine coverage of 75% with ORI in week 1
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", vacc_week == 1) %>%
  format_median_iqr()
# percent case averted (PCA) at week 1
onset_impact %>% 
  filter(variable == "pct_reduc_case", vacc_week == 1) %>%
  format_median_iqr()
# CA At week 9
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", vacc_week == 9) %>%
  format_median_iqr()
# PCA At week 9
onset_impact %>% 
  filter(variable == "pct_reduc_case", vacc_week == 9) %>%
  format_median_iqr()

# weekly reduction rate
fit_decay[["9"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# half_life                  
fit_decay[["9"]] %>%
  summarize_quantiles(half_life) %>% 
  format_median_iqr(digits=1)

# case averted per 1000 OCVs (CAPD) at week 1
onset_impact %>% 
  filter(variable == "case_averted_per_1000_OCV", vacc_week == 1) %>%
  format_median_iqr(digits=2)
onset_impact %>% 
  filter(variable == "cost_per_case_averted", vacc_week == 1) %>%
  format_median_iqr(digits=0)

onset_impact %>% 
  filter(variable == "case_averted_per_1000_OCV", vacc_week == 9) %>%
  format_median_iqr(digits=2)
onset_impact %>% 
  filter(variable == "cost_per_case_averted", vacc_week == 9) %>%
  format_median_iqr(digits=0)

# fold increase in case averted per ORI by prioritization
ca <- onset_impact %>% 
  filter(variable == "s_ch_averted_tot") %>% 
  pull(q500)

ca_top_q_size <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_size_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, pct_reduc_case) %>% 
  filter(variable == "s_ch_averted_tot") %>% 
  pull(q500)

paste0("fold increase: ", ca_top_q_size / ca)

# fold increase in case averted per 1000 OCVs
capd_top_q_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV) %>% 
  pull(q500)

# fold increase in case averted per 1000 OCVs
capd_top_10_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_10_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV) %>% 
  pull(q500)

capd <- onset_impact %>% 
  filter(variable == "case_averted_per_1000_OCV") %>% 
  pull(q500) 

paste0("fold increase in case averted per 1,000 OCVs: ", 
       capd_top_q_ar / capd)

paste0("fold increase in case averted per 1,000 OCVs: ", 
       capd_top_10_ar / capd)

# fold increase in averted per 1000 OCVs
# top quartile
dapd_top_q_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(daly_averted_per_1000_OCV) %>%  
  pull(q500)

# top 10%
dapd_top_10_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_10_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(daly_averted_per_1000_OCV) %>%  
  pull(q500)

dapd <- onset_impact %>% 
  filter(variable == "daly_averted_per_1000_OCV") %>% 
  pull(q500) 

paste0("fold increase in DALY averted per 1,000 OCVs: ",
       dapd_top_q_ar / dapd)


paste0("fold increase in DALY averted per 1,000 OCVs: ",
       dapd_top_10_ar / dapd)

# cost per DALY averted
# top quartile
cpda_top_q_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted) %>% 
  pull(q500)
# top 10%
cpda_top_10_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_10_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted) %>% 
  pull(q500)

cpda <- onset_impact %>% 
  filter(variable == "cost_per_daly_averted") %>% 
  pull(q500) 

paste0("fold decrease in cost per DALY averted: ", cpda/cpda_top_q_ar)
paste0("fold decrease in cost per DALY averted: ", cpda/cpda_top_10_ar)
# subset
cea_subsets %>% 
  mutate(prob_i = as.integer(round(100*priority_prob))) %>% 
  filter(value_column == "case_averted_per_1000_OCV", vacc_week == 9, 
         prob_i == 0) %>% 
  summarize_quantiles(q500) %>% 
  format_median_iqr(digits=2)

cea_subsets %>% 
  mutate(prob_i = as.integer(round(100*priority_prob))) %>% 
  filter(value_column == "case_averted_per_1000_OCV", vacc_week == 1, 
         vacc_priority == "AR", prob_i == 30) %>% 
  summarize_quantiles(q500) %>% 
  format_median_iqr(digits=2)
```

# Introduction

Values used in the abstract
```{r}
# number of OCVs doses shipped from OCV stockpile (million doses)
ocv_use_dashboard
# proportion approved by ICG (reactive) by 2024
ocv_use_dashboard$cumulative_icg[2] / ocv_use_dashboard$cumulative_total[2]
# outbreaks span from 2010 to 2020
unique(outbk_data_summary$year)
outbk_data_summary %>% filter(year == 2020)
```

# Plots

## Figure 1. Outbreaks and % case reduction 

The right panel of Figure 1 in the main text was generated using the following code. The left panel, a flowchart illustrating outbreak identification and selection for vaccine impact modeling, was created in PowerPoint.

```{r}
# Global parameters ----------------------------------------------------
text_size <- 5          # Default text size for annotations
line_height <- 0.95     # Line spacing for text annotation
pattern_dens <- 0.1     # Density of pattern filling
pattern_spacing <- 0.1  # Spacing between patterns
step_linewidth <- 0.7   # Width of step lines
arrow_vertical_len <- 14 # Vertical length of arrows

# Axis labels
ylabel <- "Weekly suspected cholera cases"
xlabel <- "Weeks since outbreak onset"

# Key timeline parameters
week_or <- 3      # Outbreak response at week 3  
week_vax <- 6.5   # Vaccination started
week_vax_eff <- 8.5 # Vaccine takes effect

# Rectangle dimensions configuration ------------------------------------
# This section makes it easy to adjust all rectangles at once

# Base rectangle dimensions
rect_width <- 0.8
rect_height <- 6

# Spacing between rectangles
gap_horizontal <- 1.9   # Horizontal gap between rectangle groups
gap_vertical <- 12      # Vertical gap between top and bottom rectangles

# Font sizes for elements
eq_symbol_size <- 8     # Size for equation symbols (<, +)
paren_size <- 9         # Size for parentheses
plus_size <- 5          # Size for plus symbols

# Vertical offset for second equation from first
eq2_vertical_offset <- 24

# Helper functions ----------------------------------------------------
create_filled_square <- function(x1, y1, x2, y2, fill_color, 
                                 color='black', alpha=0.5) {
  geom_rect(xmin=x1, xmax=x2, ymin=y1, ymax=y2,
            color=color, fill=fill_color, alpha=alpha, inherit.aes = FALSE)
}

create_striped_square <- function(x1, y1, x2, y2, fill_color, color='black',
                                  dens=0.1, angle=30) {
  geom_rect_pattern(
    xmin=x1, xmax=x2, ymin=y1, ymax=y2,
    pattern = 'stripe',
    fill = NA,
    colour = color,
    pattern_density = dens,
    pattern_spacing = 0.025,
    pattern_fill = fill_color,
    pattern_colour = NA,
    pattern_angle = angle,
    inherit.aes = FALSE)
}

create_divider <- function(x1, x2, y, extension=0.2, linewidth=0.7) {
  annotate("segment", 
           x = x1 - extension, 
           y = y, 
           xend = x2 + extension, 
           yend = y, 
           linewidth = linewidth)
}

theme_upper_panel <- function() {
  theme_classic() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 16),
      axis.text.y = element_text(size = 14),
      plot.caption = element_text(size = 14),
      legend.position = "none"
    )
}

# Data preparation ----------------------------------------------------
# Create synthetic example of an outbreak trajectory with and without vaccination
d <- data.frame(
  date = seq(as.Date("2010-06-15"), by = "week", length.out = 24),
  week = 0:23,
  # No vaccination scenario
  s_ch = c(2, 2, 3, 4, 10, 12, 29, 100, 120, 150, 140, 
           75, 68, 47, 24, 5, 12, 5, 3, 2, 4, 1, 0, 0)
)

# Add vaccination scenario (40% of cases after vaccine takes effect)
d$s_ch_vacc <- c(
  2, 2, 3, 4, 10, 12, 29, 100, 120, 
  round(0.4 * c(150, 140, 75, 68, 47, 24, 5, 12, 5, 3, 2, 4, 1, 0, 0))
)
d$s_ch_cumul <- cumsum(d$s_ch)

# Rename columns for clarity
names(d) <- c("date", "week", "No vacc", "Vacc", "No vacc cumulative")

# Reshape data for plotting
dlong <- pivot_longer(d, cols = c("No vacc", "Vacc"))

# Define time weeks for additional scenarios
weeks <- 0:25

# Generate additional outbreak scenarios
# Smallest outbreak (ended before vaccination)
smallest_no_vax <- rpois(length(weeks), lambda=30 * exp(-0.1 * (weeks - 2)^2))
smallest_no_vax[(week_vax-1):length(smallest_no_vax)] <- 0

# Smaller outbreak (ended at vaccine takes effect)
smaller_no_vax <- rpois(length(weeks), lambda=50 * exp(-0.1 * (weeks - 3)^2))
smaller_no_vax[week_vax_eff:length(smaller_no_vax)] <- 0

# Small outbreak (peak at week 8)
small_no_vax <- rpois(length(weeks), lambda=70 * exp(-0.08 * (weeks - 8)^2))
small_with_vax <- small_no_vax
small_with_vax[weeks > week_vax_eff] <- round(small_with_vax[weeks > week_vax_eff] * 0.5)  # 50% reduction

# Large outbreak (peak at week 12)
large_no_vax <- rpois(length(weeks), lambda=120 * exp(-0.04 * (weeks - 12)^2))
large_with_vax <- large_no_vax
large_with_vax[weeks > week_vax_eff] <- round(large_with_vax[weeks > week_vax_eff] * 0.5)  # 50% reduction

# Combine into a data frame
df <- tibble(
  week = rep(weeks, 6),
  cases = c(smallest_no_vax, smaller_no_vax, small_no_vax, small_with_vax, 
            large_no_vax, large_with_vax),
  outbreak = rep(c("Smallest", "Smaller", "Small", "Small", "Large", "Large"), 
                 each = length(weeks)),
  scenario = c(rep("No Vaccine", 2*length(weeks)), 
               rep(c("No Vaccine", "With Vaccine"), times=2, each=length(weeks)))
)

# Create wide version of the data
df_wide <- df %>%
  pivot_wider(names_from = scenario, values_from = cases) %>%
  mutate(reduction = `No Vaccine` - `With Vaccine`)

# Rectangle positions for equation 1 ----------------------------------
# Starting position for first equation
base_x1 <- 15
base_y1 <- 65

# a/b fraction
a_x1 <- base_x1
a_y1 <- base_y1
a_x2 <- a_x1 + rect_width
a_y2 <- a_y1 + rect_height

b_x1 <- a_x1
b_y1 <- a_y1 - gap_vertical
b_x2 <- b_x1 + rect_width
b_y2 <- b_y1 + rect_height

ab_div_y <- (a_y1 + b_y2) / 2

# c/d fraction
c_x1 <- a_x1 + gap_horizontal
c_y1 <- a_y1
c_x2 <- c_x1 + rect_width
c_y2 <- c_y1 + rect_height

d_x1 <- c_x1
d_y1 <- b_y1
d_x2 <- d_x1 + rect_width
d_y2 <- d_y1 + rect_height

cd_div_y <- (c_y1 + d_y2) / 2

# Inequality symbol
eq1_lt_x <- (a_x2 + c_x1) / 2
eq1_lt_y <- (a_y1 + b_y2) / 2

# Equation label
eq1_label_x <- a_x1 - 1.2
eq1_label_y <- cd_div_y

# Rectangle positions for equation 2 ----------------------------------
# Left side: 1/2(a/b+c/d)
# a/b part
eq2_a_x1 <- a_x1 + 0.2
eq2_a_y1 <- a_y1 - eq2_vertical_offset
eq2_a_x2 <- eq2_a_x1 + rect_width
eq2_a_y2 <- eq2_a_y1 + rect_height

eq2_b_x1 <- eq2_a_x1
eq2_b_y1 <- eq2_a_y1 - gap_vertical
eq2_b_x2 <- eq2_b_x1 + rect_width
eq2_b_y2 <- eq2_b_y1 + rect_height

# c/d part
eq2_c_x1 <- eq2_a_x1 + gap_horizontal
eq2_c_y1 <- eq2_a_y1
eq2_c_x2 <- eq2_c_x1 + rect_width
eq2_c_y2 <- eq2_c_y1 + rect_height

eq2_d_x1 <- eq2_c_x1
eq2_d_y1 <- eq2_b_y1
eq2_d_x2 <- eq2_d_x1 + rect_width
eq2_d_y2 <- eq2_d_y1 + rect_height

# Parentheses
left_paren_x <- eq2_a_x1 - 0.3
right_paren_x <- eq2_c_x2 + 0.3
paren_y <- (eq2_a_y1 + eq2_b_y2) / 2

# 1/2 fraction position
eq2_frac_x <- right_paren_x + 0.6
eq2_frac_y_num <- (eq2_a_y1 + eq2_a_y2)/2
eq2_frac_y_denom <- (eq2_b_y1 + eq2_b_y2)/2
eq2_frac_y_line <- (eq2_a_y1 + eq2_b_y2)/2

# Plus symbol position
eq2_plus1_x <- (eq2_a_x2 + eq2_c_x1) / 2
eq2_plus1_y <- (eq2_a_y1 + eq2_b_y2) / 2

# Less than symbol for second equation
eq2_lt_x <- eq2_frac_x + 0.8
eq2_lt_y <- (eq2_a_y1 + eq2_b_y2)/2

# Right side: (a+c)/(b+d)
# Numerator (a+c)
eq2_right_a_x1 <- eq2_lt_x + 0.8
eq2_right_a_y1 <- eq2_a_y1
eq2_right_a_x2 <- eq2_right_a_x1 + rect_width
eq2_right_a_y2 <- eq2_right_a_y1 + rect_height

eq2_right_c_x1 <- eq2_right_a_x1 + gap_horizontal
eq2_right_c_y1 <- eq2_a_y1
eq2_right_c_x2 <- eq2_right_c_x1 + rect_width
eq2_right_c_y2 <- eq2_right_c_y1 + rect_height

eq2_plus2_x <- (eq2_right_a_x2 + eq2_right_c_x1)/2
eq2_plus2_y <- (eq2_right_a_y1 + eq2_right_a_y2)/2

# Denominator (b+d)
eq2_right_b_x1 <- eq2_right_a_x1
eq2_right_b_y1 <- eq2_b_y1
eq2_right_b_x2 <- eq2_right_b_x1 + rect_width
eq2_right_b_y2 <- eq2_right_b_y1 + rect_height

eq2_right_d_x1 <- eq2_right_c_x1
eq2_right_d_y1 <- eq2_right_b_y1
eq2_right_d_x2 <- eq2_right_d_x1 + rect_width
eq2_right_d_y2 <- eq2_right_d_y1 + rect_height

eq2_plus3_x <- eq2_plus2_x
eq2_plus3_y <- (eq2_right_b_y1 + eq2_right_b_y2)/2

# Division line
eq2_right_div_x1 <- eq2_right_a_x1
eq2_right_div_x2 <- eq2_right_c_x2
eq2_right_div_y <- (eq2_right_a_y1 + eq2_right_b_y2) / 2

# Equation 2 label
eq2_label_x <- eq1_label_x
eq2_label_y <- a_y1 - eq2_vertical_offset

# Create plots --------------------------------------------------------
# Panel 1: Main example with vaccine effect
plt1a <- ggplot(dlong, aes(week, value, linetype = name, shape = name, group = name)) +
  # Background stripes for no vaccination scenario
  geom_rect_pattern(
    data = subset(dlong, name == "No vacc"),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=value),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_colour = NA,
    pattern_density = pattern_dens,
    pattern_fill = 'steelblue',
    pattern_angle = 30) +
  
  # Vaccination impact rectangle
  geom_rect(
    data = d,
    aes(xmin=week, xmax=lead(week), ymin=`Vacc`, ymax=`No vacc`), 
    fill='steelblue', 
    alpha=0.5,
    color=NA,
    inherit.aes = FALSE) +
  
  # Step lines for clarity
  geom_step(linewidth=step_linewidth) +
  
  # Outbreak response triggered arrow and label
  geom_segment(
    x=week_or, y=10+arrow_vertical_len, xend=week_or, yend=10,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_or-0.5, y=38, 
    label = "Outbreak\nresponse triggered", 
    hjust=0.5, size=text_size, lineheight=line_height
  ) +
  
  # Vaccination started arrow and label
  geom_segment(
    x=week_vax, y=32+arrow_vertical_len, xend=week_vax, yend=32,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_vax, y=60, 
    label = "Vaccination\nstarted", hjust=1,
    size=text_size, lineheight=line_height
  ) +
  
  # Vaccine takes effect arrow and label
  geom_segment(
    x=week_vax_eff, y=122+arrow_vertical_len, xend=week_vax_eff, yend = 122,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_vax_eff, y=147, 
    label="Vaccine\ntakes effect", hjust=1,
    size=text_size, lineheight=line_height
  ) +
  
  # Scenario labels
  annotate(
    "text", x=11.5, y=120, 
    label = "Without\nvaccination", hjust=0, 
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=9.5, y=12, 
    label = "With\nvaccination", hjust=0,
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=16, y=75, 
    label = "% case\nreduction", 
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=19.2, y=75, 
    label = bquote("= 100 \u00D7"), 
    size=text_size, lineheight=line_height
  ) +
  
  # Fraction expressed using squares
  create_divider(x1=21.1-0.2, x2=22.1+0.2, y=75) +
  create_filled_square(
    x1=21.1, x2=22.1, y1=77, y2=85, 
    color='black', fill="steelblue", alpha=0.5
  ) +
  create_striped_square(
    x1=21.1, x2=22.1, y1=64, y2=72,
    fill_color='steelblue', color='black'
  ) +
  
  # Colored lines on the horizontal axis marking periods
  annotate("segment", x=0, y=0, xend=3, yend=0, linewidth=2, color="firebrick") +
  annotate("segment", x=3, y=0, xend=7, yend=0, linewidth=2, color="forestgreen") +
  annotate("segment", x=7, y=0, xend=9, yend=0, linewidth=2, color="purple") +
  
  scale_x_continuous(limits=c(0,25)) +
  labs(x="", y=ylabel) +
  theme_upper_panel()

# Panel 2: Smaller outbreaks
plt1b <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot smaller outbreaks
  geom_step(
    data = subset(df, outbreak == "Smallest"),
    aes(linetype=scenario),
    linewidth=step_linewidth, 
    color='grey'
  ) +
  geom_step(
    data = subset(df, outbreak == "Smaller"),
    aes(linetype=scenario),
    linewidth=step_linewidth, 
    color='black'
  ) +
  
  labs(x=NULL, y=ylabel) +
  theme_upper_panel()

# Panel 3: Small outbreak with equations
plt1c <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot small outbreak
  geom_step(
    data = subset(df, outbreak == "Small"), 
    aes(linetype=scenario),
    linewidth=step_linewidth
  ) +
  
  # Pattern for no vaccination scenario
  geom_rect_pattern(
    data = subset(df, (scenario == "No Vaccine" & outbreak == "Small")),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=cases),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_density = pattern_dens*0.5,
    pattern_fill = 'steelblue',
    pattern_colour = NA,
    pattern_angle = 30
  ) +
  
  # Vaccination impact area
  geom_rect(
    data=subset(df_wide, outbreak == "Small"),
    aes(xmin=week, xmax=lead(week), ymin=`With Vaccine`, ymax=`No Vaccine`),
    fill="steelblue", 
    alpha=0.5, 
    inherit.aes=FALSE
  ) +
  
  # EQUATION 1: a/b < c/d
  # a/b fraction
  create_filled_square(a_x1, a_y1, a_x2, a_y2, "steelblue") +
  create_striped_square(b_x1, b_y1, b_x2, b_y2, "steelblue", dens=pattern_dens*0.5) +
  create_divider(a_x1, a_x2, ab_div_y) +
  
  # Less than symbol
  annotate("text", x=eq1_lt_x, y=eq1_lt_y, label= "<", size=eq_symbol_size) +
  
  # c/d fraction
  create_filled_square(c_x1, c_y1, c_x2, c_y2, "firebrick") +
  create_striped_square(d_x1, d_y1, d_x2, d_y2, "firebrick", dens=pattern_dens*0.5) +
  create_divider(c_x1, c_x2, cd_div_y) +
  
  # Label for equation 1
  annotate("text", x=eq1_label_x, y=eq1_label_y, label="1.", size=5, hjust=0) +
  
  # EQUATION 2: 1/2(a/b+c/d) < (a+c)/(b+d)
  # Left side: 1/2(a/b+c/d)
  
  # a/b part
  create_filled_square(eq2_a_x1, eq2_a_y1, eq2_a_x2, eq2_a_y2, "steelblue") +
  create_striped_square(eq2_b_x1, eq2_b_y1, eq2_b_x2, eq2_b_y2, "steelblue", dens=pattern_dens*0.5) +
  create_divider(eq2_a_x1, eq2_a_x2, (eq2_a_y1 + eq2_b_y2)/2) +
  
  # Plus symbol
  annotate("text", x=eq2_plus1_x, y=eq2_plus1_y, label="+", size=plus_size) +
  
  # c/d part
  create_filled_square(eq2_c_x1, eq2_c_y1, eq2_c_x2, eq2_c_y2, "firebrick") +
  create_striped_square(eq2_d_x1, eq2_d_y1, eq2_d_x2, eq2_d_y2, "firebrick", dens=pattern_dens*0.5) +
  create_divider(eq2_c_x1, eq2_c_x2, (eq2_c_y1 + eq2_d_y2)/2) +
  
  # Parentheses
  annotate("text", x=left_paren_x, y=paren_y, label="(", size=paren_size) +
  annotate("text", x=right_paren_x, y=paren_y, label=")", size=paren_size) +
  
  # 1/2 fraction
  annotate("text", x=eq2_frac_x, y=eq2_frac_y_num, label="1", size=5) +
  annotate("text", x=eq2_frac_x, y=eq2_frac_y_denom, label="2", size=5) +
  annotate("segment", 
           x=eq2_frac_x-0.3, y=eq2_frac_y_line, 
           xend=eq2_frac_x+0.3, yend=eq2_frac_y_line, 
           linewidth=0.8) +
  
  # Less than symbol for second equation
  annotate("text", x=eq2_lt_x, y=eq2_lt_y, label="<", size=eq_symbol_size) +
  
  # Right side: (a+c)/(b+d)
  # Numerator (a+c)
  create_filled_square(eq2_right_a_x1, eq2_right_a_y1, eq2_right_a_x2, eq2_right_a_y2, "steelblue") +
  annotate("text", x=eq2_plus2_x, y=eq2_plus2_y, label="+", size=plus_size) +
  create_filled_square(eq2_right_c_x1, eq2_right_c_y1, eq2_right_c_x2, eq2_right_c_y2, "firebrick") +
  
  # Division line
  create_divider(eq2_right_div_x1, eq2_right_div_x2, eq2_right_div_y) +
  
  # Denominator (b+d)
  create_striped_square(eq2_right_b_x1, eq2_right_b_y1, eq2_right_b_x2, eq2_right_b_y2, 
                        "steelblue", dens=pattern_dens*0.5) +
  annotate("text", x=eq2_plus3_x, y=eq2_plus3_y, label="+", size=plus_size) +
  create_striped_square(eq2_right_d_x1, eq2_right_d_y1, eq2_right_d_x2, eq2_right_d_y2, 
                        "firebrick", dens=pattern_dens*0.5) +
  
  # Label for equation 2
  annotate("text", x=eq2_label_x, y=eq2_label_y, label="2.", size=5, hjust=0) +
  
  labs(x=NULL, y=ylabel) +
  theme_upper_panel()

# Panel 4: Large outbreak  
plt1d <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot large outbreak
  geom_step(
    data = subset(df, outbreak == "Large"), 
    aes(linetype=scenario),
    linewidth=step_linewidth
  ) +
  
  # Pattern for no vaccination scenario
  geom_rect_pattern(
    data = subset(df, (scenario == "No Vaccine" & outbreak == "Large")),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=cases),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_density = pattern_dens*0.5,
    pattern_fill = 'firebrick',
    pattern_colour = NA,
    pattern_angle = 30
  ) +
  
  # Vaccination impact area
  geom_rect(
    data=subset(df_wide, outbreak == "Large"),
    aes(xmin=week, xmax=lead(week), ymin=`With Vaccine`, ymax=`No Vaccine`),
    fill="firebrick", 
    alpha=0.5, 
    inherit.aes=FALSE
  ) +
  
  # Axis and labels
  labs(x=xlabel, y=ylabel) +
  scale_x_continuous(limits=c(0,25)) +
  theme_upper_panel() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    legend.position = "none"
  )

# Combine all plots
# Right panel of Figure 1
combined_plot <- (plt1a / plt1b / plt1c / plt1d) + 
  plot_layout(
    heights = c(0.8, 0.4, 0.4, 0.4), 
    axis_titles = 'collect',
    guides = 'collect'
  )

# Save plot
if (save_plot) {
  fac <- 1.8
  ggsave(paste0("plots/fig1_vacc_eff_schematic_", tstamp(), ".pdf"),
        combined_plot,
        device = cairo_pdf,
        width=1.25*fgw1*fac,
        height=1.6*fgw1*fac,
        units = "mm")
}
```

## Figure 2: Outbreak Description

Outbreak ids by confirmation status
```{r}
paste0(length(unique(outbreak_data_all$id_outbreak)), " outbreaks total")

ids_excl <- outbreak_data_all |> 
  filter(!(id_outbreak %in% outbreak_data$id_outbreak)) |>  
  pull(id_outbreak) |>
  unique()

# Identify confirmed outbreaks (with laboratory confirmation)
ids_conf <- outbreak_data |> 
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique()

# Identify suspected (not confirmed) outbreaks
ids_susp <- outbreak_data |>
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |>
  filter(!(c_ch_sum > 0)) |>
  pull(id_outbreak) |>
  unique()
```

```{r}
# Load and prepare country-region mapping data
country_region <- fread("data/paper/africa_subregion_country_names.csv")
dt <- country_region
country_region <- country_region[Subregion == "Middle Africa", Subregion := "Central Africa"]
# change Middle to Central
country_region %>% mutate()
country_region$iso3c <- sapply(country_region$Country, 
                               function(x) countrycode(x, origin = "country.name.en", 
                                                     destination = "iso3c"))

# Join outbreak data with regional information
outbreak_data$iso3c <- outbreak_data$country
outbreak_data_region <- left_join(outbreak_data, country_region, by = "iso3c")

# =============================================================================
# PLOT 2A: TEMPORAL TRENDS BY SUBREGION (STACKED AREA CHART)
# =============================================================================

# Aggregate data by month and subregion
outbreak_month <- outbreak_data_region %>% 
  mutate(date = as.Date(date),
         date_bin = floor_date(date, "1 month"))

# Calculate total cases per subregion per month
outbreak_month_summary <- outbreak_month %>%
  group_by(date_bin, Subregion) %>%
  summarise(total_cases = sum(s_ch, na.rm = TRUE), .groups = "drop")

# Create complete time series grid to fill gaps
all_subregions <- unique(outbreak_month_summary$Subregion)
all_date_bins <- seq(min(outbreak_month_summary$date_bin), 
                     max(outbreak_month_summary$date_bin), 
                     by = "1 month")
full_grid <- expand_grid(date_bin = all_date_bins, Subregion = all_subregions)

# Merge and fill missing values with zero
df_complete <- full_grid %>%
  left_join(outbreak_month_summary, by = c("date_bin", "Subregion")) %>%
  mutate(total_cases = replace_na(total_cases, 0))

# Create stacked area plot
plt_2a <- ggplot(df_complete, aes(x = date_bin, y = total_cases, 
                                  fill = Subregion)) +
  geom_area(position = "stack", alpha = 0.8) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  scale_y_continuous(labels = comma_format()) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  labs(x = "", y = "Monthly suspected cases", fill = "") +
  theme_light() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    plot.margin = margin(0,0,0,0,"pt")
  )

# =============================================================================
# PLOT 2B: SPATIAL DISTRIBUTION MAP
# =============================================================================
# Prepare outbreak data for mapping
dt <- outbreak_data %>% rename(lctn_pr = location_period_id)
first_rows_df <- dt %>% 
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

# Ensure consistent data types for joining
shp$lctn_pr <- as.character(shp$lctn_pr)

# Join outbreak data with spatial data
map_data <- left_join(first_rows_df, shp, by = "lctn_pr")
map_data %<>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE))

# Convert to spatial features
map_data <- st_as_sf(map_data)

islands_to_exclude <- c("CPV",  # Cape Verde
                        "SYC",  # Seychelles
                        "MYT",  # Mayotte
                        "REU",  # Réunion
                        "SHN")  # Saint Helena, Ascension and Tristan da Cunha

# Exclude these from your Africa shapefile
africa_main <- africa %>% dplyr::filter(!ISO3 %in% islands_to_exclude)
bbox   <- st_bbox(africa_main)       # much tighter now

plt_2b <- ggplot() +
  geom_sf(data = africa_main,  fill = NA, colour = "black", size = .3) +
  geom_sf(data = map_data, aes(colour = is_confirmed, fill = is_confirmed),
          alpha = .7, size = .8) +
  coord_sf(xlim = c(bbox["xmin"], bbox["xmax"]-10),
           ylim = c(bbox["ymin"]+12, bbox["ymax"]),
           expand = FALSE, clip = "on") +    # clip cuts everything outside
  scale_colour_manual("", values = c("steelblue4", "firebrick4"),
                      labels  = c("Not confirmed", "Confirmed")) +
  scale_fill_manual("",   values = c("steelblue4", "firebrick4"),
                    labels  = c("Not confirmed", "Confirmed")) +
  theme_void() +                                   
  theme(legend.position = "top",
        plot.margin     = margin(0,0,0,0,"pt"),
        panel.spacing = unit(0, "pt"))

# plt_2b <- plt_2b + theme(aspect.ratio = 0.7)

# =============================================================================
# PLOT 2C: EPIDEMIC CURVES WITH INSET
# =============================================================================

# Prepare weekly case data
week_sch <- outbreak_data[, c("id_outbreak", "week", "s_ch")]
setDT(week_sch)
max_week <- week_sch[, max(week)]

# Create complete grid for all outbreaks up to max week
all_weeks <- CJ(id_outbreak = unique(week_sch$id_outbreak), week = 1:max_week)

# Merge and fill missing values with zero
week_sch_full <- merge(all_weeks, week_sch, 
                       by = c("id_outbreak", "week"), all.x = TRUE)
week_sch_full[is.na(s_ch), s_ch := 0]

# Calculate summary statistics across all outbreaks
week_sch_summary <- week_sch_full[, .(
  sum_sch = sum(s_ch, na.rm = TRUE),
  median = median(s_ch, na.rm = TRUE),
  p025 = quantile(s_ch, 0.025, na.rm = TRUE),
  p25  = quantile(s_ch, 0.25, na.rm = TRUE),
  p75  = quantile(s_ch, 0.75, na.rm = TRUE),
  p975 = quantile(s_ch, 0.975, na.rm = TRUE)
), by = week]


# Main epidemic curve plot
plt2c <- ggplot(week_sch_summary, aes(x = week)) +
  geom_ribbon(aes(ymin = p025, ymax = p975), fill = set1[3], alpha = 0.45) +
  geom_ribbon(aes(ymin = p25, ymax = p75), fill = set1[3], alpha = 0.6) +
  geom_line(aes(y = median), color = "black", linewidth = 1) +
  labs(title = "", x = "Epidemic week", y = "Suspected cholera cases") +
  scale_x_continuous(limits = c(0, 50), breaks = seq(0, 50, 10)) +
  theme_light() +
  theme(
    plot.margin = margin(0,1,0,1, "pt")
  )

# time to peak
outbk_data_summary %>% 
  summarize(mean_time_to_peak = mean(time_to_peak_week),
            median_time_to_peak = median(time_to_peak_week))

# percentage of samples that are shorter than or equal to 10 weeks
sum(outbk_data_summary$time_to_peak_week <= 10) /
  length(outbk_data_summary$time_to_peak_week)


plt2c_inset <- ggplot(outbk_data_summary, aes(x = time_to_peak_week)) +
  stat_ecdf(geom = "step", size = 0.8) +
  labs(x = "Time to peak (weeks)", 
       y = expression(P(X <= x)))+
  scale_x_continuous(limits=c(0,30), breaks = seq(0,30,10), labels = seq(0,30,10))+
  scale_y_continuous(limits=c(0,1), breaks = seq(0,1,0.5), labels = seq(0,1,0.5))+
  theme_light()+
  theme_tight()+
  theme(plot.margin = margin(0,0,0,0,"pt"))


# Convert inset to grob and add to main plot with better positioning
grob_inset <- ggplotGrob(plt2c_inset)

plt_2c_combined <- plt2c +
  annotation_custom(
    grob = grob_inset,
    xmin = 20, xmax = 50,  # Moved to right side with more space
    ymin = 100, ymax = 340   # Increased height for better visibility
  )

# =============================================================================
# PLOT 2D: OUTBREAK SIZE vs DURATION
# =============================================================================

plt_2d <- outbk_data_summary %>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE)) %>% 
  ggplot() +
  geom_point(aes(x = total_s_ch, y = duration_weeks, color = is_confirmed), 
             alpha = 0.7, size = 1.2) +
  scale_y_log10(labels = comma_format()) +
  scale_x_log10(labels = comma_format()) +
  scale_color_manual("", 
                     labels = c("Not confirmed", "Confirmed"),
                     values = c("steelblue4", "firebrick4")) +
  labs(title = "", x = "Total suspected cases", y = "Outbreak duration (weeks)") +
  theme_light() +
  theme(legend.position = "top", plot.margin = margin(0,1,0,0,"pt"))

# =============================================================================
# PLOT 2E: OUTBREAK SIZE vs ATTACK RATE
# =============================================================================

summary(outbk_data_summary$duration_weeks) 
summary(outbk_data_summary$total_s_ch)
summary(outbk_data_summary$ar_per_10e3)

summary(outbk_data_summary[id_outbreak %in% ids_conf]$duration_weeks)
summary(outbk_data_summary[id_outbreak %in% ids_conf]$total_s_ch)
summary(outbk_data_summary[id_outbreak %in% ids_conf]$ar_per_10e3)

plt_2e <- outbk_data_summary %>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE)) %>% 
  ggplot() +
  geom_point(aes(x = total_s_ch, y = ar_per_10e3, color = is_confirmed), 
             alpha = 0.7, size = 1.2) +
  scale_y_log10(labels = comma_format()) +
  scale_x_log10(labels = comma_format()) +
  scale_color_manual("", 
                     labels = c("Not confirmed", "Confirmed"),
                     values = c("steelblue4", "firebrick4")) +  
  labs(title="", x = "Total suspected cases", y = "Attack rate (per 10,000)") +
  theme_light() +
  theme(legend.position="top",  plot.margin = margin(0,0,0,1,"pt"))



# =============================================================================
# FINAL COMBINED PLOT WITH IMPROVED LAYOUT
# =============================================================================

# Create combined plot with better proportions
# Row 1: Temporal trends (full width)
# Row 2: Map (60%) + Epidemic curve (40%) 
# Row 3: Scatter plots (50% each)
combined_plot <- plt_2a /
  (plt_2b + plt_2c_combined + plot_layout(widths = c(0.6, 1))) /
  (plt_2d + plt_2e + plot_layout(widths = c(1, 1), guides="collect") +
     plot_annotation(theme = theme(legend.position = "top")))+
  plot_layout(heights = c(1,1.3,1)) +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.margin = margin(0,0,0,0,"pt"),
        plot.tag.position = c(0.2, 1),
        plot.tag = element_text(size = 13))

# Display the combined plot
# combined_plot
if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/fig2_outbreak_summary_", tstamp(), ".pdf"), combined_plot,
         device = cairo_pdf, width = fgw2 * fac, height = fgw2 * fac, units = "mm")
}
```

## Figure S1. Indirect Vaccine Effectiveness 

## Figure S2. All outbreaks (eligible and ineligible)

```{r figS1, eval=FALSE}
# Create time series plot with different colors by outbreak type
plt_S1a <- ggplot() +
  geom_line(data = filter(outbreak_data_all, id_outbreak %in% ids_excl),
            aes(date, s_ch, group = id_outbreak, color = "Excluded"), 
            alpha = 0.5) +
  geom_line(data = filter(outbreak_data, !(id_outbreak %in% ids_conf)),
            aes(date, s_ch, group = id_outbreak, color = "Suspected"), 
            alpha = 0.5) +
  geom_line(data = filter(outbreak_data, id_outbreak %in% ids_conf), 
            aes(date, s_ch, group = id_outbreak, color = "Confirmed"),
            alpha = 0.5) +
  scale_color_manual("", values = c("Excluded" = "yellow4",
                                  "Suspected" = "steelblue4", 
                                  "Confirmed" = "firebrick4")) +
  labs(x = "", y = "Weekly suspected cases") +
  theme_light() +
  theme_tight() +
  theme(legend.position = "top")

# (B) Outbreak size vs. duration by type
outbreak_data_all |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                  as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d_all

outbreak_data |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                  as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d

if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1a_outbreak_ts_", tstamp(), ".pdf"), plt_S1a,
         device = cairo_pdf, width = fgw2 * fac, height = fgw1 * fac, units = "mm")
}

plt_S1b <- ggplot() +
  geom_point(data = filter(d_all, id_outbreak %in% ids_excl), 
             aes(duration, total_s_ch, color = "Excluded"), alpha = 0.5) +
  geom_point(data = filter(d, id_outbreak %in% ids_susp), 
             aes(duration, total_s_ch, color = "Suspected"), alpha = 0.5) +
  geom_point(data = filter(d, id_outbreak %in% ids_conf), 
             aes(duration, total_s_ch, color = "Confirmed"), alpha = 0.5) +
  scale_x_log10() +
  scale_y_log10() +
  theme_light() +
  theme_tight() +
  scale_color_manual("", values = c("Excluded" = "yellow4", 
                                  "Suspected" = "steelblue4", 
                                  "Confirmed" = "firebrick4")) +
  labs(x = "Outbreak duration (days)", y = "Total suspected cases") +
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1b_outbreak_size_dur_", tstamp(), ".pdf"), plt_S1b,
         device = cairo_pdf, width = fgw1 * fac, height = fgw1 * fac, units = "mm")
}

# (C) Map of outbreaks
dt <- outbreak_data 
dt_all <- outbreak_data_all 

# Prepare data for mapping
dt <- dt %>% rename(lctn_pr = location_period_id)
dt_all <- dt_all %>% rename(lctn_pr = location_period_id)

first_rows_df <- dt %>% 
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

first_rows_df_all <- dt_all %>%
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

first_rows_df$lctn_pr <- as.numeric(first_rows_df$lctn_pr)
first_rows_df_all$lctn_pr <- as.numeric(first_rows_df_all$lctn_pr)

# Join with spatial data
shp$lctn_pr <- as.character(shp$lctn_pr)
first_rows_df$lctn_pr <- as.character(first_rows_df$lctn_pr)
first_rows_df_all$lctn_pr <- as.character(first_rows_df_all$lctn_pr)

d <- left_join(first_rows_df, shp, by = "lctn_pr")
d_all <- left_join(first_rows_df_all, shp, by = "lctn_pr")

# Create indicator variables for outbreak types
d <- d %>%
  mutate(id_conf = ifelse(id_outbreak %in% ids_conf, 1, 0),
         id_sus = ifelse(!id_outbreak %in% ids_conf, 1, 0))

# Subset data by confirmation status
d_conf <- d %>% filter(id_conf == 1)
d_sus <- d %>% filter(id_sus == 1)

# Convert to spatial data
d_conf <- st_as_sf(d_conf)
d_sus <- st_as_sf(d_sus)
d_all <- d_all %>% mutate(id_excl = ifelse(id_outbreak %in% ids_excl, 1, 0))
d <- st_as_sf(d)
d_all <- st_as_sf(d_all)

# Create map
plt_S1c <- ggplot() +
  geom_sf(data = africa, fill = NA, color = "black") +  # Country outlines
  geom_sf(data = d_all, aes(fill = id_excl), color = "black", 
          fill = "yellow4", alpha = 0.5) +
  geom_sf(data = d_sus, aes(fill = id_sus), color = "steelblue4", 
          fill = "steelblue4", alpha = 0.5) +
  geom_sf(data = d_conf, aes(fill = id_conf), color = "firebrick4", 
          fill = "firebrick4", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Map of Cholera Outbreaks")


if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1c_outbreak_map_", tstamp(), ".pdf"), plt_S1c,
         device = cairo_pdf, width = fgw1 * fac, height = fgw1 * fac, units = "mm")
}
```

## Figure 3: Impact across delays and prioritization

# Summary values in results section
```{r fig3, eval=FALSE}
onset_impact <- cea_ori %>%
  filter(vacc_cov %in% c(0.6, 0.75, 0.9), vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, 
                      pct_reduc_case, 
                      case_averted_per_1000_OCV, 
                      daly_averted_per_1000_OCV,
                      cost_per_case_averted, 
                      cost_per_daly_averted)

# cases averted (CA) for vaccine coverage of 75% with ORI in week 1
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", 
         vacc_week == 1, vacc_cov == 0.75) %>%
  format_median_iqr()
# percent case averted (PCA) at week 1
onset_impact %>% 
  filter(variable == "pct_reduc_case", 
         vacc_week == 1, vacc_cov == 0.75) %>%
  format_median_iqr()

# CA by vaccine coverage in week 1
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", vacc_week == 1) %>%
  group_by(vacc_cov) %>% 
  format_median_iqr()
# CA by vaccine coverage in week 9
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", vacc_week == 9) %>%
  group_by(vacc_cov) %>% 
  format_median_iqr()

# PCA by vaccine coverage in week 9
onset_impact %>% 
  filter(variable == "pct_reduc_case", vacc_week == 9) %>%
  group_by(vacc_cov) %>% 
  format_median_iqr()


# outbrreaks lasting >=7 weeks
sum(outbk_data_summary$duration_weeks >= 7) / nrow(outbk_data_summary)

quantile(fit_decay[["9"]]$r_squared, probs=c(0.25,0.5,0.75))

# instantaneous decay rate
fit_decay[["9"]] %>% 
  summarize_quantiles(k_hat) %>% 
  format_median_iqr(digits=2)

# weekly reduction rate
fit_decay[["9"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# half_life                  
fit_decay[["9"]] %>%
  summarize_quantiles(half_life) %>% 
  format_median_iqr(digits=1)

# weekly reduction rate for all outbreaks
fit_decay[["5"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# weekly reduction rate for all outbreaks
fit_decay[["14"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# weekly reduction rate for all outbreaks
fit_decay[["21"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# outbrreaks lasting >=7 weeks
sum(outbk_data_summary$duration_weeks >= 19) / nrow(outbk_data_summary)

# linear decrease in proportion of ORI-eligible outbreaks
round(abs(coef(fit_lm)[[2]] * 100), digits=1)
# observed at week 9
1 - fit_lm$model %>% filter(vacc_week == 9) %>% pull(prop_included_dur)
# predicted at week 9
1 - predict(fit_lm, newdata=data.frame(vacc_week=9))

cea_prioritization %>%
  mutate(prob_i = as.integer(round(100 * prob))) %>% 
  filter(metric == "med_s_ch_averted_tot", 
         percentile_by =="size", prob_i %in% c(0, 90), 
         vacc_week %in% c(1,9)) %>% 
  select(vacc_week, prob_i, median_value)

capd <- cea_prioritization %>%
  mutate(prob_i = as.integer(round(100 * prob))) %>% 
  filter(metric == "med_case_averted_per_1000_OCV", 
         percentile_by =="ar", prob_i %in% c(0, 90), 
         vacc_week %in% c(1,9)) %>% 
  select(vacc_week, prob_i, median_value)

# fold increase
capd$median_value[3] / capd$median_value[1]
capd$median_value[4] / capd$median_value[2]

dalyapd <- cea_prioritization %>%
  mutate(prob_i = as.integer(round(100 * prob))) %>% 
  filter(metric == "med_daly_averted_per_1000_OCV", 
         percentile_by =="ar", prob_i %in% c(0, 90), 
         vacc_week %in% c(1,9)) %>% 
  select(vacc_week, prob_i, median_value)

# fold increase
dalyapd$median_value[3] / dalyapd$median_value[1]
dalyapd$median_value[4] / dalyapd$median_value[2]

cea_prioritization %>%
  mutate(prob_i = as.integer(round(100 * prob))) %>% 
  filter(metric == "med_pct_reduc_case", 
         prob_i %in% c(90), 
         vacc_week %in% c(1,9)) %>% 
  select(vacc_week, percentile_by, median_value)

# Cost per case averted (CPCA) for ORI with vaccine coverage of 75%
onset_impact %>% 
  filter(variable == "cost_per_case_averted", 
         vacc_week %in% c(1,9), vacc_cov == 0.75) %>%
  format_median_iqr(digits=0)

onset_impact %>% 
  filter(variable == "cost_per_daly_averted", 
         vacc_week %in% c(1,9), vacc_cov == 0.75) %>%
  format_median_iqr(digits=0)

# gdp data from World Bank during 2010-2020
gdp_data <- read_xls("data/paper/GDP_WorldBank.xls")
gdp_2010_2020 <- gdp_data %>% 
  filter(`Country Name` == "Sub-Saharan Africa (IDA & IBRD countries)") %>%
  select(c(`2010`:`2020`)) |> 
  as.numeric() 

# PCA in the top quartile by duration
cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_dur_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(pct_reduc_case) %>%  
  select(vacc_week,q250,q500,q750)

# Cost per DALY averted
cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_10_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted) %>%  
  select(vacc_week,q250,q500,q750) -> cpdalya_top10
# ratio to the median gdp
round(cpdalya_top10$q500/median(gdp_2010_2020), digits=1) 

# Cost per DALY averted
cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted) %>%  
  select(vacc_week,q250,q500,q750) -> cpdalya

# % reduction in cost per DALY averted 
100*(1 - cpdalya_top10$q500/cpdalya$q500)

# exponential decay
fit_decay_ct[["1"]] %>%
  summarize_quantiles(r_squared) %>% 
  format_median_iqr(digits=2)

# instantaneous decay rate across cases
fit_decay_ct[["1"]] %>%
  summarize_quantiles(k_hat) %>% 
  format_median_iqr(digits=4)

# % reduction for the 10 additional missing cases for ORIs implemented in week 1
fit_decay_ct[["1"]] %>% 
  mutate(rate_reduc_10 = 1 - exp(-k_hat*10)) %>% 
  summarize_quantiles(rate_reduc_10*100) %>% 
  format_median_iqr(digits=1)

# % reduction for the 10 additional missing cases for ORIs implemented in week 1
fit_decay_ct[["1"]] %>% 
  mutate(rate_reduc_10 = 1 - exp(-k_hat*100)) %>% 
  summarize_quantiles(rate_reduc_10*100) %>% 
  format_median_iqr(digits=1)

# half_life                  
fit_decay_ct[["1"]] %>%
  summarize_quantiles(half_life) %>% 
  format_median_iqr(digits=1)

# % reduction for the 10 additional missing cases for ORIs implemented in week 1
fit_decay_ct[["5"]] %>% 
  mutate(rate_reduc_10 = 1 - exp(-k_hat*10)) %>% 
  summarize_quantiles(rate_reduc_10*100) %>% 
  format_median_iqr(digits=1)

# % reduction for the 100 additional missing cases for ORIs implemented in week 1
fit_decay_ct[["5"]] %>% 
  mutate(rate_reduc_10 = 1 - exp(-k_hat*100)) %>% 
  summarize_quantiles(rate_reduc_10*100) %>% 
  format_median_iqr(digits=1)
# half_life                  
fit_decay_ct[["5"]] %>%
  summarize_quantiles(half_life) %>% 
  format_median_iqr(digits=1)
```

### Left panel

Percent cases averted

```{r}
ymean  <- 95 
ymedian<- 87.5
y50    <- c(82, 93)
y95    <- c(75, 100)

VW_plot_seq <- seq(1,15,2)
# percentage case averted 
p1 <- cea_ori_pca %>%
  filter(vacc_week %in% VW_plot_seq) %>%
  ggplot(aes(x = vacc_week, color = as.factor(vacc_cov), 
             group = as.factor(vacc_cov))) +
  ci_layers_from_summary(
    x_discrete = TRUE,
    dodge_width = 1.4,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = 2.4,
    mean_size = x_size,
    mean_stroke = x_stroke
  ) +
  add_summary_braces(
    x0 = 13,
    ymean = ymean, ymedian = ymedian,
    y50 = y50, y95 = y95,
    pt_size_median = 3,
    pt_size_mean = x_size,
    pt_stroke_mean = x_stroke,
    dx50 = 0.5, dx95 = 0.8,
    text_50_x = 13.6,
    text_50_y = y50[2],
    text_95_x = 14.85,
    text_95_y = y95[2],
    brace_50_x = 13.2,
    brace_95_x = 14.4,
    text_offset_mean = 0.3,
    lw_50 = 0.6, lw_95 = 0.6,
    whisker_50 = line_width_50, 
    whisker_95 = line_width_95
  ) +
  scale_y_continuous("% cases averted",
                     labels = scales::percent_format(scale = 1), 
                     limits=c(0,100)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[1:3],
    labels = c(paste0(format_num(VC3*100, 1), "%")))+
  theme_light() +
  theme_tight() +
  theme_x_blank()

# plot 2
# exponential decay curve
set.seed(2025)
runid_sample <- 1
n_sample <- 20

x_range <- seq(1, 15, by = 0.2) # x-axis grid
t0 <- 1  # baseline week for normalization
## ---- normalized "spaghetti" using fitted params ----
# use fits longer than 9 weeks
ids_wk <- fit_decay$`9` %>% 
  filter(id_outbreak %in% unique(fit_decay$`9`$id_outbreak)) %>%
  pull(id_outbreak) %>% unique()

id_outbreak_sample_wk <- sample(ids_wk, size = n_sample)

fitted_models <- fit_decay$`9` %>% 
  filter(runid %in% runid_sample,
         id_outbreak %in% id_outbreak_sample_wk)

## ---- normalized "spaghetti" using fitted params ----
# On the normalized scale, A0 cancels; z_hat(t) = exp(-k * (t - t0))
spaghetti_data_norm_wk <- fitted_models %>%
  mutate(vacc_week = list(x_range)) %>%
  unnest(vacc_week) %>%
  mutate(
    # z_hat = exp(-decay_rate_k * (vacc_week - t0))
    z_hat = exp(- k_hat * (vacc_week - t0))
  )

spaghetti_data_norm <- spaghetti_data_norm_wk
id_outbreak_sample <- id_outbreak_sample_wk

## ---- raw data, normalized by week-1 value per series ----
# Expecting 'cea_ori' with columns: vacc_cov, ori_occurred, vacc_week, runid, id_outbreak, pct_reduc_case
raw_data_points_norm <- cea_weekly_delay_2_ori %>%
  filter(vacc_cov == 0.75,
         vacc_week %in% 1:15, 
         runid %in% runid_sample,
         id_outbreak %in% id_outbreak_sample) %>%
  arrange(runid, id_outbreak, vacc_week) %>%
  group_by(runid, id_outbreak) %>%
  mutate(
    y0 = first(pct_reduc_case),
    z  = ifelse(y0 > 0, pct_reduc_case / y0, NA_real_)
  ) %>%
  ungroup()

## ---- summary curves from k only (median + IQR) ----
# Use the full data set
# k_q <- quantile(fit_decay$`5`$decay_rate_k, 
#                 probs = c(0.025, .25, .5, .75, 0.975), na.rm = TRUE)
k_q <- quantile(fit_decay$`9`$k_hat, 
                probs = c(0.025, .25, .5, .75, 0.975), na.rm = TRUE)
k_q025 <- k_q[[1]]; k_q25 <- k_q[[2]]; k_med <- k_q[[3]]; 
k_q75 <- k_q[[4]]; k_q975 <- k_q[[5]]

summary_curves_norm <- tibble(
  vacc_week = x_range,
  med = exp(-k_med * (x_range - t0)),
  lo  = exp(-k_q75 * (x_range - t0)),  # faster decay => lower envelope
  hi  = exp(-k_q25 * (x_range - t0)),   # slower decay => higher envelope
  q975   = exp(-k_q975 * (x_range - t0)),  # faster decay => lower envelope
  q025  = exp(-k_q025 * (x_range - t0))   # slower decay => higher envelope
)

## ---- pretty label (normalized) ----
format_k <- paste0(round(k_med, digits = 3))
eq_label_norm <- paste0("$\\hat{z}(t)=\\exp\\{-", format_k, "(t-1)\\}$")

## ---- plot (normalized scale 0-1) ----
p2 <- ggplot() +
  # Individual normalized fitted curves
  geom_line(
    data = spaghetti_data_norm,
    aes(x = vacc_week, y = z_hat, group = id_outbreak, color = id_outbreak),
    alpha = 0.60
  ) +
  # Raw points, normalized by their own week-1
  geom_point(
    data = raw_data_points_norm,
    aes(x = vacc_week, y = z, color = id_outbreak),
    alpha = 0.60, size = 1
  ) +
  # Median + IQR summary on normalized scale
  geom_ribbon(
    data = summary_curves_norm,
    aes(x = vacc_week, ymin = lo, ymax = hi),
    inherit.aes = FALSE,
    fill = "firebrick",
    alpha = 0.30
  ) +
  geom_ribbon(
    data = summary_curves_norm,
    aes(x = vacc_week, ymin = q025, ymax = q975),
    inherit.aes = FALSE,
    fill = "firebrick",
    alpha = 0.20
  ) +
  geom_line(
    data = summary_curves_norm,
    aes(x = vacc_week, y = med),
    color = "firebrick",
    linewidth = 1.2
  ) +
  annotate(
    "latex",
    x = 15, y = 1,
    label = eq_label_norm,
    size = 4.5, hjust = 1, vjust = 1
  ) +
  # scale_y_continuous("Normalized reduction", limits = c(0, 1)) +
  scale_y_continuous("Normalized reduction") +
  guides(color = "none") +
  theme_light() +
  theme_tight() +
  theme_x_blank()

# plot 3
co <- coef(fit_lm)
alpha <- unname(co[1])   # intercept
beta  <- unname(co[2])   # slope (expected to be negative)
r2    <- summary(fit_lm)$r.squared

# --- prediction grid & values ---
x_grid <- seq(min(fit_lm$model$vacc_week), 
              max(fit_lm$model$vacc_week), length.out = 200)
pred   <- predict(fit_lm, newdata = data.frame(vacc_week = x_grid))

paste0("Eveary week, additional ", 
       format_num(abs(100*beta), digits=1), 
       "% outbreaks ended before ORI")
wk <- 9
paste0("Week ", wk, ", ", 
       format_num(100*(1-pred[min(which(x_grid>wk))]), digits=1), 
       "% outbreaks ended before ORI")
# --- a tidy equation label using LaTeX (percent scale) ---
# round for display
alpha_lab <- format(round(alpha, 2), nsmall = 2)
beta_lab  <- format(round(beta, 3), nsmall = 3)
r2_lab    <- format(round(r2, 3), nsmall = 3)

eq_label <- paste0(
  "$\\hat{y}(t) = ", alpha_lab, beta_lab, "\\, t$"
)

# add 1 for weeks 1-3
df <- data.frame(vacc_week = 1:3, 
                 prop_included_dur=1)
df <- rbind(df, fit_lm$model)

p3 <- ggplot() +
  # raw data (black)
  geom_point(
    data = df,
    aes(vacc_week, prop_included_dur),
    color = "black", size = 2
  ) +
  # fitted line (firebrick)
  geom_line(
    aes(x_grid, pred),
    color = "firebrick", linewidth = 1.2
  ) +
  # equation label (LaTeX)
  annotate(
    "latex",
    x = 15, y = 1,
    label = eq_label,
    size = 4,
    hjust = 1, vjust = 1
  ) +
  # scale_x_continuous("Delay in ORI implementation (weeks)", 
  #                    breaks = seq(1, 15, 2), limits=c(1,15)) +
  scale_y_continuous("Proportion included") +
  theme_light() +
  theme_tight()

```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- p1 +
  annotate("text", x = -Inf, y = Inf, label = "A",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2), limits=c(0.2,15.5))+
  coord_cartesian(clip = "off")

p2_tag <- p2 +
  annotate("text", x = -Inf, y = Inf, label = "B",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2), limits=c(0.2,15.5))+
  coord_cartesian(clip = "off")

p3_tag <- p3 +
  annotate("text", x = -Inf, y = Inf, label = "C",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Week of ORI implementation",
                     breaks=seq(1,15,2), limits=c(0.2,15.5))+
  coord_cartesian(clip = "off")

plts_pca <- (p1_tag / p2_tag / p3_tag) + plot_layout(ncol=1)

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig3_onset_triggered_left_", tstamp(), ".pdf"),
         plts_pca,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### Right panel

```{r}
# Create the color palette function (your code was correct here)
myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))

y_labs <- c("Attack rate percentile", "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
                   "Prioritizing longer outbreaks",
                   "Prioritizing larger outbreaks",
                   "Prioritizing higher CFR outbreaks")

plot_data <- cea_prioritization %>% filter(metric == "med_s_ch_averted_tot")

plot_data %>%
  filter(vacc_cov == 0.75, vacc_week == 9, percentile_by == "size") %>%
  group_by(prob) %>% 
  summarize(med_case_averted = median_value) 

# Create the color palette function (your code was correct here)
myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))

fill_limits <- c(3, 1750)
brks <- c(10, 100, 1000)

legend_title <- "Cases averted" 
pct_by <- c("ar", "duration", "size", "cfr")

plts_ca <- vector('list', 4)

q_labels <- c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))
i <- 3
plt_1  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = q_labels) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    annotate(
      "segment",
      x = 9, y = 0,
      xend = c(9, 1), # End points for the three arrows
      yend = c(0.9, 0),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4
    ) +
    annotate(
      "point",
      x = c(9), # The four key coordinates
      y = c(0),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "segment",
      x = 8, y = 0.5,
      xend = c(8, 1), # End points for the three arrows
      yend = c(0.9, 0.5),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4,
      linetype = "dashed" # Changed to dashed lines
    ) +
    annotate(
      "point",
      x = c(8), # The four key coordinates
      y = c(0.5),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "text",
      x = c(9.5, 5), # x-positions for the two labels
      y = c(0.5, 0.05), # y-positions for the two labels
      label = c(y_annots[i], "Reducing ORI delay"),
      angle = c(90, 0), # Rotate the vertical label
      color = "black",
      # fontface = "bold",
      size = 4,
      hjust = 0.5
    ) +
    labs(
      x = "Week of ORI implementation",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme_x_blank() +
    theme(legend.position = "top")

# Cases averted per 1,000 OCVs
plot_data <- cea_prioritization %>% 
  filter(metric == "med_case_averted_per_1000_OCV")

fill_limits <- c(0.01, 10)
brks <- c(1, 5, 10)

legend_title <- "Cases averted per 1,000 OCVs" 
pct_by <- c("ar", "duration", "size", "cfr")


i <- 1
plt_2  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = q_labels) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    annotate(
      "segment",
      x = 9, y = 0,
      xend = c(9, 1), # End points for the three arrows
      yend = c(0.9, 0),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4
    ) +
    annotate(
      "point",
      x = c(9), # The four key coordinates
      y = c(0),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "segment",
      x = 8, y = 0.5,
      xend = c(8, 1), # End points for the three arrows
      yend = c(0.9, 0.5),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4,
      linetype = "dashed" # Changed to dashed lines
    ) +
    annotate(
      "point",
      x = c(8), # The four key coordinates
      y = c(0.5),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "text",
      x = c(9.5, 5), # x-positions for the two labels
      y = c(0.5, 0.05), # y-positions for the two labels
      label = c(y_annots[i], "Reducing ORI delay"),
      angle = c(90, 0), # Rotate the vertical label
      color = "black",
      # fontface = "bold",
      size = 4,
      hjust = 0.5
    ) +
    labs(
      x = "Week of ORI implementation",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme_x_blank() +
    theme(legend.position = "top")

# DALYs averted per 1,000 OCVs
plot_data <- 
  cea_prioritization %>% filter(metric == "med_daly_averted_per_1000_OCV")

fill_limits <- c(1e-5, 1.3)
brks <- c(0.1, 0.5, 1)
legend_title <- "DALYs averted per 1,000 OCVs" 

i <- 1
plt_3  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = q_labels) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    annotate(
      "segment",
      x = 9, y = 0,
      xend = c(9, 1), # End points for the three arrows
      yend = c(0.9, 0),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4
    ) +
    annotate(
      "point",
      x = c(9), # The four key coordinates
      y = c(0),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "segment",
      x = 8, y = 0.5,
      xend = c(8, 1), # End points for the three arrows
      yend = c(0.9, 0.5),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4,
      linetype = "dashed" # Changed to dashed lines
    ) +
    annotate(
      "point",
      x = c(8), # The four key coordinates
      y = c(0.5),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "text",
      x = c(9.5, 5), # x-positions for the two labels
      y = c(0.5, 0.05), # y-positions for the two labels
      label = c(y_annots[i], "Reducing ORI delay"),
      angle = c(90, 0), # Rotate the vertical label
      color = "black",
      # fontface = "bold",
      size = 4,
      hjust = 0.5
    ) +
    labs(
      x = "Week of ORI implementation",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(legend.position = "top")
```

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plt_1 +
  annotate("text", x = -Inf, y = Inf, label = "D",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

p2_tag <- plt_2 +
  annotate("text", x = -Inf, y = Inf, label = "E",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

p3_tag <- plt_3 +
  annotate("text", x = -Inf, y = Inf, label = "F",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Week of ORI implementation",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

plts_right <- (p1_tag / p2_tag / p3_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")
```

### Combine left and right

```{r}
plts_pca_capd <- (plts_pca | plts_right) + 
  plot_layout(ncol=2) + 
  plot_layout(axis_titles = "collect_x") &
  theme(
    legend.position = "top", # put legends at top
    legend.justification = "center", # center horizontally
    legend.box.just = "center",
    legend.box = "horizontal", # arrange horizontally
    legend.text          = element_text(size = 13),
    legend.title         = element_text(size = 13),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 3, l = 0)
  )

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig3_pca_capd_", tstamp(), ".pdf"),
         plts_pca_capd,
         device = cairo_pdf, width = fgw2*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

## Figure S3

Median impact data with 75% coverage

### A Number of cases averted

```{r}
plot_data <- cea_prioritization %>% filter(metric == "med_s_ch_averted_tot")

plot_data %>% 
  filter(vacc_cov == 0.75) %>%
  pull(median_value) %>% range()

# 1. Create the color palette function (your code was correct here)
myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))

y_labs <- c("Attack rate percentile", 
            "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
              "Prioritizing longer outbreaks",
              "Prioritizing larger outbreaks",
              "Prioritizing higher CFR outbreaks")

fill_limits <- c(3, 1750)
brks <- c(10, 100, 1000)

legend_title <- "Cases averted" 
pct_by <- c("ar", "duration", "size", "cfr")

plts_ca <- vector('list', 4)

for (i in 1:4) {
  plts_ca[[i]]  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    labs(
      x = "Delay in ORI implementation (weeks)",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(
      legend.position = "top",
    ) 
}
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plts_ca[[1]] +
  annotate("text", x = -Inf, y = Inf, label = "A",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p2_tag <- plts_ca[[2]] +
  # annotate("text", x = -Inf, y = Inf, label = "B",
  #          hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p3_tag <- plts_ca[[3]] +
  # annotate("text", x = -Inf, y = Inf, label = "C",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p4_tag <- plts_ca[[4]] +
  # annotate("text", x = -Inf, y = Inf, label = "D",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Week of ORI implementation",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

plts_combined <- (p1_tag / p2_tag / p3_tag / p4_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")+
  plot_layout(guides = "collect") &                    
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/ca_delay_pri_", tstamp(), ".pdf"),
         plts_combined,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### B. Percent cases averted

```{r}
plot_data <- cea_prioritization %>% filter(metric == "med_pct_reduc_case")

plot_data %>% 
  filter(vacc_cov == 0.75) %>%
  pull(median_value) %>% range()

myPalette <- colorRampPalette(viridisLite::plasma(9))

y_labs <- c("Attack rate percentile", "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
                   "Prioritizing longer outbreaks",
                   "Prioritizing larger outbreaks",
                   "Prioritizing higher CFR outbreaks")

fill_limits <- c(2, 85)
brks <- c(20, 40, 60, 80)

legend_title <- "% cases averted" 
pct_by <- c("ar", "duration", "size", "cfr")

plts_pca <- vector('list', 4)

for (i in 1:4) {
  plts_pca[[i]]  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      # trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    labs(
      x = "Delay in ORI implementation (weeks)",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(
      legend.position = "top",
    ) 
}
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plts_pca[[1]] +
  annotate("text", x = -Inf, y = Inf, label = "B",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p2_tag <- plts_pca[[2]] +
  # annotate("text", x = -Inf, y = Inf, label = "B",
  #          hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p3_tag <- plts_pca[[3]] +
  # annotate("text", x = -Inf, y = Inf, label = "C",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous( breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p4_tag <- plts_pca[[4]] +
  # annotate("text", x = -Inf, y = Inf, label = "D",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Delay in ORI implementation (weeks)",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")


plts_combined <- (p1_tag / p2_tag / p3_tag / p4_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")+
  plot_layout(guides = "collect") &                    
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/pca_delay_pri_", tstamp(), ".pdf"),
         plts_combined,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### C. Cases averted per 1,000 OCVs

```{r}
plot_data <- 
  cea_prioritization %>% filter(metric == "med_case_averted_per_1000_OCV")

plot_data %>% 
  filter(vacc_cov == 0.75) %>%
  pull(median_value) %>% range()

myPalette <- colorRampPalette(viridisLite::cividis(9))

y_labs <- c("Attack rate percentile", "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
                   "Prioritizing longer outbreaks",
                   "Prioritizing larger outbreaks",
                   "Prioritizing higher CFR outbreaks")

fill_limits <- c(0.01, 10)
brks <- c(1, 5, 10)

legend_title <- "Cases averted per 1,000 OCVs" 
pct_by <- c("ar", "duration", "size", "cfr")

plts_capd <- vector('list', 4)

for (i in 1:4) {
  plts_capd[[i]]  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    labs(
      x = "Delay in ORI implementation (weeks)",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(
      legend.position = "top",
    ) 
}

```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plts_capd[[1]] +
  annotate("text", x = -Inf, y = Inf, label = "C",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p2_tag <- plts_capd[[2]] +
  # annotate("text", x = -Inf, y = Inf, label = "B",
  #          hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p3_tag <- plts_capd[[3]] +
  # annotate("text", x = -Inf, y = Inf, label = "C",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous( breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p4_tag <- plts_capd[[4]] +
  # annotate("text", x = -Inf, y = Inf, label = "D",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Delay in ORI implementation (weeks)",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")


plts_combined <- (p1_tag / p2_tag / p3_tag / p4_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")+
  plot_layout(guides = "collect") &                    
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/capd_delay_pri_", tstamp(), ".pdf"),
         plts_combined,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### D. DALYs averted per 1,000 OCVs

```{r}
plot_data <- 
  cea_prioritization %>% filter(metric == "med_daly_averted_per_1000_OCV")

plot_data %>% 
  filter(vacc_cov == 0.75) %>%
  pull(median_value) %>% range()

# Create the color palette function (your code was correct here)
# myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))
# myPalette <- colorRampPalette(viridisLite::plasma(9))
# myPalette <- colorRampPalette(viridisLite::cividis(9))
myPalette <- colorRampPalette(scico::scico(9, palette = "batlow"))
# myPalette <- colorRampPalette(RColorBrewer::brewer.pal(9, "YlGnBu"))
# myPalette <- colorRampPalette(viridisLite::inferno(9))
# myPalette <- colorRampPalette(scico::scico(9, palette = "lajolla"))

y_labs <- c("Attack rate percentile", "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
                   "Prioritizing longer outbreaks",
                   "Prioritizing larger outbreaks",
                   "Prioritizing higher CFR outbreaks")

fill_limits <- c(1e-5, 1.3)
brks <- c(0.1, 0.5, 1)

legend_title <- "DALYs averted per 1,000 OCVs" 
pct_by <- c("ar", "duration", "size", "cfr")

legend_title <- "DALYs averted per 1,000 OCVs" 

plts_dapd <- vector('list', 4)

for (i in 1:4) {
  plts_dapd[[i]]  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))) +
      
    scale_x_continuous(expand = c(0, 0), 
                       breaks = seq(1, 15, 2)) +
    labs(
      x = "Delay in ORI implementation (weeks)",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(
      legend.position = "top",
    ) 
}
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plts_dapd[[1]] +
  annotate("text", x = -Inf, y = Inf, label = "D",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p2_tag <- plts_dapd[[2]] +
  # annotate("text", x = -Inf, y = Inf, label = "B",
  #          hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p3_tag <- plts_dapd[[3]] +
  # annotate("text", x = -Inf, y = Inf, label = "C",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous( breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p4_tag <- plts_dapd[[4]] +
  # annotate("text", x = -Inf, y = Inf, label = "D",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Delay in ORI implementation (weeks)",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")


plts_combined <- (p1_tag / p2_tag / p3_tag / p4_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")+
  plot_layout(guides = "collect") &                    
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/dalyapd_delay_pri_", tstamp(), ".pdf"),
         plts_combined,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

## Figure 4: Cumulative impact during OCV stockpile deployment

Plotting data
```{r}
# summary by subset_id in addition to the grouping variables # 
# grouping variables \in c("vacc_cov","vacc_week","target_vacc_doses",
# "vacc_priority","priority_prob", "subset_id","pp_r")
# columns you pivot on / keep
keys <- c("vacc_cov","vacc_week","target_vacc_doses",
          "vacc_priority","priority_prob","subset_id","pp_r")
value_col <- c("doses_used", "outbreak_count", 
               "s_ch_averted_tot",
               "death_averted_tot",
               "daly_averted_tot", 
               "case_averted_per_1000_OCV",
               "death_averted_per_1000_OCV",
               "daly_averted_per_1000_OCV", 
               "cost_per_case_averted", 
               "cost_per_death_averted", 
               "cost_per_daly_averted")

# --- configuration for the four scenarios ---
x_prob  <- 0.3                # “top 30%” means keep >= 70th percentile
x_pct   <- as.integer(x_prob * 100)
vars_3  <- c("AR","dur","size")

# optional: a tiny helper to map to the exact labels you want
make_scenario_label <- function(vacc_priority, pp_r) {
  if (pp_r == 0) return("All")
  if (pp_r == x_prob && vacc_priority %in% vars_3) {
    switch(vacc_priority,
      "AR"   = sprintf("Q[%d%%](AR)",   x_pct),
      "dur"  = sprintf("Q[%d%%](Dur)",  x_pct),
      "size" = sprintf("Q[%d%%](Size)", x_pct)
    )
  } else NA_character_
}

# filter to the four scenarios and build the plotting factor
d <- cea_subsets %>%
  mutate(pp_r = round(priority_prob, 1)) %>%
  filter(
    # vacc_week == VW,
    vacc_cov  == VC,
    value_column %in% value_col,
    # keep either "All" (pp=0) for AR (pick any) OR the x=70% slice for the three variables
    (pp_r == 0 & vacc_priority  == "AR") | (pp_r == x_prob & vacc_priority %in% vars_3)
  ) %>%
  mutate(
    scenario = mapply(make_scenario_label, vacc_priority, pp_r)
  ) %>%
  filter(!is.na(scenario)) %>%
  mutate(
    # order in legend: AAR, dur, size, All
    scenario = factor(
      scenario,
      levels = c(sprintf("Q[%d%%](AR)", x_pct),
                 sprintf("Q[%d%%](Dur)", x_pct),
                 sprintf("Q[%d%%](Size)", x_pct),
                 "All")
    )
  )
unique(d[,c("pp_r", "vacc_priority")])
# split and join to bring doses_used alongside outcomes
dr <- d %>%
  filter(value_column == value_col[1]) %>%
  select(any_of(keys), scenario, q500)

names(dr) <- c(keys, "scenario", "doses_used")

dw <- d %>%
  filter(value_column %in% value_col[2:length(value_col)]) %>%
  left_join(dr, by = c(keys, "scenario"))
levs <- levels(dw$scenario)  # expects: Q[30%](AR), Q[30%](Dur), Q[30%](Size), All
unique(dw[,c("vacc_priority","pp_r","scenario")])

expr_labels <- setNames(
  list(
    bquote(Q[30*"%"](AR)),
    bquote(Q[30*"%"](Dur)),
    bquote(Q[30*"%"](Size)),
    expression(All)
  ),
  c(paste0("Q[", x_pct, "%](AR)"), 
    paste0("Q[", x_pct, "%](Dur)"),
    paste0("Q[", x_pct, "%](Size)"),
    "All")
)
# 
# # set.seed(42)
# # n_subset_plot <- 30
# # ids <- sample(200, n_subset_plot) # just to reduce plotting load
# ids <- 1:200
# # create nudges to prevent too much overlap between subset id's
# dw[, doses_used_plot := doses_used * rnorm(length(dw$doses_used), 
#                                          mean=1, sd=1.3e-1)]
v_week <- c(1, 9)
alph <- 0.4

# dw_7 <- dw
```

Summary values in the main text
```{r}
# outbreak count
# outbreak count does not vary within each subset (while impact metrics may vary across runid's) and therefore, you could select any values among q025, q250, ..., q975 and mean.
dw %>%
  filter(value_column == "outbreak_count", 
         vacc_week == 1, 
         scenario == "All") %>%
  group_by(target_vacc_doses) %>% 
  summarise(
    across(c(q500),
           list(q500 = ~quantile(.x, probs=0.5),
                q250 = ~quantile(.x, probs=0.25),
                q750 = ~quantile(.x, probs=0.75),
                min = ~min(.x),
                max = ~max(.x))),
    .groups = "drop"
  )

dw %>%
  filter(value_column == "outbreak_count", 
         vacc_week == 1, 
         target_vacc_doses == 200000000) %>%
  group_by(scenario) %>% 
  summarise(
    across(c(q500),
           list(q500 = ~quantile(.x, probs=0.5),
                q250 = ~quantile(.x, probs=0.25),
                q750 = ~quantile(.x, probs=0.75),
                min = ~min(.x),
                max = ~max(.x))),
    .groups = "drop"
  )

# cases averted 
ca <- dw %>%
  filter(value_column == "s_ch_averted_tot") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

# in the order of Q[30%](AR), Q[30%](Dur), Q[30%](Size), and All 
ca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 0)

ca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits=0)

capd <- dw %>%
  filter(value_column == "case_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

capd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 3)

capd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 3)

# compare with the analyses at the individual level
# median values are smaller at the individual level
cea_ori %>% 
  filter(vacc_week %in% c(1,9), vacc_cov == 0.75) %>% 
  group_by(vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV) 

# variation of vaccine impact and efficiency
cea_ori %>% 
  filter(vacc_week %in% c(1,9), vacc_cov == 0.75) %>% 
  group_by(vacc_week) %>%
  summarize(min = min(pct_reduc_case),
            max = max(pct_reduc_case))

cea_ori %>% 
  filter(vacc_week %in% c(1,9), vacc_cov == 0.75) %>% 
  group_by(vacc_week) %>%
  summarize(min = min(case_averted_per_1000_OCV),
            max = max(case_averted_per_1000_OCV))

dapd <- dw %>%
  filter(value_column == "daly_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

dapd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 3)

dapd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 3)

cpca <- dw %>%
  filter(value_column == "cost_per_case_averted") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

cpca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 0)

cpca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 0)

# cost per DALY averted
cpda <- dw %>%
  filter(value_column == "cost_per_daly_averted") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

cpda %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 0)

cpda %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 0)
```

### Left panel

Plotting parameters
```{r}
pt_size <- 1.35

make_log_offsets_span <- 
  function(x, group, levels = NULL, base = 10, span_factor = 1.5) {
    if (!is.null(levels)) group <- factor(group, levels = levels, ordered = TRUE)
    k <- nlevels(factor(group))
    if (k == 1) return(x)
    step_dex <- log(span_factor, base = 10) / (k - 1) # per-step dex from total span
    idx <- as.integer(factor(group)) - (k + 1) / 2
    x * (base^(step_dex * idx))
  }
span_fctr <- 1.4
shp_wk1 <- 17
shp_wk9 <- 1
```

Number of targeted outbreaks
```{r}
ocv_dose_2020 <- global_ocv_use$cumulative_icg[1] * 1e6
ocv_dose_2024 <- global_ocv_use$cumulative_icg[2] * 1e6

dose_marks <- tibble::tibble(
  x = c(ocv_dose_2020, ocv_dose_2024),
  lab = c("2020", "2024")
)

cnt_iqr_summary <- dw %>%
  filter(value_column == "outbreak_count") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

cnt_iqr_summary_off <- cnt_iqr_summary %>%
  filter(vacc_week == 1) %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_1 <- cnt_iqr_summary_off %>% 
  filter(vacc_week == 1) %>%
  ggplot(aes(x = x_off, 
             y = q500, color = scenario)) +
  geom_linerange(aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(aes(y=q500), size=pt_size, shape=1) +
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  geom_text(
    data = dose_marks,
    inherit.aes = FALSE,
    aes(x = x, y = Inf, label = lab),
    vjust = 1.5,           # push slightly down from the top
    hjust = 1.1,
    nudge_x = 0,           # no horizontal nudge
    size = 3
  )+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_log10("Number of targeted outbreaks") +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  theme_light() +
  theme_x_blank() +
  theme(legend.position = "top") +
  theme_tight_legend(box_space_pt = 3.5,
                     key_space_cm = 0.03)
# plt_1
```

Total suspected cases averted (s_ch_averted_tot)
```{r}
df <- data.frame(x=sort(unique(dw$doses_used_plot)))
df$y <- df$x

ca_iqr_summary <- dw %>%
  filter(value_column == "s_ch_averted_tot") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # `q500` is also used as a variable inside the summarize_quantile() function
  summarize_quantiles(x)

ca_iqr_summary_off <- ca_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_2 <- 
  ggplot(ca_iqr_summary_off, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(ca_iqr_summary_off, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=ca_iqr_summary_off, 
             aes(y=q500, shape=as.factor(vacc_week)), size=pt_size)+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_log10("Total cases averted", 
                breaks = c(1e1,1e2,1e3,1e4,1e5,1e6),
                labels = c(1e1,1e2,1e3,1e4,1e5,1e6)) +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  scale_shape_manual("",
    values = c("1" = shp_wk1, "9" = shp_wk9),
    labels = c("1 week", "9 weeks")
  )+
  theme_light() +
  theme_x_blank() +
  guides(color = "none", shape="none")

# plt_2
```

Case averted per 1,000 OCV doses (case_averted_per_dose)

```{r}
capd_iqr_summary <- dw %>%
  filter(value_column == "case_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

capd_iqr_summary <- capd_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_3 <- 
  ggplot(capd_iqr_summary, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(capd_iqr_summary, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=capd_iqr_summary, 
             aes(y=q500, shape=as.factor(vacc_week)), size=pt_size)+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_continuous("Case averted per 1,000 OCVs", 
                     breaks=c(0:5,7.5,10,12.5,15),
                     labels=c(0:5,"",10,"",15),
                     minor_breaks = NULL) +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  scale_shape_manual("",
    values = c("1" = shp_wk1, "9" = shp_wk9),
    labels = c("1 week", "9 weeks")
  )+
  theme_light() +
  guides(color = "none", shape="none")

# plt_3
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5
# x <- 0 because of log scale
# A, B, and C annotations
plt_1_tag <- plt_1 +
  annotate("text", x = 0, y = Inf, label = "A",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_log10("")+
  coord_cartesian(clip = "off")


plt_2_tag <- plt_2 +
  annotate("text", x = 0, y = Inf, label = "B",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_log10("")+
  coord_cartesian(clip = "off")

plt_3_tag <- plt_3 +
  annotate("text", x = 0, y = Inf, label = "C",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_log10("Vaccine doses deployed")+
  coord_cartesian(clip = "off")

plts_ca_agg <- (plt_1_tag / plt_2_tag / plt_3_tag) +
  plot_layout(ncol = 1) &
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig4_agg_left_", tstamp(), ".pdf"),
         plts_ca_agg,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### Right panel

DALY averted per 1,000 OCV doses (case_averte_per_dose)

```{r}
dapd_iqr_summary <- dw %>%
  filter(value_column == "daly_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

dapd_iqr_summary <- dapd_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_4 <- 
  ggplot(dapd_iqr_summary, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(dapd_iqr_summary, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=dapd_iqr_summary, 
             aes(y=q500, shape=as.factor(vacc_week)),
             size=pt_size)+
  geom_text(
    data = dose_marks,
    inherit.aes = FALSE,
    aes(x = x, y = Inf, label = lab),
    vjust = 1.5,           # push slightly down from the top
    hjust = 1.1, 
    nudge_x = 0,           # no horizontal nudge
    size = 3
  )+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_continuous("DALY averted per 1,000 OCVs") +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  scale_shape_manual("",
    values = c("1" = shp_wk1, "9" = shp_wk9),
    labels = c("week 1", "week 9")
  )+
  theme_light() +
  guides(color = "none")+
  theme_x_blank()+
  theme(legend.position = "top") +
  theme_tight_legend(box_space_pt = 3,
                     key_space_cm = 0.03)

# plt_4
```

Cost per case averted (cost_per_case_averted)

```{r}
cpca_iqr_summary <- dw %>%
  filter(value_column == "cost_per_case_averted") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

cpca_iqr_summary <- cpca_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_5 <- 
  ggplot(cpca_iqr_summary, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(cpca_iqr_summary, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=cpca_iqr_summary, 
             aes(y=q500, shape=as.factor(vacc_week)), size=pt_size)+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_log10("Cost per case averted") +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
   scale_shape_manual("",
    values = c("1" = 17, "9" = 1),
    labels = c("1 week", "9 weeks")
  )+
  theme_light() +
  theme_x_blank() +
  guides(color = "none", shape="none")

# plt_5
```

Cost per DALY averted (cost_per_daly_averted)
```{r}
cpda_iqr_summary <- dw %>%
  filter(value_column == "cost_per_daly_averted") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

cpda_iqr_summary <- cpda_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_6 <-
  ggplot(cpda_iqr_summary, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(cpda_iqr_summary, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=cpda_iqr_summary, 
             aes(y=q500, shape=as.factor(vacc_week)), size=pt_size)+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_log10("Cost per DALY averted") +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  scale_shape_manual("",
    values = c("1" = shp_wk1, "9" = shp_wk9),
    labels = c("1 week", "9 weeks")
  )+
  theme_light() +
  guides(color = "none", shape="none")
 
# plt_6
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5
# x <- 0 because of log scale
# A, B, and C annotations
plt_4_tag <- plt_4 +
  annotate("text", x = 0, y = Inf, label = "D",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_log10("")+
  coord_cartesian(clip = "off")

plt_5_tag <-  plt_5 +
  annotate("text", x = 0, y = Inf, label = "E",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_log10("")+
  coord_cartesian(clip = "off")

plt_6_tag <-  plt_6 +
  annotate("text", x = 0, y = Inf, label = "F",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_log10("Vaccine doses deployed")+
  coord_cartesian(clip = "off")

plts_cpca_agg <- (plt_4_tag / plt_5_tag / plt_6_tag) + 
  plot_layout(ncol = 1) &
  theme(legend.position = "top",
        legend.box = "vertical",
        legend.spacing.y = unit(3,"pt")) # <-- space between the two legend boxes)

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig4_agg_right_", tstamp(), ".pdf"),
         plts_cpca_agg,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### Combine

Left and right panels
```{r}
plts_pca_cpca <- (plts_ca_agg | plts_cpca_agg) + 
  plot_layout(ncol=2) + 
  plot_layout(axis_titles = "collect_x") &
  theme(
    legend.position = "top", # put legends at top
    legend.justification = "center", # center horizontally
    legend.box.just = "center",
    legend.box = "horizontal", # arrange horizontally
    legend.text          = element_text(size = 13),
    legend.title         = element_text(size = 13),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 3, l = 0)
  )
   
update_geom_defaults("point", list(size = 4))
leg_pt <- guide_legend(override.aes = list(size = 4))
# Optional: larger legend symbols in the final combined figure
plts_pca_cpca <- plts_pca_cpca + guides(color = leg_pt, shape = leg_pt)

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig4_pca_cpca_agg_", tstamp(), ".pdf"),
         plts_pca_cpca,
         device = cairo_pdf, width = fgw2*fac,
         height = fgw2*fac*fac, units = "mm")
}
```


# Result Tables

## Table 3. Impact under the outbreak onset-triggered scenario

All and top 10% by attack rate, duration, and size
```{r table3}
# Create summary for time-triggered vaccination at different times
cea_ori[(vacc_cov == VC & vacc_week %in% c(1, 13))] %>% 
  group_by(vacc_week) %>% 
  summarize_quantiles(case_averted_per_1000_OCV) -> capd

format_median_iqr(capd, row=1, digits=3)
format_median_iqr(capd, row=2, digits=3)

cea_ori[(vacc_cov == VC & vacc_week %in% c(1, 13))] %>% 
  group_by(vacc_week) %>% 
  summarize_quantiles(cost_per_case_averted) -> cpca

format_median_iqr(cpca, row=1, digits=0)
format_median_iqr(cpca, row=2, digits=0)

##
wks <- c(1, 5, 9, 13)
tab <- data.frame()

for (wk in wks) {
  d <- cea_ori[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp)
}

# top 10% in attack rate
dat <- cea_ori %>% filter(id_outbreak %in% top_10_ar_ids)
for (wk in wks) {
  d <- dat[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp) 
}

# top 10% in duration
dat <- cea_ori %>% filter(id_outbreak %in% top_10_dur_ids)
for (wk in wks) {
  d <- dat[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp)
}

# top 10% in size
dat <- cea_ori %>% filter(id_outbreak %in% top_10_size_ids)
for (wk in wks) {
  d <- dat[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp)
}

tab$week <- rep(wks, 4)
tab$group <- rep(c("All","Top 10% attack rate",
                    "Top 10% duration","Top 10% size"), each = 4)

# format number of outbreaks with percentages
tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/length(unique(cea$id_outbreak))*100, digits=1), "%)") 

if (save_table) {
  fwrite(tab, 
         paste0("outputs/paper/table3_onset_triggered_", 
                tstamp(), ".csv"), 
            row.names = FALSE)
}
```

## Table S1. Impact under the outbreak onset-triggered scenario

Broader vaccine coverage and ORI implementation delays

```{r tableS1}
# Create summary for time-triggered vaccination at different times
wks <- c(1, 5, 9, 13)
tab <- data.frame()

for (vc in VC3) {
  for (wk in wks) {
    d <- cea_ori[(vacc_cov == vc & vacc_week == wk)]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
    tab <- rbind(tab, tab_temp)
  }  
}
tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/length(unique(cea$id_outbreak))*100, digits=1), "%)")

tab$week <- rep(wks, length(VC3))
tab$vc <- rep(VC3, each=length(wks))

if (save_table) {
  fwrite(tab, paste0("outputs/paper/tableS1_onset_triggered_vc_vw_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```

## Table 4. Case-triggered scenario

```{r table4}
# Create summary for case-triggered vaccination at different thresholds
vwks <- c(1, 9)
ctgs <- c(1, 10, 50, 140)
tab <- data.frame()

for (wk in vwks) {
  for (ctg in ctgs) {
    d <- cea_ct_ori[(vacc_cov == VC & vacc_week == wk & 
                       case_trigger == ctg)]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
    tab <- rbind(tab, tab_temp)
  }
}

tab$case_trigger <- rep(ctgs, length(vwks))
tab$vacc_week <- rep(vwks, each=length(ctgs))

tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/nrow(outbk_data_summary)*100, digits=1), "%)")

if (save_table) {
  fwrite(tab, paste0("outputs/paper/table4_case_triggered_", tstamp(), ".csv"),
            row.names = FALSE)
}
```

## Table S2. Onset-triggered ORIs across varying delays to stop ORI after outbreak end

```{r}
# Create summary for case-triggered vaccination at different thresholds
vwks <- c(1, 5, 9, 13)
lags <- c(1, 3, 5, 7)
tab <- data.frame()

for (wk in vwks) {
  for (lag in lags) {
    d <- cea_weekly[(ori_occurred & 
                     vacc_week == wk & 
                       no_ori_delay_outbreak_end == lag)]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
    tab <- rbind(tab, tab_temp)
  }
}

tab$no_ori_lag <- rep(lags, length(vwks))
tab$vacc_week <- rep(vwks, each=length(lags))

tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/length(unique(cea_weekly$id_outbreak))*100, digits=1), "%)")

if (save_table) {
  fwrite(tab, paste0("outputs/paper/tableS2_no_ori_lag_", tstamp(), ".csv"),
            row.names = FALSE)
}
```

## Table S3: Onset-triggered ORIs across alternative outbreak dataset (expanding size limit)

```{r tableS1}
# 75% vaccine coverage only
unique(cea_alt_data$vacc_cov)

wks <- c(1, 5, 9, 13)
tab <- data.frame()
cea_alt_ori <- cea_alt_data[(ori_occurred == TRUE)]

for (wk in wks) {
  d <- cea_alt_ori[vacc_week == wk]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp)
}

tab$week <- wks
# format number of outbreaks with percentages
tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/length(unique(cea_alt_data$id_outbreak))*100, digits=1), "%)") 
if (save_table) {
  fwrite(tab, paste0("outputs/paper/tableS3_alt_data_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```



# Discussion

```{r}
impact <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,5,9,13)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, pct_reduc_case, 
                      case_averted_per_1000_OCV,
                      daly_averted_per_1000_OCV,
                      cost_per_case_averted, 
                      cost_per_daly_averted)


impact %>% filter(variable == "pct_reduc_case") 
impact %>% filter(variable == "case_averted_per_1000_OCV")
impact %>% filter(variable == "daly_averted_per_1000_OCV")
impact %>% filter(variable == "cost_per_case_averted")
impact %>% filter(variable == "cost_per_daly_averted")
# 9 weeks or longer as being representative 
fit_decay[["9"]] %>%
  mutate(rate_reduc = 100*(1 - exp(- k_hat))) %>% 
  summarize_group(vars=c("rate_reduc"), group_name="> 9 weeks") %>% 
  format_median_iqr(digits=1)

fit_decay[["9"]] %>% 
  summarize_group(vars=c("half_life"), group_name="> 9 weeks") %>% 
  format_median_iqr(digits=1)

impact %>% filter(variable == "pct_reduc_case") 
# top quartile

# top quartile in attack rate
cea_ori %>% 
  filter(vacc_cov %in% c(0.75), 
         vacc_week %in% c(5)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(pct_reduc_case)

cea_ori %>% 
  filter(id_outbreak %in% top_q_dur_ids, 
         vacc_cov %in% c(0.75), 
         vacc_week %in% c(13)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(pct_reduc_case)

cea_ori %>% 
  filter(vacc_cov %in% c(0.75), 
         vacc_week %in% c(1,5,9,13)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV)

cea_ori %>% 
  filter(id_outbreak %in% top_q_ar_ids, 
         vacc_cov %in% c(0.75), 
         vacc_week %in% c(1,5,9,13)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV)
# dw from Figure 4
ca <- dw %>%
  filter(value_column == "s_ch_averted_tot") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

ca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 0)

ca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 0)

capd <- dw %>%
  filter(value_column == "case_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

capd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 3)

capd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 3)


# Cost per DALY averted (CPDALYA)
cea_ori %>% 
  filter(vacc_cov %in% c(0.75), 
         vacc_week %in% c(1)) %>%
  summarize_quantiles(cost_per_daly_averted)

# CPDALYA in 
cea_ori %>% 
  filter(id_outbreak %in% top_q_ar_ids, 
         vacc_cov %in% c(0.75), 
         vacc_week %in% c(1)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted)  
  

```
