---
title: "Cholera Vaccine Impact Modeling: Visualization and Tables"
author: "Jong-Hoon Kim"
date: "2025-10-31"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

# Introduction

This document creates visualizations and tables to summarize the results from 
cholera vaccine impact simulations. 

# Setup

```{r setup}
#| warning: false
#| message: false

# Load required packages
source("R/utils.R")  # utility functions
library(data.table)
library(tidyverse)
library(magrittr)
library(purrr)
library(broom)
library(effsize)
library(randomForest)
library(rstatix)
library(ggridges)
library(readxl)
library(janitor)
library(scales)
library(viridis)
library(sf)
library(metR)
library(googlesheets4)
library(MASS)
library(patchwork)
library(ggpattern)
library(gridExtra)
library(corrr)
library(RColorBrewer)
library(countrycode)
library(conflicted)
library(xdvir)

# Resolve package conflicts in favor of dplyr
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflicts_prefer(data.table::first)
conflicts_prefer(ggplot2::margin)
# Output control flags
save_plot <- FALSE
save_table <- FALSE

# Figure dimensions
figure_size <- data.table(
  journal = c("Lancet"),
  single = c(85),
  double = c(178)
)
fgw1 <- figure_size[journal == "Lancet", single]
fgw2 <- figure_size[journal == "Lancet", double]

# Color palette for consistent visualization
set1 <- brewer.pal(9, "Set1")  # color sets
# my_discrete_colors <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00")

# Define standard probability values for quantiles
PROBS <- c(0.025, 0.25, 0.5, 0.75, 0.975)  # 2.5%, 25%, 50%, 75%, 97.5% quantiles

# Reference vaccination week
VW <- 9  # representative vaccination week from GTFCC data

# Sequence of vaccination weeks to evaluate
VW_seq <- seq(1, 15, 2)

# Three representative vaccination coverage rates
VC3 <- c(0.6, 0.75, 0.9)  # 60%, 75%, 90%

# Reference vaccination coverage rate
VC <- VC3[2]  # 75% coverage

# Thresholds for case-triggered vaccination to analyze
cts_sel <- c(1, seq(10, 100, 10))  # case triggers selected

# Latest date for which OCV campaign data is available
YEAR <- as.Date("2020-12-28")

# Load geographic data for mapping
shp <- st_read("data/paper/total_shp_0427.shp", quiet = TRUE)
africa <- st_read("data/paper/afr_g2014_2013_0.shp", quiet = TRUE)

# only a subset of sample runid's is used for plotting for convenience
set.seed(42)
# sample_runids <- sample(1:200, size=30)

# Plot parameters
line_width_50 <- 1.2
line_width_95 <- 0.5
line_interval <- 0.7
point_size <- 0.5
x_size <- 1.5
x_stroke <- 1
```

# Load data

```{r load-data}
#| warning: false
# Load outbreak data
# All outbreaks including those affected by OCVs
outbreak_data_all <- 
  as.data.frame(fread("outputs/paper/outbreak_data_all_20250228.csv"))

# Outbreaks included in vaccine impact modeling (not affected by OCVs)
outbreak_data <- fread("outputs/paper/outbreak_data_20250228.csv") 

# outbreaks filtered
# This includes outbreaks that are finally analyzed and excludes
# outbreaks that were duplicated
outbreaks_filtered <-
  readRDS("outputs/paper/outbreak_summary_filtered_20250418.rds")

outbreak_data <- outbreak_data %>%
  filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 
# outbreaks excluded
outbreaks_excluded <- read_rds("outputs/paper/outbreaks_excluded_20250612.rds")

# Historical OCV deployment data
# data were collected based on the website: https://apps.epicentre-msf.org/public/app/gtfcc
# The number of doses shipped by 2020 and 2024 can be obtained by adjusting the shipment period
global_ocv_use <-
  clean_names(fread("data/paper/OCV_Use_GTFCC_Dashboard_20251031.csv"))

# Load model parameters
dis_params <- as.data.frame(fread("data/paper/parameters.csv"))

# Load Bayesian beta regression results for the indirect vaccine effectiveness
ive_fit <- readRDS("outputs/paper/ive_yhat_yrep_20250919.rds")
ive_data <- ive_fit$yrep
rm(ive_fit)  # Free memory by removing large object

# Load outbreak summary data
ds <- clean_names(as.data.frame(fread("data/paper/outbreak_files_Oct2024/outbreak_summary.csv")))
# message(paste0(length(unique(ds$country)), " countries"))

# Create unique outbreak identifiers
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# Join outbreak data with summary information
outbk_data_summary <- 
  left_join(
    filter(outbreak_data, week == 1), 
    ds[, c("outbreak_pop", "total_s_ch", "total_c_ch", "total_deaths", "duration_weeks", 
           "peak_week_incidence", "time_to_peak_week", "peak_week_incidence_per_100percent", 
           "peak_week_incidence_per_10e3", "mean_r0_1st_week", "area_per_1km2", 
           "rural_urban", "ar_per_10e3", "cfr_per_100percent", "id_outbreak")], 
    by = "id_outbreak"
  )

outbk_data_summary %<>% 
  filter(id_outbreak %in% outbreaks_filtered$id_outbreak)

outbk_data_summary <- fread("outputs/paper/outbreak_data_summary_20251022.csv")
# Summarize attack rates and population sizes
# summary(outbk_data_summary$ar_per_10e3)
# summary(outbk_data_summary$outbreak_pop)

# outbreaks in the top quartile by attack rate (ar), duration (dur), and size
n25 <- floor(0.25 * nrow(outbk_data_summary))

top_q_ar_ids <- outbk_data_summary %>% 
  arrange(desc(ar_per_10e3)) %>% slice(1:n25) %>% pull(id_outbreak)
top_q_dur_ids <- outbk_data_summary %>%
  arrange(desc(duration_weeks)) %>% slice(1:n25) %>% pull(id_outbreak)
top_q_size_ids <- outbk_data_summary %>% 
  arrange(desc(total_s_ch)) %>% slice(1:n25) %>% pull(id_outbreak)


n10 <- floor(0.1 * nrow(outbk_data_summary))

top_10_ar_ids <- outbk_data_summary %>% 
  arrange(desc(ar_per_10e3)) %>% slice(1:n10) %>% pull(id_outbreak)
top_10_dur_ids <- outbk_data_summary %>%
  arrange(desc(duration_weeks)) %>% slice(1:n10) %>% pull(id_outbreak)
top_10_size_ids <- outbk_data_summary %>% 
  arrange(desc(total_s_ch)) %>% slice(1:n10) %>% pull(id_outbreak)
```

# Load vaccine impact simulation results

```{r load-results}
# Outbreak onset-triggered vaccination results
cea_files <- c("outputs/paper/cea_onset_trigger_20260612.rds")
cea <- map_dfr(cea_files, read_rds)
cea_ori <- cea %>% filter(ori_occurred == TRUE)

# percentage case averted summary for Figure 3
cea_ori_pca <- read_rds("outputs/paper/cea_pca_summary_20251031.rds")
# impact summary for Figure 3
cea_prioritization <-
  read_rds("outputs/paper/summary_impact_0.75_prioritization_20251031.rds")

cea_weekly <- read_rds("outputs/paper/cea_delay_onset_trigger_20260612.rds")
cea_weekly_delay_2 <- 
  cea_weekly %>% filter(no_ori_delay_outbreak_end == 2)

cea_weekly_delay_2_ori <- cea_weekly %>% 
  filter(no_ori_delay_outbreak_end == 2, ori_occurred == TRUE)

cea_alt_data <- 
  read_rds("outputs/paper/cea_alt_data_onset_trigger_20260612.rds")

# weekly decline
fit_decay <- read_rds("outputs/paper/fit_exp_decay_20251031.rds")
fit_decay_ct <- 
  read_rds("outputs/paper/fit_exp_decay_case_trigger_20251031.rds")
fit_lm <- read_rds("outputs/paper/fit_lm_20251031.rds")
# Case-triggered vaccination results
cea_ct <- read_rds("outputs/paper/cea_case_trigger_20260612.rds")
cea_ct_ori <- cea_ct %>% filter(ori_occurred == TRUE)
# outbreak subsets for given number of OCVs
subset_files <- c(
  "outputs/paper/cea_summary_subsets_vw_1_20260619.rds",
  "outputs/paper/cea_summary_subsets_vw_9_20260619.rds"
)
cea_subsets <- map_dfr(subset_files, read_rds)
```

# Abstract

Values used in the abstract
```{r}
onset_impact <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, 
                      pct_reduc_case, 
                      case_averted_per_1000_OCV, 
                      daly_averted_per_1000_OCV,
                      cost_per_case_averted, 
                      cost_per_daly_averted)

# cases averted (CA) for vaccine coverage of 75% with ORI in week 1
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", vacc_week == 1) %>%
  format_median_iqr()
# percent case averted (PCA) at week 1
onset_impact %>% 
  filter(variable == "pct_reduc_case", vacc_week == 1) %>%
  format_median_iqr()
# CA At week 9
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", vacc_week == 9) %>%
  format_median_iqr()
# PCA At week 9
onset_impact %>% 
  filter(variable == "pct_reduc_case", vacc_week == 9) %>%
  format_median_iqr()

# weekly reduction rate
fit_decay[["9"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# half_life                  
fit_decay[["9"]] %>%
  summarize_quantiles(half_life) %>% 
  format_median_iqr(digits=1)

# case averted per 1000 OCVs (CAPD) at week 1
onset_impact %>% 
  filter(variable == "case_averted_per_1000_OCV", vacc_week == 1) %>%
  format_median_iqr(digits=2)
onset_impact %>% 
  filter(variable == "cost_per_case_averted", vacc_week == 1) %>%
  format_median_iqr(digits=0)

onset_impact %>% 
  filter(variable == "case_averted_per_1000_OCV", vacc_week == 9) %>%
  format_median_iqr(digits=2)
onset_impact %>% 
  filter(variable == "cost_per_case_averted", vacc_week == 9) %>%
  format_median_iqr(digits=0)

# fold increase in case averted per ORI by prioritization
ca <- onset_impact %>% 
  filter(variable == "s_ch_averted_tot") %>% 
  pull(q500)

ca_top_q_size <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_size_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, pct_reduc_case) %>% 
  filter(variable == "s_ch_averted_tot") %>% 
  pull(q500)

paste0("fold increase: ", ca_top_q_size / ca)

# fold increase in case averted per 1000 OCVs
capd_top_q_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV) %>% 
  pull(q500)

# fold increase in case averted per 1000 OCVs
capd_top_10_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_10_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV) %>% 
  pull(q500)

capd <- onset_impact %>% 
  filter(variable == "case_averted_per_1000_OCV") %>% 
  pull(q500) 

paste0("fold increase in case averted per 1,000 OCVs: ", 
       capd_top_q_ar / capd)

paste0("fold increase in case averted per 1,000 OCVs: ", 
       capd_top_10_ar / capd)

# fold increase in averted per 1000 OCVs
# top quartile
dapd_top_q_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(daly_averted_per_1000_OCV) %>%  
  pull(q500)

# top 10%
dapd_top_10_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_10_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(daly_averted_per_1000_OCV) %>%  
  pull(q500)

dapd <- onset_impact %>% 
  filter(variable == "daly_averted_per_1000_OCV") %>% 
  pull(q500) 

paste0("fold increase in DALY averted per 1,000 OCVs: ",
       dapd_top_q_ar / dapd)


paste0("fold increase in DALY averted per 1,000 OCVs: ",
       dapd_top_10_ar / dapd)

# cost per DALY averted
# top quartile
cpda_top_q_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted) %>% 
  pull(q500)
# top 10%
cpda_top_10_ar <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_10_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted) %>% 
  pull(q500)

cpda <- onset_impact %>% 
  filter(variable == "cost_per_daly_averted") %>% 
  pull(q500) 

paste0("fold decrease in cost per DALY averted: ", cpda/cpda_top_q_ar)
paste0("fold decrease in cost per DALY averted: ", cpda/cpda_top_10_ar)
# subset
cea_subsets %>% 
  mutate(prob_i = as.integer(round(100*priority_prob))) %>% 
  filter(value_column == "case_averted_per_1000_OCV", vacc_week == 9, 
         prob_i == 0) %>% 
  summarize_quantiles(q500) %>% 
  format_median_iqr(digits=2)

cea_subsets %>% 
  mutate(prob_i = as.integer(round(100*priority_prob))) %>% 
  filter(value_column == "case_averted_per_1000_OCV", vacc_week == 1, 
         vacc_priority == "AR", prob_i == 30) %>% 
  summarize_quantiles(q500) %>% 
  format_median_iqr(digits=2)
```

# Introduction

Values used in the abstract
```{r}
# number of OCVs doses shipped from OCV stockpile (million doses)
global_ocv_use
# proportion approved by ICG (reactive) by 2024
global_ocv_use$cumulative_icg[2] / global_ocv_use$cumulative_total[2]
# outbreaks span from 2010 to 2020
unique(outbk_data_summary$year)
outbk_data_summary %>% filter(year == 2020)
```

# Plots

## Figure 1. Outbreaks and % case reduction 

The right panel of Figure 1 in the main text was generated using the following code. The left panel, a flowchart illustrating outbreak identification and selection for vaccine impact modeling, was created in PowerPoint.

```{r}
# Global parameters ----------------------------------------------------
text_size <- 5          # Default text size for annotations
line_height <- 0.95     # Line spacing for text annotation
pattern_dens <- 0.1     # Density of pattern filling
pattern_spacing <- 0.1  # Spacing between patterns
step_linewidth <- 0.7   # Width of step lines
arrow_vertical_len <- 14 # Vertical length of arrows

# Axis labels
ylabel <- "Weekly suspected cholera cases"
xlabel <- "Weeks since outbreak onset"

# Key timeline parameters
week_or <- 3      # Outbreak response at week 3  
week_vax <- 6.5   # Vaccination started
week_vax_eff <- 8.5 # Vaccine takes effect

# Rectangle dimensions configuration ------------------------------------
# This section makes it easy to adjust all rectangles at once

# Base rectangle dimensions
rect_width <- 0.8
rect_height <- 6

# Spacing between rectangles
gap_horizontal <- 1.9   # Horizontal gap between rectangle groups
gap_vertical <- 12      # Vertical gap between top and bottom rectangles

# Font sizes for elements
eq_symbol_size <- 8     # Size for equation symbols (<, +)
paren_size <- 9         # Size for parentheses
plus_size <- 5          # Size for plus symbols

# Vertical offset for second equation from first
eq2_vertical_offset <- 24

# Helper functions ----------------------------------------------------
create_filled_square <- function(x1, y1, x2, y2, fill_color, 
                                 color='black', alpha=0.5) {
  geom_rect(xmin=x1, xmax=x2, ymin=y1, ymax=y2,
            color=color, fill=fill_color, alpha=alpha, inherit.aes = FALSE)
}

create_striped_square <- function(x1, y1, x2, y2, fill_color, color='black',
                                  dens=0.1, angle=30) {
  geom_rect_pattern(
    xmin=x1, xmax=x2, ymin=y1, ymax=y2,
    pattern = 'stripe',
    fill = NA,
    colour = color,
    pattern_density = dens,
    pattern_spacing = 0.025,
    pattern_fill = fill_color,
    pattern_colour = NA,
    pattern_angle = angle,
    inherit.aes = FALSE)
}

create_divider <- function(x1, x2, y, extension=0.2, linewidth=0.7) {
  annotate("segment", 
           x = x1 - extension, 
           y = y, 
           xend = x2 + extension, 
           yend = y, 
           linewidth = linewidth)
}

theme_upper_panel <- function() {
  theme_classic() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 16),
      axis.text.y = element_text(size = 14),
      plot.caption = element_text(size = 14),
      legend.position = "none"
    )
}

# Data preparation ----------------------------------------------------
# Create synthetic example of an outbreak trajectory with and without vaccination
d <- data.frame(
  date = seq(as.Date("2010-06-15"), by = "week", length.out = 24),
  week = 0:23,
  # No vaccination scenario
  s_ch = c(2, 2, 3, 4, 10, 12, 29, 100, 120, 150, 140, 
           75, 68, 47, 24, 5, 12, 5, 3, 2, 4, 1, 0, 0)
)

# Add vaccination scenario (40% of cases after vaccine takes effect)
d$s_ch_vacc <- c(
  2, 2, 3, 4, 10, 12, 29, 100, 120, 
  round(0.4 * c(150, 140, 75, 68, 47, 24, 5, 12, 5, 3, 2, 4, 1, 0, 0))
)
d$s_ch_cumul <- cumsum(d$s_ch)

# Rename columns for clarity
names(d) <- c("date", "week", "No vacc", "Vacc", "No vacc cumulative")

# Reshape data for plotting
dlong <- pivot_longer(d, cols = c("No vacc", "Vacc"))

# Define time weeks for additional scenarios
weeks <- 0:25

# Generate additional outbreak scenarios
# Smallest outbreak (ended before vaccination)
smallest_no_vax <- rpois(length(weeks), lambda=30 * exp(-0.1 * (weeks - 2)^2))
smallest_no_vax[(week_vax-1):length(smallest_no_vax)] <- 0

# Smaller outbreak (ended at vaccine takes effect)
smaller_no_vax <- rpois(length(weeks), lambda=50 * exp(-0.1 * (weeks - 3)^2))
smaller_no_vax[week_vax_eff:length(smaller_no_vax)] <- 0

# Small outbreak (peak at week 8)
small_no_vax <- rpois(length(weeks), lambda=70 * exp(-0.08 * (weeks - 8)^2))
small_with_vax <- small_no_vax
small_with_vax[weeks > week_vax_eff] <- round(small_with_vax[weeks > week_vax_eff] * 0.5)  # 50% reduction

# Large outbreak (peak at week 12)
large_no_vax <- rpois(length(weeks), lambda=120 * exp(-0.04 * (weeks - 12)^2))
large_with_vax <- large_no_vax
large_with_vax[weeks > week_vax_eff] <- round(large_with_vax[weeks > week_vax_eff] * 0.5)  # 50% reduction

# Combine into a data frame
df <- tibble(
  week = rep(weeks, 6),
  cases = c(smallest_no_vax, smaller_no_vax, small_no_vax, small_with_vax, 
            large_no_vax, large_with_vax),
  outbreak = rep(c("Smallest", "Smaller", "Small", "Small", "Large", "Large"), 
                 each = length(weeks)),
  scenario = c(rep("No Vaccine", 2*length(weeks)), 
               rep(c("No Vaccine", "With Vaccine"), times=2, each=length(weeks)))
)

# Create wide version of the data
df_wide <- df %>%
  pivot_wider(names_from = scenario, values_from = cases) %>%
  mutate(reduction = `No Vaccine` - `With Vaccine`)

# Rectangle positions for equation 1 ----------------------------------
# Starting position for first equation
base_x1 <- 15
base_y1 <- 65

# a/b fraction
a_x1 <- base_x1
a_y1 <- base_y1
a_x2 <- a_x1 + rect_width
a_y2 <- a_y1 + rect_height

b_x1 <- a_x1
b_y1 <- a_y1 - gap_vertical
b_x2 <- b_x1 + rect_width
b_y2 <- b_y1 + rect_height

ab_div_y <- (a_y1 + b_y2) / 2

# c/d fraction
c_x1 <- a_x1 + gap_horizontal
c_y1 <- a_y1
c_x2 <- c_x1 + rect_width
c_y2 <- c_y1 + rect_height

d_x1 <- c_x1
d_y1 <- b_y1
d_x2 <- d_x1 + rect_width
d_y2 <- d_y1 + rect_height

cd_div_y <- (c_y1 + d_y2) / 2

# Inequality symbol
eq1_lt_x <- (a_x2 + c_x1) / 2
eq1_lt_y <- (a_y1 + b_y2) / 2

# Equation label
eq1_label_x <- a_x1 - 1.2
eq1_label_y <- cd_div_y

# Rectangle positions for equation 2 ----------------------------------
# Left side: 1/2(a/b+c/d)
# a/b part
eq2_a_x1 <- a_x1 + 0.2
eq2_a_y1 <- a_y1 - eq2_vertical_offset
eq2_a_x2 <- eq2_a_x1 + rect_width
eq2_a_y2 <- eq2_a_y1 + rect_height

eq2_b_x1 <- eq2_a_x1
eq2_b_y1 <- eq2_a_y1 - gap_vertical
eq2_b_x2 <- eq2_b_x1 + rect_width
eq2_b_y2 <- eq2_b_y1 + rect_height

# c/d part
eq2_c_x1 <- eq2_a_x1 + gap_horizontal
eq2_c_y1 <- eq2_a_y1
eq2_c_x2 <- eq2_c_x1 + rect_width
eq2_c_y2 <- eq2_c_y1 + rect_height

eq2_d_x1 <- eq2_c_x1
eq2_d_y1 <- eq2_b_y1
eq2_d_x2 <- eq2_d_x1 + rect_width
eq2_d_y2 <- eq2_d_y1 + rect_height

# Parentheses
left_paren_x <- eq2_a_x1 - 0.3
right_paren_x <- eq2_c_x2 + 0.3
paren_y <- (eq2_a_y1 + eq2_b_y2) / 2

# 1/2 fraction position
eq2_frac_x <- right_paren_x + 0.6
eq2_frac_y_num <- (eq2_a_y1 + eq2_a_y2)/2
eq2_frac_y_denom <- (eq2_b_y1 + eq2_b_y2)/2
eq2_frac_y_line <- (eq2_a_y1 + eq2_b_y2)/2

# Plus symbol position
eq2_plus1_x <- (eq2_a_x2 + eq2_c_x1) / 2
eq2_plus1_y <- (eq2_a_y1 + eq2_b_y2) / 2

# Less than symbol for second equation
eq2_lt_x <- eq2_frac_x + 0.8
eq2_lt_y <- (eq2_a_y1 + eq2_b_y2)/2

# Right side: (a+c)/(b+d)
# Numerator (a+c)
eq2_right_a_x1 <- eq2_lt_x + 0.8
eq2_right_a_y1 <- eq2_a_y1
eq2_right_a_x2 <- eq2_right_a_x1 + rect_width
eq2_right_a_y2 <- eq2_right_a_y1 + rect_height

eq2_right_c_x1 <- eq2_right_a_x1 + gap_horizontal
eq2_right_c_y1 <- eq2_a_y1
eq2_right_c_x2 <- eq2_right_c_x1 + rect_width
eq2_right_c_y2 <- eq2_right_c_y1 + rect_height

eq2_plus2_x <- (eq2_right_a_x2 + eq2_right_c_x1)/2
eq2_plus2_y <- (eq2_right_a_y1 + eq2_right_a_y2)/2

# Denominator (b+d)
eq2_right_b_x1 <- eq2_right_a_x1
eq2_right_b_y1 <- eq2_b_y1
eq2_right_b_x2 <- eq2_right_b_x1 + rect_width
eq2_right_b_y2 <- eq2_right_b_y1 + rect_height

eq2_right_d_x1 <- eq2_right_c_x1
eq2_right_d_y1 <- eq2_right_b_y1
eq2_right_d_x2 <- eq2_right_d_x1 + rect_width
eq2_right_d_y2 <- eq2_right_d_y1 + rect_height

eq2_plus3_x <- eq2_plus2_x
eq2_plus3_y <- (eq2_right_b_y1 + eq2_right_b_y2)/2

# Division line
eq2_right_div_x1 <- eq2_right_a_x1
eq2_right_div_x2 <- eq2_right_c_x2
eq2_right_div_y <- (eq2_right_a_y1 + eq2_right_b_y2) / 2

# Equation 2 label
eq2_label_x <- eq1_label_x
eq2_label_y <- a_y1 - eq2_vertical_offset

# Create plots --------------------------------------------------------
# Panel 1: Main example with vaccine effect
plt1a <- ggplot(dlong, aes(week, value, linetype = name, shape = name, group = name)) +
  # Background stripes for no vaccination scenario
  geom_rect_pattern(
    data = subset(dlong, name == "No vacc"),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=value),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_colour = NA,
    pattern_density = pattern_dens,
    pattern_fill = 'steelblue',
    pattern_angle = 30) +
  
  # Vaccination impact rectangle
  geom_rect(
    data = d,
    aes(xmin=week, xmax=lead(week), ymin=`Vacc`, ymax=`No vacc`), 
    fill='steelblue', 
    alpha=0.5,
    color=NA,
    inherit.aes = FALSE) +
  
  # Step lines for clarity
  geom_step(linewidth=step_linewidth) +
  
  # Outbreak response triggered arrow and label
  geom_segment(
    x=week_or, y=10+arrow_vertical_len, xend=week_or, yend=10,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_or-0.5, y=38, 
    label = "Outbreak\nresponse triggered", 
    hjust=0.5, size=text_size, lineheight=line_height
  ) +
  
  # Vaccination started arrow and label
  geom_segment(
    x=week_vax, y=32+arrow_vertical_len, xend=week_vax, yend=32,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_vax, y=60, 
    label = "Vaccination\nstarted", hjust=1,
    size=text_size, lineheight=line_height
  ) +
  
  # Vaccine takes effect arrow and label
  geom_segment(
    x=week_vax_eff, y=122+arrow_vertical_len, xend=week_vax_eff, yend = 122,
    arrow = arrow(type = "closed", angle = 10, length = unit(3, "mm")),
    inherit.aes = FALSE
  ) +
  annotate(
    "text", x=week_vax_eff, y=147, 
    label="Vaccine\ntakes effect", hjust=1,
    size=text_size, lineheight=line_height
  ) +
  
  # Scenario labels
  annotate(
    "text", x=11.5, y=120, 
    label = "Without\nvaccination", hjust=0, 
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=9.5, y=12, 
    label = "With\nvaccination", hjust=0,
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=16, y=75, 
    label = "% case\nreduction", 
    size=text_size, lineheight=line_height
  ) +
  annotate(
    "text", x=19.2, y=75, 
    label = bquote("= 100 \u00D7"), 
    size=text_size, lineheight=line_height
  ) +
  
  # Fraction expressed using squares
  create_divider(x1=21.1-0.2, x2=22.1+0.2, y=75) +
  create_filled_square(
    x1=21.1, x2=22.1, y1=77, y2=85, 
    color='black', fill="steelblue", alpha=0.5
  ) +
  create_striped_square(
    x1=21.1, x2=22.1, y1=64, y2=72,
    fill_color='steelblue', color='black'
  ) +
  
  # Colored lines on the horizontal axis marking periods
  annotate("segment", x=0, y=0, xend=3, yend=0, linewidth=2, color="firebrick") +
  annotate("segment", x=3, y=0, xend=7, yend=0, linewidth=2, color="forestgreen") +
  annotate("segment", x=7, y=0, xend=9, yend=0, linewidth=2, color="purple") +
  
  scale_x_continuous(limits=c(0,25)) +
  labs(x="", y=ylabel) +
  theme_upper_panel()

# Panel 2: Smaller outbreaks
plt1b <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot smaller outbreaks
  geom_step(
    data = subset(df, outbreak == "Smallest"),
    aes(linetype=scenario),
    linewidth=step_linewidth, 
    color='grey'
  ) +
  geom_step(
    data = subset(df, outbreak == "Smaller"),
    aes(linetype=scenario),
    linewidth=step_linewidth, 
    color='black'
  ) +
  
  labs(x=NULL, y=ylabel) +
  theme_upper_panel()

# Panel 3: Small outbreak with equations
plt1c <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot small outbreak
  geom_step(
    data = subset(df, outbreak == "Small"), 
    aes(linetype=scenario),
    linewidth=step_linewidth
  ) +
  
  # Pattern for no vaccination scenario
  geom_rect_pattern(
    data = subset(df, (scenario == "No Vaccine" & outbreak == "Small")),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=cases),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_density = pattern_dens*0.5,
    pattern_fill = 'steelblue',
    pattern_colour = NA,
    pattern_angle = 30
  ) +
  
  # Vaccination impact area
  geom_rect(
    data=subset(df_wide, outbreak == "Small"),
    aes(xmin=week, xmax=lead(week), ymin=`With Vaccine`, ymax=`No Vaccine`),
    fill="steelblue", 
    alpha=0.5, 
    inherit.aes=FALSE
  ) +
  
  # EQUATION 1: a/b < c/d
  # a/b fraction
  create_filled_square(a_x1, a_y1, a_x2, a_y2, "steelblue") +
  create_striped_square(b_x1, b_y1, b_x2, b_y2, "steelblue", dens=pattern_dens*0.5) +
  create_divider(a_x1, a_x2, ab_div_y) +
  
  # Less than symbol
  annotate("text", x=eq1_lt_x, y=eq1_lt_y, label= "<", size=eq_symbol_size) +
  
  # c/d fraction
  create_filled_square(c_x1, c_y1, c_x2, c_y2, "firebrick") +
  create_striped_square(d_x1, d_y1, d_x2, d_y2, "firebrick", dens=pattern_dens*0.5) +
  create_divider(c_x1, c_x2, cd_div_y) +
  
  # Label for equation 1
  annotate("text", x=eq1_label_x, y=eq1_label_y, label="1.", size=5, hjust=0) +
  
  # EQUATION 2: 1/2(a/b+c/d) < (a+c)/(b+d)
  # Left side: 1/2(a/b+c/d)
  
  # a/b part
  create_filled_square(eq2_a_x1, eq2_a_y1, eq2_a_x2, eq2_a_y2, "steelblue") +
  create_striped_square(eq2_b_x1, eq2_b_y1, eq2_b_x2, eq2_b_y2, "steelblue", dens=pattern_dens*0.5) +
  create_divider(eq2_a_x1, eq2_a_x2, (eq2_a_y1 + eq2_b_y2)/2) +
  
  # Plus symbol
  annotate("text", x=eq2_plus1_x, y=eq2_plus1_y, label="+", size=plus_size) +
  
  # c/d part
  create_filled_square(eq2_c_x1, eq2_c_y1, eq2_c_x2, eq2_c_y2, "firebrick") +
  create_striped_square(eq2_d_x1, eq2_d_y1, eq2_d_x2, eq2_d_y2, "firebrick", dens=pattern_dens*0.5) +
  create_divider(eq2_c_x1, eq2_c_x2, (eq2_c_y1 + eq2_d_y2)/2) +
  
  # Parentheses
  annotate("text", x=left_paren_x, y=paren_y, label="(", size=paren_size) +
  annotate("text", x=right_paren_x, y=paren_y, label=")", size=paren_size) +
  
  # 1/2 fraction
  annotate("text", x=eq2_frac_x, y=eq2_frac_y_num, label="1", size=5) +
  annotate("text", x=eq2_frac_x, y=eq2_frac_y_denom, label="2", size=5) +
  annotate("segment", 
           x=eq2_frac_x-0.3, y=eq2_frac_y_line, 
           xend=eq2_frac_x+0.3, yend=eq2_frac_y_line, 
           linewidth=0.8) +
  
  # Less than symbol for second equation
  annotate("text", x=eq2_lt_x, y=eq2_lt_y, label="<", size=eq_symbol_size) +
  
  # Right side: (a+c)/(b+d)
  # Numerator (a+c)
  create_filled_square(eq2_right_a_x1, eq2_right_a_y1, eq2_right_a_x2, eq2_right_a_y2, "steelblue") +
  annotate("text", x=eq2_plus2_x, y=eq2_plus2_y, label="+", size=plus_size) +
  create_filled_square(eq2_right_c_x1, eq2_right_c_y1, eq2_right_c_x2, eq2_right_c_y2, "firebrick") +
  
  # Division line
  create_divider(eq2_right_div_x1, eq2_right_div_x2, eq2_right_div_y) +
  
  # Denominator (b+d)
  create_striped_square(eq2_right_b_x1, eq2_right_b_y1, eq2_right_b_x2, eq2_right_b_y2, 
                        "steelblue", dens=pattern_dens*0.5) +
  annotate("text", x=eq2_plus3_x, y=eq2_plus3_y, label="+", size=plus_size) +
  create_striped_square(eq2_right_d_x1, eq2_right_d_y1, eq2_right_d_x2, eq2_right_d_y2, 
                        "firebrick", dens=pattern_dens*0.5) +
  
  # Label for equation 2
  annotate("text", x=eq2_label_x, y=eq2_label_y, label="2.", size=5, hjust=0) +
  
  labs(x=NULL, y=ylabel) +
  theme_upper_panel()

# Panel 4: Large outbreak  
plt1d <- ggplot(dat=df, aes(x=week, y=cases)) +
  # Time markers
  annotate("segment", x=week_or, y=0, xend=week_or, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax, y=0, xend=week_vax, yend=Inf, linetype="dotted", linewidth=0.6) +
  annotate("segment", x=week_vax_eff, y=0, xend=week_vax_eff, yend=Inf, linetype="dotted", linewidth=0.6) +
  
  # Plot large outbreak
  geom_step(
    data = subset(df, outbreak == "Large"), 
    aes(linetype=scenario),
    linewidth=step_linewidth
  ) +
  
  # Pattern for no vaccination scenario
  geom_rect_pattern(
    data = subset(df, (scenario == "No Vaccine" & outbreak == "Large")),
    aes(xmin=week, xmax=lead(week), ymin=0, ymax=cases),
    pattern = 'stripe',
    fill = NA,
    colour = NA,
    pattern_density = pattern_dens*0.5,
    pattern_fill = 'firebrick',
    pattern_colour = NA,
    pattern_angle = 30
  ) +
  
  # Vaccination impact area
  geom_rect(
    data=subset(df_wide, outbreak == "Large"),
    aes(xmin=week, xmax=lead(week), ymin=`With Vaccine`, ymax=`No Vaccine`),
    fill="firebrick", 
    alpha=0.5, 
    inherit.aes=FALSE
  ) +
  
  # Axis and labels
  labs(x=xlabel, y=ylabel) +
  scale_x_continuous(limits=c(0,25)) +
  theme_upper_panel() +
  theme(
    axis.text.x = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    legend.position = "none"
  )

# Combine all plots
# Right panel of Figure 1
combined_plot <- (plt1a / plt1b / plt1c / plt1d) + 
  plot_layout(
    heights = c(0.8, 0.4, 0.4, 0.4), 
    axis_titles = 'collect',
    guides = 'collect'
  )

# Save plot
if (save_plot) {
  fac <- 1.8
  ggsave(paste0("plots/fig1_vacc_eff_schematic_", tstamp(), ".pdf"),
        combined_plot,
        device = cairo_pdf,
        width=1.25*fgw1*fac,
        height=1.6*fgw1*fac,
        units = "mm")
}
```

## Figure 2: Outbreak Description

Outbreak ids by confirmation status
```{r}
paste0(length(unique(outbreak_data_all$id_outbreak)), " outbreaks total")

ids_excl <- outbreak_data_all |> 
  filter(!(id_outbreak %in% outbreak_data$id_outbreak)) |>  
  pull(id_outbreak) |>
  unique()

# Identify confirmed outbreaks (with laboratory confirmation)
ids_conf <- outbreak_data |> 
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique()

# Identify suspected (not confirmed) outbreaks
ids_susp <- outbreak_data |>
  group_by(id_outbreak) |>
  mutate(c_ch_sum = sum(c_ch)) |>
  filter(!(c_ch_sum > 0)) |>
  pull(id_outbreak) |>
  unique()
```

```{r}
# Load and prepare country-region mapping data
country_region <- fread("data/paper/africa_subregion_country_names.csv")
dt <- country_region
country_region <- country_region[Subregion == "Middle Africa", Subregion := "Central Africa"]
# change Middle to Central
country_region %>% mutate()
country_region$iso3c <- sapply(country_region$Country, 
                               function(x) countrycode(x, origin = "country.name.en", 
                                                     destination = "iso3c"))

# Join outbreak data with regional information
outbreak_data$iso3c <- outbreak_data$country
outbreak_data_region <- left_join(outbreak_data, country_region, by = "iso3c")

# =============================================================================
# PLOT 2A: TEMPORAL TRENDS BY SUBREGION (STACKED AREA CHART)
# =============================================================================

# Aggregate data by month and subregion
outbreak_month <- outbreak_data_region %>% 
  mutate(date = as.Date(date),
         date_bin = floor_date(date, "1 month"))

# Calculate total cases per subregion per month
outbreak_month_summary <- outbreak_month %>%
  group_by(date_bin, Subregion) %>%
  summarise(total_cases = sum(s_ch, na.rm = TRUE), .groups = "drop")

# Create complete time series grid to fill gaps
all_subregions <- unique(outbreak_month_summary$Subregion)
all_date_bins <- seq(min(outbreak_month_summary$date_bin), 
                     max(outbreak_month_summary$date_bin), 
                     by = "1 month")
full_grid <- expand_grid(date_bin = all_date_bins, Subregion = all_subregions)

# Merge and fill missing values with zero
df_complete <- full_grid %>%
  left_join(outbreak_month_summary, by = c("date_bin", "Subregion")) %>%
  mutate(total_cases = replace_na(total_cases, 0))

# Create stacked area plot
plt_2a <- ggplot(df_complete, aes(x = date_bin, y = total_cases, 
                                  fill = Subregion)) +
  geom_area(position = "stack", alpha = 0.8) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  scale_y_continuous(labels = comma_format()) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  labs(x = "", y = "Monthly suspected cases", fill = "") +
  theme_light() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    plot.margin = margin(0,0,0,0,"pt")
  )

# =============================================================================
# PLOT 2B: SPATIAL DISTRIBUTION MAP
# =============================================================================
# Prepare outbreak data for mapping
dt <- outbreak_data %>% rename(lctn_pr = location_period_id)
first_rows_df <- dt %>% 
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

# Ensure consistent data types for joining
shp$lctn_pr <- as.character(shp$lctn_pr)

# Join outbreak data with spatial data
map_data <- left_join(first_rows_df, shp, by = "lctn_pr")
map_data %<>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE))

# Convert to spatial features
map_data <- st_as_sf(map_data)

islands_to_exclude <- c("CPV",  # Cape Verde
                        "SYC",  # Seychelles
                        "MYT",  # Mayotte
                        "REU",  # Réunion
                        "SHN")  # Saint Helena, Ascension and Tristan da Cunha

# Exclude these from your Africa shapefile
africa_main <- africa %>% dplyr::filter(!ISO3 %in% islands_to_exclude)
bbox   <- st_bbox(africa_main)       # much tighter now

plt_2b <- ggplot() +
  geom_sf(data = africa_main,  fill = NA, colour = "black", size = .3) +
  geom_sf(data = map_data, aes(colour = is_confirmed, fill = is_confirmed),
          alpha = .7, size = .8) +
  coord_sf(xlim = c(bbox["xmin"], bbox["xmax"]-10),
           ylim = c(bbox["ymin"]+12, bbox["ymax"]),
           expand = FALSE, clip = "on") +    # clip cuts everything outside
  scale_colour_manual("", values = c("steelblue4", "firebrick4"),
                      labels  = c("Not confirmed", "Confirmed")) +
  scale_fill_manual("",   values = c("steelblue4", "firebrick4"),
                    labels  = c("Not confirmed", "Confirmed")) +
  theme_void() +                                   
  theme(legend.position = "top",
        plot.margin     = margin(0,0,0,0,"pt"),
        panel.spacing = unit(0, "pt"))

# plt_2b <- plt_2b + theme(aspect.ratio = 0.7)

# =============================================================================
# PLOT 2C: EPIDEMIC CURVES WITH INSET
# =============================================================================

# Prepare weekly case data
week_sch <- outbreak_data[, c("id_outbreak", "week", "s_ch")]
setDT(week_sch)
max_week <- week_sch[, max(week)]

# Create complete grid for all outbreaks up to max week
all_weeks <- CJ(id_outbreak = unique(week_sch$id_outbreak), week = 1:max_week)

# Merge and fill missing values with zero
week_sch_full <- merge(all_weeks, week_sch, 
                       by = c("id_outbreak", "week"), all.x = TRUE)
week_sch_full[is.na(s_ch), s_ch := 0]

# Calculate summary statistics across all outbreaks
week_sch_summary <- week_sch_full[, .(
  sum_sch = sum(s_ch, na.rm = TRUE),
  median = median(s_ch, na.rm = TRUE),
  p025 = quantile(s_ch, 0.025, na.rm = TRUE),
  p25  = quantile(s_ch, 0.25, na.rm = TRUE),
  p75  = quantile(s_ch, 0.75, na.rm = TRUE),
  p975 = quantile(s_ch, 0.975, na.rm = TRUE)
), by = week]


# Main epidemic curve plot
plt2c <- ggplot(week_sch_summary, aes(x = week)) +
  geom_ribbon(aes(ymin = p025, ymax = p975), fill = set1[3], alpha = 0.45) +
  geom_ribbon(aes(ymin = p25, ymax = p75), fill = set1[3], alpha = 0.6) +
  geom_line(aes(y = median), color = "black", linewidth = 1) +
  labs(title = "", x = "Epidemic week", y = "Suspected cholera cases") +
  scale_x_continuous(limits = c(0, 50), breaks = seq(0, 50, 10)) +
  theme_light() +
  theme(
    plot.margin = margin(0,1,0,1, "pt")
  )

# time to peak
outbk_data_summary %>% 
  summarize(mean_time_to_peak = mean(time_to_peak_week),
            median_time_to_peak = median(time_to_peak_week))

# percentage of samples that are shorter than or equal to 10 weeks
sum(outbk_data_summary$time_to_peak_week <= 10) /
  length(outbk_data_summary$time_to_peak_week)


plt2c_inset <- ggplot(outbk_data_summary, aes(x = time_to_peak_week)) +
  stat_ecdf(geom = "step", size = 0.8) +
  labs(x = "Time to peak (weeks)", 
       y = expression(P(X <= x)))+
  scale_x_continuous(limits=c(0,30), breaks = seq(0,30,10), labels = seq(0,30,10))+
  scale_y_continuous(limits=c(0,1), breaks = seq(0,1,0.5), labels = seq(0,1,0.5))+
  theme_light()+
  theme_tight()+
  theme(plot.margin = margin(0,0,0,0,"pt"))


# Convert inset to grob and add to main plot with better positioning
grob_inset <- ggplotGrob(plt2c_inset)

plt_2c_combined <- plt2c +
  annotation_custom(
    grob = grob_inset,
    xmin = 20, xmax = 50,  # Moved to right side with more space
    ymin = 100, ymax = 340   # Increased height for better visibility
  )

# =============================================================================
# PLOT 2D: OUTBREAK SIZE vs DURATION
# =============================================================================

plt_2d <- outbk_data_summary %>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE)) %>% 
  ggplot() +
  geom_point(aes(x = total_s_ch, y = duration_weeks, color = is_confirmed), 
             alpha = 0.7, size = 1.2) +
  scale_y_log10(labels = comma_format()) +
  scale_x_log10(labels = comma_format()) +
  scale_color_manual("", 
                     labels = c("Not confirmed", "Confirmed"),
                     values = c("steelblue4", "firebrick4")) +
  labs(title = "", x = "Total suspected cases", y = "Outbreak duration (weeks)") +
  theme_light() +
  theme(legend.position = "top", plot.margin = margin(0,1,0,0,"pt"))

# =============================================================================
# PLOT 2E: OUTBREAK SIZE vs ATTACK RATE
# =============================================================================

summary(outbk_data_summary$duration_weeks) 
summary(outbk_data_summary$total_s_ch)
summary(outbk_data_summary$ar_per_10e3)

summary(outbk_data_summary[id_outbreak %in% ids_conf]$duration_weeks)
summary(outbk_data_summary[id_outbreak %in% ids_conf]$total_s_ch)
summary(outbk_data_summary[id_outbreak %in% ids_conf]$ar_per_10e3)

plt_2e <- outbk_data_summary %>% 
  mutate(is_confirmed = ifelse(id_outbreak %in% ids_conf, TRUE, FALSE)) %>% 
  ggplot() +
  geom_point(aes(x = total_s_ch, y = ar_per_10e3, color = is_confirmed), 
             alpha = 0.7, size = 1.2) +
  scale_y_log10(labels = comma_format()) +
  scale_x_log10(labels = comma_format()) +
  scale_color_manual("", 
                     labels = c("Not confirmed", "Confirmed"),
                     values = c("steelblue4", "firebrick4")) +  
  labs(title="", x = "Total suspected cases", y = "Attack rate (per 10,000)") +
  theme_light() +
  theme(legend.position="top",  plot.margin = margin(0,0,0,1,"pt"))



# =============================================================================
# FINAL COMBINED PLOT WITH IMPROVED LAYOUT
# =============================================================================

# Create combined plot with better proportions
# Row 1: Temporal trends (full width)
# Row 2: Map (60%) + Epidemic curve (40%) 
# Row 3: Scatter plots (50% each)
combined_plot <- plt_2a /
  (plt_2b + plt_2c_combined + plot_layout(widths = c(0.6, 1))) /
  (plt_2d + plt_2e + plot_layout(widths = c(1, 1), guides="collect") +
     plot_annotation(theme = theme(legend.position = "top")))+
  plot_layout(heights = c(1,1.3,1)) +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.margin = margin(0,0,0,0,"pt"),
        plot.tag.position = c(0.2, 1),
        plot.tag = element_text(size = 13))

# Display the combined plot
# combined_plot
if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/fig2_outbreak_summary_", tstamp(), ".pdf"), combined_plot,
         device = cairo_pdf, width = fgw2 * fac, height = fgw2 * fac, units = "mm")
}
```

## Figure S1. Indirect Vaccine Effectiveness 

## Figure S2. All outbreaks (eligible and ineligible)

```{r figS2bc, eval=TRUE}
# ---- Figure S2 (B,C): all outbreaks, excluded ones colored by reason ----

s2_levels <- c("Included", "Too small or large", "Nested",
               "No OCV data", "Affected by OCV campaign")

# Okabe-Ito (colorblind-safe) for reasons; neutral gray for the included baseline
s2_pal <- c(
  "Included"                 = "gray75",
  "Too small or large"       = "#E69F00",
  "Nested"                   = "#0072B2",
  "No OCV data"              = "#009E73",
  "Affected by OCV campaign" = "#D55E00"
)

# Included: ds restricted to the analyzed set (consistent w/ excluded data)
s2_included <- ds %>%
  dplyr::filter(id_outbreak %in% outbreaks_filtered$id_outbreak) %>%
  dplyr::transmute(id_outbreak, total_s_ch, duration_weeks, ar_per_10e3,
                   category = "Included")

# Excluded: recode filter_reason to display labels
s2_excluded <- outbreaks_excluded %>%
  dplyr::transmute(id_outbreak, total_s_ch, duration_weeks, ar_per_10e3,
    category = dplyr::recode(filter_reason,
      "Too small or too large" = "Too small or large",
      "Nested outbreak"        = "Nested",
      "Outbreak after 2020"    = "No OCV data",
      "OCV-affected outbreak"  = "Affected by OCV campaign"))

s2_df <- dplyr::bind_rows(s2_included, s2_excluded) %>%
  dplyr::mutate(category = factor(category, levels = s2_levels))

# table(s2_df$category)  # expect 1192 Included + 472 across 4 reasons = 1664

s2_layers <- function(yvar, ylab, mar) {
  list(
    geom_point(data = dplyr::filter(s2_df, category == "Included"),
               aes(x = total_s_ch, y = .data[[yvar]], color = category),
               alpha = 0.5, size = 1.2),
    geom_point(data = dplyr::filter(s2_df, category != "Included"),
               aes(x = total_s_ch, y = .data[[yvar]], color = category),
               alpha = 0.7, size = 1.2),
    scale_y_log10(labels = comma_format()),
    scale_x_log10(labels = comma_format()),
    scale_color_manual("", values = s2_pal, limits = s2_levels, drop = FALSE),
    labs(title = "", x = "Total suspected cases", y = ylab),
    theme_light(),
    theme(legend.position = "top", plot.margin = mar),
    guides(color = guide_legend(nrow = 2, byrow = TRUE,
                                override.aes = list(alpha = 1, size = 2)))
  )
}

plt_S2b <- ggplot() + s2_layers("duration_weeks", "Outbreak duration (weeks)",
                                margin(0,1,0,0,"pt"))
plt_S2c <- ggplot() + s2_layers("ar_per_10e3", "Attack rate (per 10,000)",
                                margin(0,0,0,1,"pt"))

plt_S2bc <- plt_S2b + plt_S2c +
  plot_layout(widths = c(1, 1), guides = "collect") &
  theme(legend.position = "top")

plt_S2bc

if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS2bc_outbreak_size_dur_ar_", tstamp(), ".pdf"),
         plt_S2bc, device = cairo_pdf,
         width = fgw2 * fac, height = fgw1 * fac, units = "mm")
}
```

```{r figS1, eval=FALSE}
# Create time series plot with different colors by outbreak type
plt_S1a <- ggplot() +
  geom_line(data = filter(outbreak_data_all, id_outbreak %in% ids_excl),
            aes(date, s_ch, group = id_outbreak, color = "Excluded"), 
            alpha = 0.5) +
  geom_line(data = filter(outbreak_data, !(id_outbreak %in% ids_conf)),
            aes(date, s_ch, group = id_outbreak, color = "Suspected"), 
            alpha = 0.5) +
  geom_line(data = filter(outbreak_data, id_outbreak %in% ids_conf), 
            aes(date, s_ch, group = id_outbreak, color = "Confirmed"),
            alpha = 0.5) +
  scale_color_manual("", values = c("Excluded" = "yellow4",
                                  "Suspected" = "steelblue4", 
                                  "Confirmed" = "firebrick4")) +
  labs(x = "", y = "Weekly suspected cases") +
  theme_light() +
  theme_tight() +
  theme(legend.position = "top")

# (B) Outbreak size vs. duration by type
outbreak_data_all |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                  as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d_all

outbreak_data |> 
  group_by(id_outbreak) |> 
  summarize(duration = as.integer(as.Date(end_date_outbreak[1]) - 
                                  as.Date(start_date_outbreak[1])),
            total_s_ch = sum(s_ch),
            total_c_ch = sum(c_ch)) -> d

if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1a_outbreak_ts_", tstamp(), ".pdf"), plt_S1a,
         device = cairo_pdf, width = fgw2 * fac, height = fgw1 * fac, units = "mm")
}

plt_S1b <- ggplot() +
  geom_point(data = filter(d_all, id_outbreak %in% ids_excl), 
             aes(duration, total_s_ch, color = "Excluded"), alpha = 0.5) +
  geom_point(data = filter(d, id_outbreak %in% ids_susp), 
             aes(duration, total_s_ch, color = "Suspected"), alpha = 0.5) +
  geom_point(data = filter(d, id_outbreak %in% ids_conf), 
             aes(duration, total_s_ch, color = "Confirmed"), alpha = 0.5) +
  scale_x_log10() +
  scale_y_log10() +
  theme_light() +
  theme_tight() +
  scale_color_manual("", values = c("Excluded" = "yellow4", 
                                  "Suspected" = "steelblue4", 
                                  "Confirmed" = "firebrick4")) +
  labs(x = "Outbreak duration (days)", y = "Total suspected cases") +
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1b_outbreak_size_dur_", tstamp(), ".pdf"), plt_S1b,
         device = cairo_pdf, width = fgw1 * fac, height = fgw1 * fac, units = "mm")
}

# (C) Map of outbreaks
dt <- outbreak_data 
dt_all <- outbreak_data_all 

# Prepare data for mapping
dt <- dt %>% rename(lctn_pr = location_period_id)
dt_all <- dt_all %>% rename(lctn_pr = location_period_id)

first_rows_df <- dt %>% 
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

first_rows_df_all <- dt_all %>%
  group_by(id_outbreak) %>% 
  slice(1) %>%
  ungroup()

first_rows_df$lctn_pr <- as.numeric(first_rows_df$lctn_pr)
first_rows_df_all$lctn_pr <- as.numeric(first_rows_df_all$lctn_pr)

# Join with spatial data
shp$lctn_pr <- as.character(shp$lctn_pr)
first_rows_df$lctn_pr <- as.character(first_rows_df$lctn_pr)
first_rows_df_all$lctn_pr <- as.character(first_rows_df_all$lctn_pr)

d <- left_join(first_rows_df, shp, by = "lctn_pr")
d_all <- left_join(first_rows_df_all, shp, by = "lctn_pr")

# Create indicator variables for outbreak types
d <- d %>%
  mutate(id_conf = ifelse(id_outbreak %in% ids_conf, 1, 0),
         id_sus = ifelse(!id_outbreak %in% ids_conf, 1, 0))

# Subset data by confirmation status
d_conf <- d %>% filter(id_conf == 1)
d_sus <- d %>% filter(id_sus == 1)

# Convert to spatial data
d_conf <- st_as_sf(d_conf)
d_sus <- st_as_sf(d_sus)
d_all <- d_all %>% mutate(id_excl = ifelse(id_outbreak %in% ids_excl, 1, 0))
d <- st_as_sf(d)
d_all <- st_as_sf(d_all)

# Create map
plt_S1c <- ggplot() +
  geom_sf(data = africa, fill = NA, color = "black") +  # Country outlines
  geom_sf(data = d_all, aes(fill = id_excl), color = "black", 
          fill = "yellow4", alpha = 0.5) +
  geom_sf(data = d_sus, aes(fill = id_sus), color = "steelblue4", 
          fill = "steelblue4", alpha = 0.5) +
  geom_sf(data = d_conf, aes(fill = id_conf), color = "firebrick4", 
          fill = "firebrick4", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Map of Cholera Outbreaks")


if (save_plot) {
  fac <- 1.2
  ggsave(paste0("plots/figS1c_outbreak_map_", tstamp(), ".pdf"), plt_S1c,
         device = cairo_pdf, width = fgw1 * fac, height = fgw1 * fac, units = "mm")
}
```

## Figure 3: Impact across delays and prioritization

# Summary values in results section
```{r fig3, eval=FALSE}
onset_impact <- cea_ori %>%
  filter(vacc_cov %in% c(0.6, 0.75, 0.9), vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, 
                      pct_reduc_case, 
                      case_averted_per_1000_OCV, 
                      daly_averted_per_1000_OCV,
                      cost_per_case_averted, 
                      cost_per_daly_averted)

# cases averted (CA) for vaccine coverage of 75% with ORI in week 1
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", 
         vacc_week == 1, vacc_cov == 0.75) %>%
  format_median_iqr()
# percent case averted (PCA) at week 1
onset_impact %>% 
  filter(variable == "pct_reduc_case", 
         vacc_week == 1, vacc_cov == 0.75) %>%
  format_median_iqr()

# CA by vaccine coverage in week 1
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", vacc_week == 1) %>%
  group_by(vacc_cov) %>% 
  format_median_iqr()
# CA by vaccine coverage in week 9
onset_impact %>% 
  filter(variable == "s_ch_averted_tot", vacc_week == 9) %>%
  group_by(vacc_cov) %>% 
  format_median_iqr()

# PCA by vaccine coverage in week 9
onset_impact %>% 
  filter(variable == "pct_reduc_case", vacc_week == 9) %>%
  group_by(vacc_cov) %>% 
  format_median_iqr()


# outbrreaks lasting >=7 weeks
sum(outbk_data_summary$duration_weeks >= 7) / nrow(outbk_data_summary)

quantile(fit_decay[["9"]]$r_squared, probs=c(0.25,0.5,0.75))

# instantaneous decay rate
fit_decay[["9"]] %>% 
  summarize_quantiles(k_hat) %>% 
  format_median_iqr(digits=2)

# weekly reduction rate
fit_decay[["9"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# half_life                  
fit_decay[["9"]] %>%
  summarize_quantiles(half_life) %>% 
  format_median_iqr(digits=1)

# weekly reduction rate for all outbreaks
fit_decay[["5"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# weekly reduction rate for all outbreaks
fit_decay[["14"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# weekly reduction rate for all outbreaks
fit_decay[["21"]] %>% 
  mutate(rate_reduc = 1 - exp(-k_hat)) %>% 
  summarize_quantiles(rate_reduc*100) %>% 
  format_median_iqr(digits=1)

# outbrreaks lasting >=7 weeks
sum(outbk_data_summary$duration_weeks >= 19) / nrow(outbk_data_summary)

# linear decrease in proportion of ORI-eligible outbreaks
round(abs(coef(fit_lm)[[2]] * 100), digits=1)
# observed at week 9
1 - fit_lm$model %>% filter(vacc_week == 9) %>% pull(prop_included_dur)
# predicted at week 9
1 - predict(fit_lm, newdata=data.frame(vacc_week=9))

cea_prioritization %>%
  mutate(prob_i = as.integer(round(100 * prob))) %>% 
  filter(metric == "med_s_ch_averted_tot", 
         percentile_by =="size", prob_i %in% c(0, 90), 
         vacc_week %in% c(1,9)) %>% 
  select(vacc_week, prob_i, median_value)

capd <- cea_prioritization %>%
  mutate(prob_i = as.integer(round(100 * prob))) %>% 
  filter(metric == "med_case_averted_per_1000_OCV", 
         percentile_by =="ar", prob_i %in% c(0, 90), 
         vacc_week %in% c(1,9)) %>% 
  select(vacc_week, prob_i, median_value)

# fold increase
capd$median_value[3] / capd$median_value[1]
capd$median_value[4] / capd$median_value[2]

dalyapd <- cea_prioritization %>%
  mutate(prob_i = as.integer(round(100 * prob))) %>% 
  filter(metric == "med_daly_averted_per_1000_OCV", 
         percentile_by =="ar", prob_i %in% c(0, 90), 
         vacc_week %in% c(1,9)) %>% 
  select(vacc_week, prob_i, median_value)

# fold increase
dalyapd$median_value[3] / dalyapd$median_value[1]
dalyapd$median_value[4] / dalyapd$median_value[2]

cea_prioritization %>%
  mutate(prob_i = as.integer(round(100 * prob))) %>% 
  filter(metric == "med_pct_reduc_case", 
         prob_i %in% c(90), 
         vacc_week %in% c(1,9)) %>% 
  select(vacc_week, percentile_by, median_value)

# Cost per case averted (CPCA) for ORI with vaccine coverage of 75%
onset_impact %>% 
  filter(variable == "cost_per_case_averted", 
         vacc_week %in% c(1,9), vacc_cov == 0.75) %>%
  format_median_iqr(digits=0)

onset_impact %>% 
  filter(variable == "cost_per_daly_averted", 
         vacc_week %in% c(1,9), vacc_cov == 0.75) %>%
  format_median_iqr(digits=0)

# gdp data from World Bank during 2010-2020
gdp_data <- read_xls("data/paper/GDP_WorldBank.xls")
gdp_2010_2020 <- gdp_data %>% 
  filter(`Country Name` == "Sub-Saharan Africa (IDA & IBRD countries)") %>%
  select(c(`2010`:`2020`)) |> 
  as.numeric() 

# PCA in the top quartile by duration
cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_q_dur_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(pct_reduc_case) %>%  
  select(vacc_week,q250,q500,q750)

# Cost per DALY averted
cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  filter(id_outbreak %in% top_10_ar_ids) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted) %>%  
  select(vacc_week,q250,q500,q750) -> cpdalya_top10
# ratio to the median gdp
round(cpdalya_top10$q500/median(gdp_2010_2020), digits=1) 

# Cost per DALY averted
cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted) %>%  
  select(vacc_week,q250,q500,q750) -> cpdalya

# % reduction in cost per DALY averted 
100*(1 - cpdalya_top10$q500/cpdalya$q500)

# exponential decay
fit_decay_ct[["1"]] %>%
  summarize_quantiles(r_squared) %>% 
  format_median_iqr(digits=2)

# instantaneous decay rate across cases
fit_decay_ct[["1"]] %>%
  summarize_quantiles(k_hat) %>% 
  format_median_iqr(digits=4)

# % reduction for the 10 additional missing cases for ORIs implemented in week 1
fit_decay_ct[["1"]] %>% 
  mutate(rate_reduc_10 = 1 - exp(-k_hat*10)) %>% 
  summarize_quantiles(rate_reduc_10*100) %>% 
  format_median_iqr(digits=1)

# % reduction for the 10 additional missing cases for ORIs implemented in week 1
fit_decay_ct[["1"]] %>% 
  mutate(rate_reduc_10 = 1 - exp(-k_hat*100)) %>% 
  summarize_quantiles(rate_reduc_10*100) %>% 
  format_median_iqr(digits=1)

# half_life                  
fit_decay_ct[["1"]] %>%
  summarize_quantiles(half_life) %>% 
  format_median_iqr(digits=1)

# % reduction for the 10 additional missing cases for ORIs implemented in week 1
fit_decay_ct[["5"]] %>% 
  mutate(rate_reduc_10 = 1 - exp(-k_hat*10)) %>% 
  summarize_quantiles(rate_reduc_10*100) %>% 
  format_median_iqr(digits=1)

# % reduction for the 100 additional missing cases for ORIs implemented in week 1
fit_decay_ct[["5"]] %>% 
  mutate(rate_reduc_10 = 1 - exp(-k_hat*100)) %>% 
  summarize_quantiles(rate_reduc_10*100) %>% 
  format_median_iqr(digits=1)
# half_life                  
fit_decay_ct[["5"]] %>%
  summarize_quantiles(half_life) %>% 
  format_median_iqr(digits=1)
```

### Left panel

Percent cases averted

```{r}
ymean  <- 95 
ymedian<- 87.5
y50    <- c(82, 93)
y95    <- c(75, 100)

VW_plot_seq <- seq(1,15,2)
# percentage case averted 
p1 <- cea_ori_pca %>%
  filter(vacc_week %in% VW_plot_seq) %>%
  ggplot(aes(x = vacc_week, color = as.factor(vacc_cov), 
             group = as.factor(vacc_cov))) +
  ci_layers_from_summary(
    x_discrete = TRUE,
    dodge_width = 1.4,
    lw95 = line_width_95,
    lw50 = line_width_50,
    pr_size = 2.4,
    mean_size = x_size,
    mean_stroke = x_stroke
  ) +
  add_summary_braces(
    x0 = 13,
    ymean = ymean, ymedian = ymedian,
    y50 = y50, y95 = y95,
    pt_size_median = 3,
    pt_size_mean = x_size,
    pt_stroke_mean = x_stroke,
    dx50 = 0.5, dx95 = 0.8,
    text_50_x = 13.6,
    text_50_y = y50[2],
    text_95_x = 14.85,
    text_95_y = y95[2],
    brace_50_x = 13.2,
    brace_95_x = 14.4,
    text_offset_mean = 0.3,
    lw_50 = 0.6, lw_95 = 0.6,
    whisker_50 = line_width_50, 
    whisker_95 = line_width_95
  ) +
  scale_y_continuous("% cases averted",
                     labels = scales::percent_format(scale = 1), 
                     limits=c(0,100)) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[1:3],
    labels = c(paste0(format_num(VC3*100, 1), "%")))+
  theme_light() +
  theme_tight() +
  theme_x_blank()

# plot 2
# exponential decay curve
set.seed(2025)
runid_sample <- 1
n_sample <- 20

x_range <- seq(1, 15, by = 0.2) # x-axis grid
t0 <- 1  # baseline week for normalization
## ---- normalized "spaghetti" using fitted params ----
# use fits longer than 9 weeks
ids_wk <- fit_decay$`9` %>% 
  filter(id_outbreak %in% unique(fit_decay$`9`$id_outbreak)) %>%
  pull(id_outbreak) %>% unique()

id_outbreak_sample_wk <- sample(ids_wk, size = n_sample)

fitted_models <- fit_decay$`9` %>% 
  filter(runid %in% runid_sample,
         id_outbreak %in% id_outbreak_sample_wk)

## ---- normalized "spaghetti" using fitted params ----
# On the normalized scale, A0 cancels; z_hat(t) = exp(-k * (t - t0))
spaghetti_data_norm_wk <- fitted_models %>%
  mutate(vacc_week = list(x_range)) %>%
  unnest(vacc_week) %>%
  mutate(
    # z_hat = exp(-decay_rate_k * (vacc_week - t0))
    z_hat = exp(- k_hat * (vacc_week - t0))
  )

spaghetti_data_norm <- spaghetti_data_norm_wk
id_outbreak_sample <- id_outbreak_sample_wk

## ---- raw data, normalized by week-1 value per series ----
# Expecting 'cea_ori' with columns: vacc_cov, ori_occurred, vacc_week, runid, id_outbreak, pct_reduc_case
raw_data_points_norm <- cea_weekly_delay_2_ori %>%
  filter(vacc_cov == 0.75,
         vacc_week %in% 1:15, 
         runid %in% runid_sample,
         id_outbreak %in% id_outbreak_sample) %>%
  arrange(runid, id_outbreak, vacc_week) %>%
  group_by(runid, id_outbreak) %>%
  mutate(
    y0 = first(pct_reduc_case),
    z  = ifelse(y0 > 0, pct_reduc_case / y0, NA_real_)
  ) %>%
  ungroup()

## ---- summary curves from k only (median + IQR) ----
# Use the full data set
# k_q <- quantile(fit_decay$`5`$decay_rate_k, 
#                 probs = c(0.025, .25, .5, .75, 0.975), na.rm = TRUE)
k_q <- quantile(fit_decay$`9`$k_hat, 
                probs = c(0.025, .25, .5, .75, 0.975), na.rm = TRUE)
k_q025 <- k_q[[1]]; k_q25 <- k_q[[2]]; k_med <- k_q[[3]]; 
k_q75 <- k_q[[4]]; k_q975 <- k_q[[5]]

summary_curves_norm <- tibble(
  vacc_week = x_range,
  med = exp(-k_med * (x_range - t0)),
  lo  = exp(-k_q75 * (x_range - t0)),  # faster decay => lower envelope
  hi  = exp(-k_q25 * (x_range - t0)),   # slower decay => higher envelope
  q975   = exp(-k_q975 * (x_range - t0)),  # faster decay => lower envelope
  q025  = exp(-k_q025 * (x_range - t0))   # slower decay => higher envelope
)

## ---- pretty label (normalized) ----
format_k <- paste0(round(k_med, digits = 3))
eq_label_norm <- paste0("$\\hat{z}(t)=\\exp\\{-", format_k, "(t-1)\\}$")

## ---- plot (normalized scale 0-1) ----
p2 <- ggplot() +
  # Individual normalized fitted curves
  geom_line(
    data = spaghetti_data_norm,
    aes(x = vacc_week, y = z_hat, group = id_outbreak, color = id_outbreak),
    alpha = 0.60
  ) +
  # Raw points, normalized by their own week-1
  geom_point(
    data = raw_data_points_norm,
    aes(x = vacc_week, y = z, color = id_outbreak),
    alpha = 0.60, size = 1
  ) +
  # Median + IQR summary on normalized scale
  geom_ribbon(
    data = summary_curves_norm,
    aes(x = vacc_week, ymin = lo, ymax = hi),
    inherit.aes = FALSE,
    fill = "firebrick",
    alpha = 0.30
  ) +
  geom_ribbon(
    data = summary_curves_norm,
    aes(x = vacc_week, ymin = q025, ymax = q975),
    inherit.aes = FALSE,
    fill = "firebrick",
    alpha = 0.20
  ) +
  geom_line(
    data = summary_curves_norm,
    aes(x = vacc_week, y = med),
    color = "firebrick",
    linewidth = 1.2
  ) +
  annotate(
    "latex",
    x = 15, y = 1,
    label = eq_label_norm,
    size = 4.5, hjust = 1, vjust = 1
  ) +
  # scale_y_continuous("Normalized reduction", limits = c(0, 1)) +
  scale_y_continuous("Normalized reduction") +
  guides(color = "none") +
  theme_light() +
  theme_tight() +
  theme_x_blank()

# plot 3
co <- coef(fit_lm)
alpha <- unname(co[1])   # intercept
beta  <- unname(co[2])   # slope (expected to be negative)
r2    <- summary(fit_lm)$r.squared

# --- prediction grid & values ---
x_grid <- seq(min(fit_lm$model$vacc_week), 
              max(fit_lm$model$vacc_week), length.out = 200)
pred   <- predict(fit_lm, newdata = data.frame(vacc_week = x_grid))

paste0("Eveary week, additional ", 
       format_num(abs(100*beta), digits=1), 
       "% outbreaks ended before ORI")
wk <- 9
paste0("Week ", wk, ", ", 
       format_num(100*(1-pred[min(which(x_grid>wk))]), digits=1), 
       "% outbreaks ended before ORI")
# --- a tidy equation label using LaTeX (percent scale) ---
# round for display
alpha_lab <- format(round(alpha, 2), nsmall = 2)
beta_lab  <- format(round(beta, 3), nsmall = 3)
r2_lab    <- format(round(r2, 3), nsmall = 3)

eq_label <- paste0(
  "$\\hat{y}(t) = ", alpha_lab, beta_lab, "\\, t$"
)

# add 1 for weeks 1-3
df <- data.frame(vacc_week = 1:3, 
                 prop_included_dur=1)
df <- rbind(df, fit_lm$model)

p3 <- ggplot() +
  # raw data (black)
  geom_point(
    data = df,
    aes(vacc_week, prop_included_dur),
    color = "black", size = 2
  ) +
  # fitted line (firebrick)
  geom_line(
    aes(x_grid, pred),
    color = "firebrick", linewidth = 1.2
  ) +
  # equation label (LaTeX)
  annotate(
    "latex",
    x = 15, y = 1,
    label = eq_label,
    size = 4,
    hjust = 1, vjust = 1
  ) +
  # scale_x_continuous("Delay in ORI implementation (weeks)", 
  #                    breaks = seq(1, 15, 2), limits=c(1,15)) +
  scale_y_continuous("Proportion included") +
  theme_light() +
  theme_tight()

```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- p1 +
  annotate("text", x = -Inf, y = Inf, label = "A",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2), limits=c(0.2,15.5))+
  coord_cartesian(clip = "off")

p2_tag <- p2 +
  annotate("text", x = -Inf, y = Inf, label = "B",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2), limits=c(0.2,15.5))+
  coord_cartesian(clip = "off")

p3_tag <- p3 +
  annotate("text", x = -Inf, y = Inf, label = "C",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Week of ORI implementation",
                     breaks=seq(1,15,2), limits=c(0.2,15.5))+
  coord_cartesian(clip = "off")

plts_pca <- (p1_tag / p2_tag / p3_tag) + plot_layout(ncol=1)

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig3_onset_triggered_left_", tstamp(), ".pdf"),
         plts_pca,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### Right panel

```{r}
# Create the color palette function (your code was correct here)
myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))

y_labs <- c("Attack rate percentile", "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
                   "Prioritizing longer outbreaks",
                   "Prioritizing larger outbreaks",
                   "Prioritizing higher CFR outbreaks")

plot_data <- cea_prioritization %>% filter(metric == "med_s_ch_averted_tot")

plot_data %>%
  filter(vacc_cov == 0.75, vacc_week == 9, percentile_by == "size") %>%
  group_by(prob) %>% 
  summarize(med_case_averted = median_value) 

# Create the color palette function (your code was correct here)
myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))

fill_limits <- c(3, 1750)
brks <- c(10, 100, 1000)

legend_title <- "Cases averted" 
pct_by <- c("ar", "duration", "size", "cfr")

plts_ca <- vector('list', 4)

q_labels <- c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))
i <- 3
plt_1  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = q_labels) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    annotate(
      "segment",
      x = 9, y = 0,
      xend = c(9, 1), # End points for the three arrows
      yend = c(0.9, 0),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4
    ) +
    annotate(
      "point",
      x = c(9), # The four key coordinates
      y = c(0),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "segment",
      x = 8, y = 0.5,
      xend = c(8, 1), # End points for the three arrows
      yend = c(0.9, 0.5),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4,
      linetype = "dashed" # Changed to dashed lines
    ) +
    annotate(
      "point",
      x = c(8), # The four key coordinates
      y = c(0.5),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "text",
      x = c(9.5, 5), # x-positions for the two labels
      y = c(0.5, 0.05), # y-positions for the two labels
      label = c(y_annots[i], "Reducing ORI delay"),
      angle = c(90, 0), # Rotate the vertical label
      color = "black",
      # fontface = "bold",
      size = 4,
      hjust = 0.5
    ) +
    labs(
      x = "Week of ORI implementation",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme_x_blank() +
    theme(legend.position = "top")

# Cases averted per 1,000 OCVs
plot_data <- cea_prioritization %>% 
  filter(metric == "med_case_averted_per_1000_OCV")

fill_limits <- c(0.01, 10)
brks <- c(1, 5, 10)

legend_title <- "Cases averted per 1,000 OCVs" 
pct_by <- c("ar", "duration", "size", "cfr")


i <- 1
plt_2  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = q_labels) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    annotate(
      "segment",
      x = 9, y = 0,
      xend = c(9, 1), # End points for the three arrows
      yend = c(0.9, 0),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4
    ) +
    annotate(
      "point",
      x = c(9), # The four key coordinates
      y = c(0),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "segment",
      x = 8, y = 0.5,
      xend = c(8, 1), # End points for the three arrows
      yend = c(0.9, 0.5),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4,
      linetype = "dashed" # Changed to dashed lines
    ) +
    annotate(
      "point",
      x = c(8), # The four key coordinates
      y = c(0.5),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "text",
      x = c(9.5, 5), # x-positions for the two labels
      y = c(0.5, 0.05), # y-positions for the two labels
      label = c(y_annots[i], "Reducing ORI delay"),
      angle = c(90, 0), # Rotate the vertical label
      color = "black",
      # fontface = "bold",
      size = 4,
      hjust = 0.5
    ) +
    labs(
      x = "Week of ORI implementation",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme_x_blank() +
    theme(legend.position = "top")

# DALYs averted per 1,000 OCVs
plot_data <- 
  cea_prioritization %>% filter(metric == "med_daly_averted_per_1000_OCV")

fill_limits <- c(1e-5, 1.3)
brks <- c(0.1, 0.5, 1)
legend_title <- "DALYs averted per 1,000 OCVs" 

i <- 1
plt_3  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = q_labels) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    annotate(
      "segment",
      x = 9, y = 0,
      xend = c(9, 1), # End points for the three arrows
      yend = c(0.9, 0),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4
    ) +
    annotate(
      "point",
      x = c(9), # The four key coordinates
      y = c(0),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "segment",
      x = 8, y = 0.5,
      xend = c(8, 1), # End points for the three arrows
      yend = c(0.9, 0.5),
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black",
      linewidth = 0.4,
      linetype = "dashed" # Changed to dashed lines
    ) +
    annotate(
      "point",
      x = c(8), # The four key coordinates
      y = c(0.5),
      size = 3,
      shape = 21,
      fill = "white",
      color = "black",
      stroke = 1.2
    ) +
    annotate(
      "text",
      x = c(9.5, 5), # x-positions for the two labels
      y = c(0.5, 0.05), # y-positions for the two labels
      label = c(y_annots[i], "Reducing ORI delay"),
      angle = c(90, 0), # Rotate the vertical label
      color = "black",
      # fontface = "bold",
      size = 4,
      hjust = 0.5
    ) +
    labs(
      x = "Week of ORI implementation",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(legend.position = "top")
```

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plt_1 +
  annotate("text", x = -Inf, y = Inf, label = "D",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

p2_tag <- plt_2 +
  annotate("text", x = -Inf, y = Inf, label = "E",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

p3_tag <- plt_3 +
  annotate("text", x = -Inf, y = Inf, label = "F",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Week of ORI implementation",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

plts_right <- (p1_tag / p2_tag / p3_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")
```

### Combine left and right

```{r}
plts_pca_capd <- (plts_pca | plts_right) + 
  plot_layout(ncol=2) + 
  plot_layout(axis_titles = "collect_x") &
  theme(
    legend.position = "top", # put legends at top
    legend.justification = "center", # center horizontally
    legend.box.just = "center",
    legend.box = "horizontal", # arrange horizontally
    legend.text          = element_text(size = 13),
    legend.title         = element_text(size = 13),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 3, l = 0)
  )

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig3_pca_capd_", tstamp(), ".pdf"),
         plts_pca_capd,
         device = cairo_pdf, width = fgw2*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

## Figure S3

Median impact data with 75% coverage

### A Number of cases averted

```{r}
plot_data <- cea_prioritization %>% filter(metric == "med_s_ch_averted_tot")

plot_data %>% 
  filter(vacc_cov == 0.75) %>%
  pull(median_value) %>% range()

# 1. Create the color palette function (your code was correct here)
myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))

y_labs <- c("Attack rate percentile", 
            "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
              "Prioritizing longer outbreaks",
              "Prioritizing larger outbreaks",
              "Prioritizing higher CFR outbreaks")

fill_limits <- c(3, 1750)
brks <- c(10, 100, 1000)

legend_title <- "Cases averted" 
pct_by <- c("ar", "duration", "size", "cfr")

plts_ca <- vector('list', 4)

for (i in 1:4) {
  plts_ca[[i]]  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    labs(
      x = "Delay in ORI implementation (weeks)",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(
      legend.position = "top",
    ) 
}
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plts_ca[[1]] +
  annotate("text", x = -Inf, y = Inf, label = "A",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p2_tag <- plts_ca[[2]] +
  # annotate("text", x = -Inf, y = Inf, label = "B",
  #          hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p3_tag <- plts_ca[[3]] +
  # annotate("text", x = -Inf, y = Inf, label = "C",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p4_tag <- plts_ca[[4]] +
  # annotate("text", x = -Inf, y = Inf, label = "D",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Week of ORI implementation",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")

plts_combined <- (p1_tag / p2_tag / p3_tag / p4_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")+
  plot_layout(guides = "collect") &                    
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/ca_delay_pri_", tstamp(), ".pdf"),
         plts_combined,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### B. Percent cases averted

```{r}
plot_data <- cea_prioritization %>% filter(metric == "med_pct_reduc_case")

plot_data %>% 
  filter(vacc_cov == 0.75) %>%
  pull(median_value) %>% range()

myPalette <- colorRampPalette(viridisLite::plasma(9))

y_labs <- c("Attack rate percentile", "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
                   "Prioritizing longer outbreaks",
                   "Prioritizing larger outbreaks",
                   "Prioritizing higher CFR outbreaks")

fill_limits <- c(2, 85)
brks <- c(20, 40, 60, 80)

legend_title <- "% cases averted" 
pct_by <- c("ar", "duration", "size", "cfr")

plts_pca <- vector('list', 4)

for (i in 1:4) {
  plts_pca[[i]]  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      # trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    labs(
      x = "Delay in ORI implementation (weeks)",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(
      legend.position = "top",
    ) 
}
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plts_pca[[1]] +
  annotate("text", x = -Inf, y = Inf, label = "B",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p2_tag <- plts_pca[[2]] +
  # annotate("text", x = -Inf, y = Inf, label = "B",
  #          hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p3_tag <- plts_pca[[3]] +
  # annotate("text", x = -Inf, y = Inf, label = "C",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous( breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p4_tag <- plts_pca[[4]] +
  # annotate("text", x = -Inf, y = Inf, label = "D",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Delay in ORI implementation (weeks)",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")


plts_combined <- (p1_tag / p2_tag / p3_tag / p4_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")+
  plot_layout(guides = "collect") &                    
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/pca_delay_pri_", tstamp(), ".pdf"),
         plts_combined,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### C. Cases averted per 1,000 OCVs

```{r}
plot_data <- 
  cea_prioritization %>% filter(metric == "med_case_averted_per_1000_OCV")

plot_data %>% 
  filter(vacc_cov == 0.75) %>%
  pull(median_value) %>% range()

myPalette <- colorRampPalette(viridisLite::cividis(9))

y_labs <- c("Attack rate percentile", "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
                   "Prioritizing longer outbreaks",
                   "Prioritizing larger outbreaks",
                   "Prioritizing higher CFR outbreaks")

fill_limits <- c(0.01, 10)
brks <- c(1, 5, 10)

legend_title <- "Cases averted per 1,000 OCVs" 
pct_by <- c("ar", "duration", "size", "cfr")

plts_capd <- vector('list', 4)

for (i in 1:4) {
  plts_capd[[i]]  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))) +
      
    scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
    labs(
      x = "Delay in ORI implementation (weeks)",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(
      legend.position = "top",
    ) 
}

```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plts_capd[[1]] +
  annotate("text", x = -Inf, y = Inf, label = "C",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p2_tag <- plts_capd[[2]] +
  # annotate("text", x = -Inf, y = Inf, label = "B",
  #          hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p3_tag <- plts_capd[[3]] +
  # annotate("text", x = -Inf, y = Inf, label = "C",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous( breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p4_tag <- plts_capd[[4]] +
  # annotate("text", x = -Inf, y = Inf, label = "D",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Delay in ORI implementation (weeks)",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")


plts_combined <- (p1_tag / p2_tag / p3_tag / p4_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")+
  plot_layout(guides = "collect") &                    
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/capd_delay_pri_", tstamp(), ".pdf"),
         plts_combined,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### D. DALYs averted per 1,000 OCVs

```{r}
plot_data <- 
  cea_prioritization %>% filter(metric == "med_daly_averted_per_1000_OCV")

plot_data %>% 
  filter(vacc_cov == 0.75) %>%
  pull(median_value) %>% range()

# Create the color palette function (your code was correct here)
# myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))
# myPalette <- colorRampPalette(viridisLite::plasma(9))
# myPalette <- colorRampPalette(viridisLite::cividis(9))
myPalette <- colorRampPalette(scico::scico(9, palette = "batlow"))
# myPalette <- colorRampPalette(RColorBrewer::brewer.pal(9, "YlGnBu"))
# myPalette <- colorRampPalette(viridisLite::inferno(9))
# myPalette <- colorRampPalette(scico::scico(9, palette = "lajolla"))

y_labs <- c("Attack rate percentile", "Outbreak duration percentile", 
            "Outbreak size percentile", 
            "CFR percentile")

y_annots <- c("Prioritizing higher AR outbreaks", 
                   "Prioritizing longer outbreaks",
                   "Prioritizing larger outbreaks",
                   "Prioritizing higher CFR outbreaks")

fill_limits <- c(1e-5, 1.3)
brks <- c(0.1, 0.5, 1)

legend_title <- "DALYs averted per 1,000 OCVs" 
pct_by <- c("ar", "duration", "size", "cfr")

legend_title <- "DALYs averted per 1,000 OCVs" 

plts_dapd <- vector('list', 4)

for (i in 1:4) {
  plts_dapd[[i]]  <- plot_data %>% 
    filter(percentile_by == pct_by[i], vacc_cov == 0.75) %>%
    ggplot(aes(x = vacc_week, y = prob, 
                 fill = median_value)) +
    geom_tile(color = "white", linewidth = 0.25) +
    scale_fill_gradientn(
      trans = scales::pseudo_log_trans(),
      colours = myPalette(100), # Generate a smooth ramp of 100 colors
      breaks = brks,
      limits = fill_limits, 
      name = legend_title # Legend title
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      breaks = seq(0, 0.8, 0.2),
      labels = c(expression(Q[0*"%"]),
                 expression(Q[20*"%"]),
                 expression(Q[40*"%"]),
                 expression(Q[60*"%"]),
                 expression(Q[80*"%"]))) +
      
    scale_x_continuous(expand = c(0, 0), 
                       breaks = seq(1, 15, 2)) +
    labs(
      x = "Delay in ORI implementation (weeks)",
      y = y_labs[i]
    ) +
    theme_light() +
    theme_tight() +
    theme(
      legend.position = "top",
    ) 
}
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5

# A, B, and C annotations
p1_tag <- plts_dapd[[1]] +
  annotate("text", x = -Inf, y = Inf, label = "D",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p2_tag <- plts_dapd[[2]] +
  # annotate("text", x = -Inf, y = Inf, label = "B",
  #          hjust = hj, vjust = vj, size = sz) +
  scale_x_continuous(breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p3_tag <- plts_dapd[[3]] +
  # annotate("text", x = -Inf, y = Inf, label = "C",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous( breaks=seq(1,15,2))+
  theme_x_blank()+
  coord_cartesian(clip = "off")

p4_tag <- plts_dapd[[4]] +
  # annotate("text", x = -Inf, y = Inf, label = "D",
  #          hjust = hj, vjust = vj, size = sz) + 
  scale_x_continuous("Delay in ORI implementation (weeks)",
                     breaks=seq(1,15,2))+
  coord_cartesian(clip = "off")


plts_combined <- (p1_tag / p2_tag / p3_tag / p4_tag) + 
  plot_layout(ncol=1) +
  plot_layout(axis_titles = "collect_x")+
  plot_layout(guides = "collect") &                    
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/dalyapd_delay_pri_", tstamp(), ".pdf"),
         plts_combined,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

## Figure 4: Cumulative impact during OCV stockpile deployment

Plotting data
```{r}
# summary by subset_id in addition to the grouping variables # 
# grouping variables \in c("vacc_cov","vacc_week","target_vacc_doses",
# "vacc_priority","priority_prob", "subset_id","pp_r")
# columns you pivot on / keep
keys <- c("vacc_cov","vacc_week","target_vacc_doses",
          "vacc_priority","priority_prob","subset_id","pp_r")
value_col <- c("doses_used", "outbreak_count", 
               "s_ch_averted_tot",
               "death_averted_tot",
               "daly_averted_tot", 
               "case_averted_per_1000_OCV",
               "death_averted_per_1000_OCV",
               "daly_averted_per_1000_OCV", 
               "cost_per_case_averted", 
               "cost_per_death_averted", 
               "cost_per_daly_averted")

# --- configuration for the four scenarios ---
x_prob  <- 0.3                # “top 30%” means keep >= 70th percentile
x_pct   <- as.integer(x_prob * 100)
vars_3  <- c("AR","dur","size")

# optional: a tiny helper to map to the exact labels you want
make_scenario_label <- function(vacc_priority, pp_r) {
  if (pp_r == 0) return("All")
  if (pp_r == x_prob && vacc_priority %in% vars_3) {
    switch(vacc_priority,
      "AR"   = sprintf("Q[%d%%](AR)",   x_pct),
      "dur"  = sprintf("Q[%d%%](Dur)",  x_pct),
      "size" = sprintf("Q[%d%%](Size)", x_pct)
    )
  } else NA_character_
}

# filter to the four scenarios and build the plotting factor
d <- cea_subsets %>%
  mutate(pp_r = round(priority_prob, 1)) %>%
  filter(
    # vacc_week == VW,
    vacc_cov  == VC,
    value_column %in% value_col,
    # keep either "All" (pp=0) for AR (pick any) OR the x=70% slice for the three variables
    (pp_r == 0 & vacc_priority  == "AR") | (pp_r == x_prob & vacc_priority %in% vars_3)
  ) %>%
  mutate(
    scenario = mapply(make_scenario_label, vacc_priority, pp_r)
  ) %>%
  filter(!is.na(scenario)) %>%
  mutate(
    # order in legend: AAR, dur, size, All
    scenario = factor(
      scenario,
      levels = c(sprintf("Q[%d%%](AR)", x_pct),
                 sprintf("Q[%d%%](Dur)", x_pct),
                 sprintf("Q[%d%%](Size)", x_pct),
                 "All")
    )
  )
unique(d[,c("pp_r", "vacc_priority")])
# split and join to bring doses_used alongside outcomes
dr <- d %>%
  filter(value_column == value_col[1]) %>%
  select(any_of(keys), scenario, q500)

names(dr) <- c(keys, "scenario", "doses_used")

dw <- d %>%
  filter(value_column %in% value_col[2:length(value_col)]) %>%
  left_join(dr, by = c(keys, "scenario"))
levs <- levels(dw$scenario)  # expects: Q[30%](AR), Q[30%](Dur), Q[30%](Size), All
unique(dw[,c("vacc_priority","pp_r","scenario")])

expr_labels <- setNames(
  list(
    bquote(Q[30*"%"](AR)),
    bquote(Q[30*"%"](Dur)),
    bquote(Q[30*"%"](Size)),
    expression(All)
  ),
  c(paste0("Q[", x_pct, "%](AR)"), 
    paste0("Q[", x_pct, "%](Dur)"),
    paste0("Q[", x_pct, "%](Size)"),
    "All")
)
# 
# # set.seed(42)
# # n_subset_plot <- 30
# # ids <- sample(200, n_subset_plot) # just to reduce plotting load
# ids <- 1:200
# # create nudges to prevent too much overlap between subset id's
# dw[, doses_used_plot := doses_used * rnorm(length(dw$doses_used), 
#                                          mean=1, sd=1.3e-1)]
v_week <- c(1, 9)
alph <- 0.4

# dw_7 <- dw
```

Summary values in the main text
```{r}
# outbreak count
# outbreak count does not vary within each subset (while impact metrics may vary across runid's) and therefore, you could select any values among q025, q250, ..., q975 and mean.
dw %>%
  filter(value_column == "outbreak_count", 
         vacc_week == 1, 
         scenario == "All") %>%
  group_by(target_vacc_doses) %>% 
  summarise(
    across(c(q500),
           list(q500 = ~quantile(.x, probs=0.5),
                q250 = ~quantile(.x, probs=0.25),
                q750 = ~quantile(.x, probs=0.75),
                min = ~min(.x),
                max = ~max(.x))),
    .groups = "drop"
  )

dw %>%
  filter(value_column == "outbreak_count", 
         vacc_week == 1, 
         target_vacc_doses == 200000000) %>%
  group_by(scenario) %>% 
  summarise(
    across(c(q500),
           list(q500 = ~quantile(.x, probs=0.5),
                q250 = ~quantile(.x, probs=0.25),
                q750 = ~quantile(.x, probs=0.75),
                min = ~min(.x),
                max = ~max(.x))),
    .groups = "drop"
  )

# cases averted 
ca <- dw %>%
  filter(value_column == "s_ch_averted_tot") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

# in the order of Q[30%](AR), Q[30%](Dur), Q[30%](Size), and All 
ca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 0)

ca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits=0)

capd <- dw %>%
  filter(value_column == "case_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

capd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 3)

capd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 3)

# compare with the analyses at the individual level
# median values are smaller at the individual level
cea_ori %>% 
  filter(vacc_week %in% c(1,9), vacc_cov == 0.75) %>% 
  group_by(vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV) 

# variation of vaccine impact and efficiency
cea_ori %>% 
  filter(vacc_week %in% c(1,9), vacc_cov == 0.75) %>% 
  group_by(vacc_week) %>%
  summarize(min = min(pct_reduc_case),
            max = max(pct_reduc_case))

cea_ori %>% 
  filter(vacc_week %in% c(1,9), vacc_cov == 0.75) %>% 
  group_by(vacc_week) %>%
  summarize(min = min(case_averted_per_1000_OCV),
            max = max(case_averted_per_1000_OCV))

dapd <- dw %>%
  filter(value_column == "daly_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

dapd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 3)

dapd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 3)

cpca <- dw %>%
  filter(value_column == "cost_per_case_averted") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

cpca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 0)

cpca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 0)

# cost per DALY averted
cpda <- dw %>%
  filter(value_column == "cost_per_daly_averted") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

cpda %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 0)

cpda %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 0)
```

### Left panel

Plotting parameters
```{r}
pt_size <- 1.35

make_log_offsets_span <- 
  function(x, group, levels = NULL, base = 10, span_factor = 1.5) {
    if (!is.null(levels)) group <- factor(group, levels = levels, ordered = TRUE)
    k <- nlevels(factor(group))
    if (k == 1) return(x)
    step_dex <- log(span_factor, base = 10) / (k - 1) # per-step dex from total span
    idx <- as.integer(factor(group)) - (k + 1) / 2
    x * (base^(step_dex * idx))
  }
span_fctr <- 1.4
shp_wk1 <- 17
shp_wk9 <- 1
```

Number of targeted outbreaks
```{r}
ocv_dose_2020 <- global_ocv_use$cumulative_icg[1] * 1e6
ocv_dose_2024 <- global_ocv_use$cumulative_icg[2] * 1e6

dose_marks <- tibble::tibble(
  x = c(ocv_dose_2020, ocv_dose_2024),
  lab = c("2020", "2024")
)

cnt_iqr_summary <- dw %>%
  filter(value_column == "outbreak_count") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

cnt_iqr_summary_off <- cnt_iqr_summary %>%
  filter(vacc_week == 1) %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_1 <- cnt_iqr_summary_off %>% 
  filter(vacc_week == 1) %>%
  ggplot(aes(x = x_off, 
             y = q500, color = scenario)) +
  geom_linerange(aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(aes(y=q500), size=pt_size, shape=1) +
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  geom_text(
    data = dose_marks,
    inherit.aes = FALSE,
    aes(x = x, y = Inf, label = lab),
    vjust = 1.5,           # push slightly down from the top
    hjust = 1.1,
    nudge_x = 0,           # no horizontal nudge
    size = 3
  )+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_log10("Number of targeted outbreaks") +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  theme_light() +
  theme_x_blank() +
  theme(legend.position = "top") +
  theme_tight_legend(box_space_pt = 3.5,
                     key_space_cm = 0.03)
# plt_1
```

Total suspected cases averted (s_ch_averted_tot)
```{r}
df <- data.frame(x=sort(unique(dw$doses_used_plot)))
df$y <- df$x

ca_iqr_summary <- dw %>%
  filter(value_column == "s_ch_averted_tot") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # `q500` is also used as a variable inside the summarize_quantile() function
  summarize_quantiles(x)

ca_iqr_summary_off <- ca_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_2 <- 
  ggplot(ca_iqr_summary_off, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(ca_iqr_summary_off, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=ca_iqr_summary_off, 
             aes(y=q500, shape=as.factor(vacc_week)), size=pt_size)+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_log10("Total cases averted", 
                breaks = c(1e1,1e2,1e3,1e4,1e5,1e6),
                labels = c(1e1,1e2,1e3,1e4,1e5,1e6)) +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  scale_shape_manual("",
    values = c("1" = shp_wk1, "9" = shp_wk9),
    labels = c("1 week", "9 weeks")
  )+
  theme_light() +
  theme_x_blank() +
  guides(color = "none", shape="none")

# plt_2
```

Case averted per 1,000 OCV doses (case_averted_per_dose)

```{r}
capd_iqr_summary <- dw %>%
  filter(value_column == "case_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

capd_iqr_summary <- capd_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_3 <- 
  ggplot(capd_iqr_summary, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(capd_iqr_summary, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=capd_iqr_summary, 
             aes(y=q500, shape=as.factor(vacc_week)), size=pt_size)+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_continuous("Case averted per 1,000 OCVs", 
                     breaks=c(0:5,7.5,10,12.5,15),
                     labels=c(0:5,"",10,"",15),
                     minor_breaks = NULL) +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  scale_shape_manual("",
    values = c("1" = shp_wk1, "9" = shp_wk9),
    labels = c("1 week", "9 weeks")
  )+
  theme_light() +
  guides(color = "none", shape="none")

# plt_3
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5
# x <- 0 because of log scale
# A, B, and C annotations
plt_1_tag <- plt_1 +
  annotate("text", x = 0, y = Inf, label = "A",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_log10("")+
  coord_cartesian(clip = "off")


plt_2_tag <- plt_2 +
  annotate("text", x = 0, y = Inf, label = "B",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_log10("")+
  coord_cartesian(clip = "off")

plt_3_tag <- plt_3 +
  annotate("text", x = 0, y = Inf, label = "C",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_log10("Vaccine doses deployed")+
  coord_cartesian(clip = "off")

plts_ca_agg <- (plt_1_tag / plt_2_tag / plt_3_tag) +
  plot_layout(ncol = 1) &
  theme(legend.position = "top")

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig4_agg_left_", tstamp(), ".pdf"),
         plts_ca_agg,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### Right panel

DALY averted per 1,000 OCV doses (case_averte_per_dose)

```{r}
dapd_iqr_summary <- dw %>%
  filter(value_column == "daly_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

dapd_iqr_summary <- dapd_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_4 <- 
  ggplot(dapd_iqr_summary, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(dapd_iqr_summary, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=dapd_iqr_summary, 
             aes(y=q500, shape=as.factor(vacc_week)),
             size=pt_size)+
  geom_text(
    data = dose_marks,
    inherit.aes = FALSE,
    aes(x = x, y = Inf, label = lab),
    vjust = 1.5,           # push slightly down from the top
    hjust = 1.1, 
    nudge_x = 0,           # no horizontal nudge
    size = 3
  )+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_continuous("DALY averted per 1,000 OCVs") +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  scale_shape_manual("",
    values = c("1" = shp_wk1, "9" = shp_wk9),
    labels = c("week 1", "week 9")
  )+
  theme_light() +
  guides(color = "none")+
  theme_x_blank()+
  theme(legend.position = "top") +
  theme_tight_legend(box_space_pt = 3,
                     key_space_cm = 0.03)

# plt_4
```

Cost per case averted (cost_per_case_averted)

```{r}
cpca_iqr_summary <- dw %>%
  filter(value_column == "cost_per_case_averted") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

cpca_iqr_summary <- cpca_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_5 <- 
  ggplot(cpca_iqr_summary, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(cpca_iqr_summary, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=cpca_iqr_summary, 
             aes(y=q500, shape=as.factor(vacc_week)), size=pt_size)+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_log10("Cost per case averted") +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
   scale_shape_manual("",
    values = c("1" = 17, "9" = 1),
    labels = c("1 week", "9 weeks")
  )+
  theme_light() +
  theme_x_blank() +
  guides(color = "none", shape="none")

# plt_5
```

Cost per DALY averted (cost_per_daly_averted)
```{r}
cpda_iqr_summary <- dw %>%
  filter(value_column == "cost_per_daly_averted") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x)

cpda_iqr_summary <- cpda_iqr_summary %>%
  mutate(
    scenario = factor(scenario, levels = levs),
    x_off = make_log_offsets_span(target_vacc_doses, 
                                  scenario, levels = levs, 
                                  span_factor = span_fctr)
)

plt_6 <-
  ggplot(cpda_iqr_summary, aes(x=x_off, color=scenario)) +
  geom_linerange(data=subset(cpda_iqr_summary, vacc_week == 9), 
                 aes(ymin=q250, ymax=q750), 
                 position=position_identity(), 
                 linewidth=line_width_95) + 
  geom_point(data=cpda_iqr_summary, 
             aes(y=q500, shape=as.factor(vacc_week)), size=pt_size)+
  geom_vline(xintercept = ocv_dose_2020, 
             linetype = "dashed", color = "firebrick")+
  geom_vline(xintercept = ocv_dose_2024, 
             linetype = "dashed", color = "firebrick")+
  scale_x_log10("Vaccine doses deployed") +
  scale_y_log10("Cost per DALY averted") +
  scale_color_manual("",
    breaks = levs,
    labels = unname(expr_labels[levs]),
    values = my_discrete_colors[1:4]
  )+
  scale_shape_manual("",
    values = c("1" = shp_wk1, "9" = shp_wk9),
    labels = c("1 week", "9 weeks")
  )+
  theme_light() +
  guides(color = "none", shape="none")
 
# plt_6
```

Combine, tag, and save

```{r}
hj <- 3.4
vj <- -0.7
sz <- 5
# x <- 0 because of log scale
# A, B, and C annotations
plt_4_tag <- plt_4 +
  annotate("text", x = 0, y = Inf, label = "D",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_log10("")+
  coord_cartesian(clip = "off")

plt_5_tag <-  plt_5 +
  annotate("text", x = 0, y = Inf, label = "E",
           hjust = hj, vjust = vj, size = sz) +
  scale_x_log10("")+
  coord_cartesian(clip = "off")

plt_6_tag <-  plt_6 +
  annotate("text", x = 0, y = Inf, label = "F",
           hjust = hj, vjust = vj, size = sz) + 
  scale_x_log10("Vaccine doses deployed")+
  coord_cartesian(clip = "off")

plts_cpca_agg <- (plt_4_tag / plt_5_tag / plt_6_tag) + 
  plot_layout(ncol = 1) &
  theme(legend.position = "top",
        legend.box = "vertical",
        legend.spacing.y = unit(3,"pt")) # <-- space between the two legend boxes)

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig4_agg_right_", tstamp(), ".pdf"),
         plts_cpca_agg,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw2*fac*fac, units = "mm")
}
```

### Combine

Left and right panels
```{r}
plts_pca_cpca <- (plts_ca_agg | plts_cpca_agg) + 
  plot_layout(ncol=2) + 
  plot_layout(axis_titles = "collect_x") &
  theme(
    legend.position = "top", # put legends at top
    legend.justification = "center", # center horizontally
    legend.box.just = "center",
    legend.box = "horizontal", # arrange horizontally
    legend.text          = element_text(size = 13),
    legend.title         = element_text(size = 13),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 3, l = 0)
  )
   
update_geom_defaults("point", list(size = 4))
leg_pt <- guide_legend(override.aes = list(size = 4))
# Optional: larger legend symbols in the final combined figure
plts_pca_cpca <- plts_pca_cpca + guides(color = leg_pt, shape = leg_pt)

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/fig4_pca_cpca_agg_", tstamp(), ".pdf"),
         plts_pca_cpca,
         device = cairo_pdf, width = fgw2*fac,
         height = fgw2*fac*fac, units = "mm")
}
```


## Figure S6. Dynamic model fits 

Illustrate model fits that are good or bad and
Model fits were assumed to be poor if:

  1. Parameter values are too close to the bound 
  2. The number of data points are fewer than 8, two times the number of parameters to      be estimated (i.e., 4)
  3. Modeled trajectories didn't capture the dynamics of outbreaks
  
```{r}
# dynamic model fits
fit <- fread("outputs/paper/fit_4p_fA010_20241203T18.csv")
fit_ndata <- fread("outputs/paper/fit_params_ndata_20241206T08.csv")
fit_vis <- read_xlsx("outputs/paper/vacc_fit_params_20241203T2122.xlsx")

max_R0 <- 30 # maximum value for R0, above which the value is suspicious
n_crit <- 8 # minimum number of data points to ensure stability in parameter estimates
get_ids_extreme_est <- function(x, er=0.05) {
  # susceptible pop is too close to 1 or 0
  # R0 is too close to a upper bound 30
  ids <- ifelse(x$s0 < 1*er | x$s0 > 1*(1-er) | x$R0 > max_R0*(1-er), TRUE, FALSE) 
  x$id_outbreak[ids]
}

get_ids_low_ndata <- function(x, threshold=8) {
  # number of data points are too small, may have resulted in ustable parameter values.
  ids <- ifelse(x$ndata < threshold, TRUE, FALSE)
  x$id_outbreak[ids]
}

get_ids_poor_fit <- function(x) {
  # visually assessed -- dynamic model outputs reasonally capture data?
  ids <- ifelse(is.na(fit_vis$Good_Fit_Visual), FALSE, TRUE)
  x$id_outbreak[ids]
}

# ids_robust <- unique(c(ids1 %in%, ids2))
ids1 <- get_ids_extreme_est(fit, er=0.2)
ids2 <- get_ids_low_ndata(fit_ndata, threshold=8)
ids3 <- get_ids_poor_fit(fit_vis)

ids_poor_fit <- unique(c(ids1, ids2, ids3))
ids_robust <- setdiff(fit$id_outbreak, ids_poor_fit)
```

plot

```{r}
inc <- read_csv("outputs/paper/incidence_comparison_20241231T15.csv")
inc <- inc %>% 
  group_by(id_outbreak) %>%
  mutate(week = row_number())
             
# bad fit
# good fit
p1 <- inc %>% 
  filter(id_outbreak == ids_robust[1]) %>% 
  ggplot(aes(x = week)) +
  geom_point(aes(y = inc_data, color = "Data"),
             size = 2) +
  geom_line(aes(y = inc_model, color = "Model"),
            linewidth = 1) +
  scale_color_manual(
    name = NULL,
    values = c(
      "Data" = "black",
      "Model" = "firebrick"
    )
  ) +
  labs(
    x = "Epidemic week",
    y = "Weekly suspected cholera"
  ) +
  scale_x_continuous(breaks=seq(0,10,2))+
  theme_tight()+
  theme_light(base_size = 12) +
  theme(
    legend.position = "top",
    legend.justification = "center"
  )

# bad fit
p2 <- inc %>% 
  filter(id_outbreak == ids_poor_fit[2]) %>% 
  ggplot(aes(x = week)) +
  geom_point(aes(y = inc_data, color = "Data"),
             size = 2) +
  geom_line(aes(y = inc_model, color = "Model"),
            linewidth = 1) +
  scale_color_manual(
    name = NULL,
    values = c(
      "Data" = "black",
      "Model" = "firebrick"
    )
  ) +
  labs(
    x = "Epidemic week",
    y = ""
  ) +
  theme_tight()+
  theme_light(base_size = 12) +
  guides(color="none")
  

p12 <- (p1 | p2) +
  plot_layout(axis_titles = "collect_x") &
  theme(
    legend.position = "top", # put legends at top
    legend.justification = "center", # center horizontally
    legend.box.just = "center",
    legend.box = "horizontal", # arrange horizontally
    legend.text          = element_text(size = 13),
    legend.title         = element_text(size = 13),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 3, l = 0)
  )

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/figS6_ts_data_dynamic_model_", tstamp(), ".pdf"),
         p12,
         device = cairo_pdf, width = fgw2*fac,
         height = fgw1*fac*fac, units = "mm")
}
```

## Figure S7: Overall vaccine effectiveness between dynamic and static models

```{r}
# static model
# OVE are based on the definition
# DVE and IVE are given
# TVE = 1 - (1-DVE)(1-IVE)
# OVE = 1 - (vc*(1-DVE)(1-IVE) + (1-vc)*(1-IVE))
setDT(outbreak_data)
pop_by_age_wpp <- as.data.frame(fread("outputs/paper/wpp_pop_by_age_20241022.csv"))
age_dist <- pop_by_age_wpp
# extract outbreak data, can be done by including only the first rwo
od <- outbreak_data[week == 1]
od$prop_u5 <- NA
for (i in 1:nrow(od)) {
  od$prop_u5[i] <- 
    age_dist[(age_dist$`ISO3 Alpha-code` == od$country[i] &
                as.character(age_dist$Year) == as.character(od$year[i])), "prop_u5"]
}

params <- as.data.frame(fread("outputs/paper/params_20251020.csv"))
vacc_par <- params[,c("vacc_effect_direct_u5", "vacc_effect_direct_5p", "vacc_effect_indirect_id")]
increment <- 0.02 # increment used in posterior predictive values

lst <- vector('list', nrow(od))
vacc_eff_static <- data.frame()
vcs <- c(0.6, 0.75, 0.9)
nsample <- 200
for (vc in vcs) {
  for (i in 1:nrow(od)) {
    d <- od[i, ]
    d <- rbindlist(lapply(1:100, function(x) rbind(d,d)))
    d$runid <- NA
    d$vacc_cov <- vc
    d$ve_u5 <- NA
    d$ve_5p <- NA
    d$dve <- NA
    d$ive <- NA
    d$tve <- NA
    d$ove <- NA
    
    for (j in 1:nsample) {
      d$runid[j] <- j
      d$ve_u5[j] <- vacc_par$vacc_effect_direct_u5[j]
      d$ve_5p[j] <- vacc_par$vacc_effect_direct_5p[j]
      d$dve[j] <- d$prop_u5[j]*d$ve_u5[j] + (1-d$prop_u5[j])*d$ve_5p[j]
      # d$ive[j] <- vacc_par$ive[j]
      d$ive[j] <- as.numeric(ive_data[vacc_par$vacc_effect_indirect_id[j], (round(vc/increment)+1)])
      d$tve[j] <- 1 - (1-d$dve[j])*(1-d$ive[j])
      d$ove[j] <- 1 - (vc*(1-d$dve[j])*(1-d$ive[j]) + (1-vc)*(1-d$ive[j]))
    }
    lst[[i]] <- d
  }
  vacc_eff_static <- rbindlist(list(vacc_eff_static, rbindlist(lst)))
}


# dynamic model
# Vaccine impact simulation results for the dynamic models
d <- fread("outputs/paper/vacc_fit_params_by_vacc_state_20241206T08.csv")

vcs <- c(0.6, 0.75, 0.9)

lst <- vector('list', length(vcs))
for (i in 1:length(vcs)) {
  vc <- vcs[i]
  df <- d[, c("runid","data_id","id_outbreak")]
  v <- round(100*vc, digits=0)
  df$vacc_cov <- vc
  df$dve <- 1 - (d[[paste0("CI_VC", v, "_vacc")]]/vc)/(d[[paste0("CI_VC", v, "_novacc")]]/(1-vc))
  df$ive <- 1 - (d[[paste0("CI_VC", v, "_novacc")]]/(1-vc))/d[[paste0("CI_novacc")]]
  df$tve <- 1 - (d[[paste0("CI_VC", v, "_vacc")]]/vc)/d[[paste0("CI_novacc")]]
  df$ove <- 1 - (d[[paste0("CI_VC", v, "_vacc")]] + d[[paste0("CI_VC", v, "_novacc")]])/d[[paste0("CI_novacc")]]
  
  lst[[i]] <- df
}
vacc_eff_dynamic <- rbindlist(lst)

vacc_eff_static_dynamic <- 
  left_join(vacc_eff_static, vacc_eff_dynamic, 
            by=c("id_outbreak", "vacc_cov", "runid"))

vacc_eff_static_dynamic <- vacc_eff_static_dynamic %>% 
  mutate(diff_ove = ove.y - ove.x) 

vacc_eff_static_dynamic %>% 
  filter(id_outbreak %in% ids_robust, vacc_cov == 0.9) -> vesd_temp

summary(vesd_temp$diff_ove, na.rm=T) 
quantile(vesd_temp$diff_ove, probs=c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=T)
hist(vesd_temp$diff_ove)


#  fit
plt <- vacc_eff_static_dynamic %>% 
  filter(id_outbreak %in% ids_robust, vacc_cov == 0.9) %>% 
  ggplot(aes(x = diff_ove)) +
  geom_histogram() +
  labs(
    x = "Difference",
    y = "Count"
  ) +
  theme_tight()+
  theme_light(base_size = 12)
  

if (save_plot) {
  fac <- 1.35
  ggsave(paste0("plots/figS7_diff_static_dynamic_", tstamp(), ".pdf"),
         plt,
         device = cairo_pdf, width = fgw1*fac*fac,
         height = fgw1*fac*fac, units = "mm")
}
```

# Result Tables

## Table 3. Impact under the outbreak onset-triggered scenario

All and top 10% by attack rate, duration, and size
```{r table3}
# Create summary for time-triggered vaccination at different times
cea_ori[(vacc_cov == VC & vacc_week %in% c(1, 13))] %>% 
  group_by(vacc_week) %>% 
  summarize_quantiles(case_averted_per_1000_OCV) -> capd

format_median_iqr(capd, row=1, digits=3)
format_median_iqr(capd, row=2, digits=3)

cea_ori[(vacc_cov == VC & vacc_week %in% c(1, 13))] %>% 
  group_by(vacc_week) %>% 
  summarize_quantiles(cost_per_case_averted) -> cpca

format_median_iqr(cpca, row=1, digits=0)
format_median_iqr(cpca, row=2, digits=0)

##
wks <- c(1, 5, 9, 13)
tab <- data.frame()

for (wk in wks) {
  d <- cea_ori[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp)
}

# top 10% in attack rate
dat <- cea_ori %>% filter(id_outbreak %in% top_10_ar_ids)
for (wk in wks) {
  d <- dat[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp) 
}

# top 10% in duration
dat <- cea_ori %>% filter(id_outbreak %in% top_10_dur_ids)
for (wk in wks) {
  d <- dat[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp)
}

# top 10% in size
dat <- cea_ori %>% filter(id_outbreak %in% top_10_size_ids)
for (wk in wks) {
  d <- dat[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp)
}

tab$week <- rep(wks, 4)
tab$group <- rep(c("All","Top 10% attack rate",
                    "Top 10% duration","Top 10% size"), each = 4)

# format number of outbreaks with percentages
tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/length(unique(cea$id_outbreak))*100, digits=1), "%)") 

if (save_table) {
  fwrite(tab, 
         paste0("outputs/paper/table3_onset_triggered_", 
                tstamp(), ".csv"), 
            row.names = FALSE)
}
```

## Table S1. Impact under the outbreak onset-triggered scenario

Broader vaccine coverage and ORI implementation delays

```{r tableS1}
# Create summary for time-triggered vaccination at different times
wks <- c(1, 5, 9, 13)
tab <- data.frame()

for (vc in VC3) {
  for (wk in wks) {
    d <- cea_ori[(vacc_cov == vc & vacc_week == wk)]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
    tab <- rbind(tab, tab_temp)
  }  
}
tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/length(unique(cea$id_outbreak))*100, digits=1), "%)")

tab$week <- rep(wks, length(VC3))
tab$vc <- rep(VC3, each=length(wks))

if (save_table) {
  fwrite(tab, paste0("outputs/paper/tableS1_onset_triggered_vc_vw_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```

## Table 4. Case-triggered scenario

```{r table4}
# Create summary for case-triggered vaccination at different thresholds
vwks <- c(1, 9)
ctgs <- c(1, 10, 50, 140)
tab <- data.frame()

for (wk in vwks) {
  for (ctg in ctgs) {
    d <- cea_ct_ori[(vacc_cov == VC & vacc_week == wk & 
                       case_trigger == ctg)]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
    tab <- rbind(tab, tab_temp)
  }
}

tab$case_trigger <- rep(ctgs, length(vwks))
tab$vacc_week <- rep(vwks, each=length(ctgs))

tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/nrow(outbk_data_summary)*100, digits=1), "%)")

if (save_table) {
  fwrite(tab, paste0("outputs/paper/table4_case_triggered_", tstamp(), ".csv"),
            row.names = FALSE)
}
```

## Table S2. Onset-triggered ORIs across varying delays to stop ORI after outbreak end

```{r}
# Create summary for case-triggered vaccination at different thresholds
vwks <- c(1, 5, 9, 13)
lags <- c(1, 3, 5, 7)
tab <- data.frame()

for (wk in vwks) {
  for (lag in lags) {
    d <- cea_weekly[(ori_occurred & 
                     vacc_week == wk & 
                       no_ori_delay_outbreak_end == lag)]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
    tab <- rbind(tab, tab_temp)
  }
}

tab$no_ori_lag <- rep(lags, length(vwks))
tab$vacc_week <- rep(vwks, each=length(lags))

tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/length(unique(cea_weekly$id_outbreak))*100, digits=1), "%)")

if (save_table) {
  fwrite(tab, paste0("outputs/paper/tableS2_no_ori_lag_", tstamp(), ".csv"),
            row.names = FALSE)
}
```

## Table S3: Onset-triggered ORIs across alternative outbreak dataset (expanding size limit)

```{r tableS1}
# 75% vaccine coverage only
unique(cea_alt_data$vacc_cov)

wks <- c(1, 5, 9, 13)
tab <- data.frame()
cea_alt_ori <- cea_alt_data[(ori_occurred == TRUE)]

for (wk in wks) {
  d <- cea_alt_ori[vacc_week == wk]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- length(unique(d$id_outbreak))
  tab <- rbind(tab, tab_temp)
}

tab$week <- wks
# format number of outbreaks with percentages
tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/length(unique(cea_alt_data$id_outbreak))*100, digits=1), "%)") 
if (save_table) {
  fwrite(tab, paste0("outputs/paper/tableS3_alt_data_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```



# Discussion

```{r}
impact <- cea_ori %>%
  filter(vacc_cov %in% c(0.75), vacc_week %in% c(1,5,9,13)) %>%
  group_by(vacc_cov, vacc_week) %>%
  summarize_quantiles(s_ch_averted_tot, pct_reduc_case, 
                      case_averted_per_1000_OCV,
                      daly_averted_per_1000_OCV,
                      cost_per_case_averted, 
                      cost_per_daly_averted)


impact %>% filter(variable == "pct_reduc_case") 
impact %>% filter(variable == "case_averted_per_1000_OCV")
impact %>% filter(variable == "daly_averted_per_1000_OCV")
impact %>% filter(variable == "cost_per_case_averted")
impact %>% filter(variable == "cost_per_daly_averted")
# 9 weeks or longer as being representative 
fit_decay[["9"]] %>%
  mutate(rate_reduc = 100*(1 - exp(- k_hat))) %>% 
  summarize_group(vars=c("rate_reduc"), group_name="> 9 weeks") %>% 
  format_median_iqr(digits=1)

fit_decay[["9"]] %>% 
  summarize_group(vars=c("half_life"), group_name="> 9 weeks") %>% 
  format_median_iqr(digits=1)

impact %>% filter(variable == "pct_reduc_case") 
# top quartile

# top quartile in attack rate
cea_ori %>% 
  filter(vacc_cov %in% c(0.75), 
         vacc_week %in% c(5)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(pct_reduc_case)

cea_ori %>% 
  filter(id_outbreak %in% top_q_dur_ids, 
         vacc_cov %in% c(0.75), 
         vacc_week %in% c(13)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(pct_reduc_case)

cea_ori %>% 
  filter(vacc_cov %in% c(0.75), 
         vacc_week %in% c(1,5,9,13)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV)

cea_ori %>% 
  filter(id_outbreak %in% top_q_ar_ids, 
         vacc_cov %in% c(0.75), 
         vacc_week %in% c(1,5,9,13)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(case_averted_per_1000_OCV)
# dw from Figure 4
ca <- dw %>%
  filter(value_column == "s_ch_averted_tot") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

ca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 0)

ca %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 0)

capd <- dw %>%
  filter(value_column == "case_averted_per_1000_OCV") %>%
  group_by(vacc_week, scenario, target_vacc_doses) %>%
  mutate(x = q500) %>% # new variable to reduce confusion for the following calculations
  summarize_quantiles(x) 

capd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 9) %>% 
  format_median_iqr(digits = 3)

capd %>% 
  filter(target_vacc_doses == 200e6, vacc_week == 1) %>% 
  format_median_iqr(digits = 3)


# Cost per DALY averted (CPDALYA)
cea_ori %>% 
  filter(vacc_cov %in% c(0.75), 
         vacc_week %in% c(1)) %>%
  summarize_quantiles(cost_per_daly_averted)

# CPDALYA in
cea_ori %>%
  filter(id_outbreak %in% top_q_ar_ids,
         vacc_cov %in% c(0.75),
         vacc_week %in% c(1)) %>%
  group_by(vacc_week) %>%
  summarize_quantiles(cost_per_daly_averted)


```

## Table S4 / Figure S8. Post-hoc ρ/κ observed-to-infection sensitivity band

The model runs in observed suspected-case units. The gap to true *symptomatic*
infections is a single deterministic scalar `m = ρ/κ`, where ρ is the positive
predictive value (PPV) of the suspected-case definition and κ is the
care-seeking proportion. As established in the methods, `m` enters no dynamic,
cost-per-unit, effectiveness, or targeting calculation, so this is a pure
post-hoc rescale — NOT a new simulation. This block reads the stored CEA output
and reports an `m` sensitivity band; it does not alter the model run.

```{r supp-rho-kappa-band}
#| warning: false
#| message: false

set.seed(20240612)

# ---------------------------------------------------------------------------
# 1. ρ and κ as Beta distributions (mean + approximate 95% interval).
#    Fitted by moment matching: SD implied by the 95% interval width, then
#    Beta(α, β) from mean and variance. Sources:
#      ρ  PPV of suspected-case definition: 0.52 (95% CrI 0.24-0.80), Wiens 2023
#      κ  care-seeking proportion (Wiens 2025), two scenarios:
#         "cholera-relevant"  0.586 (95% CI 0.399-0.752)  -> m < 1
#         "general-diarrhea"  0.328 (95% CI 0.281-0.379)  -> m > 1
# ---------------------------------------------------------------------------
fit_beta_mom <- function(mean, lo, hi, z = qnorm(0.975)) {
  v <- ((hi - lo) / (2 * z))^2          # variance implied by the 95% interval
  if (v >= mean * (1 - mean))
    stop("Interval too wide for a Beta with this mean.")
  k <- mean * (1 - mean) / v - 1        # Beta "concentration"
  c(alpha = mean * k, beta = (1 - mean) * k)
}

rho_spec   <- list(mean = 0.52,  lo = 0.24,  hi = 0.80)            # Wiens 2023
kappa_spec <- list(
  `cholera-relevant` = list(mean = 0.586, lo = 0.399, hi = 0.752), # Wiens 2025
  `general-diarrhea` = list(mean = 0.328, lo = 0.281, hi = 0.379)  # Wiens 2025
)

rho_ab   <- fit_beta_mom(rho_spec$mean, rho_spec$lo, rho_spec$hi)
kappa_ab <- lapply(kappa_spec, function(s) fit_beta_mom(s$mean, s$lo, s$hi))

# Print fitted parameters and the Beta-implied 95% interval to check the fit
beta_check <- function(lbl, ab, spec) {
  qi <- qbeta(c(0.025, 0.975), ab["alpha"], ab["beta"])
  data.frame(
    quantity   = lbl,
    alpha      = round(ab["alpha"], 3),
    beta       = round(ab["beta"], 3),
    target_mean = spec$mean,
    fitted_mean = round(ab["alpha"] / (ab["alpha"] + ab["beta"]), 3),
    target_95   = sprintf("%.3f-%.3f", spec$lo, spec$hi),
    fitted_95   = sprintf("%.3f-%.3f", qi[1], qi[2]),
    row.names = NULL)
}
beta_fits <- rbind(
  beta_check("rho (PPV)", rho_ab, rho_spec),
  beta_check("kappa (cholera-relevant)", kappa_ab[["cholera-relevant"]],
             kappa_spec[["cholera-relevant"]]),
  beta_check("kappa (general-diarrhea)", kappa_ab[["general-diarrhea"]],
             kappa_spec[["general-diarrhea"]]))
print(beta_fits)

# ---------------------------------------------------------------------------
# 2. Draw N samples; m = ρ/κ per draw (ρ and κ independent). Summarize m.
# ---------------------------------------------------------------------------
N <- 1e4
band <- function(x) quantile(x, c(0.5, 0.025, 0.975), names = FALSE, na.rm = TRUE)

rho_draws <- rbeta(N, rho_ab["alpha"], rho_ab["beta"])
m_draws   <- lapply(kappa_ab, function(ab) rho_draws / rbeta(N, ab["alpha"], ab["beta"]))

m_summary <- do.call(rbind, lapply(names(m_draws), function(s) {
  q <- band(m_draws[[s]])
  data.frame(scenario = s,
             kappa_beta = sprintf("Beta(%.1f, %.1f)",
                                  kappa_ab[[s]]["alpha"], kappa_ab[[s]]["beta"]),
             m_median = round(q[1], 3),
             m_lo = round(q[2], 3), m_hi = round(q[3], 3),
             row.names = NULL)
}))
print(m_summary)

# ---------------------------------------------------------------------------
# 3. Reference (static, m = 1) absolute outputs from the stored onset CEA
#    output, primary scenario: onset-triggered, vacc_cov = 0.75, vacc_week = 1.
#    Self-contained: reuse cea_ori if present, else read the stored file.
# ---------------------------------------------------------------------------
if (exists("cea_ori")) {
  cea_src <- data.table::as.data.table(cea_ori)
} else {
  cea_src <- data.table::as.data.table(
    readRDS("outputs/paper/cea_onset_trigger_20260612.rds"))
  cea_src <- cea_src[ori_occurred == TRUE]
}
ref <- cea_src[ori_occurred == TRUE & vacc_cov == 0.75 & vacc_week == 1]

ref_inf   <- median(ref$s_ch_averted_tot,        na.rm = TRUE)  # infections averted
ref_per1k <- median(ref$case_averted_per_1000_OCV, na.rm = TRUE)
ref_cost  <- median(ref$cost_per_case_averted,   na.rm = TRUE)

# m multiplies the absolute infection COUNTS (count = m x observed). Cost per
# symptomatic infection averted relabels only the denominator to infections,
# so it scales as 1/m (numerator net_cost stays anchored to treated cases).
#
# INVARIANT TO m BY CONSTRUCTION (left untouched, not rescaled here):
#   pct_reduc_case (ratio of cases), the delay-impact exponential decay
#   (fit on pct_reduc_case), and all attack-rate / outbreak-size rankings
#   (quantile cutoffs are preserved under any constant monotonic m).
adj_count <- function(ref, m) band(ref * m)   # infections, per-1000 doses
adj_cost  <- function(ref, m) band(ref / m)   # cost per infection averted

fmt <- function(v, digits = 0)
  sprintf("%s [%s–%s]",
          formatC(v[1], format = "f", digits = digits, big.mark = ","),
          formatC(v[2], format = "f", digits = digits, big.mark = ","),
          formatC(v[3], format = "f", digits = digits, big.mark = ","))

row_for <- function(label, ref_val, adj_fun, digits) {
  out <- data.frame(Output = label,
                    `Reference (m = 1)` =
                      formatC(ref_val, format = "f", digits = digits,
                              big.mark = ","),
                    check.names = FALSE)
  for (s in names(m_draws))
    out[[s]] <- fmt(adj_fun(ref_val, m_draws[[s]]), digits)
  out
}

tableS4 <- rbind(
  row_for("Symptomatic infections averted (per ORI)", ref_inf,   adj_count, 0),
  row_for("Infections averted per 1,000 doses",       ref_per1k, adj_count, 2),
  row_for("Cost per symptomatic infection averted (US$)", ref_cost, adj_cost, 0))

knitr::kable(
  tableS4,
  caption = paste0(
    "Table S4. Post-hoc ρ/κ observed-to-infection rescale of the ",
    "absolute outputs (onset-triggered, vacc_cov 0.75, vaccination in week 1). ",
    "Reference is the static m = 1 model median; scenario columns give the ",
    "m-adjusted median [95% band] from ρ/κ uncertainty only."))

# ---------------------------------------------------------------------------
# 4. Figure S8: forest of fold-change vs the m = 1 reference (log scale).
#    Counts scale by m; cost per infection scales by 1/m. The two impact
#    metrics share the SAME fold-change (both = m) and are shown as one row.
# ---------------------------------------------------------------------------
lab_impact <- "Symptomatic infections averted\n(and per 1,000 doses)"
lab_cost   <- "Cost per symptomatic\ninfection averted"

# Impact fold-change = m; the two impact metrics are identical -> one row.
fold_impact <- do.call(rbind, lapply(names(m_draws), function(s) {
  q <- band(m_draws[[s]])
  data.frame(Output = lab_impact, scenario = s,
             mid = q[1], lo = q[2], hi = q[3], row.names = NULL)
}))
# Cost per infection averted scales as 1/m: the EXACT reciprocal of the impact
# fold-change (invert the median; swap-and-invert the CI bounds).
fold_cost <- data.frame(Output = lab_cost, scenario = fold_impact$scenario,
                        mid = 1 / fold_impact$mid,
                        lo  = 1 / fold_impact$hi,
                        hi  = 1 / fold_impact$lo, row.names = NULL)
fold_df <- rbind(fold_impact, fold_cost)
# cost as the bottom row, impact above it (first factor level = bottom)
fold_df$Output <- factor(fold_df$Output, levels = c(lab_cost, lab_impact))

# (6) Sanity check: cost fold-change is exactly the reciprocal of the impact one
stopifnot(all(abs(fold_cost$mid - 1 / fold_impact$mid) < 1e-12))
cat(sprintf("reciprocal check: impact %.2f / %.2f  ->  cost %.2f / %.2f\n",
            fold_impact$mid[1], fold_impact$mid[2],
            fold_cost$mid[1], fold_cost$mid[2]))

# Colourblind-safe Okabe-Ito palette
okabe_ito <- c(`cholera-relevant` = "#009E73", `general-diarrhea` = "#E69F00")

p_forest <- ggplot(fold_df, aes(x = mid, y = Output, colour = scenario)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(xmin = lo, xmax = hi),
                  position = position_dodge(width = 0.5)) +
  geom_text(aes(label = sprintf("%.2f", mid)),
            position = position_dodge(width = 0.5),
            vjust = -0.9, size = 3, show.legend = FALSE) +
  scale_x_log10(breaks = c(0.3, 0.5, 1, 2, 3),
                limits = c(0.3, 3),
                expand = expansion(mult = c(0.03, 0.03))) +
  scale_y_discrete(expand = expansion(add = c(0.6, 0.6))) +
  scale_colour_manual(values = okabe_ito,
                      name = "Care-seeking (κ) scenario") +
  guides(colour = guide_legend(title.position = "top", title.hjust = 0)) +
  labs(x = "Fold-change vs. observed (m = 1) reference", y = NULL) +
  theme_light() +                       # project-standard theme (white panel,
  theme_tight() +                       # no grey background) + tight legend
  theme(legend.position = "top",        # smaller text so it fits 120 mm wide
        text = element_text(size = 9),
        legend.text = element_text(size = 8))
print(p_forest)

if (save_plot) {
  base <- paste0("plots/figS8_rho_kappa_band_", tstamp())
  ggsave(paste0(base, ".pdf"), p_forest, width = 120, height = 74,
         units = "mm", device = cairo_pdf)            # vector, renders κ
  ggsave(paste0(base, ".png"), p_forest, width = 120, height = 74,
         units = "mm", dpi = 300, type = "cairo")      # 300-dpi raster, κ-safe
}
```

Figure S8 / Table S4 caption. The factor `m = ρ/κ` converts observed suspected
cases to true symptomatic infections; ρ is the PPV of the suspected-case
definition (Wiens 2023) and κ the care-seeking proportion (Wiens 2025). This is
a deterministic post-hoc rescale of the stored model outputs, **not** a new
simulation run, and the reported band reflects ρ/κ uncertainty alone. The two
κ scenarios bracket the direction of the correction: a cholera-relevant
care-seeking proportion gives m < 1 (median ≈ 0.89), modestly *reducing* the
absolute infection counts and *raising* cost per infection averted, whereas a
general-diarrhea care-seeking proportion gives m > 1 (median ≈ 1.59),
*inflating* the counts and *lowering* cost per infection averted. Because m is a
single constant multiplier, it leaves every relative and ranking result
unchanged — the percentage of cases averted, the delay–impact decay, and the
attack-rate/outbreak-size prioritization are invariant to m by construction and
are therefore not rescaled here.
