---
title: "trash"
output: html_document
date: "2025-08-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Results

## Result Summary Tables

```{r result-tables}
# Load the simulation results
cea_vc60 <- readRDS("outputs/cea_0.6_time_trigger_20250304.rds")
cea_vc75 <- readRDS("outputs/cea_0.75_time_trigger_20250304.rds")
cea_vc90 <- readRDS("outputs/cea_0.9_time_trigger_20250304.rds")

# Combine the results into a single dataset
cea <- rbind(cea_vc60, cea_vc75)
cea <- rbind(cea, cea_vc90)

# Load case-triggered results
cea_ct_vc60 <- readRDS("outputs/cea_0.6_case_trigger_20250311.rds")
cea_ct_vc75 <- readRDS("outputs/cea_0.75_case_trigger_20250311.rds")
cea_ct_vc90 <- readRDS("outputs/cea_0.9_case_trigger_20250311.rds")

# Combine case-triggered results
cea_ct <- rbind(cea_ct_vc60, cea_ct_vc75)
cea_ct <- rbind(cea_ct, cea_ct_vc90)

# Load preemptive vaccination results
cea_pe_vc60 <- readRDS("outputs/cea_0.6_preemptive_20250314.rds")
cea_pe_vc75 <- readRDS("outputs/cea_0.75_preemptive_20250314.rds")
cea_pe_vc90 <- readRDS("outputs/cea_0.9_preemptive_20250314.rds")

# Combine preemptive results
cea_pe <- rbind(cea_pe_vc60, cea_pe_vc75)
cea_pe <- rbind(cea_pe, cea_pe_vc90)

# Create summary statistics table for time-triggered vaccination
tab <- data.frame()
wks <- c(1, 5, 9, 13)
for (wk in wks) {
  d <- cea[(vacc_cov == VC & vacc_week == wk)]
  tab <- rbind(tab, impact_summary(d))
}
tab$week <- wks

if (save_table) {
  fwrite(tab, paste0("outputs/svim_time_trigger_tab_", tstamp(), ".csv"))
}

# Create summary statistics table for case-triggered vaccination
tab_ct <- data.frame()
ctgs <- c(1, 50, 100, 300)
for (ctg in ctgs) {
  d <- cea_ct[(vacc_cov == VC & case_trigger == ctg)]
  tab_ct <- rbind(tab_ct, impact_summary(d))
}
tab_ct$case_trigger <- ctgs
if (save_table) {
  fwrite(tab_ct, paste0("outputs/svim_case_trigger_tab_", tstamp(), ".csv"))
}
```


## Subset Analysis

```{r subset-analysis, fig.width=8, fig.height=10}
# Load subset analysis results
cea_subsets <- readRDS("outputs/cea_0.75_time_tigger_subsets_by_year_20250306.rds")
cea_subsets_case_wt <- readRDS("outputs/cea_0.75_time_trigger_case_weighted_subsets_by_year_20250313.rds")

# Filter results for easier visualization
cea_subsets |> filter(vacc_week == 9, vacc_cov == 0.75) -> dat
cea_subsets_case_wt |> filter(vacc_week == 9, vacc_cov == 0.75) -> dat_wt

# Reference values for individual assessment
cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week) |>
  summarize(vacc_cov = vacc_cov[1],
            count = n(),
            median_prc = median(pct_reduc_case, na.rm = TRUE),
            median_capd = median(case_averted_per_1000_OCV, na.rm = TRUE),
            median_cpca = median(cost_per_case_averted, na.rm = TRUE)) -> ref_val

# Combine random and case-weighted results
dat$sampling <- "Random"
dat_wt$sampling <- "Case-weighted"
d <- rbind(dat, dat_wt)

# Plot 1: OCV deployment over time
plt_a1 <- ocv_year |>
  filter(year <= 2022) |>
  ggplot(aes(x = as.factor(year), y = cum_num_ocv)) +
  geom_point() +
  labs(x = "", y = "Cumulative\nOCVs deployed") +
  theme_light() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# Plot 2: Cases averted
plt_b1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = s_ch_averted_tot + 1, color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7),
               outlier.size = 0.6) +
  labs(x = "", y = "Cumulative\ncases averted") +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  theme_light() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# Plot 3: Percent case reduction
plt_c1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = pct_reduc_case, color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7),
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_prc),
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "% reduction\nin cases") +
  theme_light() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# Plot 4: Cases averted per 1000 OCVs
plt_d1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = case_averted_per_1000_OCV,
             color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7),
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_capd),
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_continuous(limits = c(0, 2)) +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "Cases averted\nper 1,000 OCVs") +
  theme_light() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# Plot 5: Cost per case averted
plt_e1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = cost_per_case_averted, color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7),
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_cpca),
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_log10(limits = c(1, 1e5)) +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "Cost per case\naverted (US$)") +
  theme_light()

# Combine all plots
plt <- wrap_plots(list(plt_a1, plt_b1, plt_c1, plt_d1, plt_e1), ncol = 1,
                  guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0.13, 0.92),
        plot.tag = element_text(size = 12),
        legend.position = "top")

if (save_plot) {
  ggsave(paste0("plots/fig4_cumulative_impact_", tstamp(), ".pdf"),
         plt, device = cairo_pdf, width = 70*2.4, height = 70*2.4, units = "mm")
}

plt  # Display the plot
```

## Visualize Percent Case Reduction

```{r plot-case-reduction}
# Filter data for plotting
svim_tt_med |> filter(vacc_week %in% VW_seq) -> svim_tt_med_vw
svim_ct_med |> filter(case_trigger %in% cts_sel) -> svim_ct_med_c

# Plot time-triggered case reduction
plt3a <- cea |>
  filter(ori_occurred == TRUE) |>
  filter(vacc_cov %in% VC3, vacc_week %in% VW_seq) |>
  ggplot(aes(x = as.factor(vacc_week), y = pct_reduc_case,
             color = as.factor(vacc_cov))) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7),
               outlier.size = 0.6) +
  geom_hline(data = svim_pe_summary, aes(yintercept = max_pct_reduc_case,
                                         color = as.factor(vacc_cov)), linetype = "dashed") +
  geom_line(data = svim_tt_med_vw, aes(x = 1:8, y = fitted),
            color = "black", linewidth = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_color_manual(name = "Vaccine coverage",
                     values = c("#E41A1C", "#377EB8", "#4DAF4A"),
                     labels = c(paste0(format_num(VC3*100, 1), "%"))) +
  labs(x = "", y = "Case reduction") +
  theme_light() +
  theme(legend.position = "top")

# Plot case-triggered reduction
plt3b <- cea_ct |>
  filter(vacc_cov %in% VC3, case_trigger %in% cts_sel) |>
  mutate(case_trigger = factor(case_trigger, levels = cts_sel),
         vacc_cov = factor(vacc_cov, levels = c(0.6, 0.75, 0.9))) |>
  ggplot() +
  geom_boxplot(aes(x = case_trigger, y = pct_reduc_case,
                   group = interaction(case_trigger, vacc_cov),
                   color = vacc_cov),
               width = 0.6, position = position_dodge(width = 0.7),
               outlier.size = 0.6) +
  geom_hline(data = svim_pe_summary, aes(yintercept = max_pct_reduc_case,
                                         color = as.factor(vacc_cov)), linetype = "dashed") +
  geom_line(data = svim_ct_med_c, aes(x = 1:8, y = fitted),
            color = "black", linewidth = 1) +
  scale_y_continuous(labels = NULL) +
  scale_color_manual(name = "Vaccine coverage",
                     values = c("#E41A1C", "#377EB8", "#4DAF4A"),
                     labels = c("60%", "75%", "90%")) +
  labs(x = "", y = "") +
  theme_light() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank()) +
  theme(legend.position = "top")

# Calculate percentiles for outbreak size groups
probs <- seq(0, 1, by = 0.1)
unq_data <- filter(cea, runid == 1 & vacc_cov == 0.75 & vacc_week == 1)
size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)

# Analyze impact by outbreak size for time-triggered vaccination
d <- cea |>
  group_by(vacc_week, vacc_cov, id_outbreak) |>
  summarize(s_ch_tot = s_ch_tot[1],
            pct_reduc_case = median(pct_reduc_case))  # median across runid's

d$size_grp <- sapply(d$s_ch_tot, function(x) get_group_id(x, size_grp_cutoff))

lst <- vector("list", length(size_grp_cutoff) - 1)
d <- setDT(d)
for(i in 1:length(lst)){
  lst[[i]] <- d[size_grp >= i]
}

res_list <- vector('list', length(size_grp_cutoff) - 1)
for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    size_grp = min(size_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm = TRUE)),
    by = .(vacc_week, vacc_cov)]
}
res_rbind <- rbindlist(res_list)

# Plot heatmap of case reduction by outbreak size and vaccination week
plt3c <- res_rbind |>
  filter(vacc_cov %in% VC, vacc_week %in% VW_seq) |>
  ggplot(aes(x = vacc_week, y = size_grp)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case)) +
  metR::geom_contour2(aes(z = pct_reduc_case, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median case reduction", limits = c(0, 100),
                       breaks = seq(0, 100, 50),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Vaccination week since the outbreak began",
       y = "Total case percentile") +
  scale_x_continuous(breaks = VW_seq) +
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                                expression(">="*P[30*"%"]),
                                expression(">="*P[60*"%"]),
                                expression(">="*P[90*"%"]))) +
  theme_light() +
  theme(legend.position = "bottom")

# Analyze impact by outbreak size for case-triggered vaccination
d <- cea_ct |>
  group_by(case_trigger, vacc_cov, id_outbreak) |>
  summarize(s_ch_tot = s_ch_tot[1],
            pct_reduc_case = median(pct_reduc_case))  # median across runid's

d$size_grp <- sapply(d$s_ch_tot, function(x) get_group_id(x, size_grp_cutoff))

lst <- vector("list", length(size_grp_cutoff) - 1)
d <- setDT(d)
for(i in 1:length(lst)){
  lst[[i]] <- d[size_grp >= i]
}

res_list <- vector('list', length(size_grp_cutoff) - 1)
for (i in 1:length(lst)) {
  d <- lst[[i]]
  res_list[[i]] <- d[, .(
    size_grp = min(size_grp),
    pct_reduc_case = median(pct_reduc_case, na.rm = TRUE)),
    by = .(case_trigger, vacc_cov)]
}
res_rbind <- rbindlist(res_list)

# Plot heatmap of case reduction by outbreak size and case trigger
plt3d <- res_rbind |>
  filter(vacc_cov %in% VC, case_trigger %in% cts_sel) |>
  ggplot(aes(x = case_trigger, y = size_grp)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case)) +
  metR::geom_contour2(aes(z = pct_reduc_case, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median case reduction", limits = c(0, 100),
                       breaks = seq(0, 100, 50),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Cumulative cases at vaccination", y = "") +
  scale_x_continuous(breaks = cts_sel) +
  scale_y_continuous(breaks = c(1, 4, 7, 10), labels = NULL) +
  theme_light() +
  theme(legend.position = "bottom")

# Combine all four plots
plt3 <- (plt3a + plt3b + plt3c + plt3d) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              guides = "collect",
              heights = c(1, 1)) &
  theme(plot.tag.position = c(0, 0.9),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Adjust tag positions for top panels
for (i in c(1, 2)) {
  plt3[[i]] <- plt3[[i]] + theme(plot.tag.position = c(0, 0.86))
}

# Save the plot
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig3_ori_all_pcr_outbk_size_", tstamp(hour = TRUE), ".pdf"),
         plt3,
         device = cairo_pdf, width = fgw2*fac, height = fgw2*1.15*fac, units = "mm")
}
```

## Exponential Impact Model Fit

```{r exponential-fit}
# Select case triggers for analysis
cts_sel <- c(1, seq(40, 280, 40))  # case triggers selected

# Prepare data for time-triggered vaccination
cea |>
  filter(vacc_cov %in% VC, vacc_week <= max(VW_seq)) |>
  group_by(vacc_week) |>
  summarize(vacc_cov = vacc_cov[1],
            count = n(),
            median_ca = quantile(s_ch_averted_tot, probs = c(0.5)),
            median_prc = quantile(pct_reduc_case, probs = c(0.5)),
            median_capd = quantile(case_averted_per_1000_OCV, probs = c(0.5), na.rm = TRUE),
            median_cpca = quantile(cost_per_case_averted, probs = c(0.5), na.rm = TRUE)) -> svim_tt_med

# Define exponential decay function
f <- function(t, Asym = 0, lrc, y0) {
  Asym + (y0 - Asym) * exp(-exp(lrc) * t)
}

# Fit exponential model for percent case reduction
fit <- nls(median_prc ~ f(t = vacc_week, Asym = 0, lrc, y0),
           data = svim_tt_med, start = list(lrc = 0.1, y0 = 100))

# Weekly reduction rate
exp_rate <- exp(as.numeric(coef(fit)["lrc"]))
# Half-life (weeks)
half_life <- log(2)/abs(exp(as.numeric(coef(fit)["lrc"])))

# Add fitted values to data
svim_tt_med$fitted <- fitted(fit)

# Similar fits for other metrics
fit2 <- nls(median_ca ~ f(t = vacc_week, Asym = 0, lrc, y0),
            data = svim_tt_med, start = list(lrc = 0.1, y0 = 100))

fit3 <- nls(median_capd ~ f(t = vacc_week, Asym = 0, lrc, y0),
            data = svim_tt_med, start = list(lrc = 0.1, y0 = 2))

# Linear fit for cost per case averted (which increases over time)
fit4 <- lm(median_cpca ~ vacc_week, data = svim_tt_med)

# Prepare data for case-triggered vaccination
cea_ct |>
  filter(vacc_cov %in% VC, case_trigger <= max(cts_sel)) |>
  group_by(case_trigger) |>
  summarize(vacc_cov = vacc_cov[1],
            count = n(),
            median_prc = quantile(pct_reduc_case, probs = c(0.5))) -> svim_ct_med

# Fit exponential model for case-triggered vaccination
fit_ct <- nls(median_prc ~ f(t = case_trigger, Asym = 0, lrc, y0),
              data = svim_ct_med, start = list(lrc = log(0.01), y0 = 100))

# Reduction rate by unit (a single case)
case_rate <- exp(as.numeric(coef(fit_ct)["lrc"]))
# Half-decrease point (cases)
case_half <- log(2)/abs(exp(as.numeric(coef(fit_ct)["lrc"])))

paste0("Median case reduction (%) decreases with increasing number of cumulative cases by ",
       format_num(case_rate*100, digits = 1), "% per case, halving every ",
       format_num(case_half, digits = 1), " additional cases.")

# Add fitted values to data
svim_ct_med$fitted <- fitted(fit_ct)
```

## Impact from Preemptive Vaccination

```{r preemptive-impact}
# Calculate summary statistics for preemptive vaccination by coverage level
cea_pe |>
  filter(vacc_cov %in% VC3) |>
  group_by(vacc_cov) |>
  summarize(mean_pct_reduc_case = mean(pct_reduc_case),
            q500_pct_reduc_case = quantile(pct_reduc_case, probs = c(0.5)),
            q025_pct_reduc_case = quantile(pct_reduc_case, probs = c(0.025)),
            q975_pct_reduc_case = quantile(pct_reduc_case, probs = c(0.975)),
            max_pct_reduc_case = max(pct_reduc_case),
            min_pct_reduc_case = min(pct_reduc_case)) -> svim_pe_summary

paste0("Maximum percent case reduction from pre-emptive vaccination: ",
       format_num(svim_pe_summary$max_pct_reduc_case[1], digits = 1), "%, ",
       format_num(svim_pe_summary$max_pct_reduc_case[2], digits = 1), "%, and ",
       format_num(svim_pe_summary$max_pct_reduc_case[3], digits = 1),
       "% for vaccine coverage rates of 60%, 75%, and 90%, respectively.")
```

# Preemptive vaccination results
cea_pe_vc60 <- read_rds("outputs/cea_0.6_preemptive_20250314.rds")
cea_pe_vc75 <- read_rds("outputs/cea_0.75_preemptive_20250314.rds")
cea_pe_vc90 <- read_rds("outputs/cea_0.9_preemptive_20250314.rds")

# Combine preemptive results
cea_pe <- rbind(cea_pe_vc60, cea_pe_vc75)
rm(cea_pe_vc60, cea_pe_vc75)  # Free memory
cea_pe <- rbind(cea_pe, cea_pe_vc90)
rm(cea_pe_vc90)
cea_pe <- cea_pe %>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak)

# Preemptive vaccination results
cea_pe_vc60 <- read_rds("outputs/cea_0.6_preemptive_20250314.rds")
cea_pe_vc75 <- read_rds("outputs/cea_0.75_preemptive_20250314.rds")
cea_pe_vc90 <- read_rds("outputs/cea_0.9_preemptive_20250314.rds")

# Combine preemptive results
cea_pe <- rbind(cea_pe_vc60, cea_pe_vc75)
rm(cea_pe_vc60, cea_pe_vc75)  # Free memory
cea_pe <- rbind(cea_pe, cea_pe_vc90)
rm(cea_pe_vc90)
cea_pe <- cea_pe %>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak)

# ```{r fig3-prep}
# ---- Calculate impact from preemptive vaccination ----
cea_pe |>
  filter(vacc_cov %in% VC3) |>
  group_by(vacc_cov) |>
  summarize(mean_pct_reduc_case = mean(pct_reduc_case),
            q500_pct_reduc_case = quantile(pct_reduc_case, probs = c(0.5)),
            q025_pct_reduc_case = quantile(pct_reduc_case, probs = c(0.025)),
            q975_pct_reduc_case = quantile(pct_reduc_case, probs = c(0.975)),
            max_pct_reduc_case = max(pct_reduc_case),
            min_pct_reduc_case = min(pct_reduc_case)) -> svim_pe_summary

# Report maximum case reduction potential
message(paste0("Maximum percent case reduction from pre-emptive vaccination: ",
               format_num(svim_pe_summary$max_pct_reduc_case[1], digits = 1), "%, ",
               format_num(svim_pe_summary$max_pct_reduc_case[2], digits = 1), "%, and ",
               format_num(svim_pe_summary$max_pct_reduc_case[3], digits = 1),
               "% for vaccine coverage rates of 60%, 75%, and 90%, respectively."))

# ---- Fit exponential model to time-triggered vaccination results ----
# Prepare data for time-triggered vaccination at 75% coverage
cea |>
  # filter(ori_occurred == TRUE) |> # Only outbreaks that actually occurred
  filter(vacc_cov %in% VC, vacc_week <= max(VW_seq)) |>
  group_by(vacc_week) |>
  summarize(vacc_cov = vacc_cov[1],
            count = n(),
            median_ca = quantile(s_ch_averted_tot, probs = c(0.5)),
            median_prc = quantile(pct_reduc_case, probs = c(0.5)),
            median_capd = quantile(case_averted_per_1000_OCV, probs = c(0.5),
                                   na.rm = TRUE),
            median_cpca = quantile(cost_per_case_averted, probs = c(0.5),
                                   na.rm = TRUE)) -> svim_tt_med

# Define exponential decay function for modeling impact decay over time
f <- function(t, Asym = 0, lrc, y0) {
  Asym + (y0 - Asym) * exp(-exp(lrc) * t)
}

# Fit exponential model for percent case reduction over time
fit <- nls(median_prc ~ f(t = vacc_week, Asym = 0, lrc, y0),
           data = svim_tt_med, start = list(lrc = 0.1, y0 = 100))

# Calculate key decay parameters
(exp_rate <- exp(as.numeric(coef(fit)["lrc"])))  # Weekly reduction rate
(half_life <- log(2)/abs(exp(as.numeric(coef(fit)["lrc"]))))  # Half-life in weeks

# Add fitted values to data
svim_tt_med$fitted <- fitted(fit)

# Fit similar models for other impact metrics
fit2 <- nls(median_ca ~ f(t = vacc_week, Asym = 0, lrc, y0),
            data = svim_tt_med, start = list(lrc = 0.1, y0 = 100))

fit3 <- nls(median_capd ~ f(t = vacc_week, Asym = 0, lrc, y0),
            data = svim_tt_med, start = list(lrc = 0.1, y0 = 2))

# Linear fit for cost per case averted (which increases over time)
fit4 <- lm(median_cpca ~ vacc_week, data = svim_tt_med)

# ---- Fit exponential model to case-triggered vaccination results ----
# Prepare data for case-triggered vaccination at 75% coverage
cea_ct |>
  # filter(ori_occurred == TRUE) |> # Only outbreaks that actually occurred
  filter(vacc_cov %in% VC, case_trigger <= max(cts_sel)) |>
  filter(case_trigger %in% c(10)) |>
  group_by(vacc_week) |>
  summarize(vacc_cov = vacc_cov[1],
            case_trigger = case_trigger[1],
            count = n(),
            median_ca = quantile(s_ch_averted_tot, probs = c(0.5)),
            median_prc = quantile(pct_reduc_case, probs = c(0.5)),
            median_capd = quantile(case_averted_per_1000_OCV, probs = c(0.5),
                                   na.rm = TRUE),
            median_cpca = quantile(cost_per_case_averted, probs = c(0.5),
                                   na.rm = TRUE)) -> svim_ct_med


# Fit exponential model for case-triggered vaccination
fit_ct <- nls(median_prc ~ f(t = vacc_week, Asym = 0, lrc, y0),
              data = svim_ct_med, start = list(lrc = log(0.01), y0 = 100))

# Calculate key decay parameters
(case_rate <- exp(as.numeric(coef(fit_ct)["lrc"])))  # Per-case reduction rate
(case_half <- log(2)/abs(exp(as.numeric(coef(fit_ct)["lrc"]))))  # Half-decrease point

# Report findings
message(paste0("Median case reduction decreases with increasing number of cumulative cases by ",
               format_num(case_rate*100, digits = 1), "% per case, halving every ",
               format_num(case_half, digits = 1), " additional cases."))

# Add fitted values to data
svim_ct_med$fitted <- fitted(fit_ct)
# ```


## Figure 0

Correlation between the cumulative cases at week 9 and the total
cumulative cases and the ORI impact

```{r}
outbreak_data_all %>%
  group_by(id_outbreak) %>%
  mutate(cum_s_ch = cumsum(s_ch)) -> outbreak_data_all

ts_wk_9 <- outbreak_data_all %>%
  filter(week == 9) %>%
  select(id_outbreak, cum_s_ch)

cea_wk_9 <- cea %>%
  filter(vacc_week == 9) %>%
  select(runid, id_outbreak, vacc_week, vacc_cov, s_ch_tot, pct_reduc_case,
         s_ch_averted_tot)

d <- left_join(cea_wk_9, ts_wk_9, by='id_outbreak')

d %>%
  filter(runid == 1, !is.na(cum_s_ch)) %>%
  ggplot(aes(cum_s_ch, s_ch_tot))+
  geom_point()

d %>%
  filter(runid == 1, !is.na(cum_s_ch), vacc_cov == 0.75) %>%
  ggplot(aes(cum_s_ch, pct_reduc_case))+
  geom_point()+
  scale_x_log10()

d %>%
  filter(runid == 1, !is.na(cum_s_ch), vacc_cov == 0.75) %>%
  ggplot(aes(cum_s_ch, s_ch_averted_tot))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()

```


```{r}
# d0 %>%
#   filter(Subregion %in% c("Eastern Africa","Western Africa","Middle Africa")) %>% # included studies
#   filter(id_outbreak %in% outbreak_data$id_outbreak) %>% # included studies
#   ggplot() +
#   geom_line(aes(date, s_ch, group = id_outbreak, color=Subregion),
#             alpha=0.3)+
#   scale_color_manual("", values=my_discrete_colors[1:5])+
#   theme_minimal()+
#   theme(legend.position = "top")


# # Calculate proportions (always includes all subregions)
# df_prop <- df_complete %>%
#   group_by(date_bin) %>%
#   mutate(prop_cases = total_cases / sum(total_cases, na.rm = TRUE)) %>%
#   ungroup()


# Plot
ggplot(df_prop, aes(x = date_bin, y = prop_cases, fill = Subregion)) +
  geom_area(position = "stack", color = NA, alpha=0.8) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Proportional Contribution of Subregions Over Time",
    x = "Date (Grouped by 2 Months)",
    y = "Proportion of Cases",
    fill = "Subregion"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

d0 %>%
  filter(id_outbreak %in% outbreak_data$id_outbreak) %>% # included studies
  ggplot() +
  geom_point(aes(duration, total_s_ch), alpha = 0.5) +
  # geom_point(data = filter(d, id_outbreak %in% ids_susp),
  #            aes(duration, total_s_ch, color = "Suspected"), alpha = 0.5) +
  # geom_point(data = filter(d, id_outbreak %in% ids_conf),
  #            aes(duration, total_s_ch, color = "Confirmed"), alpha = 0.5)





  d0$outbreak_group <- ifelse(d0$id_outbreak %in% ids_excl, "Excluded",
                              ifelse(d0$id_outbreak %in% ids_conf, "Confirmed",
                                     "Suspected"))

d0$ocv_affected <- ifelse(d0$id_outbreak %in% docv$id_outbreak, "Yes", "No")
d0$excluded_step <- ifelse(d0$id_outbreak %in% excluded1$id_outbreak, "1",
                           ifelse(d0$id_outbreak %in% excluded2$id_outbreak, "2", ifelse(d0$id_outbreak %in% excluded3$id_outbreak, "3", ifelse(d0$id_outbreak %in% excluded4$id_outbreak, "4", "NA"))))


d0 %>%
  group_by(excluded_step) %>%
  summarize(n_id_outbreak = length(unique(id_outbreak)),
            total_s_ch = sum(s_ch),
            total_s_ch_per_outbreak = total_s_ch / n_id_outbreak)

# Create time series plot with different colors by outbreak type
ggplot(d0) +
  geom_line(aes(date, s_ch, group = id_outbreak, color=outbreak_group),
            linewidth=0.1, alpha = 0.5)+
  facet_wrap(~Subregion, ncol=1, scale="free_y")+
  theme_light()

ggplot(d0) +
  geom_line(aes(date, s_ch, group = id_outbreak, color=ocv_affected),
            linewidth=0.1, alpha = 0.5)+
  facet_wrap(~Subregion, ncol=1, scale="free_y")+
  theme_light()

ggplot(d0) +
  geom_line(aes(date, s_ch, group = id_outbreak, color=excluded_step),
            linewidth=0.1, alpha = 0.5)+
  facet_wrap(~Subregion, ncol=1, scale="free_y")+
  theme_light()



geom_line(data = filter(d0, id_outbreak %in% ids_excl),
          aes(date, s_ch, group = id_outbreak, color = "Excluded"),
          alpha = 0.5) +
  geom_line(data = filter(d0, id_outbreak %in% ids_susp),
            aes(date, s_ch, group = id_outbreak, color = "Suspected"),
            alpha = 0.5) +
  geom_line(data = filter(d0, id_outbreak %in% ids_conf),
            aes(date, s_ch, group = id_outbreak, color = "Confirmed"),
            alpha = 0.5) +
  scale_color_manual("", values = c("Excluded" = "yellow4",
                                    "Suspected" = "steelblue4",
                                    "Confirmed" = "firebrick4")) +
  labs(x = "", y = "Weekly suspected cases") +
  theme_light() +
  theme_tight() +
  theme(legend.position = "top") +
  facet_wrap(dSubregion, ncol=1)
```


## Table 4: Impact of Preemptive Vaccination

```{r table4}
# Create summary for preemptive vaccination at different coverage levels
tab_pe <- data.frame()

for (vc in VC3) {
  d <- cea_pe[vacc_cov == vc]
  tab_pe <- rbind(tab_pe, impact_summary(d))
}

tab_pe$vacc_cov <- VC3 * 100  # Convert to percentage

# Format the table with key impact metrics
table4 <- data.frame(
  "Vaccine Coverage %" = tab_pe$vacc_cov,
  "Cases Averted (95% CI)" = paste0(
    format(tab_pe$median_cases_averted, big.mark = ",", digits = 0),
    " (",
    format(tab_pe$lb_cases_averted, big.mark = ",", digits = 0),
    "-",
    format(tab_pe$ub_cases_averted, big.mark = ",", digits = 0),
    ")"
  ),
  "Case Reduction %" = paste0(
    format(tab_pe$median_pct_reduc_case, digits = 1),
    "% (",
    format(tab_pe$lb_pct_reduc_case, digits = 1),
    "-",
    format(tab_pe$ub_pct_reduc_case, digits = 1),
    ")"
  ),
  "Cases Averted per 1,000 OCVs" = paste0(
    format(tab_pe$median_case_averted_per_1000_OCV, digits = 2),
    " (",
    format(tab_pe$lb_case_averted_per_1000_OCV, digits = 2),
    "-",
    format(tab_pe$ub_case_averted_per_1000_OCV, digits = 2),
    ")"
  ),
  "Cost per Case Averted (USD)" = paste0(
    "$", format(tab_pe$median_cost_per_case_averted, big.mark = ",", digits = 0),
    " ($", format(tab_pe$lb_cost_per_case_averted, big.mark = ",", digits = 0),
    "-$", format(tab_pe$ub_cost_per_case_averted, big.mark = ",", digits = 0),
    ")"
  )
)

# Display table
knitr::kable(table4, caption = "Impact of preemptive vaccination at different coverage levels")

if (save_table) {
  write.csv(table4, paste0("outputs/table4_preemptive_impact_", tstamp(), ".csv"),
            row.names = FALSE)
}
```

# Combine into a single table
table5 <- data.frame(
  "Targeting Strategy" = c("Random", "Case-weighted"),
  "Cases Averted (95% CI)" = c(
    paste0(
      format(random_summary$median_cases_averted, big.mark = ",", digits = 0),
      " (",
      format(random_summary$lb_cases_averted, big.mark = ",", digits = 0),
      "-",
      format(random_summary$ub_cases_averted, big.mark = ",", digits = 0),
      ")"
    ),
    paste0(
      format(case_weighted_summary$median_cases_averted, big.mark = ",", digits = 0),
      " (",
      format(case_weighted_summary$lb_cases_averted, big.mark = ",", digits = 0),
      "-",
      format(case_weighted_summary$ub_cases_averted, big.mark = ",", digits = 0),
      ")"
    )
  ),
  "Case Reduction %" = c(
    paste0(
      format(random_summary$median_pct_reduc_case, digits = 1),
      "% (",
      format(random_summary$lb_pct_reduc_case, digits = 1),
      "-",
      format(random_summary$ub_pct_reduc_case, digits = 1),
      ")"
    ),
    paste0(
      format(case_weighted_summary$median_pct_reduc_case, digits = 1),
      "% (",
      format(case_weighted_summary$lb_pct_reduc_case, digits = 1),
      "-",
      format(case_weighted_summary$ub_pct_reduc_case, digits = 1),
      ")"
    )
  ),
  "Cases Averted per 1,000 OCVs" = c(
    paste0(
      format(random_summary$median_case_averted_per_1000_OCV, digits = 2),
      " (",
      format(random_summary$lb_case_averted_per_1000_OCV, digits = 2),
      "-",
      format(random_summary$ub_case_averted_per_1000_OCV, digits = 2),
      ")"
    ),
    paste0(
      format(case_weighted_summary$median_case_averted_per_1000_OCV, digits = 2),
      " (",
      format(case_weighted_summary$lb_case_averted_per_1000_OCV, digits = 2),
      "-",
      format(case_weighted_summary$ub_case_averted_per_1000_OCV, digits = 2),
      ")"
    )
  ),
  "Cost per Case Averted (USD)" = c(
    paste0(
      "$", format(random_summary$median_cost_per_case_averted, big.mark = ",", digits = 0),
      " ($", format(random_summary$lb_cost_per_case_averted, big.mark = ",", digits = 0),
      "-$", format(random_summary$ub_cost_per_case_averted, big.mark = ",", digits = 0),
      ")"
    ),
    paste0(
      "$", format(case_weighted_summary$median_cost_per_case_averted, big.mark = ",", digits = 0),
      " ($", format(case_weighted_summary$lb_cost_per_case_averted, big.mark = ",", digits = 0),
      "-$", format(case_weighted_summary$ub_cost_per_case_averted, big.mark = ",", digits = 0),
      ")"
    )
  )
)

# Display table
knitr::kable(table5, caption = "Comparison of OCV targeting strategies (2022 data, 75% coverage)")


This section visualizes the impact of vaccination on case reduction under different strategies and coverage levels.

## Descriptive statistics of outbreaks
```{r}
outbreak_data %>%
  ggplot(aes(week, s_ch, group=id_outbreak))+
  geom_line(color="steelblue", alpha=0.5)+
  labs(x = "Epidemic week", y = "Suspected cases")+
  theme_light()+
  theme_tight() -> plt

if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig0_outbreaks_ts_", tstamp(), ".pdf"),
         plt,
         device = cairo_pdf, width = fgw1*fac*1.15,
         height = fgw1*fac, units = "mm")
}

names(outbk_data_summary)
unique(outbk_data_summary$country)
unique(outbk_data_summary$year)
table(outbk_data_summary$rural_urban)
probs <- c(0,0.025,0.25,0.5,0.75,0.975,1)

summarize_quantiles(outbk_data_summary, total_s_ch, probs=probs)
summarize_quantiles(outbk_data_summary, total_c_ch, probs=probs)
summarize_quantiles(outbk_data_summary, cfr_per_100percent, probs=probs)
summarize_quantiles(outbk_data_summary, ar_per_10e3, probs=probs)
summarize_quantiles(outbk_data_summary, duration_weeks, probs=probs)

set.seed(42)

# sampling because large dataset makes plotting slow
cea_s <- filter(cea, runid %in% sample_runids) # cea subset
cea_s %>%
  filter(vacc_cov == 0.75 & vacc_week %in% VW_seq) %>%
  group_by(vacc_cov, vacc_week, id_outbreak) %>%
  summarize(med_pct_reduc_case = median(pct_reduc_case,na.rm=T),
            .groups = "drop") -> pcr_wk # median across runid's

# order outbreaks
desired_outbreak_order <- pcr_wk %>%
  filter(vacc_week == 1) %>%
  arrange(desc(med_pct_reduc_case)) %>%
  pull(id_outbreak)

pcr_wk_ordered <- pcr_wk %>%
  mutate(
    # Convert id_outbreak to a factor, applying rev() to the levels
    id_outbreak = factor(id_outbreak, levels = rev(desired_outbreak_order)),
    vacc_week = factor(vacc_week)
  )

# pcr_wk_ordered %>% filter(vacc_week == 1) %>% summarize(max_pcr = max(med_pct_reduc_case))

pcr_wk_ordered |>
  ggplot(aes(x = vacc_week, y = id_outbreak, fill = med_pct_reduc_case)) +
  geom_tile()+
  scale_y_discrete(breaks = NULL, labels = NULL) +
  scale_x_discrete() +
  scale_fill_gradient2(name = "Cases averted (%)   ",
                       labels = function(x) paste0(x, "%"),
                       limits = c(0,100),
                       low = "white",
                       mid = "orange",
                       high = "darkred",
                       midpoint = 45)+
  labs(x = "Delay in ORI implementation (weeks)",
       y = "Outbreaks") +  # Optional title
  theme_minimal() +
  theme_tight()+
  theme(plot.margin = margin(0,0,0,0)) -> plt_a

d <- pcr_wk_ordered %>% filter(vacc_week == 1)

outbk_data_summary_ordered <-
  outbk_data_summary %>%
  mutate(id_outbreak = factor(id_outbreak, levels = rev(desired_outbreak_order)))

# outbk_data_summary_ordered <- left_join(d, outbk_data_summary, by="id_outbreak")
# outbk_data_summary_ordered$id_ordered <- 1192:1

outbk_data_summary_ordered %>%
  ggplot(aes(id_outbreak, duration_weeks))+
  geom_col(fill="steelblue")+
  scale_x_discrete(breaks = NULL, labels = NULL)+
  # scale_color_manual(name = "Outbreak duration (week)", values="steelblue")+
  coord_flip()+
  labs(title="", y="Outbreak duration (weeks)", x="") +
  theme_minimal()+
  theme_tight()+
  theme(plot.margin = margin(0,0,0,0)) -> plt_b

outbk_data_summary_ordered %>%
  ggplot(aes(id_outbreak, total_s_ch))+
  geom_col(fill="firebrick")+
  scale_x_discrete(breaks = NULL, labels = NULL)+
  # scale_y_continuous(limits=c(0,7000))+
  scale_y_log10(limits=c(1,7000))+
  labs(title="", y="Outbreak size", x = "") +
  coord_flip()+
  theme_minimal()+
  theme_tight()+
  theme(plot.margin = margin(0,0,0,0)) -> plt_c

# Combine all four plots into a single figure
plt_0 <- (plt_a + plt_b + plt_c) +
  plot_layout(ncol = 3, widths = c(1, 0.7, 0.7)) &
  theme(legend.position = "top",
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        plot.caption = element_text(size = 14))

# plt_0

if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig0_individual_ori_impact_size_duration_", tstamp(), ".pdf"),
         plt_0,
         device = cairo_pdf, width = fgw2*1.3,
         height = fgw2, units = "mm")
}

pcr_wk_ordered %>% filter(vacc_week == 1) -> df
df <- left_join(df,
                outbk_data_summary_ordered[, c("total_s_ch", "duration_weeks",
                                               "ar_per_10e3", "id_outbreak")],
                by = "id_outbreak")

# 1. Compute correlation coefficients
# Between med_pct_reduc_case and total_s_ch (original scale)
cor_original <- cor(df$med_pct_reduc_case, df$total_s_ch, use = "complete.obs")

# Between med_pct_reduc_case and log10(total_s_ch)
cor_log10 <- cor(df$med_pct_reduc_case, log10(df$total_s_ch), use = "complete.obs")

# Between med_pct_reduc_case and duration_weeks
cor_duration <- cor(df$med_pct_reduc_case, df$duration_weeks, use = "complete.obs")

# Between med_pct_reduc_case and ar_per_10e3
cor_ar <- cor(df$med_pct_reduc_case, df$ar_per_10e3, use = "complete.obs")

# Print correlation coefficients
cat("Correlation coefficients:\n")
cat("med_pct_reduc_case vs total_s_ch (original):", round(cor_original, 3), "\n")
cat("med_pct_reduc_case vs log10(total_s_ch):", round(cor_log10, 3), "\n")
cat("med_pct_reduc_case vs duration_weeks:", round(cor_duration, 3), "\n")
cat("med_pct_reduc_case vs ar_per_10e3:", round(cor_ar, 3), "\n\n")

# 2. Create scatter plots
# Plot 1: med_pct_reduc_case vs total_s_ch (original scale)
p1 <- ggplot(df, aes(x = total_s_ch, y = med_pct_reduc_case)) +
  geom_point(color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(x = "Outbreak size",
       y = "Median % Reduction in Cases",
       title = paste("Original scale (r =", round(cor_original, 3), ")")) +
  theme_minimal()

# Plot 2: med_pct_reduc_case vs log10(total_s_ch)
p2 <- ggplot(df, aes(x = log10(total_s_ch), y = med_pct_reduc_case)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(x = "Log10(Total Secondary Cases)",
       y = "Median % Reduction in Cases",
       title = paste("Log10 Scale (r =", round(cor_log10, 3), ")")) +
  theme_minimal()

# Plot 3: med_pct_reduc_case vs duration_weeks
p3 <- ggplot(df, aes(x = duration_weeks, y = med_pct_reduc_case)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(x = "Duration (weeks)",
       y = "Median % Reduction in Cases",
       title = paste("Duration vs Reduction (r =", round(cor_duration, 3), ")")) +
  theme_minimal()

# Plot 4: med_pct_reduc_case vs ar_per_10e3
p4 <- ggplot(df, aes(x = ar_per_10e3, y = med_pct_reduc_case)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(x = "Attack Rate per 10,000",
       y = "Median % Reduction in Cases",
       title = paste("Attack Rate vs Reduction (r =", round(cor_ar, 3), ")")) +
  theme_minimal()

# Display plots in a 2x2 grid
grid.arrange(p1, p2, p3, p4, ncol = 2)

# 3. Statistical significance tests for correlations
# Correlation test for original total_s_ch
test_original <- cor.test(df$med_pct_reduc_case, df$total_s_ch)
cat("Correlation test (original scale):\n")
print(test_original)

# Correlation test for log10(total_s_ch)
test_log10 <- cor.test(df$med_pct_reduc_case, log10(df$total_s_ch))
cat("\nCorrelation test (log10 scale):\n")
print(test_log10)

# Correlation test for duration_weeks
test_duration <- cor.test(df$med_pct_reduc_case, df$duration_weeks)
cat("\nCorrelation test (duration):\n")
print(test_duration)


# Correlation test for log10(duration_weeks)
test_log_duration <- cor.test(df$med_pct_reduc_case, log10(df$duration_weeks))
cat("\nCorrelation test (duration):\n")
print(test_log_duration)

# Correlation test for ar_per_10e3
test_ar <- cor.test(df$med_pct_reduc_case, df$ar_per_10e3)
cat("\nCorrelation test (attack rate):\n")
print(test_ar)

```
# # Prepare data by outbreak size group
# d <- cea_s |>
#   group_by(vacc_week, vacc_cov, id_outbreak) |>
#   summarize(s_ch_tot = s_ch_tot[1],
#             pct_reduc_case = median(pct_reduc_case)) # median across simulation runs

# Assign size group to each outbreak
# d$size_grp <- sapply(d$s_ch_tot, function(x) get_group_id(x, size_grp_cutoff))


# # Create nested data by size group
# d <- setDT(d) # calculate_impact_by_subset uses data.table functionalities
# res <- rbindlist(lapply(1:10, function(i) calculate_impact_by_subset(d, i)))


## Table 1: Summary Statistics of the Cholera Outbreaks

```{r table1}
# Create summary statistics for outbreaks
outbk_summary <- outbk_data_summary |>
  filter(id_outbreak %in% outbreaks_filtered$id_outbreak) %>%
  summarize(
    n_outbreaks = n(),
    median_pop = median(pop, na.rm = TRUE),
    q25_pop = quantile(pop, 0.25, na.rm = TRUE),
    q75_pop = quantile(pop, 0.75, na.rm = TRUE),
    median_ar = median(ar_per_10e3, na.rm = TRUE),
    q25_ar = quantile(ar_per_10e3, 0.25, na.rm = TRUE),
    q75_ar = quantile(ar_per_10e3, 0.75, na.rm = TRUE),
    median_duration = median(duration_weeks, na.rm = TRUE),
    q25_duration = quantile(duration_weeks, 0.25, na.rm = TRUE),
    q75_duration = quantile(duration_weeks, 0.75, na.rm = TRUE),
    median_size = median(total_s_ch, na.rm = TRUE),
    q25_size = quantile(total_s_ch, 0.25, na.rm = TRUE),
    q75_size = quantile(total_s_ch, 0.75, na.rm = TRUE)
  )

# Format table for display
table1 <- data.frame(
  Characteristic = c("Number of outbreaks",
                     "Population size",
                     "Attack rate (per 10,000)",
                     "Outbreak duration (weeks)",
                     "Total suspected cases"),
  Value = c(
    as.character(outbk_summary$n_outbreaks),
    paste0(format(outbk_summary$median_pop, big.mark = ",", scientific = FALSE),
           " (", format(outbk_summary$q25_pop, big.mark = ",", scientific = FALSE),
           "-", format(outbk_summary$q75_pop, big.mark = ",", scientific = FALSE), ")"),
    paste0(format(outbk_summary$median_ar, digits = 1),
           " (", format(outbk_summary$q25_ar, digits = 1),
           "-", format(outbk_summary$q75_ar, digits = 1), ")"),
    paste0(format(outbk_summary$median_duration, digits = 1),
           " (", format(outbk_summary$q25_duration, digits = 1),
           "-", format(outbk_summary$q75_duration, digits = 1), ")"),
    paste0(format(outbk_summary$median_size, big.mark = ",", scientific = FALSE),
           " (", format(outbk_summary$q25_size, big.mark = ",", scientific = FALSE),
           "-", format(outbk_summary$q75_size, big.mark = ",", scientific = FALSE), ")")
  )
)

# Display table
knitr::kable(table1, caption = "Summary statistics of cholera outbreaks included in the analysis")

if (save_table) {
  fwrite(table1, paste0("outputs/table1_outbreak_summary_", tstamp(), ".csv"),
         row.names = FALSE)
}
```

## Preemptive Vaccination Strategy

This section models vaccination done before outbreaks begin.

```{r preemptive-sim, eval=FALSE}
tic <- Sys.time()
dat <- outbreak_data

# Settings for preemptive vaccination
vacc_covs <- c(0.6, 0.75, 0.9)
vacc_weeks <- -52  # Negative value ensures pre-emptive vaccination (well before outbreak)
nruns <- 200       # Number of simulation runs

# Create and register parallel cluster
cl <- makeCluster(getOption("cl.cores", detectCores() - 6))
registerDoParallel(cl)

# Run simulations in parallel
svim_week_list <-
  foreach(i = 1:nruns, .packages = c("data.table")) %dopar% {
    run_vacc_impact_outbreak_weekly(
      outbreak_data = dat,
      ive_data = ive_data,
      parameters = params,
      age_dist = age_dist,
      vacc_week = vacc_weeks,    # Preemptive vaccination
      vacc_cov = vacc_covs,
      runid = i)
  }
stopCluster(cl)

# Track execution time
toc <- Sys.time()
message(paste("Execution time:", toc - tic))

# Save results
fwrite(rbindlist(svim_week_list),
       paste0("outputs/svim_week_preemptive_", tstamp(), ".csv"))
```
## Calculate Impact Metrics for Preemptive Vaccination

```{r cea-preemptive}
# Load summarized preemptive vaccination results
d <- fread("outputs/svim_preemptive_sum_over_week_runid_1_200_20250302.csv")

# Process each vaccine coverage level
for (vc in VC3) {
  message(paste("Processing results for vaccine coverage =", vc))

  # Full cost-effectiveness analysis
  cea <- add_cea_variables(d[vacc_cov == vc], mean_age_inf = mean_age_inf)
  cea <- add_cea_results(cea)

  # Save detailed results (as RDS for larger files)
  saveRDS(cea, paste0("outputs/cea_", vc, "_preemptive_", tstamp(), ".rds"))

  # Calculate mean across simulation runs
  m <- mean_across_runid(d[vacc_cov == vc])
  m_cea <- add_cea_variables(m, mean_age_inf = mean_age_inf)
  m_cea <- add_cea_results(m_cea)

  # Save average results
  saveRDS(m_cea,
          paste0("outputs/cea_", vc, "_preemptive_mean_across_runid_",
                 tstamp(), ".rds"))

  # Clean up to save memory
  rm(cea, m, m_cea)
}
```


### Left panel 2
% decline across # cases that accumulated before the outbreak response is triggered 
```{r}
# Calculate proportion of outbreaks included
# outbreaks missed by delayed vaccination
wks_delay <- c(1, 5, 9, 13)
# ct <- c(1, 5, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)
ct_values <- c(1, seq(20, 100, 20))
vax_missed_id <- vector("list", length(wks_delay))

for (w in seq_along(wks_delay)) {
  # Inner list: one element per ct
  vax_missed_id[[w]] <- vector("list", length(ct))
  
  for (i in seq_along(ct_values)) {
    vax_missed_id[[w]][[i]] <- cea_ct %>%
      filter(vacc_week == wks_delay[w] & runid == 1 & vacc_cov == 0.75) %>%
      filter(case_trigger == ct_values[i]) %>%
      filter(!ori_occurred) %>%
      pull(id_outbreak)
  }
}

# Count number of missed outbreaks for each [wks_delay, ct] pair
n_outbreaks_missed <- map(vax_missed_id, function(inner_list) {
  map_int(inner_list, length)   # get length of each vector inside
})

# Total number of outbreaks
n_outbreaks <- nrow(outbk_data_summary)

# Build a data frame in long format
prop_df <- map2_df(
  n_outbreaks_missed,
  wks_delay,
  function(inner_counts, delay_val) {
    tibble(
      wks_delay = delay_val,
      case_trigger = ct,
      prop_included_vax = 100 * (1 - inner_counts / n_outbreaks)
    )
  }
)

p1 <- cea_ct %>% 
  filter(ori_occurred == TRUE) %>%
  filter(vacc_cov == VC, vacc_week %in% wks_delay, case_trigger %in% ct_values) %>%
  ggplot(aes(x = as.factor(case_trigger), 
             y = pct_reduc_case,
             color = as.factor(vacc_week))) +
  # 95% central range (thin line + median point)
  stat_summary(
    fun = median,
    fun.min = function(z) quantile(z, 0.025),
    fun.max = function(z) quantile(z, 0.975),
    geom = "pointrange",
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thick line only, no point)
  stat_summary(
    fun.min = function(z) quantile(z, 0.25),
    fun.max = function(z) quantile(z, 0.75),
    geom = "errorbar",
    position = position_dodge(width = line_interval),
    linewidth = line_width_50,
    width = 0  # vertical line only, no horizontal cap
  ) +
  # Mean marked with bold red X
  stat_summary(
    aes(group = as.factor(vacc_week), color = as.factor(vacc_week)),
    fun = mean,
    geom = "point",
    shape = 4,        # X shape
    size = x_size,
    stroke = x_stroke,
    position = position_dodge(width = line_interval)
  ) +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:7],
    labels = c("1 week", paste0(seq(5,13,4), " weeks"))) +
  labs(x = NULL, y = "% cases averted") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))

# Bottom plot: proportion of outbreaks included
p2 <- prop_df %>% 
  filter(wks_delay %in% wks_delay) %>%
  ggplot(aes(x = as.factor(case_trigger), y = prop_included_vax, group = 
               as.factor(wks_delay), color = as.factor(wks_delay))) +
  geom_line() +
  geom_point() +
  # geom_col(fill="steelblue") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(x = "# cases that triggered the ORI",
       y = "% included") +
  scale_color_manual(
    name = "",
    values = my_discrete_colors[4:7],
    labels = NULL) +
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0))

cea_ct %>%
  filter(ori_occurred) %>% 
  filter(vacc_cov %in% VC, vacc_week %in% wks_delay, case_trigger %in% c(1,100)) %>% 
  group_by(vacc_week, case_trigger) %>% 
  summarize_quantiles(pct_reduc_case)
```

```{r}
# save_plot <- TRUE
# set.seed(42)
# sample_runid <- sample(1:200, 30)
probs <- seq(0, 1, by = 0.1)
unq_data <- filter(cea, vacc_week == 1, vacc_cov == VC, runid == 1)
size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)
ar_grp_cutoff <- quantile(unq_data$ar_per_10e3, probs = probs)
dur_grp_cutoff <- quantile(unq_data$duration_weeks, probs = probs)

d <- left_join(cea_ct, 
               outbk_data_summary[,c("spatial_scale", "rural_urban", 
                                     "duration_weeks", "ar_per_10e3", 
                                     "id_outbreak")], 
               by="id_outbreak")

d %>% group_by(vacc_cov, vacc_week, case_trigger) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d0
d %>% group_by(spatial_scale, vacc_cov, vacc_week, case_trigger) %>%
  summarize_quantiles(variable=pct_reduc_case) -> d1
d %>% group_by(rural_urban, vacc_cov, vacc_week, case_trigger) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d2
d %>% group_by(vacc_cov, vacc_week, case_trigger) %>% 
  filter(ar_per_10e3 >= ar_grp_cutoff[10]) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d3
d %>% group_by(vacc_cov, vacc_week, case_trigger) %>% 
  filter(duration_weeks >= dur_grp_cutoff[10]) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d4
d %>% group_by(vacc_cov, vacc_week, case_trigger) %>% 
  filter(s_ch_tot >= size_grp_cutoff[10]) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d5

d0 <- data.frame(group="all", d0)
names(d1) <- c("group", names(d1)[2:ncol(d1)])
names(d2) <- c("group", names(d1)[2:ncol(d1)])
d3 <- data.frame(group="top_10_ar", d3)
d4 <- data.frame(group="top_10_dur", d4)
d5 <- data.frame(group="top_10_size", d5)

df <- rbindlist(list(d0,d1,d2,d3,d4,d5))

df$group <- factor(
  df$group,
  levels = c("all", "top_10_ar", "top_10_dur", "top_10_size",
             "country", "admin1", "admin2", "admin3",
             "rural", "urban"))
# Set vacc_cov as factor to match discrete color mapping
df$vacc_cov <- factor(df$vacc_cov)
df$vacc_week <- factor(df$vacc_week)
df$case_trigger <- factor(df$case_trigger)

p1 <- df %>% 
  filter(group == "all", vacc_cov %in% VC3, vacc_week %in% c(1,5,9)) %>%
  ggplot(aes(x = case_trigger, color = vacc_week)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = vacc_week, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_color_manual(
    name = NULL,
    values = my_discrete_colors[4:6],  # same as before
    labels = paste("Week", c(1, 5, 9))
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(x = "Number of cases that triggered the ORI", y = "% case averted") +
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0))

# a week delay at c(1,5,9) is recorded at c(2,6,10)
p2 <- df %>% 
  filter(vacc_cov == VC, vacc_week %in% c(9), 
         case_trigger %in% c(1, 10, 50)) %>%
  ggplot(aes(x = group, color = case_trigger)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = case_trigger, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_x_discrete(labels = c("All", 
                              expression(">="*P[90*"%"]*"(AR)"),
                              expression(">="*P[90*"%"]*"(Dur)"),
                              expression(">="*P[90*"%"]*"(Size)"),
                              "Country", "Adm1", "Adm2", "Adm3",
                              "Rural", "Urban")) +
  scale_color_manual(
    name = NULL,
    values = my_discrete_colors[4:6],  # same as before
    labels = paste0("Week ", c(1,5,9))
  ) +
  guides(color = "none") + 
  labs(x = NULL, y = "% case averted") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  theme(plot.margin = margin(0,0,0,0))
```

```{r}
# save_plot <- TRUE
# set.seed(42)
# sample_runid <- sample(1:200, 30)
probs <- seq(0, 1, by = 0.1)
unq_data <- filter(cea, vacc_week == 1, vacc_cov == VC, runid == 1)
size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)
ar_grp_cutoff <- quantile(unq_data$ar_per_10e3, probs = probs)
dur_grp_cutoff <- quantile(unq_data$duration_weeks, probs = probs)

d <- left_join(cea_ct, 
               outbk_data_summary[,c("spatial_scale", "rural_urban", 
                                     "duration_weeks", "ar_per_10e3", 
                                     "id_outbreak")], 
               by="id_outbreak")

d %>% group_by(vacc_cov, vacc_week, case_trigger) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d0
d %>% group_by(spatial_scale, vacc_cov, vacc_week, case_trigger) %>%
  summarize_quantiles(variable=pct_reduc_case) -> d1
d %>% group_by(rural_urban, vacc_cov, vacc_week, case_trigger) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d2
d %>% group_by(vacc_cov, vacc_week, case_trigger) %>% 
  filter(ar_per_10e3 >= ar_grp_cutoff[10]) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d3
d %>% group_by(vacc_cov, vacc_week, case_trigger) %>% 
  filter(duration_weeks >= dur_grp_cutoff[10]) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d4
d %>% group_by(vacc_cov, vacc_week, case_trigger) %>% 
  filter(s_ch_tot >= size_grp_cutoff[10]) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d5

d0 <- data.frame(group="all", d0)
names(d1) <- c("group", names(d1)[2:ncol(d1)])
names(d2) <- c("group", names(d1)[2:ncol(d1)])
d3 <- data.frame(group="top_10_ar", d3)
d4 <- data.frame(group="top_10_dur", d4)
d5 <- data.frame(group="top_10_size", d5)

df <- rbindlist(list(d0,d1,d2,d3,d4,d5))

df$group <- factor(
  df$group,
  levels = c("all", "top_10_ar", "top_10_dur", "top_10_size",
             "country", "admin1", "admin2", "admin3",
             "rural", "urban"))
# Set vacc_cov as factor to match discrete color mapping
df$vacc_cov <- factor(df$vacc_cov)
df$vacc_week <- factor(df$vacc_week)
df$case_trigger <- factor(df$case_trigger)

p1 <- df %>% 
  filter(group == "all", vacc_cov %in% VC3, vacc_week %in% c(1,5,9)) %>%
  ggplot(aes(x = case_trigger, color = vacc_week)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = vacc_week, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_color_manual(
    name = NULL,
    values = my_discrete_colors[4:6],  # same as before
    labels = paste("Week", c(1, 5, 9))
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(x = "Number of cases that triggered the ORI", y = "% case averted") +
  theme_light() +
  theme_tight() +
  theme(plot.margin = margin(0,0,0,0))

# a week delay at c(1,5,9) is recorded at c(2,6,10)
p2 <- df %>% 
  filter(vacc_cov == VC, vacc_week %in% c(9), 
         case_trigger %in% c(1, 10, 50)) %>%
  ggplot(aes(x = group, color = case_trigger)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = case_trigger, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_x_discrete(labels = c("All", 
                              expression(">="*P[90*"%"]*"(AR)"),
                              expression(">="*P[90*"%"]*"(Dur)"),
                              expression(">="*P[90*"%"]*"(Size)"),
                              "Country", "Adm1", "Adm2", "Adm3",
                              "Rural", "Urban")) +
  scale_color_manual(
    name = NULL,
    values = my_discrete_colors[4:6],  # same as before
    labels = paste0("Week ", c(1,5,9))
  ) +
  guides(color = "none") + 
  labs(x = NULL, y = "% case averted") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  theme(plot.margin = margin(0,0,0,0))
```

### Size and duration by admin  
```{r}
d %>% group_by(vacc_cov, vacc_week, spatial_scale, case_trigger) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d0
d %>% group_by(vacc_cov, vacc_week, spatial_scale, case_trigger) %>% 
  filter(ar_per_10e3 >= ar_grp_cutoff[10]) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d1
d %>% group_by(vacc_cov, vacc_week, spatial_scale, case_trigger) %>% 
  filter(duration_weeks >= dur_grp_cutoff[10]) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d2
d %>% group_by(vacc_cov, vacc_week, spatial_scale, case_trigger) %>% 
  filter(s_ch_tot >= size_grp_cutoff[10]) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> d3

d0 <- data.frame(group="all", d0)
d1 <- data.frame(group="top_10_ar", d1)
d2 <- data.frame(group="top_10_dur", d2)
d3 <- data.frame(group="top_10_size", d3)

df <- rbindlist(list(d0,d1,d2,d3))

df$group <- factor(
  df$group,
  levels = c("all", "top_10_ar", "top_10_dur", "top_10_size"))
# Set vacc_cov as factor to match discrete color mapping
df$vacc_cov <- factor(df$vacc_cov)
df$vacc_week <- factor(df$vacc_week)
df$case_trigger <- factor(df$case_trigger)
df$spatial_scale <- factor(df$spatial_scale)

p3 <- df %>% 
  filter(vacc_cov == VC, vacc_week %in% c(9), 
         case_trigger %in% c(1, 10, 50)) %>%
  ggplot(aes(x = group, color = case_trigger)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = case_trigger, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_x_discrete(labels = c("All", 
                              expression(">="*P[90*"%"]*"(AR)"),
                              expression(">="*P[90*"%"]*"(Dur)"),
                              expression(">="*P[90*"%"]*"(Size)"),
                              "Country", "Adm1", "Adm2", "Adm3",
                              "Rural", "Urban")) +
  scale_color_manual(
    name = NULL,
    values = my_discrete_colors[7:9],  # same as before
    labels = paste0("Week ", c(1,5,9))
  ) +
  guides(color = "none") + 
  labs(x = NULL, y = "% case averted") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  theme(plot.margin = margin(0,0,0,0)) +
  facet_wrap(~spatial_scale)
```

#### Previous plots
```{r}
wks_delay <- c(1,5,9)

cea_ct %>%
  filter(ori_occurred) %>% 
  filter(vacc_cov %in% VC, vacc_week %in% wks_delay, case_trigger %in% c(1,100)) %>% 
  group_by(vacc_week, case_trigger) %>% 
  summarize_quantiles(pct_reduc_case)

cea_ct_s <- filter(cea_ct, runid %in% sample_runids)

plt4a <- cea_ct_s %>%
  filter(vacc_cov %in% VC, vacc_week %in% wks_delay, 
         case_trigger %in% c(1, seq(20,100,20))) %>%
  ggplot() +
  geom_boxplot(aes(x = as.factor(case_trigger), y = pct_reduc_case,
                   color = as.factor(vacc_week)),
               width = 0.6, position = position_dodge(width = 0.7),
               outlier.size = 0.6) +
  scale_y_continuous(limits=c(0,100)) +
  scale_color_manual(name = "Delay in ORI implementation (weeks)",
                     values = my_discrete_colors[4:6],
                     labels = as.character(wks)) + 
  labs(x = "Number of cases that triggered the ORI", y = "Cases averted (%)") + 
  theme_light() +
  theme_tight() 

# Impact by outbreak size for case-triggered vaccination

probs <- seq(0, 1, by = 0.1)
d <- cea_ct
d <- d[ori_occurred == TRUE]
size_grp_cutoff <- quantile(d$s_ch_averted_tot, probs = probs)
d$size_grp <- sapply(d$s_ch_tot, function(x) get_group_id(x, size_grp_cutoff))

res <- rbindlist(lapply(1:10, function(i) 
  calculate_impact_by_subset(d, i, subset_var = "size_grp", 
                             grouping_vars = c("vacc_week","vacc_cov","case_trigger"))))

res %>% 
  filter(vacc_week %in% c(5), vacc_cov == 0.75, subset_id %in% c(1, 10), 
         case_trigger %in% c(1,100)) %>%
  group_by(subset_id, case_trigger) %>%
  summarize(mprc = pct_reduc_case_median)

# Create heatmap of impact by outbreak size and case trigger
plt4b <- res |>
  filter(vacc_cov %in% VC, vacc_week %in% c(5)) |>
  ggplot(aes(x = case_trigger, y = subset_id)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case_median)) +
  metR::geom_contour2(aes(z = pct_reduc_case_median, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)", limits = c(0, 60),
                       breaks = seq(0, 60, 20),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Number of cases that triggered the ORI", y = "Total cases percentile") +
  scale_x_continuous()+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                                expression(">="*P[30*"%"]),
                                expression(">="*P[60*"%"]),
                                expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() +
  theme(legend.position = "bottom")


# Combine all four plots into a single figure
plt4 <- (plt4a + plt4b) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              nrow = 1,
              guides = "collect") &
  theme(plot.tag.position = c(0, 0.92),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig4_case_trigger_outbk_size_", tstamp(), ".pdf"),
         plt4,
         device = cairo_pdf, width = fgw2*fac, 
         height = fgw1*1.15*fac, units = "mm")
}
```

```{r}
# save_plot <- TRUE
# set.seed(42)
# sample_runid <- sample(1:200, 30)

wks_delay <- c(1,5,9)

cea_ct %>%
  filter(ori_occurred) %>% 
  filter(vacc_cov %in% VC, vacc_week %in% wks_delay, case_trigger %in% c(1,100)) %>% 
  group_by(vacc_week, case_trigger) %>% 
  summarize_quantiles(pct_reduc_case)

cea_ct_s <- filter(cea_ct, runid %in% sample_runids)

plt4a <- cea_ct_s %>%
  filter(vacc_cov %in% VC, vacc_week %in% wks_delay, 
         case_trigger %in% c(1, seq(20,100,20))) %>%
  ggplot() +
  geom_boxplot(aes(x = as.factor(case_trigger), y = pct_reduc_case,
                   color = as.factor(vacc_week)),
               width = 0.6, position = position_dodge(width = 0.7),
               outlier.size = 0.6) +
  scale_y_continuous(limits=c(0,100)) +
  scale_color_manual(name = "Delay in ORI implementation (weeks)",
                     values = my_discrete_colors[4:6],
                     labels = as.character(wks)) + 
  labs(x = "Number of cases that triggered the ORI", y = "Cases averted (%)") + 
  theme_light() +
  theme_tight() 

# Impact by outbreak size for case-triggered vaccination

probs <- seq(0, 1, by = 0.1)
d <- cea_ct
d <- d[ori_occurred == TRUE]
size_grp_cutoff <- quantile(d$s_ch_averted_tot, probs = probs)
d$size_grp <- sapply(d$s_ch_tot, function(x) get_group_id(x, size_grp_cutoff))

res <- rbindlist(lapply(1:10, function(i) 
  calculate_impact_by_subset(d, i, subset_var = "size_grp", 
                             grouping_vars = c("vacc_week","vacc_cov","case_trigger"))))

res %>% 
  filter(vacc_week %in% c(5), vacc_cov == 0.75, subset_id %in% c(1, 10), 
         case_trigger %in% c(1,100)) %>%
  group_by(subset_id, case_trigger) %>%
  summarize(mprc = pct_reduc_case_median)

# Create heatmap of impact by outbreak size and case trigger
plt4b <- res |>
  filter(vacc_cov %in% VC, vacc_week %in% c(5)) |>
  ggplot(aes(x = case_trigger, y = subset_id)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case_median)) +
  metR::geom_contour2(aes(z = pct_reduc_case_median, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)", limits = c(0, 60),
                       breaks = seq(0, 60, 20),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Number of cases that triggered the ORI", y = "Total cases percentile") +
  scale_x_continuous()+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                                expression(">="*P[30*"%"]),
                                expression(">="*P[60*"%"]),
                                expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() +
  theme(legend.position = "bottom")


# Combine all four plots into a single figure
plt4 <- (plt4a + plt4b) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              nrow = 1,
              guides = "collect") &
  theme(plot.tag.position = c(0, 0.92),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig4_case_trigger_outbk_size_", tstamp(), ".pdf"),
         plt4,
         device = cairo_pdf, width = fgw2*fac, 
         height = fgw1*1.15*fac, units = "mm")
}
```


## Create Outbreak Subsets

```{r create-subsets}
# Prepare outbreak summary data
# ---- 1. Random sampling of outbreaks by year ----
# This simulates targeting outbreaks randomly, matched to historical OCV deployment
DF <- data.frame()

# ---- 2. Case-weighted sampling of outbreaks by year ----
# This simulates targeting outbreaks with more cases, matched to historical OCV deployment
names(dat) <- c("id_outbreak", "year", "rural_urban", "ar", "cfr", "spatial_scale",
                "pop", "duration", "cases")

set.seed(42)
DF <- data.frame()
for (j in 1:200) {
  set.seed(j)
  message(paste("Simulation run", j))
  df <- data.frame()
  for (i in 2013:2022) {
    yr <- i  # 2013 = the first year when OCVs were deployed
    # Sample outbreaks weighted by case count
    smpl <- sample_outbreaks(
      outbreak_data = filter(dat, year == yr), 
      target_population = ocv_year[year == yr, num_ocv],
      sampling_method = "cases")
    
    df <- rbind(df, data.frame(year = smpl$sampled_data$year, 
                         id_outbreak = smpl$sampled_data$id_outbreak))
  }
  df$subset_id <- j
  DF <- rbind(DF, df)
}

# Save case-weighted sampling results
fwrite(DF, paste0("outputs/case_weighted_subsets_by_year_", tstamp(), ".csv"))
```

### Statistical tests for the difference between top and bottom quartiles in terms of % cases averted and % weekly decline in cases averted
```{r}
set.seed(123)

# df from the Right panel code block for Figure 3
outbreak_df <- df
names(outbreak_df) <- c("id_outbreak", "rate_reduc", "half_life", 
               "attack_rate", "duration_weeks", "spatial_scale", 
               "rural_urban", "pop", "outbreak_size", "pcr", "row_id") 

# ---- 2. Classify into quartiles ----
outbreak_df <- outbreak_df %>% mutate(
  impact_group = case_when(
    row_id %in% top_pcr_ids ~ "Top",
    row_id %in% bottom_pcr_ids  ~ "Bottom",
    TRUE ~ "Middle"), 
  decline_group = case_when(
    row_id %in% top_rate_ids ~ "Top",
    row_id %in% bottom_rate_ids ~ "Bottom",
    TRUE ~ "Middle")
)

# ---- 3. Log-transform skewed variables ----
outbreak_df <- outbreak_df %>%
  mutate(
    log_attack_rate = log1p(attack_rate),
    log_outbreak_size = log1p(outbreak_size),
    log_pop = log1p(pop)
  )

# ---- 4. Wilcoxon tests ----
vars_to_test <- c("duration_weeks", "log_attack_rate", 
                  "log_outbreak_size", "log_pop", 
                  "spatial_scale", 
                  "rural_urban")

test_results <- lapply(vars_to_test, function(var) {
  wilcox_test <- wilcox.test(
    formula(paste(var, "~ impact_group")),
    data = outbreak_df %>% filter(impact_group %in% c("Top", "Bottom"))
  )
  t_stat <- tidy(wilcox_test)
  t_stat$variable <- var
  t_stat
}) %>% bind_rows()

# plot (example of log_attack_rate)
ggplot(outbreak_df %>% filter(impact_group %in% c("Top", "Bottom")),
       aes(x = impact_group, y = log_attack_rate)) +
  geom_boxplot()

# ---- 5. Effect sizes ----
effect_sizes <- lapply(vars_to_test, function(var) {
  eff <- wilcox_effsize(
    formula(paste(var, "~ impact_group")),
    data = outbreak_df %>% filter(impact_group %in% c("Top", "Bottom"))
  )
  eff$variable <- var
  eff
}) %>% bind_rows()

# ---- 6. Logistic regression ----
# log_model <- glm(
#   impact_group ~ log_attack_rate + log_outbreak_size + log_pop + duration_weeks + rural_urban,
#   data = outbreak_df %>% filter(impact_group %in% c("Top", "Bottom")),
#   family = binomial
# )

# Filter and ensure binary factor
df_log <- outbreak_df %>%
  filter(impact_group %in% c("Top", "Bottom")) %>%
  mutate(impact_group = factor(impact_group, levels = c("Bottom", "Top")))

# Logistic regression
log_model <- glm(
  impact_group ~ log_attack_rate + log_outbreak_size + log_pop + duration_weeks + rural_urban + spatial_scale,
  data = df_log,
  family = binomial
)
summary(log_model)

library(car)
vif(log_model)

# ---- 7. Random forest variable importance ----
rf_model <- randomForest(
  factor(impact_group) ~ log_attack_rate + log_outbreak_size + log_pop + duration_weeks + rural_urban + spatial_scale,
  data = outbreak_df %>% filter(impact_group %in% c("Top", "Bottom")),
  ntree = 500,
  importance = TRUE
)

importance(rf_model)
varImpPlot(rf_model, type = 1)  # type = 1 for MeanDecreaseAccuracy
varImpPlot(rf_model, type = 2)  # type = 2 for MeanDecreaseGini
# ---- 8. Spearman correlations ----
cor_results <- outbreak_df %>%
  select(log_attack_rate, log_outbreak_size, log_pop, duration_weeks, pcr, rate_reduc) %>%
  cor(method = "spearman", use = "pairwise.complete.obs")

# ---- 9. Outputs ----
list(
  test_results = test_results,
  effect_sizes = effect_sizes,
  logistic_regression = summary(log_model),
  rf_importance = importance(rf_model),
  correlations = cor_results
)

```

```{r}
# --- Packages ---
# library(dplyr)
# library(ggplot2)
library(broom)
library(rstatix)        # wilcox_effsize
library(randomForest)
library(car)            # vif
# library(purrr)
# library(tidyr)
library(glue)

set.seed(123)

# --- 0) Prepare data (your object 'df' from the right panel of Fig. 3) ---
outbreak_df <- df_original_scale
outbreak_df <- outbreak_df %>% mutate(row_id = row_number())

names(outbreak_df) <- c("id_outbreak", "rate_reduc", "half_life", 
                        "attack_rate", "duration_weeks", "spatial_scale", 
                        "rural_urban", "pop", "outbreak_size", "pcr", "row_id")

# If the Top/Bottom ID vectors are not defined yet, create them from quartiles
if (!exists("top_pcr_ids") || !exists("bottom_pcr_ids")) {
  qs_pcr <- quantile(outbreak_df$pcr, c(.25, .75), na.rm = TRUE)
  bottom_pcr_ids <- outbreak_df$row_id[outbreak_df$pcr <= qs_pcr[1]]
  top_pcr_ids    <- outbreak_df$row_id[outbreak_df$pcr >= qs_pcr[2]]
}
if (!exists("top_rate_ids") || !exists("bottom_rate_ids")) {
  qs_rate <- quantile(outbreak_df$rate_reduc, c(.25, .75), na.rm = TRUE)
  bottom_rate_ids <- outbreak_df$row_id[outbreak_df$rate_reduc <= qs_rate[1]]
  top_rate_ids    <- outbreak_df$row_id[outbreak_df$rate_reduc >= qs_rate[2]]
}

# --- 1) Define groups (impact: % cases averted; decline: % weekly decline) ---
outbreak_df <- outbreak_df %>%
  mutate(
    impact_group = case_when(
      row_id %in% top_pcr_ids ~ "Top",
      row_id %in% bottom_pcr_ids ~ "Bottom",
      TRUE ~ "Middle"
    ),
    decline_group = case_when(
      row_id %in% top_rate_ids ~ "Top",
      row_id %in% bottom_rate_ids ~ "Bottom",
      TRUE ~ "Middle"
    )
  )

# --- 2) Transform skewed vars; coerce categorical types ---
outbreak_df <- outbreak_df %>%
  mutate(
    log_attack_rate  = log1p(attack_rate),
    log_outbreak_size = log1p(outbreak_size),
    log_pop = log1p(pop),
    spatial_scale = as.factor(spatial_scale),
    rural_urban   = as.factor(rural_urban)
  )

# ------------------------------ Helper functions ------------------------------

# (A) Univariate Wilcoxon + effect size for continuous/ordinal variables
run_wilcoxon_block <- function(dat, group_var, vars_cont) {
  dat2 <- dat %>% filter(.data[[group_var]] %in% c("Top","Bottom")) %>%
    mutate(group = factor(.data[[group_var]], levels = c("Bottom","Top")))
  
  # Wilcoxon w/ FDR
  wcx <- map_dfr(vars_cont, function(v) {
    wt <- wilcox.test(reformulate("group", v), data = dat2, exact = FALSE)
    tidy(wt) %>% mutate(variable = v)
  }) %>%
    group_by(variable) %>%
    mutate(p_adj = p.adjust(p.value, method = "fdr")) %>%
    ungroup()
  
  # Effect sizes (r from rstatix)
  eff <- map_dfr(vars_cont, function(v) {
    es <- wilcox_effsize(reformulate("group", v), data = dat2)
    es$variable <- v
    es
  })
  
  # Group summaries (median, IQR) for reporting
  summ <- dat2 %>%
    select(group, all_of(vars_cont)) %>%
    pivot_longer(cols = -group, names_to = "variable", values_to = "value") %>%
    group_by(variable, group) %>%
    summarise(
      n = sum(!is.na(value)),
      median = median(value, na.rm = TRUE),
      q25 = quantile(value, .25, na.rm = TRUE),
      q75 = quantile(value, .75, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    pivot_wider(
      id_cols = variable,
      names_from = group,
      values_from = c(n, median, q25, q75),
      names_sep = "."
    )
  
  out <- wcx %>%
    left_join(eff %>% select(variable, effect.size = effsize, magnitude), by = "variable") %>%
    left_join(summ, by = "variable") %>%
    select(variable, statistic, p.value, p_adj, effect.size, magnitude,
           starts_with("n_"), starts_with("median"), starts_with("q25"), starts_with("q75"))
  out
}

# (B) Fisher's exact test for binary categorical variables
run_fisher_block <- function(dat, group_var, vars_cat) {
  dat2 <- dat %>% filter(.data[[group_var]] %in% c("Top","Bottom")) %>%
    mutate(group = factor(.data[[group_var]], levels = c("Bottom","Top")))
  
  map_dfr(vars_cat, function(v) {
    tab <- table(dat2$group, dat2[[v]])
    ft  <- stats::fisher.test(tab)
    broom::tidy(ft) %>%
      mutate(variable = v,
             # add simple proportion by group for context
             prop_top    = mean(dat2[[v]] == levels(dat2[[v]])[2] & dat2$group=="Top",  na.rm = TRUE),
             prop_bottom = mean(dat2[[v]] == levels(dat2[[v]])[2] & dat2$group=="Bottom",na.rm = TRUE))
  }) %>%
    group_by(variable) %>%
    mutate(p_adj = p.adjust(p.value, method = "fdr")) %>%
    ungroup()
}

# (C) Multivariable logistic regression + VIF and OR table
run_logistic <- function(dat, group_var, predictors) {
  dat2 <- dat %>% filter(.data[[group_var]] %in% c("Top","Bottom")) %>%
    mutate(group = factor(.data[[group_var]], levels = c("Bottom","Top")))
  
  f <- reformulate(predictors, "group")
  fit <- glm(f, data = dat2, family = binomial)
  vif_tbl <- tryCatch({
    tibble(term = names(vif(fit)), vif = as.numeric(vif(fit)))
  }, error = function(e) tibble(term = NA_character_, vif = NA_real_))
  
  or_tbl <- broom::tidy(fit, conf.int = TRUE, exponentiate = TRUE) %>%
    mutate(across(c(estimate, conf.low, conf.high), ~round(.x, 3))) %>%
    rename(OR = estimate, CI_low = conf.low, CI_high = conf.high)
  
  list(model = fit, or_table = or_tbl, vif = vif_tbl)
}

# (D) Random forest importance (for robustness / nonlinearity)
run_rf <- function(dat, group_var, predictors, ntree = 500) {
  dat2 <- dat %>% filter(.data[[group_var]] %in% c("Top","Bottom")) %>%
    mutate(group = factor(.data[[group_var]]))
  # randomForest requires numeric features; convert factors via model.matrix
  mm <- model.matrix(~ . - 1, data = dat2[, predictors])
  rf <- randomForest(x = mm, y = dat2$group, ntree = ntree, importance = TRUE)
  imp <- importance(rf)
  list(model = rf, importance = imp)
}

# (E) Spearman correlations among continuous variables (context)
run_spearman <- function(dat, vars) {
  dat %>%
    select(all_of(vars)) %>%
    cor(method = "spearman", use = "pairwise.complete.obs")
}

# Master wrapper to run an entire block for a given grouping
run_block <- function(dat, group_var) {
  cont_vars <- c("duration_weeks", "log_attack_rate", "log_outbreak_size", "log_pop")
  cat_vars  <- c("rural_urban", "spatial_scale")  # treated as factors
  
  # Univariate
  wilcox_out <- run_wilcoxon_block(dat, group_var, cont_vars)
  fisher_out <- run_fisher_block(dat, group_var, cat_vars)
  
  # Multivariable (exclude the defining metric itself)
  predictors <- c("log_attack_rate", "log_outbreak_size", "log_pop",
                  "duration_weeks", "rural_urban", "spatial_scale")
  logit_out  <- run_logistic(dat, group_var, predictors)
  rf_out     <- run_rf(dat, group_var, predictors, ntree = 500)
  
  # Correlations (context)
  cor_out <- run_spearman(dat, vars = c("log_attack_rate", "log_outbreak_size",
                                        "log_pop", "duration_weeks", "pcr", "rate_reduc"))
  
  list(
    wilcoxon = wilcox_out,
    fisher   = fisher_out,
    logistic_or = logit_out$or_table,
    logistic_vif = logit_out$vif,
    rf_importance = rf_out$importance,
    correlations  = cor_out
  )
}

# --- 3) RUN: (A) impact contrast, (B) rate-of-decline contrast ---
impact_results  <- run_block(outbreak_df, group_var = "impact_group")
decline_results <- run_block(outbreak_df, group_var = "decline_group")

# --- 4) Example: boxplot for attack rate by decline_group (Top vs Bottom) ---
ggplot(outbreak_df %>% filter(decline_group %in% c("Top","Bottom")),
       aes(x = decline_group, y = log_attack_rate)) +
  geom_boxplot(width = 0.6) +
  labs(x = "Decline group (Top vs Bottom quartile of weekly % decline)",
       y = "log(1 + attack rate)")

# --- 5) Compact tables for Supplement ---


supp_table_decline <- decline_results$wilcoxon %>%
  transmute(
    Variable = variable,
    `Median (IQR) - Bottom` = paste0(round(median.Bottom, 2), " (", round(q25.Bottom, 2), ", ", round(q75.Bottom, 2), ")"),
    `Median (IQR) - Top`    = paste0(round(median.Top, 2), " (", round(q25.Top, 2), ", ", round(q75.Top, 2), ")"),
    `W` = round(statistic, 1),
    `P (FDR)` = signif(p_adj, 3),
    `Effect size r` = round(effect.size, 3),
    `Magnitude` = magnitude
  )

supp_table_decline_cat <- decline_results$fisher %>%
  transmute(
    Variable = variable,
    `P (FDR)` = signif(p_adj, 3)
  )

# Print to console when knitting
supp_table_decline
supp_table_decline_cat
decline_results$logistic_or
decline_results$logistic_vif

```

###### Previous plots

```{r}
# (C) Impact by outbreak size for outbreak onset-triggered vaccination
# Calculate percentiles for outbreak size groups
probs <- seq(0, 1, by = 0.1)
runids <- unique(cea_s$runid)
unq_data <- filter(cea_s, runid == runids[1] & vacc_cov == 0.75 & vacc_week == 1)
size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)
cea_s$size_grp <- sapply(cea_s$s_ch_tot, function(x) get_group_id(x, size_grp_cutoff))

res <- rbindlist(lapply(1:10, function(i) calculate_impact_by_subset(cea_s, i)))

# Create heatmap of impact by outbreak size and vaccination week
plt3c <- res |>
  filter(vacc_cov %in% VC, vacc_week %in% VW_seq) |>
  ggplot(aes(x = vacc_week, y = subset_id)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case_median)) +
  metR::geom_contour2(aes(z = pct_reduc_case_median, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)  ", limits = c(0, 100),
                       breaks = seq(0, 100, 50),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Delay in ORI implementation (weeks)", 
       y = "Total case percentile") +
  scale_x_continuous(breaks = seq(1, 15, 2))+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() +
  theme(legend.position = "bottom")

ar_grp_cutoff <- quantile(outbk_data_summary$ar_per_10e3, probs = probs)
dur_grp_cutoff <- quantile(outbk_data_summary$duration_weeks, probs = probs)

cea_s %<>%
  group_by(runid, id_outbreak, vacc_cov) %>%
  mutate(pct_reduc_case_frac = c(1, 1-frac_diff(pct_reduc_case, lag=1)),
         pct_reduc_case_diff = c(1, -diff(pct_reduc_case, lag=1)))

cea_s %>% filter(vacc_cov == 0.75) -> d1
d1$grp <- "All"
cea_s %>% filter(ar_per_10e3 > ar_grp_cutoff[10]) -> d2 # top 10%
d2$grp <- "AR"
cea_s %>% filter(duration_weeks > dur_grp_cutoff[10]) -> d3 
d3$grp <- "Dur"
cea_s %>% filter(s_ch_tot > size_grp_cutoff[10]) -> d4 
d4$grp <- "Size"
d <- rbind(d1,d2)
d_ <- rbind(d3,d4)
d <- rbind(d, d_)

plt3d <- d %>% 
  filter(vacc_cov == VC, vacc_week > 1, vacc_week <= 15) %>%
  ggplot(aes(x = as.factor(grp), y = 100*pct_reduc_case_frac)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_x_discrete(labels = c(expression("All"),
                              expression(">="*P[90*"%"]*"(AR)"),
                              expression(">="*P[90*"%"]*"(Dur)"),
                              expression(">="*P[90*"%"]*"(Size)"))) +
  labs(x = "", y = "Weekly reduction in % case averted") +
  theme_light() +
  theme_tight()

# Combine all four plots into a single figure
plt3 <- (plt3a + plt3b + plt3c + plt3d) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              guides = "collect", 
              heights = c(1, 1)) &
  theme(plot.tag.position = c(0.05, 1),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Adjust tag positions for top panels
for (i in c(1, 2)) {
  plt3[[i]] <- plt3[[i]] + theme(plot.tag.position = c(0.05, 0.9))
}

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig3_onset_triggered_", tstamp(), ".pdf"),
         plt3,
         device = cairo_pdf, width = fgw2*fac, 
         height = fgw2*1.15*fac, units = "mm")
}
```

###### Across attack rate, duration, and outbreak size  

```{r}
d <- left_join(cea, 
               outbk_data_summary[,c("spatial_scale", "rural_urban",
                                            "id_outbreak")], 
               by="id_outbreak")

probs <- seq(0, 1, by = 0.1)
unq_data <- filter(d, vacc_week == 1, vacc_cov == VC, runid == 1)
size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)
ar_grp_cutoff <- quantile(unq_data$ar_per_10e3, probs = probs)
dur_grp_cutoff <- quantile(unq_data$duration_weeks, probs = probs)

d$ar_cat <- cut(d$ar_per_10e3, breaks = c(-Inf, ar_grp_cutoff[2:11]), 
                  labels=1:(length(ar_grp_cutoff)-1))
d$dur_cat <- cut(d$duration_weeks, breaks = c(-Inf, dur_grp_cutoff[2:11]), 
                  labels=1:(length(dur_grp_cutoff)-1))
d$size_cat <- cut(d$s_ch_tot, breaks = c(-Inf,size_grp_cutoff[2:11]), 
                  labels=1:(length(size_grp_cutoff)-1))


d %>% group_by(vacc_cov, vacc_week, size_cat) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> df

# Set vacc_cov as factor to match discrete color mapping
df$vacc_cov <- factor(df$vacc_cov)
df$vacc_week <- factor(df$vacc_week)
df$size_cat <- factor(df$size_cat)

df %>% 
  filter(vacc_cov == VC, vacc_week %in% c(1,5,9)) %>%
  ggplot(aes(x = size_cat, color = vacc_week)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = vacc_week, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_color_manual(
    name = NULL,
    values = my_discrete_colors[4:6],  # same as before
    labels = paste("Week", c(1, 5, 9))
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(x = NULL, y = "% case averted") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))


d %>% group_by(vacc_cov, vacc_week, dur_cat) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> df

# Set vacc_cov as factor to match discrete color mapping
df$vacc_cov <- factor(df$vacc_cov)
df$vacc_week <- factor(df$vacc_week)
df$dur_cat <- factor(df$dur_cat)

df %>%
  filter(vacc_cov == VC, vacc_week %in% c(1,5,9)) %>%
  ggplot(aes(x = dur_cat, color = vacc_week)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = vacc_week, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_color_manual(
    name = NULL,
    values = my_discrete_colors[4:6],  # same as before
    labels = paste("Week", c(1, 5, 9))
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(x = NULL, y = "% case averted") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))

d %>% group_by(vacc_cov, vacc_week, ar_cat) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> df
# Set vacc_cov as factor to match discrete color mapping
df$vacc_cov <- factor(df$vacc_cov)
df$vacc_week <- factor(df$vacc_week)
df$ar_cat <- factor(df$ar_cat)

df %>%
  filter(vacc_cov == VC, vacc_week %in% c(1,5,9)) %>%
  ggplot(aes(x = ar_cat, color = vacc_week)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = vacc_week, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_color_manual(
    name = NULL,
    values = my_discrete_colors[4:6],  # same as before
    labels = paste("Week", c(1, 5, 9))
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(x = NULL, y = "% case averted") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))
```


```{r fig4, eval=FALSE}
# Filter subset analysis results for easier visualization
# cea_subsets |> filter(vacc_week == 9, vacc_cov == 0.75) -> dat
# cea_subsets_case_wt |> filter(vacc_week == 9, vacc_cov == 0.75) -> dat_wt

# Reference values for individual assessment
cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(s_ch_averted_tot) 

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(case_averted_per_1000_OCV)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(cost_per_daly_averted)

cea |>
  filter(vacc_cov == 0.75, vacc_week == 9) |>
  group_by(vacc_week, vacc_cov) |>
  summarize_quantiles(cost_per_case_averted)


# Combine random and case-weighted results
dat$sampling <- "Random"
dat_wt$sampling <- "Case-weighted"
d <- rbind(dat, dat_wt)

d <- left_join(subsets, cea, by="id_outbreak")
# (A) OCV deployment over time
plt_a1 <- ocv_year |>
  filter(year <= 2022) |> 
  ggplot(aes(x = as.factor(year), y = cum_num_ocv)) +
  geom_point() +
  labs(x = "", y = "Cumulative\nOCVs deployed") +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (B) Cases averted
plt_b1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = s_ch_averted_tot + 1, color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  labs(x = "", y = "Cumulative\ncases averted") +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (C) Percent case reduction
plt_c1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = pct_reduc_case, color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_prc), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_color_manual(name = "",
                     values = c("firebrick", "steelblue"),
                     labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "% reduction\nin cases") +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (D) Cases averted per 1000 OCVs
plt_d1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = case_averted_per_1000_OCV, 
             color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_capd), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  # scale_y_continuous(limits = c(0, 2)) +
  scale_color_manual(name = "",
                   values = c("firebrick", "steelblue"),
                   labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "Cases averted\nper 1,000 OCVs") +
  theme_light() +
  theme_tight() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# (E) Cost per case averted
plt_e1 <- d |>
  ggplot(aes(x = as.factor(year_max), y = cost_per_case_averted, color = sampling)) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.7), 
               outlier.size = 0.6) +
  geom_hline(data = ref_val, aes(yintercept = median_cpca), 
             color = "firebrick", linetype = "dotted", linewidth = 1) +
  scale_y_log10(limits = c(1, 1e5)) +
  scale_color_manual(name = "",
                   values = c("firebrick", "steelblue"),
                   labels = c("Case-weighted", "Random")) +
  labs(x = "", y = "Cost per case\naverted (US$)") +
  theme_light() +
  theme_tight()

# Combine all panels into a single figure
plt6 <- wrap_plots(list(plt_a1, plt_b1, plt_c1, plt_d1, plt_e1), ncol = 1,
                  guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0.13, 0.92),
        plot.tag = element_text(size = 12),
        legend.position = "top")

# Save figure
if (save_plot) {
  ggsave(paste0("plots/fig4_cumulative_impact_", tstamp(), ".pdf"),
         plt6, device = cairo_pdf, width = 70*2.4, height = 70*2.4, units = "mm")
}

d %>% 
  filter(year_max %in% c(2013, 2022)) %>%
  group_by(year_max, sampling) %>% 
  summarize_quantiles(variable=s_ch_averted_tot) -> ca1

format_median_iqr(ca1, row=2)
format_median_iqr(ca1, row=4)

d %>% 
  filter(year_max %in% c(2013, 2022)) %>%
  group_by(year_max, sampling) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> pcr1

format_median_iqr(pcr1, row=2)
format_median_iqr(pcr1, row=4)

d %>% 
  filter(year_max %in% c(2013, 2022)) %>%
  group_by(year_max, sampling) %>% 
  summarize_quantiles(variable=case_averted_per_1000_OCV) -> capd1

format_median_iqr(capd1, row=2)
format_median_iqr(capd1, row=4)

d %>% group_by(year_max, sampling) %>% 
  summarize(case_averted_per_1000_OCV = median(case_averted_per_1000_OCV))
  
```

###### Across population size

```{r}
d <- left_join(cea, 
               outbk_data_summary[,c("spatial_scale", "rural_urban",
                                            "id_outbreak")], 
               by="id_outbreak")

d$pop_cat <- cut(d$pop, breaks = 10)

breaks <- c(0, 1e5, 3e5, 1e6, Inf)     # cut-points
labels <- c("Small (<100k)",
            "Medium (100–300k)",
            "Large (300k–1M)",
            "Very large (>1M)")

d <- d %>% 
  mutate(pop_cat = cut(pop, breaks = breaks, labels = labels, right = FALSE))


d %>% group_by(vacc_cov, vacc_week, pop_cat) %>% 
  summarize_quantiles(variable=pct_reduc_case) -> df

# Set vacc_cov as factor to match discrete color mapping
df$vacc_cov <- factor(df$vacc_cov)
df$vacc_week <- factor(df$vacc_week)
df$pop_cat <- factor(df$pop_cat)

df %>% 
  filter(vacc_cov == VC, vacc_week %in% c(1,5,9)) %>%
  ggplot(aes(x = pop_cat, color = vacc_week)) +
  # 95% central range (thin line + median point)
  geom_pointrange(
    aes(y = q500, ymin = q025, ymax = q975),
    position = position_dodge(width = line_interval),
    linewidth = line_width_95,
    size = point_size
  ) +
  # 50% central range (thicker line only)
  geom_errorbar(
    aes(ymin = q250, ymax = q750),
    linewidth = line_width_50,
    width = 0,
    position = position_dodge(width = line_interval)
  ) +
  # Mean marked with red X
  geom_point(
    aes(group = vacc_week, y = mean),
    shape = 4,  # X shape
    size = x_size,
    stroke = x_stroke,
    # color = "red",
    position = position_dodge(width = line_interval)
  ) +
  scale_color_manual(
    name = NULL,
    values = my_discrete_colors[4:6],  # same as before
    labels = paste("Week", c(1, 5, 9))
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(x = NULL, y = "% case averted") +
  theme_light() +
  theme_tight() +
  theme_x_blank() +
  theme(plot.margin = margin(0,0,0,0))
```


## Figure S2: Case reduction across attack rate and duration

```{r fig3, eval=FALSE}
# Calculate percentiles for outbreak attack rate groups
probs <- seq(0, 1, by = 0.1)
runids <- unique(cea_s$runid)
unq_data <- filter(cea_s, runid == runids[1] & vacc_cov == 0.75 & vacc_week == 1)
# size_grp_cutoff <- quantile(unq_data$s_ch_tot, probs = probs)
ar_grp_cutoff <- quantile(unq_data$ar_per_10e3, probs = probs)
cea_s$ar_grp <- sapply(cea_s$ar_per_10e3, function(x) get_group_id(x, ar_grp_cutoff))

res <- rbindlist(lapply(1:10, function(i)
  calculate_impact_by_subset(cea_s, i, subset_var = "ar_grp")))

plt3sa <- res |>
  filter(vacc_cov %in% VC, vacc_week %in% VW_seq) |>
  ggplot(aes(x = vacc_week, y = subset_id)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case_median)) +
  metR::geom_contour2(aes(z = pct_reduc_case_median, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)  ", limits = c(0, 100),
                       breaks = seq(0, 100, 50),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Delay in ORI implementation (weeks)", 
       y = "Attack rate percentile") +
  scale_x_continuous(breaks = seq(1, 15, 2))+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() +
  theme(legend.position = "bottom")

# duration group 
dur_grp_cutoff <- quantile(unq_data$duration_weeks, probs = probs)
cea_s$dur_grp <- 
  sapply(cea_s$duration_weeks, function(x) get_group_id(x, dur_grp_cutoff))

res <- rbindlist(lapply(1:10, function(i) 
  calculate_impact_by_subset(cea_s, i, subset_var = "dur_grp")))

plt3sb <- res |>
  filter(vacc_cov %in% VC, vacc_week %in% VW_seq) |>
  ggplot(aes(x = vacc_week, y = subset_id)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case_median)) +
  metR::geom_contour2(aes(z = pct_reduc_case_median, label = stat(level)),
                      linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)  ", limits=c(0, 100),
                       breaks = seq(0, 100, 50),
                       labels = scales::percent_format(scale=1)) +
  labs(x = "Delay in ORI implementation (weeks)", 
       y = "Duration percentile") +
  scale_x_continuous(breaks = seq(1, 15, 2))+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() +
  theme(legend.position = "bottom")

# Combine two plots into a single figure
plt3s <- (plt3sa + plt3sb) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              nrow = 1,
              guides = "collect") &
  theme(plot.tag.position = c(0.05, 1),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Adjust tag positions for top panels
for (i in c(1, 2)) {
  plt3s[[i]] <- plt3s[[i]] + theme(plot.tag.position = c(0.05, 0.9))
}

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/figS2_onset_triggered_AR_Dur_", tstamp(), ".pdf"),
         plt3s,
         device = cairo_pdf, width = fgw2*fac, 
         height = fgw1*1.15*fac, units = "mm")
}
```


## Figure S4: Case Reduction in Case-Triggered Scenario across attack rate and duration

```{r}
save_plot <- TRUE
set.seed(42)
sample_runid <- sample(1:200, 30)
cea_ct_s <- filter(cea_ct, runid %in% sample_runid)

wks <- c(1,5,9)
# 
# cea_ct_s %>%
# cea_ct_s <- left_join(cea_ct_s,
#                  outbk_data_summary[,c("id_outbreak",
#                                        "ar_per_10e3","duration_weeks")], 
#                  by = "id_outbreak")
cea_ct %>%  
  filter(vacc_cov %in% VC, vacc_week %in% wks, case_trigger %in% c(1,100)) %>% 
  group_by(vacc_week, case_trigger) %>% 
  summarize_quantiles(pct_reduc_case)

cea_ct <- 
  left_join(cea_ct,
    outbk_data_summary[,c("id_outbreak", "ar_per_10e3","duration_weeks")], 
     by = "id_outbreak")

# Impact by duration size for outbreak onset-triggered vaccination
# Calculate percentiles for attack rate groups
probs <- seq(0, 1, by = 0.1)
runids <- unique(cea_ct_s$runid)
unq_data <- filter(cea_ct_s, runid == runids[1] & vacc_cov == VC & vacc_week == 1)
ar_grp_cutoff <- quantile(outbk_data_summary$ar_per_10e3, probs = probs)
dur_grp_cutoff <- quantile(outbk_data_summary$duration_weeks, probs = probs)

# Prepare data by outbreak size group
d <- cea_ct_s |> 
  group_by(vacc_week, case_trigger, vacc_cov, id_outbreak) |> 
  summarize(ar_per_10e3 = ar_per_10e3[1],
            duration_weeks = duration_weeks[1],
            med_pct_reduc_case = median(pct_reduc_case)) # median across simulation runs

# Assign size group to each outbreak
d$ar_grp <- sapply(d$ar_per_10e3, function(x) get_group_id(x, ar_grp_cutoff))
d$dur_grp <- sapply(d$duration_weeks, function(x) get_group_id(x, dur_grp_cutoff))


# Create nested data by size group
lst <- vector("list", length(ar_grp_cutoff) - 1)
d <- setDT(d)
for(i in 1:length(lst)){
  lst[[i]] <- d[ar_grp >= i]
}

d %>%
  filter(vacc_cov %in% VC, vacc_week %in% c(5), case_trigger %in% c(1,100), 
         ar_grp %in% c(1,10)) %>% 
  group_by(vacc_week, case_trigger) %>% 
  summarize_quantiles(med_pct_reduc_case)

# Calculate median impact by size group
res_list <- vector('list', length(ar_grp_cutoff) - 1)
for (i in 1:length(lst)) {
  el <- lst[[i]]
  res_list[[i]] <- el[, .(
    ar_grp = min(ar_grp),
    med_pct_reduc_case = median(med_pct_reduc_case, na.rm = TRUE)),
    by = .(vacc_week, case_trigger, vacc_cov)]
}
res_rbind <- rbindlist(res_list)

res_rbind %>%
  group_by(vacc_week, vacc_cov, ar_grp) %>%
  mutate(reference = med_pct_reduc_case[case_trigger == 1]) %>%
  mutate(rel_med_pct_reduc_case = med_pct_reduc_case/reference) -> tmp

tmp %>% 
  filter(vacc_week %in% c(5), vacc_cov ==0.75, ar_grp %in% c(1, 10), 
         case_trigger %in% c(1,100)) %>%
  group_by(ar_grp, case_trigger) %>%
  summarize(mprc = med_pct_reduc_case)



# 
# probs <- seq(0, 1, by = 0.1)
d <- cea_ct
d <- d[ori_occurred == TRUE]
ar_grp_cutoff <- quantile(d$ar_per_10e3, probs = probs)
d$ar_grp <- sapply(d$ar_per_10e3, function(x) get_group_id(x, ar_grp_cutoff))

res <- rbindlist(lapply(1:10, function(i) 
  calculate_impact_by_subset(d, i, subset_var = "ar_grp", 
          grouping_vars = c("vacc_week","vacc_cov","case_trigger"))))

res %>% 
  filter(vacc_week %in% c(5), vacc_cov == 0.75, subset_id %in% c(1, 10), 
         case_trigger %in% c(1,100)) %>%
  group_by(subset_id, case_trigger) %>%
  summarize(mprc = pct_reduc_case_median)



# Create heatmap of impact by outbreak size and case trigger
plt4as <- res |>
  filter(vacc_cov %in% VC, vacc_week %in% c(5)) |>
  ggplot(aes(x = case_trigger, y = subset_id)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case_median)) +
  metR::geom_contour2(aes(z = pct_reduc_case_median, label = stat(level)),
                     linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)", limits = c(0, 70),
                       breaks = seq(0, 60, 20),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Number of cases that triggered the ORI", y = "Attack rate percentile") +
  scale_x_continuous()+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light() +
  theme_tight() +
  theme(legend.position = "bottom")

# Duration group
dur_grp_cutoff <- quantile(d$duration_weeks, probs = probs)
d$dur_grp <- sapply(d$duration_weeks, function(x) get_group_id(x, dur_grp_cutoff))

res <- rbindlist(lapply(1:10, function(i) 
  calculate_impact_by_subset(d, i, subset_var = "dur_grp", 
          grouping_vars = c("vacc_week","vacc_cov","case_trigger"))))

res %>% 
  filter(vacc_week %in% c(5), vacc_cov == 0.75, subset_id %in% c(1, 10), 
         case_trigger %in% c(1,100)) %>%
  group_by(subset_id, case_trigger) %>%
  summarize(mprc = pct_reduc_case_median)

plt4bs <- res |>
  filter(vacc_cov %in% VC, vacc_week %in% c(5)) |>
  ggplot(aes(x = case_trigger, y = subset_id)) +
  metR::geom_contour_fill(aes(z = pct_reduc_case_median)) +
  metR::geom_contour2(aes(z = pct_reduc_case_median, label = stat(level)),
                     linetype = "dotted") +
  scale_fill_viridis_c("Median cases averted (%)", limits = c(0,70),
                       breaks = seq(0, 60, 20),
                       labels = scales::percent_format(scale = 1)) +
  labs(x = "Number of cases that triggered the ORI", y = "Duration percentile") +
  scale_x_continuous()+
  scale_y_continuous(breaks = c(1, 4, 7, 10),
                     labels = c(expression(">="*P[0*"%"]),
                              expression(">="*P[30*"%"]),
                              expression(">="*P[60*"%"]),
                              expression(">="*P[90*"%"]))) +
  theme_light()+
  theme_tight()+
  theme(legend.position = "bottom")

# Combine all four plots into a single figure
plt4s <- (plt4as + plt4bs) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2,
              nrow = 1,
              guides = "collect") &
  theme(plot.tag.position = c(0, 0.9),
        plot.tag = element_text(size = 13),
        legend.position = "top")

# Save figure
if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig4S_case_trigger_ar_dur_grp_", tstamp(), ".pdf"),
         plt4s,
         device = cairo_pdf, width = fgw2*fac, 
         height = fgw1*1.15*fac, units = "mm")
}

 
```


### The same set of outbreaks across ORI delays (Not included in the paper)

```{r}
# look for ids that are consistent across 
wks <- c(1, 5, 9, 13)
ids0 <- cea %>% 
  filter(vacc_week == max(wks) & runid == 1 & vacc_cov == 0.75) %>% 
  filter(ori_occurred) %>% 
  pull(id_outbreak)

vax_missed_id <- vector('list', length(wks))

for (i in 1:length(wks)) {
  cea %>% 
    filter(id_outbreak %in% ids0) %>%
    filter(vacc_week == wks[i] & runid == 1 & vacc_cov == 0.75) %>% 
    filter(!ori_occurred) %>% 
    pull(id_outbreak) -> vax_missed_id[[i]]
}
map_vec(vax_missed_id, function(i) length(i))

# Create summary for time-triggered vaccination at different times
wks <- c(1, 5, 9, 13)
tab <- data.frame()
interval_est <- "IQR"
cea_ids0 <- cea %>% filter(id_outbreak %in% ids0)
for (wk in wks) {
  d <- cea_ids0[(vacc_cov == VC & vacc_week == wk)]
  tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
  tab_temp$n_outbreaks <- nrow(d[d$runid == 1])
  tab <- rbind(tab, tab_temp)
}
tab$week <- wks
tab$group <- "All"
tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/1192*100, digits=1), "%)") 

if (save_table) {
  fwrite(tab, paste0("outputs/table3_ids0_onset_triggered_impact_IQR_", 
                     tstamp(), ".csv"), row.names = FALSE)
}
```
## Impact of outbreak subsets by size

This was not included in the manuscript
```{r tableS3}
sample_size_values <- c(1,10,50,100) 
wks <- c(1, 5, 9, 13)
# Calculate summary statistics
tab <- data.frame()

for (ss in sample_size_values) {
  for (wk in wks) {
    d <- cea_subsets_by_size_q2[vacc_week == wk & vacc_cov == VC & sample_size == ss]
    # tab <- rbind(tab, impact_summary(d, probs=c(0.25, 0.5, 0.75)))
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- median(d$n_outbreaks) # mean across runids
    tab <- rbind(tab, tab_temp)
  }
}
tab$week <- rep(wks, length(sample_size_values))
tab$sample_size <- rep(sample_size_values, each=length(wks))

tab1 <- data.frame()
for (ss in sample_size_values) {
  for (wk in wks) {
    d <- cea_subsets_by_size_q3[vacc_week == wk & vacc_cov == VC & sample_size == ss]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- median(d$n_outbreaks) # mean across runids
    tab1 <- rbind(tab1, tab_temp)
  }
}
tab1$week <- rep(wks, length(sample_size_values))
tab1$sample_size <- rep(sample_size_values, each=length(wks))

tab2 <- data.frame()
for (ss in sample_size_values) {
  for (wk in wks) {
    d <- cea_subsets_by_size_q4[vacc_week == wk & vacc_cov == VC & sample_size == ss]
    # tab <- rbind(tab, impact_summary(d, probs=c(0.25, 0.5, 0.75)))
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- median(d$n_outbreaks) # mean across runids
    tab2 <- rbind(tab2, tab_temp)
  }
}
tab2$week <- rep(wks, length(sample_size_values))
tab2$sample_size <- rep(sample_size_values, each=length(wks))

tabs <- rbindlist(list(tab, tab1, tab2))
tabs$subsets <- rep(paste0("Q", 1:3), each=nrow(tab))

if (save_table) {
  write.csv(tabs, paste0("outputs/tableS2_outbreak_subsets_by_size_IQR_", 
                        tstamp(), ".csv"), 
            row.names = FALSE)
}
```
Outbreak data sets vary across case trigger thresholds and implementation delays
```{r table3}
# Create summary for case-triggered vaccination at different thresholds
vwks <- c(1, 5, 9, 13)
ctgs <- c(1, 10, 50, 100)
tab <- data.frame()

cea_ct_ori <- cea_ct[ori_occurred == TRUE]

for (wk in vwks) {
  for (ctg in ctgs) {
    d <- cea_ct_ori[(vacc_cov == VC & vacc_week == wk & case_trigger == ctg)]
    tab_temp <- impact_summary(d, probs=c(0.25, 0.5, 0.75))
    tab_temp$n_outbreaks <- nrow(d[d$runid == 1])
    tab <- rbind(tab, tab_temp)
  }
}
tab$case_trigger <- rep(ctgs, length(vwks))
tab$vacc_week <- rep(vwks, each=length(ctgs))

tab$n_outbreaks <- paste0(format_num(tab$n_outbreaks, big.mark=","), " (", round(tab$n_outbreaks/1192*100, digits=1), "%)")

if (save_table) {
  fwrite(tab, paste0("outputs/table4_case_triggered_impact_IQR_", tstamp(), ".csv"), 
            row.names = FALSE)
}
```


