---
title: "Vaccine Impact Simulation"
author: "Jong-Hoon Kim"
date: "2025-06-18"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

# Introduction

This document provides a comprehensive analysis of oral cholera vaccine (OCV) impact using data from outbreaks in sub-Saharan Africa. We model three vaccination strategies:

1. **Outbreak onset-triggered vaccination**: Outbreak response immunization (ORI)  is triggered at outbreak onset but can be implemented after a delay.
2. **Case-triggered vaccination**: ORI is triggered only after the cumulative cases reach specific thresholds and can be implemented again after additional delay

We evaluate these strategies at various coverage levels (60%, 75%, and 90%) and quantify impacts in terms of:
- Cases averted
- Deaths averted
- Disability-adjusted life years (DALYs) averted
- Cost-effectiveness metrics

This analysis is organized into sections for parameter preparation, data loading, simulation execution, and results analysis.

# Setup

## Load Packages and Utility Functions

```{r setup}
#| warning: false
#| message: false

# Core packages
library(data.table)
library(magrittr)
library(tidyverse)
library(readxl)
library(janitor)
library(ggplot2)
library(ggridges)
library(patchwork)
library(conflicted)
library(scales)
library(viridis)
library(doParallel)  # For parallel processing
library(MASS)

# Resolve conflicts in favor of dplyr functions
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflicts_prefer(data.table::first)
conflicts_prefer(data.table::year)

# Load utility functions
source("utils.R")  # Custom utility functions for vaccine impact calculations

# Global parameters
VC3 <- c(0.6, 0.75, 0.9)      # Standard vaccine coverage values (60%, 75%, 90%)
VC <- VC3[2]                  # Reference coverage (75%) 
VW <- 9                       # Reference vaccination week (median from GTFCC data)
VW_seq <- seq(1, 19, 2)       # Sequence of vaccination weeks to evaluate
PROBS <- c(0.025, 0.25, 0.5, 0.75, 0.975)  # Probability quantiles for uncertainty
YEAR <- as.Date("2020-12-28") # Most recent date with OCV campaign data

# Flags for output control
save_csv <- FALSE             # Whether to save results as CSV
save_rds <- FALSE             # Whether to save results as RDS
save_plot <- FALSE            # Whether to save generated plots
save_table <- FALSE           # Whether to save generated tables

# Figure dimensions
figure_size <- data.frame(
  journal = c("Lancet"),
  single = c(85),
  double = c(178)
)
fgw1 <- figure_size[figure_size$journal == "Lancet",]$single
fgw2 <- figure_size[figure_size$journal == "Lancet",]$double
```

# Parameter Preparation

## Parameter Samples for Simulation

```{r params-sample}
# Load parameter definitions
dat_params <- data.table::fread("data/parameters.csv")

# Draw parameter samples for multiple simulation runs
nruns <- 200  # Number of simulation runs
params <- draw_parameter_samples(nruns = nruns, parameter_data = dat_params)

# Save parameter samples
if(save_csv) {
  data.table::fwrite(params, paste0("outputs/params_", tstamp(), ".csv"))
}

```

## Vaccine Coverage Data

```{r vacc-coverage}
# Vaccine coverage data from Table 2 in Pezzoli et al. 2020
vacc_cov_data <- 
  data.frame(setting = c("Endemic", "Humanitarian crisis", "Outbreak"),
           admin_cov_first_dose = c(93.6, 82.9, 98.8)/100,
           admin_cov_second_dose = c(92.4, 82.0, 93.0)/100, 
           estimate_cov_at_least_one = c(85.1, 73.2, 86.3)/100)

# Calculate first-round coverage for different settings
vacc_cov_data |> 
  filter(setting != "Endemic") |>
  group_by(setting) |>
  summarize(
    frac_effect = estimate_cov_at_least_one / 
                 (1 - (1 - admin_cov_first_dose) * (1 - admin_cov_second_dose)),
    vacc_cov_r1 = admin_cov_first_dose * frac_effect
  ) -> vacc_cov_r1

# We use the vaccine coverage for "Outbreak" setting rather than "Humanitarian crisis"
```

## Indirect Vaccine Effectiveness

```{r ive-model}
# Load previously fitted Bayesian model for indirect vaccine effectiveness (IVE)
ive_fit <- readRDS("outputs/ive_yrep_20241010.rds")
ive_data <- ive_fit$yrep
rm(ive_fit)  # Remove large object to save memory

# Example of how we compute vaccine impact
# The impact of vaccination at coverage proportion π is:
# P_{i,π} = 1-[π(1-DVE_i)(1-IVE) + (1-π)(1-IVE)]
# where DVE_i is direct vaccine efficacy for age group i
# and IVE is indirect vaccine effectiveness
```

# Load Outbreak Data

## Outbreak and OCV Campaign Data

```{r load-data}
#| warning: false
# Load outbreak time series data
outbreak_data_all <- as.data.frame(fread("outputs/outbreak_data_all_20250228.csv"))

# Load outbreaks not affected by OCVs (for modeling)
outbreak_data <- fread("outputs/outbreak_data_20250228.csv") 
outbreaks_filtered <-
  readRDS("outputs/outbreak_summary_filtered_20250418.rds")

# Load OCV usage data
ocv_campaign_events <- fread("data/GTFCC_Dashboard_23April2025.csv")
# Filter outbreak-response (i.e., ICG-approved) events
ocv_campaign_events %>% 
  filter(APPROVED_BY == "ICG") %>% 
  select(COUNTRY, DATE_ROUND_ONE, DOSE_DELIVERED_ROUND_ONE) %>% 
  mutate(DATE_ROUND_ONE = as.Date(DATE_ROUND_ONE)) -> ori_events

ori_events <- ori_events[complete.cases(ori_events),]
ori_events <- clean_names(ori_events)
ori_events$year <- lubridate::year(ori_events$date_round_one)

# Calculate delay statistics between request and vaccination
delay_req_r1 <- as.integer(ocv_campaign_events$DATE_ROUND_ONE - ocv_campaign_events$DATE_REQUEST)
ocv_delay <- quantile(delay_req_r1, probs=c(0.25,0.5,0.75,0.9), na.rm=TRUE)

# Historical OCV doses deployed by year
d <- clean_names(fread("data/OCV_Use_GTFCC_Dashboard.csv"))
d |> 
  filter(shipped == "Shipped", type == "Reactive") |> 
  group_by(year) |> 
  summarize(num_ocv = sum(dose)) |>
  setDT() -> ocv_year

ocv_year$cum_num_ocv <- cumsum(ocv_year$num_ocv)

# OCV campaign data
docv_ <- clean_names(fread("data/outbreak_files_Oct2024/ocv_long_dataset.csv"))
docv <- docv_[(ocv_before_1 == TRUE | ocv_before_2 == TRUE | 
              ocv_before_3 == TRUE | ocv_during == TRUE)]

# Get outbreak summary data
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Oct2024/outbreak_summary.csv")))

# Create unique IDs for each outbreak
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# Join outbreak data with summary details
outbk_data_summary <- 
  left_join(filter(outbreak_data, week == 1), 
            ds[, c("outbreak_pop", "total_s_ch", "total_c_ch", "total_deaths", "duration_weeks", 
                  "peak_week_incidence", "time_to_peak_week", "peak_week_incidence_per_100percent", 
                  "peak_week_incidence_per_10e3", "mean_r0_1st_week", "area_per_1km2", 
                  "rural_urban", "ar_per_10e3", "cfr_per_100percent", "id_outbreak")], 
            by = "id_outbreak")

outbk_data_summary %>% 
  group_by(spatial_scale) %>% 
  summarize_quantiles(variable=cfr_per_100percent)
# outbreaks that do not provide deaths information
sum(is.na(outbreaks_filtered$total_deaths))
```

## Load Economic and Demographic Data

```{r econ-demo-data}
# Load life expectancy and GDP data
life_exp_data <- as.data.frame(fread("outputs/wpp_life_expectancy_20241022.csv"))
life_exp_data %>% 
  filter(Year >= 2010, Year <= 2020, `Region, subregion, country or area *` == "Sub-Saharan Africa") %>% 
  pull(`25`) %>%  
  as.numeric() -> life_exp_25

summary(life_exp_25)

# Load population by age data 
age_dist <- as.data.frame(fread("outputs/wpp_pop_by_age_20241022.csv"))
age_dist %>% 
  filter(Year >= 2010, Year <= 2020, `Region, subregion, country or area *` == "Sub-Saharan Africa") %>% 
  pull(prop_u5) -> prop_U5

summary(prop_U5)

gdp_data <- read_xls("data/GDP_WorldBank.xls")
gdp_data %>% 
  filter(`Country Name` == "Sub-Saharan Africa (IDA & IBRD countries)") %>%
  select(c(`2010`:`2020`)) |> 
  as.numeric() -> gdp_ssa

summary(gdp_ssa)

workforce_data <- read_xls("data/Workforce_Worldbank.xls")
workforce_data %>% 
  filter(`Country Name` == "Sub-Saharan Africa (excluding high income)")%>%
  select(c(`2010`:`2020`)) |> 
  as.numeric() -> wf_ssa

summary(wf_ssa)

# Consumer price index for cost adjustment (2015 to 2023)
dis_params <- as.data.frame(data.table::fread("data/parameters.csv"))

# Standardize country code variable names
names(life_exp_data)[names(life_exp_data) == "ISO3 Alpha-code"] <- "country"
names(life_exp_data)[names(life_exp_data) == "Year"] <- "year"
names(workforce_data)[names(workforce_data) == "Country Code"] <- "country"
names(gdp_data)[names(gdp_data) == "Country Code"] <- "country"

# Format workforce data into long format
workforce_long <- 
  workforce_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "pct_workforce")
workforce_long$year <- as.integer(workforce_long$year)

# Format GDP data into long format
gdp_long <- 
  gdp_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "gdp")
gdp_long$year <- as.integer(gdp_long$year)
```

## Cost-Effectiveness Parameters

```{r cost-params}
# Disease-specific parameters
# convert to data.frame because extracting a column of the data.table as numbers gives me troubles
parms <- as.data.frame(fread("data/parameters.csv"))
parms <- filter(parms, Disease == "Cholera")
# GBD data deleted and replaced with Mbewe (2025) Open Forum Infect Dis data
parms <- filter(parms, !(Parameter == "Prop_Moderate" & Value == 0.289))
parms <- filter(parms, !(Parameter == "Prop_Severe" & Value == 0.069))

# Disease burden parameters
# pr_asymp <- parms[Parameter == "Prop_Asymptomatic", "Value"]
# pr_mild <- parms[Parameter == "Prop_Mild", "Value"]
pr_moderate <- parms[parms$Parameter == "Prop_Moderate", "Value"]
pr_severe <- parms[parms$Parameter == "Prop_Severe", "Value"]
dur <- parms[parms$Parameter == "Duration_Illness", "Value"]
# wt_asymp <- parms[Parameter == "Disability_Weight_Asymptomatic", "Value"]
# wt_mild <- parms[Parameter == "Disability_Weight_Mild", "Value"]
wt_moderate <- parms[parms$Parameter == "Disability_Weight_Moderate", "Value"]
wt_severe <- parms[parms$Parameter == "Disability_Weight_Severe", "Value"]

# Vaccination costs
vacc_cost_per_dose <- parms[parms$Parameter == "Vaccine_Cost", "Value"]
vacc_delivery_cost <- parms[parms$Parameter == "Vaccine_Delivery_Cost", "Value"]
vacc_shipping_cost <- parms[parms$Parameter == "Vaccine_Shipping_Cost", "Value"]

# Direct medical costs
patient_cost_hosp <- parms[parms$Parameter == "Patient_Cost_Hosp", "Value"]
patient_cost_outpt <- parms[parms$Parameter == "Patient_Cost_Outpt", "Value"]
public_cost_hosp <- parms[parms$Parameter == "Public_Cost_Hosp", "Value"]
public_cost_outpt <- parms[parms$Parameter == "Public_Cost_Outpt", "Value"] 

# Productivity costs
patient_workday_lost <- parms[parms$Parameter == "Pt_Workdays_Lost", "Value"]
caregiver_workday_lost <- parms[parms$Parameter == "Caregiver_Workdays_Lost", "Value"]
mean_age_inf <- parms[parms$Parameter == "Mean_Age_Infection", "Value"]

# Vaccination regimen
dose_regimen <- 1  # Single-dose regimen
vacc_cost_per_person <- vacc_cost_per_dose * dose_regimen

# Calculate cost of illness per patient (weighted average)
pr_tot <- pr_moderate + pr_severe
coi_per_patient <- 
    pr_moderate / pr_tot * (patient_cost_outpt + public_cost_outpt) + 
    pr_severe / pr_tot * (patient_cost_hosp + public_cost_hosp)

# Adjust costs for inflation (2015 to 2023)
cpi_percent <- read_xls("data/imf-dm-export-20241028.xls")
cpi_percent <- cpi_percent %>%
  pivot_longer(cols = `2015`:`2029`, names_to = "year", values_to = "value")
cpi_percent$year <- as.numeric(cpi_percent$year)

# Adjust medical costs (2015 to 2023)
cpi_med <- cpi_percent %>% filter(year >= 2016 & year <= 2023)
xi_med <- (cpi_med$value/100 + 1)
inflation_factor_med <- prod(xi_med)

patient_cost_hosp <- patient_cost_hosp * inflation_factor_med
patient_cost_outpt <- patient_cost_outpt * inflation_factor_med
public_cost_hosp <- public_cost_hosp * inflation_factor_med
public_cost_outpt <- public_cost_outpt * inflation_factor_med

# Adjust vaccine delivery cost (2016 to 2023)
cpi_vac <- cpi_percent %>% filter(year >= 2017 & year <= 2023)
xi_vac <- (cpi_vac$value/100 + 1)
inflation_factor_vac <- prod(xi_vac)
vacc_delivery_cost <- vacc_delivery_cost * inflation_factor_vac
```

# Vaccine Impact Simulation

## Outbreak Onset-Triggered Vaccination Strategy 

This section models vaccination initiated at specific weeks after outbreak onset.

```{r onset-trigger-sim, eval=FALSE}
tic <- Sys.time()  # Start timer
# dat <- outbreak_data %>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 

# sensitivity analyses 
outbreaks_excl <- read_rds("outputs/outbreaks_excluded_20250612.rds")
dat <- outbreak_data_all %>% filter(id_outbreak %in% outbreaks_excl$id_outbreak) 

# Setup for parallel processing
vacc_covs <- c(0.6, 0.75, 0.9)    # Vaccine coverage levels to simulate
vacc_weeks <- seq(1, 15, by = 2)  # Vaccination weeks to simulate
nruns <- 200                      # Number of simulation runs per configuration
batch_size <- 10                  # Number of runs per parallel batch

# Loop through batches of runs for parallel execution
for (batch_id in seq(1, nruns, by = batch_size)) {
  message(paste("Processing batch ID:", batch_id))
  run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in current batch

  # Create and register parallel cluster
  cl <- makeCluster(getOption("cl.cores", detectCores() - 3))
  registerDoParallel(cl)

  # Run simulations in parallel
  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = dat,
        ive_data = ive_data,
        parameters = params,
        age_dist = age_dist,
        vacc_week = vacc_weeks,    # Outbreak onset-triggered vaccination
        case_trigger = NULL,
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)

  # Save batch results
  fwrite(rbindlist(svim_week_list), 
         paste0("outputs/svim_week_list_vacc_week_batch_", batch_id, "_", tstamp(), ".csv"))
}
toc <- Sys.time()
message(paste("Total execution time:", toc - tic))
```

## Case-Triggered Vaccination Strategy

```{r case-trigger-sim, eval=FALSE}
tic <- Sys.time()  # Start timer
dat <- outbreak_data
dat <- dat %>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak)

# Define case thresholds that trigger vaccination
case_triggers <- c(1, 10, 50, 100)
vacc_covs <- c(0.75, 0.9)
vacc_weeks <- seq(1, 15, 2)
nruns <- 200                # Number of simulation runs
batch_size <- 10            # Number of runs per batch

# Loop through batches of runs for parallel execution
for (batch_id in seq(1, nruns, by = batch_size)) {
  message(paste("Processing batch ID:", batch_id))
  run_ids <- batch_id:(batch_id + batch_size - 1)

  # Create and register parallel cluster
  cl <- makeCluster(getOption("cl.cores", detectCores() - 2))
  registerDoParallel(cl)

  # Run simulations in parallel
  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = dat,
        ive_data = ive_data,
        parameters = params,
        age_dist = age_dist,
        vacc_week = vacc_weeks,
        case_trigger = case_triggers,  # Case-triggered vaccination
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)
  
  # Save batch results
  saveRDS(rbindlist(svim_week_list),
          paste0("outputs/svim_week_case_trigger_batch_", 
                 batch_id, "_", tstamp(), ".rds"))
}
toc <- Sys.time()
message(paste("Total execution time:", round(toc - tic, digits=2), "minutes"))
```

## Combine Simulation Results

This section combines results from batch processing and organizes them for analysis.

```{r combine-results}
# ---- 1. Time-triggered vaccination results ----
id_begin <- 1
id_end <- 200
batch_size <- 10
batch_ids <- seq(id_begin, id_end, by = batch_size)

svim_df <- data.frame()
for (i in 1:length(batch_ids)) {
  message(paste("Processing batch", i, "of", length(batch_ids)))
  # d <- read_rds(paste0("outputs/svim_week_list_vacc_week_batch_",
  #                      batch_ids[i], "_20250514.rds"))
  d <- fread(paste0("outputs/svim_week_list_vacc_week_batch_",
                       batch_ids[i], "_20250514.csv"))
  svim_df <- rbindlist(list(svim_df, sum_over_week(d)))
  rm(d)  # Free memory
}

# Save combined results
saveRDS(svim_df, 
        paste0("outputs/svim_onset_trigger_sum_over_week_runid_1_200_",
               tstamp(), ".rds"))

# ---- 2. Case-triggered vaccination results ----
id_begin <- 1
id_end <- 200
batch_size <- 10
batch_ids <- seq(id_begin, id_end, by = batch_size)

svim_df <- data.frame()
for (i in 1:length(batch_ids)) {
  message(paste("Processing batch", i, "of", length(batch_ids)))
  d <- read_rds(paste0("outputs/svim_week_case_trigger_batch_",
                   batch_ids[i], "_20250702.rds"))
  svim_df <- rbindlist(list(svim_df, sum_over_week(d, case_trigger = TRUE)))
  rm(d) 
}

# Save combined results
saveRDS(svim_df, 
        paste0("outputs/svim_case_trigger_sum_over_week_runid_1_200_",
               tstamp(), ".rds"))
```

# Cost-Effectiveness Analysis

## Calculate Impact Metrics for Outbreak Onset-Triggered Vaccination

```{r cea-time-trigger}
# Load summarized time-triggered vaccination results
d <- read_rds("outputs/svim_onset_trigger_sum_over_week_runid_1_200_20250716.rds")
# d <- read_rds("outputs/svim_onset_trigger_sum_over_week_runid_1_200_20250520.rds")
# Process each vaccine coverage level
for (vc in VC3) {
  message(paste("Processing results for vaccine coverage =", vc))
  
  # Full cost-effectiveness analysis
  svim_tt_cea <- add_cea_variables(d[vacc_cov == vc], 
                                   mean_age_inf = mean_age_inf)
  svim_tt_cea <- add_cea_results(svim_tt_cea)
  
  # Save detailed results
  saveRDS(svim_tt_cea,
         paste0("outputs/cea_", vc, "_onset_trigger_", 
                tstamp(), ".rds"))
  
  rm(svim_tt_cea)
  
  # # Calculate mean across simulation runs
  # svim_tt_mean <- mean_across_runid(d[vacc_cov == vc])
  # svim_tt_mean_cea <- add_cea_variables(svim_tt_mean, 
  #                                       mean_age_inf = mean_age_inf)
  # svim_tt_mean_cea <- add_cea_results(svim_tt_mean_cea)
  # 
  # # Save average results
  # saveRDS(svim_tt_mean_cea,
  #        paste0("outputs/cea_", vc, "_onset_trigger_mean_across_runid_", 
  #               tstamp(), ".rds"))
  
  # Clean up to save memory
  # rm(svim_tt_cea, svim_tt_mean, svim_tt_mean_cea)
}
rm(d)  # Free memory
```

## Calculate Impact Metrics for Case-Triggered Vaccination

```{r cea-case-trigger}
# Load summarized case-triggered vaccination results
d_ct <- 
  read_rds("outputs/svim_case_trigger_sum_over_week_runid_1_200_20250523.rds")
# d_ct <- 
#   read_rds("outputs/svim_case_trigger_sum_over_week_runid_1_200_20250716.rds")

# Process each vaccine coverage level
# for (vc in VC3) {
for (vc in unique(d_ct$vacc_cov)) {  
  message(paste("Processing results for vaccine coverage =", vc))
  
  # Full cost-effectiveness analysis
  svim_ct_cea <- add_cea_variables(d_ct[vacc_cov == vc], 
                                   mean_age_inf = mean_age_inf)
  svim_ct_cea <- add_cea_results(svim_ct_cea)
  # Save detailed results
  saveRDS(svim_ct_cea,
         paste0("outputs/cea_", vc, "_case_trigger_", 
                tstamp(), ".rds"))
  rm(svim_ct_cea) 
}
```

# Subset Analysis

This section creates and analyzes subsets of outbreaks to simulate different targeting strategies.

## Create Outbreak Subsets: sample until the target number of outbreaks is reached.

```{r}
outbreaks_filtered |>
  group_by(id_outbreak) |> 
  summarize(year = data.table::year(start_date),
            rural_urban = rural_urban,
            ar = ar_per_10e3,
            cfr = cfr_per_100percent,
            spatial_scale = spatial_scale,
            pop = outbreak_pop,
            duration = duration_weeks,
            size = total_s_ch) -> outbk_dat

nsubset <- 200  # Number of simulated subsets
ssize <- seq(1, 9, by=1) # sample size for each subset

for (j in 1:length(ssize)){
  outbreak_subsets <- vector('list', nsubset)
  for (i in 1:nsubset) {
    outbreak_subsets[[i]] <- 
      sample_outbreaks(outbreak_data = outbk_dat,
                       target_count = ssize[j],
                       target_population = Inf,
                       sampling_method = "random",
                       seed = i,
                       max_attempts = 1000)
  
  }
  saveRDS(outbreak_subsets, 
        paste0("outputs/outbreak_subsets_", ssize[j], "_", tstamp(), ".rds"))
}
```

## Create Outbreak Subsets for Global OCV Stockpile Use

```{r}
outbreaks_filtered |>
  group_by(id_outbreak) |> 
  summarize(year = data.table::year(start_date),
            rural_urban = rural_urban,
            ar = ar_per_10e3,
            cfr = cfr_per_100percent,
            spatial_scale = spatial_scale,
            pop = outbreak_pop,
            duration = duration_weeks,
            size = total_s_ch) -> outbk_dat

nsubset <- 200  # Number of simulated subsets

# For 2013 and 2014, we don't have OCV deployment records by event and 
# so outbreaks are randomly sampled to meet the total number of OCVs

DF1 <- data.frame()
for (j in 1:nsubset) {
  set.seed(j)
  message(paste("Simulation run", j))
  df <- data.frame()
  for (i in 2013:2014) {
    yr <- i  # 2013 = the first year when OCVs were deployed
    # Sample outbreaks in proportion to OCVs deployed that year
    smpl <- sample_outbreaks(
      outbreak_data = filter(outbk_dat, year == yr), 
      target_population = ocv_year[year == yr, num_ocv],
      sampling_method = "random")
    
    df <- rbind(df, data.frame(year = smpl$sampled_data$year, 
                              id_outbreak = smpl$sampled_data$id_outbreak))
  }
  df$subset_id <- j
  DF1 <- rbind(DF1, df)
}

# This simulates targeting outbreaks randomly, matched to historical OCV deployment
outbk_dat2 <- outbk_dat[, c("id_outbreak", "pop", "year")]
names(outbk_dat2) <- c("id_outbreak", "population_size", "year")

DF2 <- data.frame()

for (j in 1:nsubset) {
  set.seed(j)
  message(paste("Simulation run", j))
  df <- data.frame()
  for (i in 2015:2020) {
    yr <- i  # 2013 = the first year when OCVs were deployed
    # Sample outbreaks in proportion to OCVs deployed that year
    smpl <- match_ocv_to_outbreaks(
      ocv_deployments = 1000*ori_events[year == yr, dose_delivered_round_one], 
      outbreak_populations = outbk_dat2[outbk_dat$year == yr,],
      error_tolerance = 0.2, maximize_coverage = FALSE)
      
    df <- rbind(df, smpl$valid_matches)
  }

  DF2 <- rbind(DF2,  cbind(df, data.frame(subset_id = j)))
}

DF <- rbind(DF1, DF2[, c("year", "id_outbreak", "subset_id")])

saveRDS(DF, paste0("outputs/random_subsets_by_year_", tstamp(), ".rds"))
```

## Create Outbreak Subsets

```{r create-subsets}
# Prepare outbreak summary data
# ---- 1. Random sampling of outbreaks by year ----
# This simulates targeting outbreaks randomly, matched to historical OCV deployment
DF <- data.frame()

# ---- 2. Case-weighted sampling of outbreaks by year ----
# This simulates targeting outbreaks with more cases, matched to historical OCV deployment
names(dat) <- c("id_outbreak", "year", "rural_urban", "ar", "cfr", "spatial_scale",
                "pop", "duration", "cases")

set.seed(42)
DF <- data.frame()
for (j in 1:200) {
  set.seed(j)
  message(paste("Simulation run", j))
  df <- data.frame()
  for (i in 2013:2022) {
    yr <- i  # 2013 = the first year when OCVs were deployed
    # Sample outbreaks weighted by case count
    smpl <- sample_outbreaks(
      outbreak_data = filter(dat, year == yr), 
      target_population = ocv_year[year == yr, num_ocv],
      sampling_method = "cases")
    
    df <- rbind(df, data.frame(year = smpl$sampled_data$year, 
                         id_outbreak = smpl$sampled_data$id_outbreak))
  }
  df$subset_id <- j
  DF <- rbind(DF, df)
}

# Save case-weighted sampling results
fwrite(DF, paste0("outputs/case_weighted_subsets_by_year_", tstamp(), ".csv"))
```

## Analyze Outbreak Onset-Triggered Vaccination on Subsets

```{r subset-time-trigger}
# Load outbreak subset definitions
DF <- readRDS("outputs/random_subsets_by_year_20250605.rds")

# Create analysis combinations (subset × year)
df <- data.frame(expand.grid(subset_id = 1:200, year_max = 2013:2020))
# Initialize results list
lst <- vector('list', nrow(df))

# Process each subset-year combination
for (i in 1:nrow(df)) {
  # Get outbreaks in this subset up to specified year
  d <- DF[(DF$year <= df$year_max[i] & DF$subset_id == df$subset_id[i]),]
  
  # Extract impact results for these outbreaks
  dat <- cea[id_outbreak %in% d$id_outbreak]
  
  # Sum impact across outbreaks
  datall <- sum_over_outbreaks_by_runid(dat, ori_occurred = T)
  
  # Tag with subset metadata
  datall$subset_id <- df$subset_id[i]
  datall$year_max <- df$year_max[i]
  
  # Store results
  lst[[i]] <- datall
}

# Combine all results
dfall <- rbindlist(lst)

# Save results
saveRDS(dfall, paste0("outputs/cea_0.75_onset_trigger_random_subsets_by_year_",
                     tstamp(), ".rds"))
```

## Analysis over subsets, 1, ..., 100

```{r}
# Filter subset analysis results for easier visualization
set.seed(42)
DF <- data.frame()
ssize <- c(1:9, seq(10,100,10))
for (n in ssize) {
  subset <- read_rds(paste0("outputs/outbreak_subsets_", n, ".rds"))
  for(i in 1:length(subset)){
    d <- subset[[i]]$sampled_data
    d$subset_id <- i
    d$sample_size <- n
    
    DF <- rbind(DF, d)
  }
}

# Create analysis combinations (subset × year)
df <- data.frame(expand.grid(subset_id = 1:200, 
                             sample_size = c(1:10,seq(20,100,10))))
# Initialize results list
lst <- vector('list', nrow(df))

# Process each subset-year combination
for (i in 1:nrow(df)) {
  # Get outbreaks in this subset up to specified year
  d <- DF[(DF$sample_size == df$sample_size[i] & DF$subset_id == df$subset_id[i]),]
  
  # Extract impact results for these outbreaks
  dat <- cea[id_outbreak %in% d$id_outbreak]
  
  # Sum impact across outbreaks
  # datall <- sum_over_outbreaks_by_runid(dat)
  datall <- sum_over_outbreaks_by_runid(dat, ori_occurred = T)
  # Tag with subset metadata
  datall$subset_id <- df$subset_id[i]
  datall$sample_size <- df$sample_size[i]
  
  # Store results
  lst[[i]] <- datall
}

# Combine all results
dfall <- rbindlist(lst)

# Save results
saveRDS(dfall, paste0("outputs/cea_onset_trigger_subsets_by_size_",
                     tstamp(), ".rds"))
```


## Analyze Case-Triggered Vaccination on Subsets

```{r subset-case-trigger}
# Load outbreak subset definitions
DF <- fread("outputs/case_weighted_subsets_by_year_20250313.csv")

# Load vaccination impact results
cea_ct <- fread("outputs/cea_0.75_case_trigger_20250311.csv")

# Create analysis combinations (subset × year)
df <- data.frame(expand.grid(subset_id = 1:200, 
                             year_max = 2013:2022))

# Initialize results list
lst <- vector('list', nrow(df))

# Process each subset-year combination
for (i in 1:nrow(df)) {
  # Get outbreaks in this subset up to specified year
  d <- DF[(DF$year <= df$year_max[i] & DF$subset_id == df$subset_id[i]),]
  
  # Extract impact results for these outbreaks
  dat <- cea_ct[id_outbreak %in% d$id_outbreak]
  
  # Sum impact across outbreaks (specify case_trigger=TRUE for correct handling)
  datall <- sum_over_outbreaks_by_runid(dat, case_trigger = TRUE)
  
  # Tag with subset metadata
  datall$subset_id <- df$subset_id[i]
  datall$year_max <- df$year_max[i]
  
  # Store results
  lst[[i]] <- datall
}

# Combine all results
dfall <- rbindlist(lst)

# Save results
fwrite(dfall, 
       paste0("outputs/cea_0.75_case_trigger_case_weighted_subsets_by_year_", 
              tstamp(), ".csv"))
```


