---
title: "Vaccine Impact Simulation"
author: "Jong-Hoon Kim"
date: "2025-06-18"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

# Introduction

This document provides a comprehensive analysis of oral cholera vaccine (OCV) impact using data from outbreaks in sub-Saharan Africa. We model three vaccination strategies:

1. **Outbreak onset-triggered vaccination**: Outbreak response immunization (ORI)  is triggered at outbreak onset but can be implemented after a delay.
2. **Case-triggered vaccination**: ORI is triggered only after the cumulative cases reach specific thresholds and can be implemented again after additional delay

We evaluate these strategies at various coverage levels (60%, 75%, and 90%) and quantify impacts in terms of:
- Cases averted
- Deaths averted
- Disability-adjusted life years (DALYs) averted
- Cost-effectiveness metrics

This analysis is organized into sections for parameter preparation, data loading, simulation execution, and results analysis.

# Setup

## Load Packages and Utility Functions

```{r setup}
#| warning: false
#| message: false

# Core packages
library(data.table)
library(magrittr)
library(tidyverse)
library(readxl)
library(janitor)
library(ggplot2)
library(ggridges)
library(patchwork)
library(conflicted)
library(scales)
library(viridis)
library(doParallel)  # For parallel processing
library(MASS)
library(purrr)
library(furrr) # For parallel processing

# Resolve conflicts in favor of dplyr functions
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflicts_prefer(data.table::first)
conflicts_prefer(data.table::year)

# Load utility functions
source("R/utils.R")  # Custom utility functions for vaccine impact calculations

# Global parameters
VC3 <- c(0.6, 0.75, 0.9)      # Standard vaccine coverage values (60%, 75%, 90%)
VC <- VC3[2]                  # Reference coverage (75%) 
VW <- 9                       # Reference vaccination week (median from GTFCC data)
VW_seq <- seq(1, 19, 2)       # Sequence of vaccination weeks to evaluate
PROBS <- c(0.025, 0.25, 0.5, 0.75, 0.975)  # Probability quantiles for uncertainty
YEAR <- as.Date("2020-12-28") # Most recent date with OCV campaign data

# Flags for output control
save_csv <- FALSE             # Whether to save results as CSV
save_rds <- FALSE             # Whether to save results as RDS
save_plot <- FALSE            # Whether to save generated plots
save_table <- FALSE           # Whether to save generated tables

# Figure dimensions
figure_size <- data.frame(
  journal = c("Lancet"),
  single = c(85),
  double = c(178)
)
fgw1 <- figure_size[figure_size$journal == "Lancet",]$single
fgw2 <- figure_size[figure_size$journal == "Lancet",]$double
```

# Parameter Preparation

## Parameter Samples for Simulation

```{r params-sample}
# Load parameter definitions
dat_params <- data.table::fread("data/paper/parameters.csv")

# Draw parameter samples for multiple simulation runs
nruns <- 200  # Number of simulation runs
params <- draw_parameter_samples(nruns = nruns, parameter_data = dat_params)

# Save parameter samples
if(save_csv) {
  data.table::fwrite(params, paste0("outputs/params_", tstamp(), ".csv"))
}

```

## Vaccine Coverage Data

```{r vacc-coverage}
# Vaccine coverage data from Table 2 in Pezzoli et al. 2020
vacc_cov_data <- 
  data.frame(setting = c("Endemic", "Humanitarian crisis", "Outbreak"),
           admin_cov_first_dose = c(93.6, 82.9, 98.8)/100,
           admin_cov_second_dose = c(92.4, 82.0, 93.0)/100, 
           estimate_cov_at_least_one = c(85.1, 73.2, 86.3)/100)

# Calculate first-round coverage for different settings
vacc_cov_data |> 
  filter(setting != "Endemic") |>
  group_by(setting) |>
  summarize(
    frac_effect = estimate_cov_at_least_one / 
                 (1 - (1 - admin_cov_first_dose) * (1 - admin_cov_second_dose)),
    vacc_cov_r1 = admin_cov_first_dose * frac_effect
  ) -> vacc_cov_r1

# We use the vaccine coverage for "Outbreak" setting rather than "Humanitarian crisis"
```

## Indirect Vaccine Effectiveness

```{r ive-model}
# Load previously fitted Bayesian model for indirect vaccine effectiveness (IVE)
# ive_fit <- readRDS("outputs/ive_yrep_20241010.rds")
# ive_data2 <- ive_fit$yrep
# rm(ive_fit)  # Remove large object to save memory

ive_fit <- readRDS("outputs/paper/ive_yhat_yrep_20250919.rds")

# 50 colums from 0 to 0.98 with the increment of 0.2
# 60% coverage
# > quantile(ive_fit$yhat[,31], probs=c(0.025, 0.5, 0.975), na.rm=T)
#      2.5%       50%     97.5% 
# 0.5451031 0.7794489 0.9104249 

# 90% coverage
# > quantile(ive_fit$yhat[,46], probs=c(0.025, 0.5, 0.975), na.rm=T)
#      2.5%       50%     97.5% 
# 0.5835098 0.9500404 0.9963310 

ive_data <- ive_fit$yhat
rm(ive_fit)

# Example of how we compute vaccine impact
# The impact of vaccination at coverage proportion π is:
# P_{i,π} = 1-[π(1-DVE_i)(1-IVE) + (1-π)(1-IVE)]
# where DVE_i is direct vaccine efficacy for age group i
# and IVE is indirect vaccine effectiveness
```

# Load Outbreak Data

## Outbreak and OCV Campaign Data

```{r load-data}
#| warning: false
# Load outbreak time series data
outbreak_data_all <- as.data.frame(fread("outputs/paper/outbreak_data_all_20250228.csv"))

# Load outbreaks not affected by OCVs (for modeling)
outbreak_data <- fread("outputs/paper/outbreak_data_20250228.csv") 
outbreaks_filtered <-
  readRDS("outputs/paper/outbreak_summary_filtered_20250418.rds")

# Load OCV usage data
ocv_campaign_events <- fread("data/paper/GTFCC_Dashboard_23April2025.csv")
# Filter outbreak-response (i.e., ICG-approved) events
ocv_campaign_events %>% 
  filter(APPROVED_BY == "ICG") %>% 
  select(COUNTRY, DATE_ROUND_ONE, DOSE_DELIVERED_ROUND_ONE) %>% 
  mutate(DATE_ROUND_ONE = as.Date(DATE_ROUND_ONE)) -> ori_events

ori_events <- ori_events[complete.cases(ori_events),]
ori_events <- clean_names(ori_events)
ori_events$year <- lubridate::year(ori_events$date_round_one)

# Calculate delay statistics between request and vaccination
delay_req_r1 <- as.integer(ocv_campaign_events$DATE_ROUND_ONE - ocv_campaign_events$DATE_REQUEST)
ocv_delay <- quantile(delay_req_r1, probs=c(0.25,0.5,0.75,0.9), na.rm=TRUE)

# Historical OCV doses deployed by year
d <- clean_names(fread("data/paper/OCV_Use_GTFCC_Dashboard.csv"))
d |> 
  filter(shipped == "Shipped", type == "Reactive") |> 
  group_by(year) |> 
  summarize(num_ocv = sum(dose)) |>
  setDT() -> ocv_year

ocv_year$cum_num_ocv <- cumsum(ocv_year$num_ocv)

# OCV campaign data
docv_ <- clean_names(fread("data/paper/outbreak_files_Oct2024/ocv_long_dataset.csv"))
docv <- docv_[(ocv_before_1 == TRUE | ocv_before_2 == TRUE | 
              ocv_before_3 == TRUE | ocv_during == TRUE)]

# Get outbreak summary data
ds <- clean_names(as.data.frame(fread("data/paper/outbreak_files_Oct2024/outbreak_summary.csv")))

# Create unique IDs for each outbreak
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# Join outbreak data with summary details
outbk_data_summary <- 
  left_join(filter(outbreak_data, week == 1), 
            ds[, c("outbreak_pop", "total_s_ch", "total_c_ch", "total_deaths", "duration_weeks", 
                  "peak_week_incidence", "time_to_peak_week", "peak_week_incidence_per_100percent", 
                  "peak_week_incidence_per_10e3", "mean_r0_1st_week", "area_per_1km2", 
                  "rural_urban", "ar_per_10e3", "cfr_per_100percent", "id_outbreak")], 
            by = "id_outbreak")

outbk_data_summary %>% 
  group_by(spatial_scale) %>% 
  summarize_quantiles(variable=cfr_per_100percent)
# outbreaks that do not provide deaths information
sum(is.na(outbreaks_filtered$total_deaths))
```

## Load Economic and Demographic Data

```{r econ-demo-data}
# Load life expectancy and GDP data
life_exp_data <- as.data.frame(fread("outputs/paper/wpp_life_expectancy_20241022.csv"))
life_exp_data %>% 
  filter(Year >= 2010, Year <= 2020, `Region, subregion, country or area *` == "Sub-Saharan Africa") %>% 
  pull(`25`) %>%  
  as.numeric() -> life_exp_25

summary(life_exp_25)

# Load population by age data 
age_dist <- as.data.frame(fread("outputs/paper/wpp_pop_by_age_20241022.csv"))
age_dist %>% 
  filter(Year >= 2010, Year <= 2020, `Region, subregion, country or area *` == "Sub-Saharan Africa") %>% 
  pull(prop_u5) -> prop_U5

summary(prop_U5)

gdp_data <- read_xls("data/paper/GDP_WorldBank.xls")
gdp_data %>% 
  filter(`Country Name` == "Sub-Saharan Africa (IDA & IBRD countries)") %>%
  select(c(`2010`:`2020`)) |> 
  as.numeric() -> gdp_ssa

summary(gdp_ssa)

workforce_data <- read_xls("data/paper/Workforce_Worldbank.xls")
workforce_data %>% 
  filter(`Country Name` == "Sub-Saharan Africa (excluding high income)")%>%
  select(c(`2010`:`2020`)) |> 
  as.numeric() -> wf_ssa

summary(wf_ssa)

# Consumer price index for cost adjustment (2015 to 2023)
dis_params <- as.data.frame(data.table::fread("data/paper/parameters.csv"))

# Standardize country code variable names
names(life_exp_data)[names(life_exp_data) == "ISO3 Alpha-code"] <- "country"
names(life_exp_data)[names(life_exp_data) == "Year"] <- "year"
names(workforce_data)[names(workforce_data) == "Country Code"] <- "country"
names(gdp_data)[names(gdp_data) == "Country Code"] <- "country"

# Format workforce data into long format
workforce_long <- 
  workforce_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "pct_workforce")
workforce_long$year <- as.integer(workforce_long$year)

# Format GDP data into long format
gdp_long <- 
  gdp_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "gdp")
gdp_long$year <- as.integer(gdp_long$year)
```

## Cost-Effectiveness Parameters

```{r cost-params}
# Disease-specific parameters
# convert to data.frame because extracting a column of the data.table as numbers gives me troubles
parms <- as.data.frame(fread("data/paper/parameters.csv"))
parms <- filter(parms, Disease == "Cholera")
# GBD data deleted and replaced with Mbewe (2025) Open Forum Infect Dis data
parms <- filter(parms, !(Parameter == "Prop_Moderate" & Value == 0.289))
parms <- filter(parms, !(Parameter == "Prop_Severe" & Value == 0.069))

# Disease burden parameters
# pr_asymp <- parms[Parameter == "Prop_Asymptomatic", "Value"]
# pr_mild <- parms[Parameter == "Prop_Mild", "Value"]
pr_moderate <- parms[parms$Parameter == "Prop_Moderate", "Value"]
pr_severe <- parms[parms$Parameter == "Prop_Severe", "Value"]
dur <- parms[parms$Parameter == "Duration_Illness", "Value"]
# wt_asymp <- parms[Parameter == "Disability_Weight_Asymptomatic", "Value"]
# wt_mild <- parms[Parameter == "Disability_Weight_Mild", "Value"]
wt_moderate <- parms[parms$Parameter == "Disability_Weight_Moderate", "Value"]
wt_severe <- parms[parms$Parameter == "Disability_Weight_Severe", "Value"]

# Vaccination costs
vacc_cost_per_dose <- parms[parms$Parameter == "Vaccine_Cost", "Value"]
vacc_delivery_cost <- parms[parms$Parameter == "Vaccine_Delivery_Cost", "Value"]
vacc_shipping_cost <- parms[parms$Parameter == "Vaccine_Shipping_Cost", "Value"]

# Direct medical costs
patient_cost_hosp <- parms[parms$Parameter == "Patient_Cost_Hosp", "Value"]
patient_cost_outpt <- parms[parms$Parameter == "Patient_Cost_Outpt", "Value"]
public_cost_hosp <- parms[parms$Parameter == "Public_Cost_Hosp", "Value"]
public_cost_outpt <- parms[parms$Parameter == "Public_Cost_Outpt", "Value"] 

# Productivity costs
patient_workday_lost <- parms[parms$Parameter == "Pt_Workdays_Lost", "Value"]
caregiver_workday_lost <- parms[parms$Parameter == "Caregiver_Workdays_Lost", "Value"]
mean_age_inf <- parms[parms$Parameter == "Mean_Age_Infection", "Value"]

# Vaccination regimen
dose_regimen <- 1  # Single-dose regimen
vacc_cost_per_person <- vacc_cost_per_dose * dose_regimen

# Calculate cost of illness per patient (weighted average)
pr_tot <- pr_moderate + pr_severe
coi_per_patient <- 
    pr_moderate / pr_tot * (patient_cost_outpt + public_cost_outpt) + 
    pr_severe / pr_tot * (patient_cost_hosp + public_cost_hosp)

# Adjust costs for inflation (2015 to 2023)
cpi_percent <- read_xls("data/paper/imf-dm-export-20241028.xls")
cpi_percent <- cpi_percent %>%
  pivot_longer(cols = `2015`:`2029`, names_to = "year", values_to = "value")
cpi_percent$year <- as.numeric(cpi_percent$year)

# Adjust medical costs (2015 to 2023)
cpi_med <- cpi_percent %>% filter(year >= 2016 & year <= 2023)
xi_med <- (cpi_med$value/100 + 1)
inflation_factor_med <- prod(xi_med)

patient_cost_hosp <- patient_cost_hosp * inflation_factor_med
patient_cost_outpt <- patient_cost_outpt * inflation_factor_med
public_cost_hosp <- public_cost_hosp * inflation_factor_med
public_cost_outpt <- public_cost_outpt * inflation_factor_med

# Adjust vaccine delivery cost (2016 to 2023)
cpi_vac <- cpi_percent %>% filter(year >= 2017 & year <= 2023)
xi_vac <- (cpi_vac$value/100 + 1)
inflation_factor_vac <- prod(xi_vac)
vacc_delivery_cost <- vacc_delivery_cost * inflation_factor_vac
```

# Vaccine Impact Simulation

## Outbreak Onset-Triggered Vaccination Strategy 

This section models vaccination initiated at specific weeks after outbreak onset.

```{r onset-trigger-sim, eval=FALSE}
tic <- Sys.time()  # Start timer
# dat <- outbreak_data %>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak) 

# sensitivity analyses 
outbreaks_excl <- read_rds("outputs/outbreaks_excluded_20250612.rds")
dat <- outbreak_data_all %>% filter(id_outbreak %in% outbreaks_excl$id_outbreak) 

# Setup for parallel processing
vacc_covs <- c(0.6, 0.75, 0.9)    # Vaccine coverage levels to simulate
vacc_weeks <- seq(1, 15, by = 2)  # Vaccination weeks to simulate
nruns <- 200                      # Number of simulation runs per configuration
batch_size <- 10                  # Number of runs per parallel batch

# Loop through batches of runs for parallel execution
for (batch_id in seq(1, nruns, by = batch_size)) {
  message(paste("Processing batch ID:", batch_id))
  run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in current batch

  # Create and register parallel cluster
  cl <- makeCluster(getOption("cl.cores", detectCores() - 3))
  registerDoParallel(cl)

  # Run simulations in parallel
  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = dat,
        ive_data = ive_data,
        parameters = params,
        age_dist = age_dist,
        vacc_week = vacc_weeks,    # Outbreak onset-triggered vaccination
        case_trigger = NULL,
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)

  # Save batch results
  fwrite(rbindlist(svim_week_list), 
         paste0("outputs/svim_week_list_vacc_week_batch_", batch_id, "_", tstamp(), ".csv"))
}
toc <- Sys.time()
message(paste("Total execution time:", toc - tic))
```

## Case-Triggered Vaccination Strategy

```{r case-trigger-sim, eval=FALSE}
tic <- Sys.time()  # Start timer
dat <- outbreak_data
dat <- dat %>% filter(id_outbreak %in% outbreaks_filtered$id_outbreak)

# Define case thresholds that trigger vaccination
case_triggers <- c(1, 5, seq(10,100,10), seq(150, 300,50))
vacc_covs <- c(0.75)
# vacc_weeks <- c(1, 5, 9, 13)
vacc_weeks <- c(3, 7, 11, 15)
nruns <- 200               # Number of simulation runs
batch_size <- 10            # Number of runs per batch

# Loop through batches of runs for parallel execution
for (batch_id in seq(41, nruns, by = batch_size)) {
  message(paste("Processing batch ID:", batch_id))
  run_ids <- batch_id:(batch_id + batch_size - 1)

  # Create and register parallel cluster
  ncores <- detectCores()
  cl <- makeCluster(getOption("cl.cores", ncores - 1))
  registerDoParallel(cl)

  # Run simulations in parallel
  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = dat,
        ive_data = ive_data,
        parameters = params,
        age_dist = age_dist,
        vacc_week = vacc_weeks,
        case_trigger = case_triggers,  # Case-triggered vaccination
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)
  
  # Save batch results
  saveRDS(rbindlist(svim_week_list),
          paste0("outputs/svim_week_case_trigger_batch_", 
                 batch_id, "_", tstamp(), ".rds"))
}
toc <- Sys.time()
message(paste("Total execution time:", round(toc - tic, digits=2), "minutes"))
```

## Combine Simulation Results

This section combines results from batch processing and organizes them for analysis.

```{r combine-results}
# ---- 1. Time-triggered vaccination results ----
id_begin <- 1
id_end <- 200
batch_size <- 10
batch_ids <- seq(id_begin, id_end, by = batch_size)

svim_df <- data.frame()
for (i in 1:length(batch_ids)) {
  message(paste("Processing batch", i, "of", length(batch_ids)))
  # d <- read_rds(paste0("outputs/svim_week_list_vacc_week_batch_",
  #                      batch_ids[i], "_20250514.rds"))
  d <- fread(paste0("outputs/svim_week_list_vacc_week_batch_",
                       batch_ids[i], "_20250514.csv"))
  svim_df <- rbindlist(list(svim_df, sum_over_week(d)))
  rm(d)  # Free memory
}

# Save combined results
saveRDS(svim_df, 
        paste0("outputs/svim_onset_trigger_sum_over_week_runid_1_200_",
               tstamp(), ".rds"))

# ---- 2. Case-triggered vaccination results ----
id_begin <- 1
id_end <- 200
batch_size <- 10
batch_ids <- seq(id_begin, id_end, by = batch_size)

svim_df <- data.frame()
for (i in 1:length(batch_ids)) {
  message(paste("Processing batch", i, "of", length(batch_ids)))
  d <- read_rds(paste0("outputs/svim_week_case_trigger_batch_",
                   batch_ids[i], "_20250810.rds"))
  svim_df <- rbindlist(list(svim_df, sum_over_week(d, case_trigger = TRUE)))
  rm(d) 
}

# Save combined results
saveRDS(svim_df, 
        paste0("outputs/svim_case_trigger_sum_over_week_runid_1_200_",
               tstamp(), ".rds"))
```

# Cost-Effectiveness Analysis

## Calculate Impact Metrics for Outbreak Onset-Triggered Vaccination

```{r cea-time-trigger}
# Load summarized time-triggered vaccination results
d <- read_rds("outputs/svim_onset_trigger_sum_over_week_runid_1_200_20250716.rds")
# d <- read_rds("outputs/svim_onset_trigger_sum_over_week_runid_1_200_20250520.rds")
# Process each vaccine coverage level
for (vc in VC3) {
  message(paste("Processing results for vaccine coverage =", vc))
  
  # Full cost-effectiveness analysis
  svim_tt_cea <- add_cea_variables(d[vacc_cov == vc], 
                                   mean_age_inf = mean_age_inf)
  svim_tt_cea <- add_cea_results(svim_tt_cea)
  
  # Save detailed results
  saveRDS(svim_tt_cea,
         paste0("outputs/cea_", vc, "_onset_trigger_", 
                tstamp(), ".rds"))
  
  rm(svim_tt_cea)
  
  # # Calculate mean across simulation runs
  # svim_tt_mean <- mean_across_runid(d[vacc_cov == vc])
  # svim_tt_mean_cea <- add_cea_variables(svim_tt_mean, 
  #                                       mean_age_inf = mean_age_inf)
  # svim_tt_mean_cea <- add_cea_results(svim_tt_mean_cea)
  # 
  # # Save average results
  # saveRDS(svim_tt_mean_cea,
  #        paste0("outputs/cea_", vc, "_onset_trigger_mean_across_runid_", 
  #               tstamp(), ".rds"))
  
  # Clean up to save memory
  # rm(svim_tt_cea, svim_tt_mean, svim_tt_mean_cea)
}
rm(d)  # Free memory
```

## Calculate Impact Metrics for Case-Triggered Vaccination

```{r cea-case-trigger}
# Load summarized case-triggered vaccination results
d_ct <- 
  read_rds("outputs/svim_case_trigger_sum_over_week_runid_1_200_20250810.rds")
# d_ct <- 
#   read_rds("outputs/svim_case_trigger_sum_over_week_runid_1_200_20250809.rds")
# d_ct <- 
#   read_rds("outputs/svim_case_trigger_sum_over_week_runid_1_200_20250523.rds")
# d_ct <- 
#   read_rds("outputs/svim_case_trigger_sum_over_week_runid_1_200_20250716.rds")

# Process each vaccine coverage level
# for (vc in VC3) {
for (vc in unique(d_ct$vacc_cov)) {  
  message(paste("Processing results for vaccine coverage =", vc))
  
  # Full cost-effectiveness analysis
  svim_ct_cea <- add_cea_variables(d_ct[vacc_cov == vc], 
                                   mean_age_inf = mean_age_inf)
  svim_ct_cea <- add_cea_results(svim_ct_cea)
  # Save detailed results
  saveRDS(svim_ct_cea,
         paste0("outputs/cea_", vc, "_case_trigger_", 
                tstamp(), ".rds"))
  rm(svim_ct_cea) 
}

# cea_ct <- read_rds("outputs/cea_0.75_case_trigger_20250809.rds")
# cea_ct2 <- read_rds("outputs/cea_0.75_case_trigger_20250810.rds")
# cea_ct <- rbindlist(list(cea_ct, cea_ct2))
# saveRDS(cea_ct,
#          paste0("outputs/cea_", vc, "_case_trigger_", 
#                 tstamp(), ".rds"))
```

# Subset Analysis

This section creates and analyzes subsets of outbreaks to simulate different targeting strategies.

## Create Outbreak Subsets: sample until the target number of outbreaks is reached.

```{r}
outbreaks_filtered |>
  group_by(id_outbreak) |> 
  summarize(year = data.table::year(start_date),
            rural_urban = rural_urban,
            ar = ar_per_10e3,
            cfr = cfr_per_100percent,
            spatial_scale = spatial_scale,
            pop = outbreak_pop,
            duration = duration_weeks,
            size = total_s_ch) -> outbk_dat

summary(outbk_dat$duration)

# outbreaks that are longer than the quantile(outbk_dat$duration, probs=probs)

probs <- c(0, 0.25, 0.5, 0.75)

nsubset <- 200  # Number of simulated subsets
ssize <- c(1:10, seq(20,100,10))
  
for (pr in probs) {
  
  eligible_ids <- outbk_dat %>% 
    filter(duration > quantile(outbk_dat$duration, probs=pr)) %>% 
    pull(id_outbreak) 
  
  dat <- outbk_dat %>% filter(id_outbreak %in% eligible_ids)
  
  for (j in 1:length(ssize)){
    outbreak_subsets <- vector('list', nsubset)
    for (i in 1:nsubset) {
      outbreak_subsets[[i]] <- 
        sample_outbreaks(outbreak_data = dat,
                         target_count = ssize[j],
                         target_population = Inf,
                         sampling_method = "random",
                         seed = i,
                         max_attempts = 1000)
    
    }
    saveRDS(outbreak_subsets, 
          paste0("outputs/outbreak_", pr, "_subsets_size_", ssize[j], "_", tstamp(), ".rds"))
  }
}
```


## Outbreak subsets: sample until the target number of OCV doses is reached

```{r}
# --- 1. Improved Data Preparation ---
# Use standard assignment and be explicit about summarization.
outbk_dat <- outbreaks_filtered |>
  summarize(
    year = first(data.table::year(start_date)),
    rural_urban = first(rural_urban),
    ar = first(ar_per_10e3),
    cfr = first(cfr_per_100percent),
    spatial_scale = first(spatial_scale),
    pop = first(outbreak_pop),
    duration = first(duration_weeks),
    size = first(total_s_ch),
    .by = id_outbreak
  )

# --- 2. Setup Simulation Parameters ---
OCV_doses <- 1000000
nsubset <- 200
probs <- seq(0, 0.9, 0.1)

# --- 3. Run Simulations Functionally and in Parallel ---

# Set up a parallel processing plan. This uses multiple CPU cores.
# The number of workers can be adjusted, e.g., plan(multisession, workers = 4)
plan(multisession)

# The outer map_dfr iterates over each probability cutoff.
# The inner future_map_dfr iterates over each simulation run in parallel.
# The .dfr suffix automatically row-binds the results into a single data frame.
DF <- map_dfr(probs, ~{
  pr <- .x # Assign to a more readable variable name inside the loop
  
  # Filter data for the current probability cutoff
  dat_filtered <- outbk_dat |>
    filter(size > quantile(outbk_dat$size, probs = pr))
  
  # Run nsubset simulations for this cutoff *in parallel*
  future_map_dfr(1:nsubset, ~{
    j <- .x
    set.seed(j) # Ensures reproducibility for each simulation run
    
    # Run the sampling function
    smpl <- sample_outbreaks(
      outbreak_data = dat_filtered,
      target_population = OCV_doses,
      target_count = NULL,
      sampling_method = "random"
    )
    
    # Return a data frame for this single run
    tibble(
      id_outbreak = smpl$sampled_data$id_outbreak,
      id_sample = j,
      size_prob_cutoff = pr
    )
  }, .options = furrr_options(seed = TRUE)) # Ensures seeds are handled correctly in parallel
  
}, .progress = TRUE) # Adds a progress bar for the outer loop

# --- 4. Save the final results ---
# (Assuming tstamp() is a user-defined function for timestamps)
# tstamp <- function() format(Sys.time(), "%Y%m%d_%H%M%S")
saveRDS(DF, paste0("outputs/subsets_", round(OCV_doses/1e6), 
                   "M_size_", tstamp(), ".rds"))

# Clean up parallel workers
plan(sequential)
```

Analysis

```{r}
df <- expand.grid(
  size_cut = seq(0, 1, 0.1),
  vacc_week = 1:15
)
# Ensure consistent results for the example
set.seed(42) 
df$val <- rnorm(nrow(df), 0, 2)

heatmap_plot <- ggplot(df, aes(x = vacc_week, y = size_cut, fill = val)) +
  geom_tile(color = "white", linewidth = 0.5) + # Add white borders for clarity
  
  # Use a diverging color scale for data centered around 0
  scale_fill_gradient2(
    low = "steelblue",      # Color for low values
    mid = "white",      # Color for midpoint
    high = "firebrick",     # Color for high values
    midpoint = 0,         # The center of your data
    name = "Value"
  ) +
  
  # Make sure axes are treated correctly
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 1, 0.2)) +
  scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
  
  # Add informative labels and a clean theme
  labs(
    title = "Impact of Vaccination Timing and Outbreak Prioritization",
    x ="Vaccination Week\n    *slower implementation*", y = NULL 
  ) +
  
  # # --- Add custom annotations for the x-axis ---
  # annotate("text", x = 8, y = -0.15, # Position below the plot
  #          label = "Vaccination Week\n    *slower implementation*", 
  #          hjust = 0.5) +
  annotate("segment", x = 6, xend = 7.8, y=-0.25, yend = -0.25, # Draw the arrow
           arrow = arrow(length = unit(0.2, "cm")), lineend = "round") +
  
  # --- Add custom annotations for the y-axis ---
  annotate("text", x = -3, y = 0.5, # Position left of the plot
           label = "Size Cutoff\n    *prioritizing larger outbreaks*", 
           hjust = 0.5, angle = 90) +
  annotate("segment", x = -2, xend = -2, y = 0.9, yend = 1, # Draw the arrow
           arrow = arrow(length = unit(0.2, "cm")), lineend = "round") +
  
  theme_minimal(base_size = 14) +
  theme(
    plot.title.position = "plot",
    panel.grid = element_blank()
  )

print(heatmap_plot)
```

Analysis

```{r}
# Load outbreak subset definitions
# DF <- readRDS("outputs/random_subsets_by_year_20250605.rds")
cea <- read_rds("outputs/cea_0.6_onset_trigger_20250716.rds")
cea <- rbind(cea, read_rds("outputs/cea_0.75_onset_trigger_20250716.rds"))
cea <- rbind(cea, read_rds("outputs/cea_0.9_onset_trigger_20250716.rds"))

# Initialize parameters
probs <- seq(0, 0.9, 0.1)  # Define probability sequence
DF <- readRDS("outputs/subsets_1M_size_20250921.rds")
dt <- data.table(expand.grid(subset_id = 1:200, prob = probs))

# Function to process each probability level
process_prob <- function(pr, dt_subset, DF, cea) {
  # Filter for specific prob
  dt_prob <- dt_subset[prob == pr]
  
  # Process each subset
  results <- map(dt_prob$subset_id, function(id) {
    # Filter DF for matching conditions
    d <- DF[size_cutoff == pr & id_sample == id]
    
    # Filter cea and sum impacts
    dat <- cea[id_outbreak %in% d$id_outbreak]
    datall <- sum_over_outbreaks_by_runid(dat, ori_occurred = TRUE)
    
    # Add subset metadata
    datall[, subset_id := id]
    return(datall)
  })
  
  # Combine results for this probability
  rbindlist(results)
}

# Process all probabilities using purrr::map
lst <- map(probs, ~process_prob(.x, dt, DF, cea))
names(lst) <- as.character(probs)
df <- rbindlist(lst, use.names=TRUE, fill=TRUE, idcol="size_cutoff")

if (save_rds) {
  saveRDS(df, paste0("outputs/cea_onset_trigger_1m_doses_size_cutoff_",
                     tstamp(), ".rds"))
}  
```

```{r}
df |>
  filter(vacc_cov == 0.75) |>
  group_by(vacc_week, vacc_cov, size_cutoff) |>
  summarize_quantiles(pct_reduc_case) 
```


## Create outbreak subsets for global OCV stockpile use

```{r}
outbreaks_filtered |>
  group_by(id_outbreak) |> 
  summarize(year = data.table::year(start_date),
            rural_urban = rural_urban,
            ar = ar_per_10e3,
            cfr = cfr_per_100percent,
            spatial_scale = spatial_scale,
            pop = outbreak_pop,
            duration = duration_weeks,
            size = total_s_ch) -> outbk_dat

# For 2013 and 2014, we don't have OCV deployment records by event and 
# so outbreaks are randomly sampled to meet the total number of OCVs
nsubset <- 200
# probs <- c(0, 0.25, 0.5, 0.75)
probs <- seq(0, 0.9, 0.1)

for (pr in probs) {
  # # duration
  # eligible_ids <- outbk_dat %>% 
  #   filter(duration > quantile(outbk_dat$duration, probs=pr)) %>% 
  #   pull(id_outbreak) 
  
  eligible_ids <- outbk_dat %>% 
    filter(size > quantile(outbk_dat$size, probs=pr)) %>% 
    pull(id_outbreak) 
    
  dat <- outbk_dat %>% filter(id_outbreak %in% eligible_ids)
  
  DF1 <- data.frame()
  for (j in 1:nsubset) {
    set.seed(j)
    message(paste("Simulation run", j))
    df <- data.frame()
    for (i in 2013:2014) {
      yr <- i  # 2013 = the first year when OCVs were deployed
      # Sample outbreaks in proportion to OCVs deployed that year
      smpl <- sample_outbreaks(
        outbreak_data = filter(dat, year == yr), 
        target_population = ocv_year[year == yr, num_ocv],
        sampling_method = "random")
      
      df <- rbind(df, data.frame(year = smpl$sampled_data$year, 
                                id_outbreak = smpl$sampled_data$id_outbreak))
    }
    df$subset_id <- j
    DF1 <- rbind(DF1, df)
  }
  
  # This simulates targeting outbreaks randomly, matched to historical OCV deployment
  dat2 <- dat[, c("id_outbreak", "pop", "year")]
  names(dat2) <- c("id_outbreak", "population_size", "year")
  
  DF2 <- data.frame()
  
  for (j in 1:nsubset) {
    set.seed(j)
    message(paste("Simulation run", j))
    df <- data.frame()
    for (i in 2015:2020) {
      yr <- i  # 2013 = the first year when OCVs were deployed
      # Sample outbreaks in proportion to OCVs deployed that year
      smpl <- match_ocv_to_outbreaks(
        ocv_deployments = 1000*ori_events[year == yr, dose_delivered_round_one], 
        outbreak_populations = dat2[dat2$year == yr,],
        error_tolerance = 0.2, maximize_coverage = FALSE)
        
      df <- rbind(df, smpl$valid_matches)
    }
  
    DF2 <- rbind(DF2,  cbind(df, data.frame(subset_id = j)))
  }
  
  DF <- rbind(DF1, DF2[, c("year", "id_outbreak", "subset_id")])
  
  saveRDS(DF, paste0("outputs/random_subsets_size_", pr, "_by_year_", tstamp(), ".rds"))
}
```

## Analyis of outbreak onset-triggered vaccination on subsets

```{r subset-time-trigger}
# Load outbreak subset definitions
# DF <- readRDS("outputs/random_subsets_by_year_20250605.rds")
cea <- read_rds("outputs/cea_0.6_onset_trigger_20250716.rds")
cea <- rbind(cea, read_rds("outputs/cea_0.75_onset_trigger_20250716.rds"))
cea <- rbind(cea, read_rds("outputs/cea_0.9_onset_trigger_20250716.rds"))

probs <- c(0, 0.25, 0.5, 0.75)

for (pr in probs) {
  
  DF <- readRDS(paste0("outputs/random_subsets_", pr, "_by_year_20250828.rds"))
  
  # Create analysis combinations (subset × year)
  df <- data.frame(expand.grid(subset_id = 1:200, year_max = 2013:2020))
  # Initialize results list
  lst <- vector('list', nrow(df))
  
  # Process each subset-year combination
  for (i in 1:nrow(df)) {
    # Get outbreaks in this subset up to specified year
    d <- DF[(DF$year <= df$year_max[i] & DF$subset_id == df$subset_id[i]),]
    
    # Extract impact results for these outbreaks
    dat <- cea[id_outbreak %in% d$id_outbreak]
    
    # Sum impact across outbreaks
    datall <- sum_over_outbreaks_by_runid(dat, ori_occurred = T)
    
    # Tag with subset metadata
    datall$subset_id <- df$subset_id[i]
    datall$year_max <- df$year_max[i]
    
    # Store results
    lst[[i]] <- datall
  }
  
  # Combine all results
  dfall <- rbindlist(lst)
  
  # Save results
  saveRDS(dfall, paste0("outputs/cea_stockpile_", pr, "_onset_trigger_by_year_",
                       tstamp(), ".rds"))
}

```



# Subset Analysis

This section creates and analyzes subsets of outbreaks to simulate different targeting strategies.

## Create Outbreak Subsets: sample until the target number of outbreaks is reached.

```{r}
# outbreaks_filtered |>
#   group_by(id_outbreak) |> 
#   summarize(year = data.table::year(start_date),
#             rural_urban = rural_urban,
#             ar = ar_per_10e3,
#             cfr = cfr_per_100percent,
#             spatial_scale = spatial_scale,
#             pop = outbreak_pop,
#             duration = duration_weeks,
#             size = total_s_ch) -> outbk_dat
# 
# summary(outbk_dat$duration)
# 
# # outbreaks that are longer than the quantile(outbk_dat$duration, probs=probs)
# 
# probs <- c(0, 0.25, 0.5, 0.75)
# 
# nsubset <- 200  # Number of simulated subsets
# ssize <- c(1:10, seq(20,100,10))
#   
# for (pr in probs) {
#   
#   eligible_ids <- outbk_dat %>% 
#     filter(duration > quantile(outbk_dat$duration, probs=pr)) %>% 
#     pull(id_outbreak) 
#   
#   dat <- outbk_dat %>% filter(id_outbreak %in% eligible_ids)
#   
#   for (j in 1:length(ssize)){
#     outbreak_subsets <- vector('list', nsubset)
#     for (i in 1:nsubset) {
#       outbreak_subsets[[i]] <- 
#         sample_outbreaks(outbreak_data = dat,
#                          target_count = ssize[j],
#                          target_population = Inf,
#                          sampling_method = "random",
#                          seed = i,
#                          max_attempts = 1000)
#     
#     }
#     saveRDS(outbreak_subsets, 
#           paste0("outputs/outbreak_", pr, "_subsets_size_", ssize[j], "_", tstamp(), ".rds"))
#   }
# }
```


## Create Outbreak Subsets: sample until the target number of outbreaks is reached.

```{r}
outbk_dat <- outbreaks_filtered |>
  summarize(
    year = first(data.table::year(start_date)),
    rural_urban = first(rural_urban),
    ar = first(ar_per_10e3),
    cfr = first(cfr_per_100percent),
    spatial_scale = first(spatial_scale),
    pop = first(outbreak_pop),
    duration = first(duration_weeks),
    size = first(total_s_ch),
    .by = id_outbreak
  )

# --- 2. Setup Simulation Parameters ---
priorities <- c("size", "dur", "AR")
# ocv_doses <- c(1e6, 2e6, 10e6)
n_outbreaks <- c(1, 5, 10, 50)
nsubset <- 200
probs <- seq(0, 0.9, 0.1)

# --- 3. Run Simulations Functionally and in Parallel ---

# Set up a parallel processing plan. This uses multiple CPU cores.
# The number of workers can be adjusted, e.g., plan(multisession, workers = 4)
plan(multisession)

for (priority in priorities) {
  for (n_outbk in n_outbreaks) {
  # The outer map_dfr iterates over each probability cutoff.
  # The inner future_map_dfr iterates over each simulation run in parallel.
  # The .dfr suffix automatically row-binds the results into a single data frame.
  DF <- map_dfr(probs, ~{
    pr <- .x # Assign to a more readable variable name inside the loop
    
    # Filter data for the current probability cutoff
    if (priority  == "size") {
      dat_filtered <- outbk_dat |>
        filter(size > quantile(outbk_dat$size, probs = pr))
    }
    else if (priority  == "dur") {
      dat_filtered <- outbk_dat |>
        filter(duration > quantile(outbk_dat$duration, probs = pr))
    }
    else if (priority  == "AR") {
      dat_filtered <- outbk_dat |>
        filter(ar > quantile(outbk_dat$ar, probs = pr))
    }
    # Run nsubset simulations for this cutoff *in parallel*
    future_map_dfr(1:nsubset, ~{
      j <- .x
      set.seed(j) # Ensures reproducibility for each simulation run
      
      # Run the sampling function
      smpl <- sample_outbreaks(
        outbreak_data = dat_filtered,
        target_population = NULL,
        target_count = n_outbk,
        sampling_method = "random"
      )
      
      # Return a data frame for this single run
      tibble(
        id_outbreak = smpl$sampled_data$id_outbreak,
        id_sample = j,
        prob_cutoff = pr
      )
    }, .options = furrr_options(seed = TRUE)) # Ensures seeds are handled correctly in parallel
    
  }, .progress = TRUE) # Adds a progress bar for the outer loop
  
  # --- 4. Save the final results ---
  saveRDS(DF, paste0("outputs/subsets_num_outbk_", n_outbk, 
                     "_", priority, "_", tstamp(), ".rds"))
  }
}
# Clean up parallel workers
plan(sequential)
```

## Create Outbreak Subsets: sample until the target number of OCV doses is reached.

```{r}
# Outbreak onset-triggered vaccination results
cea_files <- 
  c("outputs/paper/cea_0.75_onset_trigger_20250919.rds")

cea <- map_dfr(cea_files, read_rds)

outbk_dat <- outbreaks_filtered |>
  summarize(
    year = first(data.table::year(start_date)),
    rural_urban = first(rural_urban),
    ar = first(ar_per_10e3),
    cfr = first(cfr_per_100percent),
    spatial_scale = first(spatial_scale),
    pop = first(outbreak_pop),
    duration = first(duration_weeks),
    size = first(total_s_ch),
    .by = id_outbreak
  )

# exclude additionally where gdp is not available
cea %>% filter(is.na(gdp)) %>% pull(id_outbreak) %>% unique() -> ids_na
outbk_dat <- outbk_dat[!(outbk_dat$id_outbreak %in% ids_na), ]

priorities <- c("size", "dur", "AR")
# ocv_doses  <- c(1e5, 2e5, 5e5, 1e6, 2e6, 5e6, 10e6, 20e6, 50e6, 100e6, 200e6)
ocv_doses  <- c(1e5, 2e5, 5e5, 1e6)
nsubset    <- 200
probs      <- seq(0, 0.9, 0.1)
cov        <- 0.75                     # vaccine coverage (fraction)
# 
# priorities <- c("size")
# ocv_doses  <- c(1e6)
# nsubset    <- 10
# probs      <- seq(0, 0.9, 0.1)[1]
# cov        <- 0.75   

set.seed(42)

plan(multisession)

# helper to filter by priority + pr
filter_by_priority <- function(dat, priority, pr) {
  if (priority == "size") {
    thr <- quantile(dat$size, probs = pr, na.rm = TRUE)
    dat %>% filter(size >= thr)
  } else if (priority == "dur") {
    thr <- quantile(dat$duration, probs = pr, na.rm = TRUE)
    dat %>% filter(duration >= thr)
  } else if (priority == "AR") {
    thr <- quantile(dat$ar, probs = pr, na.rm = TRUE)
    dat %>% filter(ar >= thr)
  } else stop("Unknown priority")
}

# safe sampler (identical to yours, but minimal)
run_one_sample <- function(dat, ocv_dose, j) {
  set.seed(j)
  smpl <- sample_outbreaks(
    outbreak_data     = dat,
    target_population = ocv_dose,
    target_count      = NULL,
    sampling_method   = "random"
  )
  tibble(
    id_outbreak = smpl$sampled_data$id_outbreak,
    pop = smpl$sampled_data$pop,
    id_sample   = j
  )
}

for (priority in priorities) {
  for (ocv_dose in ocv_doses) {
    outbreaks_sampled <- map_dfr(probs, function(pr) {
      dat_filtered <- filter_by_priority(outbk_dat, priority, pr)
      # capacity = eligible pop * coverage (single dose assumed)
      cap <- sum(dat_filtered$pop, na.rm = TRUE) * cov
      cap_ok <- is.finite(cap) && cap > 0 && ocv_dose <= cap

      if (!cap_ok) {
        # Skip sampling; return a single status row (no per-replicate noise)
        return(tibble(
          id_outbreak      = NA_character_,
          id_sample        = NA_integer_,
          prob             = pr,
          vacc_priority    = priority,
          ocv_dose         = ocv_dose,
          capacity         = cap,
          vacc_cov         = cov,
          sampled          = FALSE,
          reason           = ifelse(nrow(dat_filtered) == 0L,
                                    "no_outbreaks_after_filter",
                                    "dose_exceeds_capacity")
        ))
      }

      # Run nsubset samples in parallel
      future_map_dfr(
        1:nsubset,
        ~ run_one_sample(dat_filtered, ocv_dose, .x),
        .options = furrr_options(seed = TRUE)
      ) %>%
        mutate(
          prob             = pr,
          vacc_priority    = priority,
          ocv_dose         = ocv_dose,
          capacity         = cap,
          vacc_cov         = cov,
          sampled          = TRUE,
          reason           = NA_character_,
          .before = 1
        )
    }, .progress = TRUE)

    nm <- paste0("outputs/subsets_", 
             round(ocv_dose/1e6), "M_", priority, "_", tstamp(), ".rds")
    
    if (ocv_dose < 1e6) {
       nm <- paste0("outputs/subsets_", 
             round(ocv_dose/1e3), "K_", priority, "_", tstamp(), ".rds")
    }
    saveRDS(outbreaks_sampled, nm)
  }
}

plan(sequential)
# DF %>% 
#   group_by(id_sample, prob) %>% 
#   summarize(total_pop = sum(pop))
```

combine datasets

```{r}
subset_files <- c(
  "outputs/subsets_1M_AR_20251002.rds",
  "outputs/subsets_2M_AR_20251002.rds",
  "outputs/subsets_5M_AR_20251002.rds",
  "outputs/subsets_10M_AR_20251002.rds",
  "outputs/subsets_20M_AR_20251002.rds",
  "outputs/subsets_50M_AR_20251002.rds",
  "outputs/subsets_100M_AR_20251002.rds",
  "outputs/subsets_200M_AR_20251002.rds",
  "outputs/subsets_500M_AR_20251002.rds",
  "outputs/subsets_1000M_AR_20251002.rds",
  "outputs/subsets_1M_dur_20251002.rds",
  "outputs/subsets_2M_dur_20251002.rds",
  "outputs/subsets_5M_dur_20251002.rds",
  "outputs/subsets_10M_dur_20251002.rds",
  "outputs/subsets_20M_dur_20251002.rds",
  "outputs/subsets_50M_dur_20251002.rds",
  "outputs/subsets_100M_dur_20251002.rds",
  "outputs/subsets_200M_dur_20251002.rds",
  "outputs/subsets_500M_dur_20251002.rds",
  "outputs/subsets_1000M_dur_20251002.rds",
  "outputs/subsets_1M_size_20251002.rds",
  "outputs/subsets_2M_size_20251002.rds",
  "outputs/subsets_5M_size_20251002.rds",
  "outputs/subsets_10M_size_20251002.rds",
  "outputs/subsets_20M_size_20251002.rds",
  "outputs/subsets_50M_size_20251002.rds",
  "outputs/subsets_100M_size_20251002.rds",
  "outputs/subsets_200M_size_20251002.rds",
  "outputs/subsets_500M_size_20251002.rds",
  "outputs/subsets_1000M_size_20251002.rds")

subset_data <- map_dfr(subset_files, read_rds)
saveRDS(subset_data, paste0("outputs/paper/subsets_dose_proritization_", tstamp(), ".rds"))
```


Analysis
```{r}
cea_data <- read_rds("outputs/paper/cea_0.75_onset_trigger_20250919.rds")
cea_data %>% filter(vacc_week == 1, runid == 1) -> d
d1 <- d[d$gdp > 0, ]
length(d1$id_outbreak)
d1[is.na(d$net_cost), ]$id_outbreak
d2 <- d1[!complete.cases(d1),]
unique(d2$death_tot)
unique(d2$death_averted_tot)

d2[(death_tot > 0 & death_averted_tot == 0), ]


subset_data <- 
  read_rds("outputs/paper/subsets_dose_proritization_20251002.rds")

# --- 2. Perform Analysis with a Single, Vectorized Operation ---

# Convert input data.frames to data.tables
setDT(subset_data)

subset_dt <- subset_data %>% 
  filter(sampled == TRUE, vacc_priority == "AR", id_sample == 1)

# Perform the operations using data.table

res_list <- vector('list', 200)
vacc_cov <- 0.75

for (i in 1:length(res_list)) {
  
  subset_dt <- subset_data %>% 
    select(prob, vacc_priority, ocv_dose, sampled, id_outbreak, pop, id_sample) %>% 
    filter(sampled == TRUE, !is.na(id_sample), 
           vacc_priority == "AR", id_sample == i)
  
  res_list[[i]] <- 
    cea_data[subset_dt, on = .(id_outbreak), allow.cartesian = TRUE][, .(
      sum_pop = sum(pop),
      sum_vacc_dose = sum(vacc_dose),
      sum_s_ch = sum(s_ch_tot),
      sum_s_ch_averted = sum(s_ch_averted_tot),
      sum_death_averted = sum(death_averted_tot),
      percent_case_averted = 100 * sum(s_ch_averted_tot) / sum(s_ch_tot),
      case_averted_per_1000_OCV = 
        1000 * sum(s_ch_averted_tot) / sum(vacc_dose),
      
      cost_per_daly_averted = sum(net_cost, na.rm=TRUE) / 
        sum(daly_averted, na.rm=TRUE), 
      cost_per_case_averted = sum(net_cost, na.rm=TRUE) / 
        sum(s_ch_averted_tot, na.rm=TRUE),
      n_outbreaks_ori = sum(ori_occurred),
      n_outbreaks_in_subset = .N
    ),

    by = .(prob, id_sample, ocv_dose, runid, vacc_week)
  ]
}
res1 <- rbindlist(res_list)

# Ensure the result is not grouped (data.table automatically drops grouping)
setkey(res1, NULL)
# Ensure res1 is a data.table. This modifies res1 in place for efficiency.
# setDT(res1)

# summary across runid's and id_sample's
res <- res1[, .(
  pop_q500 = median(sum_pop),
  vacc_dose_q500 = median(sum_vacc_dose),
  s_ch_q500 = median(sum_s_ch),
  s_ch_averted_q500 = median(sum_s_ch_averted),
  pca_q500 = median(percent_case_averted),
  pca_q025 = quantile(percent_case_averted, probs=0.025),
  pca_q975 = quantile(percent_case_averted, probs=0.975),
  pca_q250 = quantile(percent_case_averted, probs=0.25),
  pca_q750 = quantile(percent_case_averted, probs=0.75),
  capd_q500 = median(case_averted_per_1000_OCV, na.rm = TRUE),
  capd_q025 = quantile(case_averted_per_1000_OCV, na.rm = TRUE, probs=0.025),
  capd_q975 = quantile(case_averted_per_1000_OCV, na.rm = TRUE, probs=0.975),
  capd_q250 = quantile(case_averted_per_1000_OCV, na.rm = TRUE, probs=0.25),
  capd_q750 = quantile(case_averted_per_1000_OCV, na.rm = TRUE, probs=0.75),
  n_outbreaks_in_subset_q500 = median(n_outbreaks_in_subset)
  ), 
  by = .(prob, ocv_dose, vacc_week)
]


# --- 3. Save the Final Output ---
if (save_rds) {
  saveRDS(res, paste0("outputs/paper/cea_dose_ar_priority_", tstamp(), ".rds"))
}

# View a sample of the results
print(res)
res <- read_rds("outputs/paper/cea_dose_ar_priority_20251002.rds")
res %>% filter(ocv_dose == 10e6, vacc_week %in% c(1,9), prob %in% c(0.0,0.9))
```


Analysis
```{r}
# --- 1. Load and Combine Data Efficiently ---

# Load the cost-effectiveness analysis (CEA) results
# Use map_dfr for a clean, scalable way to load and row-bind multiple files.
cea_files <- c(
  "outputs/paper/cea_0.6_onset_trigger_20250919.rds",
  "outputs/paper/cea_0.75_onset_trigger_20250919.rds",
  "outputs/paper/cea_0.9_onset_trigger_20250919.rds"
)
cea_data <- map_dfr(cea_files, read_rds)

# Load the outbreak subset definitions
# subsets_df <- readRDS("outputs/subsets_1M_size_20250922.rds")
# subsets_df <- readRDS("outputs/subsets_10M_AR_20250922.rds")
# subsets_df <- readRDS("outputs/subsets_10M_size_20250922.rds")
subsets_df <- readRDS("outputs/paper/subsets_10M_dur_20250922.rds")

# --- 2. Perform Analysis with a Single, Vectorized Operation ---

# Convert input data.frames to data.tables
setDT(subsets_df)
setDT(cea_data)

subsets_df <- subsets_df %>% filter(id_sample %in% 1:30)

# Perform the operations using data.table
res1 <- cea_data[subsets_df, on = .(id_outbreak), allow.cartesian = TRUE][
  , .(
    sum_vacc_dose = sum(vacc_dose),
    sum_s_ch = sum(s_ch_tot),
    sum_s_ch_averted = sum(s_ch_averted_tot),
    case_averted_per_1000_OCV = 1000 * sum(s_ch_averted_tot) / sum(vacc_dose),
    n_outbreaks_in_subset = .N
  ),
  by = .(size_prob_cutoff, id_sample, runid, vacc_week, vacc_cov)
]

# Ensure the result is not grouped (data.table automatically drops grouping)
setkey(res1, NULL)
# Ensure res1 is a data.table. This modifies res1 in place for efficiency.
# setDT(res1)

# summary across runid's and id_sample's
res <- res1[, .(
  vacc_dose_q500 = median(sum_vacc_dose),
  s_ch_q500 = median(sum_s_ch),
  s_ch_averted_q500 = median(sum_s_ch_averted),
  case_averted_per_1000_OCV_q500 = median(case_averted_per_1000_OCV, na.rm = TRUE),
  n_outbreaks_in_subset_q500 = median(n_outbreaks_in_subset)
), by = .(size_prob_cutoff, vacc_week, vacc_cov)]


# --- 3. Save the Final Output ---

if (save_rds) {
  # (Assuming tstamp() is a user-defined function for timestamps)
  # tstamp <- function() format(Sys.time(), "%Y%m%d_%H%M%S")
  saveRDS(res,
    paste0("outputs/cea_onset_trigger_10m_doses_dur_cutoff_", tstamp(), ".rds")
  )
}

# View a sample of the results
print(res)
res <- read_rds("outputs/cea_onset_trigger_10m_doses_size_cutoff_20250922.rds")
res %>% filter(vacc_cov == 0.75, vacc_week %in% c(1,9), 
               size_prob_cutoff %in% c(0,0.9))
```


Analysis of median with IQR
```{r}
cea_files <- c("outputs/cea_0.75_onset_trigger_20250919.rds")
cea_data <- map_dfr(cea_files, read_rds)

# Load the outbreak subset definitions
subsets_df <- readRDS("outputs/subsets_10M_AR_20250922.rds")
# subsets_df <- readRDS("outputs/subsets_10M_size_20250922.rds")
# subsets_df <- readRDS("outputs/subsets_10M_dur_20250922.rds")

# --- 2. Perform Analysis with a Single, Vectorized Operation ---

# Convert input data.frames to data.tables
setDT(subsets_df)
setDT(cea_data)

# Perform the operations using data.table
res1 <- cea_data[subsets_df, on = .(id_outbreak), allow.cartesian = TRUE][
  , .(
    sum_vacc_dose = sum(vacc_dose),
    sum_s_ch = sum(s_ch_tot),
    sum_s_ch_averted = sum(s_ch_averted_tot),
    case_averted_per_1000_OCV = 1000 * sum(s_ch_averted_tot) / sum(vacc_dose),
    n_outbreaks_in_subset = .N
  ),
  by = .(size_prob_cutoff, id_sample, runid, vacc_week, vacc_cov)
]

# Ensure the result is not grouped (data.table automatically drops grouping)
setkey(res1, NULL)
# Ensure res1 is a data.table. This modifies res1 in place for efficiency.
# setDT(res1)

# summary across runid's and id_sample's
res <- res1[, .(
  vacc_dose_q500 = median(sum_vacc_dose),
  s_ch_q500 = median(sum_s_ch),
  s_ch_averted_q500 = median(sum_s_ch_averted),
  case_averted_per_1000_OCV_q500 = median(case_averted_per_1000_OCV, na.rm = TRUE),
  case_averted_per_1000_OCV_q250 = quantile(case_averted_per_1000_OCV, probs=0.25, na.rm = TRUE),
  case_averted_per_1000_OCV_q750 = quantile(case_averted_per_1000_OCV, probs=0.75, na.rm = TRUE),
  
  n_outbreaks_in_subset_q500 = median(n_outbreaks_in_subset)
), by = .(size_prob_cutoff, vacc_week, vacc_cov)]

res %>% filter(vacc_cov == 0.75, vacc_week %in% c(1,9), 
               size_prob_cutoff %in% c(0,0.9))
```

Plot

```{r}
# res <- read_rds("outputs/cea_onset_trigger_10m_doses_ar_cutoff_20250922.rds")
# res <- read_rds("outputs/cea_onset_trigger_10m_doses_size_cutoff_20250922.rds")
res <- read_rds("outputs/cea_onset_trigger_10m_doses_dur_cutoff_20250922.rds")
# 1. Create the color palette function (your code was correct here)
myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))

# 3. Build the complete, polished plot
heatmap_plot <- res %>% filter(vacc_cov == 0.75) %>% 
  ggplot(aes(x = vacc_week, y = size_prob_cutoff, 
             fill = case_averted_per_1000_OCV_q500)) +
  geom_tile(color = "white", linewidth = 0.25) +
  # --- CORRECTED COLOR SCALE ---
  # Provide a vector of colors to the 'colours' argument
  scale_fill_gradientn(
    trans = scales::pseudo_log_trans(),
    colours = myPalette(100), # Generate a smooth ramp of 100 colors
    breaks = c(1,2,5,10),
    limits = c(0.2, 15), 
    # Handles values outside the limits gracefully
    # oob = scales::squish, 
    name = "Cases Averted\nper 1,000 OCVs" # Legend title
  ) +
  # --- POLISHED LABELS AND AXES ---
  # scale_y_continuous(
  #   expand = c(0, 0), 
  #   breaks = seq(0, 1, 0.2),
  #   labels = scales::percent_format(accuracy = 1) # Formats 0.2 as "20%"
  # ) +
  scale_y_continuous(
    expand = c(0, 0), 
    breaks = seq(0, 0.8, 0.2),
    labels = c(expression(">="*P[0*"%"]),
               expression(">="*P[20*"%"]),
               expression(">="*P[40*"%"]),
               expression(">="*P[60*"%"]),
               expression(">="*P[80*"%"]))) +
  
  scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
  # 1. Add the arrows
  #    The origin point for all arrows is (x=9, y=0)
  annotate(
    "segment",
    x = 9, y = 0,
    xend = c(9, 1), # End points for the three arrows
    yend = c(0.9, 0),
    arrow = arrow(length = unit(0.3, "cm")),
    color = "black",
    linewidth = 0.4
  ) +

  # 2. Add the four points on top of the arrows
  #    Using shape 21 allows for a fill and a border color
  annotate(
    "point",
    x = c(9), # The four key coordinates
    y = c(0),
    size = 3,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +

  # 3. Add text labels next to the arrows
  #    Coordinates are chosen to be midway along the arrow, with a slight offset
  annotate(
    "text",
    x = c(9.5, 5), # x-positions for the two labels
    y = c(0.5, 0.05), # y-positions for the two labels
    # label = c("Prioritizing high AR outbreaks", "Reducing ORI delay"),
    # label = c("Prioritizing larger outbreaks", "Reducing ORI delay"),
    label = c("Prioritizing longer outbreaks", "Reducing ORI delay"),
    angle = c(90, 0), # Rotate the vertical label
    color = "black",
    # fontface = "bold",
    size = 4,
    hjust = 0.5
  ) +
  
  # labs(
  #   x = "vaccination week<br>(*slower Implementation →*)",
  #   y = "attack rate % cutoff<br>(*prioritizing high AR outbreaks →*)"
  # ) +
  labs(
    x = "Delay in ORI implementation (weeks)",
    y = "Outbreak duration percentile"
    # y = "Outbreak size percentile"
    # y = "Attack rate percentile"
  ) +
  theme_light() +
  theme_tight() +
  theme(
    axis.title.x = element_markdown(),
    axis.title.y = element_markdown(),
    legend.position = "top",
    plot.margin = margin(0,0.03,0,0)
  )
  
 #  # --- CLEAN THEME ---
 # theme_minimal(base_size = 14) +
 #  theme(
 #    plot.title.position = "plot",
 #    panel.grid = element_blank(),
 #    strip.background = element_rect(fill = "grey90", color = "grey90"),
 #    strip.text = element_text(face = "bold"),
 #    
 #    # 3. Tell ggplot to render the axis titles as Markdown
 #    axis.title.x = element_markdown(lineheight = 1, margin = margin(t = 10)),
 #    axis.title.y = element_markdown(),
 #    legend.position = "top"
 #  )
# print(heatmap_plot)

if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig_dur_delay_", tstamp(), ".pdf"),
         heatmap_plot,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw1*fac, units = "mm")
}
```

Plot 2
```{r}
res <- read_rds("outputs/cea_onset_trigger_10m_doses_ar_cutoff_20250922.rds")
# res <- read_rds("outputs/cea_onset_trigger_10m_doses_size_cutoff_20250922.rds")
# res <- read_rds("outputs/cea_onset_trigger_10m_doses_dur_cutoff_20250922.rds")
# 1. Create the color palette function (your code was correct here)
myPalette <- colorRampPalette(brewer.pal(9, "YlOrBr"))

# 3. Build the complete, polished plot
heatmap_plot <- res %>% filter(vacc_cov == 0.75) %>% 
  ggplot(aes(x = vacc_week, y = size_prob_cutoff, 
             fill = case_averted_per_1000_OCV_q500)) +
  geom_tile(color = "white", linewidth = 0.25) +
  # --- CORRECTED COLOR SCALE ---
  # Provide a vector of colors to the 'colours' argument
  scale_fill_gradientn(
    trans = scales::pseudo_log_trans(),
    colours = myPalette(100), # Generate a smooth ramp of 100 colors
    breaks = c(1,2,5,10),
    limits = c(0.2, 15), 
    # Handles values outside the limits gracefully
    # oob = scales::squish, 
    name = "Cases Averted\nper 1,000 OCVs" # Legend title
  ) +
  # --- POLISHED LABELS AND AXES ---
  # scale_y_continuous(
  #   expand = c(0, 0), 
  #   breaks = seq(0, 1, 0.2),
  #   labels = scales::percent_format(accuracy = 1) # Formats 0.2 as "20%"
  # ) +
  scale_y_continuous(
    expand = c(0, 0), 
    breaks = seq(0, 0.8, 0.2),
    labels = c(expression(">="*P[0*"%"]),
               expression(">="*P[20*"%"]),
               expression(">="*P[40*"%"]),
               expression(">="*P[60*"%"]),
               expression(">="*P[80*"%"]))) +
  
  scale_x_continuous(expand = c(0, 0), breaks = seq(1, 15, 2)) +
  # 1. Add the arrows
  #    The origin point for all arrows is (x=9, y=0)
  annotate(
    "segment",
    x = 9, y = 0,
    xend = c(9, 1), # End points for the three arrows
    yend = c(0.9, 0),
    arrow = arrow(length = unit(0.3, "cm")),
    color = "black",
    linewidth = 0.4
  ) +
  # 2. Add the four points on top of the arrows
  #    Using shape 21 allows for a fill and a border color
  annotate(
    "point",
    x = c(9), # The four key coordinates
    y = c(0),
    size = 3,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +

  annotate(
    "segment",
    x = 8, y = 0.5,
    xend = c(8, 1), # End points for the three arrows
    yend = c(0.9, 0.5),
    arrow = arrow(length = unit(0.3, "cm")),
    color = "black",
    linewidth = 0.4,
    linetype = "dashed" # Changed to dashed lines
  ) +
   annotate(
    "point",
    x = c(8), # The four key coordinates
    y = c(0.5),
    size = 3,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  # 3. Add text labels next to the arrows
  #    Coordinates are chosen to be midway along the arrow, with a slight offset
  annotate(
    "text",
    x = c(9.5, 5), # x-positions for the two labels
    y = c(0.5, 0.05), # y-positions for the two labels
    label = c("Prioritizing high AR outbreaks", "Reducing ORI delay"),
    # label = c("Prioritizing larger outbreaks", "Reducing ORI delay"),
    # label = c("Prioritizing longer outbreaks", "Reducing ORI delay"),
    angle = c(90, 0), # Rotate the vertical label
    color = "black",
    # fontface = "bold",
    size = 4,
    hjust = 0.5
  ) +
  
  # labs(
  #   x = "vaccination week<br>(*slower Implementation →*)",
  #   y = "attack rate % cutoff<br>(*prioritizing high AR outbreaks →*)"
  # ) +
  labs(
    x = "Delay in ORI implementation (weeks)",
    # y = "Outbreak duration percentile"
    # y = "Outbreak size percentile"
    y = "Attack rate percentile"
  ) +
  theme_light() +
  theme_tight() +
  theme(
    # axis.title.x = element_markdown(),
    # axis.title.y = element_markdown(),
    legend.position = "top",
    plot.margin = margin(0,0.03,0,0)
  )
  
 #  # --- CLEAN THEME ---
 # theme_minimal(base_size = 14) +
 #  theme(
 #    plot.title.position = "plot",
 #    panel.grid = element_blank(),
 #    strip.background = element_rect(fill = "grey90", color = "grey90"),
 #    strip.text = element_text(face = "bold"),
 #    
 #    # 3. Tell ggplot to render the axis titles as Markdown
 #    axis.title.x = element_markdown(lineheight = 1, margin = margin(t = 10)),
 #    axis.title.y = element_markdown(),
 #    legend.position = "top"
 #  )
# print(heatmap_plot)

if (save_plot) {
  fac <- 1.3
  ggsave(paste0("plots/fig_ar_delay_2_", tstamp(), ".pdf"),
         heatmap_plot,
         device = cairo_pdf, width = fgw1*fac,
         height = fgw1*fac, units = "mm")
}
```

## Create outbreak subsets for global OCV stockpile use

```{r}
outbreaks_filtered |>
  group_by(id_outbreak) |> 
  summarize(year = data.table::year(start_date),
            rural_urban = rural_urban,
            ar = ar_per_10e3,
            cfr = cfr_per_100percent,
            spatial_scale = spatial_scale,
            pop = outbreak_pop,
            duration = duration_weeks,
            size = total_s_ch) -> outbk_dat

# Define cutoff values
ar_cutoffs <- c(1:10)

# Use purrr::map to filter outbreak IDs for each cutoff
ids_outbreaks_ar <- map(ar_cutoffs, function(cutoff) {
  outbk_dat %>%
    filter(ar > cutoff) %>%
    pull(id_outbreak)
}) %>% setNames(as.character(ar_cutoffs))

saveRDS(ids_outbreaks_ar, paste0("outputs/ids_outbreaks_ar_", tstamp(), ".rds"))

summary(outbk_dat$year)
summary(outbk_dat$ar)
summary(outbk_dat$size)
summary(outbk_dat$duration)
summary(outbk_dat$duration)
sum(outbk_dat$ar > 1)
sum(outbk_dat$ar > 2)
sum(outbk_dat$ar > 10)
sum(outbk_dat$size > 300)
sum(outbk_dat$size > 1000)
sum(outbk_dat$size > 2000)
sum(outbk_dat$duration > 13)
sum(outbk_dat$duration > 26)
sum(outbk_dat$duration > 39)
```

```{r}
nsubset <- 200  # Number of simulated subsets

# For 2013 and 2014, we don't have OCV deployment records by event and 
# so outbreaks are randomly sampled to meet the total number of OCVs
# probs <- c(0, 0.25, 0.5, 0.75)
# probs <- seq(0.8, 0.9, 0.1)

ar_cutoffs <- c(1:10)
# for (pr in probs) {
id_sets <- read_rds("outputs/ids_outbreaks_ar_20250923.rds")

for (cutoff in ar_cutoffs) {
  
  eligible_ids <- id_sets[[as.character(cutoff)]]
   
  # eligible_ids <- outbk_dat %>%
  #   filter(duration > quantile(outbk_dat$duration, probs=pr)) %>%
  #   pull(id_outbreak)
  
  # eligible_ids <- outbk_dat %>% 
  #   filter(ar > quantile(outbk_dat$ar, probs=pr)) %>% 
  #   pull(id_outbreak) 
   
  # eligible_ids <- outbk_dat %>% 
  #   filter(size > quantile(outbk_dat$size, probs=pr)) %>% 
  #   pull(id_outbreak) 
    
  dat <- outbk_dat %>% filter(id_outbreak %in% eligible_ids)
  
  DF1 <- data.frame()
  for (j in 1:nsubset) {
    set.seed(j)
    message(paste("Simulation run", j))
    df <- data.frame()
    for (i in 2013:2014) {
      yr <- i  # 2013 = the first year when OCVs were deployed
      # Sample outbreaks in proportion to OCVs deployed that year
      smpl <- sample_outbreaks(
        outbreak_data = filter(dat, year == yr), 
        target_population = ocv_year[year == yr, num_ocv],
        target_count = NULL,
        sampling_method = "random")
      
      df <- rbind(df, data.frame(year = smpl$sampled_data$year, 
                                id_outbreak = smpl$sampled_data$id_outbreak))
    }
    df$subset_id <- j
    DF1 <- rbind(DF1, df)
  }
  
  # This simulates targeting outbreaks randomly, matched to historical OCV deployment
  dat2 <- dat[, c("id_outbreak", "pop", "year")]
  names(dat2) <- c("id_outbreak", "population_size", "year")
  
  DF2 <- data.frame()
  
  for (j in 1:nsubset) {
    set.seed(j)
    message(paste("Simulation run", j))
    df <- data.frame()
    for (i in 2015:2020) {
      yr <- i  # 2013 = the first year when OCVs were deployed
      # Sample outbreaks in proportion to OCVs deployed that year
      smpl <- match_ocv_to_outbreaks(
        ocv_deployments = 1000*ori_events[year == yr, dose_delivered_round_one], 
        outbreak_populations = dat2[dat2$year == yr,],
        error_tolerance = 0.2, maximize_coverage = FALSE)
        
      df <- rbind(df, smpl$valid_matches)
    }
  
    DF2 <- rbind(DF2,  cbind(df, data.frame(subset_id = j)))
  }
  
  DF <- rbind(DF1, DF2[, c("year", "id_outbreak", "subset_id")])

  saveRDS(DF, paste0("outputs/random_subsets_ar_cutoff_", cutoff, "_by_year_", tstamp(), ".rds"))
    
  # saveRDS(DF, paste0("outputs/random_subsets_dur_", pr, "_by_year_", tstamp(), ".rds"))
  # saveRDS(DF, paste0("outputs/random_subsets_ar_", pr, "_by_year_", tstamp(), ".rds"))
  # saveRDS(DF, paste0("outputs/random_subsets_size_", pr, "_by_year_", tstamp(), ".rds"))
}
```

## Analyis of outbreaks onset-triggered vaccination on subsets

```{r subset-time-trigger}
# Load outbreak subset definitions
# DF <- readRDS("outputs/random_subsets_by_year_20250605.rds")
cea <- read_rds("outputs/cea_0.6_onset_trigger_20250716.rds")
cea <- rbind(cea, read_rds("outputs/cea_0.75_onset_trigger_20250716.rds"))
cea <- rbind(cea, read_rds("outputs/cea_0.9_onset_trigger_20250716.rds"))


probs <- c(0, 0.25, 0.5, 0.75)

for (pr in probs) {
  
  DF <- readRDS(paste0("outputs/random_subsets_", pr, "_by_year_20250828.rds"))
  
  # Create analysis combinations (subset × year)
  df <- data.frame(expand.grid(subset_id = 1:200, year_max = 2013:2020))
  # Initialize results list
  lst <- vector('list', nrow(df))
  
  # Process each subset-year combination
  for (i in 1:nrow(df)) {
    # Get outbreaks in this subset up to specified year
    d <- DF[(DF$year <= df$year_max[i] & DF$subset_id == df$subset_id[i]),]
    
    # Extract impact results for these outbreaks
    dat <- cea[id_outbreak %in% d$id_outbreak]
    
    # Sum impact across outbreaks
    datall <- sum_over_outbreaks_by_runid(dat, ori_occurred = T)
    
    # Tag with subset metadata
    datall$subset_id <- df$subset_id[i]
    datall$year_max <- df$year_max[i]
    
    # Store results
    lst[[i]] <- datall
  }
  
  # Combine all results
  dfall <- rbindlist(lst)
  
  # Save results
  saveRDS(dfall, paste0("outputs/cea_stockpile_", pr, "_onset_trigger_by_year_",
                       tstamp(), ".rds"))
}
```


## Analaysis of subsets by AR cutoffs

```{r subset-time-trigger}
# Load outbreak subset definitions
# DF <- readRDS("outputs/random_subsets_by_year_20250605.rds")
cea <- read_rds("outputs/cea_0.75_onset_trigger_20250919.rds")

ar_cutoffs <- 1:3
res_lst <- vector('list', length(ar_cutoffs))

for (ar in ar_cutoffs) {
  
  DF <- readRDS(paste0("outputs/random_subsets_ar_cutoff_", ar, "_by_year_20250923.rds"))
  # Create analysis combinations (subset × year)
  df <- data.frame(expand.grid(subset_id = 1:200, year_max = 2013:2020))
  # Initialize results list
  lst <- vector('list', nrow(df))
  # Process each subset-year combination
  for (i in 1:nrow(df)) {
    # Get outbreaks in this subset up to specified year
    d <- DF[(DF$year <= df$year_max[i] & DF$subset_id == df$subset_id[i]),]
    # Extract impact results for these outbreaks
    dat <- cea[id_outbreak %in% d$id_outbreak]
    
    # Sum impact across outbreaks
    datall <- sum_over_outbreaks_by_runid(dat, ori_occurred = T)
    
    # Tag with subset metadata
    datall$subset_id <- df$subset_id[i]
    datall$year_max <- df$year_max[i]
    
    # Store results
    lst[[i]] <- datall
  }
  # Combine all results
  res_lst[[as.character(ar)]] <- rbindlist(lst)
}  

saveRDS(res_lst, 
        paste0("outputs/cea_stockpile_ar_cutoff_onset_trigger_by_year_",
               tstamp(), ".rds"))
```


## Analysis of outbreak subsets by size

```{r}
probs <- c(0, 0.25, 0.5, 0.75)
ssize <- c(1:10, seq(20,100,10))
  
for (pr in probs) {
  
  DF <- data.frame()
  for (n in ssize) {
    subset <- read_rds(paste0("outputs/outbreak_", pr, "_subsets_size_", 
                              n, "_20250827.rds"))
    for (i in 1:length(subset)) {
      d <- subset[[i]]$sampled_data
      d$subset_id <- i
      d$sample_size <- n
      
      DF <- rbind(DF, d)
    }
  }
  
  # Create analysis combinations (subset × year)
  df <- data.frame(expand.grid(subset_id = 1:200, sample_size = ssize))
  # Initialize results list
  lst <- vector('list', nrow(df))
  
  # Process each subset-year combination
  for (i in 1:nrow(df)) {
    # Get outbreaks in this subset up to specified year
    d <- DF[(DF$sample_size == df$sample_size[i] & DF$subset_id == df$subset_id[i]),]
    # Extract impact results for these outbreaks
    dat <- cea[id_outbreak %in% d$id_outbreak]
    # Sum impact across outbreaks    
    res <- sum_over_outbreaks_by_runid(dat, ori_occurred = TRUE)
    # Tag with subset metadata
    res$subset_id <- df$subset_id[i]
    res$sample_size <- df$sample_size[i]
  
    # Store results
    lst[[i]] <- res
  }
  
  # Combine all results
  res_df <- rbindlist(lst)
  
  # Save results
  saveRDS(res_df, paste0("outputs/cea_onset_trigger_subsets_", pr, "_by_size_",
                       tstamp(), ".rds"))
}
```

## Analyze Case-Triggered Vaccination on Subsets

```{r subset-case-trigger}
# Load outbreak subset definitions
DF <- fread("outputs/case_weighted_subsets_by_year_20250313.csv")

# Load vaccination impact results
cea_ct <- fread("outputs/cea_0.75_case_trigger_20250311.csv")

# Create analysis combinations (subset × year)
df <- data.frame(expand.grid(subset_id = 1:200, 
                             year_max = 2013:2022))

# Initialize results list
lst <- vector('list', nrow(df))

# Process each subset-year combination
for (i in 1:nrow(df)) {
  # Get outbreaks in this subset up to specified year
  d <- DF[(DF$year <= df$year_max[i] & DF$subset_id == df$subset_id[i]),]
  
  # Extract impact results for these outbreaks
  dat <- cea_ct[id_outbreak %in% d$id_outbreak]
  
  # Sum impact across outbreaks (specify case_trigger=TRUE for correct handling)
  datall <- sum_over_outbreaks_by_runid(dat, case_trigger = TRUE)
  
  # Tag with subset metadata
  datall$subset_id <- df$subset_id[i]
  datall$year_max <- df$year_max[i]
  
  # Store results
  lst[[i]] <- datall
}

# Combine all results
dfall <- rbindlist(lst)

# Save results
fwrite(dfall, 
       paste0("outputs/cea_0.75_case_trigger_case_weighted_subsets_by_year_", 
              tstamp(), ".csv"))
```

## Exponential decy fit in percentage cases averted

### Outbreak onset-triggered
```{r}
dat <- left_join(cea[, c("vacc_week", "pct_reduc_case", "id_outbreak",
                         "week_delay_to_vacc_effect", "runid",
                       "vacc_cov")],
               outbk_data_summary[, c("ar_per_10e3","duration_weeks",
                                  "spatial_scale", "rural_urban",
                                  "id_outbreak")],
               by="id_outbreak")


# some outbreaks show no decrease and we treat them separate from fitting a decay model
is_constant_impact <- function(df, y_var = "pct_reduc_case") {
  sd(df[[y_var]], na.rm = TRUE) < 1e-6
}

fit_or_constant <- function(df) {
  if (is_constant_impact(df)) {
    tibble(k_hat = 0, half_life = Inf, constant = TRUE)
  } else {
    fit <- fit_decay_model(df, y_var = "pct_reduc_case", x_var = "vacc_week")
    fit$constant <- FALSE
    fit
  }
}

summary(outbk_data_summary$duration_weeks)
cut_points <- c(5, 9, 14, 21) # min, Q1,Q2,Q3

res <- list()
for (wk_cut in cut_points) {

  eligible_ids <- dat %>%
    filter(vacc_week <= wk_cut) %>%
    group_by(id_outbreak) %>%
    summarise(
      max_delay = max(week_delay_to_vacc_effect, na.rm = TRUE),
      duration = first(duration_weeks)
    ) %>%
    filter(duration > max_delay) %>%
    pull(id_outbreak)

  subset_df <- dat %>%
    filter(vacc_week <= wk_cut, id_outbreak %in% eligible_ids) %>%
    # Group by outbreak characteristics
    group_by(runid, id_outbreak, vacc_cov) %>%
    nest() %>%

    # Apply decay model to each group
    # mutate(fit = map(data, fit_decay_model)) %>%
    mutate(fit = map(data, fit_or_constant)) %>%
    unnest(fit) %>%
    ungroup()

  # Store results for the current cut point
  res[[as.character(wk_cut)]] <- subset_df
}

# res$`15` %>% filter(vacc_cov == 0.9) -> x
# 
# for (i in 1:200) {
#   x %>% filter(runid == i) %>% summarize(n_c = sum(constant)) -> nc
#   if (nc$n_c != 0) cat(i, nc$n_c, "\n")
# }
# 
# res$`15` %>% filter(vacc_cov == 0.9, runid == 2) -> x
# which(x$constant == TRUE)

saveRDS(res, paste0("outputs/fit_decay_model_", tstamp(), ".rds"))
```

### Case-triggered
```{r}
dat_ct <- left_join(cea_ct[, c("vacc_week", "pct_reduc_case", "id_outbreak", 
                      "runid", "case_trigger", "ori_occurred", "vacc_cov",
                      "week_delay_to_vacc_effect")],
               outbk_data_summary[, c("ar_per_10e3", "duration_weeks", 
                                  "spatial_scale", "rural_urban", 
                                  "id_outbreak")],
               by="id_outbreak")


# some outbreaks show no decrease and we treat them separate from fitting a decay model
is_constant_impact <- function(df, y_var = "pct_reduc_case") {
  sd(df[[y_var]], na.rm = TRUE) < 1e-6
}

fit_or_constant <- function(df) {
  if (is_constant_impact(df)) {
    tibble(k_hat = 0, half_life = Inf, constant = TRUE)
  } else {
    fit <- fit_decay_model(df, y_var = "pct_reduc_case", x_var = "vacc_week")
    fit$constant <- FALSE
    fit
  }
}

vacc_wks <- VW_seq[1:5]
# ct_values <- c(1, 5, seq(10, 100, by=10))  # decay across first 50 cases
ct_values <- c(1, 10, 100)
res <- list()

# for (vacc_wk in vacc_wks) {
for (ct_val in ct_values) {
  # Step 1: Identify outbreak IDs that are long enough for the max case_trigger (50)
  eligible_ids <- dat_ct %>%
    filter(vacc_week <= max(vacc_wks), case_trigger == ct_val) %>%
    group_by(id_outbreak) %>%
    summarise(
      max_delay = max(week_delay_to_vacc_effect, na.rm = TRUE),
      duration = first(duration_weeks)
    ) %>%
    filter(duration > max_delay) %>%
    pull(id_outbreak)

    # eligible_ids <- dat %>%
    # filter(vacc_week <= wk_cut) %>%
    # group_by(id_outbreak) %>%
    # summarise(
    #   max_delay = max(week_delay_to_vacc_effect, na.rm = TRUE),
    #   duration = first(duration_weeks)
    # ) %>%
    # filter(duration > max_delay) %>%
    # pull(id_outbreak)
  
  # Step 2: Subset to eligible outbreaks for this vacc_wk and ct range
  subset_df <- dat_ct %>%
    filter(vacc_week %in% vacc_wks,
           id_outbreak %in% eligible_ids,
           case_trigger == ct_val) %>%

    # Step 3: Group by run and outbreak
    group_by(runid, id_outbreak) %>%
    nest() %>%
    # Step 4: Apply decay model across case_trigger
    # mutate(fit = map(data, ~ fit_decay_model(.x, y_var = "pct_reduc_case",
    #                                          x_var = "case_trigger"))) %>%
    mutate(fit = map(data, fit_or_constant)) %>%
    unnest(fit) %>%
    ungroup() %>%

    # Optional: annotate vacc_week
    mutate(case_trigger = ct_val,
           n_obs = map_int(data, nrow))

  # Store in results list
  res[[as.character(ct_val)]] <- subset_df
}
saveRDS(res, paste0("outputs/fit_decay_model_case_trigger_vw_9_", tstamp(), ".rds"))
```

## Build plot summary
```{r}
# If you already define these elsewhere, remove/adjust here
VC3    <- c(0.6, 0.75, 0.9)
VW_seq <- 1:15

build_plot_summary <- function(data, value_col,
                               by = c("vacc_cov", "vacc_week"),
                               round_digits = 3) {
  val <- rlang::enquo(value_col)
  
  data %>%
    group_by(across(all_of(by))) %>%
    summarise(
      mean = mean(!!val, na.rm = TRUE),
      q025 = quantile(!!val, 0.025, na.rm = TRUE),
      q250 = quantile(!!val, 0.25,  na.rm = TRUE),
      q500 = quantile(!!val, 0.50,  na.rm = TRUE),
      q750 = quantile(!!val, 0.75,  na.rm = TRUE),
      q975 = quantile(!!val, 0.975, na.rm = TRUE),
      n    = sum(!is.na(!!val)),
      .groups = "drop"
    ) %>%
    mutate(across(c(mean, q025, q250, q500, q750, q975),
                  ~ round(.x, round_digits)))
}

# =============================================================================
# 1) BUILD & SAVE SUMMARIES  (tiny files you can track in Git instead of cea_ori)
# =============================================================================

# Example: summarize % cases averted by week & coverage
cea_sum <- cea_ori %>%
  filter(vacc_cov %in% VC3, vacc_week %in% VW_seq) %>%
  build_plot_summary(pct_reduc_case, by = c("vacc_cov", "vacc_week"))

saveRDS(cea_sum, paste0("outputs/paper/cea_ori_plot_summary_", tstamp(), ".rds"))
```

## cea summary by attack rate, duration, and size percentile

```{r}
outbk_dat <- outbreaks_filtered |>
  summarize(
    year = first(data.table::year(start_date)),
    rural_urban = first(rural_urban),
    ar = first(ar_per_10e3),
    cfr = first(cfr_per_100percent),
    spatial_scale = first(spatial_scale),
    pop = first(outbreak_pop),
    duration = first(duration_weeks),
    size = first(total_s_ch),
    .by = id_outbreak
  )

# Percentile grid (edit as you like; for full 0–1 include 1)
probs <- seq(0, 0.9, by = 0.1)

# ---- Helpers ----
# 1) IDs in the top p-quantile (e.g., p=0.9 for top 10%)
top_p_ids <- function(data, var, p = 0.9) {
  thr <- stats::quantile(dplyr::pull(data, {{ var }}), probs = p, na.rm = TRUE)
  data %>%
    filter({{ var }} >= thr) %>%
    pull(id_outbreak) %>%
    unique()
}

# lapply(probs, function(p) top_p_ids(outbk_dat, ar, p = p))
ids_list_ar <- map(probs, ~ top_p_ids(outbk_dat, ar, p = .x))
ids_list_dur <- map(probs, ~ top_p_ids(outbk_dat, duration, p = .x))
ids_list_size <- map(probs, ~ top_p_ids(outbk_dat, size, p = .x))
ids_list_cfr <- map(probs, ~ top_p_ids(outbk_dat, cfr, p = .x))
```

```{r}
cea <- read_rds("outputs/paper/cea_0.75_onset_trigger_20250919.rds")
cea_ori <- cea %>% filter(ori_occurred == TRUE)

summarize_by_ids <- function(cea, ids_list, probs, metrics,
                             id_col = "id_outbreak") {
  stopifnot(length(ids_list) == length(probs))

  map2_dfr(probs, ids_list, function(p, ids) {
    if (length(ids) == 0L) {
      # Return an empty tibble with expected columns (so bind_rows works)
      return(tibble(
        prob = numeric(0),
        vacc_cov = numeric(0),
        vacc_week = numeric(0)
      ))
    }

    cea %>%
      filter(.data[[id_col]] %in% ids) %>%
      group_by(vacc_cov, vacc_week) %>%
      summarise(
        across(
          all_of(metrics),
          ~ median(.x, na.rm = TRUE),
          .names = "med_{.col}"
        ),
        # n = n(),                     # how many rows contributed
        n = length(ids),                     # how many rows contributed
        .groups = "drop"
      ) %>%
      mutate(prob = p, .before = 1)
  })
}

cea_ori$daly_averted_per_1000_OCV <- 1000 * cea_ori$daly_averted / 
  cea_ori$vacc_dose

cea_metrics <- c("s_ch_averted_tot", "pct_reduc_case",
                 "case_averted_per_1000_OCV", "daly_averted_per_1000_OCV")  # add 
med_ar   <- summarize_by_ids(cea_ori, ids_list_ar, probs, 
                             metrics = cea_metrics)
med_dur  <- summarize_by_ids(cea_ori, ids_list_dur, probs, 
                             metrics = cea_metrics)
med_size <- summarize_by_ids(cea_ori, ids_list_size, probs, 
                             metrics = cea_metrics)
med_cfr <- summarize_by_ids(cea_ori, ids_list_cfr, probs, 
                            metrics = cea_metrics)
# Optional: label and combine into one tidy table
med_all <- bind_rows(
  mutate(med_ar,   percentile_by = "ar"),
  mutate(med_dur,  percentile_by = "duration"),
  mutate(med_size, percentile_by = "size"),
  mutate(med_cfr, percentile_by = "cfr")) %>%
  relocate(percentile_by, prob, vacc_cov, vacc_week)

med_all_long <- med_all %>%
  pivot_longer(
    cols = starts_with("med_"),
    names_to = "metric",
    values_to = "median_value")

saveRDS(med_all_long, 
        paste0("outputs/paper/summary_impact_delay_prioritization_", 
               tstamp(), ".rds"))
```

