---
title: "Vaccine impact modeling"
format: html
editor_options: 
  code_folding: hide
  chunk_output_type: console
---

## Load packages and source `utils.R`
```{r}
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
source("utils.R") # utility functions
library(conflicted)
conflict_prefer("filter", "dplyr")
# global parameters
VC3 <- c(0.6, 0.75, 0.9) # representative vaccine coverage rates
VC <- VC3[2] # most representative
VW <- 9 # median vaccination week from the GTFCC dashboard
VW_seq <- seq(1,19,2)
PROBS <- c(0.025,0.25,0.5,0.75,0.975)
save_csv <- FALSE
save_rds <- FALSE
```


## Load outbreak datasets

```{r}
# Outbreak summary
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Oct2024/outbreak_summary.csv")))
paste0(length(unique(ds$country)), " countries")
# create 'id_outbreak' variables to link with the other dataset
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# OCV campaign records
docv <- clean_names(fread("data/outbreak_files_Oct2024/ocv_long_dataset.csv"))
ocv_before <- clean_names(fread("data/outbreak_files_Oct2024/ocv_before_outbreak_list.csv"))
ocv_during <- clean_names(fread("data/outbreak_files_Oct2024/ocv_during_outbreak_list.csv"))
ocv_during$start_date_old <- ocv_during$start_date
ocv_during$start_date <- format(as.Date(ocv_during$start_date, format="%m/%d/%y"), "%Y-%m-%d")
sum(docv$ocv_during, na.rm = T) == nrow(ocv_during) 

# the most recent year in which OCV campaign information is available  
YEAR <- max(docv$start_date,na.rm=T)

docv |> 
  pivot_longer(cols=ocv_before_1:ocv_after_1, names_to="ocv_timing") |>
  filter(value == T) |> 
  group_by(ocv_timing) |> 
  count()

# time-series data
ts <- clean_names(readRDS("data/outbreak_files_Oct2024/time_series_outbreak_extraction.rds")) 
ts |>
  group_by(location, outbreak_number) |> # identify unique outbreaks
  mutate(week = row_number(),
         date = tl + 3, # pick the date for the midpoint for each week
         start_date_outbreak = min(tl), 
         end_date_outbreak = max(tr),
         id_outbreak = paste0(location, "-", start_date_outbreak, "-",
                              end_date_outbreak)) |>
  ungroup() -> ts # time series with unique ids

paste0(length(unique(ts$id_outbreak)), " unique outbreaks")
paste0(sum(unique(ts$id_outbreak) %in% ds$id_outbreak), 
       " unique outbreaks matched between time series and outbreak summary")

paste0(length(unique(ts$country)), " countries")

# ids for the outbreaks with confirmed dataset
ts |> 
  group_by(id_outbreak) |> 
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique() -> ids_conf

paste0(length(ids_conf), " confirmed outbreaks")
paste0(length(unique(ts$id_outbreak)) - length(ids_conf), " suspected outbreaks")
# data.table::fwrite(ts, paste0("outputs/outbreak_data_all_", tstamp(), ".csv"))
if (save_csv) {
  fwrite(ts, paste0("outputs/outbreak_data_all_", tstamp(), ".csv"))
}

paste0(sum((unique(ts$id_outbreak)) %in% docv$id_outbreak), 
      " outbreaks were affected by OCVs")

ts_noocv <- ts[!(ts$id_outbreak %in% docv$id_outbreak), ]

if (save_csv) {
  fwrite(ts_noocv, paste0("outputs/outbreak_data_", tstamp(), ".csv"))
}
paste0(sum((unique(ts_noocv$id_outbreak)) %in% docv$id_outbreak), 
      " outbreaks from modeled data were affected by OCVs")

ts_noocv |> 
  group_by(id_outbreak) |> 
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique() -> ids_conf_noocv

paste0(length(ids_conf_noocv), " confirmed outbreaks out of ", length(unique(ts_noocv$id_outbreak)), " outbreaks unaffected by OCVs")
paste0(length(unique(ts_noocv$id_outbreak)) - length(ids_conf_noocv), " suspected outbreaks")

outbreak_data <- as.data.frame(fread("outputs/outbreak_data_20241121.csv"))
outbreak_data_all <- as.data.frame(fread("outputs/outbreak_data_all_20241121.csv"))
```

## Parameter sets
```{r}
## simulation parameters
params <- as.data.frame(fread("outputs/params_20241017.csv"))
# IVE posterior predictive values from Bayesian beta regression (see parameters.qmd)
ive_fit <- read_rds("outputs/ive_yrep_20241010.rds")
ive_data <- ive_fit$yrep
rm(ive_fit)
pop_by_age_wpp <- as.data.frame(fread("outputs/wpp_pop_by_age_20241022.csv"))
life_exp_data <- as.data.frame(fread("outputs/wpp_life_expectancy_20241022.csv"))
gdp_data <- read_xls("data/GDP_WorldBank.xls")
workforce_data <- read_xls("data/Workforce_Worldbank.xls")

## Cost-effectiveness analysis parameters
# Disease symptoms, vaccine costs, etc.
parms <- as.data.frame(fread("data/parameters.csv"))
parms <- filter(parms, Disease == "Cholera")
pr_asymp <- parms[parms$Parameter == "Prop_Asymptomatic", "Value"]
pr_mild <- parms[parms$Parameter == "Prop_Mild", "Value"]
pr_moderate <- parms[parms$Parameter == "Prop_Moderate", "Value"]
pr_severe <- parms[parms$Parameter == "Prop_Severe", "Value"]
dur <- parms[parms$Parameter == "Duration_Illness", "Value"]
wt_asymp <- parms[parms$Parameter == "Disability_Weight_Asymptomatic", "Value"]
wt_mild <- parms[parms$Parameter == "Disability_Weight_Mild", "Value"]
wt_moderate <- parms[parms$Parameter == "Disability_Weight_Moderate", "Value"]
wt_severe <- parms[parms$Parameter == "Disability_Weight_Severe", "Value"]

vacc_cost_per_dose <- parms[parms$Parameter == "Vaccine_Cost", "Value"]
vacc_delivery_cost <- parms[parms$Parameter == "Vaccine_Delivery_Cost", "Value"]
patient_cost_hosp <- parms[parms$Parameter == "Patient_Cost_Hosp", "Value"]
patient_cost_outpt <- parms[parms$Parameter == "Patient_Cost_Outpt", "Value"]
public_cost_hosp <- parms[parms$Parameter == "Public_Cost_Hosp", "Value"]
public_cost_outpt <-parms[parms$Parameter == "Public_Cost_Outpt", "Value"] 
patient_workday_lost <- parms[parms$Parameter == "Pt_Workdays_Lost", "Value"]
caregiver_workday_lost <- parms[parms$Parameter == "Caregiver_Workdays_Lost", "Value"]
mean_age_inf <- parms[parms$Parameter == "Mean_Age_Infection", "Value"]
vacc_shipping_cost <- parms[parms$Parameter == "Vaccine_Shipping_Cost", "Value"]

dose_regimen <- 1
vacc_cost_per_person <- vacc_cost_per_dose * dose_regimen

# assume that reported cases are either moderate or severe cases
pr_tot <- pr_moderate + pr_severe
(coi_per_patient <- 
    pr_moderate / pr_tot * (patient_cost_outpt + public_cost_outpt) + 
    pr_severe / pr_tot * (patient_cost_hosp + public_cost_hosp))

avg_age_inf <- parms[parms$Parameter == "Mean_Age_Infection", "Value"]

## Cost adjustment 2015 to 2023
# consumer price index for cost adjustment (2015 to 2023)
cpi_data <- read_xls("data/API_FP.CPI.TOTL.ZG_DS2_en_excel_v2_9839.xls")
# another gdp_data used previously
gdp_data2 <- read_xls("data/API_NY.GDP.DEFL.KD.ZG_DS2_en_excel_v2_9995.xls")

cpi1 <- cpi_data %>%
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "value")
cpi1$year <- as.numeric(cpi1$year)
cpi1 <- cpi1 %>% filter(year >= 2015 & year <= 2023)

gdp1 <- gdp_data2 %>%
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "value")
gdp1$year <- as.numeric(gdp1$year)
gdp1 <- gdp1 %>% filter(year >= 2015 & year <= 2023)

# Use % change in CPI for sub-Saharan African region, which is available

cpi_percent <- read_xls("data/imf-dm-export-20241028.xls")

cpi_percent <- cpi_percent %>%
  pivot_longer(
    cols = `2015`:`2029`,         
    names_to = "year",             
    values_to = "value"            
  )

cpi_percent$year <- as.numeric(cpi_percent$year)

# year starts at the t + 1 year of reported cost, so if 2015 cost then should use
# % change CPI starting 2016
cpi_percent <- cpi_percent %>% 
  filter(year >= 2016 & year <= 2023)

xi <- (cpi_percent$value/100 + 1)
# cumulative_inflation_factor <- xi[1] * xi[2] * xi[3] * xi[4] * xi[5] * xi[6] * xi[7] * xi[8]
cumulative_inflation_factor <- prod(xi)
patient_cost_hosp_adj <- patient_cost_hosp * cumulative_inflation_factor
patient_cost_outpt_adj <- patient_cost_outpt * cumulative_inflation_factor
public_cost_hosp_adj <- public_cost_hosp * cumulative_inflation_factor
public_cost_outpt_adj <- public_cost_outpt * cumulative_inflation_factor

# vaccine delivery cost was reported in 2016 USD$
cpi_percent <- cpi_percent %>% 
  filter(year >= 2017 & year <= 2023)

xi <- (cpi_percent$value/100 + 1)
# cumulative_inflation_factor <- xi[1] * xi[2] * xi[3] * xi[4] * xi[5] * xi[6] * xi[7]
cumulative_inflation_factor <- prod(xi)
vacc_delivery_cost_adj <- vacc_delivery_cost * cumulative_inflation_factor

patient_cost_hosp <- patient_cost_hosp_adj
patient_cost_outpt <- patient_cost_outpt_adj
public_cost_hosp <- public_cost_hosp_adj
public_cost_outpt <- public_cost_outpt_adj
vacc_delivery_cost <- vacc_delivery_cost_adj

## Life expectancy and other cost data
# the following gdp is different from the overall sub-Saharan Africa gdp
# gdp_data2 <- read_xls("data/GDP_WorldBank.xls")
names(life_exp_data)[names(life_exp_data) == "ISO3 Alpha-code"] <- "country"
names(life_exp_data)[names(life_exp_data) == "Year"] <- "year"
names(workforce_data)[names(workforce_data) == "Country Code"] <- "country"

workforce_long <- 
  workforce_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols=`2001`:`2023`, names_to="year", values_to="pct_workforce")
workforce_long$year <- as.integer(workforce_long$year)

names(gdp_data)[names(gdp_data) == "Country Code"] <- "country"
gdp_long <- 
  gdp_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols=`2001`:`2023`, names_to="year", values_to="gdp")
gdp_long$year <- as.integer(gdp_long$year)
```

## Summarize vaccine impact

### Across study variation across or among outbreaks 
```{r}
nruns <- 200
batch_size <- 5
batch_ids <- seq(1, nruns, by=batch_size)
save_rds <- TRUE

for (vc in VC3) {
  svim_df <- data.table()
  cat("vacc_cov =" , vc, "\n")
  for (i in 1:length(batch_ids)) {
    cat("batch_id =" , batch_ids[i], "\n")
    d <- rbindlist(read_rds(
      paste0("G:/My Drive/Simulation/svim_week_list_vacc_week_batch_", batch_ids[i],
             "_20241209.rds")))
    d <- d[vacc_cov == vc]
    svim_df <- rbindlist(list(svim_df, d)) 
    rm(d)
  }
  
  svim <- set_svim_raw_res(svim_df) # remove OCV-affected outbreaks, add population size, ..
  rm(svim_df)

  svim_sum_across_week_by_runid <- svim_sum_across_week(svim, case_trigger=FALSE)
  if (save_rds) {
    saveRDS(svim, 
            paste0("outputs/svim_", vc, "_time_trigger_weekly_by_runid_", tstamp(), ".rds"))
  }
  rm(svim)
  
  if (save_rds) {
    saveRDS(svim_sum_across_week_by_runid, 
            paste0("outputs/svim_", vc, "_time_trigger_sum_across_week_by_runid_", tstamp(), ".rds"))
    
  }
  
  # svim_sum_across_week_by_runid <-
  #   readRDS("outputs/svim_time_trigger_sum_across_week_by_runid_20241220.rds")
  
  svim_mean <- svim_mean_across_runid(svim_sum_across_week_by_runid)
  
  # cost-effectiveness analysis
  svim_mean_cea <- add_cea_variables(svim_mean)
  svim_mean_cea <- add_cea_results(svim_mean_cea)
  
  svim_sum_cea <- add_cea_variables(svim_sum_across_week_by_runid)
  svim_sum_cea <- add_cea_results(svim_sum_cea)
  
  if (save_rds) {
    saveRDS(svim_mean_cea,
      paste0("outputs/svim_", vc, "_time_trigger_mean_cea_across_runid_", 
             tstamp(), ".rds"))
    saveRDS(svim_sum_cea,
     paste0("outputs/svim_", vc, "_time_trigger_sum_cea_across_runid_", 
            tstamp(), ".rds"))
  }
}
rm(svim_sum_across_week_by_runid)
rm(svim_mean)
rm(svim_mean_cea)
rm(svim_sum_cea)

# case-trigger vaccination
for (vc in VC3) {
  svim_df <- data.table()
  cat("vacc_cov =" , vc, "\n")
  for (i in 1:length(batch_ids)) {
    cat("batch_id =" , batch_ids[i], "\n")
    d <- rbindlist(read_rds(
      paste0("G:/My Drive/Simulation/svim_case_trigger_week_list_batch_", batch_ids[i],
             "_20241211.rds")))
    d <- d[vacc_cov == vc]
    svim_df <- rbindlist(list(svim_df, d)) 
    rm(d)
  }
  
  svim <- set_svim_raw_res(svim_df) # remove OCV-affected outbreaks, add population size, ..
  rm(svim_df)
  
  svim_ct_sum_across_week_by_runid <- svim_sum_across_week(svim, case_trigger=TRUE)
  if (save_rds) {
    saveRDS(svim, 
            paste0("outputs/svim_", vc, "_case_trigger_weekly_by_runid_", 
                   tstamp(), ".rds"))
  }
  rm(svim)
  
  if (save_rds) {
   saveRDS(svim_ct_sum_across_week_by_runid, 
          paste0("outputs/svim_", vc, "_case_trigger_sum_across_week_by_runid_", 
                 tstamp(), ".rds"))
  }
  # svim_ct_sum_across_week_by_runid <-
  #   readRDS("outputs/svim_case_trigger_sum_across_week_by_runid_20241220.rds")
  
  svim_ct_mean <-
    svim_mean_across_runid(svim_ct_sum_across_week_by_runid, case_trigger=TRUE)
  
  svim_ct_mean <- add_summary_stats(svim_ct_mean, case_trigger=TRUE)
  
  if (save_rds) {
    saveRDS(svim_ct_mean,
      paste0("outputs/svim_", vc, "_case_trigger_mean_across_runid_", tstamp(), ".rds"))
  }
  
  # cost-effectiveness analysis
  svim_ct_sum_cea <- add_cea_variables(svim_ct_sum_across_week_by_runid)
  rm(svim_ct_sum_across_week_by_runid)
  svim_ct_sum_cea <- add_cea_results(svim_ct_sum_cea)
  svim_ct_mean_cea <- add_cea_variables(svim_ct_mean)
  rm(svim_ct_mean)
  svim_ct_mean_cea <- add_cea_results(svim_ct_mean_cea)
  
  if (save_rds) {
    saveRDS(svim_ct_sum_cea,
      paste0("outputs/svim_", vc, "_case_trigger_sum_cea_across_runid_", tstamp(), ".rds"))
    saveRDS(svim_ct_mean_cea,
      paste0("outputs/svim_", vc, "_case_trigger_mean_cea_across_runid_", tstamp(), ".rds"))
  }
  rm(svim_ct_sum_cea)
  rm(svim_ct_mean_cea)
}

# pre-emptive vaccination
svim_pe_df <- fread("outputs/svim_preemptive_raw_20241203.csv")
svim_pe <- set_svim_raw_res(svim_pe_df)
svim_pe_sum_across_week_by_runid <- svim_sum_across_week(svim_pe)
if (save_csv){
  fwrite(svim_pe_sum_across_week_by_runid,
         paste0("outputs/svim_pe_sum_across_week_by_runid_", tstamp(), ".csv"))
}
svim_pe_mean <- svim_mean_across_runid(svim_pe_sum_across_week_by_runid)
svim_pe_mean <- add_summary_stats(svim_pe_mean)
if (save_csv) {
  fwrite(svim_pe_mean,
    paste0("outputs/svim_pe_mean_across_runid_", tstamp(), ".csv"))
}
svim_pe_mean_cea <- add_cea_variables(svim_pe_mean)
svim_pe_mean_cea <- add_cea_results(svim_pe_mean_cea)
if (save_csv) {
  fwrite(svim_pe_mean_cea,
    paste0("outputs/svim_pe_mean_cea_across_runid_", tstamp(), ".csv"))
}

svim_pe_sum_across_week_by_runid |> 
  filter(vacc_cov %in% VC3) |>
  group_by(vacc_cov) |>
  summarize(mean_pcr = mean(pct_reduc_case),
            q500_pcr = quantile(pct_reduc_case, probs=0.5),
            q025_pcr = quantile(pct_reduc_case, probs=0.025),
            q975_pcr = quantile(pct_reduc_case, probs=0.975))

svim_pe_mean |> 
  filter(vacc_cov %in% VC3) |>
  group_by(vacc_cov) |>
  summarize(mean_pcr = mean(pct_reduc_case),
            q500_pcr = quantile(pct_reduc_case, probs=0.5),
            q025_pcr = quantile(pct_reduc_case, probs=0.025),
            q975_pcr = quantile(pct_reduc_case, probs=0.975))

# change 1000_OCV -> 1000_OCV
# vs <- c(0.6, 0.75, 0.9)
# for (v in vs) {
#   fn1 <- paste0("outputs/svim_", v, "_time_trigger_sum_cea_across_runid_20241222.rds")
#   d1 <- readRDS(fn1)
#   nms1 <- names(d1)
#   nms1[c(45,46)] <- c("case_averted_per_1000_OCV", "death_averted_per_1000_OCV")
#   names(d1) <- nms1
#   saveRDS(d1, fn1)
#   
#   fn2 <- paste0("outputs/svim_", v, "_time_trigger_mean_cea_across_runid_20241222.rds")
#   d2 <- readRDS(fn2)
#   nms2 <- names(d2)
#   nms2[c(44,45)] <- c("case_averted_per_1000_OCV", "death_averted_per_1000_OCV")
#   names(d2) <- nms2
#   saveRDS(d2, fn2)
# }
# 
# for (v in vs) {
#   fn1 <- paste0("outputs/svim_", v, "_case_trigger_sum_cea_across_runid_20241223.rds")
#   d1 <- readRDS(fn1)
#   nms1 <- names(d1)
#   if (nms1[45] == "case_averted_per_1000_OCVs" & 
#       nms1[46] == "death_averted_per_1000_OCVs"){
#     cat("45 & 46\n")
#     nms1[c(45,46)] <- c("case_averted_per_1000_OCV", "death_averted_per_1000_OCV")
#   }
#   names(d1) <- nms1
#   saveRDS(d1, fn1)
#   
#   fn2 <- paste0("outputs/svim_", v, "_case_trigger_mean_cea_across_runid_20241223.rds")
#   d2 <- readRDS(fn2)
#   nms2 <- names(d2)
#   if (nms2[63] == "case_averted_per_1000_OCVs" & nms2[64] == "death_averted_per_1000_OCVs"){
#     cat("63 & 64\n")
#     nms2[c(63,64)] <- c("case_averted_per_1000_OCV", "death_averted_per_1000_OCV")
#   }
#   names(d2) <- nms2
#   saveRDS(d2, fn2)
# }
```

### All outbreaks collectively

```{r}
svim_df <- fread("outputs/svim_raw_20241113.csv")
# > nrow(svim_df)
# [1] 136516800
svim_df <- svim_df[start_date_outbreak <= YEAR]
# > nrow(svim_df)
# [1] 129408000
svim <- set_svim_raw_res(svim_df)
rm(svim_df)
a <- svim_sum_across_week_all_outbreaks(svim)
as <- add_summary_stats(a)
ace <- add_cea_variables(as)
acea <- add_cea_results_all_outbreaks(ace)
#a_ <- svim_sum_across_week(svim) # the above version a lot faster
svim_all_by_runid <- svim_sum_across_outbreaks_by_runid(acea)

if (save_file) {
  # fwrite(svim_all_by_runid, 
  #        paste0("outputs/svim_all_outbreaks_by_runid_", tstamp(), ".csv"))
  fwrite(svim_all_by_runid,
    paste0("outputs/svim_all_outbreaks_cea_across_runid_", tstamp(), ".csv"))
}
```

## Outbreak subsets

```{r}
# number of doses
dat <- fread("data/OCV_Use_GTFCC_Dashboard.csv")
dat <- clean_names(dat)
# pop_size <- sum(dat[(type=="Reactive" & year < 2024), dose])
pop_size <- sum(dat[(type=="Reactive" & year <= data.table::year(YEAR)), dose])
svim <- fread("outputs/svim_mean_across_runid_20241122.csv")
d <- svim[(vacc_cov == VC & vacc_week == VW & start_date <= YEAR),
          c("id_outbreak", "population", "duration_weeks", "s_ch_tot", "ar_per_10e3")]

names(d) <- c("id_outbreak", "population_size", 
              "duration", "cumulative_cases", "attack_rate")

methods <- c("random", "attack_rate", "duration", "cases")
outbreak_subsets <- vector('list', 4)
names(outbreak_subsets) <- methods
target_pop <- pop_size

samp_size <- 200
max_attempts <- 1e5 # parameters used to limit the number of times to search outbreaks
set.seed(42)

for(i in 1:length(methods)) {
  cat("i =", methods[i], "\n")
  subset <- vector("list", samp_size)
  for (j in 1:length(subset)) {
    subset[[j]] <-
      sample_outbreaks(
        outbreak_data = d, 
        target_population = target_pop,
        sampling_method = methods[i], 
        max_attempts = max_attempts)
  }
  outbreak_subsets[[i]] <- subset
}

saveRDS(outbreak_subsets, paste0("outputs/svim_outbreak_subsets_ids_",
                                 tstamp(), ".rds"))
```

### Subsets into a single dataframe

```{r}
samp_size <- 200
methods <- c("random", "attack_rate", "duration", "cases")
lst <- vector("list", samp_size * length(methods))
for (i in 1:length(methods)) {
  for (j in 1:samp_size){
    filter(svim, id_outbreak %in% 
             outbreak_subsets[[methods[i]]][[j]]$sampled_data$id_outbreak) -> d
    d$subset_method <- methods[i]
    d$subset_id <- j
    lst[[(i-1)*samp_size + j]] <- d
  }
}
df <- rbindlist(lst)
if (save_csv) {
  fwrite(df, paste0("outputs/svim_outbreak_subsets_", tstamp(), ".csv"))
}
```

### Subset analysis - time-triggered

```{r}
# subset_ids <- readRDS("outputs/svim_outbreak_subsets_ids_20241126.rds")
subset_ids <- readRDS("outputs/svim_outbreak_subsets_ids_20250225.rds")
svim_df <- fread("outputs/svim_raw_20241113.csv")
# create svim_df subsets
nms <- c("random", "cases", "attack_rate", "duration")
df <- data.frame()
# mean across runid's and later compute statistics across subset id's
for (nm in nms) {
  for (i in 1:200) {
    cat("i = ", i,"\n")
    if (nm == "random") {
      d <- svim_df[id_outbreak %in% subset_ids$random[[i]]$sampled_data$id_outbreak]
    }
    else if(nm == "attack_rate"){
      d <- svim_df[id_outbreak %in% subset_ids$attack_rate[[i]]$sampled_data$id_outbreak]
    }
    else if(nm == "duration") {
      d <- svim_df[id_outbreak %in% subset_ids$duration[[i]]$sampled_data$id_outbreak]  
    }
    else if(nm == "cases") {
      d <- svim_df[id_outbreak %in% subset_ids$cases[[i]]$sampled_data$id_outbreak]  
    }
    
    svim <- set_svim_raw_res(d)
    rm(d)
    a <- svim_sum_across_week_all_outbreaks(svim)
    rm(svim)
    as <- add_summary_stats(a)
    rm(a)
    ace <- add_cea_variables(as)
    rm(as)
    acea <- add_cea_results_all_outbreaks(ace)
    rm(ace)
    # we could either pick one runid or mean across multiple ids
    svim_subset_by_runid <- svim_sum_across_outbreaks_by_runid(acea)
    
    svim_mean <- 
      svim_subset_by_runid[, lapply(.SD, mean), 
                           by=.(vacc_cov, vacc_week),
                           .SDcols=population:cost_per_death_averted]
    
    rm(svim_subset_by_runid)
    svim_mean$subset_id <- i
    df <- rbind(df, svim_mean)
  }
  fwrite(df, paste0("outputs/svim_mean_", nm, "_subset_ids_", tstamp(), ".csv"))
}
```


## Vaccine impact simulation

Parallel by batch `run_vacc_impact_outbreak_weekly`

```{r}
tic <- Sys.time()

library(doParallel)
# UN data on country- and year-specific age distribution (< 5 yo and 5+ yo)
# pop_by_age_wpp <- as.data.frame(fread("outputs/wpp_pop_by_age_20241017.csv"))
vacc_covs <- seq(0.6, 0.95, by=0.05)
vacc_weeks <- seq(1, 20, by=1) # vaccination can happen anytime during week 1 and week 20
outbreak_ids <- unique(outbreak_data$id_outbreak)
nruns <- 30 # number of simulation runs
batch_size <- 5

for (batch_id in seq(1, nruns, by = batch_size)) {
  cat("batch id: ", batch_id, "\n")
  run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in batches of batch_size

  cl <- makeCluster(getOption("cl.cores", detectCores() - 1))
  registerDoParallel(cl)

  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = outbreak_data,
        ive_data = ive_data,
        parameters = params,
        age_dist = pop_by_age_wpp,
        vacc_week = vacc_weeks,
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)
  saveRDS(svim_week_list, paste0("outputs/svim_week_list_vacc_week_batch_",
                                 batch_id, "_", tstamp(), ".rds"))
}
```

Parallel case_trigger by batch

```{r, eval=FALSE}
library(doParallel)
case_triggers <- c(seq(110,190, by=10), seq(200,300, by=10))
nruns <- 30
batch_size <- 5
vacc_covs <- seq(0.6, 0.95, by=0.05)
outbreak_ids <- unique(outbreak_data_all$id_outbreak)
nruns <- 30 # number of simulation runs
batch_size <- 5

cl <- makeCluster(getOption("cl.cores", detectCores() - 1))
registerDoParallel(cl)

for (batch_id in seq(1, nruns, by = batch_size)) {

  cat("batch id: ", batch_id, "\n")
  run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in batches of batch_size

  cl <- makeCluster(getOption("cl.cores", detectCores() - 4))
  registerDoParallel(cl)

  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = outbreak_data_all,
        ive_data = ive_data,
        parameters = params,
        age_dist = pop_by_age_wpp,
        case_trigger = case_triggers,
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)
  saveRDS(svim_week_list, paste0("outputs/svim_case_trigger_week_list_batch_",
                                 batch_id, "_", tstamp(), ".rds"))
}
```

Parallel preemptive vaccination `run_vacc_impact_outbreak_weekly`

```{r, eval=FALSE}
library(doParallel)

tic <- Sys.time()
# UN data on country- and year-specific age distribution (< 5 yo and 5+ yo)
# pop_by_age_wpp <- as.data.frame(fread("outputs/wpp_pop_by_age_20241022.csv"))
vacc_covs <- seq(0.6, 0.95, by=0.05)
# maximum delay possible
# max(params$dur_campaign)/7 + max(params$delay_vacc_effect)/7
vacc_weeks <- -52 # some random large number ensure than pre-emptive vaccination
nruns <- 200 # number of simulation runs

cl <- makeCluster(getOption("cl.cores", detectCores() - 6))
registerDoParallel(cl)

svim_week_list <-
  foreach(i = 1:nruns, .packages = c("data.table")) %dopar% {
    run_vacc_impact_outbreak_weekly(
      outbreak_data = outbreak_data,
      ive_data = ive_data,
      parameters = params,
      age_dist = pop_by_age_wpp,
      vacc_week = vacc_weeks,
      vacc_cov = vacc_covs,
      runid = i)
  }
stopCluster(cl)

Sys.time() - tic
# Time difference of 36.02764 mins
saveRDS(svim_week_list, paste0("outputs/svim_preemptive_list_", tstamp(), ".rds"))
```

## Summarize the simulation results

Combine the results from results from separate files across which runid's are different.
```{r}
# svim_week_list <- read_rds("outputs/svim_week_list_vacc_week_20241010.rds")
nruns <- 30
batch_size <- 5
batch_ids <- seq(1, nruns, by=batch_size)

lst <- vector('list', length(batch_ids))
for ( i in 1:length(batch_ids)) {
  lst[[i]] <- rbindlist(read_rds(
    paste0("outputs/svim_week_list_vacc_week_batch_", batch_ids[i],
           "_20241113.rds")))
}
svim_df <- rbindlist(lst)
fwrite(svim_df, paste0("outputs/svim_raw_", tstamp(), ".csv"))

lst <- vector('list', length(batch_ids))
for ( i in 1:length(batch_ids)) {
  lst[[i]] <- rbindlist(read_rds(
    paste0("outputs/svim_case_trigger_week_list_batch_", batch_ids[i],
           "_20241123.rds")))
}

svim_df <- rbindlist(lst)
fwrite(svim_df, paste0("outputs/svim_case_trigger_raw_", tstamp(), ".csv"))

# lst <- read_rds("outputs/svim_preemptive_list_20241203.rds")
svim_df <- rbindlist(lst)
# table(svim_df$ori_occurred)
# table(svim_df$week_delay_to_vacc_effect)
fwrite(svim_df, paste0("outputs/svim_preemptive_raw_", tstamp(), ".csv"))
```
