---
title: "Cholera Vaccine Impact Modeling: Simulations"
author: "Jong-Hoon Kim"
date: "2025-10-31"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

# Introduction

This document presents a comprehensive analysis of the impact of oral cholera vaccine (OCV) using outbreak data from sub-Saharan Africa (2010-2022). Two reactive vaccination strategies are evaluated:

 - **Outbreak onset-triggered** strategy: Outbreak response immunization (ORI) is triggered at outbreak onset (week 1) but implemented after a variable delay.

 - **Case-triggered** strategy: ORI is triggered once cumulative reported cases reach defined thresholds, with an additional implementation delay thereafter.

We evaluate these strategies across a range of vaccine coverage levels (60%, 75%, and 90%) and implementation delays (in 1 - 15 weeks). ORI impacts are quantified by three complementary metrics:

 - *Impact*: Cases, deaths, and disability-adjusted life years (DALYs) averted

 - *Efficiency*: impact per 1,000 OCV doses (or per 1,000 people vaccinated)

 - *Cost effectiveness*: cost (2023 US$) per unit of impact

The analysis proceeds through data preparation, simulation, and results evaluation. Results are further stratified by outbreak characteristics—attack rate, duration, and size—to explore how ORI performance varies across diverse epidemiologic contexts.

Finally, we examine outcomes in subsets of outbreaks corresponding to different target OCV stockpile volumes, reflecting realistic vaccine deployment scenarios over time. Outbreaks are sampled either across the full dataset or within strata defined by attack rate, duration, or size to evaluate the effects of prioritization strategies.

# Setup

## Load Packages and Utility Functions

```{r setup}
#| warning: false
#| message: false
# Load utility functions
source("R/utils.R")
# Core packages
library(data.table)
library(magrittr)
library(tidyverse)
library(readxl)
library(janitor)
library(ggplot2)
library(ggridges)
library(patchwork)
library(conflicted)
library(scales)
library(viridis)
library(doParallel)  # For parallel processing
library(MASS)
library(purrr)
library(furrr) # For parallel processing
library(future)
library(future.apply)
# Resolve conflicts in favor of dplyr functions
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflicts_prefer(data.table::first)
conflicts_prefer(data.table::year)

# Global parameters
VC3 <- c(0.6, 0.75, 0.9)  # Standard vaccine coverage values (60%, 75%, 90%)
VC <- VC3[2]  # Reference coverage (75%) 
VW <- 9       # Reference vaccination week (median from GTFCC data)
VW_seq <- seq(1, 19, 2)       # Sequence of vaccination weeks to evaluate
PROBS <- c(0.025, 0.25, 0.5, 0.75, 0.975)  # Probability quantiles for uncertainty
YEAR <- as.Date("2020-12-28") # Most recent date with OCV campaign data

# Flags for output control
save_csv <- FALSE             # Whether to save results as CSV
save_rds <- FALSE             # Whether to save results as RDS
save_plot <- FALSE            # Whether to save generated plots
save_table <- FALSE           # Whether to save generated tables
```

# Parameter Preparation

## Parameter samples for simulation

```{r params-sample}
# # Load parameter definitions
# dat_params <- data.table::fread("data/paper/parameters.csv")
# 
# # Draw parameter samples for multiple simulation runs
# nruns <- 200  # Number of simulation runs
# params <- draw_parameter_samples(nruns = nruns, parameter_data = dat_params)
# 
# # Save parameter samples
# if(save_csv) {
#   data.table::fwrite(params, paste0("outputs/paper/params_", tstamp(), ".csv"))
# }

params <- fread("outputs/paper/params_20260612.csv")
```

## Vaccine coverage data

We selected vaccine coverage levels to broadly reflect the estimated (rather than administrative) coverage achieved during the first round of OCV campaigns in outbreak and humanitarian settings. Reported estimates range from approximately 0.63 in humanitarian crises to 0.85 in outbreak settings. Accordingly, we considered three representative values: 60% (pessimistic), 75% (typical or most representative), and 90% (optimistic).


```{r vacc-coverage}
# Vaccine coverage data from Table 2 in Pezzoli et al. 2020
vacc_cov_data <- 
  data.frame(setting = c("Endemic", "Humanitarian crisis", "Outbreak"),
           admin_cov_first_dose = c(93.6, 82.9, 98.8)/100,
           admin_cov_second_dose = c(92.4, 82.0, 93.0)/100, 
           estimate_cov_at_least_one = c(85.1, 73.2, 86.3)/100)

# Calculate first-round coverage for different settings
vacc_cov_data |> 
  filter(setting != "Endemic") |>
  group_by(setting) |>
  summarize(
    frac_effect = estimate_cov_at_least_one / 
                 (1 - (1 - admin_cov_first_dose) * (1 - admin_cov_second_dose)),
    vacc_cov_r1 = admin_cov_first_dose * frac_effect
  )
```

## Indirect Vaccine Effectiveness

```{r ive-model}
# Load previously fitted Bayesian model for indirect vaccine effectiveness (IVE)

ive_fit <- readRDS("outputs/paper/ive_yhat_yrep_20250919.rds")
ive_data <- ive_fit$yhat
rm(ive_fit)

# Example of how we compute vaccine impact
# The impact of vaccination at coverage proportion π is:
# P_{i,π} = 1-[π(1-DVE_i)(1-IVE) + (1-π)(1-IVE)]
# where DVE_i is direct vaccine efficacy for age group i
# and IVE is indirect vaccine effectiveness
```

# Load outbreak data

## Outbreak and OCV campaign data

Around 38m and 155m doses have been deployed 2013 thorough 2020 and 2024, 
respectively.

```{r load-data}
#| warning: false
outbreak_data_all <- 
  as.data.frame(fread("outputs/paper/outbreak_data_all_20250228.csv"))

# Load outbreaks not affected by OCVs (for modeling)
# weekly data
outbreak_data <- fread("outputs/paper/outbreak_data_20250228.csv") 
# summary data
outbreaks_filtered <-
  readRDS("outputs/paper/outbreak_summary_filtered_20250418.rds")
# filter outbreaks that are unique
# outbreaks selected through a series of selection filters described in # data_prep.qmd
outbreak_data <- outbreak_data %>% 
  filter(id_outbreak %in% outbreaks_filtered$id_outbreak)

# Historical OCV doses deployed by year
d <- clean_names(fread("data/paper/OCV_Use_GTFCC_Dashboard.csv"))
d |> 
  filter(shipped == "Shipped", type == "Reactive") |> 
  group_by(year) |> 
  summarize(num_ocv = sum(dose)) |>
  setDT() -> ocv_year

ocv_year$cum_num_ocv <- cumsum(ocv_year$num_ocv)
ocv_year
# # Get outbreak summary data
# ds <- clean_names(as.data.frame(fread("data/paper/outbreak_files_Oct2024/outbreak_summary.csv")))
# 
# # Create unique IDs for each outbreak
# ds |>
#   group_by(location, outbreak_number) |>
#   mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
#   ungroup() -> ds
# 
# # Join outbreak data with summary details
# outbk_data_summary <- 
#   left_join(filter(outbreak_data, week == 1), 
#             ds[, c("outbreak_pop", "total_s_ch", "total_c_ch", "total_deaths", 
#                    "duration_weeks", 
#                   "peak_week_incidence", "time_to_peak_week", 
#                   "peak_week_incidence_per_100percent", 
#                   "peak_week_incidence_per_10e3", "mean_r0_1st_week", 
#                   "area_per_1km2", 
#                   "rural_urban", "ar_per_10e3", "cfr_per_100percent", 
#                   "id_outbreak")], 
#             by = "id_outbreak")
# 
# outbk_data_summary %>% 
#   group_by(spatial_scale) %>% 
#   summarize_quantiles(variable=cfr_per_100percent)

# fwrite(outbk_data_summary, "outputs/paper/outbreak_data_summary_20251022.csv")
outbk_data_summary <- fread("outputs/paper/outbreak_data_summary_20251022.csv")
```

## Load economic and demographic data

```{r econ-demo-data}
# Load life expectancy and GDP data
life_exp_data <- 
  as.data.frame(fread("outputs/paper/wpp_life_expectancy_20241022.csv"))
# life expectancy at age 25 during 2010-2020, mean age for cholera
life_exp_data %>% 
  filter(Year >= 2010, Year <= 2020, 
         `Region, subregion, country or area *` == "Sub-Saharan Africa") %>% 
  pull(`25`) %>%  
  as.numeric() -> life_exp_25

summary(life_exp_25)

# Load population by age data 
# prop of under 5 during 2010-2020
age_dist <- 
  as.data.frame(fread("outputs/paper/wpp_pop_by_age_20241022.csv"))
age_dist %>% 
  filter(Year >= 2010, Year <= 2020,
         `Region, subregion, country or area *` == "Sub-Saharan Africa") %>% 
  pull(prop_u5) -> prop_U5

summary(prop_U5)

# gdp data from World Bank during 2010-2020
gdp_data <- read_xls("data/paper/GDP_WorldBank.xls")
gdp_data %>% 
  filter(`Country Name` == "Sub-Saharan Africa (IDA & IBRD countries)") %>%
  select(c(`2010`:`2020`)) |> 
  as.numeric() -> gdp_ssa

summary(gdp_ssa)

# workforce data
workforce_data <- read_xls("data/paper/Workforce_Worldbank.xls")
workforce_data %>% 
  filter(`Country Name` == "Sub-Saharan Africa (excluding high income)")%>%
  select(c(`2010`:`2020`)) |> 
  as.numeric() -> wf_ssa

summary(wf_ssa)

# Standardize country code variable names
names(life_exp_data)[names(life_exp_data) == "ISO3 Alpha-code"] <- "country"
names(life_exp_data)[names(life_exp_data) == "Year"] <- "year"
names(workforce_data)[names(workforce_data) == "Country Code"] <- "country"
names(gdp_data)[names(gdp_data) == "Country Code"] <- "country"

# Format workforce data into long format
workforce_long <- 
  workforce_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "pct_workforce")
workforce_long$year <- as.integer(workforce_long$year)

# Format GDP data into long format
gdp_long <- 
  gdp_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "gdp")
gdp_long$year <- as.integer(gdp_long$year)
```

## Cost-effectiveness parameters

```{r cost-params}
parm <- fread("data/paper/parameters.csv")
parm <- parm[Disease == "Cholera"]
# GBD data deleted and replaced with Mbewe (2025) Open Forum Infect Dis data
# it is assumed that moderate and severe cases are reported 
parm <- parm[!(Parameter == "Prop_Moderate" & Value == 0.289)]
parm <- parm[!(Parameter == "Prop_Severe" & Value == 0.069)]

# Disease burden parameters
# pr_asymp <- parms[Parameter == "Prop_Asymptomatic", "Value"]
# pr_mild <- parms[Parameter == "Prop_Mild", "Value"]
pr_moderate <- parm[Parameter == "Prop_Moderate", Value]
pr_severe <- parm[Parameter == "Prop_Severe", Value]
dur <- parm[Parameter == "Duration_Illness", Value]
# wt_asymp <- parms[Parameter == "Disability_Weight_Asymptomatic", "Value"]
# wt_mild <- parms[Parameter == "Disability_Weight_Mild", "Value"]
wt_moderate <- parm[Parameter == "Disability_Weight_Moderate", Value]
wt_severe <- parm[Parameter == "Disability_Weight_Severe", Value]

# Vaccination costs
vacc_cost_per_dose <- parm[Parameter == "Vaccine_Cost", Value]
# DEPRECATED (reviewer comment 5): vacc_delivery_cost is no longer read from
# parameters.csv (Sarker 2022). It is now derived from Mogasale 2016 (SSA) in the
# delivery-cost block further below. Old line kept (commented) for diffing:
# vacc_delivery_cost <- parm[Parameter == "Vaccine_Delivery_Cost", Value]
vacc_shipping_cost <- parm[Parameter == "Vaccine_Shipping_Cost", Value]

# Direct medical costs
patient_cost_hosp <- parm[Parameter == "Patient_Cost_Hosp", Value]
patient_cost_outpt <- parm[Parameter == "Patient_Cost_Outpt", Value]
public_cost_hosp <- parm[Parameter == "Public_Cost_Hosp", Value]
public_cost_outpt <- parm[Parameter == "Public_Cost_Outpt", Value] 

# Productivity costs
patient_workday_lost <- parm[Parameter == "Pt_Workdays_Lost", Value]
caregiver_workday_lost <- parm[Parameter == "Caregiver_Workdays_Lost", Value]
mean_age_inf <- parm[Parameter == "Mean_Age_Infection", Value]

# Vaccination regimen
dose_regimen <- 1  # Single-dose regimen
vacc_cost_per_person <- vacc_cost_per_dose * dose_regimen

# Calculate cost of illness per patient (weighted average)
pr_tot <- pr_moderate + pr_severe
coi_per_patient <- 
    pr_moderate / pr_tot * (patient_cost_outpt + public_cost_outpt) + 
    pr_severe / pr_tot * (patient_cost_hosp + public_cost_hosp)

# Consumer price index for cost adjustment for inflation(2015 to 2023)
cpi_percent <- read_xls("data/paper/imf-dm-export-20241028.xls")
cpi_percent <- cpi_percent %>%
  pivot_longer(cols = `2015`:`2029`, names_to = "year", values_to = "value")
cpi_percent$year <- as.numeric(cpi_percent$year)

# Adjust medical costs (2015 to 2023)
cpi_med <- cpi_percent %>% filter(year >= 2016 & year <= 2023)
xi_med <- (cpi_med$value/100 + 1)
inflation_factor_med <- prod(xi_med)

patient_cost_hosp <- patient_cost_hosp * inflation_factor_med
patient_cost_outpt <- patient_cost_outpt * inflation_factor_med
public_cost_hosp <- public_cost_hosp * inflation_factor_med
public_cost_outpt <- public_cost_outpt * inflation_factor_med

# ============================================================================
# DEPRECATED: Sarker 2022 delivery cost (replaced per reviewer comment 5)
# Sarker et al. 2022 (Vaccine: X), Bangladesh: US$0.91/dose (2016 US$), inflated
# 2016 -> 2023 with the SSA CPI series (product of 1+rate/100 over 2017-2023,
# giving ~$2.32/dose in 2023). Kept commented for diffing against the new
# Mogasale 2016 estimate below.
#   vacc_delivery_cost <- parm[Parameter == "Vaccine_Delivery_Cost", Value]  # $0.91 (2016 US$)
#   cpi_vac <- cpi_percent %>% filter(year >= 2017 & year <= 2023)
#   xi_vac <- (cpi_vac$value/100 + 1)
#   inflation_factor_vac <- prod(xi_vac)
#   vacc_delivery_cost <- vacc_delivery_cost * inflation_factor_vac          # ~$2.32 (2023 US$)
# ============================================================================

# ----------------------------------------------------------------------------
# OCV delivery cost from Mogasale et al. 2016 (PLoS NTD), Table 2,
# "Total OCV Delivery (Unit) Cost", Method 2 = 2014 US$ per fully vaccinated
# (two-dose) person, for the three sub-Saharan-Africa Shanchol campaigns.
# ----------------------------------------------------------------------------
# (1) Per-person (two-dose) delivery cost, 2014 US$
delivery_cost_per_person_2014 <- c(
  "Guinea 2012 (reactive, rural)"   = 2.32,
  "South Sudan 2013 (refugee camp)" = 3.05,
  "South Sudan 2014 (IDP)"          = 1.72
)

# (2) Approx per-dose delivery cost = half the per-person (two-dose) cost.
#     Our ORI is single-dose; Sarker 2022 found fixed costs were ~2% of total,
#     so per-dose delivery ~= half per-person delivery. Keep both vectors.
delivery_cost_per_dose_2014 <- delivery_cost_per_person_2014 / 2

# (3) Inflate each per-dose value 2014 -> 2023 using the SSA CPI series already
#     loaded above (cpi_percent), with the SAME multiplicative method used for
#     medical and (deprecated) vaccine costs. Base year 2014 -> apply rates for
#     2015..2023 (2015 = 7.0, 2016 = 10.3 come from the IMF SSA file, not hard-coded).
cpi_deliv <- cpi_percent %>% filter(year >= 2015 & year <= 2023)
inflation_factor_deliv <- prod(cpi_deliv$value/100 + 1)
delivery_cost_per_dose_2023 <- delivery_cost_per_dose_2014 * inflation_factor_deliv

# (4) Delivery cost is now treated as a per-run random input: it is drawn
#     Uniform(min, max) over the three inflated per-dose values inside
#     draw_parameter_samples() (see params file's vacc_delivery_cost column),
#     and joined into the CEA data by run id below so add_cea_results() uses a
#     per-run cost. The scalar median here is kept as a fallback / central
#     estimate for reporting and for any run lacking the per-run column.
#     NOTE: Guinea 2012 is the ONLY *reactive* campaign of the three -- if a
#     reactive-only base case is wanted later, use the Guinea value alone.
vacc_delivery_cost     <- median(delivery_cost_per_dose_2023)
vacc_delivery_cost_min <- min(delivery_cost_per_dose_2023)
vacc_delivery_cost_max <- max(delivery_cost_per_dose_2023)

# (5) Re-derive the total per-dose vaccine cost that feeds add_cea_results().
#     Model cost term (R/utils.R ~L634), shipping now INCLUDED per reviewer comment 5:
#       (vacc_cost_per_dose + vacc_shipping_cost + vacc_delivery_cost) * vacc_dose
vacc_total_cost_per_dose <- vacc_cost_per_dose + vacc_shipping_cost + vacc_delivery_cost

# (6) Summary of the delivery-cost derivation
delivery_cost_summary <- data.frame(
  campaign        = names(delivery_cost_per_person_2014),
  per_person_2014 = as.numeric(delivery_cost_per_person_2014),
  per_dose_2014   = as.numeric(delivery_cost_per_dose_2014),
  per_dose_2023   = round(as.numeric(delivery_cost_per_dose_2023), 4),
  row.names = NULL
)
print(delivery_cost_summary)
cat(sprintf("inflation_factor_deliv (2014->2023, SSA CPI) = %.4f\n", inflation_factor_deliv))
cat(sprintf("Base-case delivery cost (median) = $%.4f/dose (2023 US$); range $%.4f - $%.4f\n",
            vacc_delivery_cost, vacc_delivery_cost_min, vacc_delivery_cost_max))
cat(sprintf("Total per-dose cost (procurement $%.2f + shipping $%.2f + delivery $%.4f) = $%.4f\n",
            vacc_cost_per_dose, vacc_shipping_cost, vacc_delivery_cost, vacc_total_cost_per_dose))
```

# Vaccine impact simulation

### Outbreak onset-triggered strategy

This section models vaccination initiated at specific weeks after outbreak onset.

```{r onset-trigger-sim, eval=FALSE}
tic <- Sys.time()  # Start timer

# Setup for parallel processing
vacc_covs <- c(0.6, 0.75, 0.9)    # Vaccine coverage levels to simulate
vacc_weeks <- seq(1, 15, by = 2)  # Vaccination weeks to simulate
nruns <- 200                      # Number of simulation runs per configuration
batch_size <- 10  # Number of runs per parallel batch
delays <- 2     # No ORI after two or more weeks after outbreak end
# Loop through batches of runs for parallel execution
for (delay in delays) {
  message(paste("delay:", delay))
  for (batch_id in seq(1, nruns, by = batch_size)) {
    message(paste("batch ID:", batch_id))
    run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in current batch
  
    # Create and register parallel cluster
    cl <- makeCluster(getOption("cl.cores", detectCores() - 3))
    registerDoParallel(cl)
  
    # Run simulations in parallel
    svim_week_list <-
      foreach(i = run_ids, .packages = c("data.table")) %dopar% {
        run_vacc_impact_outbreak_weekly(
          outbreak_data = outbreak_data,
          ive_data = ive_data,
          parameters = params,
          age_dist = age_dist,
          vacc_week = vacc_weeks,    # Outbreak onset-triggered vaccination
          case_trigger = NULL,
          vacc_cov = vacc_covs,
          runid = i,
          no_ori_delay_outbreak_end = delay)
      }
    stopCluster(cl)
  
    # Save batch results
    fwrite(rbindlist(svim_week_list), 
           paste0("outputs/svim_delay_", delay, 
                  "_week_list_vacc_week_batch_", batch_id, "_",
                  tstamp(), ".csv"))
  }
}
toc <- Sys.time()
message(paste("Total execution time:", toc - tic))
```

### Onset-triggered strategy across delays to stop ORI after outbreak end

```{r onset-trigger-sim, eval=FALSE}
tic <- Sys.time()  # Start timer
delays <- c(0:7)

# Setup for parallel processing
vacc_covs <- c(0.75)    # Vaccine coverage levels to simulate
vacc_weeks <- seq(1, 15, by = 2)  # Vaccination weeks to simulate
nruns <- 200                      # Number of simulation runs per configuration
batch_size <- 10                  # Number of runs per parallel batch

for (delay in delays) {
  # Loop through batches of runs for parallel execution
  for (batch_id in seq(1, nruns, by = batch_size)) {
    message(paste("Processing batch ID:", batch_id))
    run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in current batch
  
    # Create and register parallel cluster
    cl <- makeCluster(getOption("cl.cores", detectCores() - 3))
    registerDoParallel(cl)
  
    # Run simulations in parallel
    svim_week_list <-
      foreach(i = run_ids, .packages = c("data.table")) %dopar% {
        run_vacc_impact_outbreak_weekly(
          outbreak_data = outbreak_data,
          ive_data = ive_data,
          parameters = params,
          age_dist = age_dist,
          vacc_week = vacc_weeks,    # Outbreak onset-triggered vaccination
          case_trigger = NULL,
          vacc_cov = vacc_covs,
          runid = i,
          no_ori_delay_outbreak_end = delay)
      }
    stopCluster(cl)
  
    # Save batch results
    fwrite(rbindlist(svim_week_list), 
           paste0("outputs/svim_delay_", delay, 
                  "_week_list_vacc_week_batch_", batch_id, "_", 
                  tstamp(), ".csv"))
  }
}
toc <- Sys.time()
message(paste("Total execution time:", toc - tic))
```

### Onset-triggered strategy for the alternative dataset.

```{r onset-trigger-sim, eval=FALSE}
tic <- Sys.time()  # Start timer

# sensitivity analyses 
outbreaks_alt <- read_rds("outputs/paper/outbreaks_alt_included_20250912.rds")
dat_sim <- outbreak_data_all %>% 
  filter((id_outbreak %in% outbreaks_alt$id_outbreak) & 
           !(id_outbreak %in% outbreaks_filtered$id_outbreak)) 
delays <- 2

# Setup for parallel processing
vacc_covs <- c(0.75)    # Vaccine coverage levels to simulate
vacc_weeks <- seq(1, 15, by = 2)  # Vaccination weeks to simulate
nruns <- 200                      # Number of simulation runs per configuration
batch_size <- 10                  # Number of runs per parallel batch

for (delay in delays) {
  # Loop through batches of runs for parallel execution
  for (batch_id in seq(1, nruns, by = batch_size)) {
    message(paste("Processing batch ID:", batch_id))
    run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in current batch
  
    # Create and register parallel cluster
    cl <- makeCluster(getOption("cl.cores", detectCores() - 3))
    registerDoParallel(cl)
  
    # Run simulations in parallel
    svim_week_list <-
      foreach(i = run_ids, .packages = c("data.table")) %dopar% {
        run_vacc_impact_outbreak_weekly(
          outbreak_data = dat_sim,
          ive_data = ive_data,
          parameters = params,
          age_dist = age_dist,
          vacc_week = vacc_weeks,    # Outbreak onset-triggered vaccination
          case_trigger = NULL,
          vacc_cov = vacc_covs,
          runid = i,
          no_ori_delay_outbreak_end = delay)
      }
    stopCluster(cl)
  
    # Save batch results
    fwrite(rbindlist(svim_week_list), 
           paste0("outputs/svim_alt_data_delay_", delay, 
                  "_week_list_vacc_week_batch_", batch_id, "_", 
                  tstamp(), ".csv"))
  }
}
toc <- Sys.time()
message(paste("Total execution time:", toc - tic))
```

### Case-Triggered Strategy

```{r case-trigger-sim, eval=FALSE}
tic <- Sys.time()  # Start timer

# Define case thresholds that trigger vaccination
case_triggers <- c(1, 5, seq(10, 90, 20), seq(140, 300, 50))
vacc_covs <- c(0.75)
vacc_weeks <- seq(1, 15, by=2)

nruns <- 200               # Number of simulation runs
batch_size <- 10           # Number of runs per batch
delay <- 2                 # No ORI if two or more weeks pass since an outbreak ends.  
# Loop through batches of runs for parallel execution
for (batch_id in seq(1, nruns, by = batch_size)) {
  message(paste("Processing batch ID:", batch_id))
  run_ids <- batch_id:(batch_id + batch_size - 1)

  # Create and register parallel cluster
  ncores <- detectCores()
  cl <- makeCluster(getOption("cl.cores", ncores - 1))
  registerDoParallel(cl)

  # Run simulations in parallel
  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = outbreak_data,
        ive_data = ive_data,
        parameters = params,
        age_dist = age_dist,
        vacc_week = vacc_weeks,
        case_trigger = case_triggers,  # Case-triggered vaccination
        vacc_cov = vacc_covs,
        runid = i,
        no_ori_delay_outbreak_end = delay)
    }
  stopCluster(cl)
  
  # Save batch results
  saveRDS(rbindlist(svim_week_list),
          paste0("outputs/svim_week_case_trigger_batch_", 
                 batch_id, "_", tstamp(), ".rds"))
}
toc <- Sys.time()
message(paste("Total execution time:", round(toc - tic, digits=2), "minutes"))
```

## Combine Simulation Results

Combine results from batch processing and organizes them for analysis.

### Onset-triggered 
```{r}
id_begin <- 1
id_end <- 200
batch_size <- 10
batch_ids <- seq(id_begin, id_end, by = batch_size)

# Dates to look for
dates <- c("20251031", "20251030")

files <- unlist(lapply(batch_ids, function(id) {
  f <- paste0("outputs/svim_delay_2_week_list_vacc_week_batch_", id, "_", dates, ".csv")
  f[file.exists(f)]
}))

print(files)

df <- rbindlist(lapply(files, fread))

df_sow <- sum_over_week(df,
            group_cols = c("runid", "id_outbreak", "vacc_week", 
                         "vacc_cov", "no_ori_delay_outbreak_end"))

saveRDS(df_sow,
        paste0("outputs/onset_trigger_sum_over_week_",
               tstamp(), ".rds"))
```

### Onset-triggered across delays to stop ORI after outbreak end

```{r}
id_begin <- 1
id_end <- 200
batch_size <- 10
batch_ids <- seq(id_begin, id_end, by = batch_size)

# Dates to look for
dates <- c("20251031", "20251030")
delays <- 0:7
delays_dates <- expand.grid(delays, dates)

files <- unlist(lapply(batch_ids, function(id) {
  # f <- paste0("outputs/svim_weekly_delay_", delays, 
  #"_week_list_vacc_week_batch_", id, "_", dates, ".csv")
    f <- paste0("outputs/svim_weekly_delay_", delays_dates$Var1, 
                "_week_list_vacc_week_batch_", id, "_", delays_dates$Var2, ".csv")
  f[file.exists(f)]
}))

print(files)

df_delay <- rbindlist(lapply(files, fread))
df_delay_sow <- 
  sum_over_week(df_delay, group_cols = c("runid", "id_outbreak", "vacc_week", 
                                   "vacc_cov", "no_ori_delay_outbreak_end"))

saveRDS(df_delay_sow,
        paste0("outputs/onset_trigger_delays_sum_over_week_",
               tstamp(), ".rds"))
```

### Onset-triggered for strategy for the excluded data set.

```{r}
id_begin <- 1
id_end <- 200
batch_size <- 10
batch_ids <- seq(id_begin, id_end, by = batch_size)

# Dates to look for
dates <- c("20251106")

files <- unlist(lapply(batch_ids, function(id) {
  # f <- paste0("outputs/svim_weekly_delay_", delays, 
  #"_week_list_vacc_week_batch_", id, "_", dates, ".csv")
    f <- paste0("outputs/svim_alt_data_delay_2_week_list_vacc_week_batch_", 
                batch_id, "_", dates, ".csv")
  f[file.exists(f)]
}))

print(files)

df_alt <- rbindlist(lapply(files, fread))
df_alt_sow <- 
  sum_over_week(df_alt, group_cols = c("runid", "id_outbreak", "vacc_week", 
                                   "vacc_cov", "no_ori_delay_outbreak_end"))

saveRDS(df_alt_sow,
        paste0("outputs/onset_trigger_alt_data_sum_over_week_",
               tstamp(), ".rds"))
```

### Case-triggered strategy

```{r}
dates <- c("20251031", "20251030")

files <- unlist(lapply(batch_ids, function(id) {
  f <- paste0("outputs/svim_week_case_trigger_batch_", id, "_", dates, ".rds")
  f[file.exists(f)]
}))

print(files)

df_ct <- rbindlist(lapply(files, read_rds))
df_ct_sow <- 
  sum_over_week(df_ct,
    group_cols = c("runid", "id_outbreak", "vacc_week", "vacc_cov",
                   "no_ori_delay_outbreak_end", "case_trigger"))

saveRDS(df_ct_sow,
        paste0("outputs/case_trigger_sum_over_week_",
               tstamp(), ".rds"))
```

# Cost-Effectiveness Analysis

Calculate impact (cases, deaths, and DALYs averted), efficiency (impact per 1000 OCV), and cost effectiveness (cost [2023 US$] per unit of impact). 

## Onset-triggered strategy

```{r}
files <- list.files(
  path = "outputs",
  pattern = "onset_trigger_sum_over_week.*(20251030|20251031)",
  full.names = TRUE
)
files <- files[!grepl("delay", basename(files))]

print(files)

d <- rbindlist(lapply(files, read_rds))

d_cea_var <- add_cea_variables(d, mean_age_inf = mean_age_inf)
# Per-run OCV delivery cost (2023 US$), drawn Uniform(min,max) in
# draw_parameter_samples(); join by run id so add_cea_results() uses a per-run
# delivery cost (column) instead of the scalar median fallback.
d_cea_var <- merge(d_cea_var, params[, .(runid = run_id, vacc_delivery_cost)],
                   by = "runid", all.x = TRUE)
d_cea_res <- add_cea_results(d_cea_var, parms = as.data.frame(parm))
# #  
d_cea_res %>%
  filter(ori_occurred == TRUE, vacc_week %in% c(1,9)) %>%
  group_by(vacc_cov, vacc_week) %>% 
  summarize(pct_reduc_case_m = median(pct_reduc_case))
# 
saveRDS(d_cea_res, 
        paste0("outputs/paper/cea_onset_trigger_", tstamp(), ".rds"))
```

## Onset-triggered with delays to stop ORI after outbreak end

```{r}
files <- list.files(
  path = "outputs",
  pattern = "onset_trigger.*sum_over_week.*(20251030|20251031)",
  full.names = TRUE
)
files <- files[grepl("delay", basename(files))]

print(files)

d <- rbindlist(lapply(files, read_rds))

d_cea_var <- add_cea_variables(d, mean_age_inf = mean_age_inf)
# Per-run OCV delivery cost (2023 US$), drawn Uniform(min,max) in
# draw_parameter_samples(); join by run id so add_cea_results() uses a per-run
# delivery cost (column) instead of the scalar median fallback.
d_cea_var <- merge(d_cea_var, params[, .(runid = run_id, vacc_delivery_cost)],
                   by = "runid", all.x = TRUE)
d_cea_res <- add_cea_results(d_cea_var, parms = as.data.frame(parm))

d_cea_res %>%
  filter(ori_occurred == TRUE, vacc_cov == 0.75, vacc_week == 9) %>%
  group_by(no_ori_delay_outbreak_end) %>%
  summarize(pct_reduc_case_m = median(pct_reduc_case))

saveRDS(d_cea_res, 
        paste0("outputs/paper/cea_delay_onset_trigger_", tstamp(), ".rds"))
```


## Onset-triggered for the alternative dataset

```{r}
# alternative outbreak dataset
outbreaks_alt <- read_rds("outputs/paper/outbreaks_alt_included_20250912.rds")
# simulation results based on the original dataset
d1 <- read_rds("outputs/onset_trigger_sum_over_week_20251031.rds")
# simulation results for outbreaks that were added additionally to the alternative dataset
d2 <- read_rds("outputs/onset_trigger_alt_data_sum_over_week_20251031.rds")
# exclude outbreaks from the original dataset that are not in the alternative dataset
d1_alt <- d1 %>% filter(id_outbreak %in% outbreaks_alt$id_outbreak, 
                        vacc_cov == 0.75)
d2_alt <- d2 %>% filter(id_outbreak %in% outbreaks_alt$id_outbreak)
# > all.equal(d2_alt, d2)
# [1] TRUE
# alternative dataset by combining d1_alt and d2_alt
d_alt <- rbindlist(list(d1_alt, d2_alt))

d_alt_cea_var <- add_cea_variables(d_alt, mean_age_inf = mean_age_inf)
# Per-run OCV delivery cost (2023 US$): join by run id (see other CEA chunks).
d_alt_cea_var <- merge(d_alt_cea_var,
                       params[, .(runid = run_id, vacc_delivery_cost)],
                       by = "runid", all.x = TRUE)
d_alt_cea_res <- add_cea_results(d_alt_cea_var, parms = as.data.frame(parm))

d_alt_cea_res %>%
  filter(ori_occurred == TRUE, vacc_cov == 0.75) %>%
  group_by(vacc_week) %>%
  summarize(pct_reduc_case_m = median(pct_reduc_case))

saveRDS(d_alt_cea_res, 
        paste0("outputs/paper/cea_alt_data_onset_trigger_", tstamp(), ".rds"))
```

## Case-triggered

```{r}
files <- list.files(
  path = "outputs",
  pattern = "case_trigger.*sum_over_week.*(20251030|20251031)",
  full.names = TRUE
)

print(files)

d <- rbindlist(lapply(files, read_rds))

d_cea_var <- add_cea_variables(d, mean_age_inf = mean_age_inf)
# Per-run OCV delivery cost (2023 US$), drawn Uniform(min,max) in
# draw_parameter_samples(); join by run id so add_cea_results() uses a per-run
# delivery cost (column) instead of the scalar median fallback.
d_cea_var <- merge(d_cea_var, params[, .(runid = run_id, vacc_delivery_cost)],
                   by = "runid", all.x = TRUE)
d_cea_res <- add_cea_results(d_cea_var, parms = as.data.frame(parm))

d_cea_res %>%
  filter(ori_occurred == TRUE, vacc_cov == 0.75, vacc_week == 9) %>%
  group_by(case_trigger) %>%
  summarize(pct_reduc_case_m = median(pct_reduc_case))

saveRDS(d_cea_res, 
        paste0("outputs/paper/cea_case_trigger_", tstamp(), ".rds"))

```

# Summary results for plotting 

## Median values for plotting

```{r}
VC3    <- c(0.6, 0.75, 0.9)
VW_seq <- seq(1, 15, 2)

build_plot_summary <- function(data, value_col,
                               by = c("vacc_cov", "vacc_week"),
                               round_digits = 3) {
  val <- rlang::enquo(value_col)
  
  data %>%
    group_by(across(all_of(by))) %>%
    summarise(
      mean = mean(!!val, na.rm = TRUE),
      q025 = quantile(!!val, 0.025, na.rm = TRUE),
      q250 = quantile(!!val, 0.25,  na.rm = TRUE),
      q500 = quantile(!!val, 0.50,  na.rm = TRUE),
      q750 = quantile(!!val, 0.75,  na.rm = TRUE),
      q975 = quantile(!!val, 0.975, na.rm = TRUE),
      n    = sum(!is.na(!!val)),
      .groups = "drop"
    ) %>%
    mutate(across(c(mean, q025, q250, q500, q750, q975),
                  ~ round(.x, round_digits)))
}

# In case cea files were not loaded
cea_files <- c(
  "outputs/paper/cea_onset_trigger_20251031.rds"
)

# Example: summarize % cases averted by week & coverage
cea_summary_pca <- 
  map_dfr(cea_files, read_rds) %>% 
  filter(ori_occurred == TRUE) %>%
  filter(vacc_cov %in% VC3, vacc_week %in% VW_seq) %>%
  build_plot_summary(pct_reduc_case, by = c("vacc_cov", "vacc_week"))

saveRDS(cea_summary_pca, 
        paste0("outputs/paper/cea_pca_summary_", tstamp(), ".rds"))
```

## Median values across prioritization by attack rate, duration, and size

```{r}
outbk_dat <- outbreaks_filtered |>
  summarize(
    year = first(data.table::year(start_date)),
    rural_urban = first(rural_urban),
    ar = first(ar_per_10e3),
    cfr = first(cfr_per_100percent),
    spatial_scale = first(spatial_scale),
    pop = first(outbreak_pop),
    duration = first(duration_weeks),
    size = first(total_s_ch),
    .by = id_outbreak
  )

# Percentile grid (edit as you like; for full 0–1 include 1)
probs <- seq(0, 0.9, by = 0.1)

# ---- Helpers ----
# id_outbreak in the top p-quantile (e.g., p=0.9 for top 10%)
top_p_ids <- function(data, var, p = 0.9) {
  thr <- stats::quantile(dplyr::pull(data, {{ var }}), probs = p, na.rm = TRUE)
  data %>%
    filter({{ var }} >= thr) %>%
    pull(id_outbreak) %>%
    unique()
}

# lapply(probs, function(p) top_p_ids(outbk_dat, ar, p = p))
ids_list_ar <- map(probs, ~ top_p_ids(outbk_dat, ar, p = .x))
ids_list_dur <- map(probs, ~ top_p_ids(outbk_dat, duration, p = .x))
ids_list_size <- map(probs, ~ top_p_ids(outbk_dat, size, p = .x))
ids_list_cfr <- map(probs, ~ top_p_ids(outbk_dat, cfr, p = .x))
```

```{r}
cea_75_ori <- 
  map_dfr(cea_files, read_rds) %>% 
  filter(vacc_cov == 0.75, ori_occurred == TRUE)
  
median_by_ids <- function(cea, ids_list, probs, metrics,
                             id_col = "id_outbreak") {
  stopifnot(length(ids_list) == length(probs))

  map2_dfr(probs, ids_list, function(p, ids) {
    if (length(ids) == 0L) {
      # Return an empty tibble with expected columns (so bind_rows works)
      return(tibble(
        prob = numeric(0),
        vacc_cov = numeric(0),
        vacc_week = numeric(0)
      ))
    }

    cea %>%
      filter(.data[[id_col]] %in% ids) %>%
      group_by(vacc_cov, vacc_week) %>%
      summarise(
        across(
          all_of(metrics),
          ~ median(.x, na.rm = TRUE),
          .names = "med_{.col}"
        ),
        # n = n(),                     # how many rows contributed
        n = length(ids),               
        .groups = "drop"
      ) %>%
      mutate(prob = p, .before = 1)
  })
}

cea_metrics <- c("s_ch_averted_tot", "pct_reduc_case",
                 "case_averted_per_1000_OCV", "daly_averted_per_1000_OCV")  # add 

med_ar <- median_by_ids(cea_75_ori, ids_list_ar, probs, 
                        metrics = cea_metrics)
med_dur  <- median_by_ids(cea_75_ori, ids_list_dur, probs, 
                             metrics = cea_metrics)
med_size <- median_by_ids(cea_75_ori, ids_list_size, probs, 
                             metrics = cea_metrics)
med_cfr <- median_by_ids(cea_75_ori, ids_list_cfr, probs, 
                            metrics = cea_metrics)
# Optional: label and combine into one tidy table
med_all <- bind_rows(
  mutate(med_ar,   percentile_by = "ar"),
  mutate(med_dur,  percentile_by = "duration"),
  mutate(med_size, percentile_by = "size"),
  mutate(med_cfr, percentile_by = "cfr")) %>%
  relocate(percentile_by, prob, vacc_cov, vacc_week)

med_all_long <- med_all %>%
  pivot_longer(
    cols = starts_with("med_"),
    names_to = "metric",
    values_to = "median_value")

saveRDS(med_all_long, 
        paste0("outputs/paper/summary_impact_0.75_prioritization_", 
               tstamp(), ".rds"))
```

# Sampling outbreaks

This section creates and analyzes subsets of outbreaks to simulate different targeting strategies.

Outbreaks are randomly sampled for a given OCV budget. Outbreaks are eligible if the underlying population factored by the vaccine coverage proportion is smaller than the remaining vaccine doses. 

## Sample 

Random sampling of outbreaks from all or subsets of outbreaks
```{r}
# Default outbreak onset-triggered vaccination results (delay = 2, VC=0.75)
cea_files <- c("outputs/paper/cea_onset_trigger_20251031.rds")
cea <- map_dfr(cea_files, read_rds) %>% filter(vacc_cov == 0.75)

outbk_dat <- outbreaks_filtered |>
  summarize(
    year = first(data.table::year(start_date)),
    rural_urban = first(rural_urban),
    ar = first(ar_per_10e3),
    cfr = first(cfr_per_100percent),
    spatial_scale = first(spatial_scale),
    pop = first(outbreak_pop),
    duration = first(duration_weeks),
    size = first(total_s_ch),
    .by = id_outbreak
  )

# exclude additionally where gdp is not available
cea %>% filter(is.na(gdp)) %>% pull(id_outbreak) %>% unique() -> ids_na
outbk_dat <- outbk_dat[!(outbk_dat$id_outbreak %in% ids_na), ]

priorities <- c("size", "dur", "AR")
ocv_doses  <- c(1e5, 2e5, 5e5, 1e6, 2e6, 5e6, 10e6, 20e6, 50e6, 100e6, 200e6)
nsubset    <- 200
probs      <- c(0.25)
cov        <- 0.75                     # vaccine coverage (fraction)

set.seed(42)

plan(multisession)

# helper to filter by priority + pr
filter_by_priority <- function(dat, priority, pr) {
  if (priority == "size") {
    thr <- quantile(dat$size, probs = pr, na.rm = TRUE)
    dat %>% filter(size >= thr)
  } 
  else if (priority == "dur") {
    thr <- quantile(dat$duration, probs = pr, na.rm = TRUE)
    dat %>% filter(duration >= thr)
  } 
  else if (priority == "AR") {
    thr <- quantile(dat$ar, probs = pr, na.rm = TRUE)
    dat %>% filter(ar >= thr)
  } 
  else stop("Unknown priority")
}

# helper to filter by cutoff value for the variable
filter_by_cutoff <- function(dat, var, cutoff) {
  if (var == "size") {
    dat %>% filter(size >= cutoff)
  } else if (var == "dur") {
    dat %>% filter(duration >= cutoff)
  } else if (var == "AR") {
    dat %>% filter(ar >= cutoff)
  } else stop("Unknown variable")
}


run_one_sample <- function(dat, ocv_dose, cov, j) {
  set.seed(j)
  smpl <- sample_outbreaks(
    outbreak_data         = dat,
    target_ocv_doses      = ocv_dose,
    target_coverage       = cov,
    target_outbreak_count = NULL
  )
  tibble(
    id_outbreak = smpl$sampled_data$id_outbreak,
    pop = smpl$sampled_data$pop,
    outbreak_count = smpl$outbreak_count,
    doses_used = smpl$doses_used,
    id_sample   = j
  )
}

for (priority in priorities) {
  for (ocv_dose in ocv_doses) {
    outbreaks_sampled <- map_dfr(probs, function(pr) {
      dat_filtered <- filter_by_priority(outbk_dat, priority, pr)
      
      # Run nsubset samples in parallel
      future_map_dfr(
        1:nsubset,
        ~ run_one_sample(dat_filtered, ocv_dose, cov, .x),
        .options = furrr_options(seed = TRUE)
      ) %>%
        mutate(
          prob             = pr,
          vacc_priority    = priority,
          ocv_dose         = ocv_dose,
          vacc_cov         = cov,
          doses_used       = doses_used, 
          outbreak_count   = outbreak_count, 
          sampled          = TRUE,
          reason           = NA_character_,
          .before = 1
        )
    }, .progress = TRUE)

    nm <- paste0("outputs/subsets_", 
             round(ocv_dose/1e6), "M_", priority, "_", tstamp(), "_.rds")
    
    if (ocv_dose < 1e6) {
       nm <- paste0("outputs/subsets_", 
             round(ocv_dose/1e3), "K_", priority, "_", tstamp(), "_.rds")
    }
    saveRDS(outbreaks_sampled, nm)
  }
}

plan(sequential)
```

Summary statistics of subsets

```{r}
files <- list.files(
  path = "outputs",
  pattern = "subsets.*200M.*(AR|size|dur).*(20251104)",
  full.names = TRUE
)
files <- files[!grepl("delay", basename(files))]

print(files)

dat <- map_dfr(files, read_rds)
pr <- c(0.0, 0.3, 0.7)

d <- dat %>%
  mutate(prob_r = round(prob, digits = 1)) %>% 
  filter(prob_r %in% pr) %>% 
  group_by(vacc_priority) %>% 
  select(vacc_priority, prob, id_outbreak, outbreak_count, doses_used)

d_joined <- left_join(d, outbk_data_summary, by="id_outbreak") 
d_joined %>% 
  group_by(vacc_priority, prob) %>% 
  summarize(dose_m = median(doses_used),
            oubreaks_m = median(outbreak_count),
            ar_m = median(ar_per_10e3),
            size_m = median(total_s_ch),
            dur_m = median(duration_weeks))
```

## Aggregate 

Various impact metrics are aggregated across outbreaks within each sample using the `sum_over_outbreaks_by_runid` function

```{r}
priorities <- c("AR", "dur", "size")
ocv_doses  <- c(1e5, 2e5, 5e5, 1e6, 2e6, 5e6, 10e6, 20e6, 50e6, 100e6, 200e6)
ocv_doses  <- c(1e5)

cea_files <- 
  c("outputs/paper/cea_onset_trigger_20251031.rds")

cea_75 <- 
  map_dfr(cea_files, read_rds) %>% 
  filter(vacc_cov == 0.75, vacc_week %in% c(1,9))

##  Worker that computes one group ----------
compute_one_grp <- function(g, cea_data) {
  ids <- grp2ids[grp == g, id_outbreak][[1L]]
  if (length(ids) == 0L) return(NULL)
  # Fast keyed subset on CEA
  cea_sub <- cea_data[.(ids), nomatch = 0L]
  cae_sub_ori <- cea_sub[ori_occurred == TRUE]
  out <- sum_over_outbreaks_by_runid(cae_sub_ori, 
             group_cols <- c("runid", "id_outbreak", "vacc_week", "vacc_cov",
                    "no_ori_delay_outbreak_end"))
  
  # Attach metadata
  m <- meta[grp == g]
  out[, `:=`(
    subset_id        = m$id_sample,
    vacc_priority    = m$vacc_priority,
    priority_prob    = m$prob,
    target_vacc_doses= m$ocv_dose,
    outbreak_count   = m$outbreak_count,
    doses_used       = m$doses_used
  )]
  
  out[]
}

## ---------- 2) Prepare CEA for fast lookup ----------
vacc_weeks <- c(1, 9) 
probabilities <- c(0, 0.3)  

for (vw in vacc_weeks) {
  message(paste0("vacc week:" , vw))
  cea_data <- cea_75 %>% filter(vacc_week == vw) 
  setDT(cea_data)
  setkey(cea_data, id_outbreak)  
  
  # Grid of filenames
  grid <- CJ(vacc_priority = priorities, ocv_dose = ocv_doses)
  grid[, fn := ifelse(ocv_dose < 1e6, 
                      sprintf("outputs/subsets_%sK_%s_20251104.rds",
                              round(ocv_dose/1e3), vacc_priority),
                      sprintf("outputs/subsets_%sM_%s_20251104.rds", 
                              round(ocv_dose/1e6), vacc_priority))]
  
  # subset_list <- lapply(grid$fn, readRDS)
  subsets <- rbindlist(lapply(grid$fn, read_rds)) 
  
  setDT(subsets)
  setkey(subsets, vacc_priority, ocv_dose, vacc_cov, prob, id_sample)
  
  # Build group index & metadata
  # One group per (priority, dose, prob, id_sample)
  subsets[, grp := .GRP, by = .(vacc_priority, ocv_dose, prob, id_sample)]
  
  # Metadata to re-attach later (one row per group)
  meta <- unique(subsets[, .(grp, vacc_priority, ocv_dose, prob, id_sample,
                             outbreak_count, doses_used)])
  
  # Map each group to its vector of outbreak IDs (list column)
  grp2ids <- subsets[, .(id_outbreak = list(unique(id_outbreak))), by = grp]
    
  # Compute & save per priority (flat lapply, no nested loops) ----------
  for (pri in priorities) {
    for (p in probabilities) {
      if (p > 0 || (p == 0 & pri == "AR")) {
        message(paste0("priority: " , pri, ", prob: ", p))
        grp_p <- meta[(vacc_priority == pri & prob == p), grp]
    
        res_list <- lapply(grp_p, compute_one_grp, cea_data)
      
        subsets_cea_priority <- rbindlist(res_list, fill = TRUE)
        saveRDS(
          subsets_cea_priority,
          sprintf("outputs/cea_subsets_vacc_week_%d_%s_prob_%.1f_onset_trigger_%s.rds", 
                  vw, pri, p, tstamp())
        )
      }
    }
  }
}
```

parallel version
```{r}
# ---- setup for parallel ----
plan(multisession)  # Windows-safe; on Linux/macOS you may use plan(multicore)

priorities <- c("AR", "dur", "size")
ocv_doses  <- c(1e5, 2e5, 5e5, 1e6, 2e6, 5e6, 10e6, 20e6, 50e6, 100e6, 200e6)
cea_files  <- c("outputs/paper/cea_onset_trigger_20251031.rds")

cea_75 <- map_dfr(cea_files, readr::read_rds) %>% filter(vacc_cov == 0.75)

# ---- worker (explicit args to avoid globals) ----
compute_one_grp <- function(g, cea_data, grp2ids, meta) {
  ids <- grp2ids[grp == g, id_outbreak][[1L]]
  if (length(ids) == 0L) return(NULL)

  # keyed join (fast)
  cea_sub <- cea_data[.(ids), nomatch = 0L]
  cea_sub_ori <- cea_sub[ori_occurred == TRUE]

  # no id_outbreak in group_cols to compute across outbreaks in a subset
  out <- sum_over_outbreaks_by_runid(
    d = cea_sub_ori,
    group_cols = c("runid", "vacc_week",
                   "vacc_cov", "no_ori_delay_outbreak_end")
  )

  # attach metadata for this group
  m <- meta[grp == g]
  out[, `:=`(
    subset_id          = m$id_sample,
    vacc_priority      = m$vacc_priority,
    priority_prob      = m$prob,
    target_vacc_doses  = m$ocv_dose,
    outbreak_count     = m$outbreak_count,
    doses_used         = m$doses_used
  )]

  out[]
}
# to avoid floating point issues when comparing probabilities
canon_prob <- function(x) as.integer(round(x * 100))
# ---------- main ----------
vacc_weeks    <- c(1, 5, 9, 13)
probabilities <- seq(0, 0.9, 0.1)

for (vw in vacc_weeks) {
  message(paste0("vacc week: ", vw))

  cea_data <- cea_75 %>% filter(vacc_week == vw)
  setDT(cea_data)
  setkey(cea_data, id_outbreak)  # enables fast [.(ids)] joins

  grid <- CJ(vacc_priority = priorities, ocv_dose = ocv_doses)
  grid[, fn := ifelse(
    ocv_dose < 1e6,
    sprintf("outputs/subsets_%sK_%s_20251104.rds", round(ocv_dose/1e3), vacc_priority),
    sprintf("outputs/subsets_%sM_%s_20251104.rds", round(ocv_dose/1e6), vacc_priority)
  )]

  subsets <- rbindlist(lapply(grid$fn, readr::read_rds), use.names = TRUE)
  setDT(subsets)
  setkey(subsets, vacc_priority, ocv_dose, vacc_cov, prob, id_sample)

  # group id per (priority, dose, prob, id_sample)
  subsets[, grp := .GRP, by = .(vacc_priority, ocv_dose, prob, id_sample)]
  # metadata per group (one row each)
  meta <- unique(subsets[, .(grp, vacc_priority, ocv_dose, prob, id_sample,
                             outbreak_count, doses_used)])
  meta[, prob_i := canon_prob(prob)]
  # map each group to its outbreak id vector
  grp2ids <- subsets[, .(id_outbreak = list(unique(id_outbreak))), by = grp]
  
  subsets_cea_priority <- vector('list', length(priorities)*length(probabilities))
  ii <- 1
  for (pri in priorities) {
    for (p in probabilities) {
      p_i <- canon_prob(p)
      if (p_i > 0 || (p_i == 0 & pri == "AR")) {
        message(paste0("priority: ", pri, ", prob_i: ", p_i))
        grp_p <- meta[(meta$vacc_priority == pri & prob_i == p_i), grp]
        
        res_list <- future_lapply(
          grp_p,
          compute_one_grp,
          cea_data = cea_data,
          grp2ids  = grp2ids,
          meta     = meta,
          future.globals  = TRUE,
          future.packages = c("data.table"),
          future.seed      = TRUE,
          future.scheduling = 1
        )
        subsets_cea_priority[[ii]] <- rbindlist(res_list)
        ii <- ii + 1
        # subsets_cea_priority <- rbindlist(res_list)
        # saveRDS(
        #   subsets_cea_priority,
        #   sprintf("outputs/cea_subsets_vacc_week_%d_%s_%.1f_onset_trigger_%s.rds",
        #           vw, pri, p, tstamp())
        # )
      }
    }
  }
  saveRDS(
          subsets_cea_priority,
          sprintf("outputs/cea_subsets_vacc_week_%d_onset_trigger_%s.rds",
                  vw, tstamp())
  )
}

plan(sequential)
```

## Derived statistics 

Compute summary statistics (mean and quantiles) for variables across samples by grouping relevant variables.

We group by subset_id in addition to other key variables (vacc_cov, vacc_week, target_vacc_doses, and vacc_priority). The resulting variability in outcomes therefore reflects differences across simulation runs (runid) for the same set of outbreaks (i.e., the same subset_id). This approach ensures that comparable vaccine doses are deployed within each subset.

Test analysis
```{r}
d <- read_rds(
  "outputs/cea_subsets_vacc_week_1_AR_0.0_onset_trigger_20251105.rds"
  )
d_summary <- d %>% 
  build_summary(value_col = summary_cols, by=group_var)

names(d_summary)
unique(d_summary$value_column)

d_summary %>% 
  filter(value_column == "s_ch_averted_tot") %>% 
  group_by(vacc_week, target_vacc_doses, priority_prob) %>% 
  summarize(ca_m = median(q500))
```

```{r}
group_var <- c("vacc_cov", "vacc_week", "target_vacc_doses", 
  "vacc_priority", "priority_prob", "subset_id")

summary_cols <- c("pct_reduc_case", 
                  "n_outbreaks", "outbreak_count",
                  "vacc_dose", "doses_used",
                  "s_ch_averted_tot",  "death_averted_tot",
                  "daly_averted_tot",
                  "case_averted_per_1000_OCV", "death_averted_per_1000_OCV",
                  "daly_averted_per_1000_OCV",
                  "cost_per_daly_averted",
                  "cost_per_case_averted", "cost_per_death_averted")

priorities <- c("_AR_", "_dur_", "_size_")
dates <- c("_onset_trigger_20251105.rds")
vacc_weeks <- c(1, 9)
probabilities <- c("0.0", "0.3")

for (vw in vacc_weeks) {
  message(paste0("vacc week:", vw))
#  for (p in probabilities) {
    
    ids_grids <- expand.grid(vw, priorities, probabilities, dates)
    ids_combined <- apply(ids_grids, 1, paste, collapse = "")
  
    files <- unlist(lapply(ids_combined, function(x) {
      f <- paste0("outputs/cea_subsets_vacc_week_", x)
      f[file.exists(f)]
    }))
    print(files)
    
    ii <- 1
    cea_summary_subsets_list <- list()
    for (file in files) {
      cea_summary_subsets_list[[ii]] <- 
        map_dfr(file, read_rds) %>%
        build_summary(value_col = summary_cols, by=group_var)
      ii <- ii + 1
    }
#        cea_summary_subsets <- rbindlist(cea_summary_subsets_list) 
#        rm(cea_summary_subsets_list)
#        saveRDS(cea_summary_subsets, 
#                paste0("outputs/paper/cea_summary_subsets_vw_", vw,
#                       "_prob_", p, "_", tstamp(),".rds")
#        )
    cea_summary_subsets <- rbindlist(cea_summary_subsets_list) 
    rm(cea_summary_subsets_list)
    saveRDS(cea_summary_subsets, 
            paste0("outputs/paper/cea_summary_subsets_vw_", vw,
                    "_", tstamp(),".rds")
    )
#    }
}
```

# Exponential decy fit

Fit an exponential decay function for percentage case reduction across delay in ORI implementation

## Outbreak onset-triggered
```{r}
cea_weekly_75 <- 
  read_rds("outputs/paper/cea_delay_onset_trigger_20251031.rds")
cea_data <- 
  cea_weekly_75 %>% filter(no_ori_delay_outbreak_end == 2)
rm(cea_weekly_75)

dat <- 
  left_join(cea_data[, c("vacc_week", "pct_reduc_case", "id_outbreak", 
                         "week_delay_to_vacc_effect", "runid", "vacc_cov",
                         "ori_occurred")],
            outbk_data_summary[, c("ar_per_10e3", "duration_weeks",
                                  "spatial_scale", "rural_urban",
                                  "id_outbreak")],
            by="id_outbreak")

# some outbreaks show no decrease and we treat them separate from fitting a decay model
is_constant_impact <- function(df, y_var = "pct_reduc_case") {
  sd(df[[y_var]], na.rm = TRUE) < 1e-6
}

fit_or_constant <- function(df) {
  if (is_constant_impact(df)) {
    tibble(k_hat = 0, half_life = Inf, constant = TRUE)
  } else {
    fit <- fit_decay_model(df, 
                           y_var = "pct_reduc_case", 
                           x_var = "vacc_week")
    fit$constant <- FALSE
    fit
  }
}

summary(outbk_data_summary$duration_weeks)
vacc_week_cutoffs <- c(5, 9, 14, 21) # min, Q1,Q2,Q3

res <- vector('list', length(vacc_week_cutoffs))

for (cutoff in vacc_week_cutoffs) {

  res[[as.character(cutoff)]] <-
    dat %>%
    filter(vacc_week <= cutoff) %>%
    group_by(runid, id_outbreak) %>%
    filter(all(ori_occurred)) %>% 
    select(runid, id_outbreak, vacc_week, pct_reduc_case) %>% 
    nest() %>%
    mutate(fit = map(data, fit_or_constant)) %>%
    unnest(fit) %>%
    ungroup()
}

saveRDS(res, paste0("outputs/paper/fit_exp_decay_", tstamp(), ".rds"))
```

## Case-triggered

Fit an exponential decay function for percentage case reduction across delay in case trigger for an 9-week delay in ORI implementation

```{r}
# some outbreaks show no decrease and we treat them separate from fitting a decay model
is_constant_impact <- function(df, y_var = "pct_reduc_case") {
  sd(df[[y_var]], na.rm = TRUE) < 1e-6
}

fit_or_constant <- function(df) {
  if (is_constant_impact(df)) {
    tibble(k_hat = 0, half_life = Inf, constant = TRUE)
  } else {
    fit <- fit_decay_model(df, 
                           y_var = "pct_reduc_case", 
                           x_var = "case_trigger")
    fit$constant <- FALSE
    fit
  }
}

# Case-triggered vaccination results
cea_ct <- read_rds("outputs/paper/cea_case_trigger_20251031.rds")
dat_ct <- 
  left_join(cea_ct[, c("vacc_week", "pct_reduc_case", "id_outbreak", 
                       "runid", "case_trigger", "ori_occurred", 
                       "vacc_cov", "week_delay_to_vacc_effect")],
               outbk_data_summary[, c("ar_per_10e3", "duration_weeks", 
                                  "spatial_scale", "rural_urban", 
                                  "id_outbreak")],
               by="id_outbreak")

# exponential decay
dat_ct %>% 
  filter(vacc_cov == 0.75, vacc_week %in% c(1,5)) %>% 
  group_by(case_trigger, vacc_week) %>% 
  summarize(m_pca = median(pct_reduc_case)) %>% 
  ggplot(aes(case_trigger, m_pca))+
  geom_line()+
  facet_wrap(~vacc_week)

# fit exponential decay model
vacc_weeks <- c(1, 5)
case_cutoff <- 90 # only use

# Apply decay model to each group
res <- list()
for (vw in vacc_weeks) {
  res[[as.character(vw)]] <- 
    dat_ct %>%
    filter(vacc_week == vw, case_trigger <= case_cutoff) %>%
    group_by(runid, id_outbreak) %>%
    filter(all(ori_occurred)) %>%
    select(runid, id_outbreak, case_trigger, pct_reduc_case) %>% 
    nest() %>%
    mutate(fit = map(data, fit_or_constant)) %>%
    unnest(fit) %>%
    ungroup()
}

saveRDS(res, paste0("outputs/paper/fit_exp_decay_case_trigger_",
                    tstamp(), ".rds"))
```


# Linear decrease in the proportion of ORI-eligible outbreaks 

```{r}
wks_delay <- seq(1, 15, 1)
dur_missed_id <- vector('list', length(wks_delay))
vax_missed_id <- vector('list', length(wks_delay))

for (i in 1:length(wks_delay)) {
  outbk_data_summary %>% 
    filter(duration_weeks <= wks_delay[i]) %>%
    pull(id_outbreak) -> dur_missed_id[[i]]
  
  cea_weekly_delay_2 %>% 
    filter(vacc_week == wks_delay[i] & runid == 1) %>% 
    filter(!ori_occurred) %>% 
    pull(id_outbreak) -> vax_missed_id[[i]]
}
n_outbreaks_missed1 <- map_vec(dur_missed_id, function(i) length(i))
n_outbreaks_missed2 <- map_vec(vax_missed_id, function(i) length(i))

n_outbreaks <- nrow(outbk_data_summary)
prop_ori_outbreaks <- data.frame(
  vacc_week = seq(1, 15, 1), 
  prop_included_dur = (1 - n_outbreaks_missed1 / n_outbreaks),
  prop_included_vax = (1 - n_outbreaks_missed2 / n_outbreaks))

df <- prop_ori_outbreaks %>%
  select(vacc_week, prop_included_dur) %>%
  filter(vacc_week >= 4) 
  # longer than 5 weeks

# --- fit linear model: percent-points per week ---
fit_lm <- lm(prop_included_dur ~ vacc_week, data = df)

saveRDS(fit_lm, paste0("outputs/paper/fit_lm_", tstamp(), ".rds"))
```

