---
title: "Vaccine impact modeling"
format: html
editor_options: 
  chunk_output_type: console
---

## Load packages and source `utils.R`
```{r}
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
source("utils.R") # utility functions
library(conflicted)
conflict_prefer("filter", "dplyr")
VC3 <- c(0.6, 0.75, 0.9) # representative vaccine coverage rates
VC <- VC3[2]
VW <- 9 # representative vaccination week
VW_seq <- seq(1,19,2)
PROBS <- c(0.025,0.25,0.5,0.75,0.975)
save_csv <- FALSE
```

## Load outbreak datasets
```{r}
# extract ids for the outbreaks not affected by OCV campaigns
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Oct2024/outbreak_summary.csv")))
paste0(length(unique(ds$country)), " countries")
# create 'id_outbreak' variables to link with the 
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# data for the outbreaks affected by OCV
docv <- clean_names(fread("data/outbreak_files_Oct2024/ocv_long_dataset.csv"))
# time-series data
ts <- clean_names(readRDS("data/outbreak_files_Oct2024/time_series_outbreak_extraction.rds")) 
ts |>
  group_by(location, outbreak_number) |> # identify unique outbreaks
  mutate(week = row_number(),
         date = tl + 3, # pick the date for the midpoint for each week
         start_date_outbreak = min(tl), 
         end_date_outbreak = max(tr),
         id_outbreak = paste0(location, "-", start_date_outbreak, "-",
                              end_date_outbreak)) |>
  ungroup() -> ts # time series with unique ids

paste0(length(unique(ts$id_outbreak)), " unique outbreaks")
paste0(sum(unique(ts$id_outbreak) %in% ds$id_outbreak), 
       " unique outbreaks matched between time series and outbreak summary")

paste0(length(unique(ts$country)), " countries")

# ids for the outbreaks with confirmed dataset
ts |> 
  group_by(id_outbreak) |> 
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique() -> ids_conf

paste0(length(ids_conf), " confirmed outbreaks")
paste0(length(unique(ts$id_outbreak)) - length(ids_conf), " suspected outbreaks")
# data.table::fwrite(ts, paste0("outputs/outbreak_data_all_", tstamp(), ".csv"))
if (save_csv) {
  fwrite(ts, paste0("outputs/outbreak_data_all_", tstamp(), ".csv"))
}

paste0(sum((unique(ts$id_outbreak)) %in% docv$id_outbreak), 
      " outbreaks were affected by OCVs")

ts_noocv <- ts[!(ts$id_outbreak %in% docv$id_outbreak), ]

if (save_csv) {
  fwrite(ts_noocv, paste0("outputs/outbreak_data_", tstamp(), ".csv"))
}
paste0(sum((unique(ts_noocv$id_outbreak)) %in% docv$id_outbreak), 
      " outbreaks from modeled data were affected by OCVs")

ts_noocv |> 
  group_by(id_outbreak) |> 
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique() -> ids_conf_noocv

paste0(length(ids_conf_noocv), " confirmed outbreaks out of ", length(unique(ts_noocv$id_outbreak)), " outbreaks unaffected by OCVs")
paste0(length(unique(ts_noocv$id_outbreak)) - length(ids_conf_noocv), " suspected outbreaks")

outbreak_data_all <- fread("outputs/outbreak_data_all_20241121.csv")
outbreak_data <- fread("outputs/outbreak_data_20241121.csv")
# outbreak_data <- outbreak_data_all[!(id_outbreak %in% docv$id_outbreak)]

```

## Identify unique outbreaks
Primary analyses were 
1. overlapping outbreaks 
Find the lowest admin unit
Check if any duplicate outbreaks that reported in the same admin unit
```{r}
unique(outbreak_data$spatial_scale)
outbreak_data |> 
  group_by(spatial_scale) |> 
  summarize()

d <- outbreak_data[week == 1]
d[, .N, by="spatial_scale"]

d |> 
  filter(spatial_scale == "admin2") |> 
  pull(admin2) |> 
  unique() -> adm2

d_country <- d[spatial_scale == "country"]
d_adm1 <- d[spatial_scale == "admin1"]
d_adm2 <- d[spatial_scale == "admin2"] 
d_adm3 <- d[spatial_scale == "admin3"] 

# ids <- c()
# for (i in 1:nrow(d_adm2)) {
#   d2 <- d_adm2[i, ]
#   d3 <- d_adm3[admin2 == d2$admin2] # same admin2 
#   if (nrow(d3) > 0) {
#     min_tl <- min(d3$tl)
#     max_tr <- max(d3$tr)  
#     if (d2$tl >= min_tl & d2$tr <= max_tr) { # d_Adm3 includes d_adm2 in terms of dates
#       cat(sprintf("id: %d, nrow: %d\n", i, nrow(d3)))
#       ids <- c(ids, i)
#     }
#   }
# }
# d_adm2_dup <- d_adm2[ids,]
# 
# ids_adm1

get_duplicated_outbreaks <- function(d_upper, d_lower, string=NULL){
  indices <- c()
  for (i in 1:nrow(d_upper)) {
    dup <- d_upper[i, ]
    idx <- d_lower[[string]] == dup[[string]]
    dlow <- d_lower[idx,]
    if (nrow(dlow) > 0) {
      if (dup$tl >= min(dlow$tl) & dup$tr <= max(dlow$tr)) {
        # cat(sprintf("id: %d, nrow: %d\n", i, nrow(d3)))
        indices <- c(indices, i)
      }
    }
  }
  return(d_upper[indices,])
}

adm2_dup <- get_duplicated_outbreaks(d_upper=d_adm2, d_lower=d_adm3, string="admin2")
adm1_dup <- get_duplicated_outbreaks(d_upper=d_adm1, d_lower=d_adm3, string="admin1")
adm0_dup <- get_duplicated_outbreaks(d_upper=d_country, d_lower=d_adm3, string="country")
adm1_dup2 <- get_duplicated_outbreaks(d_upper=d_adm1, d_lower=d_adm2, string="admin1")
adm0_dup2 <- get_duplicated_outbreaks(d_upper=d_country, d_lower=d_adm2, string="country")
adm0_dup3 <- get_duplicated_outbreaks(d_upper=d_country, d_lower=d_adm1, string="country")

dup <- rbind(adm2_dup, adm1_dup, adm1_dup2, adm0_dup, adm0_dup2, adm0_dup3)
dup <- dup[!duplicated(dup),]

d <- outbreak_data[!(id_outbreak %in% dup$id_outbreak)]
table(outbreak_data[week == 1, "spatial_scale"])
table(d[week == 1, "spatial_scale"])
```


## Parameter sets need to compute the vaccine impact
```{r}
params <- as.data.frame(data.table::fread("outputs/params_20241017.csv"))
# IVE posterior predictive values from Bayesian beta regression (see parameters.qmd)
ive_fit <- read_rds("outputs/ive_yrep_20241010.rds")
ive_data <- ive_fit$yrep
rm(ive_fit)
pop_by_age_wpp <- as.data.frame(fread("outputs/wpp_pop_by_age_20241022.csv"))
life_exp_data <- as.data.frame(fread("outputs/wpp_life_expectancy_20241022.csv"))
gdp_data <- read_xls("data/GDP_WorldBank.xls")
workforce_data <- read_xls("data/Workforce_Worldbank.xls")
outbreak_data <- as.data.frame(data.table::fread("outputs/outbreak_data_20241121.csv"))
outbreak_data_all <- as.data.frame(data.table::fread("outputs/outbreak_data_all_20241121.csv"))
```

## Functions to summarize vaccine impact
```{r}
# exclude outbreaks with OCVs, add population size based on ts
# global variables need to be set: VC3, ts, docv
set_svim_raw_res <- function(svim_df) {
  svim <- svim_df[vacc_cov %in% VC3] # we only use three vacc_covs
  svim <- svim[!(id_outbreak %in% docv$id_outbreak)] # outbreaks for which no ocv's affected
  tsdt <- setDT(ts)[week == 1]
  svim <- svim[tsdt, on ="id_outbreak", population := i.pop]
  svim <- svim[, 
               c("s_ch_vacc", "outbk_yesno", "outbk_vacc_yesno") :=
                 list(s_ch - s_ch_averted,
                  ifelse(1e5 * s_ch / population > threshold, 1, 0),
                  ifelse(1e5 * (s_ch - s_ch_averted) / population > threshold, 1, 0))]
  return(svim)
}

# sum across week in terms of cases, deaths, ...
svim_sum_across_week <- function(svim, case_trigger=FALSE) {
  if (!case_trigger) {
    sum_across_week_by_runid <- svim[, .(
      country = country[1],
      start_date = as.Date(start_date_outbreak[1]),
      end_date = as.Date(end_date_outbreak[1]),
      prop_u5 = prop_u5[1],
      population = population[1],
      week_delay_to_vacc_effect = week_delay_to_vacc_effect[1],
      threshold = threshold[1],
      ori_occurred = ori_occurred[1],
      ive = ive[1],
      c_ch_tot = sum(c_ch, na.rm=T),
      death_tot = sum(deaths, na.rm=T),
      death_averted_tot = sum(deaths_averted, na.rm=T),
      pct_reduc_death = 100 * sum(deaths_averted, na.rm=T)/sum(deaths, na.rm=T),
      s_ch_averted_u5_tot = sum(s_ch_averted_u5), # weekly to cumulative
      s_ch_averted_5p_tot = sum(s_ch_averted_5p),
      s_ch_tot = sum(s_ch),
      s_ch_averted_tot = sum(s_ch_averted),
      pct_reduc_case = 100 * sum(s_ch_averted)/sum(s_ch),
      outbk_dur = sum(outbk_yesno),
      outbk_dur_vacc = sum(outbk_vacc_yesno),
      dur_reduced = sum(outbk_yesno) - sum(outbk_vacc_yesno),
      pct_reduc_dur = 100*((sum(outbk_yesno) - sum(outbk_vacc_yesno))/sum(outbk_yesno))),
      by = .(runid, id_outbreak, vacc_week, vacc_cov)]
  }
  else if (case_trigger) {
  sum_across_week_by_runid <- svim[, .(
    country = country[1],
    start_date = as.Date(start_date_outbreak[1]),
    end_date = as.Date(end_date_outbreak[1]),
    prop_u5 = prop_u5[1],
    population = population[1],
    vacc_week = vacc_week[1],
    week_delay_to_vacc_effect = week_delay_to_vacc_effect[1],
    threshold = threshold[1],
    ori_occurred = ori_occurred[1],
    ive = ive[1],
    c_ch_tot = sum(c_ch, na.rm=T),
    death_tot = sum(deaths, na.rm=T),
    death_averted_tot = sum(deaths_averted, na.rm=T),
    pct_reduc_death = 100 * sum(deaths_averted, na.rm=T)/sum(deaths, na.rm=T),
    s_ch_averted_u5_tot = sum(s_ch_averted_u5), # weekly to cumulative
    s_ch_averted_5p_tot = sum(s_ch_averted_5p),
    s_ch_tot = sum(s_ch),
    s_ch_averted_tot = sum(s_ch_averted),
    pct_reduc_case = 100 * sum(s_ch_averted)/sum(s_ch),
    outbk_dur = sum(outbk_yesno),
    outbk_dur_vacc = sum(outbk_vacc_yesno),
    dur_reduced = sum(outbk_yesno) - sum(outbk_vacc_yesno),
    pct_reduc_dur = 100*((sum(outbk_yesno) - sum(outbk_vacc_yesno))/sum(outbk_yesno))),
    by = .(runid, id_outbreak, case_trigger, vacc_cov)]
  }
  return(sum_across_week_by_runid)
}

# compute mean across runid
svim_mean_across_runid <- function(sum_across_week_by_runid, case_trigger=FALSE){
  if (!case_trigger) {
    mean_across_runid <- sum_across_week_by_runid[, .(
      country = country[1],
      start_date = start_date[1],
      end_date = end_date[1],
      prop_u5 = prop_u5[1],
      population = population[1],
      week_delay_to_vacc_effect = week_delay_to_vacc_effect[1],
      threshold = threshold[1],
      ori_occurred = ori_occurred[1],
      c_ch_tot = mean(c_ch_tot),
      ive = mean(ive),
      s_ch_averted_u5_tot = mean(s_ch_averted_u5_tot), # weekly to cumulative
      s_ch_averted_5p_tot = mean(s_ch_averted_5p_tot),
      s_ch_tot = mean(s_ch_tot),
      s_ch_averted_tot = mean(s_ch_averted_tot),
      pct_reduc_case = mean(pct_reduc_case),
      death_tot = mean(death_tot),
      death_averted_tot = mean(death_averted_tot), 
      pct_reduc_death = mean(pct_reduc_death),
      outbk_dur = mean(outbk_dur),
      outbk_dur_vacc = mean(outbk_dur_vacc),
      dur_reduced = mean(dur_reduced),
      pct_reduc_dur = mean(pct_reduc_dur)),
      by = .(id_outbreak, vacc_week, vacc_cov)]
  }
  else if (case_trigger) {
        mean_across_runid <- sum_across_week_by_runid[, .(
      country = country[1],
      start_date = start_date[1],
      end_date = end_date[1],
      prop_u5 = prop_u5[1],
      population = population[1],
      vacc_week = vacc_week[1],
      week_delay_to_vacc_effect = week_delay_to_vacc_effect[1],
      threshold = threshold[1],
      ori_occurred = ori_occurred[1],
      c_ch_tot = mean(c_ch_tot),
      ive = mean(ive),
      s_ch_averted_u5_tot = mean(s_ch_averted_u5_tot), # weekly to cumulative
      s_ch_averted_5p_tot = mean(s_ch_averted_5p_tot),
      s_ch_tot = mean(s_ch_tot),
      s_ch_averted_tot = mean(s_ch_averted_tot),
      pct_reduc_case = mean(pct_reduc_case),
      death_tot = mean(death_tot),
      death_averted_tot = mean(death_averted_tot), 
      pct_reduc_death = mean(pct_reduc_death),
      outbk_dur = mean(outbk_dur),
      outbk_dur_vacc = mean(outbk_dur_vacc),
      dur_reduced = mean(dur_reduced),
      pct_reduc_dur = mean(pct_reduc_dur)),
      by = .(id_outbreak, case_trigger, vacc_cov)]
  }
  return(mean_across_runid)
}

# compute sd across runid
svim_sd_across_runid <- function(sum_across_week_by_runid, case_trigger=FALSE) {
  
  if (!case_trigger) {
    sd_across_runid <- sum_across_week_by_runid[, .(
      country = country[1],
      start_date = start_date[1],
      end_date = end_date[1],
      prop_u5 = prop_u5[1],
      population = population[1],
      week_delay_to_vacc_effect = week_delay_to_vacc_effect[1],
      threshold = threshold[1],
      ori_occurred = ori_occurred[1],
      c_ch_tot = sd(c_ch_tot),
      ive = sd(ive),
      s_ch_averted_u5_tot = sd(s_ch_averted_u5_tot), # weekly to cumulative
      s_ch_averted_5p_tot = sd(s_ch_averted_5p_tot),
      s_ch_tot = sd(s_ch_tot),
      s_ch_averted_tot = sd(s_ch_averted_tot),
      pct_reduc_case = sd(pct_reduc_case),
      death_tot = sd(death_tot),
      death_averted_tot = sd(death_averted_tot), 
      pct_reduc_death = sd(pct_reduc_death),
      outbk_dur = sd(outbk_dur),
      outbk_dur_vacc = sd(outbk_dur_vacc),
      dur_reduced = sd(dur_reduced),
      pct_reduc_dur = sd(pct_reduc_dur)),
      by = .(id_outbreak, vacc_week, vacc_cov)]
  } 
  else if(case_trigger) {
    sd_across_runid <- sum_across_week_by_runid[, .(
      country = country[1],
      start_date = start_date[1],
      end_date = end_date[1],
      prop_u5 = prop_u5[1],
      population = population[1],
      vacc_week = vacc_week[1],
      week_delay_to_vacc_effect = week_delay_to_vacc_effect[1],
      threshold = threshold[1],
      ori_occurred = ori_occurred[1],
      c_ch_tot = sd(c_ch_tot),
      ive = sd(ive),
      s_ch_averted_u5_tot = sd(s_ch_averted_u5_tot), # weekly to cumulative
      s_ch_averted_5p_tot = sd(s_ch_averted_5p_tot),
      s_ch_tot = sd(s_ch_tot),
      s_ch_averted_tot = sd(s_ch_averted_tot),
      pct_reduc_case = sd(pct_reduc_case),
      death_tot = sd(death_tot),
      death_averted_tot = sd(death_averted_tot), 
      pct_reduc_death = sd(pct_reduc_death),
      outbk_dur = sd(outbk_dur),
      outbk_dur_vacc = sd(outbk_dur_vacc),
      dur_reduced = sd(dur_reduced),
      pct_reduc_dur = sd(pct_reduc_dur)),
      by = .(id_outbreak, case_trigger, vacc_cov)]
  }
  
  return(sd_across_runid)
}

# across outbreak week
svim_sum_across_week_all_outbreaks <- function(svim, case_trigger=FALSE) {
  # variable names change (s is droped from deaths): deaths_averted -> death_averted_tot, 
  # deaths -> death_tot
  if (!case_trigger) {
  sum_across_week_by_runid <- svim[, .(
    country = country[1],
    start_date = as.Date(start_date_outbreak[1]),
    end_date = as.Date(end_date_outbreak[1]),
    prop_u5 = prop_u5[1],
    population = population[1],
    ori_occurred = ori_occurred[1],
    c_ch_tot = sum(c_ch, na.rm=T),
    death_tot = sum(deaths, na.rm=T),
    death_averted_tot = sum(deaths_averted, na.rm=T),
    s_ch_averted_u5_tot = sum(s_ch_averted_u5), # weekly to cumulative
    s_ch_averted_5p_tot = sum(s_ch_averted_5p),
    s_ch_tot = sum(s_ch),
    s_ch_averted_tot = sum(s_ch_averted),
    outbk_dur = sum(outbk_yesno),
    outbk_dur_vacc = sum(outbk_vacc_yesno)),
    
    by = .(runid, id_outbreak, vacc_week, vacc_cov)]
  }
  else if (case_trigger) {
   sum_across_week_by_runid <- svim[, .(
    country = country[1],
    start_date = as.Date(start_date_outbreak[1]),
    end_date = as.Date(end_date_outbreak[1]),
    prop_u5 = prop_u5[1],
    population = population[1],
    ori_occurred = ori_occurred[1],
    c_ch_tot = sum(c_ch, na.rm=T),
    death_tot = sum(deaths, na.rm=T),
    death_averted_tot = sum(deaths_averted, na.rm=T),
    s_ch_averted_u5_tot = sum(s_ch_averted_u5), # weekly to cumulative
    s_ch_averted_5p_tot = sum(s_ch_averted_5p),
    s_ch_tot = sum(s_ch),
    s_ch_averted_tot = sum(s_ch_averted),
    outbk_dur = sum(outbk_yesno),
    outbk_dur_vacc = sum(outbk_vacc_yesno)),
    
    by = .(runid, id_outbreak, case_trigger, vacc_cov)]
  }
  return(sum_across_week_by_runid)
}

# cea variables for each outbreak
add_cea_results_all_outbreaks <- function(svim) {
# CEA variables are added but ratio measure need to be computed when values are summed across outbreaks
  svim |>
    mutate(yld = compute_yld(s_ch_tot, parms=parms),
           yld_averted = compute_yld(s_ch_averted_tot, parms=parms),
           yll = compute_yll(death_tot, life_exp, parms),
           yll_averted = compute_yll(death_averted_tot, life_exp, parms), 
           daly_averted = yld_averted + yll_averted,
           coi_averted = s_ch_averted_tot * coi_per_patient,
           cod_averted = death_averted_tot * gdp * life_exp,
           productivity_lost_averted = 
             s_ch_averted_tot * gdp * ((patient_workday_lost/365)*(pct_workforce/100) +
                                   (caregiver_workday_lost/365)),
           vacc_dose = ifelse(ori_occurred, population * vacc_cov * dose_regimen, 0),
           vacc_cost = (vacc_cost_per_dose + vacc_delivery_cost) * vacc_dose,
           net_cost = vacc_cost - coi_averted - cod_averted - productivity_lost_averted) -> svim_cea
  
  return(svim_cea)
}
# across outbreaks
svim_sum_across_outbreaks_by_runid <- 
  function(sum_across_week_by_runid, case_trigger=FALSE) {
    if (!case_trigger) {
      all_outbreaks_by_runid <- sum_across_week_by_runid[, .(  
        population = sum(population), 
        vacc_dose = sum(ifelse(ori_occurred, population * vacc_cov, 0)),
        c_ch_tot = sum(c_ch_tot, na.rm=T),
        death_tot = sum(death_tot, na.rm=T),
        death_averted_tot = sum(death_averted_tot, na.rm=T),
        pct_reduc_death = 100 * sum(death_averted_tot, na.rm=T)/sum(death_tot, na.rm=T),
        s_ch_averted_u5_tot = sum(s_ch_averted_u5_tot), # weekly to cumulative
        s_ch_averted_5p_tot = sum(s_ch_averted_5p_tot),
        s_ch_tot = sum(s_ch_tot),
        s_ch_averted_tot = sum(s_ch_averted_tot),
        pct_reduc_case = 100 * sum(s_ch_averted_tot) / sum(s_ch_tot),
        outbk_dur = sum(outbk_dur),
        outbk_dur_vacc = sum(outbk_dur_vacc),
        dur_reduced = sum(outbk_dur) - sum(outbk_dur_vacc),
        pct_reduc_dur = 100*((sum(outbk_dur) - sum(outbk_dur_vacc)) / sum(outbk_dur)),
        case_averted_per_1000_OCV =
          1000 * sum(s_ch_averted_tot) / sum(vacc_dose),
        death_averted_per_1000_OCV = 
          1000 * sum(death_averted_tot, na.rm=T) /sum(vacc_dose),
        cost_per_daly_averted = sum(net_cost, na.rm=T) / sum(daly_averted, na.rm=T),
        cost_per_case_averted = sum(net_cost, na.rm=T) / sum(s_ch_averted_tot, na.rm=T),
        cost_per_death_averted = sum(net_cost, na.rm=T) / sum(death_averted_tot, na.rm=T)),
        by = .(runid, vacc_week, vacc_cov)] 
    }
    else if(case_trigger) {
            all_outbreaks_by_runid <- sum_across_week_by_runid[, .(  
        population = sum(population), 
        vacc_dose = sum(ifelse(ori_occurred, population * vacc_cov, 0)),
        c_ch_tot = sum(c_ch_tot, na.rm=T),
        death_tot = sum(death_tot, na.rm=T),
        death_averted_tot = sum(death_averted_tot, na.rm=T),
        pct_reduc_death = 100 * sum(death_averted_tot, na.rm=T)/sum(death_tot, na.rm=T),
        s_ch_averted_u5_tot = sum(s_ch_averted_u5_tot), # weekly to cumulative
        s_ch_averted_5p_tot = sum(s_ch_averted_5p_tot),
        s_ch_tot = sum(s_ch_tot),
        s_ch_averted_tot = sum(s_ch_averted_tot),
        pct_reduc_case = 100 * sum(s_ch_averted_tot) / sum(s_ch_tot),
        outbk_dur = sum(outbk_dur),
        outbk_dur_vacc = sum(outbk_dur_vacc),
        dur_reduced = sum(outbk_dur) - sum(outbk_dur_vacc),
        pct_reduc_dur =
          100*((sum(outbk_dur) - sum(outbk_dur_vacc)) / sum(outbk_dur)),
        case_averted_per_1000_OCV =
          1000 * sum(s_ch_averted_tot) / sum(vacc_dose),
        death_averted_per_1000_OCV = 
          1000 * sum(death_averted_tot, na.rm=T) /sum(vacc_dose),
        cost_per_daly_averted = 
          sum(net_cost, na.rm=T) / sum(daly_averted, na.rm=T),
        cost_per_case_averted = 
          sum(net_cost, na.rm=T) / sum(s_ch_averted_tot, na.rm=T),
        cost_per_death_averted =
          sum(net_cost, na.rm=T) / sum(death_averted_tot, na.rm=T)),
        by = .(runid, case_trigger, vacc_cov)]
    }
  return(all_outbreaks_by_runid)
}

svim_summary_all_outbreaks <- function(d, case_trigger=FALSE){ 
  if (!case_trigger) {
  summary_all_outbreaks <- d[, 
    .(population = population[1],
      s_ch_tot = s_ch_tot[1],
      c_ch_tot = c_ch_tot[1],
      death_tot = death_tot[1],
      outbk_dur = outbk_dur[1],
      mean_vacc_dose = mean(vacc_dose),
      mean_CA = mean(s_ch_averted_tot), 
      q500_CA = quantile(s_ch_averted_tot, probs=0.5),
      q025_CA = quantile(s_ch_averted_tot, probs=0.025),
      q975_CA = quantile(s_ch_averted_tot, probs=0.975),
      mean_DA = mean(death_averted_tot), 
      q500_DA = quantile(death_averted_tot, probs=0.5),
      q025_DA = quantile(death_averted_tot, probs=0.025),
      q975_DA = quantile(death_averted_tot, probs=0.975),
      means_CAPD = mean(case_averted_per_1000_OCV),
      q500_CAPD = quantile(case_averted_per_1000_OCV, probs=0.5),
      q025_CAPD = quantile(case_averted_per_1000_OCV, probs=0.025),
      q975_CAPD = quantile(case_averted_per_1000_OCV, probs=0.975),
      means_DAPD = mean(death_averted_per_1000_OCV),
      q500_DAPD = quantile(death_averted_per_1000_OCV, probs=0.5),
      q025_DAPD = quantile(death_averted_per_1000_OCV, probs=0.025),
      q975_DAPD = quantile(death_averted_per_1000_OCV, probs=0.975),
      mean_PCR = mean(pct_reduc_case),
      q500_PCR = quantile(pct_reduc_case, probs=0.5),
      q025_PCR = quantile(pct_reduc_case, probs=0.025),
      q975_PCR = quantile(pct_reduc_case, probs=0.975),
      mean_PDR = mean(pct_reduc_death),
      q500_PDR = quantile(pct_reduc_death, probs=0.5),
      q025_PDR = quantile(pct_reduc_death, probs=0.025),
      q975_PDR = quantile(pct_reduc_death, probs=0.975),
      mean_durR = mean(dur_reduced),
      q500_DurR = quantile(dur_reduced, probs=0.5),
      q025_DurR = quantile(dur_reduced, probs=0.025),
      q975_DurR = quantile(dur_reduced, probs=0.975),
      mean_PDurR = mean(pct_reduc_dur),
      q500_PDurR = quantile(pct_reduc_dur, probs=0.5),
      q025_PDurR = quantile(pct_reduc_dur, probs=0.025),
      q975_PDurR = quantile(pct_reduc_dur, probs=0.975)),
    by = .(vacc_week, vacc_cov)]
  }
  else if (case_trigger){
    
  summary_all_outbreaks <- d[, 
    .(population = population[1],
      s_ch_tot = s_ch_tot[1],
      c_ch_tot = c_ch_tot[1],
      death_tot = death_tot[1],
      outbk_dur = outbk_dur[1],
      mean_vacc_dose = mean(vacc_dose),
      mean_CA = mean(s_ch_averted_tot), 
      q500_CA = quantile(s_ch_averted_tot, probs=0.5),
      q025_CA = quantile(s_ch_averted_tot, probs=0.025),
      q975_CA = quantile(s_ch_averted_tot, probs=0.975),
      mean_DA = mean(death_averted_tot), 
      q500_DA = quantile(death_averted_tot, probs=0.5),
      q025_DA = quantile(death_averted_tot, probs=0.025),
      q975_DA = quantile(death_averted_tot, probs=0.975),
      means_CAPD = mean(case_averted_per_1000_OCV),
      q500_CAPD = quantile(case_averted_per_1000_OCV, probs=0.5),
      q025_CAPD = quantile(case_averted_per_1000_OCV, probs=0.025),
      q975_CAPD = quantile(case_averted_per_1000_OCV, probs=0.975),
      means_DAPD = mean(death_averted_per_1000_OCV),
      q500_DAPD = quantile(death_averted_per_1000_OCV, probs=0.5),
      q025_DAPD = quantile(death_averted_per_1000_OCV, probs=0.025),
      q975_DAPD = quantile(death_averted_per_1000_OCV, probs=0.975),
      mean_PCR = mean(pct_reduc_case),
      q500_PCR = quantile(pct_reduc_case, probs=0.5),
      q025_PCR = quantile(pct_reduc_case, probs=0.025),
      q975_PCR = quantile(pct_reduc_case, probs=0.975),
      q500_PDR = quantile(pct_reduc_death, probs=0.5),
      q025_PDR = quantile(pct_reduc_death, probs=0.025),
      q975_PDR = quantile(pct_reduc_death, probs=0.975),
      mean_dur_reduced = mean(dur_reduced),
      q500_DR = quantile(dur_reduced, probs=0.5),
      q025_DR = quantile(dur_reduced, probs=0.025),
      q975_DR = quantile(dur_reduced, probs=0.975),
      mean_PDR = mean(pct_reduc_dur),
      q500_PDR = quantile(pct_reduc_dur, probs=0.5),
      q025_PDR = quantile(pct_reduc_dur, probs=0.025),
      q975_PDR = quantile(pct_reduc_dur, probs=0.975)),
    by = .(case_trigger, vacc_cov)]
  }
  return(summary_all_outbreaks)
}


get_groud_id <- function(x, cutoff) {
    id <- NA
    for(i in (length(cutoff)-1):1) {
      if(!is.na(x) & x >= cutoff[i]){
        id <- i
        break
      }
    }
  return(id)
}

add_summary_stats <- function(svim, case_trigger=FALSE){
  svim <- 
    left_join(svim, 
              ds[, c("location", "spatial_scale", "outbreak_pop", 
                     "weekly_threshold_per_10e5", "total_s_ch", "total_c_ch", 
                     "total_deaths", "duration_weeks", 
  "peak_week_incidence", "time_to_peak_week", "peak_week_incidence_per_100percent", 
  "peak_week_incidence_per_10e3", "mean_r0_1st_week", "area_per_1km2", 
  "ar_per_10e3", "cfr_per_100percent", "id_outbreak")], by="id_outbreak")
  
  # paste0(format_num(nrow(svim)), " rows")
  
  probs <- seq(0,1,by=0.1)
  ar_grp_cutoff <- quantile(ds$ar_per_10e3, probs=probs)
  vc <- unique(svim$vacc_cov)[1] # just to extract one condition
  if (!case_trigger) {
    vw <- unique(svim$vacc_week)[1] # just to extract one condition
    size_grp_cutoff <- quantile(svim[vacc_week == vw & vacc_cov == vc, s_ch_tot], probs=probs)
    dur_grp_cutoff <- quantile(svim[vacc_week == vw & vacc_cov == vc, outbk_dur], probs=probs, na.rm=T)
  }
  else if (case_trigger) {
    ct <- unique(svim$case_trigger)[1] # just to extract one condition
    size_grp_cutoff <- quantile(svim[case_trigger == ct & vacc_cov == vc, s_ch_tot], probs=probs)
    dur_grp_cutoff <- quantile(svim[case_trigger == ct & vacc_cov == vc, outbk_dur], probs=probs, na.rm=T)
  }
  
  svim$size_grp <- sapply(svim$s_ch_tot, function(x) get_groud_id(x, size_grp_cutoff))
  svim$dur_grp <- sapply(svim$outbk_dur, function(x) get_groud_id(x, dur_grp_cutoff))
  svim$ar_grp <- sapply(svim$ar_per_10e3, function(x) get_groud_id(x, ar_grp_cutoff))
  
  return(svim)
}

add_cea_variables <- function(svim) {
  svim$year <- data.table::year(as.Date(svim$start_date))
  svim <- left_join(svim,
            life_exp_data[, c("country","year", as.character(avg_age_inf))], 
                     by = c("country","year"))
  names(svim)[names(svim) == as.character(avg_age_inf)] <- "life_exp"
  svim$life_exp <- as.numeric(svim$life_exp)
  svim <- left_join(svim, 
                    gdp_long[, c("country", "year", "gdp")],
                    by = c("country", "year"))
  
  # sum(is.na(svim$gdp))
  # paste0("GDP unavailable for: ")
  # paste0(unique(svim[is.na(svim$gdp),]$country))
  # 
  svim <- left_join(svim, 
                     workforce_long[, c("country", "year", "pct_workforce")],
                     by = c("country", "year"))
  return(svim)
  # svim |> 
  #   group_by(id_outbreak) |> 
  #   summarize(pct_workforce = pct_workforce[1])
  # sum(is.na(svim$pct_workforce))
}

add_cea_results <- function(svim) {
  svim |>
    mutate(yld = compute_yld(s_ch_tot, parms=parms),
           yld_averted = compute_yld(s_ch_averted_tot, parms=parms),
           yll = compute_yll(death_tot, life_exp, parms),
           yll_averted = compute_yll(death_averted_tot, life_exp, parms), 
           daly_averted = yld_averted + yll_averted,
           coi_averted = s_ch_averted_tot * coi_per_patient,
           cod_averted = death_averted_tot * gdp * life_exp,
           productivity_lost_averted = 
             s_ch_averted_tot * gdp * ((patient_workday_lost/365)*(pct_workforce/100) +
                                   (caregiver_workday_lost/365)),
           vacc_dose = ifelse(ori_occurred, population * vacc_cov * dose_regimen, 0),
           vacc_cost = (vacc_cost_per_dose + vacc_delivery_cost) * vacc_dose,
           net_cost = vacc_cost - coi_averted - cod_averted - productivity_lost_averted,
           cost_per_daly_averted = 
             ifelse(daly_averted > 0, net_cost / daly_averted, NA),
           cost_per_case_averted = 
             ifelse(s_ch_averted_tot > 0, net_cost / s_ch_averted_tot, NA),
           cost_per_death_averted = 
             ifelse(death_averted_tot > 0, net_cost / death_averted_tot, NA),
           case_averted_per_1000_OCVs = 
             ifelse(vacc_dose > 0, 1000 * s_ch_averted_tot / vacc_dose, NA),
           death_averted_per_1000_OCVs = 
             ifelse(vacc_dose > 0, 1000 * death_averted_tot / vacc_dose, NA),
           cost_eff_threshold = 3 * gdp,
           ratio_cost_per_daly_averted_to_gdp = 
             ifelse(gdp > 0, cost_per_daly_averted / gdp, NA), 
           cost_effective = 
             ifelse(cost_per_daly_averted > cost_eff_threshold, 
                                 FALSE, TRUE)) -> svim_cea
  return(svim_cea)
}
```

## Cost-effectiveness analysis parameters
```{r}
# Disease symptoms, vaccine costs, etc.
parms <- as.data.frame(data.table::fread("data/parameters.csv"))
parms <- filter(parms, Disease == "Cholera")
pr_asymp <- parms[parms$Parameter == "Prop_Asymptomatic", "Value"]
pr_mild <- parms[parms$Parameter == "Prop_Mild", "Value"]
pr_moderate <- parms[parms$Parameter == "Prop_Moderate", "Value"]
pr_severe <- parms[parms$Parameter == "Prop_Severe", "Value"]
dur <- parms[parms$Parameter == "Duration_Illness", "Value"]
wt_asymp <- parms[parms$Parameter == "Disability_Weight_Asymptomatic", "Value"]
wt_mild <- parms[parms$Parameter == "Disability_Weight_Mild", "Value"]
wt_moderate <- parms[parms$Parameter == "Disability_Weight_Moderate", "Value"]
wt_severe <- parms[parms$Parameter == "Disability_Weight_Severe", "Value"]

vacc_cost_per_dose <- parms[parms$Parameter == "Vaccine_Cost", "Value"]
vacc_delivery_cost <- parms[parms$Parameter == "Vaccine_Delivery_Cost", "Value"]
patient_cost_hosp <- parms[parms$Parameter == "Patient_Cost_Hosp", "Value"]
patient_cost_outpt <- parms[parms$Parameter == "Patient_Cost_Outpt", "Value"]
public_cost_hosp <- parms[parms$Parameter == "Public_Cost_Hosp", "Value"]
public_cost_outpt <-parms[parms$Parameter == "Public_Cost_Outpt", "Value"] 
patient_workday_lost <- parms[parms$Parameter == "Pt_Workdays_Lost", "Value"]
caregiver_workday_lost <- parms[parms$Parameter == "Caregiver_Workdays_Lost", "Value"]
mean_age_inf <- parms[parms$Parameter == "Mean_Age_Infection", "Value"]
vacc_shipping_cost <- parms[parms$Parameter == "Vaccine_Shipping_Cost", "Value"]

dose_regimen <- 1
vacc_cost_per_person <- vacc_cost_per_dose * dose_regimen

# assume that reported cases are either moderate or severe cases
pr_tot <- pr_moderate + pr_severe
(coi_per_patient <- 
    pr_moderate / pr_tot * (patient_cost_outpt + public_cost_outpt) + 
    pr_severe / pr_tot * (patient_cost_hosp + public_cost_hosp))

avg_age_inf <- parms[parms$Parameter == "Mean_Age_Infection", "Value"]
```

Cost adjustment 2015 to 2023
```{r}
# consumer price index for cost adjustment (2015 to 2023)
cpi_data <- read_xls("data/API_FP.CPI.TOTL.ZG_DS2_en_excel_v2_9839.xls")
# another gdp_data used previously
gdp_data2 <- read_xls("data/API_NY.GDP.DEFL.KD.ZG_DS2_en_excel_v2_9995.xls")

cpi1 <- cpi_data %>%
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "value")
cpi1$year <- as.numeric(cpi1$year)
cpi1 <- cpi1 %>% filter(year >= 2015 & year <= 2023)

gdp1 <- gdp_data2 %>%
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "value")
gdp1$year <- as.numeric(gdp1$year)
gdp1 <- gdp1 %>% filter(year >= 2015 & year <= 2023)


# Use % change in CPI for sub-Saharan African region, which is available

cpi_percent <- read_xls("data/imf-dm-export-20241028.xls")

cpi_percent <- cpi_percent %>%
  pivot_longer(
    cols = `2015`:`2029`,         
    names_to = "year",             
    values_to = "value"            
  )

cpi_percent$year <- as.numeric(cpi_percent$year)

# year starts at the t + 1 year of reported cost, so if 2015 cost then should use
# % change CPI starting 2016
cpi_percent <- cpi_percent %>% 
  filter(year >= 2016 & year <= 2023)

xi <- (cpi_percent$value/100 + 1)
# cumulative_inflation_factor <- xi[1] * xi[2] * xi[3] * xi[4] * xi[5] * xi[6] * xi[7] * xi[8]
cumulative_inflation_factor <- prod(xi)
patient_cost_hosp_adj <- patient_cost_hosp * cumulative_inflation_factor
patient_cost_outpt_adj <- patient_cost_outpt * cumulative_inflation_factor
public_cost_hosp_adj <- public_cost_hosp * cumulative_inflation_factor
public_cost_outpt_adj <- public_cost_outpt * cumulative_inflation_factor

# vaccine delivery cost was reported in 2016 USD$
cpi_percent <- cpi_percent %>% 
  filter(year >= 2017 & year <= 2023)

xi <- (cpi_percent$value/100 + 1)
# cumulative_inflation_factor <- xi[1] * xi[2] * xi[3] * xi[4] * xi[5] * xi[6] * xi[7]
cumulative_inflation_factor <- prod(xi)
vacc_delivery_cost_adj <- vacc_delivery_cost * cumulative_inflation_factor

patient_cost_hosp <- patient_cost_hosp_adj
patient_cost_outpt <- patient_cost_outpt_adj
public_cost_hosp <- public_cost_hosp_adj
public_cost_outpt <- public_cost_outpt_adj
vacc_delivery_cost <- vacc_delivery_cost_adj
```

Life expectancy and other cost data

```{r}
# the following gdp is different from the overall sub-Saharan Africa gdp
# gdp_data2 <- read_xls("data/GDP_WorldBank.xls")
names(life_exp_data)[names(life_exp_data) == "ISO3 Alpha-code"] <- "country"
names(life_exp_data)[names(life_exp_data) == "Year"] <- "year"
names(workforce_data)[names(workforce_data) == "Country Code"] <- "country"

workforce_long <- 
  workforce_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols=`2001`:`2023`, names_to="year", values_to="pct_workforce")
workforce_long$year <- as.integer(workforce_long$year)

names(gdp_data)[names(gdp_data) == "Country Code"] <- "country"
gdp_long <- 
  gdp_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols=`2001`:`2023`, names_to="year", values_to="gdp")
gdp_long$year <- as.integer(gdp_long$year)
```

## Summarize vaccine impact

### Across study variation across or among outbreaks 
```{r}
svim_df <- fread("outputs/svim_raw_20241113.csv")
svim <- set_svim_raw_res(svim_df) # remove OCV-affected outbreaks, add population size, ..
rm(svim_df)
svim_sum_across_week_by_runid <- svim_sum_across_week(svim, case_trigger=FALSE)

if (save_csv) {
  fwrite(svim_sum_across_week_by_runid,
    paste0("outputs/svim_sum_across_week_by_runid_", tstamp(), ".csv"))
}

svim_mean <- 
  svim_mean_across_runid(svim_sum_across_week_by_runid, case_trigger=FALSE)
svim_mean <- add_summary_stats(svim_mean)

if (save_csv) {
  fwrite(svim_mean,
    paste0("outputs/svim_mean_across_runid_", tstamp(), ".csv"))
}
# cost-effectiveness analysis
svim_mean_cea <- add_cea_variables(svim_mean)
svim_mean_cea <- add_cea_results(svim_mean_cea)
if (save_csv) {
  fwrite(svim_mean_cea,
    paste0("outputs/svim_mean_cea_across_runid_", tstamp(), ".csv"))
}

# case-trigger vaccination
# svim_ct_df <- fread("outputs/svim_case_trigger_raw_20241113.csv")
svim_ct_df <- fread("outputs/svim_case_trigger_raw_20241123.csv")
svim_ct <- set_svim_raw_res(svim_ct_df)
rm(svim_ct_df)
svim_ct_sum_across_week_by_runid <- 
  svim_sum_across_week(svim_ct, case_trigger=TRUE)
rm(svim_ct)
# case-triggered simulation occurred two separate runs and to need to combine
svim_ct_df <- fread("outputs/svim_case_trigger_raw_20241113.csv")
svim_ct2 <- set_svim_raw_res(svim_ct_df)
rm(svim_ct_df)
svim_ct_sum_across_week_by_runid2 <- 
  svim_sum_across_week(svim_ct2, case_trigger=TRUE)
rm(svim_ct2)

svim_ct_sum_across_week_by_runid <-
  rbind(svim_ct_sum_across_week_by_runid, svim_ct_sum_across_week_by_runid2)
rm(svim_ct_sum_across_week_by_runid2)

if (save_csv) {
  fwrite(svim_ct_sum_across_week_by_runid,
         paste0("outputs/svim_ct_sum_across_week_by_runid_", tstamp(), ".csv"))
}
svim_ct_mean <- 
  svim_mean_across_runid(svim_ct_sum_across_week_by_runid, case_trigger=TRUE)
svim_ct_mean <- add_summary_stats(svim_ct_mean)

if (save_csv) {
  fwrite(svim_ct_mean,
    paste0("outputs/svim_ct_mean_across_runid_", tstamp(), ".csv"))
}
# cost-effectiveness analysis
svim_ct_mean_cea <- add_cea_variables(svim_ct_mean)
svim_ct_mean_cea <- add_cea_results(svim_ct_mean_cea)
if (save_csv) {
  fwrite(svim_ct_mean_cea,
    paste0("outputs/svim_ct_mean_cea_across_runid_", tstamp(), ".csv"))
}

# pre-emptive vaccination
svim_pe_df <- fread("outputs/svim_preemptive_raw_20241203.csv")
svim_pe <- set_svim_raw_res(svim_pe_df)
svim_pe_sum_across_week_by_runid <- svim_sum_across_week(svim_pe)
if (save_csv){
  fwrite(svim_pe_sum_across_week_by_runid,
         paste0("outputs/svim_pe_sum_across_week_by_runid_", tstamp(), ".csv"))
}
svim_pe_mean <- svim_mean_across_runid(svim_pe_sum_across_week_by_runid)
svim_pe_mean <- add_summary_stats(svim_pe_mean)
if (save_csv) {
  fwrite(svim_pe_mean,
    paste0("outputs/svim_pe_mean_across_runid_", tstamp(), ".csv"))
}
svim_pe_mean_cea <- add_cea_variables(svim_pe_mean)
svim_pe_mean_cea <- add_cea_results(svim_pe_mean_cea)
if (save_csv) {
  fwrite(svim_pe_mean_cea,
    paste0("outputs/svim_pe_mean_cea_across_runid_", tstamp(), ".csv"))
}


svim_pe_sum_across_week_by_runid |> 
  filter(vacc_cov %in% VC3) |>
  group_by(vacc_cov) |>
  summarize(mean_pcr = mean(pct_reduc_case),
            q500_pcr = quantile(pct_reduc_case, probs=0.5),
            q025_pcr = quantile(pct_reduc_case, probs=0.025),
            q975_pcr = quantile(pct_reduc_case, probs=0.975))

svim_pe_mean |> 
  filter(vacc_cov %in% VC3) |>
  group_by(vacc_cov) |>
  summarize(mean_pcr = mean(pct_reduc_case),
            q500_pcr = quantile(pct_reduc_case, probs=0.5),
            q025_pcr = quantile(pct_reduc_case, probs=0.025),
            q975_pcr = quantile(pct_reduc_case, probs=0.975))

```

### All outbreaks collectively

```{r}
svim_df <- fread("outputs/svim_raw_20241113.csv")
svim <- set_svim_raw_res(svim_df)
rm(svim_df)
a <- svim_sum_across_week_all_outbreaks(svim)
as <- add_summary_stats(a)
ace <- add_cea_variables(as)
acea <- add_cea_results_all_outbreaks(ace)
#a_ <- svim_sum_across_week(svim) # the above version a lot faster
svim_all_by_runid <- svim_sum_across_outbreaks_by_runid(acea)

if (save_file) {
  fwrite(svim_all_by_runid, 
         paste0("outputs/svim_all_outbreaks_by_runid_", tstamp(), ".csv"))
  fwrite(svim_all_by_runid,
    paste0("outputs/svim_all_outbreaks_cea_across_runid_", tstamp(), ".csv"))
}

dat <- svim_all_by_runid |> filter(vacc_cov == VC & vacc_week == VW)

# create tables for impact
impact_tab2 <- function(dat, probs=c(0.025,0.25,0.5,0.75,0.975)) {
  lst <- vector('list', 11)
  lst[[1]] <- quantile(dat$pct_reduc_case, probs=probs, na.rm=T)
  lst[[2]] <- quantile(dat$s_ch_averted_tot, probs=probs, na.rm=T)
  lst[[3]] <- quantile(dat$pct_reduc_death, probs=probs, na.rm=T)
  lst[[4]] <- quantile(dat$death_averted_tot, probs=probs, na.rm=T)
  lst[[5]] <- quantile(dat$case_averted_per_1000_OCV, probs=probs, na.rm=T)
  lst[[6]] <- quantile(dat$death_averted_per_1000_OCV, probs=probs, na.rm=T)
  # dose required to avert a case
  dat$dose_per_case_averted <- ifelse(dat$case_averted_per_1000_OCV > 0, 
                                      1000/dat$case_averted_per_1000_OCV, NA)
  dat$dose_per_death_averted <- ifelse(dat$death_averted_per_1000_OCV > 0,  
                                       1000/dat$death_averted_per_1000_OCV, NA)
  
  lst[[7]] <- quantile(dat$dose_per_case_averted, probs=probs, na.rm=T)
  lst[[8]] <- quantile(dat$dose_per_death_averted, probs=probs, na.rm=T)  
  lst[[9]] <- quantile(dat$cost_per_case_averted, probs=probs, na.rm=T)
  lst[[10]] <- quantile(dat$cost_per_death_averted, probs=probs, na.rm=T)
  lst[[11]] <- quantile(dat$cost_per_daly_averted, probs=probs, na.rm=T)
  
  names(lst) <- c("PCR","CA","PDR","DA","CAPD","DAPD","DPCA","DPDA",
                  "CPCA","CPDA","ICER")
  
  return(lst)
}

svim_all_by_runid |> filter(vacc_cov == VC & vacc_week == VW) -> dat
d <- impact_tab2(dat=dat)

create_table <- function(d) {
  nrow <- 1
  tab <- data.frame(
      PCR = rep(NA,nrow),
      CA = rep(NA,nrow),
      PDR = rep(NA,nrow),
      DA = rep(NA,nrow),
      CAPD = rep(NA,nrow),
      DAPD = rep(NA,nrow),
      CPCA = rep(NA,nrow),
      CPDA = rep(NA,nrow),
      ICER = rep(NA,nrow))
  i <- 1
  tab$PCR[i] <- format_median_with_95(c(d$PCR[3],d$PCR[1],d$PCR[5]), digits=1)
  tab$CA[i] <- format_median_with_95(c(d$CA[3],d$CA[1],d$CA[5]), digits=0)
  tab$PDR[i] <- format_median_with_95(c(d$PDR[3],d$PDR[1],d$PDR[5]), digits=1)
  tab$DA[i] <- format_median_with_95(c(d$DA[3],d$DA[1],d$DA[5]), digits=0)
  tab$CAPD[i] <- format_median_with_95(c(d$CAPD[3],d$CAPD[1],d$CAPD[5]), digits=3)
  tab$DAPD[i] <- format_median_with_95(c(d$DAPD[3],d$DAPD[1],d$DAPD[5]), digits=3)
  tab$CPCA[i] <- format_median_with_95(c(d$CPCA[3],d$CPCA[1],d$CPCA[5]), digits=1)
  tab$CPDA[i] <- format_median_with_95(c(d$CPDA[3],d$CPDA[1],d$CPDA[5]), digits=1)
  tab$ICER[i] <- format_median_with_95(c(d$ICER[3],d$ICER[1],d$ICER[5]), digits=1)

  return(tab)
}

vacc_weeks <- c(1, 5, 9, 13, 17)
tab <- create_table(d)
vcw <- expand.grid(VC3, vacc_weeks)
names(vcw) <- c("Vaccine coverage", "Campaign timing (week)")
df <- data.frame()
for (i in 1:nrow(vcw)) {
  svim_all_by_runid |> filter(vacc_cov == vcw[i,1] & vacc_week == vcw[i,2]) -> dat
  d <- impact_tab2(dat=dat, probs=probs)
  tab <- create_table(d)
  df <- rbind(df, tab)
}
df <- cbind(vcw, df)
if (save_csv) {
  fwrite(df, paste0("outputs/svim_summary_all_outbreaks", tstamp(), ".csv"))
}
```

## Subsets of outbreaks

```{r}
svim <- fread("outputs/svim_mean_across_runid_20241122.csv")
d <- svim |> filter(vacc_cov == VC, vacc_week == VW)
d <-  d[,c("id_outbreak", "duration_weeks", "s_ch_tot", "ar_per_10e3", "population")]
names(d) <- c("id_outbreak", "duration","cumulative_cases", "attack_rate", "population_size")

sum(d$attack_rate > 1000)
d <- d[d$attack_rate <=1000, ] # >1000 is nonsensical attack rate definition = 
methods <- c("random", "attack_rate","duration", "cases")
outbreak_subsets <- vector('list',4)
names(outbreak_subsets) <- methods
target_pop <- 135e6
samp_size <- 200
# seeds <- 1:samp_size
max_attempts <- 1e5
set.seed(42)

for(i in 1:length(methods)) {
  subset <- vector("list", samp_size)
  for (j in 1:length(subset)) {
    subset[[j]] <-
      sample_outbreaks(
        outbreak_data = d, 
        target_population = target_pop,
        sampling_method = methods[i], 
        # seed = seeds[j], 
        max_attempts = max_attempts)
  }
  outbreak_subsets[[i]] <- subset
}

saveRDS(outbreak_subsets, paste0("outputs/svim_outbreak_subsets_ids_", tstamp(), ".rds"))
```

### Subsets into a single dataframe
```{r}
samp_size <- 200
methods <- c("random", "attack_rate","duration", "cases")
lst <- vector("list", samp_size * length(methods))
for (i in 1:length(methods)) {
  for (j in 1:samp_size){
    filter(svim, id_outbreak %in% 
             outbreak_subsets[[methods[i]]][[j]]$sampled_data$id_outbreak) -> d
    d$subset_method <- methods[i]
    d$subset_id <- j
    lst[[(i-1)*samp_size + j]] <- d
  }
}
df <- rbindlist(lst)
if (save_csv) {
  fwrite(df, paste0("outputs/svim_outbreak_subsets_", tstamp(), ".csv"))
}
```

### Subset summary statistics

```{r}
df <- fread("outputs/svim_outbreak_subsets_20241126.csv")
unique(df$subset_method)
# lvs <- c("random", "attack_rate", "duration", "cases")
# df$subset_method <- factor(df$subset_method, 
#                            levels=lvs)  
  
df |> # across outbreaks in a given data subset with subset_id
  filter(vacc_cov == VC & vacc_week == VW) |> 
  group_by(subset_method, subset_id) |> # across outbreak ids
  summarize(sum_pop = sum(population),
            sum_case = sum(s_ch_tot),
            overall_ar = 1000 * sum_case / sum_pop,
            sum_dur = sum(duration_weeks)) |> 
  group_by(subset_method) |> # across subset_id 
  summarize(median_pop = quantile(sum_pop, probs=c(0.5)), 
            median_case = quantile(sum_case, probs=c(0.5)),
            median_ar = quantile(overall_ar, probs=c(0.5)),
            median_dur = quantile(sum_dur, probs=c(0.5)),
            lower_pop = quantile(sum_pop, probs=c(0.025)), 
            lower_case = quantile(sum_case, probs=c(0.025)),
            lower_ar = quantile(overall_ar, probs=c(0.025)),
            lower_dur = quantile(sum_dur, probs=c(0.025)),
            upper_pop = quantile(sum_pop, probs=c(0.975)), 
            upper_case = quantile(sum_case, probs=c(0.975)),
            upper_ar = quantile(overall_ar, probs=c(0.975)),
            upper_dur = quantile(sum_dur, probs=c(0.975))) -> dfs
            
tab <- data.frame(subset=dfs$subset_method,
                  population=rep(NA,4), cases=rep(NA,4), 
                  duration=rep(NA,4), attack_rate=rep(NA,4))
for (i in 1:4) {
  
  tab$population[i] <- format_median_with_95(c(dfs$median_pop[i],
                                          dfs$lower_pop[i],
                                          dfs$upper_pop[i]), digit=0)
  tab$cases[i] <- format_median_with_95(c(dfs$median_case[i],
                                          dfs$lower_case[i],
                                          dfs$upper_case[i]), digit=0)
  tab$duration[i] <- format_median_with_95(c(dfs$median_dur[i],
                                          dfs$lower_dur[i],
                                          dfs$upper_dur[i]), digit=0)
  tab$attack_rate[i] <- format_median_with_95(c(dfs$median_ar[i],
                                          dfs$lower_ar[i],
                                          dfs$upper_ar[i]), digit=2)
}
tab

if (save_csv) {
  fwrite(tab, paste0("outputs/subset_summary_stat_", tstamp(), ".csv"))
}
```

### Subsets analysis - time-triggered
```{r}
subset_ids <- readRDS("outputs/svim_outbreak_subsets_ids_20241126.rds")
svim_df <- fread("outputs/svim_raw_20241113.csv")
# create svim_df subsets
# nm <- "attack_rate"
# nm <- "duration"
nm <- "cases"
df <- data.frame()
# mean across runids and later compute statistics across subset ids
for (i in 1:200) {
  cat("i = ", i,"\n")
  # d <- svim_df |> 
  #   filter(id_outbreak %in% subset_ids$random[[i]]$sampled_data$id_outbreak)
  # d <- svim_df |> 
  #   filter(id_outbreak %in% subset_ids$attack_rate[[i]]$sampled_data$id_outbreak)
  # d <- svim_df[id_outbreak %in% subset_ids$duration[[i]]$sampled_data$id_outbreak]  
  d <- svim_df[id_outbreak %in% subset_ids$cases[[i]]$sampled_data$id_outbreak]  
  svim <- set_svim_raw_res(d)
  rm(d)
  a <- svim_sum_across_week_all_outbreaks(svim)
  rm(svim)
  as <- add_summary_stats(a)
  ace <- add_cea_variables(as)
  acea <- add_cea_results_all_outbreaks(ace)
  #a_ <- svim_sum_across_week(svim) # the above version a lot faster
  svim_subset_by_runid <- svim_sum_across_outbreaks_by_runid(acea)
  # mean across runid
  # svim_subset_by_runid |> 
  #   group_by(vacc_cov, vacc_week) |> 
  #   summarise(across(population:cost_per_death_averted, ~ mean(.x, na.rm = TRUE))) ->
  #   svim_mean1
  
  svim_mean <- svim_subset_by_runid[, lapply(.SD, mean), by = .(vacc_cov, vacc_week),
          .SDcols = population:cost_per_death_averted]
  
  rm(svim_subset_by_runid)
  svim_mean$subset_id <- i
  df <- rbind(df, svim_mean)
}
# fwrite(df, paste0("outputs/svim_mean_random_subset_ids_", tstamp(), ".csv"))
# fwrite(df, paste0("outputs/svim_mean_attack_rate_subset_ids_", tstamp(), ".csv"))
fwrite(df, paste0("outputs/svim_mean_", nm, "_subset_ids_", tstamp(), ".csv"))

svim_means <- df

# tab <- create_table(d)
vacc_weeks <- c(1, 5, 9, 13, 17)
vcw <- expand.grid(VC3, vacc_weeks)
names(vcw) <- c("Vaccine coverage", "Campaign timing (week)")
df <- data.frame()
for (i in 1:nrow(vcw)) {
  svim_means[vacc_cov == vcw[i,1] & vacc_week == vcw[i,2]] -> dat
  d <- impact_tab2(dat=dat, probs=PROBS)
  tab <- create_table(d)
  df <- rbind(df, tab)
}
df <- cbind(vcw, df)
df$subset_method <- nm

if (save_csv) {
  fwrite(df, paste0("outputs/svim_summary_", nm, "_subsets_", tstamp(), ".csv"))
}

df1 <- fread("outputs/svim_mean_random_subset_ids_20241126.csv")
d1 <- filter(df1, vacc_cov == VC, vacc_week == VW)
df2 <- fread("outputs/svim_mean_attack_rate_subset_ids_20241126.csv")
d2 <- filter(df2, vacc_cov == VC, vacc_week == VW)
df3 <- fread("outputs/svim_mean_cases_subset_ids_20241126.csv")
d3 <- filter(df3, vacc_cov == VC, vacc_week == VW)
df4 <- fread("outputs/svim_mean_duration_subset_ids_20241126.csv")
d4 <- filter(df4, vacc_cov == VC, vacc_week == VW)

d1 <- impact_tab2(dat=d1, probs=PROBS)
tab1 <- create_table(d1)
d2 <- impact_tab2(dat=d2, probs=PROBS)
tab2 <- create_table(d2)
d3 <- impact_tab2(dat=d3, probs=PROBS)
tab3 <- create_table(d3)
d4 <- impact_tab2(dat=d4, probs=PROBS)
tab4 <- create_table(d4)

tab <- rbind(tab1, tab2, tab3, tab4)
tab$subset_method <- c("random","attack_rate","cases","duration")
if (save_csv) {
  fwrite(tab, paste0("outputs/svim_summary_subsets_", tstamp(), ".csv"))
}
# d <- impact_tab2(dat=df, probs=PROBS)
# tab <- create_table(d)
```

### Subsets analysis - case-triggered
```{r}
subset_ids <- readRDS("outputs/svim_outbreak_subsets_ids_20241126.rds")
svim_df <- fread("outputs/svim_case_trigger_raw_20241113.csv")
nsample <- 200
# create svim_df subsets
nms <- c("random", "attack_rate", "duration", "cases")
for (nm in nms) {
  cat("subset:", nm, "\n")
  df <- data.frame()
  # mean across runids and later compute statistics across subset ids
  for (i in 1:nsample) {
    cat("i =", i,"\n")
    if (nm == "random") {
      d <- svim_df[id_outbreak %in% subset_ids$random[[i]]$sampled_data$id_outbreak]
    }
    else if (nm == "attack_rate") {
     d <- svim_df[id_outbreak %in% subset_ids$attack_rate[[i]]$sampled_data$id_outbreak]
    }
    else if (nm == "duration") {
     d <- svim_df[id_outbreak %in% subset_ids$duration[[i]]$sampled_data$id_outbreak] 
    }
    else if (nm == "cases") {
      d <- svim_df[id_outbreak %in% subset_ids$cases[[i]]$sampled_data$id_outbreak] 
    }
    #  
    #  
    svim <- set_svim_raw_res(d)
    rm(d)
    a <- svim_sum_across_week_all_outbreaks(svim, case_trigger=TRUE)
    rm(svim)
    as <- add_summary_stats(a, case_trigger=TRUE)
    ace <- add_cea_variables(as)
    acea <- add_cea_results_all_outbreaks(ace)
    #a_ <- svim_sum_across_week(svim) # the above version a lot faster
    svim_subset_by_runid <- svim_sum_across_outbreaks_by_runid(acea, case_trigger=TRUE)
    rm(acea)
    # mean across runid
    # svim_subset_by_runid |> 
    #   group_by(vacc_cov, vacc_week) |> 
    #   summarise(across(population:cost_per_death_averted, ~ mean(.x, na.rm = TRUE))) ->
    #   svim_mean1
    
    svim_mean <- 
      svim_subset_by_runid[, lapply(.SD, mean), 
                           by = .(vacc_cov, case_trigger),
                           .SDcols = population:cost_per_death_averted]
    
    rm(svim_subset_by_runid)
    svim_mean$subset_id <- i
    df <- rbind(df, svim_mean)
    rm(svim_mean)
  }
  
  fwrite(df, paste0("outputs/svim_ct_mean_", nm, "_subset_ids_", tstamp(), ".csv"))
  rm(df)
}

svim_means <- df

# tab <- create_table(d)
case_triggers <- c(5, 9, 13, 17)
vcw <- expand.grid(VC3, case_triggers)
names(vcw) <- c("Vaccine coverage", "Case trigger")
df <- data.frame()
for (i in 1:nrow(vcw)) {
  svim_means[vacc_cov == vcw[i,1] & case_trigger == vcw[i,2]] -> dat
  d <- impact_tab2(dat=dat, probs=PROBS)
  tab <- create_table(d)
  df <- rbind(df, tab)
}
df <- cbind(vcw, df)
df$subset_method <- nm

if (save_csv) {
  fwrite(df, paste0("outputs/svim_summary_", nm, "_subsets_", tstamp(), ".csv"))
}

df1 <- fread("outputs/svim_mean_random_subset_ids_20241126.csv")
d1 <- filter(df1, vacc_cov == VC, vacc_week == VW)
df2 <- fread("outputs/svim_mean_attack_rate_subset_ids_20241126.csv")
d2 <- filter(df2, vacc_cov == VC, vacc_week == VW)
df3 <- fread("outputs/svim_mean_cases_subset_ids_20241126.csv")
d3 <- filter(df3, vacc_cov == VC, vacc_week == VW)
df4 <- fread("outputs/svim_mean_duration_subset_ids_20241126.csv")
d4 <- filter(df4, vacc_cov == VC, vacc_week == VW)

d1 <- impact_tab2(dat=d1, probs=PROBS)
tab1 <- create_table(d1)
d2 <- impact_tab2(dat=d2, probs=PROBS)
tab2 <- create_table(d2)
d3 <- impact_tab2(dat=d3, probs=PROBS)
tab3 <- create_table(d3)
d4 <- impact_tab2(dat=d4, probs=PROBS)
tab4 <- create_table(d4)

tab <- rbind(tab1, tab2, tab3, tab4)
tab$subset_method <- c("random","attack_rate","cases","duration")
if (save_csv) {
  fwrite(tab, paste0("outputs/svim_summary_subsets_", tstamp(), ".csv"))
}
# d <- impact_tab2(dat=df, probs=PROBS)
# tab <- create_table(d)
```



## Vaccine impact simulation

Parallel by batch `run_vacc_impact_outbreak_weekly`

```{r}
tic <- Sys.time()

library(doParallel)
# UN data on country- and year-specific age distribution (< 5 yo and 5+ yo)
# pop_by_age_wpp <- as.data.frame(fread("outputs/wpp_pop_by_age_20241017.csv"))
vacc_covs <- seq(0.6, 0.95, by=0.05)
vacc_weeks <- seq(1, 20, by=1) # vaccination can happen anytime during week 1 and week 20
outbreak_ids <- unique(outbreak_data$id_outbreak)
nruns <- 30 # number of simulation runs
batch_size <- 5

for (batch_id in seq(1, nruns, by = batch_size)) {
  cat("batch id: ", batch_id, "\n")
  run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in batches of batch_size

  cl <- makeCluster(getOption("cl.cores", detectCores() - 1))
  registerDoParallel(cl)

  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = outbreak_data,
        ive_data = ive_data,
        parameters = params,
        age_dist = pop_by_age_wpp,
        vacc_week = vacc_weeks,
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)
  saveRDS(svim_week_list, paste0("outputs/svim_week_list_vacc_week_batch_",
                                 batch_id, "_", tstamp(), ".rds"))
}
```

Parallel case_trigger by batch

```{r, eval=FALSE}
library(doParallel)
case_triggers <- c(seq(110,190, by=10), seq(200,300, by=10))
nruns <- 30
batch_size <- 5
vacc_covs <- seq(0.6, 0.95, by=0.05)
outbreak_ids <- unique(outbreak_data_all$id_outbreak)
nruns <- 30 # number of simulation runs
batch_size <- 5

cl <- makeCluster(getOption("cl.cores", detectCores() - 1))
registerDoParallel(cl)

for (batch_id in seq(1, nruns, by = batch_size)) {

  cat("batch id: ", batch_id, "\n")
  run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in batches of batch_size

  cl <- makeCluster(getOption("cl.cores", detectCores() - 4))
  registerDoParallel(cl)

  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = outbreak_data_all,
        ive_data = ive_data,
        parameters = params,
        age_dist = pop_by_age_wpp,
        case_trigger = case_triggers,
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)
  saveRDS(svim_week_list, paste0("outputs/svim_case_trigger_week_list_batch_",
                                 batch_id, "_", tstamp(), ".rds"))
}
```

Parallel pre-emptive vaccination `run_vacc_impact_outbreak_weekly`

```{r, eval=FALSE}
library(doParallel)

tic <- Sys.time()
# UN data on country- and year-specific age distribution (< 5 yo and 5+ yo)
# pop_by_age_wpp <- as.data.frame(fread("outputs/wpp_pop_by_age_20241022.csv"))
vacc_covs <- seq(0.6, 0.95, by=0.05)
# maximum delay possible
# max(params$dur_campaign)/7 + max(params$delay_vacc_effect)/7
vacc_weeks <- -52 # some random large number ensure than pre-emptive vaccination
nruns <- 200 # number of simulation runs

cl <- makeCluster(getOption("cl.cores", detectCores() - 6))
registerDoParallel(cl)

svim_week_list <-
  foreach(i = 1:nruns, .packages = c("data.table")) %dopar% {
    run_vacc_impact_outbreak_weekly(
      outbreak_data = outbreak_data,
      ive_data = ive_data,
      parameters = params,
      age_dist = pop_by_age_wpp,
      vacc_week = vacc_weeks,
      vacc_cov = vacc_covs,
      runid = i)
  }
stopCluster(cl)

Sys.time() - tic
# Time difference of 36.02764 mins
saveRDS(svim_week_list, paste0("outputs/svim_preemptive_list_", tstamp(), ".rds"))
```

## Summarize the simulation results

Combine the results from results from separate files across which runid's are different.
```{r}
# svim_week_list <- read_rds("outputs/svim_week_list_vacc_week_20241010.rds")
nruns <- 30
batch_size <- 5
batch_ids <- seq(1, nruns, by=batch_size)

lst <- vector('list', length(batch_ids))
for ( i in 1:length(batch_ids)) {
  lst[[i]] <- rbindlist(read_rds(
    paste0("outputs/svim_week_list_vacc_week_batch_", batch_ids[i],
           "_20241113.rds")))
}
svim_df <- rbindlist(lst)
fwrite(svim_df, paste0("outputs/svim_raw_", tstamp(), ".csv"))

lst <- vector('list', length(batch_ids))
for ( i in 1:length(batch_ids)) {
  lst[[i]] <- rbindlist(read_rds(
    paste0("outputs/svim_case_trigger_week_list_batch_", batch_ids[i],
           "_20241123.rds")))
}

svim_df <- rbindlist(lst)
fwrite(svim_df, paste0("outputs/svim_case_trigger_raw_", tstamp(), ".csv"))

# lst <- read_rds("outputs/svim_preemptive_list_20241203.rds")
svim_df <- rbindlist(lst)
# table(svim_df$ori_occurred)
# table(svim_df$week_delay_to_vacc_effect)
fwrite(svim_df, paste0("outputs/svim_preemptive_raw_", tstamp(), ".csv"))
```

## Comparison with the dynamic model

Identify outbreaks (`id_outbreak`) that led to poor fits
1. Parameter values are too close to the bound 
2. The number of data points are fewer than 8, two times the number of parameters to be estimated (i.e., 4)
3. Modeled trajectories didn't capture the dynamics of outbreaks
```{r}
fit <- fread("julia/outputs/fit_4p_fA010_20241203T18.csv")
ndata <- fread("julia/outputs/fit_params_ndata_20241206T08.csv")
fit <- left_join(fit, ndata, by=c("data_id", "id_outbreak"))
er <- 0.01 # error factor 5%
fit$est_close_to_bounds <- ifelse(fit$s0 < 1*er | fit$s0 > 1*(1-er) | fit$R0 > 30*(1-er), 1, 0) 
fit$ndata_low <- ifelse(fit$ndata < 8, 1, 0)

ids_est_close_to_bounds <- fit[est_close_to_bounds == 1, id_outbreak]
ids_ndata_low <- fit[ndata_low == 1, id_outbreak]  

# vaccine efficacy
fit_vis <-  readxl::read_xlsx("julia/outputs/vacc_fit_params_20241203T2122.xlsx")
fit_vis$Good_Fit_Visual <- ifelse(is.na(fit_vis$Good_Fit_Visual), 1, 0)
fit_vis |> 
  filter(Good_Fit_Visual == 0) |>
  pull(id_outbreak) -> ids_poor_fit_visual

# ids_robust <- unique(c(ids1 %in%, ids2))
ids_poor_fit <- unique(c(ids_est_close_to_bounds, ids_ndata_low, ids_poor_fit_visual))
ids_robust <- unique(fit$id_outbreak)[!(unique(fit$id_outbreak) %in% ids_poor_fit)]
```

```{r}
svim_pe_sum <- fread("outputs/svim_pe_sum_across_week_by_runid_20241203.csv")
svim_pe_mean <- fread("outputs/svim_pe_mean_across_runid_20241203.csv")
# summary across id_outbreak and runid's
svim_pe_q <- svim_pe_sum |> 
  filter(vacc_cov == 0.75) |> 
  summarize(q025 = quantile(pct_reduc_case, probs=0.025),
            q250 = quantile(pct_reduc_case, probs=0.25),
            q500 = quantile(pct_reduc_case, probs=0.5),
            q750 = quantile(pct_reduc_case, probs=0.75),
            q975 = quantile(pct_reduc_case, probs=0.975))

svim_pe_sum_vc <- filter(svim_pe_sum, vacc_cov == 0.75)
   
d <- fread("julia/outputs/vacc_fit_params_20241203T2122.csv")

d$ove_vc60 <- (1 - d$CI_VC60 / d$CI_novacc) * 100
d$ove_vc75 <- (1 - d$CI_VC75 / d$CI_novacc) * 100
d$ove_vc90 <- (1 - d$CI_VC90 / d$CI_novacc) * 100
d_ <- d[, c("data_id","id_outbreak","ove_vc60", "ove_vc75", "ove_vc90")]
names(d_) <- c("data_id","id_outbreak", 0.6, 0.75, 0.9)
dlong <- pivot_longer(d_, cols=-c(data_id, id_outbreak))
names(dlong) <- c("data_id", "id_outbreak", "vacc_cov", "pct_reduc_case_dynamic")
dlong$vacc_cov <- as.numeric(dlong$vacc_cov)

dd <- left_join(dlong, 
                svim_pe_mean[,c("id_outbreak","vacc_cov", "vacc_week", "pct_reduc_case")],
                by=c("id_outbreak","vacc_cov"))

d2 <- readxl::read_xlsx("julia/outputs/vacc_fit_params_20241203T2122.xlsx")
d2$Good_Fit_Visual <- ifelse(is.na(d2$Good_Fit_Visual), 1, 0)
d2 |> filter(Good_Fit_Visual ==1) |> pull(id_outbreak) -> ids
dd |> 
  group_by(as.factor(vacc_cov)) |>
  filter(id_outbreak %in% ids) |>
  summarize(mean_pcr = mean(pct_reduc_case),
            mean_pcr_dynamic = mean(pct_reduc_case_dynamic)) -> x
x
dd |> 
  filter(vacc_cov == 0.75) |>
  filter(id_outbreak %in% ids) |> 
  ggplot()+
  geom_boxplot(aes(x="dynamic",y=pct_reduc_case_dynamic))+
    stat_summary(aes(x="dynamic",y=pct_reduc_case_dynamic), 
               fun=mean, colour="firebrick", geom="point", 
               shape=18, size=3, show.legend=FALSE)+
  geom_boxplot(data=svim_pe_sum_vc, 
               aes(x="static",y=pct_reduc_case))+
  stat_summary(data=svim_pe_sum_vc,
               aes(x="static",y=pct_reduc_case), 
               fun=mean, colour="firebrick", geom="point", 
               shape=18, size=3, show.legend=FALSE)+
  # geom_hline(aes(yintercept=pct_reduc_case[1]), color="forestgreen", 
  #            linewidth=1, linetype ="dotted")+
  theme_light()+
  labs(x="", y="Case reduction (%)")

```

Direct, indirect, total, and overall vaccine effectiveness computation for
the static and the dynamic model
```{r}
# static model
# effetiveness computations are based on the definition
# DVE, IVE: parameter values
# TVE = 1 - (1-DVE)(1-IVE)
# OVE = 1 - (vc*(1-DVE)(1-IVE) + (1-vc)*(1-IVE))

setDT(outbreak_data)
age_dist <- pop_by_age_wpp
# extract outbreak data, can be done by including only the first rwo
od <- outbreak_data[week == 1]
od$prop_u5 <- NA
for (i in 1:nrow(od)) {
  od$prop_u5[i] <- 
    age_dist[(age_dist$`ISO3 Alpha-code` == od$country[i] &
                as.character(age_dist$Year) == as.character(od$year[i])), "prop_u5"]
}

vacc_par <- params[,c("vacc_effect_direct_u5", "vacc_effect_direct_5p", "vacc_effect_indirect_id")]
increment <- 0.02 # increment used in posterior predictive values
vacc_par$ive <- as.numeric(ive_data[vacc_par$vacc_effect_indirect_id, (round(vc/increment)+1)])

lst <- vector('list', nrow(od))
vacc_eff_static <- data.frame()
vcs <- c(0.6, 0.75, 0.9)
nsample <- 200
for (vc in vcs) {
  for (i in 1:nrow(od)) {
    d <- od[i, ]
    d <- rbindlist(lapply(1:100, function(x) rbind(d,d)))
    d$runid <- NA
    d$vacc_cov <- vc
    d$ve_u5 <- NA
    d$ve_5p <- NA
    d$dve <- NA
    d$ive <- NA
    d$tve <- NA
    d$ove <- NA
    
    for (j in 1:nsample) {
      d$runid[j] <- j
      d$ve_u5[j] <- vacc_par$vacc_effect_direct_u5[j]
      d$ve_5p[j] <- vacc_par$vacc_effect_direct_5p[j]
      d$dve[j] <- d$prop_u5[j]*d$ve_u5[j] + (1-d$prop_u5[j])*d$ve_5p[j]
      d$ive[j] <- vacc_par$ive[j]
      d$tve[j] <- 1 - (1-d$dve[j])*(1-d$ive[j])
      d$ove[j] <- 1 - (vc*(1-d$dve[j])*(1-d$ive[j]) + (1-vc)*(1-d$ive[j]))
    }
    lst[[i]] <- d
  }
  vacc_eff_static <- rbindlist(list(vacc_eff_static, rbindlist(lst)))
}
vacc_eff_static

# dynamic model
# effectiveness computations are based on the modeled incidence 
# d <- fread("julia/outputs/vacc_fit_params_by_vacc_state_20241204T15.csv")
# d <- fread("julia/outputs/vacc_fit_params_by_vacc_state_20241205T15.csv")
# d$dve_vc60_ <- 1 - (d$CI_VC60_vacc/vc)/(d$CI_VC60_novacc/(1-vc))
# d$ive_vc60_ <- 1 - (d$CI_VC60_novacc/(1-vc)) /(d$CI_novacc)
# d$tve_vc60_ <- 1 - (d$CI_VC60_vacc/vc) /(d$CI_novacc)
# d$ove_vc60_ <- 1 - (d$CI_VC60_vacc + d$CI_VC60_novacc)/(d$CI_novacc)

d <- fread("julia/outputs/vacc_fit_params_by_vacc_state_20241206T08.csv")
# d <- fread("julia/outputs/vacc_fit_params_by_vacc_state_20241205T17.csv")
vcs <- c(0.6, 0.75, 0.9)

lst <- vector('list', length(vcs))
for (i in 1:length(vcs)) {
  vc <- vcs[i]
  df <- d[, c("runid","data_id","id_outbreak")]
  v <- round(100*vc, digits=0)
  df$vacc_cov <- vc
  df$dve <- 1 - (d[[paste0("CI_VC", v, "_vacc")]]/vc)/(d[[paste0("CI_VC", v, "_novacc")]]/(1-vc))
  df$ive <- 1 - (d[[paste0("CI_VC", v, "_novacc")]]/(1-vc))/d[[paste0("CI_novacc")]]
  df$tve <- 1 - (d[[paste0("CI_VC", v, "_vacc")]]/vc)/d[[paste0("CI_novacc")]]
  df$ove <- 1 - (d[[paste0("CI_VC", v, "_vacc")]] + d[[paste0("CI_VC", v, "_novacc")]])/d[[paste0("CI_novacc")]]
  
  lst[[i]] <- df
}
vacc_eff_dynamic <- rbindlist(lst)

# nrow(vacc_eff_dynamic)/3/200
# [1] 1406
# nrow(vacc_eff_static)/3/200
# [1] 1406
```


```{r}
# dve_cutoff <- 0.4
d1 <- vacc_eff_dynamic[,c("runid","id_outbreak","vacc_cov","dve","ive","tve","ove")]
d1$model <- "Dynamic"
# d1 <- filter(d1, d1$dve > dve_cutoff)
ids_low_dve <- d1 |> filter(dve < dve_cutoff) |> pull(id_outbreak) |> unique()

length(unique(d1$id_outbreak))
d2 <- vacc_eff_static[,c("runid","id_outbreak","vacc_cov","dve","ive","tve","ove")]
d2$model <- "Static"
d <- rbind(d1,d2) 

dlong <- pivot_longer(d, cols=c("dve","ive","tve","ove"), 
                      names_to = "vacc_eff")
dlong$vacc_eff <- as.factor(dlong$vacc_eff)
# levels(dlong$vacc_eff)
levels(dlong$vacc_eff) <- 
  c("Direct VE", "Indirect VE", "Overall VE", "Total VE")
dlong$value <- dlong$value * 100

vc <- 0.75
ids_all <- unique(fit$id_outbreak)
# ids_est_close_to_bounds
# ids_ndata_low
# ids_poor_fit_visual

# ids_robust <- ids_all[!(ids_all%in% ids_est_close_to_bounds)]
# ids_robust <- ids_all[!(ids_all%in% ids_ndata_low)]
ids_poor <- unique(c(ids_est_close_to_bounds, ids_poor_fit_visual, ids_ndata_low))
ids_robust <- ids_all[!(ids_all%in% ids_poor)]

levels(dlong$name)
dlong |> 
  filter(vacc_cov == vc) |>
  # filter(runid == 1) |>
  filter(id_outbreak %in% ids_robust) |> 
  ggplot()+
  geom_boxplot(aes(x=model, y=value))+
  stat_summary(aes(x=model, y=value), 
               fun=mean, colour="firebrick", geom="point", 
               shape=18, size=3, show.legend=FALSE)+
  facet_wrap(~vacc_eff)+
  theme_light()+
  theme_tight()+
  labs(x="", y="Vaccine effectiveness") -> plt


# ggsave(paste0("plots/vacc_eff_4_typese_", tstamp(), ".pdf"), plt,
#          device = cairo_pdf, width=fgw2, height=fgw2, units="mm")
  
# vacc_eff_dynamic |> 
#   filter(vacc_cov == vc) |>
#   filter(id_outbreak %in% ids) |> 
#   ggplot()+
#   geom_boxplot(aes(x="dynamic", y=dve))+
#   stat_summary(aes(x="dynamic",y=dve), 
#                fun=mean, colour="firebrick", geom="point", 
#                shape=18, size=3, show.legend=FALSE)+
#   geom_boxplot(data=filter(vacc_eff_static, vacc_cov==0.75), 
#                aes(x="static",y=dve))+
#   stat_summary(data=filter(vacc_eff_static, vacc_cov==0.75),
#                aes(x="static",y=dve), 
#                fun=mean, colour="firebrick", geom="point", 
#                shape=18, size=3, show.legend=FALSE)+
#   theme_light()+
#   labs(x="", y="Case reduction (%)", title="Direct VE")
# 
# 
# vacc_eff_dynamic |> 
#   filter(vacc_cov == vc) |>
#   filter(id_outbreak %in% ids) |> 
#   ggplot()+
#   geom_boxplot(aes(x="dynamic", y=ive))+
#   stat_summary(aes(x="dynamic",y=ive), 
#                fun=mean, colour="firebrick", geom="point", 
#                shape=18, size=3, show.legend=FALSE)+
#   geom_boxplot(data=filter(vacc_eff_static, vacc_cov==0.75), 
#                aes(x="static",y=ive))+
#   stat_summary(data=filter(vacc_eff_static, vacc_cov==0.75),
#                aes(x="static",y=ive), 
#                fun=mean, colour="firebrick", geom="point", 
#                shape=18, size=3, show.legend=FALSE)+
#   theme_light()+
#   labs(x="", y="Case reduction (%)", title="Indirect VE")
# 
# 
# vacc_eff_dynamic |> 
#   filter(vacc_cov == 0.75) |>
#   filter(id_outbreak %in% ids) |> 
#   ggplot()+
#   geom_boxplot(aes(x="dynamic", y=tve))+
#   stat_summary(aes(x="dynamic",y=tve), 
#                fun=mean, colour="firebrick", geom="point", 
#                shape=18, size=3, show.legend=FALSE)+
#   geom_boxplot(data=filter(vacc_eff_static, vacc_cov==0.75), 
#                aes(x="static",y=tve))+
#   stat_summary(data=filter(vacc_eff_static, vacc_cov==0.75),
#                aes(x="static",y=tve), 
#                fun=mean, colour="firebrick", geom="point", 
#                shape=18, size=3, show.legend=FALSE)+
#   theme_light()+
#   labs(x="", y="Case reduction (%)", title="Total VE")

```


## Likelihood of an outbreak size given the cumulative cases
Then, we can explore the impact of an outbreak given that
```{r}
svim <- fread("outputs/svim_mean_across_runid_20241126.csv")
svim_ct <- fread("outputs/svim_ct_mean_across_runid_20241126.csv")
svim_pe <- fread("outputs/svim_pe_mean_across_runid_20241126.csv")

svim_cea <- fread("outputs/svim_mean_cea_across_runid_20241126.csv")
svim_ct_cea <- fread("outputs/svim_ct_mean_cea_across_runid_20241126.csv")

subset_random <- fread("outputs/svim_mean_random_subset_ids_20241126.csv")
subset_attack_rate <- fread("outputs/svim_mean_attack_rate_subset_ids_20241126.csv")
subset_cases <- fread("outputs/svim_mean_cases_subset_ids_20241126.csv")
subset_duration <- fread("outputs/svim_mean_duration_subset_ids_20241126.csv")
```


```{r}
# outbreak_data |>
#   group_by(week) |>
#   summarize(s_ch_tot = sum(s_ch)) |> 
#   ggplot(aes(week, s_ch_tot))+
#   geom_line() +
#   theme_minimal()+
#   scale_x_log10()
# 
# 
# outbreak_data |>
#   ggplot(aes(week, s_ch, group=id_outbreak))+
#   geom_line(color="grey30") +
#   theme_minimal()+
#   scale_x_log10()+
#   scale_y_log10()

outbreak_data |>
  group_by(week) |>
  summarize(s_ch_tot = sum(s_ch)) -> data_tot 

outbreak_data |>
  filter(week %in% seq(1, 15, 2)) |> 
  ggplot(aes(x=s_ch+1, y=as.factor(week))) +
  geom_density_ridges(fill="steelblue", alpha=0.5) +
  geom_point(data=filter(data_tot, week %in% seq(1, 15, 2)), 
             aes(cumsum(s_ch_tot), as.factor(week)), shape=15) +
  geom_line(data=filter(data_tot, week %in% seq(1, 15, 2)), 
            aes(cumsum(s_ch_tot), as.factor(week)),
            linetype="solid", linewidth=0.7, group=1)+
  scale_x_log10() +
  coord_flip() +
  geom_line(data=filter(data_tot, week %in% seq(1, 15, 2)), 
            aes(s_ch_tot, as.factor(week)), 
            linetype="dotted", linewidth=0.7, group=1) +
  geom_point(data=filter(data_tot, week %in% seq(1, 15, 2)), 
             aes(s_ch_tot, as.factor(week)), shape=15) +
  theme_light()+
  theme_tight()+
  labs(y="Week", x="Suspected cases")


# outbreak_data[, cum_s_ch := cumsum(s_ch), by="id_outbreak"]
# cum_cases <- c(100, 200, 300)
# length(unique(outbreak_data$id_outbreak))
# outbreak_data |>
#   group_by(id_outbreak) |>
#   filter(max(cum_s_ch) > 100) -> d
# 
# length(unique(d$id_outbreak))

target_case <-  c(5,40,80,120,160,200,240,280)
lst <- vector('list', length(target_case))

for (i in 1:length(target_case)) {
  outbreak_data |>
    group_by(id_outbreak) |>
    filter(max(cum_s_ch) > target_case[i]) |>
    ungroup() |>
    filter(cum_s_ch >= target_case[i]) |>
    group_by(id_outbreak) |>
    summarize(min_week = min(week)) |>
    pull(min_week) -> lst[[i]]
}

df <- data.frame()
for (i in 1:length(target_case)) {
  d <- data.frame(cumul_case = target_case[i], week = lst[[i]])
  df <- rbind(df, d)
}
 
df |>
  ggplot(aes(cumul_case, week, group=cumul_case))+
  geom_boxplot(width=8)+
  scale_y_continuous("Weeks required to reach the case number")+
  scale_x_discrete("") +
  theme_light()+
  theme_tight()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())

hist(dw)


outbreak_data |>
  filter(week %in% seq(1, 15, 2)) |> 
  ggplot(aes(week, s_ch, group=id_outbreak))+
  geom_line(color="grey30") +
  geom_density_ridges()+
  theme_minimal()+
  scale_x_log10()+
  scale_y_log10()


outbreak_data |>
  ggplot(aes(week, s_ch, group=id_outbreak))+
  geom_point(color="grey30") +
  theme_minimal()+
  scale_x_log10()+
  scale_y_log10()

outbreak_data |>
  ggplot(aes(week, s_ch, group=id_outbreak))+
  geom_point(size=0.3) +
  theme_minimal()

outbreak_data |>
  filter(week %in% seq(1,200, by=3)) |>
  ggplot()+
  geom_boxplot(aes(as.factor(week), s_ch), outlier.size=0.6) +
  theme_minimal()

outbreak_data |>
  filter(week %in% seq(1,200, by=3)) |>
  ggplot()+
  # geom_boxplot(aes(as.factor(week), s_ch), outlier.size=0.6) +
  geom_violin(aes(as.factor(week), s_ch), outlier.size=0.6) +
  theme_minimal()

```

```{r}
# outbreaks with more than X cases

# impact of outbreaks
outbreak_data |> 
  ggplot(aes(week,s_ch))+
  geom_line(aes(group=id_outbreak), color="grey50")+
  theme_light()+
  labs(x="Week", y="Suspected cases")

outbreak_data[, cum_s_ch := cumsum(s_ch), by="id_outbreak"]
outbreak_data[, s_ch_tot := sum(s_ch), by="id_outbreak"]

outbreak_data |> 
  filter(s_ch_tot > 100)

# Let's look at the situation where we intervene
svim_ct |> 
  # filter(case_trigger %in% c(10, 100, 200, 500,  1000)) |>
  group_by(case_trigger) |> 
  filter(s_ch_tot > case_trigger) |> 
  summarize(n=n(),
            mean_pcr = mean(pct_reduc_case),
            pcr_q500 = quantile(pct_reduc_case, probs=0.5),
            pcr_q975 = quantile(pct_reduc_case, probs=0.975),
            pcr_q025 = quantile(pct_reduc_case, probs=0.025)) -> pcr_cum
  
dlong <- pivot_longer(pcr_cum, cols=c("mean_pcr","pcr_q500","pcr_q975","pcr_q025"))
dlong |> 
  filter(name == "pcr_q500") |> 
  filter(case_trigger %in% c(50, 100, 1000)) |> 
  ggplot()+
  geom_line(aes(x=case_trigger, y=value, group=name, color=name))
  # geom_violin(aes(x=case_trigger, value, group=case_trigger))
```

## Ghost outbreaks because of the lasting immunity from the previous immunization

We first identify outbreaks that might disappear if we consider lasting immunity from prior immunization activities. `find_ghost_outbreaks` function returns ids of the outbreak that might disappear.

```{r, eval=FALSE}
# 
# str <- names(svim)
# names(svim) <- gsub("mean_", "", str) # remove "mean_" to match the variable names in the existing function
# # add attack rate variable for analysis by group 
# dtmp <- ds[, c("attack_rate", "ID_outbreak")]
# names(dtmp) <- c("attack_rate", "id")
# svim <- left_join(svim, dtmp, by="id")  
# 
# svim |> filter(vacc_cov == 0.9, week_vaccination == 4) -> svim_v9w4

svim$location <- sapply(svim$id_outbreak, function(x) strsplit(x, "-")[[1]][1])
adms <- lapply(svim$id_outbreak, function(x) get_adm(x))
# create admin variables to classify outbreaks accordingly
svim$admin0 <- sapply(adms, function(x) x[[1]][2])
svim$admin1 <- sapply(adms, function(x) x[[1]][3])
svim$admin2 <- sapply(adms, function(x) x[[1]][4])
svim$admin3 <- sapply(adms, function(x) x[[1]][5])

# ghost ids are computed only for one scenario of vaccination coverage and week
# in this case, coverage of 90% and week 1
svim_v9w1 <- filter(svim, vacc_cov == 0.9 & vacc_week == 1)
svim_v9w1_sub <- svim_v9w1 |> filter(!is.na(admin0) & !is.na(admin1) & !is.na(admin2))
svim_v9w1_sub <- svim_v9w1_sub[order(svim_v9w1_sub$start_date),] 
  
svim_v9w1_sub |>
  group_by(admin0, admin1, admin2) |>
  mutate(min_end_date = min(end_date)) |> 
  ungroup() -> svim_v9w1_sub

svim_v9w1_sub |> 
  mutate(day_since_previous_outbreak = start_date - min_end_date) -> svim_v9w1_sub

length(unique(svim_v9w1_sub$id_outbreak))
mean_dur_campaign <- 6 # days duration of the vaccination campaign
mean_delay_vacc_eff <- 11 # days, mean delay before the immunity kicks in

x |>
  mutate(date_immunity = start_date + (vacc_week - 1) * 7 + 3.5 +
           mean_dur_campaign + mean_delay_vacc_eff) -> x1

x1 |>
  group_by(admin0, admin1, admin2) |>
  reframe(min_date_immunity = min(date_immunity)) -> x2

x1 <- left_join(x1, x2, by=c("admin0", "admin1", "admin2"))
x1 <- mutate(x1, date_diff = date_immunity - min_date_immunity)

# identify the first outbreak in the region (date_diff = 0) and the outbreaks
# that could disappear (ghost outbreaks, lb <= date_diff <= ub)
ghost_id <- c()
ghost_id <- c(ghost_id, x1[x1$date_diff > lb & x1$date_diff <= ub,]$id_outbreak)
length(ghost_ids)

# the following com
ghost_ids <- find_ghost_outbreaks(x=svim_v9w1_sub, lb=0, ub=365)
length(ghost_ids)
```
