---
title: "Vaccine impact modeling"
format: html
editor_options: 
  code_folding: hide
  chunk_output_type: console
---

## Load ackages and source `utils.R`
```{r}
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
source("utils.R") # utility functions
library(conflicted)
conflict_prefer("filter","dplyr")
# global parameters
VC3 <- c(0.6, 0.75, 0.9) # representative vaccine coverage
VC <- VC3[2] # most representative vaccine coverage
VW <- 9 # median vaccination week from the GTFCC dashboard
VW_seq <- seq(1,19,2)
PROBS <- c(0.025,0.25,0.5,0.75,0.975)
save_csv <- FALSE
save_rds <- FALSE
```

## Load utbreak datasets

```{r}
# Outbreak summary
ds <- clean_names(as.data.frame(fread("data/outbreak_files_Oct2024/outbreak_summary.csv")))
paste0(length(unique(ds$country)), " countries")
# create 'id_outbreak' variables to link with the other dataset
ds |>
  group_by(location, outbreak_number) |>
  mutate(id_outbreak = paste0(location, "-", start_date, "-", end_date)) |> 
  ungroup() -> ds

# OCV campaign records
docv_ <- clean_names(fread("data/outbreak_files_Oct2024/ocv_long_dataset.csv"))
docv <- docv_[(ocv_before_1 == TRUE | ocv_before_2 == TRUE | 
                ocv_before_3 == TRUE | ocv_during == TRUE)]

docv |>
  pivot_longer(cols=ocv_before_1:ocv_after_1, names_to="ocv_timing") |>
  filter(value == T) |>
  group_by(ocv_timing) |>
  count()

# time-series data
ts <- clean_names(readRDS("data/outbreak_files_Oct2024/time_series_outbreak_extraction.rds")) 
ts |>
  group_by(location, outbreak_number) |> # identify unique outbreaks
  mutate(week = row_number(),
         date = tl + 3, # pick the date for the midpoint for each week
         start_date_outbreak = min(tl), 
         end_date_outbreak = max(tr),
         id_outbreak = paste0(location, "-", start_date_outbreak, "-",
                              end_date_outbreak)) |>
  ungroup() -> ts # time series with unique ids

paste0(length(unique(ts$id_outbreak)), " unique outbreaks")
paste0(sum(unique(ts$id_outbreak) %in% ds$id_outbreak), 
       " unique outbreaks matched between time series and outbreak summary")

paste0(length(unique(ts$country)), " countries")

# ids for the outbreaks with confirmed dataset
ts |> 
  group_by(id_outbreak) |> 
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique() -> ids_conf

paste0(length(ids_conf), " confirmed outbreaks")
paste0(length(unique(ts$id_outbreak)) - length(ids_conf), " suspected outbreaks")
# data.table::fwrite(ts, paste0("outputs/outbreak_data_all_", tstamp(), ".csv"))
if (save_csv) {
  fwrite(ts, paste0("outputs/outbreak_data_all_", tstamp(), ".csv"))
}

paste0(sum((unique(ts$id_outbreak)) %in% docv_$id_outbreak), 
      " outbreaks were affected by OCVs")


ids_1 <- unique(ts$id_outbreak)[unique(ts$id_outbreak) %in% docv_$id_outbreak]
ids_2 <- unique(ts$id_outbreak)[unique(ts$id_outbreak) %in% docv$id_outbreak]
ids_0 <- setdiff(ids_1, ids_2) 

# outbreak time series not affected by OCV's
ts_noocv <- ts[!(ts$id_outbreak %in% docv$id_outbreak), ]

if (save_csv) {
  fwrite(ts_noocv, paste0("outputs/outbreak_data_", tstamp(), ".csv"))
}
paste0(sum((unique(ts_noocv$id_outbreak)) %in% docv$id_outbreak), 
      " outbreaks from modeled data were affected by OCVs")

ts_noocv |> 
  group_by(id_outbreak) |> 
  mutate(c_ch_sum = sum(c_ch)) |> 
  filter(c_ch_sum > 0) |> 
  pull(id_outbreak) |>
  unique() -> ids_conf_noocv

paste0(length(ids_conf_noocv), " confirmed outbreaks out of ", length(unique(ts_noocv$id_outbreak)), " outbreaks unaffected by OCVs")
paste0(length(unique(ts_noocv$id_outbreak)) - length(ids_conf_noocv), " suspected outbreaks")

outbreak_data <- as.data.frame(fread("outputs/outbreak_data_20250228.csv"))
outbreak_data_all <- as.data.frame(fread("outputs/outbreak_data_all_20250228.csv"))

outbk_data_summary <- 
  left_join(outbreak_data, 
            ds[,c("outbreak_pop", "total_s_ch", "total_c_ch", "total_deaths", "duration_weeks", 
"peak_week_incidence", "time_to_peak_week", "peak_week_incidence_per_100percent", 
"peak_week_incidence_per_10e3", "mean_r0_1st_week", "area_per_1km2", 
"rural_urban", "ar_per_10e3", "cfr_per_100percent", "id_outbreak"
)], by="id_outbreak")
```

## Load parameter sets
```{r}
## simulation parameters
params <- as.data.frame(fread("outputs/params_20241017.csv"))
# IVE posterior predictive values from Bayesian beta regression (see parameters.qmd)
ive_fit <- read_rds("outputs/ive_yrep_20241010.rds")
ive_data <- ive_fit$yrep
rm(ive_fit)
pop_by_age_wpp <- as.data.frame(fread("outputs/wpp_pop_by_age_20241022.csv"))
life_exp_data <- as.data.frame(fread("outputs/wpp_life_expectancy_20241022.csv"))
gdp_data <- read_xls("data/GDP_WorldBank.xls")
workforce_data <- read_xls("data/Workforce_Worldbank.xls")

## Cost-effectiveness analysis parameters
# Disease symptoms, vaccine costs, etc.
parms <- as.data.frame(fread("data/parameters.csv"))
parms <- filter(parms, Disease == "Cholera")
pr_asymp <- parms[parms$Parameter == "Prop_Asymptomatic", "Value"]
pr_mild <- parms[parms$Parameter == "Prop_Mild", "Value"]
pr_moderate <- parms[parms$Parameter == "Prop_Moderate", "Value"]
pr_severe <- parms[parms$Parameter == "Prop_Severe", "Value"]
dur <- parms[parms$Parameter == "Duration_Illness", "Value"]
wt_asymp <- parms[parms$Parameter == "Disability_Weight_Asymptomatic", "Value"]
wt_mild <- parms[parms$Parameter == "Disability_Weight_Mild", "Value"]
wt_moderate <- parms[parms$Parameter == "Disability_Weight_Moderate", "Value"]
wt_severe <- parms[parms$Parameter == "Disability_Weight_Severe", "Value"]

vacc_cost_per_dose <- parms[parms$Parameter == "Vaccine_Cost", "Value"]
vacc_delivery_cost <- parms[parms$Parameter == "Vaccine_Delivery_Cost", "Value"]
patient_cost_hosp <- parms[parms$Parameter == "Patient_Cost_Hosp", "Value"]
patient_cost_outpt <- parms[parms$Parameter == "Patient_Cost_Outpt", "Value"]
public_cost_hosp <- parms[parms$Parameter == "Public_Cost_Hosp", "Value"]
public_cost_outpt <-parms[parms$Parameter == "Public_Cost_Outpt", "Value"] 
patient_workday_lost <- parms[parms$Parameter == "Pt_Workdays_Lost", "Value"]
caregiver_workday_lost <- parms[parms$Parameter == "Caregiver_Workdays_Lost", "Value"]
mean_age_inf <- parms[parms$Parameter == "Mean_Age_Infection", "Value"]
vacc_shipping_cost <- parms[parms$Parameter == "Vaccine_Shipping_Cost", "Value"]

dose_regimen <- 1
vacc_cost_per_person <- vacc_cost_per_dose * dose_regimen

# assume that reported cases are either moderate or severe cases
pr_tot <- pr_moderate + pr_severe
(coi_per_patient <- 
    pr_moderate / pr_tot * (patient_cost_outpt + public_cost_outpt) + 
    pr_severe / pr_tot * (patient_cost_hosp + public_cost_hosp))

avg_age_inf <- parms[parms$Parameter == "Mean_Age_Infection", "Value"]

## Cost adjustment 2015 to 2023
# consumer price index for cost adjustment (2015 to 2023)
cpi_data <- read_xls("data/API_FP.CPI.TOTL.ZG_DS2_en_excel_v2_9839.xls")
# another gdp_data used previously
gdp_data2 <- read_xls("data/API_NY.GDP.DEFL.KD.ZG_DS2_en_excel_v2_9995.xls")

cpi1 <- cpi_data %>%
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "value")
cpi1$year <- as.numeric(cpi1$year)
cpi1 <- cpi1 %>% filter(year >= 2015 & year <= 2023)

gdp1 <- gdp_data2 %>%
  pivot_longer(cols = `2001`:`2023`, names_to = "year", values_to = "value")
gdp1$year <- as.numeric(gdp1$year)
gdp1 <- gdp1 %>% filter(year >= 2015 & year <= 2023)

# Use % change in CPI for sub-Saharan African region, which is available

cpi_percent <- read_xls("data/imf-dm-export-20241028.xls")

cpi_percent <- cpi_percent %>%
  pivot_longer(
    cols = `2015`:`2029`,         
    names_to = "year",             
    values_to = "value"            
  )

cpi_percent$year <- as.numeric(cpi_percent$year)

# year starts at the t + 1 year of reported cost, so if 2015 cost then should use
# % change CPI starting 2016
cpi_percent <- cpi_percent %>% 
  filter(year >= 2016 & year <= 2023)

xi <- (cpi_percent$value/100 + 1)
# cumulative_inflation_factor <- xi[1] * xi[2] * xi[3] * xi[4] * xi[5] * xi[6] * xi[7] * xi[8]
cumulative_inflation_factor <- prod(xi)
patient_cost_hosp_adj <- patient_cost_hosp * cumulative_inflation_factor
patient_cost_outpt_adj <- patient_cost_outpt * cumulative_inflation_factor
public_cost_hosp_adj <- public_cost_hosp * cumulative_inflation_factor
public_cost_outpt_adj <- public_cost_outpt * cumulative_inflation_factor

# vaccine delivery cost was reported in 2016 USD$
cpi_percent <- cpi_percent %>% 
  filter(year >= 2017 & year <= 2023)

xi <- (cpi_percent$value/100 + 1)
# cumulative_inflation_factor <- xi[1] * xi[2] * xi[3] * xi[4] * xi[5] * xi[6] * xi[7]
cumulative_inflation_factor <- prod(xi)
vacc_delivery_cost_adj <- vacc_delivery_cost * cumulative_inflation_factor

patient_cost_hosp <- patient_cost_hosp_adj
patient_cost_outpt <- patient_cost_outpt_adj
public_cost_hosp <- public_cost_hosp_adj
public_cost_outpt <- public_cost_outpt_adj
vacc_delivery_cost <- vacc_delivery_cost_adj

## Life expectancy and other cost data
# the following gdp is different from the overall sub-Saharan Africa gdp
# gdp_data2 <- read_xls("data/GDP_WorldBank.xls")
names(life_exp_data)[names(life_exp_data) == "ISO3 Alpha-code"] <- "country"
names(life_exp_data)[names(life_exp_data) == "Year"] <- "year"
names(workforce_data)[names(workforce_data) == "Country Code"] <- "country"

workforce_long <- 
  workforce_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols=`2001`:`2023`, names_to="year", values_to="pct_workforce")
workforce_long$year <- as.integer(workforce_long$year)

names(gdp_data)[names(gdp_data) == "Country Code"] <- "country"
gdp_long <- 
  gdp_data |> 
  select(c("country", `2001`:`2023`)) |>
  pivot_longer(cols=`2001`:`2023`, names_to="year", values_to="gdp")
gdp_long$year <- as.integer(gdp_long$year)
```

## Vaccine impact simulation

Parallel by batch `run_vacc_impact_outbreak_weekly`

```{r}
tic <- Sys.time()
dat <- outbreak_data
library(doParallel)
# vacc_covs <- seq(0.6, 0.95, by=0.05)
vacc_covs <- c(0.6, 0.75, 0.9)
vacc_weeks <- seq(1, 20, by=1) # vaccination can happen anytime during week 1 and week 20
nruns <- 50 # number of simulation runs
batch_size <- 10

for (batch_id in seq(1, nruns, by = batch_size)) {
  cat("batch id: ", batch_id, "\n")
  run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in batches of batch_size

  cl <- makeCluster(getOption("cl.cores", detectCores() - 1))
  registerDoParallel(cl)

  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = dat,
        ive_data = ive_data,
        parameters = params,
        age_dist = pop_by_age_wpp,
        vacc_week = vacc_weeks,
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)

  fwrite(rbindlist(svim_week_list), paste0("outputs/svim_week_list_vacc_week_batch_",
                                 batch_id, "_", tstamp(), ".csv"))
}

```

Parallel case_trigger by batch

```{r, eval=FALSE}
library(doParallel)
dat <- outbreak_data
case_triggers <- c(1,seq(10,300,by=10),seq(400,1000,by=100))
batch_size <- 10
# vacc_covs <- seq(0.6, 0.95, by=0.05)
vacc_covs <- c(0.6, 0.75, 0.9)
nruns <- 10 # number of simulation runs

for (batch_id in seq(1, nruns, by = batch_size)) {
  cat("batch id: ", batch_id, "\n")
  run_ids <- batch_id:(batch_id + batch_size - 1)  # Run IDs in batches of batch_size

  cl <- makeCluster(getOption("cl.cores", detectCores() - 4))
  registerDoParallel(cl)

  svim_week_list <-
    foreach(i = run_ids, .packages = c("data.table")) %dopar% {
      run_vacc_impact_outbreak_weekly(
        outbreak_data = dat,
        ive_data = ive_data,
        parameters = params,
        age_dist = pop_by_age_wpp,
        case_trigger = case_triggers,
        vacc_cov = vacc_covs,
        runid = i)
    }
  stopCluster(cl)
  fwrite(rbindlist(svim_week_list), paste0("outputs/svim_week_case_trigger_batch_",
                               batch_id, "_", tstamp(), ".csv"))
}
```

Parallel preemptive vaccination `run_vacc_impact_outbreak_weekly`

```{r, eval=FALSE}
library(doParallel)

tic <- Sys.time()
dat <- outbreak_data
# vacc_covs <- seq(0.6, 0.95, by=0.05)
vacc_covs <- c(0.6, 0.75, 0.9)
# maximum possible delay before vaccines take effect after vaccination
# max(params$dur_campaign)/7 + max(params$delay_vacc_effect)/7
vacc_weeks <- -52 # some random large number ensure than pre-emptive vaccination
nruns <- 200 # number of simulation runs

cl <- makeCluster(getOption("cl.cores", detectCores() - 6))
registerDoParallel(cl)

svim_week_list <-
  foreach(i = 1:nruns, .packages = c("data.table")) %dopar% {
    run_vacc_impact_outbreak_weekly(
      outbreak_data = dat,
      ive_data = ive_data,
      parameters = params,
      age_dist = pop_by_age_wpp,
      vacc_week = vacc_weeks,
      vacc_cov = vacc_covs,
      runid = i)
  }
stopCluster(cl)

Sys.time() - tic
# Time difference of 36.02764 mins
fwrite(rbindlist(svim_week_list), paste0("outputs/svim_week_preemptive_", tstamp(), ".csv"))
```

Combine the results from parallel simulation
```{r}
# library(data.table)
# source("utils.R")
id_begin <- 1
id_end <- 200
batch_size <- 10
batch_ids <- seq(id_begin, id_end, by=batch_size)

svim_df <- data.frame()
for (i in 1:length(batch_ids)) {
  d <- fread(paste0("outputs/svim_week_list_vacc_week_batch_", batch_ids[i],
           "_20250311.csv"))
  svim_df <- rbindlist(list(svim_df, sum_over_week(d)))
  rm(d)
}

fwrite(svim_df, paste0("outputs/svim_vc_85_time_trigger_sum_over_week_runid_1_200_", 
                        tstamp(), ".csv"))

# case-triggered vaccination
id_begin <- 1
id_end <- 120
batch_size <- 10
batch_ids <- seq(id_begin, id_end, by=batch_size)

svim_df <- data.frame()

for ( i in 1:length(batch_ids)) {
  cat("i =", i, "\n")
  d <- fread(paste0("outputs/svim_case_trigger_week_list_batch_",
                    batch_ids[i], "_20250312.csv"))
  svim_df <- rbindlist(list(svim_df, sum_over_week(d, case_trigger=T)))
  rm(d)
}

fwrite(svim_df, paste0("outputs/svim_vc_85_case_trigger_sum_over_week_runid_1_200_", 
                       tstamp(), ".csv"))

# preemptive vaccination
svim_df <- fread("outputs/svim_week_preemptive_20250228.csv")
svim_df <- sum_over_week(svim_df)
fwrite(svim_df, paste0("outputs/svim_preemptive_sum_over_week_runid_1_200_",
                       tstamp(), ".csv"))
```

## Summarize vaccine impact

Computes (%) case averted, case averted per 1,000 OCVs and cost effectiveness 
across vaccination timing (time-trigger or case-trigger) and vaccine coverage

### Time-triggered vaccination

```{r}
# sum over week by runid
# d <- fread("outputs/svim_time_trigger_sum_over_week_runid_1_200_20250302.csv")
d <- fread("outputs/svim_vc_85_time_trigger_sum_over_week_runid_1_200_20250313.csv")
VC3 <- 0.85
# library(tidyverse)
for (vc in VC3) {
  cat("vacc_cov =" , vc, "\n")
  # cost-effectiveness analysis
  svim_tt_cea <- add_cea_variables(d[vacc_cov == vc])
  svim_tt_cea <- add_cea_results(svim_tt_cea)
  fwrite(svim_tt_cea,
   paste0("outputs/cea_", vc, "_time_trigger_", 
          tstamp(), ".csv"))
  rm(svim_tt_cea)
  
  # calculate the mean across runids
  svim_tt_mean <- mean_across_runid(d[vacc_cov == vc])
  svim_tt_mean_cea <- add_cea_variables(svim_tt_mean)
  rm(svim_tt_mean)
  svim_tt_mean_cea <- add_cea_results(svim_tt_mean_cea)
  
  fwrite(svim_tt_mean_cea,
    paste0("outputs/cea_", vc, "_time_trigger_mean_across_runid_", 
             tstamp(), ".csv"))
  rm(svim_tt_mean_cea)
}

rm(d)

```

### case-triggered vaccination

```{r}
# case-trigger vaccination
# d_ct <- fread("outputs/svim_case_trigger_sum_over_week_runid_1_200_20250304.csv")
d_ct <- fread("outputs/svim_vc_85_case_trigger_sum_over_week_runid_1_200_20250313.csv")
for (vc in VC3) {
  cat("vacc_cov =" , vc, "\n")
  svim_ct_cea <- add_cea_variables(d_ct[vacc_cov == vc])
  svim_ct_cea <- add_cea_results(svim_ct_cea)
  
  fwrite(svim_ct_cea,
   paste0("outputs/cea_", vc, "_case_trigger_", 
          tstamp(), ".csv"))
  rm(svim_ct_cea)
    
  svim_ct_mean_cea <- add_cea_variables(d_ct[vacc_cov == vc])
  svim_ct_mean_cea <- add_cea_results(svim_ct_mean_cea)
  
  fwrite(svim_ct_mean_cea,
    paste0("outputs/cea_", vc, "_case_trigger_mean_across_runid_", 
             tstamp(), ".csv"))
  rm(svim_ct_mean_cea)
}
```

### Preemptive vaccination
```{r}
# pre-emptive vaccination
# svim_pe_df <- fread("outputs/svim_preemptive_raw_20241203.csv")
svim_pe_df <- readRDS("outputs/docv_svim_preemptive_raw_20250227.rds")
svim_pe <- set_svim_raw_res(svim_pe_df)
svim_pe_sum_across_week_by_runid <- svim_sum_across_week(svim_pe)
if (save_csv){
  saveRDS(svim_pe_sum_across_week_by_runid,
         paste0("outputs/docv_svim_pe_sum_across_week_by_runid_", tstamp(), ".rds"))
}
svim_pe_mean <- svim_mean_across_runid(svim_pe_sum_across_week_by_runid)
svim_pe_mean <- add_summary_stats(svim_pe_mean)
if (save_csv) {
  fwrite(svim_pe_mean,
    paste0("outputs/svim_pe_mean_across_runid_", tstamp(), ".csv"))
}
svim_pe_mean_cea <- add_cea_variables(svim_pe_mean)
svim_pe_mean_cea <- add_cea_results(svim_pe_mean_cea)
if (save_csv) {
  fwrite(svim_pe_mean_cea,
    paste0("outputs/svim_pe_mean_cea_across_runid_", tstamp(), ".csv"))
}
```

### All outbreaks collectively

Calculate the vaccine impact after aggregating all outbreaks.
The resulting outputs are one vaccine impact estimate for runid 
(i.e., parameter set) for each of vaccination scenarios (i.e., vaccination 
week and vaccine coverage). Therefore, 4,000 rows are expected given that 20 vaccination weeks 
(1, 2, ..., 20), one vaccine coverage (e.g., 0.75), and 200 runid's

```{r}
# inputs are outcomes at the level of individual outbreaks
d <- fread("outputs/cea_0.75_time_trigger_20250228.csv")
cea_all <- sum_over_outbreaks_by_runid(d)

fwrite(cea_all,
    paste0("outputs/cea_all_vc_75_by_runid_", 
           tstamp(), ".csv"))
```

## Outbreak subsets

```{r}
outbk_data_summary |> 
  group_by(id_outbreak) |> 
  summarize(year = year[1],
            rural_urban = rural_urban[1],
            ar = ar_per_10e3[1],
            cfr = cfr_per_100percent[1],
            spatial_scale = spatial_scale[1],
            pop = pop[1],
            # pop2 = outbreak_pop[1],
            duration = max(week),
            # duration2 = duration_weeks[1],
            size = sum(s_ch)) -> d

unique(d$spatial_scale)

cumsum(d$pop)
# Doses shipped
# 216.9M
# 166.2M (77%) shipped with ICG
dat <- clean_names(fread("data/OCV_Use_GTFCC_Dashboard.csv"))
dat |> 
  filter(shipped == "Shipped", type == "Reactive") |> 
  group_by(year) |> 
  summarize(num_ocv = sum(dose)) |>
  setDT() -> ocv_year
ocv_year$cum_num_ocv <- cumsum(ocv_year$num_ocv)

dat <- filter(d, spatial_scale %in% c("admin1","admin2", "admin3"))
nsim <- 200

lst <- vector('list', nsim) # no outbreak data for 2024 (i.e., -1)

DF <- data.frame()
for (j in 1:200) {
  set.seed(j)
  cat("j =", j, "\n")
  df <- data.frame()
  for (i in 2013:2022) {
    yr <- i  # 2013 = the first year when OCVs were deployed
    # cat("year =", yr, "\n")
    smpl <- sample_outbreaks(
      outbreak_data = filter(dat, year == yr), 
      target_population = ocv_year[year == yr, num_ocv],
      sampling_method = "random")
    
    df <- rbind(df, data.frame(year=smpl$sampled_data$year, 
                         id_outbreak = smpl$sampled_data$id_outbreak))
  }
  df$subset_id <- j
  DF <- rbind(DF, df)
}

fwrite(DF, paste0("outputs/random_subsets_by_year_", tstamp(), ".csv"))
```

### Subset analysis - time triggered

```{r}
# svim_tt <- fread("outputs/svim_time_trigger_sum_over_week_runid_1_200_20250302.csv")
DF <- fread("outputs/random_subsets_by_year_20250306.csv")
cea_tt <- fread("outputs/cea_0.75_time_trigger_20250304.csv")
df <- data.frame(expand.grid(subset_id = 1:200, 
                             year_max = 2013:2022))

lst <- vector('list', nrow(df))
for (i in 1:nrow(df)) {
  d <- DF[(DF$year <= df$year_max[i] & DF$subset_id == df$subset_id[i]),]
  dat <- cea_tt[id_outbreak %in% d$id_outbreak]
  datall <- sum_over_outbreaks_by_runid(dat)
  datall$subset_id <- df$subset_id[i]
  datall$year_max <- df$year_max[i]
  lst[[i]] <- datall
}
dfall <- rbindlist(lst)

fwrite(dfall, paste0("outputs/cea_0.75_time_trigger_subsets_by_year_", tstamp(), ".csv"))
```

### Subset analysis - case-triggered

```{r}
DF <- fread("outputs/random_subsets_by_year_20250306.csv")
cea_ct <- fread("outputs/cea_0.75_case_trigger_20250311.csv")
df <- data.frame(expand.grid(subset_id = 1:200, 
                             year_max = 2013:2022))

lst <- vector('list', nrow(df))
for (i in 1:nrow(df)) {
  d <- DF[(DF$year <= df$year_max[i] & DF$subset_id == df$subset_id[i]),]
  dat <- cea_ct[id_outbreak %in% d$id_outbreak]
  datall <- sum_over_outbreaks_by_runid(dat, case_trigger = T)
  datall$subset_id <- df$subset_id[i]
  datall$year_max <- df$year_max[i]
  lst[[i]] <- datall
}
dfall <- rbindlist(lst)

fwrite(dfall, paste0("outputs/cea_0.75_case_trigger_subsets_by_year_", tstamp(), ".csv"))
```
 

